### 2017-05-30 00-Abstract.Rmd
init

# Résumé et Abstract {-}

Écrire le résumé français ici...

Write the english abstract here...

### 2017-05-30 02-Introduction.Rmd
init

# Introduction {-}

* Introduce why tropical rainforest are such a good study case
* Define biodiversity 
* Define ecosystem functionning
* Define ecosystem services
* Introduce tropical sylviculture
* Define selective logging
* Introduce biodiversity erosion in tropical rainforests ant its impact
* Introduce Loreau partitionning
* Introduce the interest of modelling
* Introduce the idea to look at biodiversity as both a parameter and an output

Tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].

Tropical forests are primary ecosystems in terms of biodiversit and carbon storage [@Lewis2004].

High biodiversity from tropical rainforests is the source of many ecosystem functions, which support ecosystem services.

Biodiversity and ecosystem functionning field is emmering [@Loreau2000].

Selective logging is increasing in tropical forests.

Tropical logging accounts for one eight of global timber production [@Blaser2011].

Selective logging represents a major threat to biodiversity [@Gibson2013].

### 2017-05-30 03-MnMs.Rmd
init

# Materials and methods

## TROLL simulator

### Abiotic environment

### Photosynthesis

### Autotrpohic respiration

### Carbon uptake

### Tree growth

### Seed dispersion, production and recruitment

### Mortality

## TROLL sensitivity analysis

### Functional traits

### Seed rain

## Disturbance

### Disturbance module

### Design of experiment

### Outputs anlaysis ?

#### Resistance and resilience metrics

#### Biodiversity partitioning

## Selective logging

### Selective logging module

#### Designation

#### Selection

#### Rotten trees

#### Felling

#### Tracks

#### Gap damages

### Design of experiment

### Outputs analysis ?

#### Resistance and resilience metrics

#### Biodiversity partitioning

### 2017-05-30 docs/master-thesis.tex
init

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Master thesis},
            pdfauthor={Sylvain SCHMITT},
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Master thesis}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Sylvain SCHMITT}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{2017-05-30}

\usepackage{booktabs}
\usepackage{amsthm}
\makeatletter
\def\thm@space@setup{%
  \thm@preskip=8pt plus 2pt minus 4pt
  \thm@postskip=\thm@preskip
}
\makeatother

\begin{document}
\maketitle

\date{}

\thispagestyle{empty}
\begin{center}
\includegraphics{images/main.pdf}
\end{center}

\setlength{\abovedisplayskip}{-5pt}
\setlength{\abovedisplayshortskip}{-5pt}

{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{2}
\tableofcontents
}
\section*{Preface}\label{preface}
\addcontentsline{toc}{section}{Preface}

This is the online html version of my master thesis. The pdf version is
available for download with the download button on top of the windows.

Have a nice reading.

\emph{Sylvain}

\section*{Résumé et Abstract}\label{resume-et-abstract}
\addcontentsline{toc}{section}{Résumé et Abstract}

Écrire le résumé français ici\ldots{}

Write the english abstract here\ldots{}

\section*{Acknowledgments}\label{acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}

I would like to thank\ldots{}

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

\begin{itemize}
\tightlist
\item
  Introduce why tropical rainforest are such a good study case
\item
  Define biodiversity
\item
  Define ecosystem functionning
\item
  Define ecosystem services
\item
  Introduce tropical sylviculture
\item
  Define selective logging
\item
  Introduce biodiversity erosion in tropical rainforests ant its impact
\item
  Introduce Loreau partitionning
\item
  Introduce the interest of modelling
\item
  Introduce the idea to look at biodiversity as both a parameter and an
  output
\end{itemize}

Tropical rainforests have fascinated ecologists due to their outstanding
diversity \citep{connell_diversity_1978}.

Tropical forests are primary ecosystems in terms of biodiversit and
carbon storage \citep{Lewis2004}.

High biodiversity from tropical rainforests is the source of many
ecosystem functions, which support ecosystem services.

Biodiversity and ecosystem functionning field is emmering
\citep{Loreau2000}.

Selective logging is increasing in tropical forests.

Tropical logging accounts for one eight of global timber production
\citep{Blaser2011}.

Selective logging represents a major threat to biodiversity
\citep{Gibson2013}.

\section{Materials and methods}\label{materials-and-methods}

\subsection{TROLL simulator}\label{troll-simulator}

\subsubsection{Abiotic environment}\label{abiotic-environment}

\subsubsection{Photosynthesis}\label{photosynthesis}

\subsubsection{Autotrpohic respiration}\label{autotrpohic-respiration}

\subsubsection{Carbon uptake}\label{carbon-uptake}

\subsubsection{Tree growth}\label{tree-growth}

\subsubsection{Seed dispersion, production and
recruitment}\label{seed-dispersion-production-and-recruitment}

\subsubsection{Mortality}\label{mortality}

\subsection{TROLL sensitivity
analysis}\label{troll-sensitivity-analysis}

\subsubsection{Functional traits}\label{functional-traits}

\subsubsection{Seed rain}\label{seed-rain}

\subsection{Disturbance}\label{disturbance}

\subsubsection{Disturbance module}\label{disturbance-module}

\subsubsection{Design of experiment}\label{design-of-experiment}

\subsubsection{Outputs anlaysis ?}\label{outputs-anlaysis}

\paragraph{Resistance and resilience
metrics}\label{resistance-and-resilience-metrics}

\paragraph{Biodiversity partitioning}\label{biodiversity-partitioning}

\subsection{Selective logging}\label{selective-logging}

\subsubsection{Selective logging module}\label{selective-logging-module}

\paragraph{Designation}\label{designation}

\paragraph{Selection}\label{selection}

\paragraph{Rotten trees}\label{rotten-trees}

\paragraph{Felling}\label{felling}

\paragraph{Tracks}\label{tracks}

\paragraph{Gap damages}\label{gap-damages}

\subsubsection{Design of experiment}\label{design-of-experiment-1}

\subsubsection{Outputs analysis ?}\label{outputs-analysis}

\paragraph{Resistance and resilience
metrics}\label{resistance-and-resilience-metrics-1}

\paragraph{Biodiversity partitioning}\label{biodiversity-partitioning-1}

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\bibliography{bibliography/library.bib}

\end{document}

### 2017-05-30 index.Rmd
init

--- 
site: "bookdown::bookdown_site"
title: "Master thesis"
author: "Sylvain SCHMITT"
date: "`r Sys.Date()`"
output: bookdown::gitbook
csl: bibliography/mee.csl
bibliography: bibliography/library.bib
link-citations: yes
colorlinks: yes
cover-image: images/main.png
---

# Preface {-}

```{r fig.align='center', echo=FALSE, include=identical(knitr:::pandoc_to(), 'html')}
knitr::include_graphics('images/main.png', dpi = NA)
```

This is the online html version of my master thesis. The pdf version is available for download with the download button on top of the windows.

Have a nice reading.

<div align="right">  *Sylvain*

### 2017-05-30 latex/before_body.tex
init

--- 2017-05-30 docs/master-thesis.tex
@@
-\documentclass[]{article}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{unicode=true,
-            pdftitle={Master thesis},
-            pdfauthor={Sylvain SCHMITT},
-            colorlinks=true,
-            linkcolor=Maroon,
-            citecolor=Blue,
-            urlcolor=Blue,
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{natbib}
-\bibliographystyle{plainnat}
-\usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-  \title{Master thesis}
-  \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \author{Sylvain SCHMITT}
-  \preauthor{\centering\large\emph}
-  \postauthor{\par}
-  \predate{\centering\large\emph}
-  \postdate{\par}
-  \date{2017-05-30}
-
-\usepackage{booktabs}
-\usepackage{amsthm}
-\makeatletter
-\def\thm@space@setup{%
-  \thm@preskip=8pt plus 2pt minus 4pt
-  \thm@postskip=\thm@preskip
-}
-\makeatother
-
-\begin{document}
-\maketitle
-
 \date{}
@@
 \setlength{\abovedisplayshortskip}{-5pt}
-
-{
-\hypersetup{linkcolor=black}
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
-\section*{Preface}\label{preface}
-\addcontentsline{toc}{section}{Preface}
-
-This is the online html version of my master thesis. The pdf version is
-available for download with the download button on top of the windows.
-
-Have a nice reading.
-
-\emph{Sylvain}
-
-\section*{Résumé et Abstract}\label{resume-et-abstract}
-\addcontentsline{toc}{section}{Résumé et Abstract}
-
-Écrire le résumé français ici\ldots{}
-
-Write the english abstract here\ldots{}
-
-\section*{Acknowledgments}\label{acknowledgments}
-\addcontentsline{toc}{section}{Acknowledgments}
-
-I would like to thank\ldots{}
-
-\section*{Introduction}\label{introduction}
-\addcontentsline{toc}{section}{Introduction}
-
-\begin{itemize}
-\tightlist
-\item
-  Introduce why tropical rainforest are such a good study case
-\item
-  Define biodiversity
-\item
-  Define ecosystem functionning
-\item
-  Define ecosystem services
-\item
-  Introduce tropical sylviculture
-\item
-  Define selective logging
-\item
-  Introduce biodiversity erosion in tropical rainforests ant its impact
-\item
-  Introduce Loreau partitionning
-\item
-  Introduce the interest of modelling
-\item
-  Introduce the idea to look at biodiversity as both a parameter and an
-  output
-\end{itemize}
-
-Tropical rainforests have fascinated ecologists due to their outstanding
-diversity \citep{connell_diversity_1978}.
-
-Tropical forests are primary ecosystems in terms of biodiversit and
-carbon storage \citep{Lewis2004}.
-
-High biodiversity from tropical rainforests is the source of many
-ecosystem functions, which support ecosystem services.
-
-Biodiversity and ecosystem functionning field is emmering
-\citep{Loreau2000}.
-
-Selective logging is increasing in tropical forests.
-
-Tropical logging accounts for one eight of global timber production
-\citep{Blaser2011}.
-
-Selective logging represents a major threat to biodiversity
-\citep{Gibson2013}.
-
-\section{Materials and methods}\label{materials-and-methods}
-
-\subsection{TROLL simulator}\label{troll-simulator}
-
-\subsubsection{Abiotic environment}\label{abiotic-environment}
-
-\subsubsection{Photosynthesis}\label{photosynthesis}
-
-\subsubsection{Autotrpohic respiration}\label{autotrpohic-respiration}
-
-\subsubsection{Carbon uptake}\label{carbon-uptake}
-
-\subsubsection{Tree growth}\label{tree-growth}
-
-\subsubsection{Seed dispersion, production and
-recruitment}\label{seed-dispersion-production-and-recruitment}
-
-\subsubsection{Mortality}\label{mortality}
-
-\subsection{TROLL sensitivity
-analysis}\label{troll-sensitivity-analysis}
-
-\subsubsection{Functional traits}\label{functional-traits}
-
-\subsubsection{Seed rain}\label{seed-rain}
-
-\subsection{Disturbance}\label{disturbance}
-
-\subsubsection{Disturbance module}\label{disturbance-module}
-
-\subsubsection{Design of experiment}\label{design-of-experiment}
-
-\subsubsection{Outputs anlaysis ?}\label{outputs-anlaysis}
-
-\paragraph{Resistance and resilience
-metrics}\label{resistance-and-resilience-metrics}
-
-\paragraph{Biodiversity partitioning}\label{biodiversity-partitioning}
-
-\subsection{Selective logging}\label{selective-logging}
-
-\subsubsection{Selective logging module}\label{selective-logging-module}
-
-\paragraph{Designation}\label{designation}
-
-\paragraph{Selection}\label{selection}
-
-\paragraph{Rotten trees}\label{rotten-trees}
-
-\paragraph{Felling}\label{felling}
-
-\paragraph{Tracks}\label{tracks}
-
-\paragraph{Gap damages}\label{gap-damages}
-
-\subsubsection{Design of experiment}\label{design-of-experiment-1}
-
-\subsubsection{Outputs analysis ?}\label{outputs-analysis}
-
-\paragraph{Resistance and resilience
-metrics}\label{resistance-and-resilience-metrics-1}
-
-\paragraph{Biodiversity partitioning}\label{biodiversity-partitioning-1}
-
-\section{Results}\label{results}
-
-\section{Discussion}\label{discussion}
-
-\bibliography{bibliography/library.bib}
-
-\end{document}

### 2017-05-30 index.Rmd
wip

@@
 
-# Preface {-}
+`r if (knitr:::is_html_output()) '# Preface {-}'`
+
+`r if (knitr:::is_html_output()) '![stability-wip](https://img.shields.io/badge/stability-work_in_progress-lightgrey.svg)'`
 
@@
 
-This is the online html version of my master thesis. The pdf version is available for download with the download button on top of the windows.
+`r if (knitr:::is_html_output()) 'This is the online html version of my master thesis. The pdf version is available for download with the download button on top of the windows.'`
 
-Have a nice reading.
+`r if (knitr:::is_html_output()) 'Have a nice reading.'`
 
-<div align="right">  *Sylvain*
+`r if (knitr:::is_html_output()) '<div align="right">  *Sylvain*'`

### 2017-05-30 03-MnMs.Rmd
DOE plot test

@@
+```{r config, message=FALSE, warning=FALSE, include=FALSE}
+rm(list = ls()) ; invisible(gc())
+library(knitr)
+library(parallel)
+library(TROLL)
+library(RconTroll)
+library(ggplot2)
+library(ggrepel)
+library(cowplot)
+library(entropart)
+library(plotly)
+library(readr)
+opts_chunk$set(
+  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
+    cache = T, cache.lazy = F)
+```
+
 # Materials and methods
@@
 
+```{r DOE, fig.cap='Experimental design before disturbance.
+Communities are implemented along a gradient of species richness (SR) and functional dispersion (FDis) resulting in a broad range of aboveground biomass (AGB).
+FDis was caluclated based on 4 functional traits (leaf mass per area, wood specific gravity, maximum diameter, maximum height).'}
+# Data prep
+load('~/Documents/ECOFOG/Results/disturbance/maturity/matureData.Rdata')
+data <- datal$time
+data <- data[data$time > 450,]
+data <- data.frame(apply(data[c('agb','ba','n10','n30')],
+                         2, function(x) tapply(x, data$sim, mean)))
+data <- cbind(data.frame(sim = row.names(data)), data)
+data$richness <- as.factor(floor(as.numeric(as.character(data$sim))))
+species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt',
+                      header=T, dec=".", sep="",
+                      row.names = 1)[c('LMA', 'wsg', 'dmax', 'hmax')]
+species <- species[row.names(datal$species),]
+fd <- suppressWarnings(FD::dbFD(species, t(datal$species), messages = F))
+data <- cbind(data, do.call('cbind',
+                            fd[c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ')]))
+data <- cbind(data, fd$CWM)
+rm(fd, species, datal)
+
+# Graph
+g <- ggplot(data, aes(x = FDis, y = richness, 
+                      label = sim, color = agb/1000)) + 
+  geom_point(size = 3) + 
+  geom_text_repel() +
+  scale_color_continuous("Aboveground\nBiomass\n(AGB, ton C/ha)") +
+  xlab("Functional dispersion (FDis)") +
+  ylab("Species richness (SR)")
+
+g ; rm(data)
+```
+
 ### Outputs anlaysis ?

### 2017-05-30 docs/master-thesis.tex
wip
DOE plot test

@@
 }
-\section*{Preface}\label{preface}
-\addcontentsline{toc}{section}{Preface}
-
-This is the online html version of my master thesis. The pdf version is
-available for download with the download button on top of the windows.
-
-Have a nice reading.
-
-\emph{Sylvain}
-
 \section*{Résumé et Abstract}\label{resume-et-abstract}
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/DOE-1.pdf}
+\caption{\label{fig:DOE}Experimental design before disturbance. Communities
+are implemented along a gradient of species richness (SR) and functional
+dispersion (FDis) resulting in a broad range of aboveground biomass
+(AGB). FDis was caluclated based on 4 functional traits (leaf mass per
+area, wood specific gravity, maximum diameter, maximum height).}
+\end{figure}
+
 \subsubsection{Outputs anlaysis ?}\label{outputs-anlaysis}

### 2017-05-30 02-Introduction.Rmd
Intro init

@@
 
-* Introduce why tropical rainforest are such a good study case
-* Define biodiversity 
-* Define ecosystem functionning
-* Define ecosystem services
-* Introduce tropical sylviculture
-* Define selective logging
-* Introduce biodiversity erosion in tropical rainforests ant its impact
-* Introduce Loreau partitionning
-* Introduce the interest of modelling
-* Introduce the idea to look at biodiversity as both a parameter and an output
+Sutainable forest management in the tropics (i.e.
+managed selective harvesting of timber) has been widely promoted internationnaly to combat tropical deforestation and degradation [@Zimmerman2012].
+Currently tropical logging accounts for one eight of global timber production [@Blaser2011] and is still increasing.
+Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cuttint cycle [@Martin2015].
 
-Tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
+On the other hand, tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
+Effectively tropical forests host over half of the Earth's biodiversity [@Scheffers2012].
+High biodiversity from tropical rainforests is the source of many ecosystem functions.
+Amongst others, tropical forests play a key role in biogeochemical cycles, including carbone storage [@Lewis2004].
+**Add insights into carbon storage role of tropical forest.** Ecosystem functions from tropical forests support numerous ecosystem services, such as timber production and climate regulation.
 
-Tropical forests are primary ecosystems in terms of biodiversit and carbon storage [@Lewis2004].
+But several authors argue that selective logging represents a major threat to biodiversity [@Carreno-Rocabado2012; @DeAvila2015; @Gibson2013; @Martin2015; @Zimmerman2012], challenging the sustainable definition from current selective logging.
+We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better syslvicultural practives in order to reach sustainability.
 
-High biodiversity from tropical rainforests is the source of many ecosystem functions, which support ecosystem services.
+The question of selective logging impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functionning [@Loreau2000].
+Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to logging disturbance.
+And forest ecosystem response to logging disturbance will directly modify ecosystem functionning in both short and long term.
+Consequently assessing selective logging effect on tropical forest linking biodiversity and ecosystem seems an obvious and promising way [@Loreau2010].
+**Paragraph to fully review !**
 
-Biodiversity and ecosystem functionning field is emmering [@Loreau2000].
+Negative short term impacts of selective logging have been assessed [@Carreno-Rocabado2012; @DeAvila2015;  but see @Martin2015].
+Much less is known about the long term impact [@Osazuwa-Peters2015].
+The main reason is the difficulty to conduct long term empirical study [but see @Herault2010], which can be completed by the use of forest simulators [@Huth2004; @Khler2004; @Ruger2008; @Tietjen2006].
+Individual-based models of forest dynamics present the perfect framework to develop such joint biodiversity-ecosystem approaches [@Li].
+Individual-based models describe forest 'patches' accumulating carbon through time, assessing tree growth within the patch, or releasing carbon through gap opening [@Bugmann2001].
+Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit **Cite Pacala et al.
+1996**.
+Recently, the forest growth simulator TROLL [@Chave1999], an individual-based and spatially explicit forest model, was developped to introduce recent advances in plant physiological community.
+TROLL model relates physiological processes to species-specific functional traits [@Li].
+Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functionning link response to logging disturbance.
 
-Selective logging is increasing in tropical forests.
-
-Tropical logging accounts for one eight of global timber production [@Blaser2011].
-
-Selective logging represents a major threat to biodiversity [@Gibson2013].
+**Major question greater diversity (taxonomic and functional) brought a better resilience to disturbance ?**

### 2017-05-30 03-Model.Rmd
Intro init

# Model description

## Overview

## Abiotic environment

## Photosynthesis

## Autotrpohic respiration

## Carbon uptake

## Tree growth

## Seed dispersion, production and recruitment

## Mortality

### 2017-05-30 04-Sensitivity.Rmd
Intro init

# Sensitivity analysis

## Functional traits

## Seed rain

### 2017-05-30 05-Disturbance.Rmd
Intro init

--- 2017-05-30 03-MnMs.Rmd
@@
-```{r config, message=FALSE, warning=FALSE, include=FALSE}
-rm(list = ls()) ; invisible(gc())
-library(knitr)
-library(parallel)
-library(TROLL)
-library(RconTroll)
-library(ggplot2)
-library(ggrepel)
-library(cowplot)
-library(entropart)
-library(plotly)
-library(readr)
-opts_chunk$set(
-  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
-    cache = T, cache.lazy = F)
-```
+# Disturbance
 
-# Materials and methods
+## Model description
 
-## TROLL simulator
-
-### Abiotic environment
-
-### Photosynthesis
-
-### Autotrpohic respiration
-
-### Carbon uptake
-
-### Tree growth
-
-### Seed dispersion, production and recruitment
-
-### Mortality
-
-## TROLL sensitivity analysis
-
-### Functional traits
-
-### Seed rain
-
-## Disturbance
-
-### Disturbance module
-
-### Design of experiment
+## Design of experiment
 
@@
 
-### Outputs anlaysis ?
+## Outputs anlaysis ?
 
-#### Resistance and resilience metrics
+### Resistance and resilience metrics
 
-#### Biodiversity partitioning
-
-## Selective logging
-
-### Selective logging module
-
-#### Designation
-
-#### Selection
-
-#### Rotten trees
-
-#### Felling
-
-#### Tracks
-
-#### Gap damages
-
-### Design of experiment
-
-### Outputs analysis ?
-
-#### Resistance and resilience metrics
-
-#### Biodiversity partitioning
+### Biodiversity partitioning

### 2017-05-30 docs/master-thesis.tex
Intro init

@@
 
+\usepackage{amsthm}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{definition}
+\newtheorem{example}{Example}[section]
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
 \begin{document}
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Introduce why tropical rainforest are such a good study case
-\item
-  Define biodiversity
-\item
-  Define ecosystem functionning
-\item
-  Define ecosystem services
-\item
-  Introduce tropical sylviculture
-\item
-  Define selective logging
-\item
-  Introduce biodiversity erosion in tropical rainforests ant its impact
-\item
-  Introduce Loreau partitionning
-\item
-  Introduce the interest of modelling
-\item
-  Introduce the idea to look at biodiversity as both a parameter and an
-  output
-\end{itemize}
+Sutainable forest management in the tropics (i.e.~managed selective
+harvesting of timber) has been widely promoted internationnaly to combat
+tropical deforestation and degradation \citep{Zimmerman2012}. Currently
+tropical logging accounts for one eight of global timber production
+\citep{Blaser2011} and is still increasing. Most tropical timber
+production originates from selective logging, the targeted harvesting of
+timber from commercial species in a single cuttint cycle
+\citep{Martin2015}.
 
-Tropical rainforests have fascinated ecologists due to their outstanding
-diversity \citep{connell_diversity_1978}.
+On the other hand, tropical rainforests have fascinated ecologists due
+to their outstanding diversity \citep{connell_diversity_1978}.
+Effectively tropical forests host over half of the Earth's biodiversity
+\citep{Scheffers2012}. High biodiversity from tropical rainforests is
+the source of many ecosystem functions. Amongst others, tropical forests
+play a key role in biogeochemical cycles, including carbone storage
+\citep{Lewis2004}. \textbf{Add insights into carbon storage role of
+tropical forest.} Ecosystem functions from tropical forests support
+numerous ecosystem services, such as timber production and climate
+regulation.
 
-Tropical forests are primary ecosystems in terms of biodiversit and
-carbon storage \citep{Lewis2004}.
+But several authors argue that selective logging represents a major
+threat to biodiversity
+\citep{Carreno-Rocabado2012, DeAvila2015, Gibson2013, Martin2015, Zimmerman2012},
+challenging the sustainable definition from current selective logging.
+We consequently need to assess both short and long term impacts of
+selective logging on tropical forest ecosystems to implement better
+syslvicultural practives in order to reach sustainability.
 
-High biodiversity from tropical rainforests is the source of many
-ecosystem functions, which support ecosystem services.
+The question of selective logging impact on tropical forest can be
+directly related to the emerging field of biodiversity and ecosystem
+functionning \citep{Loreau2000}. Tropical forest outstanding
+biodiversity will be both a factor and a result of forest ecosystem
+response to logging disturbance. And forest ecosystem response to
+logging disturbance will directly modify ecosystem functionning in both
+short and long term. Consequently assessing selective logging effect on
+tropical forest linking biodiversity and ecosystem seems an obvious and
+promising way \citep{Loreau2010}. \textbf{Paragraph to fully review !}
 
-Biodiversity and ecosystem functionning field is emmering
-\citep{Loreau2000}.
+Negative short term impacts of selective logging have been assessed
+\citetext{\citealp{Carreno-Rocabado2012}; \citealp{DeAvila2015}; \citealp[but
+see][]{Martin2015}}. Much less is known about the long term impact
+\citep{Osazuwa-Peters2015}. The main reason is the difficulty to conduct
+long term empirical study \citep[but see][]{Herault2010}, which can be
+completed by the use of forest simulators
+\citep{Huth2004, Khler2004, Ruger2008, Tietjen2006}. Individual-based
+models of forest dynamics present the perfect framework to develop such
+joint biodiversity-ecosystem approaches \citep{Li}. Individual-based
+models describe forest `patches' accumulating carbon through time,
+assessing tree growth within the patch, or releasing carbon through gap
+opening \citep{Bugmann2001}. Up to several dozens of different Plant
+Functional Types (PFTs) are generally defined and models can sometimes
+be fully spatially explicit \textbf{Cite Pacala et al. 1996}. Recently,
+the forest growth simulator TROLL \citep{Chave1999}, an individual-based
+and spatially explicit forest model, was developped to introduce recent
+advances in plant physiological community. TROLL model relates
+physiological processes to species-specific functional traits
+\citep{Li}. Consequently, TROLL model allow to simulate fully a
+neotropical forest biodiversity to study biodiversity-ecosystem
+functionning link response to logging disturbance.
 
-Selective logging is increasing in tropical forests.
+\textbf{Major question greater diversity (taxonomic and functional)
+brought a better resilience to disturbance ?}
 
-Tropical logging accounts for one eight of global timber production
-\citep{Blaser2011}.
+\section{Model description}\label{model-description}
 
-Selective logging represents a major threat to biodiversity
-\citep{Gibson2013}.
+\subsection{Overview}\label{overview}
 
-\section{Materials and methods}\label{materials-and-methods}
+\subsection{Abiotic environment}\label{abiotic-environment}
 
-\subsection{TROLL simulator}\label{troll-simulator}
+\subsection{Photosynthesis}\label{photosynthesis}
 
-\subsubsection{Abiotic environment}\label{abiotic-environment}
+\subsection{Autotrpohic respiration}\label{autotrpohic-respiration}
 
-\subsubsection{Photosynthesis}\label{photosynthesis}
+\subsection{Carbon uptake}\label{carbon-uptake}
 
-\subsubsection{Autotrpohic respiration}\label{autotrpohic-respiration}
+\subsection{Tree growth}\label{tree-growth}
 
-\subsubsection{Carbon uptake}\label{carbon-uptake}
-
-\subsubsection{Tree growth}\label{tree-growth}
-
-\subsubsection{Seed dispersion, production and
+\subsection{Seed dispersion, production and
 recruitment}\label{seed-dispersion-production-and-recruitment}
 
-\subsubsection{Mortality}\label{mortality}
+\subsection{Mortality}\label{mortality}
 
-\subsection{TROLL sensitivity
-analysis}\label{troll-sensitivity-analysis}
+\section{Sensitivity analysis}\label{sensitivity-analysis}
 
-\subsubsection{Functional traits}\label{functional-traits}
+\subsection{Functional traits}\label{functional-traits}
 
-\subsubsection{Seed rain}\label{seed-rain}
+\subsection{Seed rain}\label{seed-rain}
 
-\subsection{Disturbance}\label{disturbance}
+\section{Disturbance}\label{disturbance}
 
-\subsubsection{Disturbance module}\label{disturbance-module}
+\subsection{Model description}\label{model-description-1}
 
-\subsubsection{Design of experiment}\label{design-of-experiment}
+\subsection{Design of experiment}\label{design-of-experiment}
 
@@
 
-\subsubsection{Outputs anlaysis ?}\label{outputs-anlaysis}
+\subsection{Outputs anlaysis ?}\label{outputs-anlaysis}
 
-\paragraph{Resistance and resilience
+\subsubsection{Resistance and resilience
 metrics}\label{resistance-and-resilience-metrics}
 
-\paragraph{Biodiversity partitioning}\label{biodiversity-partitioning}
+\subsubsection{Biodiversity
+partitioning}\label{biodiversity-partitioning}
 
-\subsection{Selective logging}\label{selective-logging}
+\section{Selective logging}\label{selective-logging}
 
-\subsubsection{Selective logging module}\label{selective-logging-module}
+\subsection{Model description}\label{model-description-2}
 
-\paragraph{Designation}\label{designation}
+\subsubsection{Designation}\label{designation}
 
-\paragraph{Selection}\label{selection}
+\subsubsection{Selection}\label{selection}
 
-\paragraph{Rotten trees}\label{rotten-trees}
+\subsubsection{Rotten trees}\label{rotten-trees}
 
-\paragraph{Felling}\label{felling}
+\subsubsection{Felling}\label{felling}
 
-\paragraph{Tracks}\label{tracks}
+\subsubsection{Tracks}\label{tracks}
 
-\paragraph{Gap damages}\label{gap-damages}
+\subsubsection{Gap damages}\label{gap-damages}
 
-\subsubsection{Design of experiment}\label{design-of-experiment-1}
+\subsection{Design of experiment}\label{design-of-experiment-1}
 
-\subsubsection{Outputs analysis ?}\label{outputs-analysis}
+\subsection{Outputs analysis ?}\label{outputs-analysis}
 
-\paragraph{Resistance and resilience
+\subsubsection{Resistance and resilience
 metrics}\label{resistance-and-resilience-metrics-1}
 
-\paragraph{Biodiversity partitioning}\label{biodiversity-partitioning-1}
+\subsubsection{Biodiversity
+partitioning}\label{biodiversity-partitioning-1}
 
@@
 
-\bibliography{bibliography/library.bib}
+\bibliography{/home/sylvain/Documents/Bibliography/library.bib}
 

### 2017-05-30 index.Rmd
DOE plot test
Intro init

@@
 output: bookdown::gitbook
-csl: bibliography/mee.csl
-bibliography: bibliography/library.bib
+csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
+bibliography: /home/sylvain/Documents/Bibliography/library.bib
 link-citations: yes

### 2017-05-30 02-Introduction.Rmd
Model description: overview & abiotic environment

@@
 Individual-based models describe forest 'patches' accumulating carbon through time, assessing tree growth within the patch, or releasing carbon through gap opening [@Bugmann2001].
-Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit **Cite Pacala et al.
-1996**.
+Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit [@Pacala1996].
 Recently, the forest growth simulator TROLL [@Chave1999], an individual-based and spatially explicit forest model, was developped to introduce recent advances in plant physiological community.

### 2017-05-30 03-Model.Rmd
Model description: overview & abiotic environment

@@
 
+TROLL model each tree indivdually in a located environment.
+Thus TROLL model, alongside with SORTIE  [@Pacala1996; @Uriarte2009] and FORMIND [@Fischer2016; @Kohler1998], can be defined as an individual-based and spatially explicit forest growth model.
+TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production.
+Trees are growing in a located light environment explicitly computed witin voxels of 1 $m^3$.
+Each tree is consistently defined by its age, diameter at brest height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA) (see figure \@ref(fig:TROLLtree)).
+Tree geometry is calculated with allometric equations but leaf area vary dinamically within each crown following carbon allocations.
+Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels.
+Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
+A species label is associated to a number of species specific parameters (see table \@ref(tab:traits)) related to functional trait values which can be sampled on the field.
+
+```{r traits, echo=FALSE}
+table <- data.frame(
+  Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
+  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breasth height threshold', 'asymptotic height', 'parameter of the tree-height-dbh allometry'),
+  Units = c('$g.m^-2$', '$mg.g^-1$', '$mg.g^-1$', '$g.cm^-3$', '$m$', '$m$', '$m$')
+)
+knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.')
+```
+
+Carbon assimilation is computed over half-hourly period of a representative day.
+Then allocation is computed to simulate tree growth from an explicit carbone balance (in contrast to previous models).
+Finally environment is updated at each timestep set to one month.
+Seedlings are not simulated explicitly but as a pool.
+In addition belowground processes, herbaceous plants, epiphytes and lianas are not simulated inside TROLL.
+The source code is written in C++ and available upon request.
+All analyses were conducted in R version 3.4.0 **Cite R, add entry in Mendeley**.
+
+```{r TROLLtree, echo=FALSE, fig.cap='Individuals tree inside TROLL explicit spatial grid from @Li.
+Tree geometry (crown radius CR, crown depth CD, height h, diameter at breast height dbh) is updated at each timestep following allometric relationship with assimilated carbon allocated to growth.
+Each tree is flagged with a species label linking to its species-specific attributes.
+Light is compputed explicitly at each timestep for each voxel.'}
+knitr::include_graphics('images/TROLLtree.png', dpi = NA)
+```
+
 ## Abiotic environment
+
+A voxel space, with a resolution of 1 $m^3$, is used to explicilty model the abiotic environment.
+For each tree crown, leaf area density is calculated on tree geometry assuming a uniform distriution across voxels occupied by the crown.
+Leaf area density is computed within each voxel summing all tree crowns inside the voxel $v$, and is denoted $LAD(v)$ (leaf area per voxel in $m².m^-3$).
+The vertical sum of $LAD$ from voxel $v$ to the ground level defines $LAI(v)$ (leaf area per fround area in $m^2.m^-2$ commonly called leaf area index):
+
+\begin{equation}
+  LAI(v) = \sum _{v'=v} ^\infty LAD(v') 
+  (\#eq:LAI)
+\end{equation}
+
+Daily variations in light intensity (photosynthetic photon flux density PPFD in $\mu mol_{photons}.m^-2.s^-1$), temperature (T in degrees Celsius), and vapor pressure deficit (VPD in $kPA$) are computed to assess carbon assimilation within each voxel of the canopy and for a representative day per month (see Appendix 1 from @Li for further details).
+Variation of PPFD Within the canopy is calculated as a loacal Beer-Lambert extinction law:
+
+\begin{equation}
+  PPFD_{max,month}(v) = PPFD_{top,max,month}*e^{-k*LAI(v)}
+  (\#eq:PPFD)
+\end{equation}
+
+The daily maximum incident PPFD at the top of canopy $PPFD_{top,max,month}$ is given as input.
+The extinction rate $k$ is assumed as constant (see table \@ref(tab:par) for parameters value), besides is variation with zenith angle and species leaf inclination angle [@Meir2000].
+Moreover only vertical light diffusion is considered ignoring lateral light diffusion, which can have an important role especially in logging gaps.
+Finally, intra-day variation at half hour time steps $t$ for a representative day every month are used to compute $PPFD_{month}(v,t)$, $T_{month}(v,t)$ and $VPD_{month}(v,t)$.
+Water and nutrient process both in soil and inside trees are not simulated.
 
@@
 
-## Autotrpohic respiration
+## Autotrophic respiration
 
@@
 
-## Seed dispersion, production and recruitment
+## Recruitment
 

### 2017-05-30 docs/master-thesis.tex
Model description: overview & abiotic environment

@@
 Functional Types (PFTs) are generally defined and models can sometimes
-be fully spatially explicit \textbf{Cite Pacala et al. 1996}. Recently,
-the forest growth simulator TROLL \citep{Chave1999}, an individual-based
-and spatially explicit forest model, was developped to introduce recent
+be fully spatially explicit \citep{Pacala1996}. Recently, the forest
+growth simulator TROLL \citep{Chave1999}, an individual-based and
+spatially explicit forest model, was developped to introduce recent
 advances in plant physiological community. TROLL model relates
@@
 
+TROLL model each tree indivdually in a located environment. Thus TROLL
+model, alongside with SORTIE \citep{Pacala1996, Uriarte2009} and FORMIND
+\citep{Fischer2016, Kohler1998}, can be defined as an individual-based
+and spatially explicit forest growth model. TROLL simulates the life
+cycle of individual trees from recruitment, with a diameter at breast
+height (dbh) above 1 cm, to death with growth and seed production. Trees
+are growing in a located light environment explicitly computed witin
+voxels of 1 \(m^3\). Each tree is consistently defined by its age,
+diameter at brest height (dbh), height (h), crown radius (CR), crown
+depth (CD) and leaf area (LA) (see figure \ref{fig:TROLLtree}). Tree
+geometry is calculated with allometric equations but leaf area vary
+dinamically within each crown following carbon allocations. Voxels
+resolution of 1 \(m^3\) allow the establishment of maximum one tree by
+1x1 m pixels. Each tree is flagged with a species label inherited from
+the parent tree through the seedling recruitment. A species label is
+associated to a number of species specific parameters (see table
+\ref{tab:traits}) related to functional trait values which can be
+sampled on the field.
+
+\begin{table}
+
+\caption{\label{tab:traits}Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.}
+\centering
+\begin{tabular}[t]{l|l|l}
+\hline
+Abbreviation & Description & Units\\
+\hline
+\$LMA\$ & leaf mass per area & \$g.m\textasciicircum{}-2\$\\
+\hline
+\$N\_m\$ & leaf nitrogen content per dry mass & \$mg.g\textasciicircum{}-1\$\\
+\hline
+\$P\_m\$ & leaf phosphorous content per dry mass & \$mg.g\textasciicircum{}-1\$\\
+\hline
+\$wsg\$ & wood specific gravity & \$g.cm\textasciicircum{}-3\$\\
+\hline
+\$dbh\_\{thresh\}\$ & diameter at breasth height threshold & \$m\$\\
+\hline
+\$h\_\{lim\}\$ & asymptotic height & \$m\$\\
+\hline
+\$a\_h\$ & parameter of the tree-height-dbh allometry & \$m\$\\
+\hline
+\end{tabular}
+\end{table}
+
+Carbon assimilation is computed over half-hourly period of a
+representative day. Then allocation is computed to simulate tree growth
+from an explicit carbone balance (in contrast to previous models).
+Finally environment is updated at each timestep set to one month.
+Seedlings are not simulated explicitly but as a pool. In addition
+belowground processes, herbaceous plants, epiphytes and lianas are not
+simulated inside TROLL. The source code is written in C++ and available
+upon request. All analyses were conducted in R version 3.4.0
+\textbf{Cite R, add entry in Mendeley}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{images/TROLLtree.png}
+\caption{\label{fig:TROLLtree}Individuals tree inside TROLL explicit spatial
+grid from \citet{Li}. Tree geometry (crown radius CR, crown depth CD,
+height h, diameter at breast height dbh) is updated at each timestep
+following allometric relationship with assimilated carbon allocated to
+growth. Each tree is flagged with a species label linking to its
+species-specific attributes. Light is compputed explicitly at each
+timestep for each voxel.}
+\end{figure}
+
 \subsection{Abiotic environment}\label{abiotic-environment}
+
+A voxel space, with a resolution of 1 \(m^3\), is used to explicilty
+model the abiotic environment. For each tree crown, leaf area density is
+calculated on tree geometry assuming a uniform distriution across voxels
+occupied by the crown. Leaf area density is computed within each voxel
+summing all tree crowns inside the voxel \(v\), and is denoted
+\(LAD(v)\) (leaf area per voxel in \(m².m^-3\)). The vertical sum of
+\(LAD\) from voxel \(v\) to the ground level defines \(LAI(v)\) (leaf
+area per fround area in \(m^2.m^-2\) commonly called leaf area index):
+
+\begin{equation}
+  LAI(v) = \sum _{v'=v} ^\infty LAD(v') 
+  \label{eq:LAI}
+\end{equation}
+
+Daily variations in light intensity (photosynthetic photon flux density
+PPFD in \(\mu mol_{photons}.m^-2.s^-1\)), temperature (T in degrees
+Celsius), and vapor pressure deficit (VPD in \(kPA\)) are computed to
+assess carbon assimilation within each voxel of the canopy and for a
+representative day per month (see Appendix 1 from \citet{Li} for further
+details). Variation of PPFD Within the canopy is calculated as a loacal
+Beer-Lambert extinction law:
+
+\begin{equation}
+  PPFD_{max,month}(v) = PPFD_{top,max,month}*e^{-k*LAI(v)}
+  \label{eq:PPFD}
+\end{equation}
+
+The daily maximum incident PPFD at the top of canopy
+\(PPFD_{top,max,month}\) is given as input. The extinction rate \(k\) is
+assumed as constant (see table \ref{tab:par} for parameters value),
+besides is variation with zenith angle and species leaf inclination
+angle \citep{Meir2000}. Moreover only vertical light diffusion is
+considered ignoring lateral light diffusion, which can have an important
+role especially in logging gaps. Finally, intra-day variation at half
+hour time steps \(t\) for a representative day every month are used to
+compute \(PPFD_{month}(v,t)\), \(T_{month}(v,t)\) and
+\(VPD_{month}(v,t)\). Water and nutrient process both in soil and inside
+trees are not simulated.
 
@@
 
-\subsection{Autotrpohic respiration}\label{autotrpohic-respiration}
+\subsection{Autotrophic respiration}\label{autotrophic-respiration}
 
@@
 
-\subsection{Seed dispersion, production and
-recruitment}\label{seed-dispersion-production-and-recruitment}
+\subsection{Recruitment}\label{recruitment}
 

### 2017-05-31 03-Model.Rmd
03-Model v1

@@
   Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breasth height threshold', 'asymptotic height', 'parameter of the tree-height-dbh allometry'),
-  Units = c('$g.m^-2$', '$mg.g^-1$', '$mg.g^-1$', '$g.cm^-3$', '$m$', '$m$', '$m$')
+  Units = c('$g.m^{-2}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$g.cm^{-3}$', '$m$', '$m$', '$m$')
 )
@@
 For each tree crown, leaf area density is calculated on tree geometry assuming a uniform distriution across voxels occupied by the crown.
-Leaf area density is computed within each voxel summing all tree crowns inside the voxel $v$, and is denoted $LAD(v)$ (leaf area per voxel in $m².m^-3$).
-The vertical sum of $LAD$ from voxel $v$ to the ground level defines $LAI(v)$ (leaf area per fround area in $m^2.m^-2$ commonly called leaf area index):
+Leaf area density is computed within each voxel summing all tree crowns inside the voxel $v$, and is denoted $LAD(v)$ (leaf area per voxel in $m².m^{-3}$).
+The vertical sum of $LAD$ from voxel $v$ to the ground level defines $LAI(v)$ (leaf area per fround area in $m^2.m^{-2}$ commonly called leaf area index):
 
@@
 
-Daily variations in light intensity (photosynthetic photon flux density PPFD in $\mu mol_{photons}.m^-2.s^-1$), temperature (T in degrees Celsius), and vapor pressure deficit (VPD in $kPA$) are computed to assess carbon assimilation within each voxel of the canopy and for a representative day per month (see Appendix 1 from @Li for further details).
+Daily variations in light intensity (photosynthetic photon flux density PPFD in $\mu mol_{photons}.m^{-2}.s^{-1}$), temperature (T in degrees Celsius), and vapor pressure deficit (VPD in $kPA$) are computed to assess carbon assimilation within each voxel of the canopy and for a representative day per month (see Appendix 1 from @Li for further details).
 Variation of PPFD Within the canopy is calculated as a loacal Beer-Lambert extinction law:
@@
 The daily maximum incident PPFD at the top of canopy $PPFD_{top,max,month}$ is given as input.
-The extinction rate $k$ is assumed as constant (see table \@ref(tab:par) for parameters value), besides is variation with zenith angle and species leaf inclination angle [@Meir2000].
+The extinction rate $k$ is assumed as constant, besides is variation with zenith angle and species leaf inclination angle [@Meir2000].
 Moreover only vertical light diffusion is considered ignoring lateral light diffusion, which can have an important role especially in logging gaps.
@@
 
+### Theory
+
+Troll simulates the carbon uptake of each individual with the Farquhar, von Caemmerer and Berry model of C3 photosynthesis [@Farquhar1980].
+Gross carbon assimilation rate ($A$ in $\mu mol~CO_2.
+m^{-2}.s^{-1}$) will be the minimum of eiter Rubisco activity ($A_v$) or RuBP generation ($A_j$):
+
+\begin{equation}
+  A=min(A_v, A_j)~|~A_v=V_{cmax}*\frac{c_i-\Gamma^*}{c_i+K_m}~;~A_j=\frac{J}{4}*\frac{c_i-\Gamma^*}{c_i+2*\Gamma^*}
+  (\#eq:A)
+\end{equation}
+
+$V_{cmax}$ is the maximum rate of carboxylation ($\mu mol~CO_2.m^{-2}.s^{-1}$).
+$c_i$ is the $CO_2$ partial pressure at carboxylation sites.
+$\Gamma^*$ is the $CO_2$ compensation point in absence of dark respiration.
+$K_m$ is the apparent knietic constant of the Rubisco.
+And $J$ is the electron transport rate ($\mu mol e^-.m^{-2}.s^{-1}$).
+$J$ depends on the light intensity with $PPFD$:
+
+\begin{equation}
+  J = \frac{1}{2*\theta}*[\alpha*PPFD+J_{max}-\sqrt{(\alpha*PPFD+J_{max})^2}-4*\theta*\alpha*PPFD*J_{max}]
+  (\#eq:J)
+\end{equation}
+
+$J_{max}$ is the maximal electron transport capacity ($\mu mol e^-.m^{-2}.s^{-1}$).
+$\theta$ is the curvature factor.
+And $\alpha$ is the apparent quantum yield to electron transport ($mole^-.mol~photons^{-1}$).
+
+Carbon assimilation by photosynthesis will then be limited by the $CO_2$ partial pressure at carboxylation sites.
+Stomata controls the gas concentration at carboxylation sites throught stomatal transport:
+
+\begin{equation}
+  A = g_s*(c_a-c_i)
+  (\#eq:Ag)
+\end{equation}
+
+$g_s$ is the stomatal conductance to $CO_2$ ($molCO_2.m^{-2}.s^{-1}$). TROLL simulates stomatal conductance $g_s$ with the model from [@Medlyn2011]:
+
+\begin{equation}
+  g_s = g_0 + (1 + \frac{g_1}{\sqrt{VPD}})*\frac{A}{c_a}
+  (\#eq:gs)
+\end{equation}
+
+$g_0$ and $g_1$ are parameters from the model. TROLL model assume $g_0 \approx 0$ (empirically tested and considered as reasonable).
+
+### Parametrization
+
+Leaf traits can be used as proxy of photosynthesis, especially leaf nutrient content which directly play a role in it [@wright_worldwide_2004].
+@Domingues2010 suggested that $V_{cmac}$ and $J_{max}$ were both limited by the leaf concentration of nitrogen $N$ and phosphorus $P$ ($mg.g^{-1}$):
+
+\begin{equation}
+  log_{10} V_{cmax-M} = min( 
+  \begin{array}{c} 
+    -1.56+0.43*log_{10} N-0.37*log_{10} LMA \\
+    -0.80+0.45*log_{10} P-0.25*log_{10} LMA 
+  \end{array} 
+  )
+  (\#eq:VcmaxM)
+\end{equation}
+
+\begin{equation}
+  log_{10} J_{max-M} = min(
+  \begin{array}{c} 
+    -1.50+0.41*log_{10} N-0.45*log_{10} LMA \\
+    -0.74+0.44*log_{10} P-0.32*log_{10} LMA 
+  \end{array}
+  )
+  (\#eq:JmaxM)
+\end{equation}
+
+$V_{cmax-M}$ and $J_{max-M}$ are the photosynthetic capacities at $25^\circ C$ of mature leaves per leaf dry mass (resp.
+$\mu mol CO_2.g^-1.s^{-1}$ and $\mu mol e^-.g^{-1}.s^{-1}$).
+$LMA$ is the leaf mass per are ($g.cm^{-2}$).
+$V_{cmax}$ and $J_{max}$ are calculated by multiplying $V_{cmax-M}$ and $J_{max-M}$ by $LMA$.
+$V_{cmax}$ and $J_{max}$ variation with temperature are caluclated with @Bernacchi2003 (see Appendix 2 from @Li for further details).
+
+TROLL computes leaf carbon assimilation $A_l$ combining equations from \@ref(eq:A) to \@ref(eq:JmaxM) for each tree crown voxel within in each crown layer $l$:
+
+\begin{equation}
+  A_l = \frac{1}{n_v*t_M} * \sum_v  \sum^{t_M}_{t=1} A(PPFD_{month}(v,t),VPD_{month}(v,t),T_{month}(v,t))
+  (\#eq:Al)
+\end{equation}
+
+$PPFD_{month}(v,t)$, $VPD_{month}(v,t)$ , and $T_{month}(v,t)$ are derived from microclimatic data.
+$n_v$ is the number of voxels within crown layer $l$.
+And the sum is calculated over the $t_M$ half-hourly intervals $t$ of a tipical day.
+
 ## Autotrophic respiration
 
-## Carbon uptake
+A large fraction of plants carbon uptake is actually used for plant maintenance and growth respiration.
+The autotrophic respiration can represents up to 65% of the gross primary productivity but varies strongly among species, sites, and environnements.
+
+TROLL uses @Atkin2015 database of mature leaf dark respiration and associated leaf traits to compute leaf maintenance respiration:
+
+\begin{equation}
+  R_{leaf-M} = 8.5431-0.1306*N-0.5670*P-0.0137*LMA+11.1*V_{cmax-M}+0.1876*N*P
+  (\#eq:Rl)
+\end{equation}
+
+$R_{leaf-M}$ si the dark respiration rate per leaf dry mass at a temperaure of $25^\circ C$ ($nmolCO_2.g^{-1}.s^{-1}$).
+The other terms are in equations \@ref(eq:VcmaxM) and \@ref(eq:JmaxM).
+TROLL assume leaf respiration during day light to be 40% of leaf dark respiration, and computes total leaf respiration by accounting for the legnth of daylight.
+
+TROLL model stem respiration ($R_{stem}$ in $\mu molC.s^{-1}$) with a constant respiration rate per volume of sapwood:
+
+\begin{equation}
+  R_{stem} = 39.6*\pi*ST*(dbh-ST)*(h-CD)
+  (\#eq:Rs)
+\end{equation}
+
+dbh, h, CD and ST are tree diameter at breast height, height, corwn depth and sapwoond thickness, respectively ($m$). TROLL assumes $ST=0.04~m$ when $dbh>30~cm$ and an increasing $ST$ for lower $dbh$.
+
+Finally, TROLL computes both fine root maintenance respiration, as half of the leaf maintenance respiration.
+Whereas coarse root and branch maintenance respiration is computed as half of the stem respiration.
+And growth respiration ($R_{growth}$) is assumed to account for 25% of the gross primary productivity minus the sum of maintenance respirations.
+
+## Net carbon uptake
+
+Net primary production of carbon for one individual $NPP_{ind}$ ($gC$) is computed by the balance between gross primary production $GPP_{ind}$ and respirations $R$:
+
+\begin{equation}
+  NPP_{ind} = GPP_{ind} - R_{maintenance} - R_{growth}
+  (\#eq:NPP)
+\end{equation}
+
+TROLL partitions individuals total leaf area $LA$ into three pools for different leaf age classes corresponding to different photosynthesis efficiency (young, mature and old leaves with $LA_{young}$, $LA_{mature}$, and $LA_{old}$ respectivelly).
+Consequently we can compute growth primary production for one individual as:
+
+\begin{equation}
+  GPP_{ind} = 189.3 * \Delta t * \sum _{l= \lfloor h-CD \rfloor +1} ^{\lfloor h \rfloor} [A_l] * (\frac{LA_{young}}{2} + LA_{mature} + \frac{LA_{old}}{2})
+  (\#eq:GPP)
+\end{equation}
+
+h and CD are tree height and crown depth, repectivelly ($m$). $\lfloor x \rfloor$ is the rounding function. $\Delta t$ is the duration of a timestep ($year$).
+
+Thus, TROLL can compute carbon allocation to wood into an increment of stem volume $\Delta V$ ($m^3$):
+
+\begin{equation}
+  \Delta V = 10^{-6} * \frac{f_{wood}*NPP_{ind}}{0.5*wsg}*Senesc(dbh)
+  (\#eq:DeltaV)
+\end{equation}
+
+$f_{wood}$ is the fixed fraction of NPP allocated to stem and branches.
+$wsg$ is the wood specific gravity ($g.cm^{-3}$, see \@ref(tab:traits)).
+TROLL assume large trees less efficient to convert NPP as growth by using a size-related growth decline with function $Senesc$ after a specific diameter at brest height threshold $dbh_{thresh}$:
+
+\begin{equation}
+  Senesc(dbh) = max(0;3-2*\frac{dbh}{dbh_{thresh}})
+  (\#eq:Senesc)
+\end{equation}
+
+Finally, TROLL can compute carbon allocation to canopy with canopy NPP fraction denoted $f_{canopy}$ and decomposed into leaf, twig and fruit production.
+Carbon allocation to leaf results in a new young leaf pool, whereas other leaf pools are updated as follow:
+
+\begin{equation}
+  \begin{array}{c} \\
+   \Delta LA_{young} = \frac{2*f_{leaves}*NPP_{ind}}{LMA}-\frac{LA_{young}}{\tau_{young}} \\
+  \Delta LA_{mature} = \frac{LA_{young}}{\tau_{young}} - \frac{LA_{mature}}{\tau_{mature}}\\
+  \Delta LA_{old} = \frac{LA_{mature}}{\tau_{mature}} - \frac{LA_{old}}{\tau_{old}}
+  \end{array}
+  (\#eq:DeltaLA)
+\end{equation}
+
+$\tau_{young}$, $\tau_{mature}$, and $\tau_{old}$ are species residence times in each leaf pools ($years$).
+The sum of residency time thus defined the leaf lifespan $LL = \tau_{young} + \tau_{mature} + \tau_{old}$ ($years$).
+$\tau_{young}$ is set to one month and $\tau_{mature}$ is set to a third of leaf lifespan $LL$.
+Previous implementation of TROLL model used @Reich1991a allometry to infer leaf lifespan $LL$ from species leaf mass per area $LMA$ [@Li].
+But the use of the allometrie from @Reich1991a with current implementation of the TROLL model resulted in an underestimation of leaf lifespan for low LMA species.
+Consequently in the following paragraph we suggest a new allometry.
+Belowground carbon allocation is not simulated inside TROLL.
+
+## Leaf lifespan
+
+The underestimation of leaf lifespan for low LMA species with the allometry from @Reich1991a resulted in indivduals unealistic early death from carbon starvation.
+We gathered data from TRY[@Kattge2011], DRYAD [@chave_towards_2009] and GLOPNET [@wright_worldwide_2004] datasets.
+We used an out of the bag method applied on a random forest to select variables with highest importance to explain leaf lifespan.
+We thus selected leaf mass per are $LMA$, leaf nitrogen content $N$ and wood specific gravity $wsg$.
+We then used a bayesian approach to test different models with growing level of complexity.
+The model with the best tradeoff between complexity (number of parameters), convergence, likelihood, and prediction quality (root mean square error of prediction RMSEP) was kept.
+We selected following model with a maximum likelihood of 13.6 and a RMSEP of 12 months:
+
+\begin{equation}
+  LL_{d} \sim \mathcal{logN}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
+  (\#eq:LLb)
+\end{equation}
+
+Leaf lifespan $LL$ follows a lognormal law with location infered from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
+Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$.
+All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without presemption following a gamma law.
+$d$ represents the dataset random effects and encompass environmental and protocol variations.
+The sampling of model \@ref(eq:LLb) resulted in the following allometry:
+
+\begin{equation}
+  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
+  (\#eq:LL)
+\end{equation}
 
@@
 
+Once the increment of stem volume $\Delta V$ calculated with equation \@ref(eq:DeltaV), TROLL convert it into an increment of tree diameter at breast height denoted $\Delta dbh$.
+TROLL infer tree height from $dbh$ using a Michaelis-Menten equation:
+
+\begin{equation}
+  h = h_{lim}*\frac{dbh}{dbh + a_h}
+  (\#eq:h)
+\end{equation}
+
+On the other hand, we have the trunk volume $V = C * \pi * (\frac{dbh}{2})^2*h$, thus:
+
+\begin{equation}
+  \begin{array}{c} \\
+    \Delta V = C*\frac{1}{2}*\pi*h*dbh*\Delta dbh + C * \pi * (\frac{dbh}{2})^2*h \\
+    \Delta V = V*\frac{\Delta dbh}{dbh}*(3-\frac{dbh}{dbh + ah})
+  \end{array}
+  (\#eq:Deltadbh)
+\end{equation}
+
+Next, TROLL used the new trunk dimension ($dbh$ and $h$) to update tree crown geometry using allometric equations [@Chave2005]:
+
+\begin{equation}
+  \begin{array}{c} \\
+    CR = 0.80 + 10.47*dbh - 3.33*dbh^2\\
+    CD = -0.48 + 0.26*h~;~CD = 0.13 + 0.17*h~(h<5~m)
+  \end{array}
+  (\#eq:C)
+\end{equation}
+
+Finally, TROLL computes the mean leaf density within the crown ($LD$ in $m^2.m^{-3}$) assuming a uniform distribution:
+
+\begin{equation}
+  LD = \frac{LA_{young}+LA_{mature}+LA_{old}}{\pi*CR^2*CD}
+  (\#eq:LD)
+\end{equation}
+
+## Mortality
+
+Mortality is partitioned in three factors inside TROLL: background death $d_b$, treefall death $d_t$ and negative density dependent death $d_{NDD}$.
+Because density dependent death $d_{NDD}$ is still in development inside TROLL we did not used it, so we will not detail is computation.
+
+@chave_towards_2009 advocated for a wood economics spectrum opposing fast growing light wood species species with high risk of mortality to slow growing dense wood species with reduced risk of mortality.
+Hence, background mortality is derived from wood specific gravity $wsg$ inside TROLL:
+
+\begin{equation}
+  d_b = m*(1-\frac{wsg}{wsg_{lim}})+d_n
+  (\#eq:db)
+\end{equation}
+
+$m$ ($events.year^{-1}$) is the reference background death rate for lighter wood species (pioneers). $d_n$ represents death by carbohydrates shortage. If the number of consecutive day with $NPP_{ind}
+< 0$ \@ref(eq:NPP) is superior to tree leaf lifespan $d_n$ is set to 1 and remains null in other cases.
+
+Mortality by treefall inside TROLL depends on a specifric stochastic threshold $\theta$:
+
+\begin{equation}
+  \theta = h_{max}*(1-v_T*|\zeta|)
+  (\#eq:theta)
+\end{equation}
+
+$h_{max}$ is the maximal tree height.
+$v_T$ is the variance term set to 0.3.
+$|\zeta|$ is the absolute value of a random centered and scaled Gaussian.
+If the tree hieght $h$ is superior to $\theta$ then the tree may fall with a probability $1-\theta/h$ [@Chave1999].
+The treefall direction is random (drawn from a uniform law ($\mathcal{U}[0,2\pi]$).
+All tree in the trajectory of the falling tree will be hurted through a variable denoted $hurt$, incremented by fallen tree height $h$.
+If a tree height is inferior than its $hurt$ values then it may die with a probability $1-\frac{1}{2}\frac{h}{hurt}$.
+$hurt$ variable is reset to null at each timestep ($month$).
+
 ## Recruitment
 
-## Mortality
+Once the tree became fertile they will start to disperse seeds. TROLL consider tree as fertile after a specific height threshold $h_{mature}$ [@Wright2005]:
+
+\begin{equation}
+  h_{mature} = -11.47+0.90*h_{max}
+  (\#eq:hmature)
+\end{equation}
+
+But TROLL is not considering seed directly through a seedbank, instead seed might be interpreted as a seedling recruitment opportunity.
+The number of reproduction opportunities per mature tree is denoted $n_s$ and set to 10 for all species.
+This assumption originates from a trade-off between seed number and seed size resulting in equivalent survival and recruitment probability.
+All $n_s$ events are dispersed with a distance randomly drawn from a Gaussian distribution.
+Additionally, TROLL model consider external seedrain through $n_{ext}$ events of seed immigration:
+
+\begin{equation}
+  n_{ext} = N_{tot}*f_{reg}*n_{ha}
+  (\#eq:next)
+\end{equation}
+
+N_{tot} is the external seedrain per hectare (number of reproduction opportunities). $f_{reg}$ is the species regional frequency. $n_{ha}$ is the simulated plot size in $ha$.
+
+Finally, a bank of seedlings to be recruited is defined for each pixel. If the ground-level light reaches a species light compensation point $LCP$ the species will be recruited:
+
+\begin{equation}
+  LCP = \frac{R_{leaf}}{\phi}
+  (\#eq:LCP)
+\end{equation}
+
+$R_{leaf}$ is the leaf respiration for maintenance (see \@ref(eq:Rl)).
+$\phi$ is the quantum yield ($\mu mol C.\mu mol~photon$) set to 0.06.
+If several species reach their $LCP$, one is picked at random.
+Seedlings are recruited with following intial geometry:
+
+\begin{equation}
+  \begin{array}{c} \\
+    dbh = \frac{a_h}{h_{max} - 1}\\
+    h = 1~m\\
+    CR = 0.5~m\\
+    CD = 0.3~m\\
+    LD = 0.8~m^2.^{-3}
+  \end{array}
+  (\#eq:C)
+\end{equation}

### 2017-05-31 05-Disturbance.Rmd
03-Model v1

@@
+```{r config5, message=FALSE, warning=FALSE, include=FALSE}
+rm(list = ls()) ; invisible(gc())
+library(ggplot2)
+library(ggrepel)
+```
+
 # Disturbance
@@
 
-```{r DOE, fig.cap='Experimental design before disturbance.
+```{r DOE, echo=FALSE, fig.cap='Experimental design before disturbance.
 Communities are implemented along a gradient of species richness (SR) and functional dispersion (FDis) resulting in a broad range of aboveground biomass (AGB).
-FDis was caluclated based on 4 functional traits (leaf mass per area, wood specific gravity, maximum diameter, maximum height).'}
+FDis was caluclated based on 4 functional traits (leaf mass per area, wood specific gravity, maximum diameter, maximum height).', cache=TRUE}
 # Data prep

### 2017-05-31 docs/master-thesis.tex
03-Model v1

@@
   \postdate{\par}
-  \date{2017-05-30}
+  \date{2017-05-31}
 
@@
 \hline
-\$LMA\$ & leaf mass per area & \$g.m\textasciicircum{}-2\$\\
+\$LMA\$ & leaf mass per area & \$g.m\textasciicircum{}\{-2\}\$\\
 \hline
-\$N\_m\$ & leaf nitrogen content per dry mass & \$mg.g\textasciicircum{}-1\$\\
+\$N\_m\$ & leaf nitrogen content per dry mass & \$mg.g\textasciicircum{}\{-1\}\$\\
 \hline
-\$P\_m\$ & leaf phosphorous content per dry mass & \$mg.g\textasciicircum{}-1\$\\
+\$P\_m\$ & leaf phosphorous content per dry mass & \$mg.g\textasciicircum{}\{-1\}\$\\
 \hline
-\$wsg\$ & wood specific gravity & \$g.cm\textasciicircum{}-3\$\\
+\$wsg\$ & wood specific gravity & \$g.cm\textasciicircum{}\{-3\}\$\\
 \hline
@@
 summing all tree crowns inside the voxel \(v\), and is denoted
-\(LAD(v)\) (leaf area per voxel in \(m².m^-3\)). The vertical sum of
+\(LAD(v)\) (leaf area per voxel in \(m².m^{-3}\)). The vertical sum of
 \(LAD\) from voxel \(v\) to the ground level defines \(LAI(v)\) (leaf
-area per fround area in \(m^2.m^-2\) commonly called leaf area index):
+area per fround area in \(m^2.m^{-2}\) commonly called leaf area index):
 
@@
 Daily variations in light intensity (photosynthetic photon flux density
-PPFD in \(\mu mol_{photons}.m^-2.s^-1\)), temperature (T in degrees
+PPFD in \(\mu mol_{photons}.m^{-2}.s^{-1}\)), temperature (T in degrees
 Celsius), and vapor pressure deficit (VPD in \(kPA\)) are computed to
@@
 \(PPFD_{top,max,month}\) is given as input. The extinction rate \(k\) is
-assumed as constant (see table \ref{tab:par} for parameters value),
-besides is variation with zenith angle and species leaf inclination
-angle \citep{Meir2000}. Moreover only vertical light diffusion is
-considered ignoring lateral light diffusion, which can have an important
-role especially in logging gaps. Finally, intra-day variation at half
-hour time steps \(t\) for a representative day every month are used to
-compute \(PPFD_{month}(v,t)\), \(T_{month}(v,t)\) and
+assumed as constant, besides is variation with zenith angle and species
+leaf inclination angle \citep{Meir2000}. Moreover only vertical light
+diffusion is considered ignoring lateral light diffusion, which can have
+an important role especially in logging gaps. Finally, intra-day
+variation at half hour time steps \(t\) for a representative day every
+month are used to compute \(PPFD_{month}(v,t)\), \(T_{month}(v,t)\) and
 \(VPD_{month}(v,t)\). Water and nutrient process both in soil and inside
@@
 
+\subsubsection{Theory}\label{theory}
+
+Troll simulates the carbon uptake of each individual with the Farquhar,
+von Caemmerer and Berry model of C3 photosynthesis \citep{Farquhar1980}.
+Gross carbon assimilation rate (\(A\) in
+\(\mu mol~CO_2. m^{-2}.s^{-1}\)) will be the minimum of eiter Rubisco
+activity (\(A_v\)) or RuBP generation (\(A_j\)):
+
+\begin{equation}
+  A=min(A_v, A_j)~|~A_v=V_{cmax}*\frac{c_i-\Gamma^*}{c_i+K_m}~;~A_j=\frac{J}{4}*\frac{c_i-\Gamma^*}{c_i+2*\Gamma^*}
+  \label{eq:A}
+\end{equation}
+
+\(V_{cmax}\) is the maximum rate of carboxylation
+(\(\mu mol~CO_2.m^{-2}.s^{-1}\)). \(c_i\) is the \(CO_2\) partial
+pressure at carboxylation sites. \(\Gamma^*\) is the \(CO_2\)
+compensation point in absence of dark respiration. \(K_m\) is the
+apparent knietic constant of the Rubisco. And \(J\) is the electron
+transport rate (\(\mu mol e^-.m^{-2}.s^{-1}\)). \(J\) depends on the
+light intensity with \(PPFD\):
+
+\begin{equation}
+  J = \frac{1}{2*\theta}*[\alpha*PPFD+J_{max}-\sqrt{(\alpha*PPFD+J_{max})^2}-4*\theta*\alpha*PPFD*J_{max}]
+  \label{eq:J}
+\end{equation}
+
+\(J_{max}\) is the maximal electron transport capacity
+(\(\mu mol e^-.m^{-2}.s^{-1}\)). \(\theta\) is the curvature factor. And
+\(\alpha\) is the apparent quantum yield to electron transport
+(\(mole^-.mol~photons^{-1}\)).
+
+Carbon assimilation by photosynthesis will then be limited by the
+\(CO_2\) partial pressure at carboxylation sites. Stomata controls the
+gas concentration at carboxylation sites throught stomatal transport:
+
+\begin{equation}
+  A = g_s*(c_a-c_i)
+  \label{eq:Ag}
+\end{equation}
+
+\(g_s\) is the stomatal conductance to \(CO_2\)
+(\(molCO_2.m^{-2}.s^{-1}\)). TROLL simulates stomatal conductance
+\(g_s\) with the model from \citep{Medlyn2011}:
+
+\begin{equation}
+  g_s = g_0 + (1 + \frac{g_1}{\sqrt{VPD}})*\frac{A}{c_a}
+  \label{eq:gs}
+\end{equation}
+
+\(g_0\) and \(g_1\) are parameters from the model. TROLL model assume
+\(g_0 \approx 0\) (empirically tested and considered as reasonable).
+
+\subsubsection{Parametrization}\label{parametrization}
+
+Leaf traits can be used as proxy of photosynthesis, especially leaf
+nutrient content which directly play a role in it
+\citep{wright_worldwide_2004}. \citet{Domingues2010} suggested that
+\(V_{cmac}\) and \(J_{max}\) were both limited by the leaf concentration
+of nitrogen \(N\) and phosphorus \(P\) (\(mg.g^{-1}\)):
+
+\begin{equation}
+  log_{10} V_{cmax-M} = min( 
+  \begin{array}{c} 
+    -1.56+0.43*log_{10} N-0.37*log_{10} LMA \\
+    -0.80+0.45*log_{10} P-0.25*log_{10} LMA 
+  \end{array} 
+  )
+  \label{eq:VcmaxM}
+\end{equation}
+
+\begin{equation}
+  log_{10} J_{max-M} = min(
+  \begin{array}{c} 
+    -1.50+0.41*log_{10} N-0.45*log_{10} LMA \\
+    -0.74+0.44*log_{10} P-0.32*log_{10} LMA 
+  \end{array}
+  )
+  \label{eq:JmaxM}
+\end{equation}
+
+\(V_{cmax-M}\) and \(J_{max-M}\) are the photosynthetic capacities at
+\(25^\circ C\) of mature leaves per leaf dry mass (resp.
+\(\mu mol CO_2.g^-1.s^{-1}\) and \(\mu mol e^-.g^{-1}.s^{-1}\)). \(LMA\)
+is the leaf mass per are (\(g.cm^{-2}\)). \(V_{cmax}\) and \(J_{max}\)
+are calculated by multiplying \(V_{cmax-M}\) and \(J_{max-M}\) by
+\(LMA\). \(V_{cmax}\) and \(J_{max}\) variation with temperature are
+caluclated with \citet{Bernacchi2003} (see Appendix 2 from \citet{Li}
+for further details).
+
+TROLL computes leaf carbon assimilation \(A_l\) combining equations from
+\eqref{eq:A} to \eqref{eq:JmaxM} for each tree crown voxel within in each
+crown layer \(l\):
+
+\begin{equation}
+  A_l = \frac{1}{n_v*t_M} * \sum_v  \sum^{t_M}_{t=1} A(PPFD_{month}(v,t),VPD_{month}(v,t),T_{month}(v,t))
+  \label{eq:Al}
+\end{equation}
+
+\(PPFD_{month}(v,t)\), \(VPD_{month}(v,t)\) , and \(T_{month}(v,t)\) are
+derived from microclimatic data. \(n_v\) is the number of voxels within
+crown layer \(l\). And the sum is calculated over the \(t_M\)
+half-hourly intervals \(t\) of a tipical day.
+
 \subsection{Autotrophic respiration}\label{autotrophic-respiration}
 
-\subsection{Carbon uptake}\label{carbon-uptake}
+A large fraction of plants carbon uptake is actually used for plant
+maintenance and growth respiration. The autotrophic respiration can
+represents up to 65\% of the gross primary productivity but varies
+strongly among species, sites, and environnements.
+
+TROLL uses \citet{Atkin2015} database of mature leaf dark respiration
+and associated leaf traits to compute leaf maintenance respiration:
+
+\begin{equation}
+  R_{leaf-M} = 8.5431-0.1306*N-0.5670*P-0.0137*LMA+11.1*V_{cmax-M}+0.1876*N*P
+  \label{eq:Rl}
+\end{equation}
+
+\(R_{leaf-M}\) si the dark respiration rate per leaf dry mass at a
+temperaure of \(25^\circ C\) (\(nmolCO_2.g^{-1}.s^{-1}\)). The other
+terms are in equations \eqref{eq:VcmaxM} and \eqref{eq:JmaxM}. TROLL assume
+leaf respiration during day light to be 40\% of leaf dark respiration,
+and computes total leaf respiration by accounting for the legnth of
+daylight.
+
+TROLL model stem respiration (\(R_{stem}\) in \(\mu molC.s^{-1}\)) with
+a constant respiration rate per volume of sapwood:
+
+\begin{equation}
+  R_{stem} = 39.6*\pi*ST*(dbh-ST)*(h-CD)
+  \label{eq:Rs}
+\end{equation}
+
+dbh, h, CD and ST are tree diameter at breast height, height, corwn
+depth and sapwoond thickness, respectively (\(m\)). TROLL assumes
+\(ST=0.04~m\) when \(dbh>30~cm\) and an increasing \(ST\) for lower
+\(dbh\).
+
+Finally, TROLL computes both fine root maintenance respiration, as half
+of the leaf maintenance respiration. Whereas coarse root and branch
+maintenance respiration is computed as half of the stem respiration. And
+growth respiration (\(R_{growth}\)) is assumed to account for 25\% of
+the gross primary productivity minus the sum of maintenance
+respirations.
+
+\subsection{Net carbon uptake}\label{net-carbon-uptake}
+
+Net primary production of carbon for one individual \(NPP_{ind}\)
+(\(gC\)) is computed by the balance between gross primary production
+\(GPP_{ind}\) and respirations \(R\):
+
+\begin{equation}
+  NPP_{ind} = GPP_{ind} - R_{maintenance} - R_{growth}
+  \label{eq:NPP}
+\end{equation}
+
+TROLL partitions individuals total leaf area \(LA\) into three pools for
+different leaf age classes corresponding to different photosynthesis
+efficiency (young, mature and old leaves with \(LA_{young}\),
+\(LA_{mature}\), and \(LA_{old}\) respectivelly). Consequently we can
+compute growth primary production for one individual as:
+
+\begin{equation}
+  GPP_{ind} = 189.3 * \Delta t * \sum _{l= \lfloor h-CD \rfloor +1} ^{\lfloor h \rfloor} [A_l] * (\frac{LA_{young}}{2} + LA_{mature} + \frac{LA_{old}}{2})
+  \label{eq:GPP}
+\end{equation}
+
+h and CD are tree height and crown depth, repectivelly (\(m\)).
+\(\lfloor x \rfloor\) is the rounding function. \(\Delta t\) is the
+duration of a timestep (\(year\)).
+
+Thus, TROLL can compute carbon allocation to wood into an increment of
+stem volume \(\Delta V\) (\(m^3\)):
+
+\begin{equation}
+  \Delta V = 10^{-6} * \frac{f_{wood}*NPP_{ind}}{0.5*wsg}*Senesc(dbh)
+  \label{eq:DeltaV}
+\end{equation}
+
+\(f_{wood}\) is the fixed fraction of NPP allocated to stem and
+branches. \(wsg\) is the wood specific gravity (\(g.cm^{-3}\), see
+\ref{tab:traits}). TROLL assume large trees less efficient to convert
+NPP as growth by using a size-related growth decline with function
+\(Senesc\) after a specific diameter at brest height threshold
+\(dbh_{thresh}\):
+
+\begin{equation}
+  Senesc(dbh) = max(0;3-2*\frac{dbh}{dbh_{thresh}})
+  \label{eq:Senesc}
+\end{equation}
+
+Finally, TROLL can compute carbon allocation to canopy with canopy NPP
+fraction denoted \(f_{canopy}\) and decomposed into leaf, twig and fruit
+production. Carbon allocation to leaf results in a new young leaf pool,
+whereas other leaf pools are updated as follow:
+
+\begin{equation}
+  \begin{array}{c} \\
+   \Delta LA_{young} = \frac{2*f_{leaves}*NPP_{ind}}{LMA}-\frac{LA_{young}}{\tau_{young}} \\
+  \Delta LA_{mature} = \frac{LA_{young}}{\tau_{young}} - \frac{LA_{mature}}{\tau_{mature}}\\
+  \Delta LA_{old} = \frac{LA_{mature}}{\tau_{mature}} - \frac{LA_{old}}{\tau_{old}}
+  \end{array}
+  \label{eq:DeltaLA}
+\end{equation}
+
+\(\tau_{young}\), \(\tau_{mature}\), and \(\tau_{old}\) are species
+residence times in each leaf pools (\(years\)). The sum of residency
+time thus defined the leaf lifespan
+\(LL = \tau_{young} + \tau_{mature} + \tau_{old}\) (\(years\)).
+\(\tau_{young}\) is set to one month and \(\tau_{mature}\) is set to a
+third of leaf lifespan \(LL\). Previous implementation of TROLL model
+used \citet{Reich1991a} allometry to infer leaf lifespan \(LL\) from
+species leaf mass per area \(LMA\) \citep{Li}. But the use of the
+allometrie from \citet{Reich1991a} with current implementation of the
+TROLL model resulted in an underestimation of leaf lifespan for low LMA
+species. Consequently in the following paragraph we suggest a new
+allometry. Belowground carbon allocation is not simulated inside TROLL.
+
+\subsection{Leaf lifespan}\label{leaf-lifespan}
+
+The underestimation of leaf lifespan for low LMA species with the
+allometry from \citet{Reich1991a} resulted in indivduals unealistic
+early death from carbon starvation. We gathered data from
+TRY\citep{Kattge2011}, DRYAD \citep{chave_towards_2009} and GLOPNET
+\citep{wright_worldwide_2004} datasets. We used an out of the bag method
+applied on a random forest to select variables with highest importance
+to explain leaf lifespan. We thus selected leaf mass per are \(LMA\),
+leaf nitrogen content \(N\) and wood specific gravity \(wsg\). We then
+used a bayesian approach to test different models with growing level of
+complexity. The model with the best tradeoff between complexity (number
+of parameters), convergence, likelihood, and prediction quality (root
+mean square error of prediction RMSEP) was kept. We selected following
+model with a maximum likelihood of 13.6 and a RMSEP of 12 months:
+
+\begin{equation}
+  LL_{d} \sim \mathcal{logN}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
+  \label{eq:LLb}
+\end{equation}
+
+Leaf lifespan \(LL\) follows a lognormal law with location infered from
+leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
+\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
+normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
+\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
+presemption following a gamma law. \(d\) represents the dataset random
+effects and encompass environmental and protocol variations. The
+sampling of model \eqref{eq:LLb} resulted in the following allometry:
+
+\begin{equation}
+  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
+  \label{eq:LL}
+\end{equation}
 
@@
 
+Once the increment of stem volume \(\Delta V\) calculated with equation
+\eqref{eq:DeltaV}, TROLL convert it into an increment of tree diameter at
+breast height denoted \(\Delta dbh\). TROLL infer tree height from
+\(dbh\) using a Michaelis-Menten equation:
+
+\begin{equation}
+  h = h_{lim}*\frac{dbh}{dbh + a_h}
+  \label{eq:h}
+\end{equation}
+
+On the other hand, we have the trunk volume
+\(V = C * \pi * (\frac{dbh}{2})^2*h\), thus:
+
+\begin{equation}
+  \begin{array}{c} \\
+    \Delta V = C*\frac{1}{2}*\pi*h*dbh*\Delta dbh + C * \pi * (\frac{dbh}{2})^2*h \\
+    \Delta V = V*\frac{\Delta dbh}{dbh}*(3-\frac{dbh}{dbh + ah})
+  \end{array}
+  \label{eq:Deltadbh}
+\end{equation}
+
+Next, TROLL used the new trunk dimension (\(dbh\) and \(h\)) to update
+tree crown geometry using allometric equations \citep{Chave2005}:
+
+\begin{equation}
+  \begin{array}{c} \\
+    CR = 0.80 + 10.47*dbh - 3.33*dbh^2\\
+    CD = -0.48 + 0.26*h~;~CD = 0.13 + 0.17*h~(h<5~m)
+  \end{array}
+  \label{eq:C}
+\end{equation}
+
+Finally, TROLL computes the mean leaf density within the crown (\(LD\)
+in \(m^2.m^{-3}\)) assuming a uniform distribution:
+
+\begin{equation}
+  LD = \frac{LA_{young}+LA_{mature}+LA_{old}}{\pi*CR^2*CD}
+  \label{eq:LD}
+\end{equation}
+
+\subsection{Mortality}\label{mortality}
+
+Mortality is partitioned in three factors inside TROLL: background death
+\(d_b\), treefall death \(d_t\) and negative density dependent death
+\(d_{NDD}\). Because density dependent death \(d_{NDD}\) is still in
+development inside TROLL we did not used it, so we will not detail is
+computation.
+
+\citet{chave_towards_2009} advocated for a wood economics spectrum
+opposing fast growing light wood species species with high risk of
+mortality to slow growing dense wood species with reduced risk of
+mortality. Hence, background mortality is derived from wood specific
+gravity \(wsg\) inside TROLL:
+
+\begin{equation}
+  d_b = m*(1-\frac{wsg}{wsg_{lim}})+d_n
+  \label{eq:db}
+\end{equation}
+
+\(m\) (\(events.year^{-1}\)) is the reference background death rate for
+lighter wood species (pioneers). \(d_n\) represents death by
+carbohydrates shortage. If the number of consecutive day with
+\(NPP_{ind} < 0\) \eqref{eq:NPP} is superior to tree leaf lifespan \(d_n\)
+is set to 1 and remains null in other cases.
+
+Mortality by treefall inside TROLL depends on a specifric stochastic
+threshold \(\theta\):
+
+\begin{equation}
+  \theta = h_{max}*(1-v_T*|\zeta|)
+  \label{eq:theta}
+\end{equation}
+
+\(h_{max}\) is the maximal tree height. \(v_T\) is the variance term set
+to 0.3. \(|\zeta|\) is the absolute value of a random centered and
+scaled Gaussian. If the tree hieght \(h\) is superior to \(\theta\) then
+the tree may fall with a probability \(1-\theta/h\) \citep{Chave1999}.
+The treefall direction is random (drawn from a uniform law
+(\(\mathcal{U}[0,2\pi]\)). All tree in the trajectory of the falling
+tree will be hurted through a variable denoted \(hurt\), incremented by
+fallen tree height \(h\). If a tree height is inferior than its \(hurt\)
+values then it may die with a probability
+\(1-\frac{1}{2}\frac{h}{hurt}\). \(hurt\) variable is reset to null at
+each timestep (\(month\)).
+
 \subsection{Recruitment}\label{recruitment}
 
-\subsection{Mortality}\label{mortality}
+Once the tree became fertile they will start to disperse seeds. TROLL
+consider tree as fertile after a specific height threshold
+\(h_{mature}\) \citep{Wright2005}:
+
+\begin{equation}
+  h_{mature} = -11.47+0.90*h_{max}
+  \label{eq:hmature}
+\end{equation}
+
+But TROLL is not considering seed directly through a seedbank, instead
+seed might be interpreted as a seedling recruitment opportunity. The
+number of reproduction opportunities per mature tree is denoted \(n_s\)
+and set to 10 for all species. This assumption originates from a
+trade-off between seed number and seed size resulting in equivalent
+survival and recruitment probability. All \(n_s\) events are dispersed
+with a distance randomly drawn from a Gaussian distribution.
+Additionally, TROLL model consider external seedrain through \(n_{ext}\)
+events of seed immigration:
+
+\begin{equation}
+  n_{ext} = N_{tot}*f_{reg}*n_{ha}
+  \label{eq:next}
+\end{equation}
+
+N\_\{tot\} is the external seedrain per hectare (number of reproduction
+opportunities). \(f_{reg}\) is the species regional frequency.
+\(n_{ha}\) is the simulated plot size in \(ha\).
+
+Finally, a bank of seedlings to be recruited is defined for each pixel.
+If the ground-level light reaches a species light compensation point
+\(LCP\) the species will be recruited:
+
+\begin{equation}
+  LCP = \frac{R_{leaf}}{\phi}
+  \label{eq:LCP}
+\end{equation}
+
+\(R_{leaf}\) is the leaf respiration for maintenance (see \eqref{eq:Rl}).
+\(\phi\) is the quantum yield (\(\mu mol C.\mu mol~photon\)) set to
+0.06. If several species reach their \(LCP\), one is picked at random.
+Seedlings are recruited with following intial geometry:
+
+\begin{equation}
+  \begin{array}{c} \\
+    dbh = \frac{a_h}{h_{max} - 1}\\
+    h = 1~m\\
+    CR = 0.5~m\\
+    CD = 0.3~m\\
+    LD = 0.8~m^2.^{-3}
+  \end{array}
+  \label{eq:C}
+\end{equation}
 
@@
 
+\subsection{Sensitivity}\label{sensitivity}
+
+\subsection{Disturbance}\label{disturbance-1}
+
+\subsection{Sylviculture}\label{sylviculture}
+
 \section{Discussion}\label{discussion}

### 2017-06-01 04-Sensitivity.Rmd
Sensitivity 1 (+M&Ms -Res)

@@
 # Sensitivity analysis
+
+@Li already assessed TROLL model sensitivity to several parameters ($k$ see \@ref(eq:PPFD), $\phi$ see \@ref(eq:LCP), $g1$ see \@ref(eq:gs), $f_{wood}$ see \@ref(eq:DeltaV), $f_{canopy}$ see \@ref(eq:DeltaLA) and $m$ see \@ref(eq:db)) which they assumed having a key role in model functioning.
+
+On the other hand, we decided to use TROLL to study resistance and resilience of ecosystem face to disturbance, highlighting the role of biodiversit.
+Consequently we particulrarly needed to assess the importance of functional traits to further better control and evaluate functional diversities.
+We also needed to assess the sensitivity of TROLL model to the seed rain constant ($n_{ext}$, see \@ref(eq:next)) because we assume it is one of the main factor of tree recruitments after disturbance in simulations.
 
@@
 
+TROLL model currenty uses leaf mass per area ($LMA$ in $g.m^{-2}$), leaf nitrogen content per dry mass ($N_m$ in $mg.g^{-1}$), leaf phosphorus content per dry mass ($P_m$ in $mg.g^{-1}$), wood specific gravity ($wsg$ in $g.cm^{-3}$), diameter at breasth height threshold ($dbh_{thresh}$ in $m$), asymptotic height ($h_{lim}$ in $m$), and parameter of the tree-height-dbh allometry ($a_h$ in $m$).
+To assess the sensitivity of TROLL model to species functionnal traits, we performed a sensitivity analysis by fixing species trait values to their mean.
+Each trait was tested independently.
+We reduce to a common mean traits with a Perason's correlation value $r \geq 0.8$ ($h_{max}$ and $a_h$ with a correlation of $r=0.98$).
+
 ## Seed rain
+
+To assess the sensitivity of TROLL model to seed rain, we performed a sensitivity analysis by fixing simulations seed rain constant to 2, 20, 200 and 2000 seeds per hectare (default value being 200 seeds per hectare).
+
+Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and 188.9 GB of memory.
+We assumed maturity of the forest after 500 years of regeneration (Maréchaux & Chave) and computed simulation 100 years after a disturbance event with 40% loss of basal area.
+Due to computer limitations we did not run replicate (besides it should be necessary to reduce simulation stochasticity).
+To assess ecosystem outputs sensitivity to studied parameters, we compared it to 100 replicates of control simulations with all parameters set to default values.
+Ecosystem outputs outside of the range of the control replicates values are significantly influenced by the studied parameter.

### 2017-06-01 docs/master-thesis.tex
Sensitivity 1 (+M&Ms -Res)

@@
   \postdate{\par}
-  \date{2017-05-31}
+  \date{2017-06-01}
 
@@
 
+\citet{Li} already assessed TROLL model sensitivity to several
+parameters (\(k\) see \eqref{eq:PPFD}, \(\phi\) see \eqref{eq:LCP}, \(g1\)
+see \eqref{eq:gs}, \(f_{wood}\) see \eqref{eq:DeltaV}, \(f_{canopy}\) see
+\eqref{eq:DeltaLA} and \(m\) see \eqref{eq:db}) which they assumed having a
+key role in model functioning.
+
+On the other hand, we decided to use TROLL to study resistance and
+resilience of ecosystem face to disturbance, highlighting the role of
+biodiversit. Consequently we particulrarly needed to assess the
+importance of functional traits to further better control and evaluate
+functional diversities. We also needed to assess the sensitivity of
+TROLL model to the seed rain constant (\(n_{ext}\), see \eqref{eq:next})
+because we assume it is one of the main factor of tree recruitments
+after disturbance in simulations.
+
 \subsection{Functional traits}\label{functional-traits}
 
+TROLL model currenty uses leaf mass per area (\(LMA\) in \(g.m^{-2}\)),
+leaf nitrogen content per dry mass (\(N_m\) in \(mg.g^{-1}\)), leaf
+phosphorus content per dry mass (\(P_m\) in \(mg.g^{-1}\)), wood
+specific gravity (\(wsg\) in \(g.cm^{-3}\)), diameter at breasth height
+threshold (\(dbh_{thresh}\) in \(m\)), asymptotic height (\(h_{lim}\) in
+\(m\)), and parameter of the tree-height-dbh allometry (\(a_h\) in
+\(m\)). To assess the sensitivity of TROLL model to species functionnal
+traits, we performed a sensitivity analysis by fixing species trait
+values to their mean. Each trait was tested independently. We reduce to
+a common mean traits with a Perason's correlation value \(r \geq 0.8\)
+(\(h_{max}\) and \(a_h\) with a correlation of \(r=0.98\)).
+
 \subsection{Seed rain}\label{seed-rain}
+
+To assess the sensitivity of TROLL model to seed rain, we performed a
+sensitivity analysis by fixing simulations seed rain constant to 2, 20,
+200 and 2000 seeds per hectare (default value being 200 seeds per
+hectare).
+
+Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and
+188.9 GB of memory. We assumed maturity of the forest after 500 years of
+regeneration (Maréchaux \& Chave) and computed simulation 100 years
+after a disturbance event with 40\% loss of basal area. Due to computer
+limitations we did not run replicate (besides it should be necessary to
+reduce simulation stochasticity). To assess ecosystem outputs
+sensitivity to studied parameters, we compared it to 100 replicates of
+control simulations with all parameters set to default values. Ecosystem
+outputs outside of the range of the control replicates values are
+significantly influenced by the studied parameter.
 

### 2017-06-01 index.Rmd
disqus test

--- 2017-05-30 index.Rmd
@@
 --- 
-site: "bookdown::bookdown_site"
 title: "Master thesis"
@@
 date: "`r Sys.Date()`"
-output: bookdown::gitbook
-csl: bibliography/mee.csl
-bibliography: bibliography/library.bib
+site: bookdown::bookdown_site
+output: 
+  bookdown::gitbook
+csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
+bibliography: /home/sylvain/Documents/Bibliography/library.bib
 link-citations: yes

### 2017-06-01 05-Disturbance.Rmd
Disturbance v1

@@
 
+Most tropical timber production originates from selective logging, the targeted harvesting of timber from species of interest.
+Consequently, tropical sylviculture can be assimilated to a disturbance.
+The main difference between a disturbance and selective logging is the targetting of both species and individuals of interest.
+So we decided to first asses unselective disturbance effect on tropical forest ecosystem to subsequently better understand selective logging effect.
+
+We first implemented a disturbance module inside TROLL model to simulate unselective disturbance.
+60 tree communities were then defined with different levels of both specific and functional richness to explore the role of biodiversity on tropical forest ecosystem answer to disturbance.
+Next, different levels of disturbance were simulated on previous mature communities.
+Finally, resilience of ecosystem major global variables (related to carbon stock, forest dynamic and florisitic structure) from simulations were partitioned between selection effects and niche complementarity [@Loreau2001].
+
 ## Model description
 
+Disturbance module was designed in the simplest way in order to relate the ecosystem answer to volume lost without any individuals nor species targetting.
+For a given iteration $disturb_{iter}$, individuals are picked randomly with a uniform law on the number of trees.
+Selected individuals are then removed without trigerring a treefall to avoid any side effect.
+The operation is repeated untill the disturbance result in a defined lost basal area ($disturb_{intensity}$ in % of BA).
+
 ## Design of experiment
+
+In order to assess the role of biodiversity in ecosystem answer to disturbance, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time.
+Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%).
+Biodiversity was integrated with two components specific and functional diversities.
+We used species richness $SR$ to represents species diversity (5, 25, and 125).
+Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean.
+Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess theim before diturbance.
+Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$.
+For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means.
+Effectivelly, we did not wanted drastic change in community means that could have more effect than functional richness itself.
+This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]).
+Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the forest mature were simulated, and thus before disturbance.
+We obtained a broad range of both functionl dispersion $FDis$ and aboveground biomass $AGB$ for simulated forest ecosystems before disturbance.
 
@@
 
-## Outputs anlaysis ?
+## Ecosystem answer anlaysis
 
-### Resistance and resilience metrics
+Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions.
+We focused on few functions and related metrics to analyse ecosystem answer to disturbance: carbon stock with aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), forest dynamic with number of stem above $10~cm$ diameter at breast height ($N10$), and floristic composition with **?** (**?** in **?**).
 
-### Biodiversity partitioning
+The resilience of metrics values post disturbance were assessed through @Henry2012 formula:
+
+\begin{equation}
+  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X(t)}{X(t_d-1)}
+  (\#eq:Resilience)
+\end{equation}
+
+The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$.
+We transformed the equation in the resilience of the system $R(t)$ at the time $t$ being the ratio of the ecosystem metric $X(t)$ at time $t$ to the ecosystem metric before the disturbance happened $X(t_d-1)$ at time $t_d-1$.
+$X(t_d-1)$ was calculated as the mean of the ecosystem output for 50 last years of the simulation of the mature forest (over 600 years).
+Then, we calculated for each simulation the recovery time $t_{recovery}$ were the ecosystem reached back its stable state defined as its state before disturbance.
+But some ecosystem metrics can reach pre disturbance value wihtout revealing ecosystem recovery.
+For instance, number of stem above $10~cm$ diameter at breast height $N10$ will first decrease due to disturbance.
+Then $N10$ will exceed its initial value due to new seedlings recruitment before decreasing towards its pre disturbanec value.
+Consequently, the first time $N10$ reaches back its pre disturbance value can not be considered as the recovery time $t_{recovery}$.
+In order to solve this issue, we considered the ecosystem as stable again when its whole set of observed metrics reach their pre disturbance values.
+
+Biodiversity is not only a facet of the experimental design and an ecosystem output through florisitic composition, but also interact on ecosystem functioning and consequently on its answer to disturbance.
+Biodiversity ecosystem functioning relation can be split in complementarity and selection effect with @Loreau2001 partitioning:
+
+\begin{equation}
+  \begin{array}{c}
+    NE = X_O - X_E = CE + SE \\
+    CE = N* \overline{\Delta RX} \overline{M}\\
+    SE = N*cov(\Delta RX,M)
+  \end{array}
+  (\#eq:BiodivPart)
+\end{equation}
+
+Biodiversity net effect $NE$ is based on the difference between ecosystem variable $X$ observed value $X_O$ within the community mixture of species and its expected value $X_E$ if species performance were equal to their performance in monocultures.
+This effect can be partitioned between complementarity effect $CE$, representing niche partitionning, positive interactions, and resource supply, and selectvie effect $SE$ due to dominant species pool driving the ecosystem.
+Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:
+
+\begin{equation}
+  \Delta RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
+  (\#eq:DeltaRY)
+\end{equation}
+
+$X_sp$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$.
+$P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance.
+Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives).
+Whereas $SE$ become positive when dominant species outperform theimselves in mixture than in monoculture, and negative when less dominant species outperform theimselves in mixture than in monoculture [@Tobner2016].
+
+Recovery trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
+In order to do that, the design of experiment was repeated for each species indivdually representing 652 simulations of monoculture.

### 2017-06-01 docs/master-thesis.tex
disqus test
Disturbance v1

@@
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
+\usepackage[unicode=true]{hyperref}
 \PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{unicode=true,
+\hypersetup{
             pdftitle={Master thesis},
@@
 \fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-  \title{Master thesis}
-  \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \author{Sylvain SCHMITT}
-  \preauthor{\centering\large\emph}
-  \postauthor{\par}
-  \predate{\centering\large\emph}
-  \postdate{\par}
-  \date{2017-06-01}
-
 \usepackage{booktabs}
@@
 \makeatother
+
+\title{Master thesis}
+\author{Sylvain SCHMITT}
+\date{2017-06-01}
 
@@
 
-\date{}
+\thispagestyle{empty}
+\begin{center}
+\includegraphics{images/main.pdf}
+\end{center}
+
+\setlength{\abovedisplayskip}{-5pt}
+\setlength{\abovedisplayshortskip}{-5pt}
 
@@
 \begin{center}
-\includegraphics{images/main.pdf}
+\includegraphics{images/second.pdf}
 \end{center}

### 2017-06-01 docs/master-thesis.tex
Disturbance v1 build

@@
 
+Most tropical timber production originates from selective logging, the
+targeted harvesting of timber from species of interest. Consequently,
+tropical sylviculture can be assimilated to a disturbance. The main
+difference between a disturbance and selective logging is the targetting
+of both species and individuals of interest. So we decided to first
+asses unselective disturbance effect on tropical forest ecosystem to
+subsequently better understand selective logging effect.
+
+We first implemented a disturbance module inside TROLL model to simulate
+unselective disturbance. 60 tree communities were then defined with
+different levels of both specific and functional richness to explore the
+role of biodiversity on tropical forest ecosystem answer to disturbance.
+Next, different levels of disturbance were simulated on previous mature
+communities. Finally, resilience of ecosystem major global variables
+(related to carbon stock, forest dynamic and florisitic structure) from
+simulations were partitioned between selection effects and niche
+complementarity \citep{Loreau2001}.
+
 \subsection{Model description}\label{model-description-1}
 
+Disturbance module was designed in the simplest way in order to relate
+the ecosystem answer to volume lost without any individuals nor species
+targetting. For a given iteration \(disturb_{iter}\), individuals are
+picked randomly with a uniform law on the number of trees. Selected
+individuals are then removed without trigerring a treefall to avoid any
+side effect. The operation is repeated untill the disturbance result in
+a defined lost basal area (\(disturb_{intensity}\) in \% of BA).
+
 \subsection{Design of experiment}\label{design-of-experiment}
+
+In order to assess the role of biodiversity in ecosystem answer to
+disturbance, we needed to create a space of experiments encompassing
+both variation of disturbance, biodiversity and time. Disturbance was
+represented by percentage of basal area loss (0\%, 25\%, 50\% and 75\%).
+Biodiversity was integrated with two components specific and functional
+diversities. We used species richness \(SR\) to represents species
+diversity (5, 25, and 125). Functional diversity can be related to
+numerous components, and \citet{Borgy2017} argued for 5: richness,
+divergence, regularity, overlap and mean. Because mature forest were
+created from a bare soil with TROLL simulations, we could not control a
+priori divergence, regularity and overlap but only assess theim before
+diturbance. Consequently, we focused on functional richness with convex
+hull volume \(CHV\) and functional mean with community weighted mean
+\(CWM\). For each level of species richness \(SR\), we selected 20
+communities with growing convex hull volume \(CHV\) but with a community
+weighted means close to the regional species pool community weighted
+means. Effectivelly, we did not wanted drastic change in community means
+that could have more effect than functional richness itself. This design
+of experiments resulted in 60 communities (\(5~SR*20~CHV\)) and 240
+simulations (\(60 ~communities*4~levels~of~disturbance\)) over 600 years
+(maturity being assumed after 500 years of regeneration \citep{Li}).
+Figure \ref{fig:DOE} presents the design of experiment for communities
+biodiversity after the forest mature were simulated, and thus before
+disturbance. We obtained a broad range of both functionl dispersion
+\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
+before disturbance.
 
@@
 
-\subsection{Outputs anlaysis ?}\label{outputs-anlaysis}
+\subsection{Ecosystem answer anlaysis}\label{ecosystem-answer-anlaysis}
 
-\subsubsection{Resistance and resilience
-metrics}\label{resistance-and-resilience-metrics}
+Tropical forest ecosystems provides numerous ecosystem services linked
+to several ecosystem functions. We focused on few functions and related
+metrics to analyse ecosystem answer to disturbance: carbon stock with
+aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), forest dynamic with
+number of stem above \(10~cm\) diameter at breast height (\(N10\)), and
+floristic composition with \textbf{?} (\textbf{?} in \textbf{?}).
 
-\subsubsection{Biodiversity
-partitioning}\label{biodiversity-partitioning}
+The resilience of metrics values post disturbance were assessed through
+\citet{Henry2012} formula:
+
+\begin{equation}
+  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X(t)}{X(t_d-1)}
+  \label{eq:Resilience}
+\end{equation}
+
+The resilience of the system \(R(t)\) at the time \(t\) is described by
+the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
+\(Loss(t_d)\) at disturbance time \(t_d\). We transformed the equation
+in the resilience of the system \(R(t)\) at the time \(t\) being the
+ratio of the ecosystem metric \(X(t)\) at time \(t\) to the ecosystem
+metric before the disturbance happened \(X(t_d-1)\) at time \(t_d-1\).
+\(X(t_d-1)\) was calculated as the mean of the ecosystem output for 50
+last years of the simulation of the mature forest (over 600 years).
+Then, we calculated for each simulation the recovery time
+\(t_{recovery}\) were the ecosystem reached back its stable state
+defined as its state before disturbance. But some ecosystem metrics can
+reach pre disturbance value wihtout revealing ecosystem recovery. For
+instance, number of stem above \(10~cm\) diameter at breast height
+\(N10\) will first decrease due to disturbance. Then \(N10\) will exceed
+its initial value due to new seedlings recruitment before decreasing
+towards its pre disturbanec value. Consequently, the first time \(N10\)
+reaches back its pre disturbance value can not be considered as the
+recovery time \(t_{recovery}\). In order to solve this issue, we
+considered the ecosystem as stable again when its whole set of observed
+metrics reach their pre disturbance values.
+
+Biodiversity is not only a facet of the experimental design and an
+ecosystem output through florisitic composition, but also interact on
+ecosystem functioning and consequently on its answer to disturbance.
+Biodiversity ecosystem functioning relation can be split in
+complementarity and selection effect with \citet{Loreau2001}
+partitioning:
+
+\begin{equation}
+  \begin{array}{c}
+    NE = X_O - X_E = CE + SE \\
+    CE = N* \overline{\Delta RX} \overline{M}\\
+    SE = N*cov(\Delta RX,M)
+  \end{array}
+  \label{eq:BiodivPart}
+\end{equation}
+
+Biodiversity net effect \(NE\) is based on the difference between
+ecosystem variable \(X\) observed value \(X_O\) within the community
+mixture of species and its expected value \(X_E\) if species performance
+were equal to their performance in monocultures. This effect can be
+partitioned between complementarity effect \(CE\), representing niche
+partitionning, positive interactions, and resource supply, and selectvie
+effect \(SE\) due to dominant species pool driving the ecosystem. Both
+metrics depend on the variation of relative ecosystem variable
+\(\Delta RX\):
+
+\begin{equation}
+  \Delta RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
+  \label{eq:DeltaRY}
+\end{equation}
+
+\(X_sp\) is the ecosystem variable value for one species either in
+mixture \(X_{sp}(mixture)\) or in monoculture \(X_{sp}(monoculture)\).
+\(P_{sp}\) is the proportion of the species in the mixture represented
+by species relative abundance. Consequently, \(CE\) averages diversity
+effects of all species presents in the mixture (both negatives and
+positives). Whereas \(SE\) become positive when dominant species
+outperform theimselves in mixture than in monoculture, and negative when
+less dominant species outperform theimselves in mixture than in
+monoculture \citep{Tobner2016}.
+
+Recovery trajectories of ecosystem variable after disturbance were
+partitioned between complementarity effect \(CE\) and selection effect
+\(SE\). In order to do that, the design of experiment was repeated for
+each species indivdually representing 652 simulations of monoculture.
 
@@
 \subsubsection{Resistance and resilience
-metrics}\label{resistance-and-resilience-metrics-1}
+metrics}\label{resistance-and-resilience-metrics}
 
 \subsubsection{Biodiversity
-partitioning}\label{biodiversity-partitioning-1}
+partitioning}\label{biodiversity-partitioning}
 

### 2017-06-01 docs/master-thesis.tex
template.tex v1

@@
+% model from https://github.com/jgm/pandoc-templates/blob/master/default.latex
+
+%%% STYLE %%%
 \documentclass[]{article}
+
+%%% PACKAGES %%%
+
+% fonts
 \usepackage{lmodern}
+
+% pdf
+\usepackage{pdfpages}
+
+% formulae
 \usepackage{amssymb,amsmath}
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{fixltx2e}
 \usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+
+% hyperlinks
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
 \IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
+\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
 \usepackage[unicode=true]{hyperref}
   \PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
+\definecolor{maroon}{cmyk}{0, 0.87, 0.68, 0.32}
 \hypersetup{
-            pdftitle={Master thesis},
-            pdfauthor={Sylvain SCHMITT},
       colorlinks=true,
-            linkcolor=Maroon,
-            citecolor=Blue,
-            urlcolor=Blue,
+    linkcolor=maroon,
+    citecolor=blue,
+    urlcolor=blue,
       breaklinks=true}
 \urlstyle{same} % don't use monospace font for urls
+
+% Bibliography
 \usepackage{natbib}
 \bibliographystyle{plainnat}
+
+% Tables
 \usepackage{longtable,booktabs}
+% Fix footnotes in tables (requires footnote package)
+\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
+
+% graphics
 \usepackage{graphicx,grffile}
@@
 \makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+
+% indent
 \IfFileExists{parskip.sty}{%
@@
 }
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
+
+% prevent overfull lines
+\setlength{\emergencystretch}{3em}  
 \providecommand{\tightlist}{%
 \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
+
+\setcounter{secnumdepth}{0}
+
+% paragraphs
 % Redefines (sub)paragraphs to behave more like sections
@@
 \fi
-\usepackage{booktabs}
-\usepackage{amsthm}
+
+% set default figure placement to htbp
 \makeatletter 
-\def\thm@space@setup{%
-  \thm@preskip=8pt plus 2pt minus 4pt
-  \thm@postskip=\thm@preskip
-}
+\def\fps@figure{htbp}
 \makeatother
 
-\title{Master thesis}
-\author{Sylvain SCHMITT}
-\date{2017-06-01}
-
+%%% BODY %%%
 \usepackage{amsthm}
@@
 \begin{document}
-\maketitle
 
-\thispagestyle{empty}
-\begin{center}
-\includegraphics{images/main.pdf}
-\end{center}
+% first pages (to wirte manually ?)
+\includepdf{images/main.pdf}
+\includepdf{images/second.pdf}
 
-\setlength{\abovedisplayskip}{-5pt}
-\setlength{\abovedisplayshortskip}{-5pt}
+\tableofcontents
 
-\thispagestyle{empty}
-\begin{center}
-\includegraphics{images/second.pdf}
-\end{center}
-
-\setlength{\abovedisplayskip}{-5pt}
-\setlength{\abovedisplayshortskip}{-5pt}
-
-{
-\hypersetup{linkcolor=black}
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
 \section*{Résumé et Abstract}\label{resume-et-abstract}
@@
 
+Once the unselective effect of disturbance on forest ecosystem studied,
+we can focus on selective logging to highlight the effect of species and
+individuals targetting on the ecosystem answer.
+
 \subsection{Model description}\label{model-description-2}
@@
 
-\subsection{Outputs analysis ?}\label{outputs-analysis}
+\subsection{Ecosystem answer
+anlaysis}\label{ecosystem-answer-anlaysis-1}
 
-\subsubsection{Resistance and resilience
-metrics}\label{resistance-and-resilience-metrics}
-
-\subsubsection{Biodiversity
-partitioning}\label{biodiversity-partitioning}
+Similar to disturbance
 
@@
 
+\listoftables
+\listoffigures
+
+\renewcommand\refname{}
 \bibliography{/home/sylvain/Documents/Bibliography/library.bib}

### 2017-06-02 03-Model.Rmd
template tex v2

@@
 )
-knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.')
+knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.', format = 'pandoc')
 ```
@@
 \begin{equation}
-  LL_{d} \sim \mathcal{logN}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
+  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
   (\#eq:LLb)

### 2017-06-02 docs/master-thesis.tex
template tex v2

@@
-% model from https://github.com/jgm/pandoc-templates/blob/master/default.latex
+% template adapted from https://github.com/jgm/pandoc-templates/blob/master/default.latex
 
@@
 % paragraphs
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+\usepackage{titlesec}
+\let\oldsection\section
+\renewcommand\section{\newpage\oldsection}
+\titleformat{\section}
+{\huge\center\scshape}{\thesection}{1em}{}[{\titlerule[0.8pt]}]
+\titleformat*{\subsection}{\LARGE\bfseries}
+\titleformat*{\subsubsection}{\Large\bfseries}
+\titleformat*{\paragraph}{\large\bfseries\itshape}
+\titleformat*{\subparagraph}{\normalsize\itshape}
 
@@
 
-% first pages (to wirte manually ?)
-\includepdf{images/main.pdf}
-\includepdf{images/second.pdf}
+% First pages
+  % First page
+  \includegraphics{images/logos}
+
+  \begin{center}
+    \vspace*{\stretch{1}}
+    \LARGE{\textbf{Mémoire de stage}} \\
+    \vspace*{\fill}
+    \large{présenté par} \\
+    \large{Sylvain SCHMITT} \\
+    \vspace*{\fill}
+    \large{pour obtenir le diplôme national de master} \\
+    \large{mention Biodiversité, écologie, évolution} \\
+    \small{parcours Biodiversité végétale et gestion des écosystèmes tropicaux (BIOGET)} \\
+    \vspace*{\fill}
+    \large{Sujet :} \\
+    \Large{\textbf{A COMPLETER}} \\
+    \vspace*{\fill}
+    \large{soutenu publiquement le XX.xxxx.201X} \\
+    \large{à Kourou} \\
+    \vspace*{\fill}
+    \large{devant le jury suivant :} \\
+    \vspace*{\fill}
+    Dr Bruno HÉRAULT  \emph{Tuteur de stage} \\
+    Titre Prénom NOM  \emph{Examinateur} \\
+    Titre Prénom NOM  \emph{Examinateur} \\
+    Dr Stéphane TRAISSAC  \emph{Enseignant-réferent} \\
+    \vspace*{\stretch{1}}
+  \end{center}
+
+  % Second page
+  \newpage
+  \vspace*{\fill}
+  \emph{Les opinions émises par les auteurs sont personnelles et n'engagent pas AgroParisTech.}
+  \newpage
 
@@
 
-\begin{table}
-
-\caption{\label{tab:traits}Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.}
-\centering
-\begin{tabular}[t]{l|l|l}
-\hline
-Abbreviation & Description & Units\\
-\hline
-\$LMA\$ & leaf mass per area & \$g.m\textasciicircum{}\{-2\}\$\\
-\hline
-\$N\_m\$ & leaf nitrogen content per dry mass & \$mg.g\textasciicircum{}\{-1\}\$\\
-\hline
-\$P\_m\$ & leaf phosphorous content per dry mass & \$mg.g\textasciicircum{}\{-1\}\$\\
-\hline
-\$wsg\$ & wood specific gravity & \$g.cm\textasciicircum{}\{-3\}\$\\
-\hline
-\$dbh\_\{thresh\}\$ & diameter at breasth height threshold & \$m\$\\
-\hline
-\$h\_\{lim\}\$ & asymptotic height & \$m\$\\
-\hline
-\$a\_h\$ & parameter of the tree-height-dbh allometry & \$m\$\\
-\hline
-\end{tabular}
-\end{table}
+\begin{longtable}[]{@{}lll@{}}
+\caption{\label{tab:traits}Species-specific parameters used in TROLL from
+\citet{Li}. Data originates from the BRIDGE \citep{Baraloto2010} and TRY
+\citep{Kattge2011} datasets.}\tabularnewline
+\toprule
+Abbreviation & Description & Units\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Abbreviation & Description & Units\tabularnewline
+\midrule
+\endhead
+\(LMA\) & leaf mass per area & \(g.m^{-2}\)\tabularnewline
+\(N_m\) & leaf nitrogen content per dry mass &
+\(mg.g^{-1}\)\tabularnewline
+\(P_m\) & leaf phosphorous content per dry mass &
+\(mg.g^{-1}\)\tabularnewline
+\(wsg\) & wood specific gravity & \(g.cm^{-3}\)\tabularnewline
+\(dbh_{thresh}\) & diameter at breasth height threshold &
+\(m\)\tabularnewline
+\(h_{lim}\) & asymptotic height & \(m\)\tabularnewline
+\(a_h\) & parameter of the tree-height-dbh allometry &
+\(m\)\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 \begin{equation}
-  LL_{d} \sim \mathcal{logN}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
+  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
   \label{eq:LLb}
@@
 
+\bibliography{/home/sylvain/Documents/Bibliography/library.bib}
 \listoftables
@@
 
-\renewcommand\refname{}
-\bibliography{/home/sylvain/Documents/Bibliography/library.bib}
+% Last pages
+  %Last page
+  \newpage
+  \paragraph{Résumé :}
+  Écrire le résumé ici...
+  \paragraph{Mots clés :} mots clés
+  \newline\newline
+  \paragraph{Abstract:}
+  Write abstract here
+  \paragraph{Keywords:} keywords
+
+  \vspace*{\fill}
+  \includegraphics{images/logo}
 

### 2017-06-02 latex/after_body.tex
template tex v2

--- 2017-06-01 docs/master-thesis.tex
@@
-% model from https://github.com/jgm/pandoc-templates/blob/master/default.latex
+%Last page
+\newpage
+\paragraph{Résumé :}
+Écrire le résumé ici...
+\paragraph{Mots clés :} mots clés
+\newline\newline
+\paragraph{Abstract:}
+Write abstract here
+\paragraph{Keywords:} keywords
 
-%%% STYLE %%%
-\documentclass[]{article}
-
-%%% PACKAGES %%%
-
-% fonts
-\usepackage{lmodern}
-
-% pdf
-\usepackage{pdfpages}
-
-% formulae
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-% hyperlinks
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
-\usepackage[unicode=true]{hyperref}
-  \PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\definecolor{maroon}{cmyk}{0, 0.87, 0.68, 0.32}
-\hypersetup{
-      colorlinks=true,
-    linkcolor=maroon,
-    citecolor=blue,
-    urlcolor=blue,
-      breaklinks=true}
-\urlstyle{same} % don't use monospace font for urls
-
-% Bibliography
-\usepackage{natbib}
-\bibliographystyle{plainnat}
-
-% Tables
-\usepackage{longtable,booktabs}
-% Fix footnotes in tables (requires footnote package)
-\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
-
-% graphics
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-
-% indent
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-
-% prevent overfull lines
-\setlength{\emergencystretch}{3em}  
-\providecommand{\tightlist}{%
-\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\setcounter{secnumdepth}{0}
-
-% paragraphs
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-% set default figure placement to htbp
-\makeatletter 
-\def\fps@figure{htbp}
-\makeatother
-
-%%% BODY %%%
-\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{proposition}{Proposition}[section]
-\theoremstyle{definition}
-\newtheorem{example}{Example}[section]
-\theoremstyle{remark}
-\newtheorem*{remark}{Remark}
-\begin{document}
-
-% first pages (to wirte manually ?)
-\includepdf{images/main.pdf}
-\includepdf{images/second.pdf}
-
-\tableofcontents
-
-\section*{Résumé et Abstract}\label{resume-et-abstract}
-\addcontentsline{toc}{section}{Résumé et Abstract}
-
-Écrire le résumé français ici\ldots{}
-
-Write the english abstract here\ldots{}
-
-\section*{Acknowledgments}\label{acknowledgments}
-\addcontentsline{toc}{section}{Acknowledgments}
-
-I would like to thank\ldots{}
-
-\section*{Introduction}\label{introduction}
-\addcontentsline{toc}{section}{Introduction}
-
-Sutainable forest management in the tropics (i.e.~managed selective
-harvesting of timber) has been widely promoted internationnaly to combat
-tropical deforestation and degradation \citep{Zimmerman2012}. Currently
-tropical logging accounts for one eight of global timber production
-\citep{Blaser2011} and is still increasing. Most tropical timber
-production originates from selective logging, the targeted harvesting of
-timber from commercial species in a single cuttint cycle
-\citep{Martin2015}.
-
-On the other hand, tropical rainforests have fascinated ecologists due
-to their outstanding diversity \citep{connell_diversity_1978}.
-Effectively tropical forests host over half of the Earth's biodiversity
-\citep{Scheffers2012}. High biodiversity from tropical rainforests is
-the source of many ecosystem functions. Amongst others, tropical forests
-play a key role in biogeochemical cycles, including carbone storage
-\citep{Lewis2004}. \textbf{Add insights into carbon storage role of
-tropical forest.} Ecosystem functions from tropical forests support
-numerous ecosystem services, such as timber production and climate
-regulation.
-
-But several authors argue that selective logging represents a major
-threat to biodiversity
-\citep{Carreno-Rocabado2012, DeAvila2015, Gibson2013, Martin2015, Zimmerman2012},
-challenging the sustainable definition from current selective logging.
-We consequently need to assess both short and long term impacts of
-selective logging on tropical forest ecosystems to implement better
-syslvicultural practives in order to reach sustainability.
-
-The question of selective logging impact on tropical forest can be
-directly related to the emerging field of biodiversity and ecosystem
-functionning \citep{Loreau2000}. Tropical forest outstanding
-biodiversity will be both a factor and a result of forest ecosystem
-response to logging disturbance. And forest ecosystem response to
-logging disturbance will directly modify ecosystem functionning in both
-short and long term. Consequently assessing selective logging effect on
-tropical forest linking biodiversity and ecosystem seems an obvious and
-promising way \citep{Loreau2010}. \textbf{Paragraph to fully review !}
-
-Negative short term impacts of selective logging have been assessed
-\citetext{\citealp{Carreno-Rocabado2012}; \citealp{DeAvila2015}; \citealp[but
-see][]{Martin2015}}. Much less is known about the long term impact
-\citep{Osazuwa-Peters2015}. The main reason is the difficulty to conduct
-long term empirical study \citep[but see][]{Herault2010}, which can be
-completed by the use of forest simulators
-\citep{Huth2004, Khler2004, Ruger2008, Tietjen2006}. Individual-based
-models of forest dynamics present the perfect framework to develop such
-joint biodiversity-ecosystem approaches \citep{Li}. Individual-based
-models describe forest `patches' accumulating carbon through time,
-assessing tree growth within the patch, or releasing carbon through gap
-opening \citep{Bugmann2001}. Up to several dozens of different Plant
-Functional Types (PFTs) are generally defined and models can sometimes
-be fully spatially explicit \citep{Pacala1996}. Recently, the forest
-growth simulator TROLL \citep{Chave1999}, an individual-based and
-spatially explicit forest model, was developped to introduce recent
-advances in plant physiological community. TROLL model relates
-physiological processes to species-specific functional traits
-\citep{Li}. Consequently, TROLL model allow to simulate fully a
-neotropical forest biodiversity to study biodiversity-ecosystem
-functionning link response to logging disturbance.
-
-\textbf{Major question greater diversity (taxonomic and functional)
-brought a better resilience to disturbance ?}
-
-\section{Model description}\label{model-description}
-
-\subsection{Overview}\label{overview}
-
-TROLL model each tree indivdually in a located environment. Thus TROLL
-model, alongside with SORTIE \citep{Pacala1996, Uriarte2009} and FORMIND
-\citep{Fischer2016, Kohler1998}, can be defined as an individual-based
-and spatially explicit forest growth model. TROLL simulates the life
-cycle of individual trees from recruitment, with a diameter at breast
-height (dbh) above 1 cm, to death with growth and seed production. Trees
-are growing in a located light environment explicitly computed witin
-voxels of 1 \(m^3\). Each tree is consistently defined by its age,
-diameter at brest height (dbh), height (h), crown radius (CR), crown
-depth (CD) and leaf area (LA) (see figure \ref{fig:TROLLtree}). Tree
-geometry is calculated with allometric equations but leaf area vary
-dinamically within each crown following carbon allocations. Voxels
-resolution of 1 \(m^3\) allow the establishment of maximum one tree by
-1x1 m pixels. Each tree is flagged with a species label inherited from
-the parent tree through the seedling recruitment. A species label is
-associated to a number of species specific parameters (see table
-\ref{tab:traits}) related to functional trait values which can be
-sampled on the field.
-
-\begin{table}
-
-\caption{\label{tab:traits}Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.}
-\centering
-\begin{tabular}[t]{l|l|l}
-\hline
-Abbreviation & Description & Units\\
-\hline
-\$LMA\$ & leaf mass per area & \$g.m\textasciicircum{}\{-2\}\$\\
-\hline
-\$N\_m\$ & leaf nitrogen content per dry mass & \$mg.g\textasciicircum{}\{-1\}\$\\
-\hline
-\$P\_m\$ & leaf phosphorous content per dry mass & \$mg.g\textasciicircum{}\{-1\}\$\\
-\hline
-\$wsg\$ & wood specific gravity & \$g.cm\textasciicircum{}\{-3\}\$\\
-\hline
-\$dbh\_\{thresh\}\$ & diameter at breasth height threshold & \$m\$\\
-\hline
-\$h\_\{lim\}\$ & asymptotic height & \$m\$\\
-\hline
-\$a\_h\$ & parameter of the tree-height-dbh allometry & \$m\$\\
-\hline
-\end{tabular}
-\end{table}
-
-Carbon assimilation is computed over half-hourly period of a
-representative day. Then allocation is computed to simulate tree growth
-from an explicit carbone balance (in contrast to previous models).
-Finally environment is updated at each timestep set to one month.
-Seedlings are not simulated explicitly but as a pool. In addition
-belowground processes, herbaceous plants, epiphytes and lianas are not
-simulated inside TROLL. The source code is written in C++ and available
-upon request. All analyses were conducted in R version 3.4.0
-\textbf{Cite R, add entry in Mendeley}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{images/TROLLtree.png}
-\caption{\label{fig:TROLLtree}Individuals tree inside TROLL explicit spatial
-grid from \citet{Li}. Tree geometry (crown radius CR, crown depth CD,
-height h, diameter at breast height dbh) is updated at each timestep
-following allometric relationship with assimilated carbon allocated to
-growth. Each tree is flagged with a species label linking to its
-species-specific attributes. Light is compputed explicitly at each
-timestep for each voxel.}
-\end{figure}
-
-\subsection{Abiotic environment}\label{abiotic-environment}
-
-A voxel space, with a resolution of 1 \(m^3\), is used to explicilty
-model the abiotic environment. For each tree crown, leaf area density is
-calculated on tree geometry assuming a uniform distriution across voxels
-occupied by the crown. Leaf area density is computed within each voxel
-summing all tree crowns inside the voxel \(v\), and is denoted
-\(LAD(v)\) (leaf area per voxel in \(m².m^{-3}\)). The vertical sum of
-\(LAD\) from voxel \(v\) to the ground level defines \(LAI(v)\) (leaf
-area per fround area in \(m^2.m^{-2}\) commonly called leaf area index):
-
-\begin{equation}
-  LAI(v) = \sum _{v'=v} ^\infty LAD(v') 
-  \label{eq:LAI}
-\end{equation}
-
-Daily variations in light intensity (photosynthetic photon flux density
-PPFD in \(\mu mol_{photons}.m^{-2}.s^{-1}\)), temperature (T in degrees
-Celsius), and vapor pressure deficit (VPD in \(kPA\)) are computed to
-assess carbon assimilation within each voxel of the canopy and for a
-representative day per month (see Appendix 1 from \citet{Li} for further
-details). Variation of PPFD Within the canopy is calculated as a loacal
-Beer-Lambert extinction law:
-
-\begin{equation}
-  PPFD_{max,month}(v) = PPFD_{top,max,month}*e^{-k*LAI(v)}
-  \label{eq:PPFD}
-\end{equation}
-
-The daily maximum incident PPFD at the top of canopy
-\(PPFD_{top,max,month}\) is given as input. The extinction rate \(k\) is
-assumed as constant, besides is variation with zenith angle and species
-leaf inclination angle \citep{Meir2000}. Moreover only vertical light
-diffusion is considered ignoring lateral light diffusion, which can have
-an important role especially in logging gaps. Finally, intra-day
-variation at half hour time steps \(t\) for a representative day every
-month are used to compute \(PPFD_{month}(v,t)\), \(T_{month}(v,t)\) and
-\(VPD_{month}(v,t)\). Water and nutrient process both in soil and inside
-trees are not simulated.
-
-\subsection{Photosynthesis}\label{photosynthesis}
-
-\subsubsection{Theory}\label{theory}
-
-Troll simulates the carbon uptake of each individual with the Farquhar,
-von Caemmerer and Berry model of C3 photosynthesis \citep{Farquhar1980}.
-Gross carbon assimilation rate (\(A\) in
-\(\mu mol~CO_2. m^{-2}.s^{-1}\)) will be the minimum of eiter Rubisco
-activity (\(A_v\)) or RuBP generation (\(A_j\)):
-
-\begin{equation}
-  A=min(A_v, A_j)~|~A_v=V_{cmax}*\frac{c_i-\Gamma^*}{c_i+K_m}~;~A_j=\frac{J}{4}*\frac{c_i-\Gamma^*}{c_i+2*\Gamma^*}
-  \label{eq:A}
-\end{equation}
-
-\(V_{cmax}\) is the maximum rate of carboxylation
-(\(\mu mol~CO_2.m^{-2}.s^{-1}\)). \(c_i\) is the \(CO_2\) partial
-pressure at carboxylation sites. \(\Gamma^*\) is the \(CO_2\)
-compensation point in absence of dark respiration. \(K_m\) is the
-apparent knietic constant of the Rubisco. And \(J\) is the electron
-transport rate (\(\mu mol e^-.m^{-2}.s^{-1}\)). \(J\) depends on the
-light intensity with \(PPFD\):
-
-\begin{equation}
-  J = \frac{1}{2*\theta}*[\alpha*PPFD+J_{max}-\sqrt{(\alpha*PPFD+J_{max})^2}-4*\theta*\alpha*PPFD*J_{max}]
-  \label{eq:J}
-\end{equation}
-
-\(J_{max}\) is the maximal electron transport capacity
-(\(\mu mol e^-.m^{-2}.s^{-1}\)). \(\theta\) is the curvature factor. And
-\(\alpha\) is the apparent quantum yield to electron transport
-(\(mole^-.mol~photons^{-1}\)).
-
-Carbon assimilation by photosynthesis will then be limited by the
-\(CO_2\) partial pressure at carboxylation sites. Stomata controls the
-gas concentration at carboxylation sites throught stomatal transport:
-
-\begin{equation}
-  A = g_s*(c_a-c_i)
-  \label{eq:Ag}
-\end{equation}
-
-\(g_s\) is the stomatal conductance to \(CO_2\)
-(\(molCO_2.m^{-2}.s^{-1}\)). TROLL simulates stomatal conductance
-\(g_s\) with the model from \citep{Medlyn2011}:
-
-\begin{equation}
-  g_s = g_0 + (1 + \frac{g_1}{\sqrt{VPD}})*\frac{A}{c_a}
-  \label{eq:gs}
-\end{equation}
-
-\(g_0\) and \(g_1\) are parameters from the model. TROLL model assume
-\(g_0 \approx 0\) (empirically tested and considered as reasonable).
-
-\subsubsection{Parametrization}\label{parametrization}
-
-Leaf traits can be used as proxy of photosynthesis, especially leaf
-nutrient content which directly play a role in it
-\citep{wright_worldwide_2004}. \citet{Domingues2010} suggested that
-\(V_{cmac}\) and \(J_{max}\) were both limited by the leaf concentration
-of nitrogen \(N\) and phosphorus \(P\) (\(mg.g^{-1}\)):
-
-\begin{equation}
-  log_{10} V_{cmax-M} = min( 
-  \begin{array}{c} 
-    -1.56+0.43*log_{10} N-0.37*log_{10} LMA \\
-    -0.80+0.45*log_{10} P-0.25*log_{10} LMA 
-  \end{array} 
-  )
-  \label{eq:VcmaxM}
-\end{equation}
-
-\begin{equation}
-  log_{10} J_{max-M} = min(
-  \begin{array}{c} 
-    -1.50+0.41*log_{10} N-0.45*log_{10} LMA \\
-    -0.74+0.44*log_{10} P-0.32*log_{10} LMA 
-  \end{array}
-  )
-  \label{eq:JmaxM}
-\end{equation}
-
-\(V_{cmax-M}\) and \(J_{max-M}\) are the photosynthetic capacities at
-\(25^\circ C\) of mature leaves per leaf dry mass (resp.
-\(\mu mol CO_2.g^-1.s^{-1}\) and \(\mu mol e^-.g^{-1}.s^{-1}\)). \(LMA\)
-is the leaf mass per are (\(g.cm^{-2}\)). \(V_{cmax}\) and \(J_{max}\)
-are calculated by multiplying \(V_{cmax-M}\) and \(J_{max-M}\) by
-\(LMA\). \(V_{cmax}\) and \(J_{max}\) variation with temperature are
-caluclated with \citet{Bernacchi2003} (see Appendix 2 from \citet{Li}
-for further details).
-
-TROLL computes leaf carbon assimilation \(A_l\) combining equations from
-\eqref{eq:A} to \eqref{eq:JmaxM} for each tree crown voxel within in each
-crown layer \(l\):
-
-\begin{equation}
-  A_l = \frac{1}{n_v*t_M} * \sum_v  \sum^{t_M}_{t=1} A(PPFD_{month}(v,t),VPD_{month}(v,t),T_{month}(v,t))
-  \label{eq:Al}
-\end{equation}
-
-\(PPFD_{month}(v,t)\), \(VPD_{month}(v,t)\) , and \(T_{month}(v,t)\) are
-derived from microclimatic data. \(n_v\) is the number of voxels within
-crown layer \(l\). And the sum is calculated over the \(t_M\)
-half-hourly intervals \(t\) of a tipical day.
-
-\subsection{Autotrophic respiration}\label{autotrophic-respiration}
-
-A large fraction of plants carbon uptake is actually used for plant
-maintenance and growth respiration. The autotrophic respiration can
-represents up to 65\% of the gross primary productivity but varies
-strongly among species, sites, and environnements.
-
-TROLL uses \citet{Atkin2015} database of mature leaf dark respiration
-and associated leaf traits to compute leaf maintenance respiration:
-
-\begin{equation}
-  R_{leaf-M} = 8.5431-0.1306*N-0.5670*P-0.0137*LMA+11.1*V_{cmax-M}+0.1876*N*P
-  \label{eq:Rl}
-\end{equation}
-
-\(R_{leaf-M}\) si the dark respiration rate per leaf dry mass at a
-temperaure of \(25^\circ C\) (\(nmolCO_2.g^{-1}.s^{-1}\)). The other
-terms are in equations \eqref{eq:VcmaxM} and \eqref{eq:JmaxM}. TROLL assume
-leaf respiration during day light to be 40\% of leaf dark respiration,
-and computes total leaf respiration by accounting for the legnth of
-daylight.
-
-TROLL model stem respiration (\(R_{stem}\) in \(\mu molC.s^{-1}\)) with
-a constant respiration rate per volume of sapwood:
-
-\begin{equation}
-  R_{stem} = 39.6*\pi*ST*(dbh-ST)*(h-CD)
-  \label{eq:Rs}
-\end{equation}
-
-dbh, h, CD and ST are tree diameter at breast height, height, corwn
-depth and sapwoond thickness, respectively (\(m\)). TROLL assumes
-\(ST=0.04~m\) when \(dbh>30~cm\) and an increasing \(ST\) for lower
-\(dbh\).
-
-Finally, TROLL computes both fine root maintenance respiration, as half
-of the leaf maintenance respiration. Whereas coarse root and branch
-maintenance respiration is computed as half of the stem respiration. And
-growth respiration (\(R_{growth}\)) is assumed to account for 25\% of
-the gross primary productivity minus the sum of maintenance
-respirations.
-
-\subsection{Net carbon uptake}\label{net-carbon-uptake}
-
-Net primary production of carbon for one individual \(NPP_{ind}\)
-(\(gC\)) is computed by the balance between gross primary production
-\(GPP_{ind}\) and respirations \(R\):
-
-\begin{equation}
-  NPP_{ind} = GPP_{ind} - R_{maintenance} - R_{growth}
-  \label{eq:NPP}
-\end{equation}
-
-TROLL partitions individuals total leaf area \(LA\) into three pools for
-different leaf age classes corresponding to different photosynthesis
-efficiency (young, mature and old leaves with \(LA_{young}\),
-\(LA_{mature}\), and \(LA_{old}\) respectivelly). Consequently we can
-compute growth primary production for one individual as:
-
-\begin{equation}
-  GPP_{ind} = 189.3 * \Delta t * \sum _{l= \lfloor h-CD \rfloor +1} ^{\lfloor h \rfloor} [A_l] * (\frac{LA_{young}}{2} + LA_{mature} + \frac{LA_{old}}{2})
-  \label{eq:GPP}
-\end{equation}
-
-h and CD are tree height and crown depth, repectivelly (\(m\)).
-\(\lfloor x \rfloor\) is the rounding function. \(\Delta t\) is the
-duration of a timestep (\(year\)).
-
-Thus, TROLL can compute carbon allocation to wood into an increment of
-stem volume \(\Delta V\) (\(m^3\)):
-
-\begin{equation}
-  \Delta V = 10^{-6} * \frac{f_{wood}*NPP_{ind}}{0.5*wsg}*Senesc(dbh)
-  \label{eq:DeltaV}
-\end{equation}
-
-\(f_{wood}\) is the fixed fraction of NPP allocated to stem and
-branches. \(wsg\) is the wood specific gravity (\(g.cm^{-3}\), see
-\ref{tab:traits}). TROLL assume large trees less efficient to convert
-NPP as growth by using a size-related growth decline with function
-\(Senesc\) after a specific diameter at brest height threshold
-\(dbh_{thresh}\):
-
-\begin{equation}
-  Senesc(dbh) = max(0;3-2*\frac{dbh}{dbh_{thresh}})
-  \label{eq:Senesc}
-\end{equation}
-
-Finally, TROLL can compute carbon allocation to canopy with canopy NPP
-fraction denoted \(f_{canopy}\) and decomposed into leaf, twig and fruit
-production. Carbon allocation to leaf results in a new young leaf pool,
-whereas other leaf pools are updated as follow:
-
-\begin{equation}
-  \begin{array}{c} \\
-   \Delta LA_{young} = \frac{2*f_{leaves}*NPP_{ind}}{LMA}-\frac{LA_{young}}{\tau_{young}} \\
-  \Delta LA_{mature} = \frac{LA_{young}}{\tau_{young}} - \frac{LA_{mature}}{\tau_{mature}}\\
-  \Delta LA_{old} = \frac{LA_{mature}}{\tau_{mature}} - \frac{LA_{old}}{\tau_{old}}
-  \end{array}
-  \label{eq:DeltaLA}
-\end{equation}
-
-\(\tau_{young}\), \(\tau_{mature}\), and \(\tau_{old}\) are species
-residence times in each leaf pools (\(years\)). The sum of residency
-time thus defined the leaf lifespan
-\(LL = \tau_{young} + \tau_{mature} + \tau_{old}\) (\(years\)).
-\(\tau_{young}\) is set to one month and \(\tau_{mature}\) is set to a
-third of leaf lifespan \(LL\). Previous implementation of TROLL model
-used \citet{Reich1991a} allometry to infer leaf lifespan \(LL\) from
-species leaf mass per area \(LMA\) \citep{Li}. But the use of the
-allometrie from \citet{Reich1991a} with current implementation of the
-TROLL model resulted in an underestimation of leaf lifespan for low LMA
-species. Consequently in the following paragraph we suggest a new
-allometry. Belowground carbon allocation is not simulated inside TROLL.
-
-\subsection{Leaf lifespan}\label{leaf-lifespan}
-
-The underestimation of leaf lifespan for low LMA species with the
-allometry from \citet{Reich1991a} resulted in indivduals unealistic
-early death from carbon starvation. We gathered data from
-TRY\citep{Kattge2011}, DRYAD \citep{chave_towards_2009} and GLOPNET
-\citep{wright_worldwide_2004} datasets. We used an out of the bag method
-applied on a random forest to select variables with highest importance
-to explain leaf lifespan. We thus selected leaf mass per are \(LMA\),
-leaf nitrogen content \(N\) and wood specific gravity \(wsg\). We then
-used a bayesian approach to test different models with growing level of
-complexity. The model with the best tradeoff between complexity (number
-of parameters), convergence, likelihood, and prediction quality (root
-mean square error of prediction RMSEP) was kept. We selected following
-model with a maximum likelihood of 13.6 and a RMSEP of 12 months:
-
-\begin{equation}
-  LL_{d} \sim \mathcal{logN}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
-  \label{eq:LLb}
-\end{equation}
-
-Leaf lifespan \(LL\) follows a lognormal law with location infered from
-leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
-\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
-normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
-\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
-presemption following a gamma law. \(d\) represents the dataset random
-effects and encompass environmental and protocol variations. The
-sampling of model \eqref{eq:LLb} resulted in the following allometry:
-
-\begin{equation}
-  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
-  \label{eq:LL}
-\end{equation}
-
-\subsection{Tree growth}\label{tree-growth}
-
-Once the increment of stem volume \(\Delta V\) calculated with equation
-\eqref{eq:DeltaV}, TROLL convert it into an increment of tree diameter at
-breast height denoted \(\Delta dbh\). TROLL infer tree height from
-\(dbh\) using a Michaelis-Menten equation:
-
-\begin{equation}
-  h = h_{lim}*\frac{dbh}{dbh + a_h}
-  \label{eq:h}
-\end{equation}
-
-On the other hand, we have the trunk volume
-\(V = C * \pi * (\frac{dbh}{2})^2*h\), thus:
-
-\begin{equation}
-  \begin{array}{c} \\
-    \Delta V = C*\frac{1}{2}*\pi*h*dbh*\Delta dbh + C * \pi * (\frac{dbh}{2})^2*h \\
-    \Delta V = V*\frac{\Delta dbh}{dbh}*(3-\frac{dbh}{dbh + ah})
-  \end{array}
-  \label{eq:Deltadbh}
-\end{equation}
-
-Next, TROLL used the new trunk dimension (\(dbh\) and \(h\)) to update
-tree crown geometry using allometric equations \citep{Chave2005}:
-
-\begin{equation}
-  \begin{array}{c} \\
-    CR = 0.80 + 10.47*dbh - 3.33*dbh^2\\
-    CD = -0.48 + 0.26*h~;~CD = 0.13 + 0.17*h~(h<5~m)
-  \end{array}
-  \label{eq:C}
-\end{equation}
-
-Finally, TROLL computes the mean leaf density within the crown (\(LD\)
-in \(m^2.m^{-3}\)) assuming a uniform distribution:
-
-\begin{equation}
-  LD = \frac{LA_{young}+LA_{mature}+LA_{old}}{\pi*CR^2*CD}
-  \label{eq:LD}
-\end{equation}
-
-\subsection{Mortality}\label{mortality}
-
-Mortality is partitioned in three factors inside TROLL: background death
-\(d_b\), treefall death \(d_t\) and negative density dependent death
-\(d_{NDD}\). Because density dependent death \(d_{NDD}\) is still in
-development inside TROLL we did not used it, so we will not detail is
-computation.
-
-\citet{chave_towards_2009} advocated for a wood economics spectrum
-opposing fast growing light wood species species with high risk of
-mortality to slow growing dense wood species with reduced risk of
-mortality. Hence, background mortality is derived from wood specific
-gravity \(wsg\) inside TROLL:
-
-\begin{equation}
-  d_b = m*(1-\frac{wsg}{wsg_{lim}})+d_n
-  \label{eq:db}
-\end{equation}
-
-\(m\) (\(events.year^{-1}\)) is the reference background death rate for
-lighter wood species (pioneers). \(d_n\) represents death by
-carbohydrates shortage. If the number of consecutive day with
-\(NPP_{ind} < 0\) \eqref{eq:NPP} is superior to tree leaf lifespan \(d_n\)
-is set to 1 and remains null in other cases.
-
-Mortality by treefall inside TROLL depends on a specifric stochastic
-threshold \(\theta\):
-
-\begin{equation}
-  \theta = h_{max}*(1-v_T*|\zeta|)
-  \label{eq:theta}
-\end{equation}
-
-\(h_{max}\) is the maximal tree height. \(v_T\) is the variance term set
-to 0.3. \(|\zeta|\) is the absolute value of a random centered and
-scaled Gaussian. If the tree hieght \(h\) is superior to \(\theta\) then
-the tree may fall with a probability \(1-\theta/h\) \citep{Chave1999}.
-The treefall direction is random (drawn from a uniform law
-(\(\mathcal{U}[0,2\pi]\)). All tree in the trajectory of the falling
-tree will be hurted through a variable denoted \(hurt\), incremented by
-fallen tree height \(h\). If a tree height is inferior than its \(hurt\)
-values then it may die with a probability
-\(1-\frac{1}{2}\frac{h}{hurt}\). \(hurt\) variable is reset to null at
-each timestep (\(month\)).
-
-\subsection{Recruitment}\label{recruitment}
-
-Once the tree became fertile they will start to disperse seeds. TROLL
-consider tree as fertile after a specific height threshold
-\(h_{mature}\) \citep{Wright2005}:
-
-\begin{equation}
-  h_{mature} = -11.47+0.90*h_{max}
-  \label{eq:hmature}
-\end{equation}
-
-But TROLL is not considering seed directly through a seedbank, instead
-seed might be interpreted as a seedling recruitment opportunity. The
-number of reproduction opportunities per mature tree is denoted \(n_s\)
-and set to 10 for all species. This assumption originates from a
-trade-off between seed number and seed size resulting in equivalent
-survival and recruitment probability. All \(n_s\) events are dispersed
-with a distance randomly drawn from a Gaussian distribution.
-Additionally, TROLL model consider external seedrain through \(n_{ext}\)
-events of seed immigration:
-
-\begin{equation}
-  n_{ext} = N_{tot}*f_{reg}*n_{ha}
-  \label{eq:next}
-\end{equation}
-
-N\_\{tot\} is the external seedrain per hectare (number of reproduction
-opportunities). \(f_{reg}\) is the species regional frequency.
-\(n_{ha}\) is the simulated plot size in \(ha\).
-
-Finally, a bank of seedlings to be recruited is defined for each pixel.
-If the ground-level light reaches a species light compensation point
-\(LCP\) the species will be recruited:
-
-\begin{equation}
-  LCP = \frac{R_{leaf}}{\phi}
-  \label{eq:LCP}
-\end{equation}
-
-\(R_{leaf}\) is the leaf respiration for maintenance (see \eqref{eq:Rl}).
-\(\phi\) is the quantum yield (\(\mu mol C.\mu mol~photon\)) set to
-0.06. If several species reach their \(LCP\), one is picked at random.
-Seedlings are recruited with following intial geometry:
-
-\begin{equation}
-  \begin{array}{c} \\
-    dbh = \frac{a_h}{h_{max} - 1}\\
-    h = 1~m\\
-    CR = 0.5~m\\
-    CD = 0.3~m\\
-    LD = 0.8~m^2.^{-3}
-  \end{array}
-  \label{eq:C}
-\end{equation}
-
-\section{Sensitivity analysis}\label{sensitivity-analysis}
-
-\citet{Li} already assessed TROLL model sensitivity to several
-parameters (\(k\) see \eqref{eq:PPFD}, \(\phi\) see \eqref{eq:LCP}, \(g1\)
-see \eqref{eq:gs}, \(f_{wood}\) see \eqref{eq:DeltaV}, \(f_{canopy}\) see
-\eqref{eq:DeltaLA} and \(m\) see \eqref{eq:db}) which they assumed having a
-key role in model functioning.
-
-On the other hand, we decided to use TROLL to study resistance and
-resilience of ecosystem face to disturbance, highlighting the role of
-biodiversit. Consequently we particulrarly needed to assess the
-importance of functional traits to further better control and evaluate
-functional diversities. We also needed to assess the sensitivity of
-TROLL model to the seed rain constant (\(n_{ext}\), see \eqref{eq:next})
-because we assume it is one of the main factor of tree recruitments
-after disturbance in simulations.
-
-\subsection{Functional traits}\label{functional-traits}
-
-TROLL model currenty uses leaf mass per area (\(LMA\) in \(g.m^{-2}\)),
-leaf nitrogen content per dry mass (\(N_m\) in \(mg.g^{-1}\)), leaf
-phosphorus content per dry mass (\(P_m\) in \(mg.g^{-1}\)), wood
-specific gravity (\(wsg\) in \(g.cm^{-3}\)), diameter at breasth height
-threshold (\(dbh_{thresh}\) in \(m\)), asymptotic height (\(h_{lim}\) in
-\(m\)), and parameter of the tree-height-dbh allometry (\(a_h\) in
-\(m\)). To assess the sensitivity of TROLL model to species functionnal
-traits, we performed a sensitivity analysis by fixing species trait
-values to their mean. Each trait was tested independently. We reduce to
-a common mean traits with a Perason's correlation value \(r \geq 0.8\)
-(\(h_{max}\) and \(a_h\) with a correlation of \(r=0.98\)).
-
-\subsection{Seed rain}\label{seed-rain}
-
-To assess the sensitivity of TROLL model to seed rain, we performed a
-sensitivity analysis by fixing simulations seed rain constant to 2, 20,
-200 and 2000 seeds per hectare (default value being 200 seeds per
-hectare).
-
-Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and
-188.9 GB of memory. We assumed maturity of the forest after 500 years of
-regeneration (Maréchaux \& Chave) and computed simulation 100 years
-after a disturbance event with 40\% loss of basal area. Due to computer
-limitations we did not run replicate (besides it should be necessary to
-reduce simulation stochasticity). To assess ecosystem outputs
-sensitivity to studied parameters, we compared it to 100 replicates of
-control simulations with all parameters set to default values. Ecosystem
-outputs outside of the range of the control replicates values are
-significantly influenced by the studied parameter.
-
-\section{Disturbance}\label{disturbance}
-
-Most tropical timber production originates from selective logging, the
-targeted harvesting of timber from species of interest. Consequently,
-tropical sylviculture can be assimilated to a disturbance. The main
-difference between a disturbance and selective logging is the targetting
-of both species and individuals of interest. So we decided to first
-asses unselective disturbance effect on tropical forest ecosystem to
-subsequently better understand selective logging effect.
-
-We first implemented a disturbance module inside TROLL model to simulate
-unselective disturbance. 60 tree communities were then defined with
-different levels of both specific and functional richness to explore the
-role of biodiversity on tropical forest ecosystem answer to disturbance.
-Next, different levels of disturbance were simulated on previous mature
-communities. Finally, resilience of ecosystem major global variables
-(related to carbon stock, forest dynamic and florisitic structure) from
-simulations were partitioned between selection effects and niche
-complementarity \citep{Loreau2001}.
-
-\subsection{Model description}\label{model-description-1}
-
-Disturbance module was designed in the simplest way in order to relate
-the ecosystem answer to volume lost without any individuals nor species
-targetting. For a given iteration \(disturb_{iter}\), individuals are
-picked randomly with a uniform law on the number of trees. Selected
-individuals are then removed without trigerring a treefall to avoid any
-side effect. The operation is repeated untill the disturbance result in
-a defined lost basal area (\(disturb_{intensity}\) in \% of BA).
-
-\subsection{Design of experiment}\label{design-of-experiment}
-
-In order to assess the role of biodiversity in ecosystem answer to
-disturbance, we needed to create a space of experiments encompassing
-both variation of disturbance, biodiversity and time. Disturbance was
-represented by percentage of basal area loss (0\%, 25\%, 50\% and 75\%).
-Biodiversity was integrated with two components specific and functional
-diversities. We used species richness \(SR\) to represents species
-diversity (5, 25, and 125). Functional diversity can be related to
-numerous components, and \citet{Borgy2017} argued for 5: richness,
-divergence, regularity, overlap and mean. Because mature forest were
-created from a bare soil with TROLL simulations, we could not control a
-priori divergence, regularity and overlap but only assess theim before
-diturbance. Consequently, we focused on functional richness with convex
-hull volume \(CHV\) and functional mean with community weighted mean
-\(CWM\). For each level of species richness \(SR\), we selected 20
-communities with growing convex hull volume \(CHV\) but with a community
-weighted means close to the regional species pool community weighted
-means. Effectivelly, we did not wanted drastic change in community means
-that could have more effect than functional richness itself. This design
-of experiments resulted in 60 communities (\(5~SR*20~CHV\)) and 240
-simulations (\(60 ~communities*4~levels~of~disturbance\)) over 600 years
-(maturity being assumed after 500 years of regeneration \citep{Li}).
-Figure \ref{fig:DOE} presents the design of experiment for communities
-biodiversity after the forest mature were simulated, and thus before
-disturbance. We obtained a broad range of both functionl dispersion
-\(FDis\) and aboveground biomass \(AGB\) for simulated forest ecosystems
-before disturbance.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{master-thesis_files/figure-latex/DOE-1.pdf}
-\caption{\label{fig:DOE}Experimental design before disturbance. Communities
-are implemented along a gradient of species richness (SR) and functional
-dispersion (FDis) resulting in a broad range of aboveground biomass
-(AGB). FDis was caluclated based on 4 functional traits (leaf mass per
-area, wood specific gravity, maximum diameter, maximum height).}
-\end{figure}
-
-\subsection{Ecosystem answer anlaysis}\label{ecosystem-answer-anlaysis}
-
-Tropical forest ecosystems provides numerous ecosystem services linked
-to several ecosystem functions. We focused on few functions and related
-metrics to analyse ecosystem answer to disturbance: carbon stock with
-aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), forest dynamic with
-number of stem above \(10~cm\) diameter at breast height (\(N10\)), and
-floristic composition with \textbf{?} (\textbf{?} in \textbf{?}).
-
-The resilience of metrics values post disturbance were assessed through
-\citet{Henry2012} formula:
-
-\begin{equation}
-  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X(t)}{X(t_d-1)}
-  \label{eq:Resilience}
-\end{equation}
-
-The resilience of the system \(R(t)\) at the time \(t\) is described by
-the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
-\(Loss(t_d)\) at disturbance time \(t_d\). We transformed the equation
-in the resilience of the system \(R(t)\) at the time \(t\) being the
-ratio of the ecosystem metric \(X(t)\) at time \(t\) to the ecosystem
-metric before the disturbance happened \(X(t_d-1)\) at time \(t_d-1\).
-\(X(t_d-1)\) was calculated as the mean of the ecosystem output for 50
-last years of the simulation of the mature forest (over 600 years).
-Then, we calculated for each simulation the recovery time
-\(t_{recovery}\) were the ecosystem reached back its stable state
-defined as its state before disturbance. But some ecosystem metrics can
-reach pre disturbance value wihtout revealing ecosystem recovery. For
-instance, number of stem above \(10~cm\) diameter at breast height
-\(N10\) will first decrease due to disturbance. Then \(N10\) will exceed
-its initial value due to new seedlings recruitment before decreasing
-towards its pre disturbanec value. Consequently, the first time \(N10\)
-reaches back its pre disturbance value can not be considered as the
-recovery time \(t_{recovery}\). In order to solve this issue, we
-considered the ecosystem as stable again when its whole set of observed
-metrics reach their pre disturbance values.
-
-Biodiversity is not only a facet of the experimental design and an
-ecosystem output through florisitic composition, but also interact on
-ecosystem functioning and consequently on its answer to disturbance.
-Biodiversity ecosystem functioning relation can be split in
-complementarity and selection effect with \citet{Loreau2001}
-partitioning:
-
-\begin{equation}
-  \begin{array}{c}
-    NE = X_O - X_E = CE + SE \\
-    CE = N* \overline{\Delta RX} \overline{M}\\
-    SE = N*cov(\Delta RX,M)
-  \end{array}
-  \label{eq:BiodivPart}
-\end{equation}
-
-Biodiversity net effect \(NE\) is based on the difference between
-ecosystem variable \(X\) observed value \(X_O\) within the community
-mixture of species and its expected value \(X_E\) if species performance
-were equal to their performance in monocultures. This effect can be
-partitioned between complementarity effect \(CE\), representing niche
-partitionning, positive interactions, and resource supply, and selectvie
-effect \(SE\) due to dominant species pool driving the ecosystem. Both
-metrics depend on the variation of relative ecosystem variable
-\(\Delta RX\):
-
-\begin{equation}
-  \Delta RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
-  \label{eq:DeltaRY}
-\end{equation}
-
-\(X_sp\) is the ecosystem variable value for one species either in
-mixture \(X_{sp}(mixture)\) or in monoculture \(X_{sp}(monoculture)\).
-\(P_{sp}\) is the proportion of the species in the mixture represented
-by species relative abundance. Consequently, \(CE\) averages diversity
-effects of all species presents in the mixture (both negatives and
-positives). Whereas \(SE\) become positive when dominant species
-outperform theimselves in mixture than in monoculture, and negative when
-less dominant species outperform theimselves in mixture than in
-monoculture \citep{Tobner2016}.
-
-Recovery trajectories of ecosystem variable after disturbance were
-partitioned between complementarity effect \(CE\) and selection effect
-\(SE\). In order to do that, the design of experiment was repeated for
-each species indivdually representing 652 simulations of monoculture.
-
-\section{Selective logging}\label{selective-logging}
-
-Once the unselective effect of disturbance on forest ecosystem studied,
-we can focus on selective logging to highlight the effect of species and
-individuals targetting on the ecosystem answer.
-
-\subsection{Model description}\label{model-description-2}
-
-\subsubsection{Designation}\label{designation}
-
-\subsubsection{Selection}\label{selection}
-
-\subsubsection{Rotten trees}\label{rotten-trees}
-
-\subsubsection{Felling}\label{felling}
-
-\subsubsection{Tracks}\label{tracks}
-
-\subsubsection{Gap damages}\label{gap-damages}
-
-\subsection{Design of experiment}\label{design-of-experiment-1}
-
-\subsection{Ecosystem answer
-anlaysis}\label{ecosystem-answer-anlaysis-1}
-
-Similar to disturbance
-
-\section{Results}\label{results}
-
-\subsection{Sensitivity}\label{sensitivity}
-
-\subsection{Disturbance}\label{disturbance-1}
-
-\subsection{Sylviculture}\label{sylviculture}
-
-\section{Discussion}\label{discussion}
-
-\listoftables
-\listoffigures
-
-\renewcommand\refname{}
-\bibliography{/home/sylvain/Documents/Bibliography/library.bib}
-
-\end{document}
+\vspace*{\fill}
+\includegraphics{images/logo}

### 2017-06-02 latex/before_body.tex
disqus test
template tex v2

@@
-\date{}
+% First page
+\includegraphics{images/logos}
 
-\thispagestyle{empty}
 \begin{center}
-\includegraphics{images/main.pdf}
+  \vspace*{\stretch{1}}
+  \LARGE{\textbf{Mémoire de stage}} \\
+  \vspace*{\fill}
+  \large{présenté par} \\
+  \large{Sylvain SCHMITT} \\
+  \vspace*{\fill}
+  \large{pour obtenir le diplôme national de master} \\
+  \large{mention Biodiversité, écologie, évolution} \\
+  \small{parcours Biodiversité végétale et gestion des écosystèmes tropicaux (BIOGET)} \\
+  \vspace*{\fill}
+  \large{Sujet :} \\
+  \Large{\textbf{A COMPLETER}} \\
+  \vspace*{\fill}
+  \large{soutenu publiquement le XX.xxxx.201X} \\
+  \large{à Kourou} \\
+  \vspace*{\fill}
+  \large{devant le jury suivant :} \\
+  \vspace*{\fill}
+  Dr Bruno HÉRAULT  \emph{Tuteur de stage} \\
+  Titre Prénom NOM  \emph{Examinateur} \\
+  Titre Prénom NOM  \emph{Examinateur} \\
+  Dr Stéphane TRAISSAC  \emph{Enseignant-réferent} \\
+  \vspace*{\stretch{1}}
 \end{center}
 
-\setlength{\abovedisplayskip}{-5pt}
-\setlength{\abovedisplayshortskip}{-5pt}
+% Second page
+\newpage
+\vspace*{\fill}
+\emph{Les opinions émises par les auteurs sont personnelles et n'engagent pas AgroParisTech.}
+\newpage

### 2017-06-02 01-Acknowledgments.Rmd
Model and M&M reorganized + sylviculture model v1

# Acknowledgments {-}

I would like to thank...

- Bruno
- Stéphane
- Laurent
- Fabian
- Isabelle
- Jérôme
- Camilla
- Aurélie
- Éric
- ...

### 2017-06-02 03-Model.Rmd
Model and M&M reorganized + sylviculture model v1

--- 2017-05-30 03-Model.Rmd
@@
 
-```{r traits, echo=FALSE}
-table <- data.frame(
-  Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
-  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breasth height threshold', 'asymptotic height', 'parameter of the tree-height-dbh allometry'),
-  Units = c('$g.m^-2$', '$mg.g^-1$', '$mg.g^-1$', '$g.cm^-3$', '$m$', '$m$', '$m$')
-)
-knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.')
+```{r TROLLtree, echo=FALSE, fig.cap='Individuals tree inside TROLL explicit spatial grid from @Li.
+Tree geometry (crown radius CR, crown depth CD, height h, diameter at breast height dbh) is updated at each timestep following allometric relationship with assimilated carbon allocated to growth.
+Each tree is flagged with a species label linking to its species-specific attributes.
+Light is compputed explicitly at each timestep for each voxel.'}
+knitr::include_graphics('images/TROLLtree.png', dpi = NA)
 ```
@@
 The source code is written in C++ and available upon request.
-All analyses were conducted in R version 3.4.0 **Cite R, add entry in Mendeley**.
+All modules of TROLL models are further detailed in [Appendix 1].
 
-```{r TROLLtree, echo=FALSE, fig.cap='Individuals tree inside TROLL explicit spatial grid from @Li.
-Tree geometry (crown radius CR, crown depth CD, height h, diameter at breast height dbh) is updated at each timestep following allometric relationship with assimilated carbon allocated to growth.
-Each tree is flagged with a species label linking to its species-specific attributes.
-Light is compputed explicitly at each timestep for each voxel.'}
-knitr::include_graphics('images/TROLLtree.png', dpi = NA)
+```{r traits, echo=FALSE}
+table <- data.frame(
+  Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
+  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breasth height threshold', 'asymptotic height', 'parameter of the tree-height-dbh allometry'),
+  Units = c('$g.m^{-2}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$g.cm^{-3}$', '$m$', '$m$', '$m$')
+)
+knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Li. Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.', format = 'pandoc')
 ```
 
-## Abiotic environment
+Previous implementation of TROLL model used @Reich1991a allometry to infer leaf lifespan $LL$ from species leaf mass per area $LMA$ [@Li, see [Appendix 1]].
+But the use of the allometrie from @Reich1991a with current implementation of the TROLL model resulted in an underestimation of leaf lifespan for low LMA species.
+Consequently in the following paragraph we suggest a new allometry.
 
-A voxel space, with a resolution of 1 $m^3$, is used to explicilty model the abiotic environment.
-For each tree crown, leaf area density is calculated on tree geometry assuming a uniform distriution across voxels occupied by the crown.
-Leaf area density is computed within each voxel summing all tree crowns inside the voxel $v$, and is denoted $LAD(v)$ (leaf area per voxel in $m².m^-3$).
-The vertical sum of $LAD$ from voxel $v$ to the ground level defines $LAI(v)$ (leaf area per fround area in $m^2.m^-2$ commonly called leaf area index):
+Selective logging is defined as the targeted harvesting of timber from species of interest.
+Consequently, tropical sylviculture can be assimilated to a disturbance.
+The main difference between a disturbance and selective logging is the targetting of both species and individuals of interest.
+So we decided to first asses unselective disturbance effect on tropical forest ecosystem to subsequently better understand selective logging effect.
+First, we implemented a disturbance module inside TROLL model to simulate unselective disturbance.
+Secondly, we implemented a sylviculture module inside TROLL model to simulate selective logging in regards to french Guiana practices.
+
+## Leaf lifespan
+
+The underestimation of leaf lifespan for low LMA species with the allometry from @Reich1991a resulted in indivduals unealistic early death from carbon starvation.
+We gathered data from TRY[@Kattge2011], DRYAD [@chave_towards_2009] and GLOPNET [@wright_worldwide_2004] datasets.
+We used an out of the bag method applied on a random forest to select variables with highest importance to explain leaf lifespan.
+We thus selected leaf mass per are $LMA$, leaf nitrogen content $N$ and wood specific gravity $wsg$.
+We then used a bayesian approach to test different models with growing level of complexity.
+The model with the best tradeoff between complexity (number of parameters), convergence, likelihood, and prediction quality (root mean square error of prediction RMSEP) was kept.
+We selected following model with a maximum likelihood of 13.6 and a RMSEP of 12 months:
 
 \begin{equation}
-  LAI(v) = \sum _{v'=v} ^\infty LAD(v') 
-  (\#eq:LAI)
+  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
+  (\#eq:LL)
 \end{equation}
 
-Daily variations in light intensity (photosynthetic photon flux density PPFD in $\mu mol_{photons}.m^-2.s^-1$), temperature (T in degrees Celsius), and vapor pressure deficit (VPD in $kPA$) are computed to assess carbon assimilation within each voxel of the canopy and for a representative day per month (see Appendix 1 from @Li for further details).
-Variation of PPFD Within the canopy is calculated as a loacal Beer-Lambert extinction law:
+Leaf lifespan $LL$ follows a lognormal law with location infered from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
+Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$.
+All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without presemption following a gamma law.
+$d$ represents the dataset random effects and encompass environmental and protocol variations.
+**Add Appendix ?**
+
+## Disturbance
+
+Disturbance module was designed in the simplest way in order to relate the ecosystem answer to volume lost without any individuals nor species targetting.
+For a given iteration $disturb_{iter}$, individuals are picked randomly with a uniform law on the number of trees.
+Selected individuals are then removed without trigerring a treefall to avoid any side effect.
+The operation is repeated untill the disturbance result in a defined lost basal area ($disturb_{intensity}$ in % of BA).
+
+## Sylviculture
+
+In french guiana context sylviculture can be narrow to selective logging, which can be split in two steps: selection and harvesting.
+Selection encompass choice of the havrestable area, tree designation by the forest office, tree selection by the harvester, and removal off tree probbed as rotten by the lumber.
+Harvesting encompass tree felling, tracks opening, and long term damages (simplified in gap damages in current TROLL implementation).
+
+### Designation and selection
+
+One major limit of current implementation of TROLL model is that it assumes a flat environment.
+Consequently the whole simulated area inside TROLL is considered has an harvestable zone.
+With all commercial species minimum and maximum harvestable diameter, TROLL calculates the total harvestable volume ${V_h}_{tot}$.
+If the total harvestable volume ${V_h}_{tot}$ exceed $30~m^3.hectare^{-1}$, commercial species minimum harvestable diameter $dbh_{min}$ is increased untill ${V_h}_{tot}$ is inferior to that upper limit.
+
+In french guiana, tree harvesters are focusing on few species with easier marketable wood, resulting in a tree harvest around $20~m^3.ha^{-1}$ (Laurent Descroix, ONF, personnal communication).
+TROLL ranks each commercial species on its economic value, and randomly remove individuals from lowest rank species untill it reaches total harvested volume ${V_{hd}}_{tot}$ (${V_{hd}}_{tot}$ was set to $25~m^3.hectare^{-1}$ in subsequent simulations).
+
+### Rotten trees
+
+20 to 30 % of designated trees are considered as rotten once probed by the lumberman.
+Rotten trees are not random and depends both on tree species and diameter.
+
+**Explain rotten bayesian model**
 
 \begin{equation}
-  PPFD_{max,month}(v) = PPFD_{top,max,month}*e^{-k*LAI(v)}
-  (\#eq:PPFD)
+  P_{rotten} = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
+  (\#eq:rotten)
 \end{equation}
 
-The daily maximum incident PPFD at the top of canopy $PPFD_{top,max,month}$ is given as input.
-The extinction rate $k$ is assumed as constant (see table \@ref(tab:par) for parameters value), besides is variation with zenith angle and species leaf inclination angle [@Meir2000].
-Moreover only vertical light diffusion is considered ignoring lateral light diffusion, which can have an important role especially in logging gaps.
-Finally, intra-day variation at half hour time steps $t$ for a representative day every month are used to compute $PPFD_{month}(v,t)$, $T_{month}(v,t)$ and $VPD_{month}(v,t)$.
-Water and nutrient process both in soil and inside trees are not simulated.
+**Add Appendix ?**
 
-## Photosynthesis
+### Harvesting
 
-## Autotrophic respiration
+Due to crown aspects, treefall from logs are often random (whereas difficult to manage, treefall can still be oriented, Laurent Descroix, ONF, personnal communication).
+Consequently, TROLL consider treefall from log as random like current natural treefall implementation inside TROLL (see code [Appendix 1]).
 
-## Carbon uptake
+Tree harvesting roads are split in three classes: truck roads, main tractor track, and secondary track.
+Because TROLL assumes a flat environment, the main track is opened starting from the middle of one side of the simulated forest and untill it reaches the center with a width of 6 meters.
+In most cases, secondary tracks are opened once trees have been designated and the geolocation taken at a maximum distance of 30 meters from designated trees (Laurent Descroix, ONF, personnal communication).
+To simulate secondary tracks, TROLL uses a loads map, measuring every trees at a distance of 30 meters for each pixel, and  a track proximity maps of the closest existing track.
+Next, the model select the pixel with the highest load and closest track, find the closest existing track and join it by removing tree in the way with a width of 5 meters.
+The operations are repeated untill no felt trees are left.
 
-## Tree growth
+### Gap damages
 
-## Recruitment
+Most of models account long term damages due to selective logging with a 10 years increased mortality [@Huth2004; @Khler2004; @Ruger2008].
+We decided to model explicitly long term logging damages because of their localised nature through a gap damages model.
+We gathered data from Paracou dataset **Cite** in cenususes between 1988 and 1992 on Paracou harvested plots.
+Individuals were categorized between alive, dead, or recruited during the period.
+We measured each individual distance to the closest gap.
+We then used a bayesian approach to test the link between tree death in the four years following the log event and distance to the closest gap.
+We adapted the model from **Cite Herault** based on a disturbance index into.
 
-## Mortality
+\begin{equation}
+  \begin{array}{c} 
+    Death_i \sim \mathcal{B}(P(Death_i)) \\
+    P(Death_i) = logit^{-1}(\theta + \beta*e^{\alpha*d_{gaps}}) = \frac{e^{\theta + \beta*e^{\alpha*d_{gaps}}}}{1 + e^{\theta + \beta*e^{\alpha*d_{gaps}}}}
+  \end{array}
+  (\#eq:death)
+\end{equation}
+
+$Death$ of a tree $i$ follows a $\mathcal{Bernoulli}$ law of probability $P(Death_i)$.
+The odds for a tree $i$ to die $O(Death_i)$ is calculated by the sum of the natural tree death odd $\theta$ and a perturbation index $\beta*e^{\alpha*d_{gaps}}$.
+The perturbation index depend on the distance $d_{gaps}$ of the tree $i$ to the closest logging gap.
+The probability for a tree $i$ to die $P(Death_i)$ is finally calculated by taking the inverse logit of the odd $logit^{-1}[O(Death_i)]$.
+**Add Appendix ?**

### 2017-06-02 04-MnMs.Rmd
Model and M&M reorganized + sylviculture model v1

--- 2017-06-01 05-Disturbance.Rmd
@@
-```{r config5, message=FALSE, warning=FALSE, include=FALSE}
-rm(list = ls()) ; invisible(gc())
-library(ggplot2)
-library(ggrepel)
-```
+# Material and Methods
 
-# Disturbance
+## Sensitivity analysis
 
-Most tropical timber production originates from selective logging, the targeted harvesting of timber from species of interest.
+@Li already assessed TROLL model sensitivity to several parameters ($k$ see \@ref(eq:PPFD), $\phi$ see \@ref(eq:LCP), $g1$ see \@ref(eq:gs), $f_{wood}$ see \@ref(eq:DeltaV), $f_{canopy}$ see \@ref(eq:DeltaLA) and $m$ see \@ref(eq:db)) which they assumed having a key role in model functioning.
+
+On the other hand, we decided to use TROLL to study resistance and resilience of ecosystem face to disturbance, highlighting the role of biodiversit.
+Consequently we particulrarly needed to assess the importance of functional traits to further better control and evaluate functional diversities.
+We also needed to assess the sensitivity of TROLL model to the seed rain constant ($n_{ext}$, see \@ref(eq:next)) because we assume it is one of the main factor of tree recruitments after disturbance in simulations.
+
+TROLL model currenty uses leaf mass per area ($LMA$ in $g.m^{-2}$), leaf nitrogen content per dry mass ($N_m$ in $mg.g^{-1}$), leaf phosphorus content per dry mass ($P_m$ in $mg.g^{-1}$), wood specific gravity ($wsg$ in $g.cm^{-3}$), diameter at breasth height threshold ($dbh_{thresh}$ in $m$), asymptotic height ($h_{lim}$ in $m$), and parameter of the tree-height-dbh allometry ($a_h$ in $m$).
+To assess the sensitivity of TROLL model to species functionnal traits, we performed a sensitivity analysis by fixing species trait values to their mean.
+Each trait was tested independently.
+We reduce to a common mean traits with a Perason's correlation value $r \geq 0.8$ ($h_{max}$ and $a_h$ with a correlation of $r=0.98$).
+
+To assess the sensitivity of TROLL model to seed rain, we performed a sensitivity analysis by fixing simulations seed rain constant to 2, 20, 200 and 2000 seeds per hectare (default value being 200 seeds per hectare).
+
+Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and 188.9 GB of memory.
+We assumed maturity of the forest after 500 years of regeneration (Maréchaux & Chave) and computed simulation 100 years after a disturbance event with 40% loss of basal area.
+Due to computer limitations we did not run replicate (besides it should be necessary to reduce simulation stochasticity).
+To assess ecosystem outputs sensitivity to studied parameters, we compared it to 100 replicates of control simulations with all parameters set to default values.
+Ecosystem outputs outside of the range of the control replicates values are significantly influenced by the studied parameter.
+
+## Disturbance
+
+*Most tropical timber production originates from selective logging, the targeted harvesting of timber from species of interest.
 Consequently, tropical sylviculture can be assimilated to a disturbance.
@@
 So we decided to first asses unselective disturbance effect on tropical forest ecosystem to subsequently better understand selective logging effect.
+We first implemented a simple disturbance module inside TROLL model to simulate unselective disturbance.* **Already inside model descripiton !**
 
@@
 
-## Model description
+## Selective logging
 
-Disturbance module was designed in the simplest way in order to relate the ecosystem answer to volume lost without any individuals nor species targetting.
-For a given iteration $disturb_{iter}$, individuals are picked randomly with a uniform law on the number of trees.
-Selected individuals are then removed without trigerring a treefall to avoid any side effect.
-The operation is repeated untill the disturbance result in a defined lost basal area ($disturb_{intensity}$ in % of BA).
+Once the unselective effect of disturbance on forest ecosystem studied, we can focus on selective logging to highlight the effect of species and individuals targetting on the ecosystem answer.
+We first implemented a sylviculture module inside TROLL model to simulate selective logging in regards to french Guiana practices.
+We used the 60 communities from disturbance simulations to study the role of biodiversity in both forest ecosystem and commercial species answer to selective logging.
+We then
 

### 2017-06-02 07-Appendix1.Rmd
Model and M&M reorganized + sylviculture model v1

# (APPENDIX) Appendix {-}

# Appendix 1

I this Appendix we further detail modules of TROLL model.

## Abiotic environment

A voxel space, with a resolution of 1 $m^3$, is used to explicilty model the abiotic environment. For each tree crown, leaf area density is calculated on tree geometry assuming a uniform distriution across voxels occupied by the crown. Leaf area density is computed within each voxel summing all tree crowns inside the voxel $v$, and is denoted $LAD(v)$ (leaf area per voxel in $m².m^{-3}$). The vertical sum of $LAD$ from voxel $v$ to the ground level defines $LAI(v)$ (leaf area per fround area in $m^2.m^{-2}$ commonly called leaf area index):

\begin{equation}
  LAI(v) = \sum _{v'=v} ^\infty LAD(v') 
  (\#eq:LAI)
\end{equation}

Daily variations in light intensity (photosynthetic photon flux density PPFD in $\mu mol_{photons}.m^{-2}.s^{-1}$), temperature (T in degrees Celsius), and vapor pressure deficit (VPD in $kPA$) are computed to assess carbon assimilation within each voxel of the canopy and for a representative day per month (see Appendix 1 from @Li for further details). Variation of PPFD Within the canopy is calculated as a loacal Beer-Lambert extinction law:

\begin{equation}
  PPFD_{max,month}(v) = PPFD_{top,max,month}*e^{-k*LAI(v)}
  (\#eq:PPFD)
\end{equation}

The daily maximum incident PPFD at the top of canopy $PPFD_{top,max,month}$ is given as input. The extinction rate $k$ is assumed as constant, besides is variation with zenith angle and species leaf inclination angle [@Meir2000]. Moreover only vertical light diffusion is considered ignoring lateral light diffusion, which can have an important role especially in logging gaps. Finally, intra-day variation at half hour time steps $t$ for a representative day every month are used to compute $PPFD_{month}(v,t)$, $T_{month}(v,t)$ and $VPD_{month}(v,t)$. Water and nutrient process both in soil and inside trees are not simulated.

## Photosynthesis

### Theory

Troll simulates the carbon uptake of each individual with the Farquhar, von Caemmerer and Berry model of C3 photosynthesis [@Farquhar1980]. Gross carbon assimilation rate ($A$ in $\mu mol~CO_2. m^{-2}.s^{-1}$) will be the minimum of eiter Rubisco activity ($A_v$) or RuBP generation ($A_j$):

\begin{equation}
  A=min(A_v, A_j)~|~A_v=V_{cmax}*\frac{c_i-\Gamma^*}{c_i+K_m}~;~A_j=\frac{J}{4}*\frac{c_i-\Gamma^*}{c_i+2*\Gamma^*}
  (\#eq:A)
\end{equation}

$V_{cmax}$ is the maximum rate of carboxylation ($\mu mol~CO_2.m^{-2}.s^{-1}$). $c_i$ is the $CO_2$ partial pressure at carboxylation sites. $\Gamma^*$ is the $CO_2$ compensation point in absence of dark respiration. $K_m$ is the apparent knietic constant of the Rubisco. And $J$ is the electron transport rate ($\mu mol e^-.m^{-2}.s^{-1}$). $J$ depends on the light intensity with $PPFD$:

\begin{equation}
  J = \frac{1}{2*\theta}*[\alpha*PPFD+J_{max}-\sqrt{(\alpha*PPFD+J_{max})^2}-4*\theta*\alpha*PPFD*J_{max}]
  (\#eq:J)
\end{equation}

$J_{max}$ is the maximal electron transport capacity ($\mu mol e^-.m^{-2}.s^{-1}$). $\theta$ is the curvature factor. And $\alpha$ is the apparent quantum yield to electron transport ($mole^-.mol~photons^{-1}$).

Carbon assimilation by photosynthesis will then be limited by the $CO_2$ partial pressure at carboxylation sites. Stomata controls the gas concentration at carboxylation sites throught stomatal transport:

\begin{equation}
  A = g_s*(c_a-c_i)
  (\#eq:Ag)
\end{equation}

$g_s$ is the stomatal conductance to $CO_2$ ($molCO_2.m^{-2}.s^{-1}$). TROLL simulates stomatal conductance $g_s$ with the model from [@Medlyn2011]:

\begin{equation}
  g_s = g_0 + (1 + \frac{g_1}{\sqrt{VPD}})*\frac{A}{c_a}
  (\#eq:gs)
\end{equation}

$g_0$ and $g_1$ are parameters from the model. TROLL model assume $g_0 \approx 0$ (empirically tested and considered as reasonable).

### Parametrization

Leaf traits can be used as proxy of photosynthesis, especially leaf nutrient content which directly play a role in it [@wright_worldwide_2004]. @Domingues2010 suggested that $V_{cmac}$ and $J_{max}$ were both limited by the leaf concentration of nitrogen $N$ and phosphorus $P$ ($mg.g^{-1}$):

\begin{equation}
  log_{10} V_{cmax-M} = min( 
  \begin{array}{c} 
    -1.56+0.43*log_{10} N-0.37*log_{10} LMA \\
    -0.80+0.45*log_{10} P-0.25*log_{10} LMA 
  \end{array} 
  )
  (\#eq:VcmaxM)
\end{equation}

\begin{equation}
  log_{10} J_{max-M} = min(
  \begin{array}{c} 
    -1.50+0.41*log_{10} N-0.45*log_{10} LMA \\
    -0.74+0.44*log_{10} P-0.32*log_{10} LMA 
  \end{array}
  )
  (\#eq:JmaxM)
\end{equation}

$V_{cmax-M}$ and $J_{max-M}$ are the photosynthetic capacities at $25^\circ C$ of mature leaves per leaf dry mass (resp. $\mu mol CO_2.g^-1.s^{-1}$ and $\mu mol e^-.g^{-1}.s^{-1}$). $LMA$ is the leaf mass per are ($g.cm^{-2}$). $V_{cmax}$ and $J_{max}$ are calculated by multiplying $V_{cmax-M}$ and $J_{max-M}$ by $LMA$. $V_{cmax}$ and $J_{max}$ variation with temperature are caluclated with @Bernacchi2003 (see Appendix 2 from @Li for further details).

TROLL computes leaf carbon assimilation $A_l$ combining equations from \@ref(eq:A) to \@ref(eq:JmaxM) for each tree crown voxel within in each crown layer $l$:

\begin{equation}
  A_l = \frac{1}{n_v*t_M} * \sum_v  \sum^{t_M}_{t=1} A(PPFD_{month}(v,t),VPD_{month}(v,t),T_{month}(v,t))
  (\#eq:Al)
\end{equation}

$PPFD_{month}(v,t)$, $VPD_{month}(v,t)$ , and $T_{month}(v,t)$ are derived from microclimatic data. $n_v$ is the number of voxels within crown layer $l$. And the sum is calculated over the $t_M$ half-hourly intervals $t$ of a tipical day.

## Autotrophic respiration

A large fraction of plants carbon uptake is actually used for plant maintenance and growth respiration. The autotrophic respiration can represents up to 65% of the gross primary productivity but varies strongly among species, sites, and environnements.

TROLL uses @Atkin2015 database of mature leaf dark respiration and associated leaf traits to compute leaf maintenance respiration:

\begin{equation}
  R_{leaf-M} = 8.5431-0.1306*N-0.5670*P-0.0137*LMA+11.1*V_{cmax-M}+0.1876*N*P
  (\#eq:Rl)
\end{equation}

$R_{leaf-M}$ si the dark respiration rate per leaf dry mass at a temperaure of $25^\circ C$ ($nmolCO_2.g^{-1}.s^{-1}$). The other terms are in equations \@ref(eq:VcmaxM) and \@ref(eq:JmaxM). TROLL assume leaf respiration during day light to be 40% of leaf dark respiration, and computes total leaf respiration by accounting for the legnth of daylight.

TROLL model stem respiration ($R_{stem}$ in $\mu molC.s^{-1}$) with a constant respiration rate per volume of sapwood:

\begin{equation}
  R_{stem} = 39.6*\pi*ST*(dbh-ST)*(h-CD)
  (\#eq:Rs)
\end{equation}

dbh, h, CD and ST are tree diameter at breast height, height, corwn depth and sapwoond thickness, respectively ($m$). TROLL assumes $ST=0.04~m$ when $dbh>30~cm$ and an increasing $ST$ for lower $dbh$.

Finally, TROLL computes both fine root maintenance respiration, as half of the leaf maintenance respiration. Whereas coarse root and branch maintenance respiration is computed as half of the stem respiration. And growth respiration ($R_{growth}$) is assumed to account for 25% of the gross primary productivity minus the sum of maintenance respirations.

## Net carbon uptake

Net primary production of carbon for one individual $NPP_{ind}$ ($gC$) is computed by the balance between gross primary production $GPP_{ind}$ and respirations $R$:

\begin{equation}
  NPP_{ind} = GPP_{ind} - R_{maintenance} - R_{growth}
  (\#eq:NPP)
\end{equation}

TROLL partitions individuals total leaf area $LA$ into three pools for different leaf age classes corresponding to different photosynthesis efficiency (young, mature and old leaves with $LA_{young}$, $LA_{mature}$, and $LA_{old}$ respectivelly). Consequently we can compute growth primary production for one individual as:

\begin{equation}
  GPP_{ind} = 189.3 * \Delta t * \sum _{l= \lfloor h-CD \rfloor +1} ^{\lfloor h \rfloor} [A_l] * (\frac{LA_{young}}{2} + LA_{mature} + \frac{LA_{old}}{2})
  (\#eq:GPP)
\end{equation}

h and CD are tree height and crown depth, repectivelly ($m$). $\lfloor x \rfloor$ is the rounding function. $\Delta t$ is the duration of a timestep ($year$).

Thus, TROLL can compute carbon allocation to wood into an increment of stem volume $\Delta V$ ($m^3$):

\begin{equation}
  \Delta V = 10^{-6} * \frac{f_{wood}*NPP_{ind}}{0.5*wsg}*Senesc(dbh)
  (\#eq:DeltaV)
\end{equation}

$f_{wood}$ is the fixed fraction of NPP allocated to stem and branches. $wsg$ is the wood specific gravity ($g.cm^{-3}$, see \@ref(tab:traits)). TROLL assume large trees less efficient to convert NPP as growth by using a size-related growth decline with function $Senesc$ after a specific diameter at brest height threshold $dbh_{thresh}$:

\begin{equation}
  Senesc(dbh) = max(0;3-2*\frac{dbh}{dbh_{thresh}})
  (\#eq:Senesc)
\end{equation}

Finally, TROLL can compute carbon allocation to canopy with canopy NPP fraction denoted $f_{canopy}$ and decomposed into leaf, twig and fruit production. Carbon allocation to leaf results in a new young leaf pool, whereas other leaf pools are updated as follow:

\begin{equation}
  \begin{array}{c} \\
   \Delta LA_{young} = \frac{2*f_{leaves}*NPP_{ind}}{LMA}-\frac{LA_{young}}{\tau_{young}} \\
  \Delta LA_{mature} = \frac{LA_{young}}{\tau_{young}} - \frac{LA_{mature}}{\tau_{mature}}\\
  \Delta LA_{old} = \frac{LA_{mature}}{\tau_{mature}} - \frac{LA_{old}}{\tau_{old}}
  \end{array}
  (\#eq:DeltaLA)
\end{equation}

$\tau_{young}$, $\tau_{mature}$, and $\tau_{old}$ are species residence times in each leaf pools ($years$). The sum of residency time thus defined the leaf lifespan $LL = \tau_{young} + \tau_{mature} + \tau_{old}$ ($years$). $\tau_{young}$ is set to one month and $\tau_{mature}$ is set to a third of leaf lifespan $LL$. Belowground carbon allocation is not simulated inside TROLL.

## Leaf lifespan

The underestimation of leaf lifespan for low LMA species with the allometry from @Reich1991a resulted in indivduals unealistic early death from carbon starvation. We gathered data from TRY[@Kattge2011], DRYAD [@chave_towards_2009] and GLOPNET [@wright_worldwide_2004] datasets. We used an out of the bag method applied on a random forest to select variables with highest importance to explain leaf lifespan. We thus selected leaf mass per are $LMA$, leaf nitrogen content $N$ and wood specific gravity $wsg$. We then used a bayesian approach to test different models with growing level of complexity. The model with the best tradeoff between complexity (number of parameters), convergence, likelihood, and prediction quality (root mean square error of prediction RMSEP) was kept. We selected following model with a maximum likelihood of 13.6 and a RMSEP of 12 months:

\begin{equation}
  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
  (\#eq:LLb)
\end{equation}

Leaf lifespan $LL$ follows a lognormal law with location infered from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$. Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$. All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without presemption following a gamma law. $d$ represents the dataset random effects and encompass environmental and protocol variations. The sampling of model \@ref(eq:LLb) resulted in the following allometry:

\begin{equation}
  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
  (\#eq:LL)
\end{equation}

## Tree growth

Once the increment of stem volume $\Delta V$ calculated with equation \@ref(eq:DeltaV), TROLL convert it into an increment of tree diameter at breast height denoted $\Delta dbh$. TROLL infer tree height from $dbh$ using a Michaelis-Menten equation:

\begin{equation}
  h = h_{lim}*\frac{dbh}{dbh + a_h}
  (\#eq:h)
\end{equation}

On the other hand, we have the trunk volume $V = C * \pi * (\frac{dbh}{2})^2*h$, thus:

\begin{equation}
  \begin{array}{c} \\
    \Delta V = C*\frac{1}{2}*\pi*h*dbh*\Delta dbh + C * \pi * (\frac{dbh}{2})^2*h \\
    \Delta V = V*\frac{\Delta dbh}{dbh}*(3-\frac{dbh}{dbh + ah})
  \end{array}
  (\#eq:Deltadbh)
\end{equation}

Next, TROLL used the new trunk dimension ($dbh$ and $h$) to update tree crown geometry using allometric equations [@Chave2005]:

\begin{equation}
  \begin{array}{c} \\
    CR = 0.80 + 10.47*dbh - 3.33*dbh^2\\
    CD = -0.48 + 0.26*h~;~CD = 0.13 + 0.17*h~(h<5~m)
  \end{array}
  (\#eq:C)
\end{equation}

Finally, TROLL computes the mean leaf density within the crown ($LD$ in $m^2.m^{-3}$) assuming a uniform distribution:

\begin{equation}
  LD = \frac{LA_{young}+LA_{mature}+LA_{old}}{\pi*CR^2*CD}
  (\#eq:LD)
\end{equation}

## Mortality

Mortality is partitioned in three factors inside TROLL: background death $d_b$, treefall death $d_t$ and negative density dependent death $d_{NDD}$. Because density dependent death $d_{NDD}$ is still in development inside TROLL we did not used it, so we will not detail is computation.

@chave_towards_2009 advocated for a wood economics spectrum opposing fast growing light wood species species with high risk of mortality to slow growing dense wood species with reduced risk of mortality. Hence, background mortality is derived from wood specific gravity $wsg$ inside TROLL:

\begin{equation}
  d_b = m*(1-\frac{wsg}{wsg_{lim}})+d_n
  (\#eq:db)
\end{equation}

$m$ ($events.year^{-1}$) is the reference background death rate for lighter wood species (pioneers). $d_n$ represents death by carbohydrates shortage. If the number of consecutive day with $NPP_{ind}
< 0$ \@ref(eq:NPP) is superior to tree leaf lifespan $d_n$ is set to 1 and remains null in other cases.

Mortality by treefall inside TROLL depends on a specifric stochastic threshold $\theta$:

\begin{equation}
  \theta = h_{max}*(1-v_T*|\zeta|)
  (\#eq:theta)
\end{equation}

$h_{max}$ is the maximal tree height. $v_T$ is the variance term set to 0.3. $|\zeta|$ is the absolute value of a random centered and scaled Gaussian. If the tree hieght $h$ is superior to $\theta$ then the tree may fall with a probability $1-\theta/h$ [@Chave1999]. The treefall direction is random (drawn from a uniform law ($\mathcal{U}[0,2\pi]$). All tree in the trajectory of the falling tree will be hurted through a variable denoted $hurt$, incremented by fallen tree height $h$. If a tree height is inferior than its $hurt$ values then it may die with a probability $1-\frac{1}{2}\frac{h}{hurt}$. $hurt$ variable is reset to null at each timestep ($month$).

## Recruitment

Once the tree became fertile they will start to disperse seeds. TROLL consider tree as fertile after a specific height threshold $h_{mature}$ [@Wright2005]:

\begin{equation}
  h_{mature} = -11.47+0.90*h_{max}
  (\#eq:hmature)
\end{equation}

But TROLL is not considering seed directly through a seedbank, instead seed might be interpreted as a seedling recruitment opportunity. The number of reproduction opportunities per mature tree is denoted $n_s$ and set to 10 for all species. This assumption originates from a trade-off between seed number and seed size resulting in equivalent survival and recruitment probability. All $n_s$ events are dispersed with a distance randomly drawn from a Gaussian distribution. Additionally, TROLL model consider external seedrain through $n_{ext}$ events of seed immigration:

\begin{equation}
  n_{ext} = N_{tot}*f_{reg}*n_{ha}
  (\#eq:next)
\end{equation}

N_{tot} is the external seedrain per hectare (number of reproduction opportunities). $f_{reg}$ is the species regional frequency. $n_{ha}$ is the simulated plot size in $ha$.

Finally, a bank of seedlings to be recruited is defined for each pixel. If the ground-level light reaches a species light compensation point $LCP$ the species will be recruited:

\begin{equation}
  LCP = \frac{R_{leaf}}{\phi}
  (\#eq:LCP)
\end{equation}

$R_{leaf}$ is the leaf respiration for maintenance (see \@ref(eq:Rl)). $\phi$ is the quantum yield ($\mu mol C.\mu mol~photon$) set to 0.06. If several species reach their $LCP$, one is picked at random. Seedlings are recruited with following intial geometry:

\begin{equation}
  \begin{array}{c} \\
    dbh = \frac{a_h}{h_{max} - 1}\\
    h = 1~m\\
    CR = 0.5~m\\
    CD = 0.3~m\\
    LD = 0.8~m^2.^{-3}
  \end{array}
  (\#eq:C)
\end{equation}

### 2017-06-02 docs/master-thesis.tex
template.tex v3 (geometry, fontsize, baselinestretch)
Model and M&M reorganized + sylviculture model v1

@@
 %%% STYLE %%%
-\documentclass[]{article}
+\documentclass[12pt,]{article}
 
@@
 \urlstyle{same} % don't use monospace font for urls
+
+% geometry
+\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
+\renewcommand{\baselinestretch}{1.1}
 
@@
 
+\begin{itemize}
+\tightlist
+\item
+  Bruno
+\item
+  Stéphane
+\item
+  Laurent
+\item
+  Fabian
+\item
+  Isabelle
+\item
+  Jérôme
+\item
+  Camilla
+\item
+  Aurélie
+\item
+  Éric
+\item
+  \ldots{}
+\end{itemize}
+
 \section*{Introduction}\label{introduction}
@@
 sampled on the field.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{images/TROLLtree.png}
+\caption{\label{fig:TROLLtree}Individuals tree inside TROLL explicit spatial
+grid from \citet{Li}. Tree geometry (crown radius CR, crown depth CD,
+height h, diameter at breast height dbh) is updated at each timestep
+following allometric relationship with assimilated carbon allocated to
+growth. Each tree is flagged with a species label linking to its
+species-specific attributes. Light is compputed explicitly at each
+timestep for each voxel.}
+\end{figure}
+
+Carbon assimilation is computed over half-hourly period of a
+representative day. Then allocation is computed to simulate tree growth
+from an explicit carbone balance (in contrast to previous models).
+Finally environment is updated at each timestep set to one month.
+Seedlings are not simulated explicitly but as a pool. In addition
+belowground processes, herbaceous plants, epiphytes and lianas are not
+simulated inside TROLL. The source code is written in C++ and available
+upon request. All modules of TROLL models are further detailed in
+\protect\hyperlink{appendix-1}{Appendix 1}.
 
@@
 
-Carbon assimilation is computed over half-hourly period of a
-representative day. Then allocation is computed to simulate tree growth
-from an explicit carbone balance (in contrast to previous models).
-Finally environment is updated at each timestep set to one month.
-Seedlings are not simulated explicitly but as a pool. In addition
-belowground processes, herbaceous plants, epiphytes and lianas are not
-simulated inside TROLL. The source code is written in C++ and available
-upon request. All analyses were conducted in R version 3.4.0
-\textbf{Cite R, add entry in Mendeley}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics{images/TROLLtree.png}
-\caption{\label{fig:TROLLtree}Individuals tree inside TROLL explicit spatial
-grid from \citet{Li}. Tree geometry (crown radius CR, crown depth CD,
-height h, diameter at breast height dbh) is updated at each timestep
-following allometric relationship with assimilated carbon allocated to
-growth. Each tree is flagged with a species label linking to its
-species-specific attributes. Light is compputed explicitly at each
-timestep for each voxel.}
-\end{figure}
-
-\subsection{Abiotic environment}\label{abiotic-environment}
-
-A voxel space, with a resolution of 1 \(m^3\), is used to explicilty
-model the abiotic environment. For each tree crown, leaf area density is
-calculated on tree geometry assuming a uniform distriution across voxels
-occupied by the crown. Leaf area density is computed within each voxel
-summing all tree crowns inside the voxel \(v\), and is denoted
-\(LAD(v)\) (leaf area per voxel in \(m².m^{-3}\)). The vertical sum of
-\(LAD\) from voxel \(v\) to the ground level defines \(LAI(v)\) (leaf
-area per fround area in \(m^2.m^{-2}\) commonly called leaf area index):
-
-\begin{equation}
-  LAI(v) = \sum _{v'=v} ^\infty LAD(v') 
-  \label{eq:LAI}
-\end{equation}
-
-Daily variations in light intensity (photosynthetic photon flux density
-PPFD in \(\mu mol_{photons}.m^{-2}.s^{-1}\)), temperature (T in degrees
-Celsius), and vapor pressure deficit (VPD in \(kPA\)) are computed to
-assess carbon assimilation within each voxel of the canopy and for a
-representative day per month (see Appendix 1 from \citet{Li} for further
-details). Variation of PPFD Within the canopy is calculated as a loacal
-Beer-Lambert extinction law:
-
-\begin{equation}
-  PPFD_{max,month}(v) = PPFD_{top,max,month}*e^{-k*LAI(v)}
-  \label{eq:PPFD}
-\end{equation}
-
-The daily maximum incident PPFD at the top of canopy
-\(PPFD_{top,max,month}\) is given as input. The extinction rate \(k\) is
-assumed as constant, besides is variation with zenith angle and species
-leaf inclination angle \citep{Meir2000}. Moreover only vertical light
-diffusion is considered ignoring lateral light diffusion, which can have
-an important role especially in logging gaps. Finally, intra-day
-variation at half hour time steps \(t\) for a representative day every
-month are used to compute \(PPFD_{month}(v,t)\), \(T_{month}(v,t)\) and
-\(VPD_{month}(v,t)\). Water and nutrient process both in soil and inside
-trees are not simulated.
-
-\subsection{Photosynthesis}\label{photosynthesis}
-
-\subsubsection{Theory}\label{theory}
-
-Troll simulates the carbon uptake of each individual with the Farquhar,
-von Caemmerer and Berry model of C3 photosynthesis \citep{Farquhar1980}.
-Gross carbon assimilation rate (\(A\) in
-\(\mu mol~CO_2. m^{-2}.s^{-1}\)) will be the minimum of eiter Rubisco
-activity (\(A_v\)) or RuBP generation (\(A_j\)):
-
-\begin{equation}
-  A=min(A_v, A_j)~|~A_v=V_{cmax}*\frac{c_i-\Gamma^*}{c_i+K_m}~;~A_j=\frac{J}{4}*\frac{c_i-\Gamma^*}{c_i+2*\Gamma^*}
-  \label{eq:A}
-\end{equation}
-
-\(V_{cmax}\) is the maximum rate of carboxylation
-(\(\mu mol~CO_2.m^{-2}.s^{-1}\)). \(c_i\) is the \(CO_2\) partial
-pressure at carboxylation sites. \(\Gamma^*\) is the \(CO_2\)
-compensation point in absence of dark respiration. \(K_m\) is the
-apparent knietic constant of the Rubisco. And \(J\) is the electron
-transport rate (\(\mu mol e^-.m^{-2}.s^{-1}\)). \(J\) depends on the
-light intensity with \(PPFD\):
-
-\begin{equation}
-  J = \frac{1}{2*\theta}*[\alpha*PPFD+J_{max}-\sqrt{(\alpha*PPFD+J_{max})^2}-4*\theta*\alpha*PPFD*J_{max}]
-  \label{eq:J}
-\end{equation}
-
-\(J_{max}\) is the maximal electron transport capacity
-(\(\mu mol e^-.m^{-2}.s^{-1}\)). \(\theta\) is the curvature factor. And
-\(\alpha\) is the apparent quantum yield to electron transport
-(\(mole^-.mol~photons^{-1}\)).
-
-Carbon assimilation by photosynthesis will then be limited by the
-\(CO_2\) partial pressure at carboxylation sites. Stomata controls the
-gas concentration at carboxylation sites throught stomatal transport:
-
-\begin{equation}
-  A = g_s*(c_a-c_i)
-  \label{eq:Ag}
-\end{equation}
-
-\(g_s\) is the stomatal conductance to \(CO_2\)
-(\(molCO_2.m^{-2}.s^{-1}\)). TROLL simulates stomatal conductance
-\(g_s\) with the model from \citep{Medlyn2011}:
-
-\begin{equation}
-  g_s = g_0 + (1 + \frac{g_1}{\sqrt{VPD}})*\frac{A}{c_a}
-  \label{eq:gs}
-\end{equation}
-
-\(g_0\) and \(g_1\) are parameters from the model. TROLL model assume
-\(g_0 \approx 0\) (empirically tested and considered as reasonable).
-
-\subsubsection{Parametrization}\label{parametrization}
-
-Leaf traits can be used as proxy of photosynthesis, especially leaf
-nutrient content which directly play a role in it
-\citep{wright_worldwide_2004}. \citet{Domingues2010} suggested that
-\(V_{cmac}\) and \(J_{max}\) were both limited by the leaf concentration
-of nitrogen \(N\) and phosphorus \(P\) (\(mg.g^{-1}\)):
-
-\begin{equation}
-  log_{10} V_{cmax-M} = min( 
-  \begin{array}{c} 
-    -1.56+0.43*log_{10} N-0.37*log_{10} LMA \\
-    -0.80+0.45*log_{10} P-0.25*log_{10} LMA 
-  \end{array} 
-  )
-  \label{eq:VcmaxM}
-\end{equation}
-
-\begin{equation}
-  log_{10} J_{max-M} = min(
-  \begin{array}{c} 
-    -1.50+0.41*log_{10} N-0.45*log_{10} LMA \\
-    -0.74+0.44*log_{10} P-0.32*log_{10} LMA 
-  \end{array}
-  )
-  \label{eq:JmaxM}
-\end{equation}
-
-\(V_{cmax-M}\) and \(J_{max-M}\) are the photosynthetic capacities at
-\(25^\circ C\) of mature leaves per leaf dry mass (resp.
-\(\mu mol CO_2.g^-1.s^{-1}\) and \(\mu mol e^-.g^{-1}.s^{-1}\)). \(LMA\)
-is the leaf mass per are (\(g.cm^{-2}\)). \(V_{cmax}\) and \(J_{max}\)
-are calculated by multiplying \(V_{cmax-M}\) and \(J_{max-M}\) by
-\(LMA\). \(V_{cmax}\) and \(J_{max}\) variation with temperature are
-caluclated with \citet{Bernacchi2003} (see Appendix 2 from \citet{Li}
-for further details).
-
-TROLL computes leaf carbon assimilation \(A_l\) combining equations from
-\eqref{eq:A} to \eqref{eq:JmaxM} for each tree crown voxel within in each
-crown layer \(l\):
-
-\begin{equation}
-  A_l = \frac{1}{n_v*t_M} * \sum_v  \sum^{t_M}_{t=1} A(PPFD_{month}(v,t),VPD_{month}(v,t),T_{month}(v,t))
-  \label{eq:Al}
-\end{equation}
-
-\(PPFD_{month}(v,t)\), \(VPD_{month}(v,t)\) , and \(T_{month}(v,t)\) are
-derived from microclimatic data. \(n_v\) is the number of voxels within
-crown layer \(l\). And the sum is calculated over the \(t_M\)
-half-hourly intervals \(t\) of a tipical day.
-
-\subsection{Autotrophic respiration}\label{autotrophic-respiration}
-
-A large fraction of plants carbon uptake is actually used for plant
-maintenance and growth respiration. The autotrophic respiration can
-represents up to 65\% of the gross primary productivity but varies
-strongly among species, sites, and environnements.
-
-TROLL uses \citet{Atkin2015} database of mature leaf dark respiration
-and associated leaf traits to compute leaf maintenance respiration:
-
-\begin{equation}
-  R_{leaf-M} = 8.5431-0.1306*N-0.5670*P-0.0137*LMA+11.1*V_{cmax-M}+0.1876*N*P
-  \label{eq:Rl}
-\end{equation}
-
-\(R_{leaf-M}\) si the dark respiration rate per leaf dry mass at a
-temperaure of \(25^\circ C\) (\(nmolCO_2.g^{-1}.s^{-1}\)). The other
-terms are in equations \eqref{eq:VcmaxM} and \eqref{eq:JmaxM}. TROLL assume
-leaf respiration during day light to be 40\% of leaf dark respiration,
-and computes total leaf respiration by accounting for the legnth of
-daylight.
-
-TROLL model stem respiration (\(R_{stem}\) in \(\mu molC.s^{-1}\)) with
-a constant respiration rate per volume of sapwood:
-
-\begin{equation}
-  R_{stem} = 39.6*\pi*ST*(dbh-ST)*(h-CD)
-  \label{eq:Rs}
-\end{equation}
-
-dbh, h, CD and ST are tree diameter at breast height, height, corwn
-depth and sapwoond thickness, respectively (\(m\)). TROLL assumes
-\(ST=0.04~m\) when \(dbh>30~cm\) and an increasing \(ST\) for lower
-\(dbh\).
-
-Finally, TROLL computes both fine root maintenance respiration, as half
-of the leaf maintenance respiration. Whereas coarse root and branch
-maintenance respiration is computed as half of the stem respiration. And
-growth respiration (\(R_{growth}\)) is assumed to account for 25\% of
-the gross primary productivity minus the sum of maintenance
-respirations.
-
-\subsection{Net carbon uptake}\label{net-carbon-uptake}
-
-Net primary production of carbon for one individual \(NPP_{ind}\)
-(\(gC\)) is computed by the balance between gross primary production
-\(GPP_{ind}\) and respirations \(R\):
-
-\begin{equation}
-  NPP_{ind} = GPP_{ind} - R_{maintenance} - R_{growth}
-  \label{eq:NPP}
-\end{equation}
-
-TROLL partitions individuals total leaf area \(LA\) into three pools for
-different leaf age classes corresponding to different photosynthesis
-efficiency (young, mature and old leaves with \(LA_{young}\),
-\(LA_{mature}\), and \(LA_{old}\) respectivelly). Consequently we can
-compute growth primary production for one individual as:
-
-\begin{equation}
-  GPP_{ind} = 189.3 * \Delta t * \sum _{l= \lfloor h-CD \rfloor +1} ^{\lfloor h \rfloor} [A_l] * (\frac{LA_{young}}{2} + LA_{mature} + \frac{LA_{old}}{2})
-  \label{eq:GPP}
-\end{equation}
-
-h and CD are tree height and crown depth, repectivelly (\(m\)).
-\(\lfloor x \rfloor\) is the rounding function. \(\Delta t\) is the
-duration of a timestep (\(year\)).
-
-Thus, TROLL can compute carbon allocation to wood into an increment of
-stem volume \(\Delta V\) (\(m^3\)):
-
-\begin{equation}
-  \Delta V = 10^{-6} * \frac{f_{wood}*NPP_{ind}}{0.5*wsg}*Senesc(dbh)
-  \label{eq:DeltaV}
-\end{equation}
-
-\(f_{wood}\) is the fixed fraction of NPP allocated to stem and
-branches. \(wsg\) is the wood specific gravity (\(g.cm^{-3}\), see
-\ref{tab:traits}). TROLL assume large trees less efficient to convert
-NPP as growth by using a size-related growth decline with function
-\(Senesc\) after a specific diameter at brest height threshold
-\(dbh_{thresh}\):
-
-\begin{equation}
-  Senesc(dbh) = max(0;3-2*\frac{dbh}{dbh_{thresh}})
-  \label{eq:Senesc}
-\end{equation}
-
-Finally, TROLL can compute carbon allocation to canopy with canopy NPP
-fraction denoted \(f_{canopy}\) and decomposed into leaf, twig and fruit
-production. Carbon allocation to leaf results in a new young leaf pool,
-whereas other leaf pools are updated as follow:
-
-\begin{equation}
-  \begin{array}{c} \\
-   \Delta LA_{young} = \frac{2*f_{leaves}*NPP_{ind}}{LMA}-\frac{LA_{young}}{\tau_{young}} \\
-  \Delta LA_{mature} = \frac{LA_{young}}{\tau_{young}} - \frac{LA_{mature}}{\tau_{mature}}\\
-  \Delta LA_{old} = \frac{LA_{mature}}{\tau_{mature}} - \frac{LA_{old}}{\tau_{old}}
-  \end{array}
-  \label{eq:DeltaLA}
-\end{equation}
-
-\(\tau_{young}\), \(\tau_{mature}\), and \(\tau_{old}\) are species
-residence times in each leaf pools (\(years\)). The sum of residency
-time thus defined the leaf lifespan
-\(LL = \tau_{young} + \tau_{mature} + \tau_{old}\) (\(years\)).
-\(\tau_{young}\) is set to one month and \(\tau_{mature}\) is set to a
-third of leaf lifespan \(LL\). Previous implementation of TROLL model
-used \citet{Reich1991a} allometry to infer leaf lifespan \(LL\) from
-species leaf mass per area \(LMA\) \citep{Li}. But the use of the
-allometrie from \citet{Reich1991a} with current implementation of the
-TROLL model resulted in an underestimation of leaf lifespan for low LMA
-species. Consequently in the following paragraph we suggest a new
-allometry. Belowground carbon allocation is not simulated inside TROLL.
+Previous implementation of TROLL model used \citet{Reich1991a} allometry
+to infer leaf lifespan \(LL\) from species leaf mass per area \(LMA\)
+\citep[see \protect\hyperlink{appendix-1}{Appendix 1}]{Li}. But the use
+of the allometrie from \citet{Reich1991a} with current implementation of
+the TROLL model resulted in an underestimation of leaf lifespan for low
+LMA species. Consequently in the following paragraph we suggest a new
+allometry.
+
+Selective logging is defined as the targeted harvesting of timber from
+species of interest. Consequently, tropical sylviculture can be
+assimilated to a disturbance. The main difference between a disturbance
+and selective logging is the targetting of both species and individuals
+of interest. So we decided to first asses unselective disturbance effect
+on tropical forest ecosystem to subsequently better understand selective
+logging effect. First, we implemented a disturbance module inside TROLL
+model to simulate unselective disturbance. Secondly, we implemented a
+sylviculture module inside TROLL model to simulate selective logging in
+regards to french Guiana practices.
 
@@
   LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
-  \label{eq:LLb}
+  \label{eq:LL}
 \end{equation}
@@
 presemption following a gamma law. \(d\) represents the dataset random
-effects and encompass environmental and protocol variations. The
-sampling of model \eqref{eq:LLb} resulted in the following allometry:
-
-\begin{equation}
-  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
-  \label{eq:LL}
-\end{equation}
-
-\subsection{Tree growth}\label{tree-growth}
-
-Once the increment of stem volume \(\Delta V\) calculated with equation
-\eqref{eq:DeltaV}, TROLL convert it into an increment of tree diameter at
-breast height denoted \(\Delta dbh\). TROLL infer tree height from
-\(dbh\) using a Michaelis-Menten equation:
-
-\begin{equation}
-  h = h_{lim}*\frac{dbh}{dbh + a_h}
-  \label{eq:h}
-\end{equation}
-
-On the other hand, we have the trunk volume
-\(V = C * \pi * (\frac{dbh}{2})^2*h\), thus:
-
-\begin{equation}
-  \begin{array}{c} \\
-    \Delta V = C*\frac{1}{2}*\pi*h*dbh*\Delta dbh + C * \pi * (\frac{dbh}{2})^2*h \\
-    \Delta V = V*\frac{\Delta dbh}{dbh}*(3-\frac{dbh}{dbh + ah})
+effects and encompass environmental and protocol variations. \textbf{Add
+Appendix ?}
+
+\subsection{Disturbance}\label{disturbance}
+
+Disturbance module was designed in the simplest way in order to relate
+the ecosystem answer to volume lost without any individuals nor species
+targetting. For a given iteration \(disturb_{iter}\), individuals are
+picked randomly with a uniform law on the number of trees. Selected
+individuals are then removed without trigerring a treefall to avoid any
+side effect. The operation is repeated untill the disturbance result in
+a defined lost basal area (\(disturb_{intensity}\) in \% of BA).
+
+\subsection{Sylviculture}\label{sylviculture}
+
+In french guiana context sylviculture can be narrow to selective
+logging, which can be split in two steps: selection and harvesting.
+Selection encompass choice of the havrestable area, tree designation by
+the forest office, tree selection by the harvester, and removal off tree
+probbed as rotten by the lumber. Harvesting encompass tree felling,
+tracks opening, and long term damages (simplified in gap damages in
+current TROLL implementation).
+
+\subsubsection{Designation and
+selection}\label{designation-and-selection}
+
+One major limit of current implementation of TROLL model is that it
+assumes a flat environment. Consequently the whole simulated area inside
+TROLL is considered has an harvestable zone. With all commercial species
+minimum and maximum harvestable diameter, TROLL calculates the total
+harvestable volume \({V_h}_{tot}\). If the total harvestable volume
+\({V_h}_{tot}\) exceed \(30~m^3.hectare^{-1}\), commercial species
+minimum harvestable diameter \(dbh_{min}\) is increased untill
+\({V_h}_{tot}\) is inferior to that upper limit.
+
+In french guiana, tree harvesters are focusing on few species with
+easier marketable wood, resulting in a tree harvest around
+\(20~m^3.ha^{-1}\) (Laurent Descroix, ONF, personnal communication).
+TROLL ranks each commercial species on its economic value, and randomly
+remove individuals from lowest rank species untill it reaches total
+harvested volume \({V_{hd}}_{tot}\) (\({V_{hd}}_{tot}\) was set to
+\(25~m^3.hectare^{-1}\) in subsequent simulations).
+
+\subsubsection{Rotten trees}\label{rotten-trees}
+
+20 to 30 \% of designated trees are considered as rotten once probed by
+the lumberman. Rotten trees are not random and depends both on tree
+species and diameter.
+
+\textbf{Explain rotten bayesian model}
+
+\begin{equation}
+  P_{rotten} = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
+  \label{eq:rotten}
+\end{equation}
+
+\textbf{Add Appendix ?}
+
+\subsubsection{Harvesting}\label{harvesting}
+
+Due to crown aspects, treefall from logs are often random (whereas
+difficult to manage, treefall can still be oriented, Laurent Descroix,
+ONF, personnal communication). Consequently, TROLL consider treefall
+from log as random like current natural treefall implementation inside
+TROLL (see code \protect\hyperlink{appendix-1}{Appendix 1}).
+
+Tree harvesting roads are split in three classes: truck roads, main
+tractor track, and secondary track. Because TROLL assumes a flat
+environment, the main track is opened starting from the middle of one
+side of the simulated forest and untill it reaches the center with a
+width of 6 meters. In most cases, secondary tracks are opened once trees
+have been designated and the geolocation taken at a maximum distance of
+30 meters from designated trees (Laurent Descroix, ONF, personnal
+communication). To simulate secondary tracks, TROLL uses a loads map,
+measuring every trees at a distance of 30 meters for each pixel, and a
+track proximity maps of the closest existing track. Next, the model
+select the pixel with the highest load and closest track, find the
+closest existing track and join it by removing tree in the way with a
+width of 5 meters. The operations are repeated untill no felt trees are
+left.
+
+\subsubsection{Gap damages}\label{gap-damages}
+
+Most of models account long term damages due to selective logging with a
+10 years increased mortality \citep{Huth2004, Khler2004, Ruger2008}. We
+decided to model explicitly long term logging damages because of their
+localised nature through a gap damages model. We gathered data from
+Paracou dataset \textbf{Cite} in cenususes between 1988 and 1992 on
+Paracou harvested plots. Individuals were categorized between alive,
+dead, or recruited during the period. We measured each individual
+distance to the closest gap. We then used a bayesian approach to test
+the link between tree death in the four years following the log event
+and distance to the closest gap. We adapted the model from \textbf{Cite
+Herault} based on a disturbance index into.
+
+\begin{equation}
+  \begin{array}{c} 
+    Death_i \sim \mathcal{B}(P(Death_i)) \\
+    P(Death_i) = logit^{-1}(\theta + \beta*e^{\alpha*d_{gaps}}) = \frac{e^{\theta + \beta*e^{\alpha*d_{gaps}}}}{1 + e^{\theta + \beta*e^{\alpha*d_{gaps}}}}
   \end{array}
-  \label{eq:Deltadbh}
-\end{equation}
-
-Next, TROLL used the new trunk dimension (\(dbh\) and \(h\)) to update
-tree crown geometry using allometric equations \citep{Chave2005}:
-
-\begin{equation}
-  \begin{array}{c} \\
-    CR = 0.80 + 10.47*dbh - 3.33*dbh^2\\
-    CD = -0.48 + 0.26*h~;~CD = 0.13 + 0.17*h~(h<5~m)
-  \end{array}
-  \label{eq:C}
-\end{equation}
-
-Finally, TROLL computes the mean leaf density within the crown (\(LD\)
-in \(m^2.m^{-3}\)) assuming a uniform distribution:
-
-\begin{equation}
-  LD = \frac{LA_{young}+LA_{mature}+LA_{old}}{\pi*CR^2*CD}
-  \label{eq:LD}
-\end{equation}
-
-\subsection{Mortality}\label{mortality}
-
-Mortality is partitioned in three factors inside TROLL: background death
-\(d_b\), treefall death \(d_t\) and negative density dependent death
-\(d_{NDD}\). Because density dependent death \(d_{NDD}\) is still in
-development inside TROLL we did not used it, so we will not detail is
-computation.
-
-\citet{chave_towards_2009} advocated for a wood economics spectrum
-opposing fast growing light wood species species with high risk of
-mortality to slow growing dense wood species with reduced risk of
-mortality. Hence, background mortality is derived from wood specific
-gravity \(wsg\) inside TROLL:
-
-\begin{equation}
-  d_b = m*(1-\frac{wsg}{wsg_{lim}})+d_n
-  \label{eq:db}
-\end{equation}
-
-\(m\) (\(events.year^{-1}\)) is the reference background death rate for
-lighter wood species (pioneers). \(d_n\) represents death by
-carbohydrates shortage. If the number of consecutive day with
-\(NPP_{ind} < 0\) \eqref{eq:NPP} is superior to tree leaf lifespan \(d_n\)
-is set to 1 and remains null in other cases.
-
-Mortality by treefall inside TROLL depends on a specifric stochastic
-threshold \(\theta\):
-
-\begin{equation}
-  \theta = h_{max}*(1-v_T*|\zeta|)
-  \label{eq:theta}
-\end{equation}
-
-\(h_{max}\) is the maximal tree height. \(v_T\) is the variance term set
-to 0.3. \(|\zeta|\) is the absolute value of a random centered and
-scaled Gaussian. If the tree hieght \(h\) is superior to \(\theta\) then
-the tree may fall with a probability \(1-\theta/h\) \citep{Chave1999}.
-The treefall direction is random (drawn from a uniform law
-(\(\mathcal{U}[0,2\pi]\)). All tree in the trajectory of the falling
-tree will be hurted through a variable denoted \(hurt\), incremented by
-fallen tree height \(h\). If a tree height is inferior than its \(hurt\)
-values then it may die with a probability
-\(1-\frac{1}{2}\frac{h}{hurt}\). \(hurt\) variable is reset to null at
-each timestep (\(month\)).
-
-\subsection{Recruitment}\label{recruitment}
-
-Once the tree became fertile they will start to disperse seeds. TROLL
-consider tree as fertile after a specific height threshold
-\(h_{mature}\) \citep{Wright2005}:
-
-\begin{equation}
-  h_{mature} = -11.47+0.90*h_{max}
-  \label{eq:hmature}
-\end{equation}
-
-But TROLL is not considering seed directly through a seedbank, instead
-seed might be interpreted as a seedling recruitment opportunity. The
-number of reproduction opportunities per mature tree is denoted \(n_s\)
-and set to 10 for all species. This assumption originates from a
-trade-off between seed number and seed size resulting in equivalent
-survival and recruitment probability. All \(n_s\) events are dispersed
-with a distance randomly drawn from a Gaussian distribution.
-Additionally, TROLL model consider external seedrain through \(n_{ext}\)
-events of seed immigration:
-
-\begin{equation}
-  n_{ext} = N_{tot}*f_{reg}*n_{ha}
-  \label{eq:next}
-\end{equation}
-
-N\_\{tot\} is the external seedrain per hectare (number of reproduction
-opportunities). \(f_{reg}\) is the species regional frequency.
-\(n_{ha}\) is the simulated plot size in \(ha\).
-
-Finally, a bank of seedlings to be recruited is defined for each pixel.
-If the ground-level light reaches a species light compensation point
-\(LCP\) the species will be recruited:
-
-\begin{equation}
-  LCP = \frac{R_{leaf}}{\phi}
-  \label{eq:LCP}
-\end{equation}
-
-\(R_{leaf}\) is the leaf respiration for maintenance (see \eqref{eq:Rl}).
-\(\phi\) is the quantum yield (\(\mu mol C.\mu mol~photon\)) set to
-0.06. If several species reach their \(LCP\), one is picked at random.
-Seedlings are recruited with following intial geometry:
-
-\begin{equation}
-  \begin{array}{c} \\
-    dbh = \frac{a_h}{h_{max} - 1}\\
-    h = 1~m\\
-    CR = 0.5~m\\
-    CD = 0.3~m\\
-    LD = 0.8~m^2.^{-3}
-  \end{array}
-  \label{eq:C}
-\end{equation}
-
-\section{Sensitivity analysis}\label{sensitivity-analysis}
+  \label{eq:death}
+\end{equation}
+
+\(Death\) of a tree \(i\) follows a \(\mathcal{Bernoulli}\) law of
+probability \(P(Death_i)\). The odds for a tree \(i\) to die
+\(O(Death_i)\) is calculated by the sum of the natural tree death odd
+\(\theta\) and a perturbation index \(\beta*e^{\alpha*d_{gaps}}\). The
+perturbation index depend on the distance \(d_{gaps}\) of the tree \(i\)
+to the closest logging gap. The probability for a tree \(i\) to die
+\(P(Death_i)\) is finally calculated by taking the inverse logit of the
+odd \(logit^{-1}[O(Death_i)]\). \textbf{Add Appendix ?}
+
+\section{Material and Methods}\label{material-and-methods}
+
+\subsection{Sensitivity analysis}\label{sensitivity-analysis}
 
@@
 after disturbance in simulations.
-
-\subsection{Functional traits}\label{functional-traits}
 
@@
 
-\subsection{Seed rain}\label{seed-rain}
-
 To assess the sensitivity of TROLL model to seed rain, we performed a
@@
 
-\section{Disturbance}\label{disturbance}
-
-Most tropical timber production originates from selective logging, the
-targeted harvesting of timber from species of interest. Consequently,
-tropical sylviculture can be assimilated to a disturbance. The main
-difference between a disturbance and selective logging is the targetting
-of both species and individuals of interest. So we decided to first
-asses unselective disturbance effect on tropical forest ecosystem to
-subsequently better understand selective logging effect.
+\subsection{Disturbance}\label{disturbance-1}
+
+\emph{Most tropical timber production originates from selective logging,
+the targeted harvesting of timber from species of interest.
+Consequently, tropical sylviculture can be assimilated to a disturbance.
+The main difference between a disturbance and selective logging is the
+targetting of both species and individuals of interest. So we decided to
+first asses unselective disturbance effect on tropical forest ecosystem
+to subsequently better understand selective logging effect. We first
+implemented a simple disturbance module inside TROLL model to simulate
+unselective disturbance.} \textbf{Already inside model descripiton !}
 
@@
 
-\subsection{Model description}\label{model-description-1}
-
-Disturbance module was designed in the simplest way in order to relate
-the ecosystem answer to volume lost without any individuals nor species
-targetting. For a given iteration \(disturb_{iter}\), individuals are
-picked randomly with a uniform law on the number of trees. Selected
-individuals are then removed without trigerring a treefall to avoid any
-side effect. The operation is repeated untill the disturbance result in
-a defined lost basal area (\(disturb_{intensity}\) in \% of BA).
+\subsection{Selective logging}\label{selective-logging}
+
+Once the unselective effect of disturbance on forest ecosystem studied,
+we can focus on selective logging to highlight the effect of species and
+individuals targetting on the ecosystem answer. We first implemented a
+sylviculture module inside TROLL model to simulate selective logging in
+regards to french Guiana practices. We used the 60 communities from
+disturbance simulations to study the role of biodiversity in both forest
+ecosystem and commercial species answer to selective logging. We then
 
@@
 
-\section{Selective logging}\label{selective-logging}
-
-Once the unselective effect of disturbance on forest ecosystem studied,
-we can focus on selective logging to highlight the effect of species and
-individuals targetting on the ecosystem answer.
-
-\subsection{Model description}\label{model-description-2}
-
-\subsubsection{Designation}\label{designation}
-
-\subsubsection{Selection}\label{selection}
-
-\subsubsection{Rotten trees}\label{rotten-trees}
-
-\subsubsection{Felling}\label{felling}
-
-\subsubsection{Tracks}\label{tracks}
-
-\subsubsection{Gap damages}\label{gap-damages}
-
-\subsection{Design of experiment}\label{design-of-experiment-1}
-
-\subsection{Ecosystem answer
-anlaysis}\label{ecosystem-answer-anlaysis-1}
-
-Similar to disturbance
-
 \section{Results}\label{results}
@@
 
-\subsection{Disturbance}\label{disturbance-1}
-
-\subsection{Sylviculture}\label{sylviculture}
+\subsection{Disturbance}\label{disturbance-2}
+
+\subsection{Sylviculture}\label{sylviculture-1}
 
@@
 
+\appendix
+
+\hypertarget{appendix-1}{\section{Appendix 1}\label{appendix-1}}
+
+I this Appendix we further detail modules of TROLL model.
+
+\subsection{Abiotic environment}\label{abiotic-environment}
+
+A voxel space, with a resolution of 1 \(m^3\), is used to explicilty
+model the abiotic environment. For each tree crown, leaf area density is
+calculated on tree geometry assuming a uniform distriution across voxels
+occupied by the crown. Leaf area density is computed within each voxel
+summing all tree crowns inside the voxel \(v\), and is denoted
+\(LAD(v)\) (leaf area per voxel in \(m².m^{-3}\)). The vertical sum of
+\(LAD\) from voxel \(v\) to the ground level defines \(LAI(v)\) (leaf
+area per fround area in \(m^2.m^{-2}\) commonly called leaf area index):
+
+\begin{equation}
+  LAI(v) = \sum _{v'=v} ^\infty LAD(v') 
+  \label{eq:LAI}
+\end{equation}
+
+Daily variations in light intensity (photosynthetic photon flux density
+PPFD in \(\mu mol_{photons}.m^{-2}.s^{-1}\)), temperature (T in degrees
+Celsius), and vapor pressure deficit (VPD in \(kPA\)) are computed to
+assess carbon assimilation within each voxel of the canopy and for a
+representative day per month (see Appendix 1 from \citet{Li} for further
+details). Variation of PPFD Within the canopy is calculated as a loacal
+Beer-Lambert extinction law:
+
+\begin{equation}
+  PPFD_{max,month}(v) = PPFD_{top,max,month}*e^{-k*LAI(v)}
+  \label{eq:PPFD}
+\end{equation}
+
+The daily maximum incident PPFD at the top of canopy
+\(PPFD_{top,max,month}\) is given as input. The extinction rate \(k\) is
+assumed as constant, besides is variation with zenith angle and species
+leaf inclination angle \citep{Meir2000}. Moreover only vertical light
+diffusion is considered ignoring lateral light diffusion, which can have
+an important role especially in logging gaps. Finally, intra-day
+variation at half hour time steps \(t\) for a representative day every
+month are used to compute \(PPFD_{month}(v,t)\), \(T_{month}(v,t)\) and
+\(VPD_{month}(v,t)\). Water and nutrient process both in soil and inside
+trees are not simulated.
+
+\subsection{Photosynthesis}\label{photosynthesis}
+
+\subsubsection{Theory}\label{theory}
+
+Troll simulates the carbon uptake of each individual with the Farquhar,
+von Caemmerer and Berry model of C3 photosynthesis \citep{Farquhar1980}.
+Gross carbon assimilation rate (\(A\) in
+\(\mu mol~CO_2. m^{-2}.s^{-1}\)) will be the minimum of eiter Rubisco
+activity (\(A_v\)) or RuBP generation (\(A_j\)):
+
+\begin{equation}
+  A=min(A_v, A_j)~|~A_v=V_{cmax}*\frac{c_i-\Gamma^*}{c_i+K_m}~;~A_j=\frac{J}{4}*\frac{c_i-\Gamma^*}{c_i+2*\Gamma^*}
+  \label{eq:A}
+\end{equation}
+
+\(V_{cmax}\) is the maximum rate of carboxylation
+(\(\mu mol~CO_2.m^{-2}.s^{-1}\)). \(c_i\) is the \(CO_2\) partial
+pressure at carboxylation sites. \(\Gamma^*\) is the \(CO_2\)
+compensation point in absence of dark respiration. \(K_m\) is the
+apparent knietic constant of the Rubisco. And \(J\) is the electron
+transport rate (\(\mu mol e^-.m^{-2}.s^{-1}\)). \(J\) depends on the
+light intensity with \(PPFD\):
+
+\begin{equation}
+  J = \frac{1}{2*\theta}*[\alpha*PPFD+J_{max}-\sqrt{(\alpha*PPFD+J_{max})^2}-4*\theta*\alpha*PPFD*J_{max}]
+  \label{eq:J}
+\end{equation}
+
+\(J_{max}\) is the maximal electron transport capacity
+(\(\mu mol e^-.m^{-2}.s^{-1}\)). \(\theta\) is the curvature factor. And
+\(\alpha\) is the apparent quantum yield to electron transport
+(\(mole^-.mol~photons^{-1}\)).
+
+Carbon assimilation by photosynthesis will then be limited by the
+\(CO_2\) partial pressure at carboxylation sites. Stomata controls the
+gas concentration at carboxylation sites throught stomatal transport:
+
+\begin{equation}
+  A = g_s*(c_a-c_i)
+  \label{eq:Ag}
+\end{equation}
+
+\(g_s\) is the stomatal conductance to \(CO_2\)
+(\(molCO_2.m^{-2}.s^{-1}\)). TROLL simulates stomatal conductance
+\(g_s\) with the model from \citep{Medlyn2011}:
+
+\begin{equation}
+  g_s = g_0 + (1 + \frac{g_1}{\sqrt{VPD}})*\frac{A}{c_a}
+  \label{eq:gs}
+\end{equation}
+
+\(g_0\) and \(g_1\) are parameters from the model. TROLL model assume
+\(g_0 \approx 0\) (empirically tested and considered as reasonable).
+
+\subsubsection{Parametrization}\label{parametrization}
+
+Leaf traits can be used as proxy of photosynthesis, especially leaf
+nutrient content which directly play a role in it
+\citep{wright_worldwide_2004}. \citet{Domingues2010} suggested that
+\(V_{cmac}\) and \(J_{max}\) were both limited by the leaf concentration
+of nitrogen \(N\) and phosphorus \(P\) (\(mg.g^{-1}\)):
+
+\begin{equation}
+  log_{10} V_{cmax-M} = min( 
+  \begin{array}{c} 
+    -1.56+0.43*log_{10} N-0.37*log_{10} LMA \\
+    -0.80+0.45*log_{10} P-0.25*log_{10} LMA 
+  \end{array} 
+  )
+  \label{eq:VcmaxM}
+\end{equation}
+
+\begin{equation}
+  log_{10} J_{max-M} = min(
+  \begin{array}{c} 
+    -1.50+0.41*log_{10} N-0.45*log_{10} LMA \\
+    -0.74+0.44*log_{10} P-0.32*log_{10} LMA 
+  \end{array}
+  )
+  \label{eq:JmaxM}
+\end{equation}
+
+\(V_{cmax-M}\) and \(J_{max-M}\) are the photosynthetic capacities at
+\(25^\circ C\) of mature leaves per leaf dry mass (resp.
+\(\mu mol CO_2.g^-1.s^{-1}\) and \(\mu mol e^-.g^{-1}.s^{-1}\)). \(LMA\)
+is the leaf mass per are (\(g.cm^{-2}\)). \(V_{cmax}\) and \(J_{max}\)
+are calculated by multiplying \(V_{cmax-M}\) and \(J_{max-M}\) by
+\(LMA\). \(V_{cmax}\) and \(J_{max}\) variation with temperature are
+caluclated with \citet{Bernacchi2003} (see Appendix 2 from \citet{Li}
+for further details).
+
+TROLL computes leaf carbon assimilation \(A_l\) combining equations from
+\eqref{eq:A} to \eqref{eq:JmaxM} for each tree crown voxel within in each
+crown layer \(l\):
+
+\begin{equation}
+  A_l = \frac{1}{n_v*t_M} * \sum_v  \sum^{t_M}_{t=1} A(PPFD_{month}(v,t),VPD_{month}(v,t),T_{month}(v,t))
+  \label{eq:Al}
+\end{equation}
+
+\(PPFD_{month}(v,t)\), \(VPD_{month}(v,t)\) , and \(T_{month}(v,t)\) are
+derived from microclimatic data. \(n_v\) is the number of voxels within
+crown layer \(l\). And the sum is calculated over the \(t_M\)
+half-hourly intervals \(t\) of a tipical day.
+
+\subsection{Autotrophic respiration}\label{autotrophic-respiration}
+
+A large fraction of plants carbon uptake is actually used for plant
+maintenance and growth respiration. The autotrophic respiration can
+represents up to 65\% of the gross primary productivity but varies
+strongly among species, sites, and environnements.
+
+TROLL uses \citet{Atkin2015} database of mature leaf dark respiration
+and associated leaf traits to compute leaf maintenance respiration:
+
+\begin{equation}
+  R_{leaf-M} = 8.5431-0.1306*N-0.5670*P-0.0137*LMA+11.1*V_{cmax-M}+0.1876*N*P
+  \label{eq:Rl}
+\end{equation}
+
+\(R_{leaf-M}\) si the dark respiration rate per leaf dry mass at a
+temperaure of \(25^\circ C\) (\(nmolCO_2.g^{-1}.s^{-1}\)). The other
+terms are in equations \eqref{eq:VcmaxM} and \eqref{eq:JmaxM}. TROLL assume
+leaf respiration during day light to be 40\% of leaf dark respiration,
+and computes total leaf respiration by accounting for the legnth of
+daylight.
+
+TROLL model stem respiration (\(R_{stem}\) in \(\mu molC.s^{-1}\)) with
+a constant respiration rate per volume of sapwood:
+
+\begin{equation}
+  R_{stem} = 39.6*\pi*ST*(dbh-ST)*(h-CD)
+  \label{eq:Rs}
+\end{equation}
+
+dbh, h, CD and ST are tree diameter at breast height, height, corwn
+depth and sapwoond thickness, respectively (\(m\)). TROLL assumes
+\(ST=0.04~m\) when \(dbh>30~cm\) and an increasing \(ST\) for lower
+\(dbh\).
+
+Finally, TROLL computes both fine root maintenance respiration, as half
+of the leaf maintenance respiration. Whereas coarse root and branch
+maintenance respiration is computed as half of the stem respiration. And
+growth respiration (\(R_{growth}\)) is assumed to account for 25\% of
+the gross primary productivity minus the sum of maintenance
+respirations.
+
+\subsection{Net carbon uptake}\label{net-carbon-uptake}
+
+Net primary production of carbon for one individual \(NPP_{ind}\)
+(\(gC\)) is computed by the balance between gross primary production
+\(GPP_{ind}\) and respirations \(R\):
+
+\begin{equation}
+  NPP_{ind} = GPP_{ind} - R_{maintenance} - R_{growth}
+  \label{eq:NPP}
+\end{equation}
+
+TROLL partitions individuals total leaf area \(LA\) into three pools for
+different leaf age classes corresponding to different photosynthesis
+efficiency (young, mature and old leaves with \(LA_{young}\),
+\(LA_{mature}\), and \(LA_{old}\) respectivelly). Consequently we can
+compute growth primary production for one individual as:
+
+\begin{equation}
+  GPP_{ind} = 189.3 * \Delta t * \sum _{l= \lfloor h-CD \rfloor +1} ^{\lfloor h \rfloor} [A_l] * (\frac{LA_{young}}{2} + LA_{mature} + \frac{LA_{old}}{2})
+  \label{eq:GPP}
+\end{equation}
+
+h and CD are tree height and crown depth, repectivelly (\(m\)).
+\(\lfloor x \rfloor\) is the rounding function. \(\Delta t\) is the
+duration of a timestep (\(year\)).
+
+Thus, TROLL can compute carbon allocation to wood into an increment of
+stem volume \(\Delta V\) (\(m^3\)):
+
+\begin{equation}
+  \Delta V = 10^{-6} * \frac{f_{wood}*NPP_{ind}}{0.5*wsg}*Senesc(dbh)
+  \label{eq:DeltaV}
+\end{equation}
+
+\(f_{wood}\) is the fixed fraction of NPP allocated to stem and
+branches. \(wsg\) is the wood specific gravity (\(g.cm^{-3}\), see
+\ref{tab:traits}). TROLL assume large trees less efficient to convert
+NPP as growth by using a size-related growth decline with function
+\(Senesc\) after a specific diameter at brest height threshold
+\(dbh_{thresh}\):
+
+\begin{equation}
+  Senesc(dbh) = max(0;3-2*\frac{dbh}{dbh_{thresh}})
+  \label{eq:Senesc}
+\end{equation}
+
+Finally, TROLL can compute carbon allocation to canopy with canopy NPP
+fraction denoted \(f_{canopy}\) and decomposed into leaf, twig and fruit
+production. Carbon allocation to leaf results in a new young leaf pool,
+whereas other leaf pools are updated as follow:
+
+\begin{equation}
+  \begin{array}{c} \\
+   \Delta LA_{young} = \frac{2*f_{leaves}*NPP_{ind}}{LMA}-\frac{LA_{young}}{\tau_{young}} \\
+  \Delta LA_{mature} = \frac{LA_{young}}{\tau_{young}} - \frac{LA_{mature}}{\tau_{mature}}\\
+  \Delta LA_{old} = \frac{LA_{mature}}{\tau_{mature}} - \frac{LA_{old}}{\tau_{old}}
+  \end{array}
+  \label{eq:DeltaLA}
+\end{equation}
+
+\(\tau_{young}\), \(\tau_{mature}\), and \(\tau_{old}\) are species
+residence times in each leaf pools (\(years\)). The sum of residency
+time thus defined the leaf lifespan
+\(LL = \tau_{young} + \tau_{mature} + \tau_{old}\) (\(years\)).
+\(\tau_{young}\) is set to one month and \(\tau_{mature}\) is set to a
+third of leaf lifespan \(LL\). Belowground carbon allocation is not
+simulated inside TROLL.
+
+\subsection{Leaf lifespan}\label{leaf-lifespan-1}
+
+The underestimation of leaf lifespan for low LMA species with the
+allometry from \citet{Reich1991a} resulted in indivduals unealistic
+early death from carbon starvation. We gathered data from
+TRY\citep{Kattge2011}, DRYAD \citep{chave_towards_2009} and GLOPNET
+\citep{wright_worldwide_2004} datasets. We used an out of the bag method
+applied on a random forest to select variables with highest importance
+to explain leaf lifespan. We thus selected leaf mass per are \(LMA\),
+leaf nitrogen content \(N\) and wood specific gravity \(wsg\). We then
+used a bayesian approach to test different models with growing level of
+complexity. The model with the best tradeoff between complexity (number
+of parameters), convergence, likelihood, and prediction quality (root
+mean square error of prediction RMSEP) was kept. We selected following
+model with a maximum likelihood of 13.6 and a RMSEP of 12 months:
+
+\begin{equation}
+  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
+  \label{eq:LLb}
+\end{equation}
+
+Leaf lifespan \(LL\) follows a lognormal law with location infered from
+leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
+\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
+normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
+\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
+presemption following a gamma law. \(d\) represents the dataset random
+effects and encompass environmental and protocol variations. The
+sampling of model \eqref{eq:LLb} resulted in the following allometry:
+
+\begin{equation}
+  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
+  \label{eq:LL}
+\end{equation}
+
+\subsection{Tree growth}\label{tree-growth}
+
+Once the increment of stem volume \(\Delta V\) calculated with equation
+\eqref{eq:DeltaV}, TROLL convert it into an increment of tree diameter at
+breast height denoted \(\Delta dbh\). TROLL infer tree height from
+\(dbh\) using a Michaelis-Menten equation:
+
+\begin{equation}
+  h = h_{lim}*\frac{dbh}{dbh + a_h}
+  \label{eq:h}
+\end{equation}
+
+On the other hand, we have the trunk volume
+\(V = C * \pi * (\frac{dbh}{2})^2*h\), thus:
+
+\begin{equation}
+  \begin{array}{c} \\
+    \Delta V = C*\frac{1}{2}*\pi*h*dbh*\Delta dbh + C * \pi * (\frac{dbh}{2})^2*h \\
+    \Delta V = V*\frac{\Delta dbh}{dbh}*(3-\frac{dbh}{dbh + ah})
+  \end{array}
+  \label{eq:Deltadbh}
+\end{equation}
+
+Next, TROLL used the new trunk dimension (\(dbh\) and \(h\)) to update
+tree crown geometry using allometric equations \citep{Chave2005}:
+
+\begin{equation}
+  \begin{array}{c} \\
+    CR = 0.80 + 10.47*dbh - 3.33*dbh^2\\
+    CD = -0.48 + 0.26*h~;~CD = 0.13 + 0.17*h~(h<5~m)
+  \end{array}
+  \label{eq:C}
+\end{equation}
+
+Finally, TROLL computes the mean leaf density within the crown (\(LD\)
+in \(m^2.m^{-3}\)) assuming a uniform distribution:
+
+\begin{equation}
+  LD = \frac{LA_{young}+LA_{mature}+LA_{old}}{\pi*CR^2*CD}
+  \label{eq:LD}
+\end{equation}
+
+\subsection{Mortality}\label{mortality}
+
+Mortality is partitioned in three factors inside TROLL: background death
+\(d_b\), treefall death \(d_t\) and negative density dependent death
+\(d_{NDD}\). Because density dependent death \(d_{NDD}\) is still in
+development inside TROLL we did not used it, so we will not detail is
+computation.
+
+\citet{chave_towards_2009} advocated for a wood economics spectrum
+opposing fast growing light wood species species with high risk of
+mortality to slow growing dense wood species with reduced risk of
+mortality. Hence, background mortality is derived from wood specific
+gravity \(wsg\) inside TROLL:
+
+\begin{equation}
+  d_b = m*(1-\frac{wsg}{wsg_{lim}})+d_n
+  \label{eq:db}
+\end{equation}
+
+\(m\) (\(events.year^{-1}\)) is the reference background death rate for
+lighter wood species (pioneers). \(d_n\) represents death by
+carbohydrates shortage. If the number of consecutive day with
+\(NPP_{ind} < 0\) \eqref{eq:NPP} is superior to tree leaf lifespan \(d_n\)
+is set to 1 and remains null in other cases.
+
+Mortality by treefall inside TROLL depends on a specifric stochastic
+threshold \(\theta\):
+
+\begin{equation}
+  \theta = h_{max}*(1-v_T*|\zeta|)
+  \label{eq:theta}
+\end{equation}
+
+\(h_{max}\) is the maximal tree height. \(v_T\) is the variance term set
+to 0.3. \(|\zeta|\) is the absolute value of a random centered and
+scaled Gaussian. If the tree hieght \(h\) is superior to \(\theta\) then
+the tree may fall with a probability \(1-\theta/h\) \citep{Chave1999}.
+The treefall direction is random (drawn from a uniform law
+(\(\mathcal{U}[0,2\pi]\)). All tree in the trajectory of the falling
+tree will be hurted through a variable denoted \(hurt\), incremented by
+fallen tree height \(h\). If a tree height is inferior than its \(hurt\)
+values then it may die with a probability
+\(1-\frac{1}{2}\frac{h}{hurt}\). \(hurt\) variable is reset to null at
+each timestep (\(month\)).
+
+\subsection{Recruitment}\label{recruitment}
+
+Once the tree became fertile they will start to disperse seeds. TROLL
+consider tree as fertile after a specific height threshold
+\(h_{mature}\) \citep{Wright2005}:
+
+\begin{equation}
+  h_{mature} = -11.47+0.90*h_{max}
+  \label{eq:hmature}
+\end{equation}
+
+But TROLL is not considering seed directly through a seedbank, instead
+seed might be interpreted as a seedling recruitment opportunity. The
+number of reproduction opportunities per mature tree is denoted \(n_s\)
+and set to 10 for all species. This assumption originates from a
+trade-off between seed number and seed size resulting in equivalent
+survival and recruitment probability. All \(n_s\) events are dispersed
+with a distance randomly drawn from a Gaussian distribution.
+Additionally, TROLL model consider external seedrain through \(n_{ext}\)
+events of seed immigration:
+
+\begin{equation}
+  n_{ext} = N_{tot}*f_{reg}*n_{ha}
+  \label{eq:next}
+\end{equation}
+
+N\_\{tot\} is the external seedrain per hectare (number of reproduction
+opportunities). \(f_{reg}\) is the species regional frequency.
+\(n_{ha}\) is the simulated plot size in \(ha\).
+
+Finally, a bank of seedlings to be recruited is defined for each pixel.
+If the ground-level light reaches a species light compensation point
+\(LCP\) the species will be recruited:
+
+\begin{equation}
+  LCP = \frac{R_{leaf}}{\phi}
+  \label{eq:LCP}
+\end{equation}
+
+\(R_{leaf}\) is the leaf respiration for maintenance (see \eqref{eq:Rl}).
+\(\phi\) is the quantum yield (\(\mu mol C.\mu mol~photon\)) set to
+0.06. If several species reach their \(LCP\), one is picked at random.
+Seedlings are recruited with following intial geometry:
+
+\begin{equation}
+  \begin{array}{c} \\
+    dbh = \frac{a_h}{h_{max} - 1}\\
+    h = 1~m\\
+    CR = 0.5~m\\
+    CD = 0.3~m\\
+    LD = 0.8~m^2.^{-3}
+  \end{array}
+  \label{eq:C}
+\end{equation}
+
+\addcontentsline{toc}{section}{References}
 \bibliography{/home/sylvain/Documents/Bibliography/library.bib}

### 2017-06-08 03-Model.Rmd
Model and Appendix

@@
 The source code is written in C++ and available upon request.
-All modules of TROLL models are further detailed in [Appendix 1].
+All modules of TROLL models are further detailed in [Appendix 1: TROLL model].
 
@@
 
-Previous implementation of TROLL model used @Reich1991a allometry to infer leaf lifespan $LL$ from species leaf mass per area $LMA$ [@Li, see [Appendix 1]].
+Previous implementation of TROLL model used @Reich1991a allometry to infer leaf lifespan $LL$ from species leaf mass per area $LMA$ [@Li, see [Appendix 1: TROLL model]].
 But the use of the allometrie from @Reich1991a with current implementation of the TROLL model resulted in an underestimation of leaf lifespan for low LMA species.
@@
 All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without presemption following a gamma law.
-$d$ represents the dataset random effects and encompass environmental and protocol variations.
-**Add Appendix ?**
+$d$ represents the dataset random effects and encompass environmental and protocol variations (see [Appendix 2: Leaf lifespan model] for more details).
 
@@
 
-**Add Appendix ?**
+(see [Appendix 3: Rotten tree model] for more details)
 
@@
 The perturbation index depend on the distance $d_{gaps}$ of the tree $i$ to the closest logging gap.
-The probability for a tree $i$ to die $P(Death_i)$ is finally calculated by taking the inverse logit of the odd $logit^{-1}[O(Death_i)]$.
-**Add Appendix ?**
+The probability for a tree $i$ to die $P(Death_i)$ is finally calculated by taking the inverse logit of the odd $logit^{-1}[O(Death_i)]$ (see [Appendix 4: Gap damages model] for more details).

### 2017-06-08 04-MnMs.Rmd
Model and Appendix

@@
 @Li already assessed TROLL model sensitivity to several parameters ($k$ see \@ref(eq:PPFD), $\phi$ see \@ref(eq:LCP), $g1$ see \@ref(eq:gs), $f_{wood}$ see \@ref(eq:DeltaV), $f_{canopy}$ see \@ref(eq:DeltaLA) and $m$ see \@ref(eq:db)) which they assumed having a key role in model functioning.
-
-On the other hand, we decided to use TROLL to study resistance and resilience of ecosystem face to disturbance, highlighting the role of biodiversit.
+On the other hand, we decided to use TROLL to study resistance and resilience of ecosystem face to disturbance, highlighting the role of biodiversity.
 Consequently we particulrarly needed to assess the importance of functional traits to further better control and evaluate functional diversities.
-We also needed to assess the sensitivity of TROLL model to the seed rain constant ($n_{ext}$, see \@ref(eq:next)) because we assume it is one of the main factor of tree recruitments after disturbance in simulations.
+We also needed to assess the sensitivity of TROLL model to the seed rain constant ($n_{ext}$, see \@ref(eq:next)) because we assumed it as one of the main factors of tree recruitments after disturbance within simulations.
 
@@
 Each trait was tested independently.
-We reduce to a common mean traits with a Perason's correlation value $r \geq 0.8$ ($h_{max}$ and $a_h$ with a correlation of $r=0.98$).
-
-To assess the sensitivity of TROLL model to seed rain, we performed a sensitivity analysis by fixing simulations seed rain constant to 2, 20, 200 and 2000 seeds per hectare (default value being 200 seeds per hectare).
+We reduce to a common mean traits with a Pearson's correlation value $r \geq 0.8$ ($h_{max}$ and $a_h$ with a correlation of $r=0.98$).
+To assess the sensitivity of TROLL model to seed rain, we performed a sensitivity analysis by fixing simulations seed rain constant to 2, 20, 200 and 2000 seeds per hectare.
 
@@
 
-## Disturbance
-
-*Most tropical timber production originates from selective logging, the targeted harvesting of timber from species of interest.
-Consequently, tropical sylviculture can be assimilated to a disturbance.
-The main difference between a disturbance and selective logging is the targetting of both species and individuals of interest.
-So we decided to first asses unselective disturbance effect on tropical forest ecosystem to subsequently better understand selective logging effect.
-We first implemented a simple disturbance module inside TROLL model to simulate unselective disturbance.* **Already inside model descripiton !**
-
-We first implemented a disturbance module inside TROLL model to simulate unselective disturbance.
-60 tree communities were then defined with different levels of both specific and functional richness to explore the role of biodiversity on tropical forest ecosystem answer to disturbance.
-Next, different levels of disturbance were simulated on previous mature communities.
-Finally, resilience of ecosystem major global variables (related to carbon stock, forest dynamic and florisitic structure) from simulations were partitioned between selection effects and niche complementarity [@Loreau2001].
-
-## Selective logging
-
-Once the unselective effect of disturbance on forest ecosystem studied, we can focus on selective logging to highlight the effect of species and individuals targetting on the ecosystem answer.
-We first implemented a sylviculture module inside TROLL model to simulate selective logging in regards to french Guiana practices.
-We used the 60 communities from disturbance simulations to study the role of biodiversity in both forest ecosystem and commercial species answer to selective logging.
-We then
-
 ## Design of experiment
 
-In order to assess the role of biodiversity in ecosystem answer to disturbance, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time.
-Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%).
-Biodiversity was integrated with two components specific and functional diversities.
-We used species richness $SR$ to represents species diversity (5, 25, and 125).
+In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time.
+Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation.
+Biodiversity was integrated with two components of its components: taxonomic and functional diversities.
+We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species).
 Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean.
@@
 This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]).
-Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the forest mature were simulated, and thus before disturbance.
+Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the mature forest were simulated, and thus before disturbance.
 We obtained a broad range of both functionl dispersion $FDis$ and aboveground biomass $AGB$ for simulated forest ecosystems before disturbance.
@@
 Communities are implemented along a gradient of species richness (SR) and functional dispersion (FDis) resulting in a broad range of aboveground biomass (AGB).
-FDis was caluclated based on 4 functional traits (leaf mass per area, wood specific gravity, maximum diameter, maximum height).', cache=TRUE}
+FDis was caluclated based on 4 functional traits (leaf mass per area, wood specific gravity, maximum diameter, and maximum height).', cache=TRUE}
 # Data prep
@@
 
-## Ecosystem answer anlaysis
+## Ecosystem answer analysis
+
+### Resilience
 
 Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions.
-We focused on few functions and related metrics to analyse ecosystem answer to disturbance: carbon stock with aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), forest dynamic with number of stem above $10~cm$ diameter at breast height ($N10$), and floristic composition with **?** (**?** in **?**).
+We decided to describe simiulated tropical forests in three major functions: forest dynamic, forest production, and forest diversity.
+Forest dynamic was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$).
+Forest production was represented by growth pimary productivity ($GPP$ in $MgC.ha^{-1}$), net pimary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
+Finally, forest diversity was represented by **??**.
 
@@
 \begin{equation}
-  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X(t)}{X(t_d-1)}
+  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X_T(t)}{X_C(t)}
   (\#eq:Resilience)
@@
 The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$.
-We transformed the equation in the resilience of the system $R(t)$ at the time $t$ being the ratio of the ecosystem metric $X(t)$ at time $t$ to the ecosystem metric before the disturbance happened $X(t_d-1)$ at time $t_d-1$.
-$X(t_d-1)$ was calculated as the mean of the ecosystem output for 50 last years of the simulation of the mature forest (over 600 years).
-Then, we calculated for each simulation the recovery time $t_{recovery}$ were the ecosystem reached back its stable state defined as its state before disturbance.
-But some ecosystem metrics can reach pre disturbance value wihtout revealing ecosystem recovery.
-For instance, number of stem above $10~cm$ diameter at breast height $N10$ will first decrease due to disturbance.
-Then $N10$ will exceed its initial value due to new seedlings recruitment before decreasing towards its pre disturbanec value.
-Consequently, the first time $N10$ reaches back its pre disturbance value can not be considered as the recovery time $t_{recovery}$.
-In order to solve this issue, we considered the ecosystem as stable again when its whole set of observed metrics reach their pre disturbance values.
+But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic.
+Consequently, to encompass undisturbed ecosystem variations throught time, we simulated an undisturbed control ecosystem $C$.
+And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the the ecosystem metric values from the control $X_C(t)$ .
+Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics.
+$R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value.
+Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$.
 
-Biodiversity is not only a facet of the experimental design and an ecosystem output through florisitic composition, but also interact on ecosystem functioning and consequently on its answer to disturbance.
+### Biodiversity effect
+
+Biodiversity is not only a facet of the experimental design and an ecosystem output through forest diversity, but also interact on ecosystem functioning and consequently on its answer to disturbance.
 Biodiversity ecosystem functioning relation can be split in complementarity and selection effect with @Loreau2001 partitioning:

### 2017-06-08 07-Appendix1.Rmd
Model and Appendix

@@
 
-# Appendix 1
+# Appendix 1: TROLL model
 
@@
 
-## Leaf lifespan
-
-The underestimation of leaf lifespan for low LMA species with the allometry from @Reich1991a resulted in indivduals unealistic early death from carbon starvation.
-We gathered data from TRY[@Kattge2011], DRYAD [@chave_towards_2009] and GLOPNET [@wright_worldwide_2004] datasets.
-We used an out of the bag method applied on a random forest to select variables with highest importance to explain leaf lifespan.
-We thus selected leaf mass per are $LMA$, leaf nitrogen content $N$ and wood specific gravity $wsg$.
-We then used a bayesian approach to test different models with growing level of complexity.
-The model with the best tradeoff between complexity (number of parameters), convergence, likelihood, and prediction quality (root mean square error of prediction RMSEP) was kept.
-We selected following model with a maximum likelihood of 13.6 and a RMSEP of 12 months:
-
-\begin{equation}
-  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
-  (\#eq:LLb)
-\end{equation}
-
-Leaf lifespan $LL$ follows a lognormal law with location infered from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
-Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$.
-All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without presemption following a gamma law.
-$d$ represents the dataset random effects and encompass environmental and protocol variations.
-The sampling of model \@ref(eq:LLb) resulted in the following allometry:
-
-\begin{equation}
-  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
-  (\#eq:LL)
-\end{equation}
-
 ## Tree growth

### 2017-06-08 08-Appendix2.Rmd
Model and Appendix

# Appendix 2: Leaf lifespan model

TROLL model previous implementation encompass Reich's 1991 and 1997 and Wright's 2004 allometries to estimate leaf lifespan with  [@Reich1991a; @Reich1997; @wright_worldwide_2004]. But we have shown that Reich's allometries are underestimating leaf lifespan for low LMA species. Moreover simulations estimated unrealistically low aboveground biomass for low LMA species. We assumed Reich's allometries underestimation of leaf lifespan for low LMA species being the source of unrealistically low aboveground biomass inside TROLL simulations. We decided to find a better allometry with @wright_worldwide_2004 GLOPNET dataset.

## Material and methods

We gathered data from TRY database to evaluate and improve previous models. We requested functional traits available in Bridge dataset and leaf lifespan for tropical species present in GLOPNET dataset and used in TROLL (see \@ref(tab:A2traits)). We compiled GLOPNET, TRY, and DRYAD databases. We kept dataset given by GLOPNET as origin dataset for observations. It correspond to leaf lifespan and also to LMA and Nmass in most of cases. We measured variable importance in functionnal traits to explain leaf lifespan with out-of bag method applied on a random forest.

```{r A2traits, echo=FALSE}
knitr::kable(data.frame(
  Name = c('LL', 'SLA', 'N', 'P', 'wsg'),
  Trait = c('Leaf lifespan (longevity)',
            'Leaf area per leaf dry mass (specific leaf area, SLA)',
            'Leaf nitrogen (N) content per leaf dry mass',
            'Leaf phosphorus (P) content per leaf dry mass',
            'Stem dry mass per stem fresh volume (stem specific density)'),
  Unit = c('month',' $m^2.kg^{-1}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$mg.mm^{-3}$'),
  TRYcode = c(12, 11, 15, 14, 4)),
  caption = 'Functional traits gathered with TRY.')
```

```{r A2data, include=FALSE, echo=FALSE, cache=TRUE}
library(readr)
library(reshape2)
library(BIOMASS)
LES <- read.csv("~/Documents/ECOFOG/TROLL/inst/extdata/LL/GLOPNET.csv", skip = 10)
LES <- LES[which(LES$BIOME == 'TROP_RF'),]
LES <- LES[which(LES$GF == 'T'),]
LES <- LES[-which(is.na(LES$log.LL)),]
data <- cbind(LES[c('Dataset', 'Species')],
              10^LES[paste0('log.',c('LL', 'LMA', 'Nmass', 'Narea',
                                     'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea'))])
names(data) <- c('Dataset', 'Species', 'LL', 'LMA', 'Nmass', 'Narea',
              'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea')
data <- droplevels(data[c('Dataset', 'Species', 'LL', 'LMA', 'Nmass')])
TRY <- read_delim("~/Documents/ECOFOG/TROLL/inst/extdata/LL/TRY3.txt",
    "\t", escape_double = FALSE, trim_ws = TRUE)
TRY <- data.frame(TRY[c('Dataset', 'SpeciesName', 'TraitName', 'StdValue')])
TRY <- TRY[-which(is.na(TRY$TraitName)),]
traits <- c('SLA', 'Nm', 'Pm', 'LL', 'wsg')
names(traits) <- unique(TRY$TraitName)
TRY$TraitName <- traits[TRY$TraitName] ; rm(traits)
K_TRY <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/LL/Kitajima.csv")
K_TRY <- data.frame(K_TRY[c('Species', 'Leaf lifespan (months)', 'LMA (g m−2)')])
names(K_TRY) <- c('SpeciesName', 'LL', 'SLA')
K_TRY$SLA <- (1/K_TRY$SLA)*1000
K_TRY <- melt(K_TRY, id.var = c('SpeciesName'))
names(K_TRY) <- c('SpeciesName', 'TraitName', 'StdValue')
Dataset <- rep('Kitajima', dim(K_TRY)[1])
K_TRY <- cbind(Dataset, K_TRY) ; rm(Dataset)
TRY <- rbind(TRY, K_TRY) ; rm(K_TRY)
TRY <- dcast(TRY, Dataset + SpeciesName ~ TraitName,
               value.var = 'StdValue', fun.aggregate = mean, na.rm = T)
TRY <- TRY[which(!is.na(TRY$LL)),]
TRY <- TRY[c('Dataset', 'SpeciesName', 'LL', 'SLA', 'Nm', 'wsg')]
names(TRY) <- c('Dataset', 'Species', 'LL', 'LMA', 'Nm', 'wsg')
TRY$LMA <- 1/TRY$LMA*1000
TRY$Nm <- TRY$Nm/10
TRY$LMA2 <- data$LMA[match(TRY$Species, data$Species)]
TRY$Nm2 <- data$Nmass[match(TRY$Species, data$Species)]
TRY$wsg2 <- apply(do.call('rbind', strsplit(TRY$Species, ' ', fixed = T)),
                   1, function(x) getWoodDensity(x[1],
                                                 x[2], region = 'SouthAmericaTrop')$meanWD)
TRY$LMA <- apply(TRY[c('LMA', 'LMA2')], 1, mean, na.rm = T)
TRY$Nm <- apply(TRY[c('Nm', 'Nm2')], 1, mean, na.rm = T)
TRY$wsg <- apply(TRY[c('wsg', 'wsg2')], 1, mean, na.rm = T)
data <- TRY[c("Dataset", "Species", "LL", "LMA", "Nm", "wsg")] ; rm(TRY)
na <- c(which(is.na(data$LMA)), which(is.na(data$Nm)), which(is.na(data$wsg)))
data <- data[-na,]; rm(na)
LMAmax <- max(data$LMA) ; Nmax <- max(data$Nm) ; wsgmax <- max(data$wsg)
data$Dataset <- as.factor(data$Dataset)
```

```{r functions, echo=FALSE}
fitgraph <- function(fit){
  pars <- fit@model_pars
  beta <- pars[grep('beta', pars)]
  beta <- beta[-grep('s', beta)]
  sigma <- pars[grep('sigma', pars)]
  posterior <- as.matrix(fit)
  r <- mcmc_areas(posterior,  pars = c(beta, sigma, 'lp__'), prob = 0.8)
  c <- mcmc_trace(posterior,  pars = c(beta, sigma, 'lp__'), n_warmup = 2000,
                   facet_args = list(nrow = 2, labeller = label_parsed))
  p <- ggpairs(data.frame(posterior[,beta]))
  return(list(r = r, c = c, p = p))
}
datasub <- function(data, dataset, wsg = TRUE){
  LMAmax <- max(data$LMA)
  Nmax <- max(data$Nm)
  if(wsg)
    wsgmax <- max(data$wsg)
  if(dataset != 'full')
    data <- data[-which(data$Dataset %in% dataset),]
  mdata <- list(S = length(levels(droplevels(data$Dataset))), 
              N = length(as.numeric(droplevels(data$Dataset))),
              J = as.numeric(droplevels(data$Dataset)),
              LL = data$LL,
              LMA = (data$LMA/LMAmax),
              Nmass = (data$Nm/Nmax))
  if(wsg)
    mdata$wsg <- data$wsg/wsgmax
  return(mdata)
}
```

We tested different models and we kept the model with the best tradeoff between model complexity, convergence, likelihood, and prediction quality (see [leaf lifespan model](LL2.html) document). We found following model, named **M1**:
$$M1~~ {LL_s}_j \sim \mathcal{logN}(\beta_0*e^{{\beta_1}_s*{LMA_s}_j - {\beta_2}_s*{Nmass_s}_j},\,\sigma)\,$$
$$s=1,...,S_{=4}~, ~~j=1,...,n_s$$
$${\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I, ~~(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}$$

Leaf lifespan ($LL$) is following a lognormal law with location ${\mu_s}_j$ and scale $\sigma$. $s$ represents the dataset used to do the fit between 1 and $S=4$, it encompass environmental and protocol variations. $j$ represents the observation for a given site $s$. Location ${\mu_s}_j$ is the exponential from the difference between LMA and Nmass with parameters ${\beta_0}_s$, ${\beta_1}_s$, and ${\beta_2}_s$. Each ${\beta_i}_s$ is following a normal law located on $\beta_i$ with scale $\sigma_i$. Finally all $beta_i$, $\sigma$, and $\sigma_i$ are assumed without presemption following gamma laws of parameters $10^-2$, $10^-2$.

We decided to add to model **M1**, wood specific gravity as a new predictor through a parameter $\beta_3$ in a new model named **M10**. We removed dataset random effect on $\beta_3$ due to the fact that all wood specific gravities were provided by DRYAD database. Due to a correlation between $\beta_0$ and $\beta_3$ in **M10** (see figure 5), we tried also a model **M11** without intercept $\beta_0$. We then sampled all models with this composite dataset, and obtained following results :

## Results

```{r OOB, echo=FALSE}
knitr::kable(randomForest::randomForest(LL ~ LMA + Nm + LMA + wsg,
                   data = data, na.action = na.omit,
                   ntree = 1000, replace = F , importance = T,
                   do.trace = F, keep.forest = T, keep.inbag = T)$importance,
             caption = 'Variable importance calculated with out-of the bag method applied on a random forest. Leaf lifespan (LL) is taken in GLOPNET database from @wright_worldwide_2004. Leaf mass per area (LMA), and leaf nitrogen content (Nmass) are taken both in TRY (https://www.try-db.org) and GLOPNET databases. Wood specific gravity (wsg) is taken both in TRY and DRYAD databases.', format = 'pandoc')
```

Table 3: Model fit results for composite dataset. *Data are taken from GLOPNET, TRY, and DRYAD datasets (see previous description). ML stand for maximum likelihood and RMSEP for root mean square error of prediction*

```{r model, echo=FALSE}
# mdata <- sapply(c('full', levels(droplevels(data$Dataset))),
#                 function(x) datasub(data, x), simplify = F)
# fits <- lapply(as.list(paste0('M', c(1, 10, 11))), function(model){
#   fit <- lapply(mdata, function(x) stan(file = file.path(mpath, paste0(model, '.stan')), data = x))
#   eval <- NA
#   fit <- list(full = fit$full, eval = eval)
#   path <- file.path(mpath, paste0(model, '.LL.Rdata'))
#   save(fit, file = path)
# })
# fits <- list()
# i <- 1
# models <- list.files(path = mpath, pattern = '.LL.Rdata')
# for(file in models){
#   load(file = file.path(mpath, file))
#   fits[[i]] <- fit ; rm(fit) ; i <- i + 1
# }
# names(fits) <- paste0('fit', c(10, 11, 1))
# g10 <- fitgraph(fits$fit10$full)
# g11 <- fitgraph(fits$fit11$full) ; rm(file, i, models, mpath)
# tab <- data.frame(
#   M = c('M1', 'M10', 'M11'),
#   Model = c('${LL_s}_j \\sim \\mathcal{logN}(\\beta_0 + {\\beta_1}_s*{LMA_s}_j - {\\beta_2}_s*{Nmass_s}_j,\\,\\sigma)\\,$',
#             '${LL_s}_j \\sim \\mathcal{logN}(\\beta_0 + {\\beta_1}_s*{LMA_s}_j - {\\beta_2}_s*{Nmass_s}_j + {\\beta_3}*{wsg}_j,\\,\\sigma)\\,$',
#             '${LL_s}_j \\sim \\mathcal{logN}({\\beta_1}_s*{LMA_s}_j - {\\beta_2}_s*{Nmass_s}_j + {\\beta_3}*{wsg}_j,\\,\\sigma)\\,$'),
#   ML = c( max(as.matrix(fits$fit1$full)[,'lp__']),
#           max(as.matrix(fits$fit10$full)[,'lp__']),
#           max(as.matrix(fits$fit11$full)[,'lp__']))
# )
# kable(tab)
```

All model have correctly converged.

```{r A2r, echo=FALSE, fig.cap='Parameters posterior ditribution. Light blue area represents 80% confidence interval, and vertical blue line the mean.'}
# g11$r
```

```{r A2c, echo=FALSE, fig.cap='Parameters markov chains. Light grey area represents warmup iterations.'}
# g11$c
```

Model **M10** have shown a correlation between parameters $\beta_0$ and $\beta_3$, responsible for the creation of model **M11** (see previous paragraphs). Model **M11** only show a slight link between $\beta_1$ and $\beta_3$ which can be considered as non significant.

```{r A2p11, echo=FALSE, fig.cap='Parameters pairs plot for model M11. Parameters density distribution, pairs plot and Pearson s coefficient of correlation.'}
# g11$p + ggtitle('M11') 
```

We can denote that retained vector of parameters with maximum likelihood for model **M11** is close to 95% confidence limit.

```{r A2graph11, echo=FALSE, fig.cap='Leaflifespan predictions for model M11 with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values. Leaf lifespan (LL) is predicted with model M10 fit. Leaf mass per area (LMA) and leaf nitrogen content (Nmass), and wood specific gravity (wsg) are taken in a composite dataset of GLOPNET, TRY and DRYAD  datasets.Warning LMA (resp. Nmass and wsg) is not constant and depend on the closest point value for right (resp. center and left) graph.'}
# beta <- as.matrix(fits$fit11$full)[which.max(as.matrix(fits$fit11$full)[,'lp__']),c('beta_1', 'beta_2', 'beta_3')]
# fit <- fits$fit10$full
# pred <- apply(as.matrix(fit), 1, function(x)
#   sapply(seq_len(dim(data)[1]), function(i)
#       rlnorm(1, x['beta_1']*(data$LMA[i]/LMAmax) - x['beta_2']*(data$Nm[i]/Nmax) + x['beta_3']*(data$wsg[i]/wsgmax), x['sigma'])
#     ))
# pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.15, 0.85)))))
# names(pred) <- c('5%', '95%')
# pred$mean <- exp((beta[1]*(data$LMA/LMAmax) - beta[2]*(data$Nm/Nmax) + beta[3]*(data$wsg/wsgmax)))
# pred$LMA <- data$LMA
# pred$Nm <- data$Nm
# pred$wsg <- data$wsg
# pred$obs <- data$LL
# lma <- ggplot(data, aes(x = LMA, y = LL)) +
#   geom_ribbon(aes(x = pred$LMA, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
#   geom_line(aes(x = pred$LMA, y = pred$`5%`, color = '5%'), linetype = 2) +
#   geom_line(aes(x = pred$LMA, y = pred$`95%`, color = '95%'), linetype = 2) +
#   geom_line(aes(x = pred$LMA, y = pred$mean, color = 'mean'), linetype = 2) +
#   geom_point() +
#   theme_bw() +
#   xlab('Leaf mass per area (LMA in g/m2)') +
#   ylab('Leaf lifespan (LL in months)') +
#   scale_color_manual(name = 'Model',
#                      values = c('5%' = 'black',
#                                 '95%' = 'black',
#                                 'mean' = 'red'))
# nm <- ggplot(data, aes(x = Nm, y = LL)) +
#   geom_ribbon(aes(x = pred$Nm, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
#   geom_line(aes(x = pred$Nm, y = pred$`5%`, color = '5%'), linetype = 2) +
#   geom_line(aes(x = pred$Nm, y = pred$`95%`, color = '95%'), linetype = 2) +
#   geom_line(aes(x = pred$Nm, y = pred$mean, color = 'mean'), linetype = 2) +
#   geom_point() +
#   theme_bw() +
#   xlab('Leaf nitrogen content (Nmass in %)') +
#   ylab('Leaf lifespan (LL in months)') +
#   scale_color_manual(name = 'Model',
#                      values = c('5%' = 'black',
#                                 '95%' = 'black',
#                                 'mean' = 'red'))
# wsg<- ggplot(data, aes(x = wsg, y = LL)) +
#   geom_ribbon(aes(x = pred$wsg, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
#   geom_line(aes(x = pred$wsg, y = pred$`5%`, color = '5%'), linetype = 2) +
#   geom_line(aes(x = pred$wsg, y = pred$`95%`, color = '95%'), linetype = 2) +
#   geom_line(aes(x = pred$wsg, y = pred$mean, color = 'mean'), linetype = 2) +
#   geom_point() +
#   theme_bw() +
#   xlab('Leaf nitrogen content (Nmass in %)') +
#   ylab('Leaf lifespan (LL in months)') +
#   scale_color_manual(name = 'Model',
#                      values = c('5%' = 'black',
#                                 '95%' = 'black',
#                                 'mean' = 'red'))
# predvsobs<- ggplot(pred, aes(x = obs, y = mean)) +
#   geom_ribbon(aes(ymin =`5%`, ymax = `95%`), color = 'grey', alpha = 0.2) +
#   geom_line(aes(y = `5%`, color = '5%'), linetype = 2) +
#   geom_line(aes(y = `95%`, color = '95%'), linetype = 2) +
#   geom_line(aes(y = mean, color = 'mean'), linetype = 2) +
#   geom_point() +
#   theme_bw() +
#   xlab('Observed leaf lifespan (LL in months)') + 
#   ylab('Predicted leaf lifespan (LL in months)') +
#   geom_abline(slope = 1) +
#   scale_color_manual(name = 'Model',
#                      values = c('5%' = 'black',
#                                 '95%' = 'black',
#                                 'mean' = 'red'))
# plot_grid(lma, nm, wsg, predvsobs, labels = c('LMA', 'Nmass', 'wsg', 'P v. O'))
```

We finally tested the new allometry obtained with the new model **M** on TROLL data (figure 7) and with TROLL simulations. Minimum leaf lifespan is now above 3 months for the new model **M** with TROLL species dataset. Moreover lowest leaf lifespan species are not the one with the smallest LMA with new model. Lastly, new model **M** variation of leaf lifespan is almost spanning all previous allometries prediciton on LMA axis.

```{r TROLL, echo=FALSE, fig.cap='Leaflifespan predictions for model different allometries with leaf mass per area (LMA), and leaf nitrogen content (Nmass), and wood specific gravity (wsg). Leaf lifespan (LL) is predicted with model M1, M10, and M11 fits or Reich s and Wright s allometries [@Reich1991a]. Leaf mass per area (LMA) leaf nitrogen content (Nmass) , and wood specific gravity (wsg) are taken in BRIDGE dataset used by TROLL [@Li].'}
# species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt', header=T, dec=".", sep="", row.names = 1)[c('LMA', 'Nmass', 'wsg')]
# species$`Reich 91` <- 1.5 + 10^(7.18+3.03*log10(species$LMA*0.0001))
# species$`Reich 97` <- 10^(2.040816*(2.579713-log10(10000.0/species$LMA)))
# species$`Wright 04` <- 0.5+10^(-2.509+1.71*log10(species$LMA))
# beta <- as.matrix(fits$fit1$full)[which.max(as.matrix(fits$fit1$full)[,'lp__']),c('beta_0', 'beta_1', 'beta_2')]
# species$Model1 <- with(species, exp((beta[1] + beta[2]*(LMA/LMAmax) - beta[3]*(Nmass*100/Nmax))))
# beta <- as.matrix(fits$fit10$full)[which.max(as.matrix(fits$fit10$full)[,'lp__']),c('beta_0', 'beta_1', 'beta_2', 'beta_3')]
# species$Model10 <- with(species, exp((beta[1] + beta[2]*(LMA/LMAmax) - beta[3]*(Nmass*100/Nmax) + beta[4]*(wsg/wsgmax))))
# beta <- as.matrix(fits$fit11$full)[which.max(as.matrix(fits$fit11$full)[,'lp__']),c('beta_1', 'beta_2', 'beta_3')]
# species$Model11 <- with(species, exp((beta[1]*(LMA/LMAmax) - beta[2]*(Nmass*100/Nmax) + beta[3]*(wsg/wsgmax))))
# data <- melt(species, id = c('LMA', 'Nmass', 'wsg'))
# names(data)[4] <- 'Allometry'
# g <- ggplot(data, aes(x = LMA, y = value, color = Allometry)) +
#   geom_point() + theme_bw() +   xlab('LMA in g/m2') +
#   ylab('Leaf lifespan (LL in months)')
# h <- ggplot(data, aes(x = Nmass, y = value, color = Allometry)) +
#   geom_point() + theme_bw() +   xlab('Nmass in %') +
#   ylab('Leaf lifespan (LL in months)')
# f <- ggplot(data, aes(x = wsg, y = value, color = Allometry)) +
#   geom_point() + theme_bw() +   xlab('wsg in mg/mm3') +
#   ylab('Leaf lifespan (LL in months)')
# plot_grid(g, h, f, labels = c('LMA', 'Nmass', 'wsg'))
```

Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early edath of individuals inside simulations. For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above X tonnes of C/ha and realistic diameter and age distribution inside the final population.

### 2017-06-08 docs/master-thesis.tex
Model and Appendix

@@
 upon request. All modules of TROLL models are further detailed in
-\protect\hyperlink{appendix-1}{Appendix 1}.
+\protect\hyperlink{appendix-1-troll-model}{Appendix 1: TROLL model}.
 
@@
 to infer leaf lifespan \(LL\) from species leaf mass per area \(LMA\)
-\citep[see \protect\hyperlink{appendix-1}{Appendix 1}]{Li}. But the use
-of the allometrie from \citet{Reich1991a} with current implementation of
-the TROLL model resulted in an underestimation of leaf lifespan for low
-LMA species. Consequently in the following paragraph we suggest a new
-allometry.
+\citep[see \protect\hyperlink{appendix-1-troll-model}{Appendix 1: TROLL
+model}]{Li}. But the use of the allometrie from \citet{Reich1991a} with
+current implementation of the TROLL model resulted in an underestimation
+of leaf lifespan for low LMA species. Consequently in the following
+paragraph we suggest a new allometry.
 
@@
 presemption following a gamma law. \(d\) represents the dataset random
-effects and encompass environmental and protocol variations. \textbf{Add
-Appendix ?}
+effects and encompass environmental and protocol variations (see
+\protect\hyperlink{appendix-2-leaf-lifespan-model}{Appendix 2: Leaf
+lifespan model} for more details).
 
@@
 
-\textbf{Add Appendix ?}
+(see \protect\hyperlink{appendix-3-rotten-tree-model}{Appendix 3: Rotten
+tree model} for more details)
 
@@
 from log as random like current natural treefall implementation inside
-TROLL (see code \protect\hyperlink{appendix-1}{Appendix 1}).
+TROLL (see code {[}Appendix 1{]}).
 
@@
 \(P(Death_i)\) is finally calculated by taking the inverse logit of the
-odd \(logit^{-1}[O(Death_i)]\). \textbf{Add Appendix ?}
+odd \(logit^{-1}[O(Death_i)]\) (see
+\protect\hyperlink{appendix-4-gap-damages-model}{Appendix 4: Gap damages
+model} for more details).
 
@@
 \eqref{eq:DeltaLA} and \(m\) see \eqref{eq:db}) which they assumed having a
-key role in model functioning.
-
-On the other hand, we decided to use TROLL to study resistance and
-resilience of ecosystem face to disturbance, highlighting the role of
-biodiversit. Consequently we particulrarly needed to assess the
-importance of functional traits to further better control and evaluate
-functional diversities. We also needed to assess the sensitivity of
-TROLL model to the seed rain constant (\(n_{ext}\), see \eqref{eq:next})
-because we assume it is one of the main factor of tree recruitments
-after disturbance in simulations.
+key role in model functioning. On the other hand, we decided to use
+TROLL to study resistance and resilience of ecosystem face to
+disturbance, highlighting the role of biodiversity. Consequently we
+particulrarly needed to assess the importance of functional traits to
+further better control and evaluate functional diversities. We also
+needed to assess the sensitivity of TROLL model to the seed rain
+constant (\(n_{ext}\), see \eqref{eq:next}) because we assumed it as one
+of the main factors of tree recruitments after disturbance within
+simulations.
 
@@
 values to their mean. Each trait was tested independently. We reduce to
-a common mean traits with a Perason's correlation value \(r \geq 0.8\)
-(\(h_{max}\) and \(a_h\) with a correlation of \(r=0.98\)).
-
-To assess the sensitivity of TROLL model to seed rain, we performed a
-sensitivity analysis by fixing simulations seed rain constant to 2, 20,
-200 and 2000 seeds per hectare (default value being 200 seeds per
-hectare).
+a common mean traits with a Pearson's correlation value \(r \geq 0.8\)
+(\(h_{max}\) and \(a_h\) with a correlation of \(r=0.98\)). To assess
+the sensitivity of TROLL model to seed rain, we performed a sensitivity
+analysis by fixing simulations seed rain constant to 2, 20, 200 and 2000
+seeds per hectare.
 
@@
 
-\subsection{Disturbance}\label{disturbance-1}
-
-\emph{Most tropical timber production originates from selective logging,
-the targeted harvesting of timber from species of interest.
-Consequently, tropical sylviculture can be assimilated to a disturbance.
-The main difference between a disturbance and selective logging is the
-targetting of both species and individuals of interest. So we decided to
-first asses unselective disturbance effect on tropical forest ecosystem
-to subsequently better understand selective logging effect. We first
-implemented a simple disturbance module inside TROLL model to simulate
-unselective disturbance.} \textbf{Already inside model descripiton !}
-
-We first implemented a disturbance module inside TROLL model to simulate
-unselective disturbance. 60 tree communities were then defined with
-different levels of both specific and functional richness to explore the
-role of biodiversity on tropical forest ecosystem answer to disturbance.
-Next, different levels of disturbance were simulated on previous mature
-communities. Finally, resilience of ecosystem major global variables
-(related to carbon stock, forest dynamic and florisitic structure) from
-simulations were partitioned between selection effects and niche
-complementarity \citep{Loreau2001}.
-
-\subsection{Selective logging}\label{selective-logging}
-
-Once the unselective effect of disturbance on forest ecosystem studied,
-we can focus on selective logging to highlight the effect of species and
-individuals targetting on the ecosystem answer. We first implemented a
-sylviculture module inside TROLL model to simulate selective logging in
-regards to french Guiana practices. We used the 60 communities from
-disturbance simulations to study the role of biodiversity in both forest
-ecosystem and commercial species answer to selective logging. We then
-
 \subsection{Design of experiment}\label{design-of-experiment}
 
-In order to assess the role of biodiversity in ecosystem answer to
-disturbance, we needed to create a space of experiments encompassing
-both variation of disturbance, biodiversity and time. Disturbance was
-represented by percentage of basal area loss (0\%, 25\%, 50\% and 75\%).
-Biodiversity was integrated with two components specific and functional
-diversities. We used species richness \(SR\) to represents species
-diversity (5, 25, and 125). Functional diversity can be related to
-numerous components, and \citet{Borgy2017} argued for 5: richness,
-divergence, regularity, overlap and mean. Because mature forest were
-created from a bare soil with TROLL simulations, we could not control a
-priori divergence, regularity and overlap but only assess theim before
-diturbance. Consequently, we focused on functional richness with convex
-hull volume \(CHV\) and functional mean with community weighted mean
-\(CWM\). For each level of species richness \(SR\), we selected 20
-communities with growing convex hull volume \(CHV\) but with a community
-weighted means close to the regional species pool community weighted
-means. Effectivelly, we did not wanted drastic change in community means
-that could have more effect than functional richness itself. This design
-of experiments resulted in 60 communities (\(5~SR*20~CHV\)) and 240
-simulations (\(60 ~communities*4~levels~of~disturbance\)) over 600 years
-(maturity being assumed after 500 years of regeneration \citep{Li}).
-Figure \ref{fig:DOE} presents the design of experiment for communities
-biodiversity after the forest mature were simulated, and thus before
+In order to assess the role of biodiversity in ecosystem answer to both
+disturbance and sylviculture, we needed to create a space of experiments
+encompassing both variation of disturbance, biodiversity and time.
+Disturbance was represented by percentage of basal area loss (0\%, 25\%,
+50\% and 75\%), or as a selective logging simulation. Biodiversity was
+integrated with two components of its components: taxonomic and
+functional diversities. We used species richness \(SR\) to represents
+taxonomic diversity (5, 25, and 125 species). Functional diversity can
+be related to numerous components, and \citet{Borgy2017} argued for 5:
+richness, divergence, regularity, overlap and mean. Because mature
+forest were created from a bare soil with TROLL simulations, we could
+not control a priori divergence, regularity and overlap but only assess
+theim before diturbance. Consequently, we focused on functional richness
+with convex hull volume \(CHV\) and functional mean with community
+weighted mean \(CWM\). For each level of species richness \(SR\), we
+selected 20 communities with growing convex hull volume \(CHV\) but with
+a community weighted means close to the regional species pool community
+weighted means. Effectivelly, we did not wanted drastic change in
+community means that could have more effect than functional richness
+itself. This design of experiments resulted in 60 communities
+(\(5~SR*20~CHV\)) and 240 simulations
+(\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
+being assumed after 500 years of regeneration \citep{Li}). Figure
+\ref{fig:DOE} presents the design of experiment for communities
+biodiversity after the mature forest were simulated, and thus before
 disturbance. We obtained a broad range of both functionl dispersion
@@
 (AGB). FDis was caluclated based on 4 functional traits (leaf mass per
-area, wood specific gravity, maximum diameter, maximum height).}
+area, wood specific gravity, maximum diameter, and maximum height).}
 \end{figure}
 
-\subsection{Ecosystem answer anlaysis}\label{ecosystem-answer-anlaysis}
+\subsection{Ecosystem answer analysis}\label{ecosystem-answer-analysis}
+
+\subsubsection{Resilience}\label{resilience}
 
 Tropical forest ecosystems provides numerous ecosystem services linked
-to several ecosystem functions. We focused on few functions and related
-metrics to analyse ecosystem answer to disturbance: carbon stock with
-aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), forest dynamic with
-number of stem above \(10~cm\) diameter at breast height (\(N10\)), and
-floristic composition with \textbf{?} (\textbf{?} in \textbf{?}).
+to several ecosystem functions. We decided to describe simiulated
+tropical forests in three major functions: forest dynamic, forest
+production, and forest diversity. Forest dynamic was represented by
+aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\)
+in \(m^2.ha^{-1}\)), total number of stem (\(N\)), number of stem above
+\(10~cm\) diameter (\(N10\)), and number of stem above \(30~cm\)
+diameter (\(N30\)). Forest production was represented by growth pimary
+productivity (\(GPP\) in \(MgC.ha^{-1}\)), net pimary productivity
+(\(NPP\) in \(MgC.ha^{-1}\)), tree autotrophic respiration in day
+(\(Rday\) in \(MgC.ha^{-1}\)) and tree autotrophic respiration in night
+(\(Rnight\) in \(MgC.ha^{-1}\)). Finally, forest diversity was
+represented by \textbf{??}.
 
@@
 \begin{equation}
-  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X(t)}{X(t_d-1)}
+  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X_T(t)}{X_C(t)}
   \label{eq:Resilience}
@@
 the ratio of recovery \(Recovery(t)\) at time \(t\) to loss suffered
-\(Loss(t_d)\) at disturbance time \(t_d\). We transformed the equation
-in the resilience of the system \(R(t)\) at the time \(t\) being the
-ratio of the ecosystem metric \(X(t)\) at time \(t\) to the ecosystem
-metric before the disturbance happened \(X(t_d-1)\) at time \(t_d-1\).
-\(X(t_d-1)\) was calculated as the mean of the ecosystem output for 50
-last years of the simulation of the mature forest (over 600 years).
-Then, we calculated for each simulation the recovery time
-\(t_{recovery}\) were the ecosystem reached back its stable state
-defined as its state before disturbance. But some ecosystem metrics can
-reach pre disturbance value wihtout revealing ecosystem recovery. For
-instance, number of stem above \(10~cm\) diameter at breast height
-\(N10\) will first decrease due to disturbance. Then \(N10\) will exceed
-its initial value due to new seedlings recruitment before decreasing
-towards its pre disturbanec value. Consequently, the first time \(N10\)
-reaches back its pre disturbance value can not be considered as the
-recovery time \(t_{recovery}\). In order to solve this issue, we
-considered the ecosystem as stable again when its whole set of observed
-metrics reach their pre disturbance values.
+\(Loss(t_d)\) at disturbance time \(t_d\). But in our peculiar case of
+tropical forest ecosystems, the equilibrium used to calculate
+\(Loss(t_d)\) can not be reduced to a specific time if the equilibrium
+is dynamic. Consequently, to encompass undisturbed ecosystem variations
+throught time, we simulated an undisturbed control ecosystem \(C\). And
+the resilience of the system \(R(t)\) at the time \(t\) was defined as
+the ratio of the ecosystem metric values in the disturbed simulation
+\(X_T(t)\) over the the ecosystem metric values from the control
+\(X_C(t)\) . Thus, the value of resilience \(R(t)\) is normalized for
+all simulations and metrics. \(R(t)\) will be equal to \(R_{eq} = 1\)
+when reaching the equilibrium value. Consequently we can calculate an
+euclidean distance to equilibrium \(d_{eq}(t)\) as
+\(d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}\).
+
+\subsubsection{Biodiversity effect}\label{biodiversity-effect}
 
 Biodiversity is not only a facet of the experimental design and an
-ecosystem output through florisitic composition, but also interact on
+ecosystem output through forest diversity, but also interact on
 ecosystem functioning and consequently on its answer to disturbance.
@@
 
-\subsection{Disturbance}\label{disturbance-2}
+\subsection{Disturbance}\label{disturbance-1}
 
@@
 
-\hypertarget{appendix-1}{\section{Appendix 1}\label{appendix-1}}
+\hypertarget{appendix-1-troll-model}{\section{Appendix 1: TROLL
+model}\label{appendix-1-troll-model}}
 
@@
 simulated inside TROLL.
-
-\subsection{Leaf lifespan}\label{leaf-lifespan-1}
-
-The underestimation of leaf lifespan for low LMA species with the
-allometry from \citet{Reich1991a} resulted in indivduals unealistic
-early death from carbon starvation. We gathered data from
-TRY\citep{Kattge2011}, DRYAD \citep{chave_towards_2009} and GLOPNET
-\citep{wright_worldwide_2004} datasets. We used an out of the bag method
-applied on a random forest to select variables with highest importance
-to explain leaf lifespan. We thus selected leaf mass per are \(LMA\),
-leaf nitrogen content \(N\) and wood specific gravity \(wsg\). We then
-used a bayesian approach to test different models with growing level of
-complexity. The model with the best tradeoff between complexity (number
-of parameters), convergence, likelihood, and prediction quality (root
-mean square error of prediction RMSEP) was kept. We selected following
-model with a maximum likelihood of 13.6 and a RMSEP of 12 months:
-
-\begin{equation}
-  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
-  \label{eq:LLb}
-\end{equation}
-
-Leaf lifespan \(LL\) follows a lognormal law with location infered from
-leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
-\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
-normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
-\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
-presemption following a gamma law. \(d\) represents the dataset random
-effects and encompass environmental and protocol variations. The
-sampling of model \eqref{eq:LLb} resulted in the following allometry:
-
-\begin{equation}
-  LL = e^{0.017*LMA - 0.103*Nmass + 1.94*wsg}
-  \label{eq:LL}
-\end{equation}
 
@@
 
+\hypertarget{appendix-2-leaf-lifespan-model}{\section{Appendix 2: Leaf
+lifespan model}\label{appendix-2-leaf-lifespan-model}}
+
+TROLL model previous implementation encompass Reich's 1991 and 1997 and
+Wright's 2004 allometries to estimate leaf lifespan with
+\citep{Reich1991a, Reich1997, wright_worldwide_2004}. But we have shown
+that Reich's allometries are underestimating leaf lifespan for low LMA
+species. Moreover simulations estimated unrealistically low aboveground
+biomass for low LMA species. We assumed Reich's allometries
+underestimation of leaf lifespan for low LMA species being the source of
+unrealistically low aboveground biomass inside TROLL simulations. We
+decided to find a better allometry with \citet{wright_worldwide_2004}
+GLOPNET dataset.
+
+\subsection{Material and methods}\label{material-and-methods-1}
+
+We gathered data from TRY database to evaluate and improve previous
+models. We requested functional traits available in Bridge dataset and
+leaf lifespan for tropical species present in GLOPNET dataset and used
+in TROLL (see \ref{tab:A2traits}). We compiled GLOPNET, TRY, and DRYAD
+databases. We kept dataset given by GLOPNET as origin dataset for
+observations. It correspond to leaf lifespan and also to LMA and Nmass
+in most of cases. We measured variable importance in functionnal traits
+to explain leaf lifespan with out-of bag method applied on a random
+forest.
+
+\begin{table}
+
+\caption{\label{tab:A2traits}Functional traits gathered with TRY.}
+\centering
+\begin{tabular}[t]{l|l|l|r}
+\hline
+Name & Trait & Unit & TRYcode\\
+\hline
+LL & Leaf lifespan (longevity) & month & 12\\
+\hline
+SLA & Leaf area per leaf dry mass (specific leaf area, SLA) & \$m\textasciicircum{}2.kg\textasciicircum{}\{-1\}\$ & 11\\
+\hline
+N & Leaf nitrogen (N) content per leaf dry mass & \$mg.g\textasciicircum{}\{-1\}\$ & 15\\
+\hline
+P & Leaf phosphorus (P) content per leaf dry mass & \$mg.g\textasciicircum{}\{-1\}\$ & 14\\
+\hline
+wsg & Stem dry mass per stem fresh volume (stem specific density) & \$mg.mm\textasciicircum{}\{-3\}\$ & 4\\
+\hline
+\end{tabular}
+\end{table}
+
+We tested different models and we kept the model with the best tradeoff
+between model complexity, convergence, likelihood, and prediction
+quality (see \href{LL2.html}{leaf lifespan model} document). We found
+following model, named \textbf{M1}:
+\[M1~~ {LL_s}_j \sim \mathcal{logN}(\beta_0*e^{{\beta_1}_s*{LMA_s}_j - {\beta_2}_s*{Nmass_s}_j},\,\sigma)\,\]
+\[s=1,...,S_{=4}~, ~~j=1,...,n_s\]
+\[{\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I, ~~(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}\]
+
+Leaf lifespan (\(LL\)) is following a lognormal law with location
+\({\mu_s}_j\) and scale \(\sigma\). \(s\) represents the dataset used to
+do the fit between 1 and \(S=4\), it encompass environmental and
+protocol variations. \(j\) represents the observation for a given site
+\(s\). Location \({\mu_s}_j\) is the exponential from the difference
+between LMA and Nmass with parameters \({\beta_0}_s\), \({\beta_1}_s\),
+and \({\beta_2}_s\). Each \({\beta_i}_s\) is following a normal law
+located on \(\beta_i\) with scale \(\sigma_i\). Finally all \(beta_i\),
+\(\sigma\), and \(\sigma_i\) are assumed without presemption following
+gamma laws of parameters \(10^-2\), \(10^-2\).
+
+We decided to add to model \textbf{M1}, wood specific gravity as a new
+predictor through a parameter \(\beta_3\) in a new model named
+\textbf{M10}. We removed dataset random effect on \(\beta_3\) due to the
+fact that all wood specific gravities were provided by DRYAD database.
+Due to a correlation between \(\beta_0\) and \(\beta_3\) in \textbf{M10}
+(see figure 5), we tried also a model \textbf{M11} without intercept
+\(\beta_0\). We then sampled all models with this composite dataset, and
+obtained following results :
+
+\subsection{Results}\label{results-1}
+
+\begin{longtable}[]{@{}lrr@{}}
+\caption{\label{tab:OOB}Variable importance calculated with out-of the bag
+method applied on a random forest. Leaf lifespan (LL) is taken in
+GLOPNET database from \citet{wright_worldwide_2004}. Leaf mass per area
+(LMA), and leaf nitrogen content (Nmass) are taken both in TRY
+(\url{https://www.try-db.org}) and GLOPNET databases. Wood specific
+gravity (wsg) is taken both in TRY and DRYAD databases.}\tabularnewline
+\toprule
+& \%IncMSE & IncNodePurity\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+& \%IncMSE & IncNodePurity\tabularnewline
+\midrule
+\endhead
+LMA & 99.69390 & 2028.079\tabularnewline
+Nm & 159.03360 & 2666.670\tabularnewline
+wsg & 50.97284 & 1475.023\tabularnewline
+\bottomrule
+\end{longtable}
+
+Table 3: Model fit results for composite dataset. \emph{Data are taken
+from GLOPNET, TRY, and DRYAD datasets (see previous description). ML
+stand for maximum likelihood and RMSEP for root mean square error of
+prediction}
+
+All model have correctly converged.
+
+Model \textbf{M10} have shown a correlation between parameters
+\(\beta_0\) and \(\beta_3\), responsible for the creation of model
+\textbf{M11} (see previous paragraphs). Model \textbf{M11} only show a
+slight link between \(\beta_1\) and \(\beta_3\) which can be considered
+as non significant.
+
+We can denote that retained vector of parameters with maximum likelihood
+for model \textbf{M11} is close to 95\% confidence limit.
+
+We finally tested the new allometry obtained with the new model
+\textbf{M} on TROLL data (figure 7) and with TROLL simulations. Minimum
+leaf lifespan is now above 3 months for the new model \textbf{M} with
+TROLL species dataset. Moreover lowest leaf lifespan species are not the
+one with the smallest LMA with new model. Lastly, new model \textbf{M}
+variation of leaf lifespan is almost spanning all previous allometries
+prediciton on LMA axis.
+
+Simulations are validating that this new allometry resolve the issue of
+unrealistically low aboveground biomass for low LMA species due to an
+early edath of individuals inside simulations. For instance with this
+allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic
+aboveground biomass above X tonnes of C/ha and realistic diameter and
+age distribution inside the final population.
+
+\hypertarget{appendix-3-rotten-tree-model}{\section{Appendix 3: Rotten
+tree model}\label{appendix-3-rotten-tree-model}}
+
+\hypertarget{appendix-4-gap-damages-model}{\section{Appendix 4: Gap
+damages model}\label{appendix-4-gap-damages-model}}
+
 \addcontentsline{toc}{section}{References}

### 2017-06-09 08-Appendix2.Rmd
Appendix2 LL

@@
 
-We gathered data from TRY database to evaluate and improve previous models.
-We requested functional traits available in Bridge dataset and leaf lifespan for tropical species present in GLOPNET dataset and used in TROLL (see \@ref(tab:A2traits)).
-We compiled GLOPNET, TRY, and DRYAD databases.
+We compiled functional traits from GLOPNET [@wright_worldwide_2004], TRY [@Kattge2011], and DRYAD [@chave_towards_2009] databases (see \@ref(tab:A2traits)).
 We kept dataset given by GLOPNET as origin dataset for observations.
-It correspond to leaf lifespan and also to LMA and Nmass in most of cases.
+Dataset defined as origin corresponded to leaf lifespan ($LL$) and most of the time to leaf mass per area ($LMA$) and leaf nitrogen content per leaf dry mass ($Nmass$).
 We measured variable importance in functionnal traits to explain leaf lifespan with out-of bag method applied on a random forest.
+Then, we used a béayesian apporach to test different models with growing level of complexity.
+We retained the model with the best tradeoff between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean sqaure error of prediciton $RMSEP$).
+We finally tested the new allometry obtained with the selected model with TROLL simulations.
 
@@
 
-```{r functions, echo=FALSE}
-fitgraph <- function(fit){
-  pars <- fit@model_pars
-  beta <- pars[grep('beta', pars)]
-  beta <- beta[-grep('s', beta)]
-  sigma <- pars[grep('sigma', pars)]
-  posterior <- as.matrix(fit)
-  r <- mcmc_areas(posterior,  pars = c(beta, sigma, 'lp__'), prob = 0.8)
-  c <- mcmc_trace(posterior,  pars = c(beta, sigma, 'lp__'), n_warmup = 2000,
-                   facet_args = list(nrow = 2, labeller = label_parsed))
-  p <- ggpairs(data.frame(posterior[,beta]))
-  return(list(r = r, c = c, p = p))
-}
-datasub <- function(data, dataset, wsg = TRUE){
-  LMAmax <- max(data$LMA)
-  Nmax <- max(data$Nm)
-  if(wsg)
-    wsgmax <- max(data$wsg)
-  if(dataset != 'full')
-    data <- data[-which(data$Dataset %in% dataset),]
-  mdata <- list(S = length(levels(droplevels(data$Dataset))), 
-              N = length(as.numeric(droplevels(data$Dataset))),
-              J = as.numeric(droplevels(data$Dataset)),
-              LL = data$LL,
-              LMA = (data$LMA/LMAmax),
-              Nmass = (data$Nm/Nmax))
-  if(wsg)
-    mdata$wsg <- data$wsg/wsgmax
-  return(mdata)
-}
-```
-
-We tested different models and we kept the model with the best tradeoff between model complexity, convergence, likelihood, and prediction quality (see [leaf lifespan model](LL2.html) document).
-We found following model, named **M1**:
-$$M1~~ {LL_s}_j \sim \mathcal{logN}(\beta_0*e^{{\beta_1}_s*{LMA_s}_j - {\beta_2}_s*{Nmass_s}_j},\,\sigma)\,$$
-$$s=1,...,S_{=4}~, ~~j=1,...,n_s$$
-$${\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I, ~~(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}$$
-
-Leaf lifespan ($LL$) is following a lognormal law with location ${\mu_s}_j$ and scale $\sigma$.
-$s$ represents the dataset used to do the fit between 1 and $S=4$, it encompass environmental and protocol variations.
-$j$ represents the observation for a given site $s$.
-Location ${\mu_s}_j$ is the exponential from the difference between LMA and Nmass with parameters ${\beta_0}_s$, ${\beta_1}_s$, and ${\beta_2}_s$.
-Each ${\beta_i}_s$ is following a normal law located on $\beta_i$ with scale $\sigma_i$.
-Finally all $beta_i$, $\sigma$, and $\sigma_i$ are assumed without presemption following gamma laws of parameters $10^-2$, $10^-2$.
-
-We decided to add to model **M1**, wood specific gravity as a new predictor through a parameter $\beta_3$ in a new model named **M10**.
-We removed dataset random effect on $\beta_3$ due to the fact that all wood specific gravities were provided by DRYAD database.
-Due to a correlation between $\beta_0$ and $\beta_3$ in **M10** (see figure 5), we tried also a model **M11** without intercept $\beta_0$.
-We then sampled all models with this composite dataset, and obtained following results :
-
 ## Results
 
-```{r OOB, echo=FALSE}
+Out of the bag method applied on a random forest highlighted the importance of leaf nitrogen content per leaf dry mass ($N_m$) to model leaf lifespan (see \@ref(tab:A2OOB)).
+$N_m$ importance was higher than leaf mass per area (158 against 96 percent of mean square error increase) which was used as a proxy for leaf lifespan in previous models.
+Finally, wood specific gravity ($wsg$) add also a significant importance in leaf lifespan estimation.
+
+```{r A2OOB, echo=FALSE}
 knitr::kable(randomForest::randomForest(LL ~ LMA + Nm + LMA + wsg,
@@
              caption = 'Variable importance calculated with out-of the bag method applied on a random forest.
+First column represents the mean decrease in mean square error (%IncMSE) whereas second column represents the total decrease in node impurities, measured by the Gini Index  (IncNodePurety).
 Leaf lifespan (LL) is taken in GLOPNET database from @wright_worldwide_2004.
@@
 
-Table 3: Model fit results for composite dataset.
-*Data are taken from GLOPNET, TRY, and DRYAD datasets (see previous description).
-ML stand for maximum likelihood and RMSEP for root mean square error of prediction*
+The selected model had a maximum likelihood of 13.6 and a RMSEP of 12 months:
 
-```{r model, echo=FALSE}
-# mdata <- sapply(c('full', levels(droplevels(data$Dataset))),
-#                 function(x) datasub(data, x), simplify = F)
-# fits <- lapply(as.list(paste0('M', c(1, 10, 11))), function(model){
-#   fit <- lapply(mdata, function(x) stan(file = file.path(mpath, paste0(model, '.stan')), data = x))
-#   eval <- NA
-#   fit <- list(full = fit$full, eval = eval)
-#   path <- file.path(mpath, paste0(model, '.LL.Rdata'))
-#   save(fit, file = path)
-# })
-# fits <- list()
-# i <- 1
-# models <- list.files(path = mpath, pattern = '.LL.Rdata')
-# for(file in models){
-#   load(file = file.path(mpath, file))
-#   fits[[i]] <- fit ; rm(fit) ; i <- i + 1
-# }
-# names(fits) <- paste0('fit', c(10, 11, 1))
-# g10 <- fitgraph(fits$fit10$full)
-# g11 <- fitgraph(fits$fit11$full) ; rm(file, i, models, mpath)
-# tab <- data.frame(
-#   M = c('M1', 'M10', 'M11'),
-#   Model = c('${LL_s}_j \\sim \\mathcal{logN}(\\beta_0 + {\\beta_1}_s*{LMA_s}_j - {\\beta_2}_s*{Nmass_s}_j,\\,\\sigma)\\,$',
-#             '${LL_s}_j \\sim \\mathcal{logN}(\\beta_0 + {\\beta_1}_s*{LMA_s}_j - {\\beta_2}_s*{Nmass_s}_j + {\\beta_3}*{wsg}_j,\\,\\sigma)\\,$',
-#             '${LL_s}_j \\sim \\mathcal{logN}({\\beta_1}_s*{LMA_s}_j - {\\beta_2}_s*{Nmass_s}_j + {\\beta_3}*{wsg}_j,\\,\\sigma)\\,$'),
-#   ML = c( max(as.matrix(fits$fit1$full)[,'lp__']),
-#           max(as.matrix(fits$fit10$full)[,'lp__']),
-#           max(as.matrix(fits$fit11$full)[,'lp__']))
-# )
-# kable(tab)
-```
+\begin{equation}
+  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
+  (\#eq:LL)
+\end{equation}
 
-All model have correctly converged.
+Leaf lifespan $LL$ follows a lognormal law with location infered from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
+Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$.
+All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without presemption following a gamma law.
+$d$ represents the dataset random effects and encompass environmental and protocol variations.
 
-```{r A2r, echo=FALSE, fig.cap='Parameters posterior ditribution. Light blue area represents 80% confidence interval, and vertical blue line the mean.'}
-# g11$r
-```
-
-```{r A2c, echo=FALSE, fig.cap='Parameters markov chains. Light grey area represents warmup iterations.'}
-# g11$c
-```
-
-Model **M10** have shown a correlation between parameters $\beta_0$ and $\beta_3$, responsible for the creation of model **M11** (see previous paragraphs).
-Model **M11** only show a slight link between $\beta_1$ and $\beta_3$ which can be considered as non significant.
-
-```{r A2p11, echo=FALSE, fig.cap='Parameters pairs plot for model M11. Parameters density distribution, pairs plot and Pearson s coefficient of correlation.'}
-# g11$p + ggtitle('M11') 
-```
-
-We can denote that retained vector of parameters with maximum likelihood for model **M11** is close to 95% confidence limit.
-
-```{r A2graph11, echo=FALSE, fig.cap='Leaflifespan predictions for model M11 with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values.
+```{r A2LLpred, echo=FALSE, fig.cap='Leaflifespan predictions for the selected model with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values.
 Leaf lifespan (LL) is predicted with model M10 fit.
@@
 Nmass and wsg) is not constant and depend on the closest point value for right (resp.
-center and left) graph.'}
-# beta <- as.matrix(fits$fit11$full)[which.max(as.matrix(fits$fit11$full)[,'lp__']),c('beta_1', 'beta_2', 'beta_3')]
-# fit <- fits$fit10$full
-# pred <- apply(as.matrix(fit), 1, function(x)
-#   sapply(seq_len(dim(data)[1]), function(i)
-#       rlnorm(1, x['beta_1']*(data$LMA[i]/LMAmax) - x['beta_2']*(data$Nm[i]/Nmax) + x['beta_3']*(data$wsg[i]/wsgmax), x['sigma'])
-#     ))
-# pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.15, 0.85)))))
-# names(pred) <- c('5%', '95%')
-# pred$mean <- exp((beta[1]*(data$LMA/LMAmax) - beta[2]*(data$Nm/Nmax) + beta[3]*(data$wsg/wsgmax)))
-# pred$LMA <- data$LMA
-# pred$Nm <- data$Nm
-# pred$wsg <- data$wsg
-# pred$obs <- data$LL
-# lma <- ggplot(data, aes(x = LMA, y = LL)) +
-#   geom_ribbon(aes(x = pred$LMA, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
-#   geom_line(aes(x = pred$LMA, y = pred$`5%`, color = '5%'), linetype = 2) +
-#   geom_line(aes(x = pred$LMA, y = pred$`95%`, color = '95%'), linetype = 2) +
-#   geom_line(aes(x = pred$LMA, y = pred$mean, color = 'mean'), linetype = 2) +
-#   geom_point() +
-#   theme_bw() +
-#   xlab('Leaf mass per area (LMA in g/m2)') +
-#   ylab('Leaf lifespan (LL in months)') +
-#   scale_color_manual(name = 'Model',
-#                      values = c('5%' = 'black',
-#                                 '95%' = 'black',
-#                                 'mean' = 'red'))
-# nm <- ggplot(data, aes(x = Nm, y = LL)) +
-#   geom_ribbon(aes(x = pred$Nm, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
-#   geom_line(aes(x = pred$Nm, y = pred$`5%`, color = '5%'), linetype = 2) +
-#   geom_line(aes(x = pred$Nm, y = pred$`95%`, color = '95%'), linetype = 2) +
-#   geom_line(aes(x = pred$Nm, y = pred$mean, color = 'mean'), linetype = 2) +
-#   geom_point() +
-#   theme_bw() +
-#   xlab('Leaf nitrogen content (Nmass in %)') +
-#   ylab('Leaf lifespan (LL in months)') +
-#   scale_color_manual(name = 'Model',
-#                      values = c('5%' = 'black',
-#                                 '95%' = 'black',
-#                                 'mean' = 'red'))
-# wsg<- ggplot(data, aes(x = wsg, y = LL)) +
-#   geom_ribbon(aes(x = pred$wsg, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
-#   geom_line(aes(x = pred$wsg, y = pred$`5%`, color = '5%'), linetype = 2) +
-#   geom_line(aes(x = pred$wsg, y = pred$`95%`, color = '95%'), linetype = 2) +
-#   geom_line(aes(x = pred$wsg, y = pred$mean, color = 'mean'), linetype = 2) +
-#   geom_point() +
-#   theme_bw() +
-#   xlab('Leaf nitrogen content (Nmass in %)') +
-#   ylab('Leaf lifespan (LL in months)') +
-#   scale_color_manual(name = 'Model',
-#                      values = c('5%' = 'black',
-#                                 '95%' = 'black',
-#                                 'mean' = 'red'))
-# predvsobs<- ggplot(pred, aes(x = obs, y = mean)) +
-#   geom_ribbon(aes(ymin =`5%`, ymax = `95%`), color = 'grey', alpha = 0.2) +
-#   geom_line(aes(y = `5%`, color = '5%'), linetype = 2) +
-#   geom_line(aes(y = `95%`, color = '95%'), linetype = 2) +
-#   geom_line(aes(y = mean, color = 'mean'), linetype = 2) +
-#   geom_point() +
-#   theme_bw() +
-#   xlab('Observed leaf lifespan (LL in months)') + 
-#   ylab('Predicted leaf lifespan (LL in months)') +
-#   geom_abline(slope = 1) +
-#   scale_color_manual(name = 'Model',
-#                      values = c('5%' = 'black',
-#                                 '95%' = 'black',
-#                                 'mean' = 'red'))
-# plot_grid(lma, nm, wsg, predvsobs, labels = c('LMA', 'Nmass', 'wsg', 'P v. O'))
-```
-
-We finally tested the new allometry obtained with the new model **M** on TROLL data (figure 7) and with TROLL simulations.
-Minimum leaf lifespan is now above 3 months for the new model **M** with TROLL species dataset.
-Moreover lowest leaf lifespan species are not the one with the smallest LMA with new model.
-Lastly, new model **M** variation of leaf lifespan is almost spanning all previous allometries prediciton on LMA axis.
-
-```{r TROLL, echo=FALSE, fig.cap='Leaflifespan predictions for model different allometries with leaf mass per area (LMA), and leaf nitrogen content (Nmass), and wood specific gravity (wsg).
-Leaf lifespan (LL) is predicted with model M1, M10, and M11 fits or Reich s and Wright s allometries [@Reich1991a].
-Leaf mass per area (LMA) leaf nitrogen content (Nmass) , and wood specific gravity (wsg) are taken in BRIDGE dataset used by TROLL [@Li].'}
-# species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt', header=T, dec=".", sep="", row.names = 1)[c('LMA', 'Nmass', 'wsg')]
-# species$`Reich 91` <- 1.5 + 10^(7.18+3.03*log10(species$LMA*0.0001))
-# species$`Reich 97` <- 10^(2.040816*(2.579713-log10(10000.0/species$LMA)))
-# species$`Wright 04` <- 0.5+10^(-2.509+1.71*log10(species$LMA))
-# beta <- as.matrix(fits$fit1$full)[which.max(as.matrix(fits$fit1$full)[,'lp__']),c('beta_0', 'beta_1', 'beta_2')]
-# species$Model1 <- with(species, exp((beta[1] + beta[2]*(LMA/LMAmax) - beta[3]*(Nmass*100/Nmax))))
-# beta <- as.matrix(fits$fit10$full)[which.max(as.matrix(fits$fit10$full)[,'lp__']),c('beta_0', 'beta_1', 'beta_2', 'beta_3')]
-# species$Model10 <- with(species, exp((beta[1] + beta[2]*(LMA/LMAmax) - beta[3]*(Nmass*100/Nmax) + beta[4]*(wsg/wsgmax))))
-# beta <- as.matrix(fits$fit11$full)[which.max(as.matrix(fits$fit11$full)[,'lp__']),c('beta_1', 'beta_2', 'beta_3')]
-# species$Model11 <- with(species, exp((beta[1]*(LMA/LMAmax) - beta[2]*(Nmass*100/Nmax) + beta[3]*(wsg/wsgmax))))
-# data <- melt(species, id = c('LMA', 'Nmass', 'wsg'))
-# names(data)[4] <- 'Allometry'
-# g <- ggplot(data, aes(x = LMA, y = value, color = Allometry)) +
-#   geom_point() + theme_bw() +   xlab('LMA in g/m2') +
-#   ylab('Leaf lifespan (LL in months)')
-# h <- ggplot(data, aes(x = Nmass, y = value, color = Allometry)) +
-#   geom_point() + theme_bw() +   xlab('Nmass in %') +
-#   ylab('Leaf lifespan (LL in months)')
-# f <- ggplot(data, aes(x = wsg, y = value, color = Allometry)) +
-#   geom_point() + theme_bw() +   xlab('wsg in mg/mm3') +
-#   ylab('Leaf lifespan (LL in months)')
-# plot_grid(g, h, f, labels = c('LMA', 'Nmass', 'wsg'))
+center and left) graph.', fig.height=12, fig.width=12}
+library(ggplot2)
+load("~/Documents/ECOFOG/TROLL/inst/doc/LL_models/M11.LL.Rdata")
+fit <- fit$full
+beta <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('beta_1', 'beta_2', 'beta_3')]
+pred <- apply(as.matrix(fit), 1, function(x)
+  sapply(seq_len(dim(data)[1]), function(i)
+      rlnorm(1, x['beta_1']*(data$LMA[i]/LMAmax) - x['beta_2']*(data$Nm[i]/Nmax) + x['beta_3']*(data$wsg[i]/wsgmax), x['sigma'])
+    ))
+pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.15, 0.85)))))
+names(pred) <- c('5%', '95%')
+pred$mean <- exp((beta[1]*(data$LMA/LMAmax) - beta[2]*(data$Nm/Nmax) + beta[3]*(data$wsg/wsgmax)))
+pred$LMA <- data$LMA
+pred$Nm <- data$Nm
+pred$wsg <- data$wsg
+pred$obs <- data$LL
+lma <- ggplot(data, aes(x = LMA, y = LL)) +
+  geom_ribbon(aes(x = pred$LMA, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
+  geom_line(aes(x = pred$LMA, y = pred$`5%`, color = '5%'), linetype = 2) +
+  geom_line(aes(x = pred$LMA, y = pred$`95%`, color = '95%'), linetype = 2) +
+  geom_line(aes(x = pred$LMA, y = pred$mean, color = 'mean'), linetype = 2) +
+  geom_point() +
+  theme_bw() +
+  xlab('Leaf mass per area (LMA in g/m2)') +
+  ylab('Leaf lifespan (LL in months)') +
+  scale_color_manual(name = 'Model',
+                     values = c('5%' = 'black',
+                                '95%' = 'black',
+                                'mean' = 'red'))
+nm <- ggplot(data, aes(x = Nm, y = LL)) +
+  geom_ribbon(aes(x = pred$Nm, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
+  geom_line(aes(x = pred$Nm, y = pred$`5%`, color = '5%'), linetype = 2) +
+  geom_line(aes(x = pred$Nm, y = pred$`95%`, color = '95%'), linetype = 2) +
+  geom_line(aes(x = pred$Nm, y = pred$mean, color = 'mean'), linetype = 2) +
+  geom_point() +
+  theme_bw() +
+  xlab('Leaf nitrogen content (Nmass in %)') +
+  ylab('Leaf lifespan (LL in months)') +
+  scale_color_manual(name = 'Model',
+                     values = c('5%' = 'black',
+                                '95%' = 'black',
+                                'mean' = 'red'))
+wsg<- ggplot(data, aes(x = wsg, y = LL)) +
+  geom_ribbon(aes(x = pred$wsg, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
+  geom_line(aes(x = pred$wsg, y = pred$`5%`, color = '5%'), linetype = 2) +
+  geom_line(aes(x = pred$wsg, y = pred$`95%`, color = '95%'), linetype = 2) +
+  geom_line(aes(x = pred$wsg, y = pred$mean, color = 'mean'), linetype = 2) +
+  geom_point() +
+  theme_bw() +
+  xlab('Leaf nitrogen content (Nmass in %)') +
+  ylab('Leaf lifespan (LL in months)') +
+  scale_color_manual(name = 'Model',
+                     values = c('5%' = 'black',
+                                '95%' = 'black',
+                                'mean' = 'red'))
+predvsobs<- ggplot(pred, aes(x = obs, y = mean)) +
+  geom_ribbon(aes(ymin =`5%`, ymax = `95%`), color = 'grey', alpha = 0.2) +
+  geom_line(aes(y = `5%`, color = '5%'), linetype = 2) +
+  geom_line(aes(y = `95%`, color = '95%'), linetype = 2) +
+  geom_line(aes(y = mean, color = 'mean'), linetype = 2) +
+  geom_point() +
+  theme_bw() +
+  xlab('Observed leaf lifespan (LL in months)') +
+  ylab('Predicted leaf lifespan (LL in months)') +
+  geom_abline(slope = 1) +
+  scale_color_manual(name = 'Model',
+                     values = c('5%' = 'black',
+                                '95%' = 'black',
+                                'mean' = 'red'))
+cowplot::plot_grid(lma, nm, wsg, predvsobs, labels = c('LMA', 'Nmass', 'wsg', 'P v. O'))
 ```
@@
 Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early edath of individuals inside simulations.
-For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above X tonnes of C/ha and realistic diameter and age distribution inside the final population.
+For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.

### 2017-06-09 docs/master-thesis.tex
A2.LL 2

@@
 
-We gathered data from TRY database to evaluate and improve previous
-models. We requested functional traits available in Bridge dataset and
-leaf lifespan for tropical species present in GLOPNET dataset and used
-in TROLL (see \ref{tab:A2traits}). We compiled GLOPNET, TRY, and DRYAD
-databases. We kept dataset given by GLOPNET as origin dataset for
-observations. It correspond to leaf lifespan and also to LMA and Nmass
-in most of cases. We measured variable importance in functionnal traits
-to explain leaf lifespan with out-of bag method applied on a random
-forest.
+We compiled functional traits from GLOPNET
+\citep{wright_worldwide_2004}, TRY \citep{Kattge2011}, and DRYAD
+\citep{chave_towards_2009} databases (see \ref{tab:A2traits}). We kept
+dataset given by GLOPNET as origin dataset for observations. Dataset
+defined as origin corresponded to leaf lifespan (\(LL\)) and most of the
+time to leaf mass per area (\(LMA\)) and leaf nitrogen content per leaf
+dry mass (\(Nmass\)). We measured variable importance in functionnal
+traits to explain leaf lifespan with out-of bag method applied on a
+random forest. Then, we used a béayesian apporach to test different
+models with growing level of complexity. We retained the model with the
+best tradeoff between model complexity (number of parameters \(K\)),
+convergence, likelihood, and prediction quality (root mean sqaure error
+of prediciton \(RMSEP\)). We finally tested the new allometry obtained
+with the selected model with TROLL simulations.
 
@@
 
-We tested different models and we kept the model with the best tradeoff
-between model complexity, convergence, likelihood, and prediction
-quality (see \href{LL2.html}{leaf lifespan model} document). We found
-following model, named \textbf{M1}:
-\[M1~~ {LL_s}_j \sim \mathcal{logN}(\beta_0*e^{{\beta_1}_s*{LMA_s}_j - {\beta_2}_s*{Nmass_s}_j},\,\sigma)\,\]
-\[s=1,...,S_{=4}~, ~~j=1,...,n_s\]
-\[{\beta_i}_s \sim \mathcal{N}({\beta_i},\,\sigma_i)\,^I, ~~(\beta_i, \sigma, \sigma_i) \sim \mathcal{\Gamma}(0.001,\,0.001)\,^{2I+1}\]
-
-Leaf lifespan (\(LL\)) is following a lognormal law with location
-\({\mu_s}_j\) and scale \(\sigma\). \(s\) represents the dataset used to
-do the fit between 1 and \(S=4\), it encompass environmental and
-protocol variations. \(j\) represents the observation for a given site
-\(s\). Location \({\mu_s}_j\) is the exponential from the difference
-between LMA and Nmass with parameters \({\beta_0}_s\), \({\beta_1}_s\),
-and \({\beta_2}_s\). Each \({\beta_i}_s\) is following a normal law
-located on \(\beta_i\) with scale \(\sigma_i\). Finally all \(beta_i\),
-\(\sigma\), and \(\sigma_i\) are assumed without presemption following
-gamma laws of parameters \(10^-2\), \(10^-2\).
-
-We decided to add to model \textbf{M1}, wood specific gravity as a new
-predictor through a parameter \(\beta_3\) in a new model named
-\textbf{M10}. We removed dataset random effect on \(\beta_3\) due to the
-fact that all wood specific gravities were provided by DRYAD database.
-Due to a correlation between \(\beta_0\) and \(\beta_3\) in \textbf{M10}
-(see figure 5), we tried also a model \textbf{M11} without intercept
-\(\beta_0\). We then sampled all models with this composite dataset, and
-obtained following results :
-
 \subsection{Results}\label{results-1}
 
+Out of the bag method applied on a random forest highlighted the
+importance of leaf nitrogen content per leaf dry mass (\(N_m\)) to model
+leaf lifespan (see \ref{tab:A2OOB}). \(N_m\) importance was higher than
+leaf mass per area (158 against 96 percent of mean square error
+increase) which was used as a proxy for leaf lifespan in previous
+models. Finally, wood specific gravity (\(wsg\)) add also a significant
+importance in leaf lifespan estimation.
+
 \begin{longtable}[]{@{}lrr@{}}
-\caption{\label{tab:OOB}Variable importance calculated with out-of the bag
-method applied on a random forest. Leaf lifespan (LL) is taken in
-GLOPNET database from \citet{wright_worldwide_2004}. Leaf mass per area
-(LMA), and leaf nitrogen content (Nmass) are taken both in TRY
+\caption{\label{tab:A2OOB}Variable importance calculated with out-of the bag
+method applied on a random forest. First column represents the mean
+decrease in mean square error (\%IncMSE) whereas second column
+represents the total decrease in node impurities, measured by the Gini
+Index (IncNodePurety). Leaf lifespan (LL) is taken in GLOPNET database
+from \citet{wright_worldwide_2004}. Leaf mass per area (LMA), and leaf
+nitrogen content (Nmass) are taken both in TRY
 (\url{https://www.try-db.org}) and GLOPNET databases. Wood specific
@@
 
-Table 3: Model fit results for composite dataset. \emph{Data are taken
-from GLOPNET, TRY, and DRYAD datasets (see previous description). ML
-stand for maximum likelihood and RMSEP for root mean square error of
-prediction}
-
-All model have correctly converged.
-
-Model \textbf{M10} have shown a correlation between parameters
-\(\beta_0\) and \(\beta_3\), responsible for the creation of model
-\textbf{M11} (see previous paragraphs). Model \textbf{M11} only show a
-slight link between \(\beta_1\) and \(\beta_3\) which can be considered
-as non significant.
-
-We can denote that retained vector of parameters with maximum likelihood
-for model \textbf{M11} is close to 95\% confidence limit.
-
-We finally tested the new allometry obtained with the new model
-\textbf{M} on TROLL data (figure 7) and with TROLL simulations. Minimum
-leaf lifespan is now above 3 months for the new model \textbf{M} with
-TROLL species dataset. Moreover lowest leaf lifespan species are not the
-one with the smallest LMA with new model. Lastly, new model \textbf{M}
-variation of leaf lifespan is almost spanning all previous allometries
-prediciton on LMA axis.
+The selected model had a maximum likelihood of 13.6 and a RMSEP of 12
+months:
+
+\begin{equation}
+  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
+  \label{eq:LL}
+\end{equation}
+
+Leaf lifespan \(LL\) follows a lognormal law with location infered from
+leaf lifespan \(LMA\), nitrogen content \(N\) and wood specific gravity
+\(wsg\) and a scale \(\sigma\). Each \({\beta_i}_d\) is following a
+normal law located on \(\beta_i\) with a scale of \(\sigma_i\). All
+\(\beta_i\), \(\sigma_i\), and \(\sigma\) are assumed without
+presemption following a gamma law. \(d\) represents the dataset random
+effects and encompass environmental and protocol variations.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A2LLpred-1.pdf}
+\caption{\label{fig:A2LLpred}Leaflifespan predictions for the selected model
+with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood
+specific gravity (wsg) and predicted versus observed values. Leaf
+lifespan (LL) is predicted with model M10 fit. Leaf mass per area (LMA)
+and leaf nitrogen content (Nmass), and wood specific gravity (wsg) are
+taken in a composite dataset of GLOPNET, TRY and DRYAD datasets.Warning
+LMA (resp. Nmass and wsg) is not constant and depend on the closest
+point value for right (resp. center and left) graph.}
+\end{figure}
 
@@
 allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic
-aboveground biomass above X tonnes of C/ha and realistic diameter and
-age distribution inside the final population.
+aboveground biomass above 400 \(tonC.ha^{-1}\) and realistic diameter
+and age distribution inside the final population.
 

### 2017-06-09 03-Model.Rmd
Appendix 3

@@
 
-20 to 30 % of designated trees are considered as rotten once probed by the lumberman.
+20 to 30 % of designated trees are considered as rotten once probed by the lumberman, and thus not harvested.
 Rotten trees are not random and depends both on tree species and diameter.
-
-**Explain rotten bayesian model**
+We gathered data from the forest office (ONF, Laurent Descroix, personnal communication) inventories precising if tree were probbed as rotten and their corresponding species and diameter.
+In addition, tree plots and sawed volume was informed.
+We then used a bayesian approach to model the link between tree species and diameter and their risk to be probbed as rotten by the lumberman.
+We test different models with growing level of complexity and kept the model with the best tradeoff between complexity (number of parameters), convergence, likelihood, and prediction quality (root mean square error of prediction RMSEP):
 
 \begin{equation}
-  P_{rotten} = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
+  \begin{array}{c} 
+    probbed~rotten \sim \mathcal{B}(P(probbed~rotten)) \\
+    P(probbed~rotten) = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
+  \end{array}
   (\#eq:rotten)
@@
 
-(see [Appendix 3: Rotten tree model] for more details)
+Tree $probbed~rotten$ follows a $Bernoulli$ law of probability $P(probbed~rotten)$.
+The odds for a tree to be probbed as rotten are calculated with the sum of a base odd to be rotten $\beta_0$ and a diameter dependent odd calculated with $\beta_1$.
+The probability for a tree to be probbed as rotten $P(probbed~rotten)$ is finally calculated by taking the inverse logit $logit^{-1}$ of the odd (see [Appendix 3: Rotten tree model] for more details).
 
@@
 We decided to model explicitly long term logging damages because of their localised nature through a gap damages model.
-We gathered data from Paracou dataset **Cite** in cenususes between 1988 and 1992 on Paracou harvested plots.
+We gathered data from Paracou dataset [@Guehl2004] in cenususes between 1988 and 1992 on Paracou harvested plots.
 Individuals were categorized between alive, dead, or recruited during the period.
@@
 We then used a bayesian approach to test the link between tree death in the four years following the log event and distance to the closest gap.
-We adapted the model from **Cite Herault** based on a disturbance index into.
+We adapted the model from @Herault2010 based on a disturbance index into:
 
@@
   \begin{array}{c} 
-    Death_i \sim \mathcal{B}(P(Death_i)) \\
-    P(Death_i) = logit^{-1}(\theta + \beta*e^{\alpha*d_{gaps}}) = \frac{e^{\theta + \beta*e^{\alpha*d_{gaps}}}}{1 + e^{\theta + \beta*e^{\alpha*d_{gaps}}}}
+    Death \sim \mathcal{B}(P(Death)) \\
+    P(Death) = logit^{-1}(\theta + \beta*e^{\alpha*d_{gaps}}) = \frac{e^{\theta + \beta*e^{\alpha*d_{gaps}}}}{1 + e^{\theta + \beta*e^{\alpha*d_{gaps}}}}
   \end{array}
@@
 
-$Death$ of a tree $i$ follows a $\mathcal{Bernoulli}$ law of probability $P(Death_i)$.
-The odds for a tree $i$ to die $O(Death_i)$ is calculated by the sum of the natural tree death odd $\theta$ and a perturbation index $\beta*e^{\alpha*d_{gaps}}$.
+$Death$ of a tree follows a $\mathcal{Bernoulli}$ law of probability $P(Death)$.
+The odds for a tree to die are calculated with the sum of the natural tree death odd $\theta$ and a perturbation index $\beta*e^{\alpha*d_{gaps}}$.
 The perturbation index depend on the distance $d_{gaps}$ of the tree $i$ to the closest logging gap.
-The probability for a tree $i$ to die $P(Death_i)$ is finally calculated by taking the inverse logit of the odd $logit^{-1}[O(Death_i)]$ (see [Appendix 4: Gap damages model] for more details).
+The probability for a tree to die $P(Death)$ is finally calculated by taking the inverse logit $logit^{-1}$ of the odd.

### 2017-06-09 08-Appendix2.Rmd
Appendix 3

@@
   TRYcode = c(12, 11, 15, 14, 4)),
-  caption = 'Functional traits gathered with TRY.')
+  caption = 'Functional traits gathered with TRY.', format = 'pandoc')
 ```
@@
 Nmass and wsg) is not constant and depend on the closest point value for right (resp.
-center and left) graph.', fig.height=12, fig.width=12}
+center and left) graph.', cahce=TRUE, fig.height=12, fig.width=12}
 library(ggplot2)

### 2017-06-09 09-Appendix3.Rmd
Appendix 3

# Appendix 3: Rotten tree model

In order to simulate sylviculture with TROLL we need to implement a new sylviculture module inside TROLL model code. A first litterature review was completed by an interview with Laurent Descroix of the Office Nationale des Forêts. We discovered that rotten trees were not random and seemed to depend both on tree species and diameter. This document presents modelling of relation between rotten trees and their species and diameter.

In fact we have two different questions:

- Predict if a tree will be probed as rotten (models **M**) 
- Predict how much of tree volume is rotten (models **N**)

First all **M** model can be written as follow:
$$Rotten_n \sim \mathcal{B}(\theta_n), ~~n \in [1,N_{=3816}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]$$
Secondly, all **N** models depend on a latent variable being the percentage of rotten wood $Pt_r$. We can assume that all trees are growing depending on species $s$ and plot $p$ fertility and are supposed to have a full healthy volume $V_h$ for a given diameter $dbh$. We obtain following model:

$$V_f \sim \mathcal{logN}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]$$

We retained following models :

```{r A3models, echo=FALSE}
mpath <- "~/Documents/ECOFOG/TROLL/inst/doc/rotten_models"
mtab <- data.frame(
  M = c('$M_{s,p}$', '$N_{s,p} + L_{s,p}$'),
  Model = c(
    "${P_{rotten}}_n \\sim \\mathcal{B}(inv_{logit}(\\beta_0 + \\beta_1*dbh_n + {\\beta_2}_p + {\\beta_3}_s))$",
    "$Volume_{of~wood} \\sim \\mathcal{logN} (log[((\\beta + \\beta_p + \\beta_s)*dbh^2)*(1 - Pr*((\\theta + \\theta_p + \\theta_s) *dbh^2))], \\sigma)$"
    )
  )
mtab <- apply(mtab, 2, as.character)
knitr::kable(mtab, caption = 'Models summary.', format = 'pandoc')
row.names(mtab) <- c('Mps', 'Nps')
```

## Probed rotten (M)

Based on complexity (number of parameters), convergence and likelihood we selected model $M_{p,s}$:

`r paste(mtab['Mps',], collapse = ': ')`

```{r A3M, include=FALSE, cache=TRUE}
library(dplyr)
data <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/Rotten/rotten_data.csv")
data <- data[c('num', 'Parcelle', 'up', 'ess', 'abattu', 'tout BE', 'sonde_creux', 'sonde_creux abatt', 'sonde_creux_refus', 'DBH_desi', 'V_Est', 'vol_BO')]
names(data) <- c('id', 'plot', 'subplot', 'spCode', 'logged', 'rotten_full', 'probed_rotten', 'probed_rotten_logged', 'probed_rotten_refused', 'dbhest', 'Vest', 'Vbo')
species <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/Rotten/rotten_species.csv")[-1,]
species <- species[c("Code \nEssence" , "Nom \nscientifique")]
names(species) <- c('spCode', 'species')
species$wsg <- apply(do.call('rbind', lapply(strsplit(species$species, ' ', fixed = T), function(x) x[1:2])),
                   1, function(x) getWoodDensity(x[1], x[2], region = 'SouthAmericaTrop')$meanWD)
data <- data.frame(left_join(data, species)) ; rm(species)
data <- data[-which(is.nan(data$wsg)),]
data$sp <- apply(data[c('plot', 'species')], 1 , paste, collapse = '.')
dbhmax <- max(data$dbhest)
data$dbh <- data$dbhest/dbhmax
wsgmax <- max(data$wsg)
load(file.path(mpath, 'Mps.Rdata'))
data2 <- group_by(data, dbhest) %>%
  summarise_each(funs(mean))
inv_logit <- function(u) {
  return(1 / (1 + exp(-u)));
}
beta <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('beta_0', 'beta_1')]
Pr <- function(dbh, dbhmax, beta0, beta1){
  theta <- beta0 + beta1*(dbh/dbhmax)
  inv_logit(theta)
}
pred <- apply(as.matrix(fit), 1, function(x)
  sapply(seq_len(dim(data2)[1]), function(i)
    Pr(data2$dbhest[i], dbhmax, x['beta_0'], x['beta_1'])
  ))
pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.05, 0.95)))))
names(pred) <- c('5%', '95%')
pred$mean <- inv_logit(beta[1] + (beta[2]/dbhmax)*data2$dbhest)
pred$dbh <- data2$dbhest
```

```{r A3Mpred, echo=FALSE, caption='Models prediction. Probability to be probed as rotten P for a given diameter at breast hieght (cm).'}
library(ggplot2)
ggplot(data2, aes(x = dbhest, y = probed_rotten)) +
  geom_ribbon(aes(x = pred$dbh, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$dbh, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$dbh, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$dbh, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  theme_bw() +
  xlab('Diameter at breast height (cm)') +
  ylab('Probability to be probed as rotten') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
```

```{r A3Mtab, echo=FALSE}
table <- data.frame(row.names = seq(45, 125, 5),
           `P` = round(sapply(seq(45, 125, 5), function(dbh) 100*inv_logit(beta[1]+beta[2]*dbh/dbhmax))))
knitr::kable(t(table), caption = 'Models prediction. Probability to be probed as rotten (P in %) for a given dbh (cm).', format = 'pandoc')
```

## Rotten volume (N)

Based on complexity (number of parameters), convergence and likelihood we selected model $N_{p,s}$ associated to hyperparameter $\rho$ with model $L_{p,s}$:

`r paste(mtab['Nps',], collapse = ': ')`

```{r A3N, include=FALSE, cache=TRUE}
data <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/Rotten/rotten_data.csv")
data <- data[c('num', 'Parcelle', 'up', 'ess', 'abattu', 'tout BE', 'sonde_creux', 'sonde_creux abatt', 'sonde_creux_refus', 'DBH_desi', 'V_Est', 'vol_BO')]
names(data) <- c('id', 'plot', 'subplot', 'spCode', 'logged', 'rotten_full', 'probed_rotten', 'probed_rotten_logged', 'probed_rotten_refused', 'dbhest', 'Vest', 'Vbo')
data <- subset(data, !is.na(Vbo))
data <- subset(data, Vbo!=0)
data <- data[c('id', 'plot', 'spCode', 'probed_rotten', 'Vbo', 'dbhest')]
species <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/Rotten/rotten_species.csv")[-1,]
species <- species[c("Code \nEssence" , "Nom \nscientifique")]
names(species) <- c('spCode', 'species')
data <- data.frame(left_join(data, species)) ; rm(species)
dbhmax <- max(data$dbhest)
data2 <- data[sample.int(dim(data)[1], 500, replace = F),]
load(file.path(mpath, 'Lps.Rdata'))
pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('beta', 'sigma', 'theta')]
Vf <- function(dbh, dbhmax, Pr, pars){
  rho <- pars[3]*(dbh/dbhmax)*(dbh/dbhmax)
  mu <- (pars[1]*(dbh/dbhmax)*(dbh/dbhmax))*(1-Pr*rho)
  rlnorm(1, log(mu), pars[2])
}
pred <- apply(as.matrix(fit), 1, function(x)
  sapply(seq_len(dim(data2)[1]), function(i)
    Vf(data2$dbhest[i], dbhmax, data2$probed_rotten[i], c(x['beta'], x['sigma'], x['theta']))
  ))
pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.05, 0.95), na.rm = T))))
names(pred) <- c('5%', '95%')
pred$mean <- (pars[1]*(data2$dbhest/dbhmax)*(data2$dbhest/dbhmax))*(1-data2$probed_rotten*(pars[3]*(data2$dbhest/dbhmax)*(data2$dbhest/dbhmax)))
pred$dbh <- data2$dbhest
```

```{r A3Npred, echo=FALSE, caption='Models prediction. Final volume of wood for a given diameter at breast hieght (cm).'}
ggplot(data2, aes(x = dbhest, y = Vbo)) +
  geom_ribbon(aes(x = pred$dbh, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$dbh, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$dbh, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$dbh, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() + geom_jitter() +
  theme_bw() +
  xlab('Estimated dbh (cm)') +
  ylab('Final softwood product (m3)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
```

```{r A3Ntab, echo=FALSE}
table <- data.frame(row.names = seq(45, 125, 5),
           Vf = round(sapply(seq(45, 125, 5), function(dbh) (pars[1]*(dbh/dbhmax)^2)*(1-pars[3]*(dbh/dbhmax)^2)),2),
           Vp = round(sapply(seq(45, 125, 5), function(dbh) 100*pars[3]*(dbh/dbhmax)^2)))
knitr::kable(t(table), caption = 'Models prediction. Final volume of wood ($V_f$ in $m^3$) and percent of rotten wood ($V_p$ in %) for a given dbh (cm) if the tree was probed rotten.', format = 'pandoc')
```

### 2017-06-09 docs/master-thesis.tex
Appendix 3

@@
 20 to 30 \% of designated trees are considered as rotten once probed by
-the lumberman. Rotten trees are not random and depends both on tree
-species and diameter.
-
-\textbf{Explain rotten bayesian model}
-
-\begin{equation}
-  P_{rotten} = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
+the lumberman, and thus not harvested. Rotten trees are not random and
+depends both on tree species and diameter. We gathered data from the
+forest office (ONF, Laurent Descroix, personnal communication)
+inventories precising if tree were probbed as rotten and their
+corresponding species and diameter. In addition, tree plots and sawed
+volume was informed. We then used a bayesian approach to model the link
+between tree species and diameter and their risk to be probbed as rotten
+by the lumberman. We test different models with growing level of
+complexity and kept the model with the best tradeoff between complexity
+(number of parameters), convergence, likelihood, and prediction quality
+(root mean square error of prediction RMSEP):
+
+\begin{equation}
+  \begin{array}{c} 
+    probbed~rotten \sim \mathcal{B}(P(probbed~rotten)) \\
+    P(probbed~rotten) = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
+  \end{array}
   \label{eq:rotten}
@@
 
-(see \protect\hyperlink{appendix-3-rotten-tree-model}{Appendix 3: Rotten
-tree model} for more details)
+Tree \(probbed~rotten\) follows a \(Bernoulli\) law of probability
+\(P(probbed~rotten)\). The odds for a tree to be probbed as rotten are
+calculated with the sum of a base odd to be rotten \(\beta_0\) and a
+diameter dependent odd calculated with \(\beta_1\). The probability for
+a tree to be probbed as rotten \(P(probbed~rotten)\) is finally
+calculated by taking the inverse logit \(logit^{-1}\) of the odd (see
+\protect\hyperlink{appendix-3-rotten-tree-model}{Appendix 3: Rotten tree
+model} for more details).
 
@@
 localised nature through a gap damages model. We gathered data from
-Paracou dataset \textbf{Cite} in cenususes between 1988 and 1992 on
+Paracou dataset \citep{Guehl2004} in cenususes between 1988 and 1992 on
 Paracou harvested plots. Individuals were categorized between alive,
@@
 the link between tree death in the four years following the log event
-and distance to the closest gap. We adapted the model from \textbf{Cite
-Herault} based on a disturbance index into.
+and distance to the closest gap. We adapted the model from
+\citet{Herault2010} based on a disturbance index into:
 
@@
 
-\(Death\) of a tree \(i\) follows a \(\mathcal{Bernoulli}\) law of
-probability \(P(Death_i)\). The odds for a tree \(i\) to die
-\(O(Death_i)\) is calculated by the sum of the natural tree death odd
-\(\theta\) and a perturbation index \(\beta*e^{\alpha*d_{gaps}}\). The
-perturbation index depend on the distance \(d_{gaps}\) of the tree \(i\)
-to the closest logging gap. The probability for a tree \(i\) to die
-\(P(Death_i)\) is finally calculated by taking the inverse logit of the
-odd \(logit^{-1}[O(Death_i)]\) (see
-\protect\hyperlink{appendix-4-gap-damages-model}{Appendix 4: Gap damages
-model} for more details).
+\(Death\) of a tree follows a \(\mathcal{Bernoulli}\) law of probability
+\(P(Death)\). The odds for a tree to die are calculated with the sum of
+the natural tree death odd \(\theta\) and a perturbation index
+\(\beta*e^{\alpha*d_{gaps}}\). The perturbation index depend on the
+distance \(d_{gaps}\) of the tree \(i\) to the closest logging gap. The
+probability for a tree to die \(P(Death)\) is finally calculated by
+taking the inverse logit \(logit^{-1}\) of the odd.
 
@@
 
-\begin{table}
-
-\caption{\label{tab:A2traits}Functional traits gathered with TRY.}
-\centering
-\begin{tabular}[t]{l|l|l|r}
-\hline
-Name & Trait & Unit & TRYcode\\
-\hline
-LL & Leaf lifespan (longevity) & month & 12\\
-\hline
-SLA & Leaf area per leaf dry mass (specific leaf area, SLA) & \$m\textasciicircum{}2.kg\textasciicircum{}\{-1\}\$ & 11\\
-\hline
-N & Leaf nitrogen (N) content per leaf dry mass & \$mg.g\textasciicircum{}\{-1\}\$ & 15\\
-\hline
-P & Leaf phosphorus (P) content per leaf dry mass & \$mg.g\textasciicircum{}\{-1\}\$ & 14\\
-\hline
-wsg & Stem dry mass per stem fresh volume (stem specific density) & \$mg.mm\textasciicircum{}\{-3\}\$ & 4\\
-\hline
-\end{tabular}
-\end{table}
+\begin{longtable}[]{@{}lllr@{}}
+\caption{\label{tab:A2traits}Functional traits gathered with
+TRY.}\tabularnewline
+\toprule
+Name & Trait & Unit & TRYcode\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+Name & Trait & Unit & TRYcode\tabularnewline
+\midrule
+\endhead
+LL & Leaf lifespan (longevity) & month & 12\tabularnewline
+SLA & Leaf area per leaf dry mass (specific leaf area, SLA) &
+\(m^2.kg^{-1}\) & 11\tabularnewline
+N & Leaf nitrogen (N) content per leaf dry mass & \(mg.g^{-1}\) &
+15\tabularnewline
+P & Leaf phosphorus (P) content per leaf dry mass & \(mg.g^{-1}\) &
+14\tabularnewline
+wsg & Stem dry mass per stem fresh volume (stem specific density) &
+\(mg.mm^{-3}\) & 4\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 
-\hypertarget{appendix-4-gap-damages-model}{\section{Appendix 4: Gap
-damages model}\label{appendix-4-gap-damages-model}}
+In order to simulate sylviculture with TROLL we need to implement a new
+sylviculture module inside TROLL model code. A first litterature review
+was completed by an interview with Laurent Descroix of the Office
+Nationale des Forêts. We discovered that rotten trees were not random
+and seemed to depend both on tree species and diameter. This document
+presents modelling of relation between rotten trees and their species
+and diameter.
+
+In fact we have two different questions:
+
+\begin{itemize}
+\tightlist
+\item
+  Predict if a tree will be probed as rotten (models \textbf{M})
+\item
+  Predict how much of tree volume is rotten (models \textbf{N})
+\end{itemize}
+
+First all \textbf{M} model can be written as follow:
+\[Rotten_n \sim \mathcal{B}(\theta_n), ~~n \in [1,N_{=3816}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]\]
+Secondly, all \textbf{N} models depend on a latent variable being the
+percentage of rotten wood \(Pt_r\). We can assume that all trees are
+growing depending on species \(s\) and plot \(p\) fertility and are
+supposed to have a full healthy volume \(V_h\) for a given diameter
+\(dbh\). We obtain following model:
+
+\[V_f \sim \mathcal{logN}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]\]
+
+We retained following models :
+
+\begin{longtable}[]{@{}ll@{}}
+\caption{\label{tab:A3models}Models summary.}\tabularnewline
+\toprule
+M & Model\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+M & Model\tabularnewline
+\midrule
+\endhead
+\(M_{s,p}\) &
+\({P_{rotten}}_n \sim \mathcal{B}(inv_{logit}(\beta_0 + \beta_1*dbh_n + {\beta_2}_p + {\beta_3}_s))\)\tabularnewline
+\(N_{s,p} + L_{s,p}\) &
+\(Volume_{of~wood} \sim \mathcal{logN} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)\tabularnewline
+\bottomrule
+\end{longtable}
+
+\subsection{Probed rotten (M)}\label{probed-rotten-m}
+
+Based on complexity (number of parameters), convergence and likelihood
+we selected model \(M_{p,s}\):
+
+\(M_{s,p}\):
+\({P_{rotten}}_n \sim \mathcal{B}(inv_{logit}(\beta_0 + \beta_1*dbh_n + {\beta_2}_p + {\beta_3}_s))\)
+
+\includegraphics{master-thesis_files/figure-latex/A3Mpred-1.pdf}
+
+\begin{longtable}[]{@{}lrrrrrrrrrrrrrrrrr@{}}
+\caption{\label{tab:A3Mtab}Models prediction. Probability to be probed as
+rotten (P in \%) for a given dbh (cm).}\tabularnewline
+\toprule
+& 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85 & 90 & 95 & 100 & 105 & 110
+& 115 & 120 & 125\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+& 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85 & 90 & 95 & 100 & 105 & 110
+& 115 & 120 & 125\tabularnewline
+\midrule
+\endhead
+P & 4 & 4 & 5 & 7 & 8 & 10 & 12 & 14 & 17 & 20 & 24 & 28 & 32 & 37 & 42
+& 47 & 52\tabularnewline
+\bottomrule
+\end{longtable}
+
+\subsection{Rotten volume (N)}\label{rotten-volume-n}
+
+Based on complexity (number of parameters), convergence and likelihood
+we selected model \(N_{p,s}\) associated to hyperparameter \(\rho\) with
+model \(L_{p,s}\):
+
+\(N_{s,p} + L_{s,p}\):
+\(Volume_{of~wood} \sim \mathcal{logN} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)
+
+\includegraphics{master-thesis_files/figure-latex/A3Npred-1.pdf}
+
+\begin{longtable}[]{@{}lrrrrrrrrrrrrrrrrr@{}}
+\caption{\label{tab:A3Ntab}Models prediction. Final volume of wood (\(V_f\)
+in \(m^3\)) and percent of rotten wood (\(V_p\) in \%) for a given dbh
+(cm) if the tree was probed rotten.}\tabularnewline
+\toprule
+& 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85 & 90 & 95 & 100 & 105 & 110
+& 115 & 120 & 125\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+& 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85 & 90 & 95 & 100 & 105 & 110
+& 115 & 120 & 125\tabularnewline
+\midrule
+\endhead
+Vf & 1.66 & 2.02 & 2.39 & 2.78 & 3.18 & 3.58 & 3.98 & 4.37 & 4.74 & 5.09
+& 5.41 & 5.68 & 5.9 & 6.06 & 6.15 & 6.16 & 6.08\tabularnewline
+Vp & 7.00 & 9.00 & 11.00 & 13.00 & 15.00 & 18.00 & 20.00 & 23.00 & 26.00
+& 29.00 & 32.00 & 36.00 & 40.0 & 43.00 & 47.00 & 52.00 &
+56.00\tabularnewline
+\bottomrule
+\end{longtable}
 

### 2017-06-09 04-MnMs.Rmd
M&Ms 2

@@
 
-### Resilience
+### Ecosystem functions
 
 Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions.
-We decided to describe simiulated tropical forests in three major functions: forest dynamic, forest production, and forest diversity.
+We decided to describe simiulated tropical forests in two major functions: forest dynamic and forest production.
 Forest dynamic was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$).
 Forest production was represented by growth pimary productivity ($GPP$ in $MgC.ha^{-1}$), net pimary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
-Finally, forest diversity was represented by **??**.
 
@@
 Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$.
+Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest dynamic (AGB, BA, N, N10, and N30) and forest production (GPP, NPP, Rday, and Rnight).
+We then used cummulative integral over time of eulcidean distance to equilibrium to assess simulations resilience.
 
@@
 
-$X_sp$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$.
+$X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$.
 $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance.
@@
 Whereas $SE$ become positive when dominant species outperform theimselves in mixture than in monoculture, and negative when less dominant species outperform theimselves in mixture than in monoculture [@Tobner2016].
+But similarly to resilience measurement, biodiversity net effect $NE$ can be in a dynamic equilibrium and vary over time without disturbance.
+So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:
 
-Recovery trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
+\begin{equation}
+  R(NE_T) = \frac{NE_T}{NE_C} = \frac{SE_T}{NE_C} + \frac{CE_T}{NE_C}
+  (\#eq:RNE)
+\end{equation}
+
+Resilience trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
 In order to do that, the design of experiment was repeated for each species indivdually representing 652 simulations of monoculture.

### 2017-06-09 05-Results.Rmd
Results: Disturbance & Appendix4

# Results

## Sensitivity

## Disturbance

### Ecosystem functions

```{r resilience, echo=FALSE}
rich <- c(5,25,125)
cvh <- 1:20
dist <- c(0,25,50,75)
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
load(file.path(path, 'disturbanceData.Rdata'))
load(file.path(path, 'disturbanceResilience.Rdata'))
```

```{r Deq, cache=TRUE, echo=FALSE}
deq <- data.frame(
  time = resilience$time,
  sim = resilience$sim,
  structure = apply(resilience[c('agb', 'ba', 'n', 'n10', 'n30')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))),
  production = apply(resilience[c('gpp', 'npp', 'Rday', 'Rnight')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2)))
)
```

```{r Ieq, cache=TRUE, echo=FALSE}
sim <- c(sapply(c(sapply(rich, function(r) paste0(r, '.', cvh))), function(x) paste0(x, '.', dist[-1])))
Ieq <- sapply(sim, function(s){
  data_s <- subset(deq, sim ==s)
  return(cbind(data_s[c(1,2)], cumsum(data_s[-c(1,2)])))
}, simplify = F) ; rm(sim)
Ieq <- data.frame(do.call('rbind', Ieq))
names(Ieq)[3:4] <- paste0('Ieq.', names(Ieq)[3:4])
```

We transformed all ecosystem outputs from the 240 disturbance simulations in resilience metrics normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A et B). We then gathered ecosystem outputs by main ecosystem functions (forest dynamic, forest production and forest diversity) to compute ecosystem distance to equilibrium (Figure \@ref(fig:datatrans) C). Finally, we integrated distance to equilibrium in a cummulative sum over time (Figure \@ref(fig:datatrans) D).

```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation. Ecosystem outputs (**A**) are normalized by the control value over time to calculate resilience (**B**); resilience of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance ot equilibirum is integrated over time in a cummulative sum (**D**). ', cache=TRUE, fig.height=8, fig.width=12}
g_agb <- ecosystem %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Aboveground biomass (tonC/ha)') +
  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
g_resilience <- resilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb, color = disturbance)) +
  geom_hline(yintercept = 1, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Resilience of aboveground biomass') +
  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
g_deq <- deq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, color = disturbance)) +
  geom_hline(yintercept = 0, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Distance to forest dynamic equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
g_Ieq <- Ieq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = Ieq.structure, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Cummulative integral of\ndistance to forest dynamic equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
```

The ranking was stable over time for the 240 simulations. So we used the cummulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience. We compared cummulative integral after 600 years to communities taxonomic and functional diversity for each leavel of disturbance (see Figure \@ref(fig:gIeq)). We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years ($Ieq_{600}$). In addition, functional eveness was complementary reducing $Ieq_{600}$. Finally species richness was not directly link to $Ieq_{600}$. Effectivelly, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$. We found similar results for all disturbance levels and other ecosystem functions (forest production and forest diversity, see [Appendix 4: Disturbance simulations]).

```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional eveness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>% 
  ggplot(aes(x = FDiv, y = Ieq.structure, label = sim)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(shape = richness, color = FEve)) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
  xlab('Functional diversity (FDiv)') +
  ylab('Cummulative integral from ecosystem 
       distance to forest dynamic equilibrium after 600 years') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\neveness\n(FEve)')
```

### Biodiversity effect

```{r BE, echo=FALSE}
load(file.path(path, 'disturbanceBE.Rdata'))
load(file.path(path, 'disturbanceBErecovery.Rdata'))
```

We measured all ecosystem outputs biodiversity net effect in disturbance simulations by comparing theim to their species corresponding monoculture simulations. The net effect was then partition between selection and complementarity effect. We normalized treatments effect by control net effect (see Table \@ref(tab:tNE)) to measure their aboveground biomass resilience to disturbance (see Figure \@ref(fig:gBE)). We found that complementarity effect was recovering net effect in the first decades untill it disappeared after a century. On the contrary selection effect was reduced or even removed by the disturbance. It increased during the whole simulation and was greater than complementarity effect only after decades. The time lag for which complementarity effect was non null and greater than selection effect was increasing with disturbance intensity. Finally 600 years after the disturbance event biodiversity net effect was still not recovered for a disturbance intensity greater than 25% of basal area. We obtained those results for aboveground biomass. We found similar results but with an amlified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 4: Disturbance simulations]). Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 4: Disturbance simulations]).

```{r tNE, echo=FALSE, cache=TRUE}
BE %>%
  mutate(agb = agb/1000) %>% 
  filter(BE == 'NE') %>%
  melt(id.vars = c('BE', 'time', 'sim')) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>%
  filter(disturbance == 0) %>%
  group_by(variable) %>% 
  summarise(mean = mean(value, na.rm = T), 
            'standard deviation' = sd(value, na.rm = T)) %>% 
  filter(variable != 'n30') %>% 
  bind_cols(., data.frame(
    name = c('aboveground biomass', 'basal area', 'number of stems', 
             'number of stems above 10cm dbh', 'growth primary production', 
             'net primary production', 'autotrophic respiration during day',
             'autotrophic respiration during night'),
    unit = c('$tonC.ha^{-1}$', '$m^2.ha^{-1}$', '$n.ha^{-1}$', '$n.ha^{-1}$', 
             '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$'))) %>% 
  kable(caption = 'Table 1: Biodiversity net effect mean value and standard deviation for different ecosystem variable.', digits = 3, format = 'pandoc')
```

```{r gBE, echo=FALSE, fig.cap='Resilience of complementarity and selection effects. Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time.', cache=TRUE, fig.width=12}
BEresilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(BE %in% c('SE', 'CE')) %>% 
  group_by(time, BE, disturbance, richness) %>%
  summarise(qt1 =quantile(agb, probs=0.05),
            qt2 =quantile(agb, probs=0.75),
            mean=mean(agb),
            n=n()) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_grid(richness ~ disturbance, labeller = 'label_both') +
  xlab('Time (years)') +
  ylab('Recovery of aboveground biomass net effect')
```

## Sylviculture

### 2017-06-09 10-Appendix4.Rmd
Results: Disturbance & Appendix4

# Appendix 4: Disturbance simulations

```{r A4.data, echo=FALSE, cache=TRUE}
rich <- c(5,25,125)
cvh <- 1:20
dist <- c(0,25,50,75)
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
load(file.path(path, 'disturbanceResilience.Rdata'))
deq <- data.frame(
  time = resilience$time,
  sim = resilience$sim,
  structure = apply(resilience[c('agb', 'ba', 'n', 'n10', 'n30')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))),
  production = apply(resilience[c('gpp', 'npp', 'Rday', 'Rnight')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))))
sim <- c(sapply(c(sapply(rich, function(r) paste0(r, '.', cvh))), function(x) paste0(x, '.', dist[-1])))
Ieq <- sapply(sim, function(s){
  data_s <- subset(deq, sim ==s)
  return(cbind(data_s[c(1,2)], cumsum(data_s[-c(1,2)])))
}, simplify = F) ; rm(sim)
Ieq <- data.frame(do.call('rbind', Ieq))
names(Ieq)[3:4] <- paste0('Ieq.', names(Ieq)[3:4])
load(file.path(path, 'disturbanceBE.Rdata'))
load(file.path(path, 'disturbanceBErecovery.Rdata'))
```

## Ecosystem functions

Appendix 4 presents ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. It encompass all functionnal diversity components [FRIC, FEve, FDiv, and FDis, @villeger_new_2008]. And it presents results for both forest dynamic (Figure \@ref(fig:A4IeqStructureAll)) and forest production (Figure \@ref(fig:A4IeqProductionAll)).

```{r A4IeqStructureAll, echo=FALSE, fig.cap='Figure S1.1: Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>%
  select(Ieq.structure, disturbance, richness, FRic, FEve, FDiv, FDis) %>% 
  melt(id.vars = c('Ieq.structure', 'disturbance', 'richness')) %>% 
  ggplot(aes(x = Ieq.structure, y = value)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(shape = richness)) +
  facet_grid(variable~disturbance, scales = "free", labeller = 'label_both') + 
  ylab('Functional diversity') +
  xlab('Cummulative integral from ecosystem 
       distance to forest dynamic equilibrium after 600 years') +
  scale_shape_discrete('Species\nrichness\n(SR)')
```

```{r A4IeqProductionAll, echo=FALSE, fig.cap='Figure S1.2: Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. Cummulative integral from ecosystem distance to forest production equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>%
  select(Ieq.production, disturbance, richness, FRic, FEve, FDiv, FDis) %>% 
  melt(id.vars = c('Ieq.production', 'disturbance', 'richness')) %>% 
  ggplot(aes(x = Ieq.production, y = value)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(shape = richness)) +
  facet_grid(variable~disturbance, scales = "free", labeller = 'label_both') + 
  ylab('Functional diversity') +
  xlab('Cummulative integral from ecosystem 
       distance to forest production equilibrium after 600 years') +
  scale_shape_discrete('Species\nrichness\n(SR)')
```

## Biodiversity effect

Figure \@ref(fig:A4allBE) presents the resilience of complementarity and selection effects for different ecosystem metrics (AGB, BA, N, GPP and NPP).

```{r A4allBE, echo=FALSE, fig.cap='Resilience of complementarity and selection effects. Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time for different ecosystem variables (AGB, BA, N, GPP).', cache=TRUE, fig.height=8, fig.width=12}
BEresilience %>% 
  select(BE, time, sim, agb, ba, n, gpp) %>% 
  melt(id.vars = c('BE', 'time', 'sim')) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>%
  filter(BE %in% c('SE', 'CE')) %>% 
  group_by(time, BE, disturbance, variable) %>%
  summarise(qt1 =quantile(value, probs=0.05, na.rm = T),
            qt2 =quantile(value, probs=0.75, na.rm = T),
            mean=mean(value, na.rm = T)) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_grid(variable ~ disturbance, scales = "free", labeller = 'label_both') +
  xlab('Time (years)') +
  ylab('Recovery of variable net effect') + 
  ylim(-2.5,2.5)
```

### 2017-06-09 docs/master-thesis.tex
Results: Disturbance & Appendix4

@@
 
-\subsubsection{Resilience}\label{resilience}
+\subsubsection{Ecosystem functions}\label{ecosystem-functions}
 
@@
 to several ecosystem functions. We decided to describe simiulated
-tropical forests in three major functions: forest dynamic, forest
-production, and forest diversity. Forest dynamic was represented by
-aboveground biomass (\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\)
-in \(m^2.ha^{-1}\)), total number of stem (\(N\)), number of stem above
-\(10~cm\) diameter (\(N10\)), and number of stem above \(30~cm\)
-diameter (\(N30\)). Forest production was represented by growth pimary
-productivity (\(GPP\) in \(MgC.ha^{-1}\)), net pimary productivity
-(\(NPP\) in \(MgC.ha^{-1}\)), tree autotrophic respiration in day
-(\(Rday\) in \(MgC.ha^{-1}\)) and tree autotrophic respiration in night
-(\(Rnight\) in \(MgC.ha^{-1}\)). Finally, forest diversity was
-represented by \textbf{??}.
+tropical forests in two major functions: forest dynamic and forest
+production. Forest dynamic was represented by aboveground biomass
+(\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\) in \(m^2.ha^{-1}\)),
+total number of stem (\(N\)), number of stem above \(10~cm\) diameter
+(\(N10\)), and number of stem above \(30~cm\) diameter (\(N30\)). Forest
+production was represented by growth pimary productivity (\(GPP\) in
+\(MgC.ha^{-1}\)), net pimary productivity (\(NPP\) in \(MgC.ha^{-1}\)),
+tree autotrophic respiration in day (\(Rday\) in \(MgC.ha^{-1}\)) and
+tree autotrophic respiration in night (\(Rnight\) in \(MgC.ha^{-1}\)).
 
@@
 euclidean distance to equilibrium \(d_{eq}(t)\) as
-\(d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}\).
+\(d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}\). Ecosystem euclidean distance
+to equilibrium was calculated in a multi-dimensional space for the two
+functions described above: forest dynamic (AGB, BA, N, N10, and N30) and
+forest production (GPP, NPP, Rday, and Rnight). We then used cummulative
+integral over time of eulcidean distance to equilibrium to assess
+simulations resilience.
 
@@
 
-\(X_sp\) is the ecosystem variable value for one species either in
+\(X_{sp}\) is the ecosystem variable value for one species either in
 mixture \(X_{sp}(mixture)\) or in monoculture \(X_{sp}(monoculture)\).
@@
 less dominant species outperform theimselves in mixture than in
-monoculture \citep{Tobner2016}.
-
-Recovery trajectories of ecosystem variable after disturbance were
+monoculture \citep{Tobner2016}. But similarly to resilience measurement,
+biodiversity net effect \(NE\) can be in a dynamic equilibrium and vary
+over time without disturbance. So in order to correctly assess selection
+and complementarity effect in answer to disturbance, we normalized it by
+undisturbed control ecosystem net effect \(NE_C\) to measure treatments
+net effect resilience \(R(NE_T)\):
+
+\begin{equation}
+  R(NE_T) = \frac{NE_T}{NE_C} = \frac{SE_T}{NE_C} + \frac{CE_T}{NE_C}
+  \label{eq:RNE}
+\end{equation}
+
+Resilience trajectories of ecosystem variable after disturbance were
 partitioned between complementarity effect \(CE\) and selection effect
@@
 \subsection{Disturbance}\label{disturbance-1}
+
+\subsubsection{Ecosystem functions}\label{ecosystem-functions-1}
+
+We transformed all ecosystem outputs from the 240 disturbance
+simulations in resilience metrics normalizing the treatment values by
+their corresponding control (Figure \ref{fig:datatrans} A et B). We then
+gathered ecosystem outputs by main ecosystem functions (forest dynamic,
+forest production and forest diversity) to compute ecosystem distance to
+equilibrium (Figure \ref{fig:datatrans} C). Finally, we integrated
+distance to equilibrium in a cummulative sum over time (Figure
+\ref{fig:datatrans} D).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/datatrans-1.pdf}
+\caption{\label{fig:datatrans}Ecosystem outputs data transformation.
+Ecosystem outputs (\textbf{A}) are normalized by the control value over
+time to calculate resilience (\textbf{B}); resilience of different
+ecosystem outputs is then used in a multidimensional space to caluclate
+ecosystem distance to equilibrium (\textbf{C}); finally distance ot
+equilibirum is integrated over time in a cummulative sum (\textbf{D}).}
+\end{figure}
+
+The ranking was stable over time for the 240 simulations. So we used the
+cummulative integral after 600 years \(Ieq_{600}\) as a measurement of
+ecosystem resilience. We compared cummulative integral after 600 years
+to communities taxonomic and functional diversity for each leavel of
+disturbance (see Figure \ref{fig:gIeq}). We found that increased
+functional diversity \citep[FDiv,][]{villeger_new_2008} was reducing
+cummulative integral from ecosystem distance to forest dynamic
+equilibrium after 600 years (\(Ieq_{600}\)). In addition, functional
+eveness was complementary reducing \(Ieq_{600}\). Finally species
+richness was not directly link to \(Ieq_{600}\). Effectivelly, low
+species richness could result in variant \(Ieq_{600}\), but increased
+species richness resulted in increased functional diversity and
+consequently lower \(Ieq_{600}\). We found similar results for all
+disturbance levels and other ecosystem functions (forest production and
+forest diversity, see
+\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
+Disturbance simulations}).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/gIeq-1.pdf}
+\caption{\label{fig:gIeq}Ecosystem resilience after 600 years with taxonomic
+and functional diversity for different levels of disturbance.
+Cummulative integral from ecosystem distance to forest dynamic
+equilibrium after 600 years was represented against functional diversity
+\citep[FDiv,][]{villeger_new_2008} for different level of disturbance
+(25, 50 and 75\% of total basal area); dot shapes represents the species
+richness whereas dot color represents functional eveness
+\citep[FEve,][]{villeger_new_2008}.}
+\end{figure}
+
+\subsubsection{Biodiversity effect}\label{biodiversity-effect-1}
+
+We measured all ecosystem outputs biodiversity net effect in disturbance
+simulations by comparing theim to their species corresponding
+monoculture simulations. The net effect was then partition between
+selection and complementarity effect. We normalized treatments effect by
+control net effect (see Table \ref{tab:tNE}) to measure their
+aboveground biomass resilience to disturbance (see Figure
+\ref{fig:gBE}). We found that complementarity effect was recovering net
+effect in the first decades untill it disappeared after a century. On
+the contrary selection effect was reduced or even removed by the
+disturbance. It increased during the whole simulation and was greater
+than complementarity effect only after decades. The time lag for which
+complementarity effect was non null and greater than selection effect
+was increasing with disturbance intensity. Finally 600 years after the
+disturbance event biodiversity net effect was still not recovered for a
+disturbance intensity greater than 25\% of basal area. We obtained those
+results for aboveground biomass. We found similar results but with an
+amlified signal for basal area (\(BA\)) and stem abundance (\(N\) but
+with an inverted signal because of the forest self thinning) (see
+\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
+Disturbance simulations}). Finally, forest growth primary productivity
+(\(GPP\)) recovered in few years (proportionnaly to disturbance), and
+its net effect was maintained by complementarity effect (see
+\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
+Disturbance simulations}).
+
+\begin{longtable}[]{@{}lrrll@{}}
+\caption{\label{tab:tNE}Table 1: Biodiversity net effect mean value and
+standard deviation for different ecosystem variable.}\tabularnewline
+\toprule
+variable & mean & standard deviation & name & unit\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+variable & mean & standard deviation & name & unit\tabularnewline
+\midrule
+\endhead
+agb & 32.721 & 17.839 & aboveground biomass &
+\(tonC.ha^{-1}\)\tabularnewline
+ba & 1.633 & 0.743 & basal area & \(m^2.ha^{-1}\)\tabularnewline
+n & 103.880 & 231.623 & number of stems & \(n.ha^{-1}\)\tabularnewline
+n10 & -6.815 & 13.003 & number of stems above 10cm dbh &
+\(n.ha^{-1}\)\tabularnewline
+gpp & 0.147 & 0.047 & growth primary production &
+\(MgC.ha^{-1}\)\tabularnewline
+npp & -0.041 & 0.036 & net primary production &
+\(MgC.ha^{-1}\)\tabularnewline
+Rday & 0.046 & 0.018 & autotrophic respiration during day &
+\(MgC.ha^{-1}\)\tabularnewline
+Rnight & 0.074 & 0.030 & autotrophic respiration during night &
+\(MgC.ha^{-1}\)\tabularnewline
+\bottomrule
+\end{longtable}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/gBE-1.pdf}
+\caption{\label{fig:gBE}Resilience of complementarity and selection effects.
+Complementarity effect (CE) and selection effect (SE) where normalized
+by control net effect (NEc), thus measuring their resilience over time.}
+\end{figure}
 
@@
 
+\hypertarget{appendix-4-disturbance-simulations}{\section{Appendix 4:
+Disturbance simulations}\label{appendix-4-disturbance-simulations}}
+
+\subsection{Ecosystem functions}\label{ecosystem-functions-2}
+
+Appendix 4 presents ecosystem resilience after 600 years with taxonomic
+and functional diversity for different levels of disturbance. It
+encompass all functionnal diversity components \citep[FRIC, FEve, FDiv,
+and FDis,][]{villeger_new_2008}. And it presents results for both forest
+dynamic (Figure \ref{fig:A4IeqStructureAll}) and forest production
+(Figure \ref{fig:A4IeqProductionAll}).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4IeqStructureAll-1.pdf}
+\caption{\label{fig:A4IeqStructureAll}Figure S1.1: Ecosystem resilience
+after 600 years with taxonomic and functional diversity for different
+levels of disturbance. Cummulative integral from ecosystem distance to
+forest dynamic equilibrium after 600 years was represented against
+functional diversity \citep[FRIC, FEve, FDiv, and
+FDis,][]{villeger_new_2008} for different level of disturbance (25, 50
+and 75\% of total basal area); dot shapes represents the species
+richness.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4IeqProductionAll-1.pdf}
+\caption{\label{fig:A4IeqProductionAll}Figure S1.2: Ecosystem resilience
+after 600 years with taxonomic and functional diversity for different
+levels of disturbance. Cummulative integral from ecosystem distance to
+forest production equilibrium after 600 years was represented against
+functional diversity \citep[FRIC, FEve, FDiv, and
+FDis,][]{villeger_new_2008} for different level of disturbance (25, 50
+and 75\% of total basal area); dot shapes represents the species
+richness.}
+\end{figure}
+
+\subsection{Biodiversity effect}\label{biodiversity-effect-2}
+
+Figure \ref{fig:A4allBE} presents the resilience of complementarity and
+selection effects for different ecosystem metrics (AGB, BA, N, GPP and
+NPP).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4allBE-1.pdf}
+\caption{\label{fig:A4allBE}Resilience of complementarity and selection
+effects. Complementarity effect (CE) and selection effect (SE) where
+normalized by control net effect (NEc), thus measuring their resilience
+over time for different ecosystem variables (AGB, BA, N, GPP).}
+\end{figure}
+
 \addcontentsline{toc}{section}{References}

### 2017-06-09 05-Results.Rmd
Steph1

@@
 Effectivelly, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$.
-We found similar results for all disturbance levels and other ecosystem functions (forest production and forest diversity, see [Appendix 4: Disturbance simulations]).
+We found similar results for all disturbance levels and other ecosystem functions (forest production and forest diversity, see [Appendix 5: Disturbance simulations]).
 
@@
 We obtained those results for aboveground biomass.
-We found similar results but with an amlified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 4: Disturbance simulations]).
-Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 4: Disturbance simulations]).
+We found similar results but with an amlified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]).
+Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 5: Disturbance simulations]).
 

### 2017-06-09 11-Appendix5.Rmd
Steph1

--- 2017-06-09 10-Appendix4.Rmd
@@
-# Appendix 4: Disturbance simulations
+# Appendix 5: Disturbance simulations
 
-```{r A4.data, echo=FALSE, cache=TRUE}
+```{r A5data, echo=FALSE, cache=TRUE}
 rich <- c(5,25,125)
@@
 
-```{r A4IeqStructureAll, echo=FALSE, fig.cap='Figure S1.1: Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
+```{r A5IeqStructureAll, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
 Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
@@
 
-```{r A4IeqProductionAll, echo=FALSE, fig.cap='Figure S1.2: Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
+```{r A5IeqProductionAll, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
 Cummulative integral from ecosystem distance to forest production equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
@@
 
-```{r A4allBE, echo=FALSE, fig.cap='Resilience of complementarity and selection effects.
+```{r A5allBE, echo=FALSE, fig.cap='Resilience of complementarity and selection effects.
 Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time for different ecosystem variables (AGB, BA, N, GPP).', cache=TRUE, fig.height=8, fig.width=12}

### 2017-06-09 docs/master-thesis.tex
Steph1

@@
 forest diversity, see
-\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
+\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
 Disturbance simulations}).
@@
 with an inverted signal because of the forest self thinning) (see
-\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
+\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
 Disturbance simulations}). Finally, forest growth primary productivity
@@
 its net effect was maintained by complementarity effect (see
-\protect\hyperlink{appendix-4-disturbance-simulations}{Appendix 4:
+\protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
 Disturbance simulations}).
@@
 
-\hypertarget{appendix-4-disturbance-simulations}{\section{Appendix 4:
-Disturbance simulations}\label{appendix-4-disturbance-simulations}}
+\section{Appendix 4: Sensitivity
+analysis}\label{appendix-4-sensitivity-analysis}
+
+\hypertarget{appendix-5-disturbance-simulations}{\section{Appendix 5:
+Disturbance simulations}\label{appendix-5-disturbance-simulations}}
 
@@
 \centering
-\includegraphics{master-thesis_files/figure-latex/A4IeqStructureAll-1.pdf}
-\caption{\label{fig:A4IeqStructureAll}Figure S1.1: Ecosystem resilience
-after 600 years with taxonomic and functional diversity for different
-levels of disturbance. Cummulative integral from ecosystem distance to
-forest dynamic equilibrium after 600 years was represented against
+\includegraphics{master-thesis_files/figure-latex/A5IeqStructureAll-1.pdf}
+\caption{\label{fig:A5IeqStructureAll}Ecosystem resilience after 600 years
+with taxonomic and functional diversity for different levels of
+disturbance. Cummulative integral from ecosystem distance to forest
+dynamic equilibrium after 600 years was represented against functional
+diversity \citep[FRIC, FEve, FDiv, and FDis,][]{villeger_new_2008} for
+different level of disturbance (25, 50 and 75\% of total basal area);
+dot shapes represents the species richness.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A5IeqProductionAll-1.pdf}
+\caption{\label{fig:A5IeqProductionAll}Ecosystem resilience after 600 years
+with taxonomic and functional diversity for different levels of
+disturbance. Cummulative integral from ecosystem distance to forest
+production equilibrium after 600 years was represented against
 functional diversity \citep[FRIC, FEve, FDiv, and
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics{master-thesis_files/figure-latex/A4IeqProductionAll-1.pdf}
-\caption{\label{fig:A4IeqProductionAll}Figure S1.2: Ecosystem resilience
-after 600 years with taxonomic and functional diversity for different
-levels of disturbance. Cummulative integral from ecosystem distance to
-forest production equilibrium after 600 years was represented against
-functional diversity \citep[FRIC, FEve, FDiv, and
-FDis,][]{villeger_new_2008} for different level of disturbance (25, 50
-and 75\% of total basal area); dot shapes represents the species
-richness.}
-\end{figure}
-
 \subsection{Biodiversity effect}\label{biodiversity-effect-2}
@@
 \centering
-\includegraphics{master-thesis_files/figure-latex/A4allBE-1.pdf}
-\caption{\label{fig:A4allBE}Resilience of complementarity and selection
+\includegraphics{master-thesis_files/figure-latex/A5allBE-1.pdf}
+\caption{\label{fig:A5allBE}Resilience of complementarity and selection
 effects. Complementarity effect (CE) and selection effect (SE) where

### 2017-06-12 04-MnMs.Rmd
A4. Sensitivity analysis

@@
   geom_point(size = 3) + 
-  geom_text_repel() +
+  ggrepel::geom_text_repel() +
   scale_color_continuous("Aboveground\nBiomass\n(AGB, ton C/ha)") +

### 2017-06-12 05-Results.Rmd
A4. Sensitivity analysis

@@
 ', cache=TRUE, fig.height=8, fig.width=12}
+library(dplyr)
 g_agb <- ecosystem %>% 
@@
 ```{r tNE, echo=FALSE, cache=TRUE}
+library(reshape2)
+library(knitr)
 BE %>%

### 2017-06-12 10-Appendix4.Rmd
A4. Sensitivity analysis

@@
-# Appendix 4: Disturbance simulations
+# Appendix 4: Sensitivity analysis
 
-```{r A4.data, echo=FALSE, cache=TRUE}
-rich <- c(5,25,125)
-cvh <- 1:20
-dist <- c(0,25,50,75)
-path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
-load(file.path(path, 'disturbanceDOE.Rdata'))
-load(file.path(path, 'disturbanceResilience.Rdata'))
-deq <- data.frame(
-  time = resilience$time,
-  sim = resilience$sim,
-  structure = apply(resilience[c('agb', 'ba', 'n', 'n10', 'n30')], 1, 
-                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))),
-  production = apply(resilience[c('gpp', 'npp', 'Rday', 'Rnight')], 1, 
-                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))))
-sim <- c(sapply(c(sapply(rich, function(r) paste0(r, '.', cvh))), function(x) paste0(x, '.', dist[-1])))
-Ieq <- sapply(sim, function(s){
-  data_s <- subset(deq, sim ==s)
-  return(cbind(data_s[c(1,2)], cumsum(data_s[-c(1,2)])))
-}, simplify = F) ; rm(sim)
-Ieq <- data.frame(do.call('rbind', Ieq))
-names(Ieq)[3:4] <- paste0('Ieq.', names(Ieq)[3:4])
-load(file.path(path, 'disturbanceBE.Rdata'))
-load(file.path(path, 'disturbanceBErecovery.Rdata'))
+To study resistance and resilience of ecosystem face to disturbance, highlighting the role of biodiversity, we decided to use TROLL model simulations [@Chave1999].
+In order to get a finer study of simulations response, we need to assess first sensitivity of the TROLL model to different parameters.
+More particularly we need to assess the importance of functional traits to further better control functional diversities in simulations.
+We also need to assess sensitivity of the model to see rain constant because we assume it is one of the main factor of tree recruitments after disturbance in the model.
+
+## Material and methods
+
+To assess the sensitivity of TROLL model to species functionnal traits, we performed a sensitivity analysis by fixing species trait values to their mean.
+Each trait was tested independently.
+We reduce to a common mean traits with a correlation $r \geq 0.8$ (see figure \@ref(fig:A4corr)).
+
+```{r A4corr, echo=FALSE, fig.cap='Correlation of functional traits within TROLL model species Blue represents negative coorelations whereas red represents positive correlations.
+Values and colour intensity represents correlation values.'}
+species  <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt', header=TRUE, dec=".", sep="")
+corrplot::corrplot.mixed(cor(species[-1]))
 ```
 
-## Ecosystem functions
+To assess the sensitivity of TROLL model to seed rain, we performed a sensitivity analysis by fixing simulations seed rain constant to 2, 20, 200 and 2000 seeds per hectare.
 
-Appendix 4 presents ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-It encompass all functionnal diversity components [FRIC, FEve, FDiv, and FDis, @villeger_new_2008].
-And it presents results for both forest dynamic (Figure \@ref(fig:A4IeqStructureAll)) and forest production (Figure \@ref(fig:A4IeqProductionAll)).
+Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and 188.9 GB of memory.
+We assumed maturity of the forest after 500 years of regeneration [@Li] and computed simulation 100 years after a disturbance event of 40% intensity.
+Due to computer limitations we did not run replicate (besides it should be necessary to reduce simulation stochasticity).
+To assess ecosystem outputs sensitivity to studied parameters, we compared it to 100 replicates of control simulations with all parameters set to default values.
+Ecosystem outputs outside of the range of the control replicates values are significantly influenced by the studied parameter.
 
-```{r A4IeqStructureAll, echo=FALSE, fig.cap='Figure S1.1: Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
-Ieq %>% 
-  left_join(., doe, by = 'sim') %>% 
-  filter(time == 600) %>%
-  select(Ieq.structure, disturbance, richness, FRic, FEve, FDiv, FDis) %>% 
-  melt(id.vars = c('Ieq.structure', 'disturbance', 'richness')) %>% 
-  ggplot(aes(x = Ieq.structure, y = value)) + 
-  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
-  geom_point(aes(shape = richness)) +
-  facet_grid(variable~disturbance, scales = "free", labeller = 'label_both') + 
-  ylab('Functional diversity') +
-  xlab('Cummulative integral from ecosystem 
-       distance to forest dynamic equilibrium after 600 years') +
-  scale_shape_discrete('Species\nrichness\n(SR)')
+## Results
+
+### Control
+
+Both disturbed ecosystem structure and functional composition corresponded to ecosystem structure and functional composition before disturbance (figure \@ref(fig:A4controlStr)).
+Consequently, we assumed that disturbance did not affect much ecosystem structure and function.
+Secondly range of values inside control replicates is low (figure \@ref(fig:A4controlVar)).
+
+```{r A4controlStr, echo=FALSE, fig.cap='Ecosystem structure before disturbance and disturbed.
+Ecosystem structure before disturbance (left) and disturbed (right) with diameter structure (A, B), diversity at different orders (C, D) and rank-abundance diagrams (E, F).', cache=TRUE, warning=FALSE, message=FALSE}
+library(entropart)
+library(RconTroll)
+library(ggplot2)
+load('~/Documents/ECOFOG/Results/sensitivity/MatureControl.Rdata')
+dbhm <- plot(mature, what = 'dbh', ggplot2 = T) +
+  scale_color_discrete(guide = F) +
+  ylab('Number of\nstem above 1 cm\ndiameter at breast height')
+dbhc <- plot(control1, what = 'distdbh', ggplot2 = T) +
+  scale_color_discrete(guide = F) +
+  ylab('Number of\nstem above 1 cm\ndiameter at breast height')
+divm <- plot(mature, what = 'diversity', ggplot2 = T) +
+  scale_color_discrete(guide = F)
+divc <- plot(control1, what = 'distdiversity', ggplot2 = T) +
+  scale_color_discrete(guide = F)
+rankm <- plot(mature, what = 'rankabund', ggplot2 = T) +
+  scale_color_discrete(guide = F) +
+  ylab('Log of\nabundance')
+rankc <- plot(control1, what = 'distrankabund', ggplot2 = T) +
+  scale_color_discrete(guide = F) +
+  ylab('Log of\nabundance')
+cowplot::plot_grid(dbhm, dbhc, divm, divc, rankm, rankc, labels = LETTERS[1:6], nrow = 3, ncol = 2)
+rm(list = ls()[-which(ls() %in% c('control1'))]) ; invisible(gc())
 ```
 
-```{r A4IeqProductionAll, echo=FALSE, fig.cap='Figure S1.2: Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest production equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
-Ieq %>% 
-  left_join(., doe, by = 'sim') %>% 
-  filter(time == 600) %>%
-  select(Ieq.production, disturbance, richness, FRic, FEve, FDiv, FDis) %>% 
-  melt(id.vars = c('Ieq.production', 'disturbance', 'richness')) %>% 
-  ggplot(aes(x = Ieq.production, y = value)) + 
-  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
-  geom_point(aes(shape = richness)) +
-  facet_grid(variable~disturbance, scales = "free", labeller = 'label_both') + 
-  ylab('Functional diversity') +
-  xlab('Cummulative integral from ecosystem 
-       distance to forest production equilibrium after 600 years') +
-  scale_shape_discrete('Species\nrichness\n(SR)')
+```{r A4controlVar, echo=FALSE, fig.cap='Control replicates variation.
+Maximum, mean and minimum number of trees with dbh above 10 cm (N10) and 30 cm (N30), above ground biomass (AGB) and basal area (BA) over simulation time.', cache=TRUE}
+path <- '~/Documents/ECOFOG/Results/sensitivity/control'
+load(file.path(path, 'control.Rdata'))
+n10 <- plot(control, what = 'abu10', ggplot2 = T) +
+  scale_color_discrete(' ') +
+  ylab('Number of stem above\n10 cm diameter\nat breast height')
+n30 <- plot(control, what = 'abu30', ggplot2 = T) +
+  scale_color_discrete(' ') +
+  ylab('Number of stem above\n30 cm diameter\nat breast height')
+agb <- plot(control, what = 'agb', ggplot2 = T) +
+  scale_color_discrete(' ') +
+  ylab('Aboveground biomass (kgC/ha)')
+ba <- plot(control, what = 'ba', ggplot2 = T) +
+  scale_color_discrete(' ') +
+  ylab('Basal area (m2/ha')
+cowplot::plot_grid(n10, n30, agb, ba, labels = c('N10', 'N30', 'AGB', 'BA'), nrow = 2, ncol = 2)
+rm(list = ls()[-which(ls() %in% c('control1', 'control'))]) ; invisible(gc())
 ```
 
-## Biodiversity effect
+### Functional traits
 
-Figure \@ref(fig:A4allBE) presents the resilience of complementarity and selection effects for different ecosystem metrics (AGB, BA, N, GPP and NPP).
+Most of functional traits had a significant long term influence on ecosystem outputs (figure \@ref(fig:A4ftVar)).
+Only **seed volume** was always in the range of variation of control replicates.
+On the other hand, few functional traits influenced final ecosystem structure (figure \@ref(fig:A4ftStr)).
+Only specific maximum diameter **dmax** add higher diversity for greater orders implying better eveness in species distributions.
+Regarding functional composition, traits fixed to mean did not change other functional traits density distribution.
 
-```{r A4allBE, echo=FALSE, fig.cap='Resilience of complementarity and selection effects.
-Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time for different ecosystem variables (AGB, BA, N, GPP).', cache=TRUE, fig.height=8, fig.width=12}
-BEresilience %>% 
-  select(BE, time, sim, agb, ba, n, gpp) %>% 
-  melt(id.vars = c('BE', 'time', 'sim')) %>% 
-  left_join(., doe, by = 'sim') %>%
-  filter(richness == 125) %>%
-  filter(BE %in% c('SE', 'CE')) %>% 
-  group_by(time, BE, disturbance, variable) %>%
-  summarise(qt1 =quantile(value, probs=0.05, na.rm = T),
-            qt2 =quantile(value, probs=0.75, na.rm = T),
-            mean=mean(value, na.rm = T)) %>% 
-  ggplot(aes(x = time)) +
-  geom_hline(yintercept = 1, color = 'red') +
-  geom_hline(yintercept = 0, color = 'blue') +
-  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
-  geom_line(aes(y = mean, color = BE)) +
-  facet_grid(variable ~ disturbance, scales = "free", labeller = 'label_both') +
-  xlab('Time (years)') +
-  ylab('Recovery of variable net effect') + 
-  ylim(-2.5,2.5)
+**ah-hmax** traits fixed to mean increased number of stems above 10 and above 30 cm dbh and basal area after disturbance (but not in long term) and did not affect aboveground biomass.
+Similarly, wood specific gravity **wsg** trait fixed to mean had exactly the same effect on number of stems above 10 and above 30 cm dbh and basal area after disturbance than **ah-hmax** but with a time lag ; and **wsg** also increased aboveground biomass.
+**dmax** trait fixed to mean slightly decreased number of stems above 10 and above 30 cm dbh over time while it increased basal area, aboveground biomass, and species eveness.
+Finally, leaf mass per area **LMA** trait fixed to mean only decreased number of stem above 10 cm dbh after ecosystem resilience to disturbance (approximately 50 years) but did not affect other ecosystem outputs.
+
+```{r A4ftVar, echo=FALSE, fig.cap='Functional traits effect on simulation ecosystem variations over time.
+Number of trees with dbh above 10 cm (N10) and 30 cm (N30), above ground biomass (AGB) and basal area (BA).
+Grey area represents the interval of control replicates whereas black line represents the mean of control replicates.', cache=TRUE, warning=FALSE}
+path <- '~/Documents/ECOFOG/Results/sensitivity/traits'
+load(file.path(path, 'traits.Rdata'))
+n10 <- plot(traits, control, what = 'abu10', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Number of stem above\n10 cm diameter\nat breast height')
+n30 <- plot(traits, control, what = 'abu30', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Number of stem above\n30 cm diameter\nat breast height')
+agb <- plot(traits, control, what = 'agb', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Aboveground biomass (kgC/ha)')
+ba <- plot(traits, control, what = 'ba', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Basal area (m2/ha')
+cowplot::plot_grid(n10, n30, agb, ba, labels = c('N10', 'N30', 'AGB', 'BA'), nrow = 2, ncol = 2)
+rm(list = ls()[-which(ls() %in% c('control1', 'control','traits'))]) ; invisible(gc())
 ```
+
+```{r A4ftStr, echo=FALSE, fig.cap='Functional traits effect on simulation ecosystem final structure.
+Tree final height histogram for traits (A) and control (B), tree final diameter histogram for traits (C) and control (D), ecosystem final diversity plot at different orders (E), and ecosystem final rank-abundance diagram (F).', cache=TRUE, warning=FALSE, message=FALSE}
+height <- plot(traits, what = 'height', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Log of\nstem number')
+heightc <- plot(control1, what = 'height', ggplot2 = T) +
+  scale_color_discrete(guide = F) +
+  ylab('Log of\nstem number')
+dbh <- plot(traits, what = 'dbh', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Number of\nstem above 1 cm\ndiameter at breast height')
+dbhc <- plot(control1, what = 'dbh', ggplot2 = T) +
+  scale_color_discrete(guide = F) +
+  ylab('Number of\nstem above 1 cm\ndiameter at breast height')
+cowplot::plot_grid(height, heightc, dbh, dbhc,
+          labels = c('A', 'B', 'C', 'D'), ncol = 2)
+rm(list = ls()[-which(ls() %in% c('control1', 'control'))]) ; invisible(gc())
+```
+
+### Seed rain
+
+Seedrain constant affected ecosystem outputs only when set lower than default value (figure \@ref(fig:A4srVar) & \@ref(fig:A4srStr)).
+Moreover, seedrain did not seem to affect aboveground biomass (figure \@ref(fig:A4srVar)) and final ecosystem height and diameter structure (figure \@ref(fig:A4srStr)).
+Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a similar effect.
+Lower seedrain implied faster decrease of stem above 10 cm dbh and higher number of stem above 30 cm dbh after ecosystem resilience to disturbance (approximately 50 years).
+Lower seedrain than default decreased basal area over time.
+In addition, lower seedrain than default decreased equitability by increasing abundance of abundant species and decreasing abundance of less abundant species.
+Seedrain constant even decreased the total number of species when fixed to 2 seed per hectare.
+Finally, seedrain constant slightly affected functional composition with higher pike on ecosystem most representatives functional trait values.
+In a nutshell, the lower is the seedrain constant the most the functional density distribution is aggregated around few functional trait values.
+
+```{r A4srVar, echo=FALSE, fig.cap='Seed rain effect on simulation ecosystem variations over time.
+Number of trees with dbh above 10 cm (N10) and 30 cm (N30), above ground biomass (AGB) and basal area (BA).
+Grey area represents the interval of control replicates whereas black line represents the mean of control replicates.', cache=TRUE, warning=FALSE}
+path <- '~/Documents/ECOFOG/Results/sensitivity/seedrain'
+load(file.path(path, 'seedrain.Rdata'))
+n10 <- plot(seedrain, control, what = 'abu10', ggplot2 = T) +
+  scale_color_discrete('Seed rain\n(nb/ha)', 
+                       labels = c(2, 20, 200)) +
+  ylab('Number of stem above\n10 cm diameter\nat breast height')
+n30 <- plot(seedrain, control, what = 'abu30', ggplot2 = T) +
+  scale_color_discrete('Seed rain\n(nb/ha)', 
+                       labels = c(2, 20, 200)) +
+  ylab('Number of stem above\n30 cm diameter\nat breast height')
+agb <- plot(seedrain, control, what = 'agb', ggplot2 = T) +
+  scale_color_discrete('Seed rain\n(nb/ha)', 
+                       labels = c(2, 20, 200)) +
+  ylab('Aboveground biomass (kgC/ha)')
+ba <- plot(seedrain, control, what = 'ba', ggplot2 = T) +
+  scale_color_discrete('Seed rain\n(nb/ha)', 
+                       labels = c(2, 20, 200)) +
+  ylab('Basal area (m2/ha')
+cowplot::plot_grid(n10, n30, agb, ba, labels = c('N10', 'N30', 'AGB', 'BA'), nrow = 2, ncol = 2)
+rm(list = ls()[-which(ls() %in% c('control1', 'control','seedrain'))]) ; invisible(gc())
+```
+
+```{r A4srStr, echo=FALSE, fig.cap='Seed rain effect on simulation ecosystem final structure.
+Tree final height histogram for traits (A) and control (B), tree final diameter histogram for traits (C) and control (D).', cache=TRUE, warning=FALSE, message=FALSE}
+height <- plot(seedrain, what = 'height', ggplot2 = T) +
+  scale_color_discrete('Seed rain\n(nb/ha)', 
+                       labels = c(2, 20, 200)) +
+  ylab('Log of\nstem number')
+heightc <- plot(control1, what = 'height', ggplot2 = T) +
+  scale_color_discrete(guide = F) +
+  ylab('Log of\nstem number')
+dbh <- plot(seedrain, what = 'dbh', ggplot2 = T) +
+  scale_color_discrete('Seed rain\n(nb/ha)', 
+                       labels = c(2, 20, 200)) +
+  ylab('Number of\nstem above 1 cm\ndiameter at breast height')
+dbhc <- plot(control1, what = 'dbh', ggplot2 = T) +
+  scale_color_discrete(guide = F) +
+  ylab('Number of\nstem above 1 cm\ndiameter at breast height')
+cowplot::plot_grid(height, heightc, dbh, dbhc,
+          labels = c('A', 'B', 'C', 'D'), ncol = 2)
+rm(list = ls()) ; invisible(gc())
+```
+
+## Discussion
+
+### Disturbance simulation
+
+Ecosystem structure, species organisation and functional composition stayed the same before and after disturbance.
+We can thus validate disturbance module in its actual state.
+Morevoer we can now consider that the composition we will initialise at the beginning of simulations will stay the same after disturbance.
+Finally control replicates has shown few stochasticity, it advocates for few or no replicates in further analysis.
+
+### Functional traits selection
+
+**ah-hmax** fixed to mean implied no high or low trees.
+Less high trees left space for more trees increasing number of stems in the ecosystem thus increasing basal area.
+Wood specific gravity **wsg** fixed to mean mainly increased wood density of light wood species.
+Globally higher wood density increased lifespan of individuals responsible for the time lag and the higher number of stems increasing basal area.
+**wsg** fiwed to mean also increased carbon capture by individuals, thus increasing aboveground biomass.
+Specific maximum diameter **dmax** fixed to mean decreased death rates.
+Decreased death rate diminished number of stems, expecially big ones thus increasing global basal area and abiveground biomass.
+**dmax** fixed to mean by keeping more small diameter stems also increased random selection of species increasing eveness in species distribution.
+Considering the high correlation between **ah** and **hmax** we could also keep only **hmax** (because of its more straightforward ecological meaning).
+
+### Seed rain constant influence
+
+Seedrain constant did not directly affect ecosystem global outputs over simulation post-disturbance time but have a major effect on species and functional composition and diversity.
+Reducing seedrain constant resulted in an ecosystem selecting few species increasing their abundance and functional dominance of their traits.
+Thus reduced seedrain constant greatly diminsihed eveness untill a decrease of total number of species for its lowest value.

### 2017-06-12 11-Appendix5.Rmd
A4. Sensitivity analysis

@@
 Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
+library(dplyr)
+library(reshape2)
 Ieq %>% 

### 2017-06-12 docs/master-thesis.tex
A4. Sensitivity analysis

@@
 
+To study resistance and resilience of ecosystem face to disturbance,
+highlighting the role of biodiversity, we decided to use TROLL model
+simulations \citep{Chave1999}. In order to get a finer study of
+simulations response, we need to assess first sensitivity of the TROLL
+model to different parameters. More particularly we need to assess the
+importance of functional traits to further better control functional
+diversities in simulations. We also need to assess sensitivity of the
+model to see rain constant because we assume it is one of the main
+factor of tree recruitments after disturbance in the model.
+
+\subsection{Material and methods}\label{material-and-methods-2}
+
+To assess the sensitivity of TROLL model to species functionnal traits,
+we performed a sensitivity analysis by fixing species trait values to
+their mean. Each trait was tested independently. We reduce to a common
+mean traits with a correlation \(r \geq 0.8\) (see figure
+\ref{fig:A4corr}).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4corr-1.pdf}
+\caption{\label{fig:A4corr}Correlation of functional traits within TROLL
+model species Blue represents negative coorelations whereas red
+represents positive correlations. Values and colour intensity represents
+correlation values.}
+\end{figure}
+
+To assess the sensitivity of TROLL model to seed rain, we performed a
+sensitivity analysis by fixing simulations seed rain constant to 2, 20,
+200 and 2000 seeds per hectare.
+
+Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and
+188.9 GB of memory. We assumed maturity of the forest after 500 years of
+regeneration \citep{Li} and computed simulation 100 years after a
+disturbance event of 40\% intensity. Due to computer limitations we did
+not run replicate (besides it should be necessary to reduce simulation
+stochasticity). To assess ecosystem outputs sensitivity to studied
+parameters, we compared it to 100 replicates of control simulations with
+all parameters set to default values. Ecosystem outputs outside of the
+range of the control replicates values are significantly influenced by
+the studied parameter.
+
+\subsection{Results}\label{results-2}
+
+\subsubsection{Control}\label{control}
+
+Both disturbed ecosystem structure and functional composition
+corresponded to ecosystem structure and functional composition before
+disturbance (figure \ref{fig:A4controlStr}). Consequently, we assumed
+that disturbance did not affect much ecosystem structure and function.
+Secondly range of values inside control replicates is low (figure
+\ref{fig:A4controlVar}).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4controlStr-1.pdf}
+\caption{\label{fig:A4controlStr}Ecosystem structure before disturbance and
+disturbed. Ecosystem structure before disturbance (left) and disturbed
+(right) with diameter structure (A, B), diversity at different orders
+(C, D) and rank-abundance diagrams (E, F).}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4controlVar-1.pdf}
+\caption{\label{fig:A4controlVar}Control replicates variation. Maximum, mean
+and minimum number of trees with dbh above 10 cm (N10) and 30 cm (N30),
+above ground biomass (AGB) and basal area (BA) over simulation time.}
+\end{figure}
+
+\subsubsection{Functional traits}\label{functional-traits}
+
+Most of functional traits had a significant long term influence on
+ecosystem outputs (figure \ref{fig:A4ftVar}). Only \textbf{seed volume}
+was always in the range of variation of control replicates. On the other
+hand, few functional traits influenced final ecosystem structure (figure
+\ref{fig:A4ftStr}). Only specific maximum diameter \textbf{dmax} add
+higher diversity for greater orders implying better eveness in species
+distributions. Regarding functional composition, traits fixed to mean
+did not change other functional traits density distribution.
+
+\textbf{ah-hmax} traits fixed to mean increased number of stems above 10
+and above 30 cm dbh and basal area after disturbance (but not in long
+term) and did not affect aboveground biomass. Similarly, wood specific
+gravity \textbf{wsg} trait fixed to mean had exactly the same effect on
+number of stems above 10 and above 30 cm dbh and basal area after
+disturbance than \textbf{ah-hmax} but with a time lag ; and \textbf{wsg}
+also increased aboveground biomass. \textbf{dmax} trait fixed to mean
+slightly decreased number of stems above 10 and above 30 cm dbh over
+time while it increased basal area, aboveground biomass, and species
+eveness. Finally, leaf mass per area \textbf{LMA} trait fixed to mean
+only decreased number of stem above 10 cm dbh after ecosystem resilience
+to disturbance (approximately 50 years) but did not affect other
+ecosystem outputs.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4ftVar-1.pdf}
+\caption{\label{fig:A4ftVar}Functional traits effect on simulation ecosystem
+variations over time. Number of trees with dbh above 10 cm (N10) and 30
+cm (N30), above ground biomass (AGB) and basal area (BA). Grey area
+represents the interval of control replicates whereas black line
+represents the mean of control replicates.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4ftStr-1.pdf}
+\caption{\label{fig:A4ftStr}Functional traits effect on simulation ecosystem
+final structure. Tree final height histogram for traits (A) and control
+(B), tree final diameter histogram for traits (C) and control (D),
+ecosystem final diversity plot at different orders (E), and ecosystem
+final rank-abundance diagram (F).}
+\end{figure}
+
+\subsubsection{Seed rain}\label{seed-rain}
+
+Seedrain constant affected ecosystem outputs only when set lower than
+default value (figure \ref{fig:A4srVar} \& \ref{fig:A4srStr}). Moreover,
+seedrain did not seem to affect aboveground biomass (figure
+\ref{fig:A4srVar}) and final ecosystem height and diameter structure
+(figure \ref{fig:A4srStr}). Seedrain constant fixed to 2 or 20 seed per
+hectare seemed to have a similar effect. Lower seedrain implied faster
+decrease of stem above 10 cm dbh and higher number of stem above 30 cm
+dbh after ecosystem resilience to disturbance (approximately 50 years).
+Lower seedrain than default decreased basal area over time. In addition,
+lower seedrain than default decreased equitability by increasing
+abundance of abundant species and decreasing abundance of less abundant
+species. Seedrain constant even decreased the total number of species
+when fixed to 2 seed per hectare. Finally, seedrain constant slightly
+affected functional composition with higher pike on ecosystem most
+representatives functional trait values. In a nutshell, the lower is the
+seedrain constant the most the functional density distribution is
+aggregated around few functional trait values.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4srVar-1.pdf}
+\caption{\label{fig:A4srVar}Seed rain effect on simulation ecosystem
+variations over time. Number of trees with dbh above 10 cm (N10) and 30
+cm (N30), above ground biomass (AGB) and basal area (BA). Grey area
+represents the interval of control replicates whereas black line
+represents the mean of control replicates.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A4srStr-1.pdf}
+\caption{\label{fig:A4srStr}Seed rain effect on simulation ecosystem final
+structure. Tree final height histogram for traits (A) and control (B),
+tree final diameter histogram for traits (C) and control (D).}
+\end{figure}
+
+\subsection{Discussion}\label{discussion-1}
+
+\subsubsection{Disturbance simulation}\label{disturbance-simulation}
+
+Ecosystem structure, species organisation and functional composition
+stayed the same before and after disturbance. We can thus validate
+disturbance module in its actual state. Morevoer we can now consider
+that the composition we will initialise at the beginning of simulations
+will stay the same after disturbance. Finally control replicates has
+shown few stochasticity, it advocates for few or no replicates in
+further analysis.
+
+\subsubsection{Functional traits
+selection}\label{functional-traits-selection}
+
+\textbf{ah-hmax} fixed to mean implied no high or low trees. Less high
+trees left space for more trees increasing number of stems in the
+ecosystem thus increasing basal area. Wood specific gravity \textbf{wsg}
+fixed to mean mainly increased wood density of light wood species.
+Globally higher wood density increased lifespan of individuals
+responsible for the time lag and the higher number of stems increasing
+basal area. \textbf{wsg} fiwed to mean also increased carbon capture by
+individuals, thus increasing aboveground biomass. Specific maximum
+diameter \textbf{dmax} fixed to mean decreased death rates. Decreased
+death rate diminished number of stems, expecially big ones thus
+increasing global basal area and abiveground biomass. \textbf{dmax}
+fixed to mean by keeping more small diameter stems also increased random
+selection of species increasing eveness in species distribution.
+Considering the high correlation between \textbf{ah} and \textbf{hmax}
+we could also keep only \textbf{hmax} (because of its more
+straightforward ecological meaning).
+
+\subsubsection{Seed rain constant
+influence}\label{seed-rain-constant-influence}
+
+Seedrain constant did not directly affect ecosystem global outputs over
+simulation post-disturbance time but have a major effect on species and
+functional composition and diversity. Reducing seedrain constant
+resulted in an ecosystem selecting few species increasing their
+abundance and functional dominance of their traits. Thus reduced
+seedrain constant greatly diminsihed eveness untill a decrease of total
+number of species for its lowest value.
+
 \hypertarget{appendix-5-disturbance-simulations}{\section{Appendix 5:
@@
 (Figure \ref{fig:A4IeqProductionAll}).
+
+\begin{verbatim}
+## 
+## Attaching package: 'dplyr'
+\end{verbatim}
+
+\begin{verbatim}
+## The following objects are masked from 'package:stats':
+## 
+##     filter, lag
+\end{verbatim}
+
+\begin{verbatim}
+## The following objects are masked from 'package:base':
+## 
+##     intersect, setdiff, setequal, union
+\end{verbatim}
 

### 2017-06-12 05-Results.Rmd
Plot messages & mistakes 1

@@
 Ecosystem outputs (**A**) are normalized by the control value over time to calculate resilience (**B**); resilience of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance ot equilibirum is integrated over time in a cummulative sum (**D**).
-', cache=TRUE, fig.height=8, fig.width=12}
+', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
 library(dplyr)

### 2017-06-12 11-Appendix5.Rmd
Plot messages & mistakes 1

--- 2017-06-09 11-Appendix5.Rmd
@@
 It encompass all functionnal diversity components [FRIC, FEve, FDiv, and FDis, @villeger_new_2008].
-And it presents results for both forest dynamic (Figure \@ref(fig:A4IeqStructureAll)) and forest production (Figure \@ref(fig:A4IeqProductionAll)).
+And it presents results for both forest dynamic (Figure \@ref(fig:A5IeqStructureAll)) and forest production (Figure \@ref(fig:A5IeqProductionAll)).
 
 ```{r A5IeqStructureAll, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
+Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
+library(dplyr)
+library(reshape2)
 Ieq %>% 
@@
 
-Figure \@ref(fig:A4allBE) presents the resilience of complementarity and selection effects for different ecosystem metrics (AGB, BA, N, GPP and NPP).
+Figure \@ref(fig:A5allBE) presents the resilience of complementarity and selection effects for different ecosystem metrics (AGB, BA, N, GPP and NPP).
 

### 2017-06-12 docs/master-thesis.tex
Plot messages & mistakes 1

@@
 
-N\_\{tot\} is the external seedrain per hectare (number of reproduction
+\(N_{tot}\) is the external seedrain per hectare (number of reproduction
 opportunities). \(f_{reg}\) is the species regional frequency.
@@
 and FDis,][]{villeger_new_2008}. And it presents results for both forest
-dynamic (Figure \ref{fig:A4IeqStructureAll}) and forest production
-(Figure \ref{fig:A4IeqProductionAll}).
-
-\begin{verbatim}
-## 
-## Attaching package: 'dplyr'
-\end{verbatim}
-
-\begin{verbatim}
-## The following objects are masked from 'package:stats':
-## 
-##     filter, lag
-\end{verbatim}
-
-\begin{verbatim}
-## The following objects are masked from 'package:base':
-## 
-##     intersect, setdiff, setequal, union
-\end{verbatim}
+dynamic (Figure \ref{fig:A5IeqStructureAll}) and forest production
+(Figure \ref{fig:A5IeqProductionAll}).
 
@@
 
-Figure \ref{fig:A4allBE} presents the resilience of complementarity and
+Figure \ref{fig:A5allBE} presents the resilience of complementarity and
 selection effects for different ecosystem metrics (AGB, BA, N, GPP and

### 2017-06-13 02-Introduction.Rmd
Results: Sensitivity 1

@@
 Currently tropical logging accounts for one eight of global timber production [@Blaser2011] and is still increasing.
-Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cuttint cycle [@Martin2015].
+Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cutting cycle [@Martin2015].
 

### 2017-06-13 05-Results.Rmd
Results: Sensitivity 1

@@
 ## Sensitivity
+
+Most of functional traits had a significant long term influence on ecosystem outputs (figure \@ref(fig:ftVar)).
+On the other hand, few functional traits influenced final ecosystem structure.
+Only specific maximum diameter **dmax** add higher diversity for greater orders implying better eveness in species distributions.
+Regarding functional composition, traits fixed to mean did not change other functional traits density distribution.
+Moreover, seedrain did not seem to affect aboveground biomass and final ecosystem height and diameter structure.
+Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a similar effect.
+Lower seedrain implied faster decrease of stem above 10 cm dbh and higher number of stem above 30 cm dbh after ecosystem resilience to disturbance (approximately 50 years).
+Lower seedrain than default decreased basal area over time.
+In addition, lower seedrain than default decreased equitability by increasing abundance of abundant species and decreasing abundance of less abundant species.
+See [Appendix 4: Sensitivity analysis] for further details.
+
+```{r ftVar, echo=FALSE, fig.cap='Functional traits effect on simulation ecosystem variations over time.
+Number of trees with dbh above 10 cm (N10) and 30 cm (N30), above ground biomass (AGB) and basal area (BA).
+Grey area represents the interval of control replicates whereas black line represents the mean of control replicates.', cache=TRUE, warning=FALSE}
+path <- '~/Documents/ECOFOG/Results/sensitivity/control'
+load(file.path(path, 'control.Rdata'))
+path <- '~/Documents/ECOFOG/Results/sensitivity/traits'
+load(file.path(path, 'traits.Rdata'))
+n10 <- plot(traits, control, what = 'abu10', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Number of stem above\n10 cm diameter\nat breast height')
+n30 <- plot(traits, control, what = 'abu30', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Number of stem above\n30 cm diameter\nat breast height')
+agb <- plot(traits, control, what = 'agb', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Aboveground biomass (kgC/ha)')
+ba <- plot(traits, control, what = 'ba', ggplot2 = T) +
+  scale_color_discrete('Trait fixed\nto mean') +
+  ylab('Basal area (m2/ha')
+cowplot::plot_grid(n10, n30, agb, ba, labels = c('N10', 'N30', 'AGB', 'BA'), nrow = 2, ncol = 2)
+rm(list = ls()) ; invisible(gc())
+```
 

### 2017-06-13 docs/master-thesis.tex
Results: Sensitivity 1

@@
 production originates from selective logging, the targeted harvesting of
-timber from commercial species in a single cuttint cycle
+timber from commercial species in a single cutting cycle
 \citep{Martin2015}.
@@
 
+Most of functional traits had a significant long term influence on
+ecosystem outputs (figure \ref{fig:ftVar}). On the other hand, few
+functional traits influenced final ecosystem structure. Only specific
+maximum diameter \textbf{dmax} add higher diversity for greater orders
+implying better eveness in species distributions. Regarding functional
+composition, traits fixed to mean did not change other functional traits
+density distribution. Moreover, seedrain did not seem to affect
+aboveground biomass and final ecosystem height and diameter structure.
+Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a
+similar effect. Lower seedrain implied faster decrease of stem above 10
+cm dbh and higher number of stem above 30 cm dbh after ecosystem
+resilience to disturbance (approximately 50 years). Lower seedrain than
+default decreased basal area over time. In addition, lower seedrain than
+default decreased equitability by increasing abundance of abundant
+species and decreasing abundance of less abundant species. See
+\protect\hyperlink{appendix-4-sensitivity-analysis}{Appendix 4:
+Sensitivity analysis} for further details.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/ftVar-1.pdf}
+\caption{\label{fig:ftVar}Functional traits effect on simulation ecosystem
+variations over time. Number of trees with dbh above 10 cm (N10) and 30
+cm (N30), above ground biomass (AGB) and basal area (BA). Grey area
+represents the interval of control replicates whereas black line
+represents the mean of control replicates.}
+\end{figure}
+
 \subsection{Disturbance}\label{disturbance-1}
@@
 
-I this Appendix we further detail modules of TROLL model.
+In this Appendix we further detail modules of TROLL model.
 
@@
 
-\section{Appendix 4: Sensitivity
-analysis}\label{appendix-4-sensitivity-analysis}
+\hypertarget{appendix-4-sensitivity-analysis}{\section{Appendix 4:
+Sensitivity analysis}\label{appendix-4-sensitivity-analysis}}
 

### 2017-06-14 05-Results.Rmd
Discussion draft 1

@@
 Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a similar effect.
-Lower seedrain implied faster decrease of stem above 10 cm dbh and higher number of stem above 30 cm dbh after ecosystem resilience to disturbance (approximately 50 years).
+Lower seedrain implied faster decrease of stem above 10 cm dbh and higher number of stem above 30 cm diameter at breast height after ecosystem resilience to disturbance (approximately 50 years).
 Lower seedrain than default decreased basal area over time.

### 2017-06-14 06-Discussion.Rmd
Discussion draft 1

# Discussion

**Take home messages:**

## "Model limits"

Discuss limits of the model with sensitivity analysis (insist in seed rain constant importance).

## Diversity and tropical forest ecosystem resilience

* Diversity have a strong role in ecosystem resilience
* Increasing functional diversity results in increased resilience
* Species and functional richnesses is not directly increasing resilience
* But increased species and functional richnesses will increase chance for high functional diversity
* Functional dispersion decreased resilience when functional eveness increased resilience
* Effectively high functional diversity needs eveness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species

## Biodiversity role in tropical forest ecosystem resilience

* Biodiversity net effect is almost always positive for ecosystems functions
* Biodiversity effects answer to disturbance is split in two parts
* First complementarity effects answer disturbance in first decades whereas the disturbance almost removed selection effect
* Then complementarity effect will diminish toward extinction when selection effect is *prendre le relais* to ensure biodiversity net effect
* This patterns are even stronger for higher disturbances
* In the case of selective logging the complementarity effect will thus be the only one playing between two *rotations*
* High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest restoration between two rotation in order to maintain both productive ecosystem and sustainable management

## Sylviculture ?

## Selective logging vs. monoculture stand ?

Results advocates for **sustainable** selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging *turn-overs*. Even if monoculture stand are *irrégularisé* they will not reach mature forest diversity, and consequently they will not reach its natural high resilience. But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each turn-overs degrading forest ecosystem functions, and resilience will become slower and slower in a *cercle vicieux*. Return toward @Zimmerman2012  discussion.

### 2017-06-14 docs/master-thesis.tex
Discussion draft 1

@@
 similar effect. Lower seedrain implied faster decrease of stem above 10
-cm dbh and higher number of stem above 30 cm dbh after ecosystem
-resilience to disturbance (approximately 50 years). Lower seedrain than
-default decreased basal area over time. In addition, lower seedrain than
-default decreased equitability by increasing abundance of abundant
-species and decreasing abundance of less abundant species. See
+cm dbh and higher number of stem above 30 cm diameter at breast height
+after ecosystem resilience to disturbance (approximately 50 years).
+Lower seedrain than default decreased basal area over time. In addition,
+lower seedrain than default decreased equitability by increasing
+abundance of abundant species and decreasing abundance of less abundant
+species. See
 \protect\hyperlink{appendix-4-sensitivity-analysis}{Appendix 4:
@@
 
+\textbf{Take home messages:}
+
+\subsection{\texorpdfstring{``Model
+limits''}{Model limits}}\label{model-limits}
+
+Discuss limits of the model with sensitivity analysis (insist in seed
+rain constant importance).
+
+\subsection{Diversity and tropical forest ecosystem
+resilience}\label{diversity-and-tropical-forest-ecosystem-resilience}
+
+\begin{itemize}
+\tightlist
+\item
+  Diversity have a strong role in ecosystem resilience
+\item
+  Increasing functional diversity results in increased resilience
+\item
+  Species and functional richnesses is not directly increasing
+  resilience
+\item
+  But increased species and functional richnesses will increase chance
+  for high functional diversity
+\item
+  Functional dispersion decreased resilience when functional eveness
+  increased resilience
+\item
+  Effectively high functional diversity needs eveness in order to better
+  answer disturbance, if not the diversity is masked by ecosystem
+  dominant species
+\end{itemize}
+
+\subsection{Biodiversity role in tropical forest ecosystem
+resilience}\label{biodiversity-role-in-tropical-forest-ecosystem-resilience}
+
+\begin{itemize}
+\tightlist
+\item
+  Biodiversity net effect is almost always positive for ecosystems
+  functions
+\item
+  Biodiversity effects answer to disturbance is split in two parts
+\item
+  First complementarity effects answer disturbance in first decades
+  whereas the disturbance almost removed selection effect
+\item
+  Then complementarity effect will diminish toward extinction when
+  selection effect is \emph{prendre le relais} to ensure biodiversity
+  net effect
+\item
+  This patterns are even stronger for higher disturbances
+\item
+  In the case of selective logging the complementarity effect will thus
+  be the only one playing between two \emph{rotations}
+\item
+  High forest diversity with important species complementarity through
+  increased functional diversity is thus an advantage for forest
+  restoration between two rotation in order to maintain both productive
+  ecosystem and sustainable management
+\end{itemize}
+
+\subsection{Sylviculture ?}\label{sylviculture-2}
+
+\subsection{Selective logging vs.~monoculture stand
+?}\label{selective-logging-vs.monoculture-stand}
+
+Results advocates for \textbf{sustainable} selective logging against
+monoculture stand: high diversity will increase forest resilience and
+thus improve logging \emph{turn-overs}. Even if monoculture stand are
+\emph{irrégularisé} they will not reach mature forest diversity, and
+consequently they will not reach its natural high resilience. But on the
+other hand selective logging in tropical forest needs to be sustainable,
+if not the diversity will slowly decrease after each turn-overs
+degrading forest ecosystem functions, and resilience will become slower
+and slower in a \emph{cercle vicieux}. Return toward
+\citet{Zimmerman2012} discussion.
+
 \appendix

### 2017-07-12 02-Introduction.Rmd
Introduction v2

@@
 # Introduction {-}
+
+Tropical forests perturbations, through deforestation and degradation, account for 8.26 billion of carbon dioxyde emissions per year [@Pearson2017].
+It represents the third source of greenhouse gas.
+Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics.
+Degradation from forest has been estimated to represents 10 time deforestation [@Herold2011] and represents 20% of greenhouse gas emissions in the brazilian Amazon [@Asner2005].
 
@@
 Amongst others, tropical forests play a key role in biogeochemical cycles, including carbone storage [@Lewis2004].
-**Add insights into carbon storage role of tropical forest.** Ecosystem functions from tropical forests support numerous ecosystem services, such as timber production and climate regulation.
+Ecosystem functions from tropical forests support numerous ecosystem services, such as timber production and climate regulation.
 
@@
 Consequently assessing selective logging effect on tropical forest linking biodiversity and ecosystem seems an obvious and promising way [@Loreau2010].
-**Paragraph to fully review !**
 
@@
 
-**Major question greater diversity (taxonomic and functional) brought a better resilience to disturbance ?**
+We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance highlighting tropical forest carbone resilience to disturbance.
+Resilience encompass several definitions but was summarized by @Oliver2015 as the degree of resistance or fast recovery from an ecosystem function to environmental disturbance.
+
+Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity.
+We assumed that when a disturbance event happen, due to a higher productivity, foerst with an increased diversity with recover quickly and thus be more resilient.
+This theoritical expectation stand on two processes: complementarity and diversity effects [@Loreau2001].
+If we take a species pools with different productivity in monoculture.
+Their assemblage will allow theim an overall higher productivity due to a better ressources acquisition through ressource partitionning and niche differentiation.
+Additionally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage.
+Complementarity effect is the addition of the better ressource acquisition and the facilitation.
+Now if we look at the evolution of the species assemblage over time, more competitive species will progressively dominate through competitive selection.
+And if competitive species are more productive, they will increase assemblage overall productivity.
+This ressources preemption by more competitive species is the selection effect.
+
+Generally, positive relationship between biodiversity and productivity were emprically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of tropical forests.
+One of the few studies was realized by @Chisholm2013 and has shwon a positive significative relationship between species richness and wood productivity on a worlwide forest network.
+But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.
+
+In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience, more particularly tropical forest carbon resilience.
+We used a simulation approach using TROLL model to assess long term processes.

### 2017-07-12 03-Model.Rmd
Introduction v2

@@
 
-Disturbance module was designed in the simplest way in order to relate the ecosystem answer to volume lost without any individuals nor species targetting.
+Disturbance module was designed in the simplest way in order to relate the ecosystem answer to volume loss without any individuals nor species targetting.
 For a given iteration $disturb_{iter}$, individuals are picked randomly with a uniform law on the number of trees.

### 2017-07-12 05-Results.Rmd
Introduction v2

@@
 We transformed all ecosystem outputs from the 240 disturbance simulations in resilience metrics normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A et B).
-We then gathered ecosystem outputs by main ecosystem functions (forest dynamic, forest production and forest diversity) to compute ecosystem distance to equilibrium (Figure \@ref(fig:datatrans) C).
+We then gathered ecosystem outputs by main ecosystem functions (forest dynamic and forest production) to compute ecosystem distance to equilibrium (Figure \@ref(fig:datatrans) C).
 Finally, we integrated distance to equilibrium in a cummulative sum over time (Figure \@ref(fig:datatrans) D).
@@
 We normalized treatments effect by control net effect (see Table \@ref(tab:tNE)) to measure their aboveground biomass resilience to disturbance (see Figure \@ref(fig:gBE)).
-We found that complementarity effect was recovering net effect in the first decades untill it disappeared after a century.
+We found that complementarity effect was recovering net effect in the first decades untill it diminished after a century.
 On the contrary selection effect was reduced or even removed by the disturbance.
 It increased during the whole simulation and was greater than complementarity effect only after decades.
-The time lag for which complementarity effect was non null and greater than selection effect was increasing with disturbance intensity.
+The time lag for which complementarity effect was greater than selection effect was increasing with disturbance intensity.
 Finally 600 years after the disturbance event biodiversity net effect was still not recovered for a disturbance intensity greater than 25% of basal area.
@@
 ```
-
-## Sylviculture

### 2017-07-12 docs/master-thesis.tex
Introduction v2

@@
 
+% citations
+\usepackage{epigraph}
+
 % set default figure placement to htbp
@@
     \large{Sujet :} \\
-    \Large{\textbf{A COMPLETER}} \\
+    \Large{\textbf{Résilience des écosystèmes forestiers tropicaux après perturbation}} \\
     \vspace*{\fill}
-    \large{soutenu publiquement le XX.xxxx.201X} \\
+    \large{soutenu publiquement le XX Juin 2017} \\
     \large{à Kourou} \\
@@
   \vspace*{\fill}
-  \emph{Les opinions émises par les auteurs sont personnelles et n'engagent pas AgroParisTech.}
+  \epigraph{Le simple est toujours faux. Ce qui ne l'est pas est inutilisable.}{\textit{Paul Valéry}}
+  \vspace*{\fill}
+  \emph{Les opinions émises par les auteurs sont personnelles et n'engagent pas AgroParisTech.}\\
+  \emph{Une version web de ce document est diponible à l'adresse suivante : \url{https://sylvainschmitt.github.io/master-thesis/}.}
   \newpage
@@
 \addcontentsline{toc}{section}{Introduction}
+
+Tropical forests perturbations, through deforestation and degradation,
+account for 8.26 billion of carbon dioxyde emissions per year
+\citep{Pearson2017}. It represents the third source of greenhouse gas.
+Besides deforestation has been rightly the focus of worldwide attention,
+degradation has been less studied and quantified through tropics.
+Degradation from forest has been estimated to represents 10 time
+deforestation \citep{Herold2011} and represents 20\% of greenhouse gas
+emissions in the brazilian Amazon \citep{Asner2005}.
 
@@
 play a key role in biogeochemical cycles, including carbone storage
-\citep{Lewis2004}. \textbf{Add insights into carbon storage role of
-tropical forest.} Ecosystem functions from tropical forests support
+\citep{Lewis2004}. Ecosystem functions from tropical forests support
 numerous ecosystem services, such as timber production and climate
@@
 tropical forest linking biodiversity and ecosystem seems an obvious and
-promising way \citep{Loreau2010}. \textbf{Paragraph to fully review !}
+promising way \citep{Loreau2010}.
 
@@
 
-\textbf{Major question greater diversity (taxonomic and functional)
-brought a better resilience to disturbance ?}
+We decided to use the forest model TROLL to study the role of
+biodiversity in forest ecosystem answer to disturbance highlighting
+tropical forest carbone resilience to disturbance. Resilience encompass
+several definitions but was summarized by \citet{Oliver2015} as the
+degree of resistance or fast recovery from an ecosystem function to
+environmental disturbance.
+
+Our work is based on the general hypothesis of a positive relationships
+between biodiversity and productivity. We assumed that when a
+disturbance event happen, due to a higher productivity, foerst with an
+increased diversity with recover quickly and thus be more resilient.
+This theoritical expectation stand on two processes: complementarity and
+diversity effects \citep{Loreau2001}. If we take a species pools with
+different productivity in monoculture. Their assemblage will allow theim
+an overall higher productivity due to a better ressources acquisition
+through ressource partitionning and niche differentiation. Additionally,
+some species will individually have an higher productivity in the
+assemblage than in monoculture due to facilitation from other species
+present in the assemblage. Complementarity effect is the addition of the
+better ressource acquisition and the facilitation. Now if we look at the
+evolution of the species assemblage over time, more competitive species
+will progressively dominate through competitive selection. And if
+competitive species are more productive, they will increase assemblage
+overall productivity. This ressources preemption by more competitive
+species is the selection effect.
+
+Generally, positive relationship between biodiversity and productivity
+were emprically and experimentally demonstrated on grassland systems
+\citep{Hooper2005a, Loreau2001a, Naeem2002}; but few studies focused on
+the case of tropical forests. One of the few studies was realized by
+\citet{Chisholm2013} and has shwon a positive significative relationship
+between species richness and wood productivity on a worlwide forest
+network. But those study still presents three major limits: (1) study
+time are inferior to a tree life time, (2) experimental network include
+scarcely disturbed plot, and (3) correlative approach does not explain
+mechanisms involved in the relationship.
+
+In the present study, we focused on mechanisms involved in the
+relationship between biodiversity and forest ecosystem resilience, more
+particularly tropical forest carbon resilience. We used a simulation
+approach using TROLL model to assess long term processes.
 
@@
 Disturbance module was designed in the simplest way in order to relate
-the ecosystem answer to volume lost without any individuals nor species
+the ecosystem answer to volume loss without any individuals nor species
 targetting. For a given iteration \(disturb_{iter}\), individuals are
@@
 Disturbance was represented by percentage of basal area loss (0\%, 25\%,
-50\% and 75\%), or as a selective logging simulation. Biodiversity was
-integrated with two components of its components: taxonomic and
-functional diversities. We used species richness \(SR\) to represents
-taxonomic diversity (5, 25, and 125 species). Functional diversity can
-be related to numerous components, and \citet{Borgy2017} argued for 5:
-richness, divergence, regularity, overlap and mean. Because mature
-forest were created from a bare soil with TROLL simulations, we could
-not control a priori divergence, regularity and overlap but only assess
-theim before diturbance. Consequently, we focused on functional richness
-with convex hull volume \(CHV\) and functional mean with community
-weighted mean \(CWM\). For each level of species richness \(SR\), we
-selected 20 communities with growing convex hull volume \(CHV\) but with
-a community weighted means close to the regional species pool community
-weighted means. Effectivelly, we did not wanted drastic change in
-community means that could have more effect than functional richness
-itself. This design of experiments resulted in 60 communities
-(\(5~SR*20~CHV\)) and 240 simulations
-(\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
-being assumed after 500 years of regeneration \citep{Li}). Figure
-\ref{fig:DOE} presents the design of experiment for communities
+50\% and 75\%), or as a selective logging simulation using default
+parameters. Biodiversity was integrated with two components of its
+components: taxonomic and functional diversities. We used species
+richness \(SR\) to represents taxonomic diversity (5, 25, and 125
+species). Functional diversity can be related to numerous components,
+and \citet{Borgy2017} argued for 5: richness, divergence, regularity,
+overlap and mean. Because mature forest were created from a bare soil
+with TROLL simulations, we could not control a priori divergence,
+regularity and overlap but only assess theim before diturbance.
+Consequently, we focused on functional richness with convex hull volume
+\(CHV\) and functional mean with community weighted mean \(CWM\). For
+each level of species richness \(SR\), we selected 20 communities with
+growing convex hull volume \(CHV\) but with a community weighted means
+close to the regional species pool community weighted means.
+Effectivelly, we did not wanted drastic change in community means that
+could have more effect than functional richness itself. This design of
+experiments resulted in 60 communities (\(5~SR*20~CHV\)) and 240
+simulations (\(60 ~communities*4~levels~of~disturbance\)) over 600 years
+(maturity being assumed after 500 years of regeneration \citep{Li}).
+Figure \ref{fig:DOE} presents the design of experiment for communities
 biodiversity after the mature forest were simulated, and thus before
@@
 their corresponding control (Figure \ref{fig:datatrans} A et B). We then
-gathered ecosystem outputs by main ecosystem functions (forest dynamic,
-forest production and forest diversity) to compute ecosystem distance to
-equilibrium (Figure \ref{fig:datatrans} C). Finally, we integrated
-distance to equilibrium in a cummulative sum over time (Figure
-\ref{fig:datatrans} D).
+gathered ecosystem outputs by main ecosystem functions (forest dynamic
+and forest production) to compute ecosystem distance to equilibrium
+(Figure \ref{fig:datatrans} C). Finally, we integrated distance to
+equilibrium in a cummulative sum over time (Figure \ref{fig:datatrans}
+D).
 
@@
 \ref{fig:gBE}). We found that complementarity effect was recovering net
-effect in the first decades untill it disappeared after a century. On
-the contrary selection effect was reduced or even removed by the
+effect in the first decades untill it diminished after a century. On the
+contrary selection effect was reduced or even removed by the
 disturbance. It increased during the whole simulation and was greater
 than complementarity effect only after decades. The time lag for which
-complementarity effect was non null and greater than selection effect
-was increasing with disturbance intensity. Finally 600 years after the
-disturbance event biodiversity net effect was still not recovered for a
-disturbance intensity greater than 25\% of basal area. We obtained those
-results for aboveground biomass. We found similar results but with an
-amlified signal for basal area (\(BA\)) and stem abundance (\(N\) but
-with an inverted signal because of the forest self thinning) (see
+complementarity effect was greater than selection effect was increasing
+with disturbance intensity. Finally 600 years after the disturbance
+event biodiversity net effect was still not recovered for a disturbance
+intensity greater than 25\% of basal area. We obtained those results for
+aboveground biomass. We found similar results but with an amlified
+signal for basal area (\(BA\)) and stem abundance (\(N\) but with an
+inverted signal because of the forest self thinning) (see
 \protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
@@
 
-\subsection{Sylviculture}\label{sylviculture-1}
-
 \section{Discussion}\label{discussion}
@@
 
-\subsection{Sylviculture ?}\label{sylviculture-2}
+\subsection{Sylviculture ?}\label{sylviculture-1}
 

### 2017-07-12 latex/before_body.tex
Introduction v2

@@
   \large{Sujet :} \\
-  \Large{\textbf{A COMPLETER}} \\
+  \Large{\textbf{Résilience des écosystèmes forestiers tropicaux après perturbation}} \\
   \vspace*{\fill}
-  \large{soutenu publiquement le XX.xxxx.201X} \\
+  \large{soutenu publiquement le XX Juin 2017} \\
   \large{à Kourou} \\
@@
 \vspace*{\fill}
-\emph{Les opinions émises par les auteurs sont personnelles et n'engagent pas AgroParisTech.}
+\epigraph{Le simple est toujours faux. Ce qui ne l'est pas est inutilisable.}{\textit{Paul Valéry}}
+\vspace*{\fill}
+\emph{Les opinions émises par les auteurs sont personnelles et n'engagent pas AgroParisTech.}\\
+\emph{Une version web de ce document est diponible à l'adresse suivante : \url{https://sylvainschmitt.github.io/master-thesis/}.}
 \newpage

### 2017-07-13 06-Discussion.Rmd
Discussion v1

@@
 
-**Take home messages:**
+In the limit of the model, we were able to show that diversity improved tropical forest resilience.
+More particularly, functional diversity and evenness are key components of diversity in forest recovery after disturbance.
+Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery.
+Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity.
+But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012.
+Because if the harvesting are not sustainable negative feedbaks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.
 
-## "Model limits"
+## TROLL limits
 
-Discuss limits of the model with sensitivity analysis (insist in seed rain constant importance).
+Belowground processes, herbaceaous plants, epyphytes and lianas are not simulated in TROLL but they not reprensent the only limit of the model.
+Other processes are simulated but simplified, and we used sensitivity analysis to assess their relative importance.
+We found that few functional traits were influencing whole forest structure and dynamic.
+In addition, we found that the seed rain constant had an important effect on species functional composition and diversity.
+High external seed rain resulted in a quick recovery of the system toward an equilibrium close to the regional species frequency levied by the seed rain.
+On the contrary, low seed rain let the simulated forest works has a closed system with more system feedback but a lower stability through time with longer species diversity transitions.
+In order to study the role of diversity in forest resilience we decided to remove the seed rain constant to get a closed system and look at the role of diversity when it maintains itself through feedbacks and not with immigration.
 
-## Diversity and tropical forest ecosystem resilience
+## Diversity improve tropical forest resilience
 
-* Diversity have a strong role in ecosystem resilience
-* Increasing functional diversity results in increased resilience
-* Species and functional richnesses is not directly increasing resilience
-* But increased species and functional richnesses will increase chance for high functional diversity
-* Functional dispersion decreased resilience when functional eveness increased resilience
-* Effectively high functional diversity needs eveness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species
+Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and eveness ($p < 0.01$ and $p < 0.05$ respectivelly).
+Thus we were able to show that diversity improve forest resilience.
+More particularly, functional diversity seems a major aspect of resilience if its enforced by a high functional eveness.
+Effectively high functional diversity needs eveness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species.
+Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes.
+Additionally, the role of eveness  confirm the review of @Zhang2012 who highlighted the role of eveness in productivity and thus resilience following our hypothesis.
+Finally, species and functional richnesses is not directly increasing resilience.
+But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998].
+An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.
 
-## Biodiversity role in tropical forest ecosystem resilience
+## Complementarity and selection insure forest resilience
 
-* Biodiversity net effect is almost always positive for ecosystems functions
-* Biodiversity effects answer to disturbance is split in two parts
-* First complementarity effects answer disturbance in first decades whereas the disturbance almost removed selection effect
-* Then complementarity effect will diminish toward extinction when selection effect is *prendre le relais* to ensure biodiversity net effect
-* This patterns are even stronger for higher disturbances
-* In the case of selective logging the complementarity effect will thus be the only one playing between two *rotations*
-* High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest restoration between two rotation in order to maintain both productive ecosystem and sustainable management
+We found that complementary effect was insuring forest resilience in the beginning of forest successions.
+We interpreted this results as the consequence of facilitation processes.
+As the only ressource simulated by TROLL is the light, the major facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps.
+But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species.
+The diminishing of complementarity effect is due to competitive selection though time in gaps succession.
+But the study scale matters as shwon by @Chisholm2013.
+Here we look at processes to a 16 ha scale.
+We do not have topography reducing micro-environment effects, but @Chisholm2013 results suggest that complementarity will be stronger at smaller scale and could explain its low value after forest recovery.
+Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada.
+They also found that selection effect was greater than complementarity in most of the case.
+In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield).
+High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two rotation in order to maintain both productive ecosystem and sustainable management.
 
-## Sylviculture ?
+## Conclusion
 
-## Selective logging vs. monoculture stand ?
-
-Results advocates for **sustainable** selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging *turn-overs*.
-Even if monoculture stand are *irrégularisé* they will not reach mature forest diversity, and consequently they will not reach its natural high resilience.
-But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each turn-overs degrading forest ecosystem functions, and resilience will become slower and slower in a *cercle vicieux*.
-Return toward @Zimmerman2012  discussion.
+We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance.
+We found that diversity improved forest resilience and that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery.
+Our results advocates for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles.
+Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience.
+But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedbacks.
+@Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces".
+Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy.
+This view correspond to the high resilience, we were able to show, of tropical forest hyperdiverse systems and let hope for a future sustainable selective logging in tropical forests.

### 2017-07-13 docs/master-thesis.tex
Discussion v1

@@
     \large{Sujet :} \\
-    \Large{\textbf{Résilience des écosystèmes forestiers tropicaux après perturbation}} \\
+    \Large{\textbf{Rôle de la biodiversité dans la résilience des écosystèmes forestiers tropicaux après perturbation}} \\
     \vspace*{\fill}
@@
 
-\textbf{Take home messages:}
-
-\subsection{\texorpdfstring{``Model
-limits''}{Model limits}}\label{model-limits}
-
-Discuss limits of the model with sensitivity analysis (insist in seed
-rain constant importance).
-
-\subsection{Diversity and tropical forest ecosystem
-resilience}\label{diversity-and-tropical-forest-ecosystem-resilience}
-
-\begin{itemize}
-\tightlist
-\item
-  Diversity have a strong role in ecosystem resilience
-\item
-  Increasing functional diversity results in increased resilience
-\item
-  Species and functional richnesses is not directly increasing
-  resilience
-\item
-  But increased species and functional richnesses will increase chance
-  for high functional diversity
-\item
-  Functional dispersion decreased resilience when functional eveness
-  increased resilience
-\item
-  Effectively high functional diversity needs eveness in order to better
-  answer disturbance, if not the diversity is masked by ecosystem
-  dominant species
-\end{itemize}
-
-\subsection{Biodiversity role in tropical forest ecosystem
-resilience}\label{biodiversity-role-in-tropical-forest-ecosystem-resilience}
-
-\begin{itemize}
-\tightlist
-\item
-  Biodiversity net effect is almost always positive for ecosystems
-  functions
-\item
-  Biodiversity effects answer to disturbance is split in two parts
-\item
-  First complementarity effects answer disturbance in first decades
-  whereas the disturbance almost removed selection effect
-\item
-  Then complementarity effect will diminish toward extinction when
-  selection effect is \emph{prendre le relais} to ensure biodiversity
-  net effect
-\item
-  This patterns are even stronger for higher disturbances
-\item
-  In the case of selective logging the complementarity effect will thus
-  be the only one playing between two \emph{rotations}
-\item
-  High forest diversity with important species complementarity through
-  increased functional diversity is thus an advantage for forest
-  restoration between two rotation in order to maintain both productive
-  ecosystem and sustainable management
-\end{itemize}
-
-\subsection{Sylviculture ?}\label{sylviculture-1}
-
-\subsection{Selective logging vs.~monoculture stand
-?}\label{selective-logging-vs.monoculture-stand}
-
-Results advocates for \textbf{sustainable} selective logging against
-monoculture stand: high diversity will increase forest resilience and
-thus improve logging \emph{turn-overs}. Even if monoculture stand are
-\emph{irrégularisé} they will not reach mature forest diversity, and
-consequently they will not reach its natural high resilience. But on the
-other hand selective logging in tropical forest needs to be sustainable,
-if not the diversity will slowly decrease after each turn-overs
-degrading forest ecosystem functions, and resilience will become slower
-and slower in a \emph{cercle vicieux}. Return toward
-\citet{Zimmerman2012} discussion.
+In the limit of the model, we were able to show that diversity improved
+tropical forest resilience. More particularly, functional diversity and
+evenness are key components of diversity in forest recovery after
+disturbance. Moreover, we found that complementarity between species was
+insuring forest recovery in forest succession start with facilitation
+before more productive species dominate the forest and insure recovery.
+Our results advocates for a sustainable harvesting of tropical forests
+through an increased resilience due to high diversity. But this
+conclusion should met a sustainable definiton of selective logging
+following \citet{Zimmerman2012}. Because if the harvesting are not
+sustainable negative feedbaks will slowly diminish diversity and its
+benefits for forest resilience, resulting in forest degradation.
+
+\subsection{TROLL limits}\label{troll-limits}
+
+Belowground processes, herbaceaous plants, epyphytes and lianas are not
+simulated in TROLL but they not reprensent the only limit of the model.
+Other processes are simulated but simplified, and we used sensitivity
+analysis to assess their relative importance. We found that few
+functional traits were influencing whole forest structure and dynamic.
+In addition, we found that the seed rain constant had an important
+effect on species functional composition and diversity. High external
+seed rain resulted in a quick recovery of the system toward an
+equilibrium close to the regional species frequency levied by the seed
+rain. On the contrary, low seed rain let the simulated forest works has
+a closed system with more system feedback but a lower stability through
+time with longer species diversity transitions. In order to study the
+role of diversity in forest resilience we decided to remove the seed
+rain constant to get a closed system and look at the role of diversity
+when it maintains itself through feedbacks and not with immigration.
+
+\subsection{Diversity improve tropical forest
+resilience}\label{diversity-improve-tropical-forest-resilience}
+
+Our results validated the hypothesis of a significative relationship
+between forest resilience and functional diversity and eveness
+(\(p < 0.01\) and \(p < 0.05\) respectivelly). Thus we were able to show
+that diversity improve forest resilience. More particularly, functional
+diversity seems a major aspect of resilience if its enforced by a high
+functional eveness. Effectively high functional diversity needs eveness
+in order to better answer disturbance, if not the diversity is masked by
+ecosystem dominant species. Those results confirms the review of
+\citet{Diaz2001} advocating for underevaluated importance of plant
+functional diversity in ecosystem processes. Additionally, the role of
+eveness confirm the review of \citet{Zhang2012} who highlighted the role
+of eveness in productivity and thus resilience following our hypothesis.
+Finally, species and functional richnesses is not directly increasing
+resilience. But increased species and functional richnesses will
+increase chance for high functional diversity through the sampling
+effect \citep{Loreau1998}. An higher sampling of regional species pool
+will allow a greater chance to pick more functionally diverse species.
+
+\subsection{Complementarity and selection insure forest
+resilience}\label{complementarity-and-selection-insure-forest-resilience}
+
+We found that complementary effect was insuring forest resilience in the
+beginning of forest successions. We interpreted this results as the
+consequence of facilitation processes. As the only ressource simulated
+by TROLL is the light, the major facilitation will be light shading of
+post-pioneer species by pioneer species in disturbance gaps. But the
+complementarity effect is reducing through time, to let the selection
+effect insure forest resilience due to an inforced dominance of more
+productive species. The diminishing of complementarity effect is due to
+competitive selection though time in gaps succession. But the study
+scale matters as shwon by \citet{Chisholm2013}. Here we look at
+processes to a 16 ha scale. We do not have topography reducing
+micro-environment effects, but \citet{Chisholm2013} results suggest that
+complementarity will be stronger at smaller scale and could explain its
+low value after forest recovery. Finally, our findings confirm results
+of \citet{Tobner2016} realized on experimental forests of 4 years in low
+diverse forest of Canada. They also found that selection effect was
+greater than complementarity in most of the case. In the case of
+selective logging the complementarity effect will thus be the major
+effect between two cutting cycles due to the cycle length (several
+decades in guyana shield). High forest diversity with important species
+complementarity through increased functional diversity is thus an
+advantage for forest recovery between two rotation in order to maintain
+both productive ecosystem and sustainable management.
+
+\subsection{Conclusion}\label{conclusion}
+
+We used closed forest system simulated with TROLL forest model to
+evaluate the role and mechanisms of biodiversity in tropical forest
+resilience to disturbance. We found that diversity improved forest
+resilience and that complementarity between species was insuring forest
+recovery in forest succession start with facilitation before more
+productive species dominate the forest and insure recovery. Our results
+advocates for sustainable selective logging against monoculture stand:
+high diversity will increase forest resilience and thus improve logging
+cycles. Even if monoculture stand are naturalized they will not reach
+mature forest diversity, and consequently they will not reach its
+natural high resilience. But on the other hand selective logging in
+tropical forest needs to be sustainable, if not the diversity will
+slowly decrease after each cycles degrading forest ecosystem functions,
+and resilience will decrease due to negative feedbacks.
+\citet{Zimmerman2012} criticized the state of sylviculture in tropics,
+suggesting that ``we have not been able to reconcile these opposing
+biological and economic forces''. Still, they advocates for a
+possibility of sustainable tropical logging, already existing in
+small-scale, that will need proper funding from the internationnal
+communtiy. This view correspond to the high resilience, we were able to
+show, of tropical forest hyperdiverse systems and let hope for a future
+sustainable selective logging in tropical forests.
 

### 2017-07-13 02-Introduction.Rmd
Thesis v1

@@
 But several authors argue that selective logging represents a major threat to biodiversity [@Carreno-Rocabado2012; @DeAvila2015; @Gibson2013; @Martin2015; @Zimmerman2012], challenging the sustainable definition from current selective logging.
-We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better syslvicultural practives in order to reach sustainability.
+We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better sylvicultural practices in order to reach sustainability.
 
@@
 Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity.
-We assumed that when a disturbance event happen, due to a higher productivity, foerst with an increased diversity with recover quickly and thus be more resilient.
+We assumed that when a disturbance event happen, due to a higher productivity, forest with an increased diversity will recover quicker and thus be more resilient.
 This theoritical expectation stand on two processes: complementarity and diversity effects [@Loreau2001].
@@
 Generally, positive relationship between biodiversity and productivity were emprically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of tropical forests.
-One of the few studies was realized by @Chisholm2013 and has shwon a positive significative relationship between species richness and wood productivity on a worlwide forest network.
+One of the few studies was realized by @Chisholm2013 and has shown a positive significative relationship between species richness and wood productivity on a worlwide forest network.
 But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.
 
-In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience, more particularly tropical forest carbon resilience.
+In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
 We used a simulation approach using TROLL model to assess long term processes.

### 2017-07-13 03-Model.Rmd
Thesis v1

@@
 Each tree is flagged with a species label linking to its species-specific attributes.
-Light is compputed explicitly at each timestep for each voxel.'}
+Light is computed explicitly at each timestep for each voxel.'}
 knitr::include_graphics('images/TROLLtree.png', dpi = NA)
@@
 
-In french guiana context sylviculture can be narrow to selective logging, which can be split in two steps: selection and harvesting.
+In french guiana context, sylviculture can be narrow to selective logging, which can be split in two steps: selection and harvesting.
 Selection encompass choice of the havrestable area, tree designation by the forest office, tree selection by the harvester, and removal off tree probbed as rotten by the lumber.
@@
 Tree harvesting roads are split in three classes: truck roads, main tractor track, and secondary track.
-Because TROLL assumes a flat environment, the main track is opened starting from the middle of one side of the simulated forest and untill it reaches the center with a width of 6 meters.
+Because TROLL assumes a flat environment, the main track is opened starting from the midle of one side of the simulated forest and untill it reaches the center with a width of 6 meters.
 In most cases, secondary tracks are opened once trees have been designated and the geolocation taken at a maximum distance of 30 meters from designated trees (Laurent Descroix, ONF, personnal communication).
@@
 
-$Death$ of a tree follows a $\mathcal{Bernoulli}$ law of probability $P(Death)$.
+$Death$ of a tree follows a $\mathcal{B}ernoulli$ law of probability $P(Death)$.
 The odds for a tree to die are calculated with the sum of the natural tree death odd $\theta$ and a perturbation index $\beta*e^{\alpha*d_{gaps}}$.

### 2017-07-13 04-MnMs.Rmd
Introduction v2
Thesis v1

@@
 Consequently we particulrarly needed to assess the importance of functional traits to further better control and evaluate functional diversities.
-We also needed to assess the sensitivity of TROLL model to the seed rain constant ($n_{ext}$, see \@ref(eq:next)) because we assumed it as one of the main factors of tree recruitments after disturbance within simulations.
+We also needed to assess the sensitivity of TROLL model to the seed rain constant ($n_{ext}$, see \@ref(eq:next)) because we assumed it was one of the main factors of tree recruitments after disturbance within simulations.
 
@@
 Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and 188.9 GB of memory.
-We assumed maturity of the forest after 500 years of regeneration (Maréchaux & Chave) and computed simulation 100 years after a disturbance event with 40% loss of basal area.
-Due to computer limitations we did not run replicate (besides it should be necessary to reduce simulation stochasticity).
+We assumed maturity of the forest after 500 years of regeneration @Li and computed simulation 100 years after a disturbance event with 40% loss of basal area.
+Due to computer limitations we did not run replicates (besides it should be necessary to reduce simulation stochasticity).
 To assess ecosystem outputs sensitivity to studied parameters, we compared it to 100 replicates of control simulations with all parameters set to default values.
@@
 In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time.
-Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation.
-Biodiversity was integrated with two components of its components: taxonomic and functional diversities.
+Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation using default parameters.
+Biodiversity was integrated with two of its components: taxonomic and functional diversities.
 We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species).
@@
 Consequently, to encompass undisturbed ecosystem variations throught time, we simulated an undisturbed control ecosystem $C$.
-And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the the ecosystem metric values from the control $X_C(t)$ .
+And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ .
 Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics.

### 2017-07-13 05-Results.Rmd
Thesis v1

@@
 On the other hand, few functional traits influenced final ecosystem structure.
-Only specific maximum diameter **dmax** add higher diversity for greater orders implying better eveness in species distributions.
+Only specific maximum diameter **dmax** add higher diversity for greater orders implying better evenness in species distributions.
 Regarding functional composition, traits fixed to mean did not change other functional traits density distribution.
@@
 We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years ($Ieq_{600}$).
-In addition, functional eveness was complementary reducing $Ieq_{600}$.
+In addition, functional evenness was complementary reducing $Ieq_{600}$.
 Finally species richness was not directly link to $Ieq_{600}$.
@@
 ```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional eveness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
+Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
 Ieq %>% 
@@
   scale_shape_discrete('Species\nrichness\n(SR)') +
-  scale_color_continuous('Functional\neveness\n(FEve)')
+  scale_color_continuous('Functional\nevenness\n(FEve)')
 ```

### 2017-07-13 06-Discussion.Rmd
Thesis v1

@@
 But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012.
-Because if the harvesting are not sustainable negative feedbaks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.
+Because if the harvesting is not sustainable negative feedbaks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.
 
@@
 
-Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and eveness ($p < 0.01$ and $p < 0.05$ respectivelly).
+Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectivelly).
 Thus we were able to show that diversity improve forest resilience.
-More particularly, functional diversity seems a major aspect of resilience if its enforced by a high functional eveness.
-Effectively high functional diversity needs eveness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species.
+More particularly, functional diversity seems a major aspect of resilience if its enforced by a high functional evenness.
+Effectively high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species.
 Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes.
-Additionally, the role of eveness  confirm the review of @Zhang2012 who highlighted the role of eveness in productivity and thus resilience following our hypothesis.
+Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis.
 Finally, species and functional richnesses is not directly increasing resilience.
@@
 We interpreted this results as the consequence of facilitation processes.
-As the only ressource simulated by TROLL is the light, the major facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps.
+As the only ressource simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps.
 But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species.
@@
 Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada.
-They also found that selection effect was greater than complementarity in most of the case.
+They also found that selection effect was greater than complementarity in most of the cases.
 In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield).
-High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two rotation in order to maintain both productive ecosystem and sustainable management.
+High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two cutting cycles in order to maintain both productive ecosystem and sustainable management.
 
@@
 We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance.
-We found that diversity improved forest resilience and that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery.
-Our results advocates for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles.
+We found that diversity improved forest resilience.
+Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery.
+Our results advocate for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles.
 Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience.
@@
 Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy.
-This view correspond to the high resilience, we were able to show, of tropical forest hyperdiverse systems and let hope for a future sustainable selective logging in tropical forests.
+This view correspond to the high resilience of tropical forest hyperdiverse systems, we were able to show, and let hope for a future sustainable selective logging in tropical forests.

### 2017-07-13 08-Appendix2.Rmd
Results: Disturbance & Appendix4
Thesis v1

@@
 Dataset defined as origin corresponded to leaf lifespan ($LL$) and most of the time to leaf mass per area ($LMA$) and leaf nitrogen content per leaf dry mass ($Nmass$).
-We measured variable importance in functionnal traits to explain leaf lifespan with out-of bag method applied on a random forest.
-Then, we used a béayesian apporach to test different models with growing level of complexity.
+We measured variable importance in functionnal traits to explain leaf lifespan with an out-of the bag method applied on a random forest.
+Then, we used a bayesian apporach to test different models with growing level of complexity.
 We retained the model with the best tradeoff between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean sqaure error of prediciton $RMSEP$).
@@
 Nmass and wsg) is not constant and depend on the closest point value for right (resp.
-center and left) graph.', cahce=TRUE, fig.height=12, fig.width=12}
+center and left) graph.', cache=TRUE, fig.height=12, fig.width=12}
 library(ggplot2)
@@
 
-Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early edath of individuals inside simulations.
+Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early death of individuals inside simulations.
 For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.

### 2017-07-13 09-Appendix3.Rmd
Thesis v1

@@
 
-In order to simulate sylviculture with TROLL we need to implement a new sylviculture module inside TROLL model code.
+In order to simulate sylviculture with TROLL we needed to implement a new sylviculture module inside TROLL model code.
 A first litterature review was completed by an interview with Laurent Descroix of the Office Nationale des Forêts.
@@
 
-In fact we have two different questions:
+In fact we have two different objectives:
 
@@
 
-$$V_f \sim \mathcal{logN}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]$$
+$$V_f \sim log\mathcal{N}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]$$
 
@@
     "${P_{rotten}}_n \\sim \\mathcal{B}(inv_{logit}(\\beta_0 + \\beta_1*dbh_n + {\\beta_2}_p + {\\beta_3}_s))$",
-    "$Volume_{of~wood} \\sim \\mathcal{logN} (log[((\\beta + \\beta_p + \\beta_s)*dbh^2)*(1 - Pr*((\\theta + \\theta_p + \\theta_s) *dbh^2))], \\sigma)$"
+    "$Volume_{of~wood} \\sim log\\mathcal{N} (log[((\\beta + \\beta_p + \\beta_s)*dbh^2)*(1 - Pr*((\\theta + \\theta_p + \\theta_s) *dbh^2))], \\sigma)$"
     )

### 2017-07-13 10-Appendix4.Rmd
Thesis v1

@@
 To study resistance and resilience of ecosystem face to disturbance, highlighting the role of biodiversity, we decided to use TROLL model simulations [@Chave1999].
-In order to get a finer study of simulations response, we need to assess first sensitivity of the TROLL model to different parameters.
-More particularly we need to assess the importance of functional traits to further better control functional diversities in simulations.
-We also need to assess sensitivity of the model to see rain constant because we assume it is one of the main factor of tree recruitments after disturbance in the model.
+In order to get a finer study of simulations response, we needed to assess sensitivity of the TROLL model to different parameters.
+More particularly we needed to assess the importance of functional traits to further better control functional diversities in simulations.
+We also needed to assess sensitivity of the model to see rain constant because we assumed it was one of the main factor of tree recruitments after disturbance in the model.
 
@@
 Each trait was tested independently.
-We reduce to a common mean traits with a correlation $r \geq 0.8$ (see figure \@ref(fig:A4corr)).
+We reduced to a common mean traits with a correlation $r \geq 0.8$ (see figure \@ref(fig:A4corr)).
 
@@
 We assumed maturity of the forest after 500 years of regeneration [@Li] and computed simulation 100 years after a disturbance event of 40% intensity.
-Due to computer limitations we did not run replicate (besides it should be necessary to reduce simulation stochasticity).
+Due to computer limitations we did not run replicates (besides it should be necessary to reduce simulation stochasticity).
 To assess ecosystem outputs sensitivity to studied parameters, we compared it to 100 replicates of control simulations with all parameters set to default values.
@@
 On the other hand, few functional traits influenced final ecosystem structure (figure \@ref(fig:A4ftStr)).
-Only specific maximum diameter **dmax** add higher diversity for greater orders implying better eveness in species distributions.
+Only specific maximum diameter **dmax** add higher diversity for greater orders implying better evenness in species distributions.
 Regarding functional composition, traits fixed to mean did not change other functional traits density distribution.
@@
 Similarly, wood specific gravity **wsg** trait fixed to mean had exactly the same effect on number of stems above 10 and above 30 cm dbh and basal area after disturbance than **ah-hmax** but with a time lag ; and **wsg** also increased aboveground biomass.
-**dmax** trait fixed to mean slightly decreased number of stems above 10 and above 30 cm dbh over time while it increased basal area, aboveground biomass, and species eveness.
+**dmax** trait fixed to mean slightly decreased number of stems above 10 and above 30 cm dbh over time while it increased basal area, aboveground biomass, and species evenness.
 Finally, leaf mass per area **LMA** trait fixed to mean only decreased number of stem above 10 cm dbh after ecosystem resilience to disturbance (approximately 50 years) but did not affect other ecosystem outputs.
@@
 Decreased death rate diminished number of stems, expecially big ones thus increasing global basal area and abiveground biomass.
-**dmax** fixed to mean by keeping more small diameter stems also increased random selection of species increasing eveness in species distribution.
+**dmax** fixed to mean by keeping more small diameter stems also increased random selection of species increasing evenness in species distribution.
 Considering the high correlation between **ah** and **hmax** we could also keep only **hmax** (because of its more straightforward ecological meaning).
@@
 Reducing seedrain constant resulted in an ecosystem selecting few species increasing their abundance and functional dominance of their traits.
-Thus reduced seedrain constant greatly diminsihed eveness untill a decrease of total number of species for its lowest value.
+Thus reduced seedrain constant greatly diminsihed evenness untill a decrease of total number of species for its lowest value.

### 2017-07-13 docs/master-thesis.tex
Thesis v1

@@
 \item
+  Maxime
+\item
+  Raphaël
+\item
+  Paracou team
+\item
   \ldots{}
@@
 selective logging on tropical forest ecosystems to implement better
-syslvicultural practives in order to reach sustainability.
+sylvicultural practices in order to reach sustainability.
 
@@
 between biodiversity and productivity. We assumed that when a
-disturbance event happen, due to a higher productivity, foerst with an
-increased diversity with recover quickly and thus be more resilient.
+disturbance event happen, due to a higher productivity, forest with an
+increased diversity will recover quicker and thus be more resilient.
 This theoritical expectation stand on two processes: complementarity and
@@
 the case of tropical forests. One of the few studies was realized by
-\citet{Chisholm2013} and has shwon a positive significative relationship
+\citet{Chisholm2013} and has shown a positive significative relationship
 between species richness and wood productivity on a worlwide forest
@@
 In the present study, we focused on mechanisms involved in the
-relationship between biodiversity and forest ecosystem resilience, more
-particularly tropical forest carbon resilience. We used a simulation
-approach using TROLL model to assess long term processes.
+relationship between biodiversity and forest ecosystem resilience. We
+used a simulation approach using TROLL model to assess long term
+processes.
 
@@
 growth. Each tree is flagged with a species label linking to its
-species-specific attributes. Light is compputed explicitly at each
+species-specific attributes. Light is computed explicitly at each
 timestep for each voxel.}
@@
 
-In french guiana context sylviculture can be narrow to selective
+In french guiana context, sylviculture can be narrow to selective
 logging, which can be split in two steps: selection and harvesting.
@@
 tractor track, and secondary track. Because TROLL assumes a flat
-environment, the main track is opened starting from the middle of one
+environment, the main track is opened starting from the midle of one
 side of the simulated forest and untill it reaches the center with a
@@
 
-\(Death\) of a tree follows a \(\mathcal{Bernoulli}\) law of probability
+\(Death\) of a tree follows a \(\mathcal{B}ernoulli\) law of probability
 \(P(Death)\). The odds for a tree to die are calculated with the sum of
@@
 needed to assess the sensitivity of TROLL model to the seed rain
-constant (\(n_{ext}\), see \eqref{eq:next}) because we assumed it as one
+constant (\(n_{ext}\), see \eqref{eq:next}) because we assumed it was one
 of the main factors of tree recruitments after disturbance within
@@
 188.9 GB of memory. We assumed maturity of the forest after 500 years of
-regeneration (Maréchaux \& Chave) and computed simulation 100 years
-after a disturbance event with 40\% loss of basal area. Due to computer
-limitations we did not run replicate (besides it should be necessary to
+regeneration \citet{Li} and computed simulation 100 years after a
+disturbance event with 40\% loss of basal area. Due to computer
+limitations we did not run replicates (besides it should be necessary to
 reduce simulation stochasticity). To assess ecosystem outputs
@@
 50\% and 75\%), or as a selective logging simulation using default
-parameters. Biodiversity was integrated with two components of its
-components: taxonomic and functional diversities. We used species
-richness \(SR\) to represents taxonomic diversity (5, 25, and 125
-species). Functional diversity can be related to numerous components,
-and \citet{Borgy2017} argued for 5: richness, divergence, regularity,
-overlap and mean. Because mature forest were created from a bare soil
-with TROLL simulations, we could not control a priori divergence,
-regularity and overlap but only assess theim before diturbance.
-Consequently, we focused on functional richness with convex hull volume
-\(CHV\) and functional mean with community weighted mean \(CWM\). For
-each level of species richness \(SR\), we selected 20 communities with
-growing convex hull volume \(CHV\) but with a community weighted means
-close to the regional species pool community weighted means.
-Effectivelly, we did not wanted drastic change in community means that
-could have more effect than functional richness itself. This design of
-experiments resulted in 60 communities (\(5~SR*20~CHV\)) and 240
-simulations (\(60 ~communities*4~levels~of~disturbance\)) over 600 years
-(maturity being assumed after 500 years of regeneration \citep{Li}).
-Figure \ref{fig:DOE} presents the design of experiment for communities
+parameters. Biodiversity was integrated with two of its components:
+taxonomic and functional diversities. We used species richness \(SR\) to
+represents taxonomic diversity (5, 25, and 125 species). Functional
+diversity can be related to numerous components, and \citet{Borgy2017}
+argued for 5: richness, divergence, regularity, overlap and mean.
+Because mature forest were created from a bare soil with TROLL
+simulations, we could not control a priori divergence, regularity and
+overlap but only assess theim before diturbance. Consequently, we
+focused on functional richness with convex hull volume \(CHV\) and
+functional mean with community weighted mean \(CWM\). For each level of
+species richness \(SR\), we selected 20 communities with growing convex
+hull volume \(CHV\) but with a community weighted means close to the
+regional species pool community weighted means. Effectivelly, we did not
+wanted drastic change in community means that could have more effect
+than functional richness itself. This design of experiments resulted in
+60 communities (\(5~SR*20~CHV\)) and 240 simulations
+(\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
+being assumed after 500 years of regeneration \citep{Li}). Figure
+\ref{fig:DOE} presents the design of experiment for communities
 biodiversity after the mature forest were simulated, and thus before
@@
 the ratio of the ecosystem metric values in the disturbed simulation
-\(X_T(t)\) over the the ecosystem metric values from the control
-\(X_C(t)\) . Thus, the value of resilience \(R(t)\) is normalized for
-all simulations and metrics. \(R(t)\) will be equal to \(R_{eq} = 1\)
-when reaching the equilibrium value. Consequently we can calculate an
+\(X_T(t)\) over the ecosystem metric values from the control \(X_C(t)\)
+. Thus, the value of resilience \(R(t)\) is normalized for all
+simulations and metrics. \(R(t)\) will be equal to \(R_{eq} = 1\) when
+reaching the equilibrium value. Consequently we can calculate an
 euclidean distance to equilibrium \(d_{eq}(t)\) as
@@
 maximum diameter \textbf{dmax} add higher diversity for greater orders
-implying better eveness in species distributions. Regarding functional
+implying better evenness in species distributions. Regarding functional
 composition, traits fixed to mean did not change other functional traits
@@
 equilibrium after 600 years (\(Ieq_{600}\)). In addition, functional
-eveness was complementary reducing \(Ieq_{600}\). Finally species
+evenness was complementary reducing \(Ieq_{600}\). Finally species
 richness was not directly link to \(Ieq_{600}\). Effectivelly, low
@@
 (25, 50 and 75\% of total basal area); dot shapes represents the species
-richness whereas dot color represents functional eveness
+richness whereas dot color represents functional evenness
 \citep[FEve,][]{villeger_new_2008}.}
@@
 conclusion should met a sustainable definiton of selective logging
-following \citet{Zimmerman2012}. Because if the harvesting are not
+following \citet{Zimmerman2012}. Because if the harvesting is not
 sustainable negative feedbaks will slowly diminish diversity and its
@@
 Our results validated the hypothesis of a significative relationship
-between forest resilience and functional diversity and eveness
+between forest resilience and functional diversity and evenness
 (\(p < 0.01\) and \(p < 0.05\) respectivelly). Thus we were able to show
@@
 diversity seems a major aspect of resilience if its enforced by a high
-functional eveness. Effectively high functional diversity needs eveness
-in order to better answer disturbance, if not the diversity is masked by
-ecosystem dominant species. Those results confirms the review of
-\citet{Diaz2001} advocating for underevaluated importance of plant
+functional evenness. Effectively high functional diversity needs
+evenness in order to better answer disturbance, if not the diversity is
+masked by ecosystem dominant species. Those results confirms the review
+of \citet{Diaz2001} advocating for underevaluated importance of plant
 functional diversity in ecosystem processes. Additionally, the role of
-eveness confirm the review of \citet{Zhang2012} who highlighted the role
-of eveness in productivity and thus resilience following our hypothesis.
-Finally, species and functional richnesses is not directly increasing
-resilience. But increased species and functional richnesses will
-increase chance for high functional diversity through the sampling
+evenness confirm the review of \citet{Zhang2012} who highlighted the
+role of evenness in productivity and thus resilience following our
+hypothesis. Finally, species and functional richnesses is not directly
+increasing resilience. But increased species and functional richnesses
+will increase chance for high functional diversity through the sampling
 effect \citep{Loreau1998}. An higher sampling of regional species pool
@@
 consequence of facilitation processes. As the only ressource simulated
-by TROLL is the light, the major facilitation will be light shading of
+by TROLL is the light, the main facilitation will be light shading of
 post-pioneer species by pioneer species in disturbance gaps. But the
@@
 diverse forest of Canada. They also found that selection effect was
-greater than complementarity in most of the case. In the case of
+greater than complementarity in most of the cases. In the case of
 selective logging the complementarity effect will thus be the major
@@
 complementarity through increased functional diversity is thus an
-advantage for forest recovery between two rotation in order to maintain
-both productive ecosystem and sustainable management.
+advantage for forest recovery between two cutting cycles in order to
+maintain both productive ecosystem and sustainable management.
 
@@
 resilience to disturbance. We found that diversity improved forest
-resilience and that complementarity between species was insuring forest
-recovery in forest succession start with facilitation before more
-productive species dominate the forest and insure recovery. Our results
-advocates for sustainable selective logging against monoculture stand:
-high diversity will increase forest resilience and thus improve logging
-cycles. Even if monoculture stand are naturalized they will not reach
-mature forest diversity, and consequently they will not reach its
-natural high resilience. But on the other hand selective logging in
-tropical forest needs to be sustainable, if not the diversity will
-slowly decrease after each cycles degrading forest ecosystem functions,
-and resilience will decrease due to negative feedbacks.
-\citet{Zimmerman2012} criticized the state of sylviculture in tropics,
-suggesting that ``we have not been able to reconcile these opposing
-biological and economic forces''. Still, they advocates for a
+resilience. Additionally we found that complementarity between species
+was insuring forest recovery in forest succession start with
+facilitation, before more productive species dominate the forest and
+insure forest recovery. Our results advocate for sustainable selective
+logging against monoculture stand: high diversity will increase forest
+resilience and thus improve logging cycles. Even if monoculture stand
+are naturalized they will not reach mature forest diversity, and
+consequently they will not reach its natural high resilience. But on the
+other hand selective logging in tropical forest needs to be sustainable,
+if not the diversity will slowly decrease after each cycles degrading
+forest ecosystem functions, and resilience will decrease due to negative
+feedbacks. \citet{Zimmerman2012} criticized the state of sylviculture in
+tropics, suggesting that ``we have not been able to reconcile these
+opposing biological and economic forces''. Still, they advocates for a
 possibility of sustainable tropical logging, already existing in
 small-scale, that will need proper funding from the internationnal
-communtiy. This view correspond to the high resilience, we were able to
-show, of tropical forest hyperdiverse systems and let hope for a future
-sustainable selective logging in tropical forests.
+communtiy. This view correspond to the high resilience of tropical
+forest hyperdiverse systems, we were able to show, and let hope for a
+future sustainable selective logging in tropical forests.
 
@@
 dry mass (\(Nmass\)). We measured variable importance in functionnal
-traits to explain leaf lifespan with out-of bag method applied on a
-random forest. Then, we used a béayesian apporach to test different
+traits to explain leaf lifespan with an out-of the bag method applied on
+a random forest. Then, we used a bayesian apporach to test different
 models with growing level of complexity. We retained the model with the
@@
 unrealistically low aboveground biomass for low LMA species due to an
-early edath of individuals inside simulations. For instance with this
+early death of individuals inside simulations. For instance with this
 allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic
@@
 
-In order to simulate sylviculture with TROLL we need to implement a new
-sylviculture module inside TROLL model code. A first litterature review
-was completed by an interview with Laurent Descroix of the Office
+In order to simulate sylviculture with TROLL we needed to implement a
+new sylviculture module inside TROLL model code. A first litterature
+review was completed by an interview with Laurent Descroix of the Office
 Nationale des Forêts. We discovered that rotten trees were not random
@@
 
-\[V_f \sim \mathcal{logN}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]\]
+\[V_f \sim log\mathcal{N}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]\]
 
@@
 \(N_{s,p} + L_{s,p}\) &
-\(Volume_{of~wood} \sim \mathcal{logN} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)\tabularnewline
+\(Volume_{of~wood} \sim log\mathcal{N} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)\tabularnewline
 \bottomrule
@@
 \(N_{s,p} + L_{s,p}\):
-\(Volume_{of~wood} \sim \mathcal{logN} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)
+\(Volume_{of~wood} \sim log\mathcal{N} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)
 
@@
 simulations \citep{Chave1999}. In order to get a finer study of
-simulations response, we need to assess first sensitivity of the TROLL
-model to different parameters. More particularly we need to assess the
+simulations response, we needed to assess sensitivity of the TROLL model
+to different parameters. More particularly we needed to assess the
 importance of functional traits to further better control functional
-diversities in simulations. We also need to assess sensitivity of the
-model to see rain constant because we assume it is one of the main
+diversities in simulations. We also needed to assess sensitivity of the
+model to see rain constant because we assumed it was one of the main
 factor of tree recruitments after disturbance in the model.
@@
 we performed a sensitivity analysis by fixing species trait values to
-their mean. Each trait was tested independently. We reduce to a common
+their mean. Each trait was tested independently. We reduced to a common
 mean traits with a correlation \(r \geq 0.8\) (see figure
@@
 disturbance event of 40\% intensity. Due to computer limitations we did
-not run replicate (besides it should be necessary to reduce simulation
+not run replicates (besides it should be necessary to reduce simulation
 stochasticity). To assess ecosystem outputs sensitivity to studied
@@
 \ref{fig:A4ftStr}). Only specific maximum diameter \textbf{dmax} add
-higher diversity for greater orders implying better eveness in species
+higher diversity for greater orders implying better evenness in species
 distributions. Regarding functional composition, traits fixed to mean
@@
 time while it increased basal area, aboveground biomass, and species
-eveness. Finally, leaf mass per area \textbf{LMA} trait fixed to mean
+evenness. Finally, leaf mass per area \textbf{LMA} trait fixed to mean
 only decreased number of stem above 10 cm dbh after ecosystem resilience
@@
 fixed to mean by keeping more small diameter stems also increased random
-selection of species increasing eveness in species distribution.
+selection of species increasing evenness in species distribution.
 Considering the high correlation between \textbf{ah} and \textbf{hmax}
@@
 abundance and functional dominance of their traits. Thus reduced
-seedrain constant greatly diminsihed eveness untill a decrease of total
+seedrain constant greatly diminsihed evenness untill a decrease of total
 number of species for its lowest value.
@@
 
-Appendix 4 presents ecosystem resilience after 600 years with taxonomic
-and functional diversity for different levels of disturbance. It
-encompass all functionnal diversity components \citep[FRIC, FEve, FDiv,
-and FDis,][]{villeger_new_2008}. And it presents results for both forest
-dynamic (Figure \ref{fig:A5IeqStructureAll}) and forest production
-(Figure \ref{fig:A5IeqProductionAll}).
+This appendix presents ecosystem resilience after 600 years with
+taxonomic and functional diversity for different levels of disturbance.
+It encompass all functionnal diversity components \citep[FRIC, FEve,
+FDiv, and FDis,][]{villeger_new_2008}. And it presents results for both
+forest dynamic (Figure \ref{fig:A5IeqStructureAll}) and forest
+production (Figure \ref{fig:A5IeqProductionAll}).
 

### 2017-07-18 02-Introduction.Rmd
Bruno 1st review

@@
 Degradation from forest has been estimated to represents 10 time deforestation [@Herold2011] and represents 20% of greenhouse gas emissions in the brazilian Amazon [@Asner2005].
+Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.
 
@@
 managed selective harvesting of timber) has been widely promoted internationnaly to combat tropical deforestation and degradation [@Zimmerman2012].
-Currently tropical logging accounts for one eight of global timber production [@Blaser2011] and is still increasing.
+Currently logging from tropical forests accounts for one eight of global timber production [@Blaser2011] and is still increasing.
 Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cutting cycle [@Martin2015].
@@
 And forest ecosystem response to logging disturbance will directly modify ecosystem functionning in both short and long term.
-Consequently assessing selective logging effect on tropical forest linking biodiversity and ecosystem seems an obvious and promising way [@Loreau2010].
+Consequently assessing selective logging effect on tropical forest linking diversity and ecosystem functionning seems an obvious and promising way [@Loreau2010].
 
@@
 Individual-based models of forest dynamics present the perfect framework to develop such joint biodiversity-ecosystem approaches [@Li].
-Individual-based models describe forest 'patches' accumulating carbon through time, assessing tree growth within the patch, or releasing carbon through gap opening [@Bugmann2001].
+Individual-based models describe forest accumulating carbon through time, assessing tree growth, or releasing carbon through gap opening [@Bugmann2001].
 Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit [@Pacala1996].

### 2017-07-18 03-Model.Rmd
Bruno 1st review

@@
 TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production.
-Trees are growing in a located light environment explicitly computed witin voxels of 1 $m^3$.
+Trees are growing in a spatialized light environment explicitly computed witin voxels of 1 $m^3$.
 Each tree is consistently defined by its age, diameter at brest height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA) (see figure \@ref(fig:TROLLtree)).
@@
 In french guiana context, sylviculture can be narrow to selective logging, which can be split in two steps: selection and harvesting.
-Selection encompass choice of the havrestable area, tree designation by the forest office, tree selection by the harvester, and removal off tree probbed as rotten by the lumber.
+Selection encompass choice of the havrestable area, harvestable tree designation by the forest office, harvested tree selection by the harvester, and removal off tree probbed as rotten by the lumber.
 Harvesting encompass tree felling, tracks opening, and long term damages (simplified in gap damages in current TROLL implementation).

### 2017-07-18 04-MnMs.Rmd
Bruno 1st review

@@
 Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean.
-Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess theim before diturbance.
+Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess them after running the simulations, i.e.
+before applying the diturbance.
 Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$.
@@
 
-## Ecosystem answer analysis
+## Ecosystem response analysis
 
@@
 Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions.
-We decided to describe simiulated tropical forests in two major functions: forest dynamic and forest production.
-Forest dynamic was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$).
-Forest production was represented by growth pimary productivity ($GPP$ in $MgC.ha^{-1}$), net pimary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
+We decided to describe simiulated tropical forests in two major functions: forest structure and forest functionning.
+Forest structure was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$).
+Forest functionning was represented by growth pimary productivity ($GPP$ in $MgC.ha^{-1}$), net pimary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
 
@@
 Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$.
-Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest dynamic (AGB, BA, N, N10, and N30) and forest production (GPP, NPP, Rday, and Rnight).
-We then used cummulative integral over time of eulcidean distance to equilibrium to assess simulations resilience.
+Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest strcuture (AGB, BA, N, N10, and N30) and forest functionning (GPP, NPP, Rday, and Rnight).
+We then used integrated eulcidean distance to equilibrium over time to assess simulations resilience.
 
@@
 Whereas $SE$ become positive when dominant species outperform theimselves in mixture than in monoculture, and negative when less dominant species outperform theimselves in mixture than in monoculture [@Tobner2016].
-But similarly to resilience measurement, biodiversity net effect $NE$ can be in a dynamic equilibrium and vary over time without disturbance.
+But similarly to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
 So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:

### 2017-07-18 05-Results.Rmd
Bruno 1st review

@@
 Most of functional traits had a significant long term influence on ecosystem outputs (figure \@ref(fig:ftVar)).
-On the other hand, few functional traits influenced final ecosystem structure.
 Only specific maximum diameter **dmax** add higher diversity for greater orders implying better evenness in species distributions.
@@
 Number of trees with dbh above 10 cm (N10) and 30 cm (N30), above ground biomass (AGB) and basal area (BA).
-Grey area represents the interval of control replicates whereas black line represents the mean of control replicates.', cache=TRUE, warning=FALSE}
+Sensitivity of model to functional traits was performed  by fixing species trait values to their mean.
+Grey area represents the interval of control replicates whereas black line represents the mean of control replicates, thus if ecosystems outputs are outside of grey area values the studied parameter is considered to have a significant influence on the model.', cache=TRUE, warning=FALSE}
 path <- '~/Documents/ECOFOG/Results/sensitivity/control'
@@
 The net effect was then partition between selection and complementarity effect.
-We normalized treatments effect by control net effect (see Table \@ref(tab:tNE)) to measure their aboveground biomass resilience to disturbance (see Figure \@ref(fig:gBE)).
-We found that complementarity effect was recovering net effect in the first decades untill it diminished after a century.
+We normalized complementarity and selection effect of disturbed simulations by biodiversity net effect of undisturbed control simulation (see Table \@ref(tab:tNE)) to measure the resilience to disturbance of their aboveground biomass (see Figure \@ref(fig:gBE)).
+We found that complementarity effect was recovering biodiversity net effect and was stronger than selection effect in the first decades.
+But after few decades the complementarity effect diminsihed toward a low value.
 On the contrary selection effect was reduced or even removed by the disturbance.
@@
 We obtained those results for aboveground biomass.
-We found similar results but with an amlified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]).
+We found similar results but with an aplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]).
 Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 5: Disturbance simulations]).

### 2017-07-18 06-Discussion.Rmd
Bruno 1st review

@@
 High external seed rain resulted in a quick recovery of the system toward an equilibrium close to the regional species frequency levied by the seed rain.
-On the contrary, low seed rain let the simulated forest works has a closed system with more system feedback but a lower stability through time with longer species diversity transitions.
+On the contrary, low seed rain let the simulated forest works as a closed system with more system feedback but a lower stability through time with longer species diversity transitions.
 In order to study the role of diversity in forest resilience we decided to remove the seed rain constant to get a closed system and look at the role of diversity when it maintains itself through feedbacks and not with immigration.
@@
 Thus we were able to show that diversity improve forest resilience.
-More particularly, functional diversity seems a major aspect of resilience if its enforced by a high functional evenness.
+More particularly, functional diversity seems a major aspect of resilience if its strengthened by a high functional evenness.
 Effectively high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species.
@@
 Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis.
-Finally, species and functional richnesses is not directly increasing resilience.
+Finally, species and functional richnesses are not directly increasing resilience.
 But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998].
@@
 As the only ressource simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps.
+We found similar results to @Morin2011 regarding complementarity and light facilitation.
 But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species.
@@
 We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance.
-We found that diversity improved forest resilience.
+We found that diversity improved forest resilience together with productivity [@Liang2016].
 Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery.

### 2017-07-18 docs/master-thesis.tex
Bruno 1st review

@@
 deforestation \citep{Herold2011} and represents 20\% of greenhouse gas
-emissions in the brazilian Amazon \citep{Asner2005}.
+emissions in the brazilian Amazon \citep{Asner2005}. Degradation has met
+numerous definitions \citep{Simula2009}, but can be defined as the
+result from a disturbance event that induced a modification of the
+forest ecosystem reducing ecosystem services while keeping the ecosystem
+as a forest, contrary to deforestation.
 
@@
 tropical deforestation and degradation \citep{Zimmerman2012}. Currently
-tropical logging accounts for one eight of global timber production
-\citep{Blaser2011} and is still increasing. Most tropical timber
-production originates from selective logging, the targeted harvesting of
-timber from commercial species in a single cutting cycle
+logging from tropical forests accounts for one eight of global timber
+production \citep{Blaser2011} and is still increasing. Most tropical
+timber production originates from selective logging, the targeted
+harvesting of timber from commercial species in a single cutting cycle
 \citep{Martin2015}.
@@
 short and long term. Consequently assessing selective logging effect on
-tropical forest linking biodiversity and ecosystem seems an obvious and
-promising way \citep{Loreau2010}.
+tropical forest linking diversity and ecosystem functionning seems an
+obvious and promising way \citep{Loreau2010}.
 
@@
 joint biodiversity-ecosystem approaches \citep{Li}. Individual-based
-models describe forest `patches' accumulating carbon through time,
-assessing tree growth within the patch, or releasing carbon through gap
-opening \citep{Bugmann2001}. Up to several dozens of different Plant
-Functional Types (PFTs) are generally defined and models can sometimes
-be fully spatially explicit \citep{Pacala1996}. Recently, the forest
-growth simulator TROLL \citep{Chave1999}, an individual-based and
-spatially explicit forest model, was developped to introduce recent
-advances in plant physiological community. TROLL model relates
-physiological processes to species-specific functional traits
-\citep{Li}. Consequently, TROLL model allow to simulate fully a
-neotropical forest biodiversity to study biodiversity-ecosystem
-functionning link response to logging disturbance.
+models describe forest accumulating carbon through time, assessing tree
+growth, or releasing carbon through gap opening \citep{Bugmann2001}. Up
+to several dozens of different Plant Functional Types (PFTs) are
+generally defined and models can sometimes be fully spatially explicit
+\citep{Pacala1996}. Recently, the forest growth simulator TROLL
+\citep{Chave1999}, an individual-based and spatially explicit forest
+model, was developped to introduce recent advances in plant
+physiological community. TROLL model relates physiological processes to
+species-specific functional traits \citep{Li}. Consequently, TROLL model
+allow to simulate fully a neotropical forest biodiversity to study
+biodiversity-ecosystem functionning link response to logging
+disturbance.
 
@@
 height (dbh) above 1 cm, to death with growth and seed production. Trees
-are growing in a located light environment explicitly computed witin
+are growing in a spatialized light environment explicitly computed witin
 voxels of 1 \(m^3\). Each tree is consistently defined by its age,
@@
 logging, which can be split in two steps: selection and harvesting.
-Selection encompass choice of the havrestable area, tree designation by
-the forest office, tree selection by the harvester, and removal off tree
-probbed as rotten by the lumber. Harvesting encompass tree felling,
-tracks opening, and long term damages (simplified in gap damages in
-current TROLL implementation).
+Selection encompass choice of the havrestable area, harvestable tree
+designation by the forest office, harvested tree selection by the
+harvester, and removal off tree probbed as rotten by the lumber.
+Harvesting encompass tree felling, tracks opening, and long term damages
+(simplified in gap damages in current TROLL implementation).
 
@@
 simulations, we could not control a priori divergence, regularity and
-overlap but only assess theim before diturbance. Consequently, we
-focused on functional richness with convex hull volume \(CHV\) and
-functional mean with community weighted mean \(CWM\). For each level of
-species richness \(SR\), we selected 20 communities with growing convex
-hull volume \(CHV\) but with a community weighted means close to the
-regional species pool community weighted means. Effectivelly, we did not
-wanted drastic change in community means that could have more effect
-than functional richness itself. This design of experiments resulted in
-60 communities (\(5~SR*20~CHV\)) and 240 simulations
+overlap but only assess them after running the simulations, i.e.~before
+applying the diturbance. Consequently, we focused on functional richness
+with convex hull volume \(CHV\) and functional mean with community
+weighted mean \(CWM\). For each level of species richness \(SR\), we
+selected 20 communities with growing convex hull volume \(CHV\) but with
+a community weighted means close to the regional species pool community
+weighted means. Effectivelly, we did not wanted drastic change in
+community means that could have more effect than functional richness
+itself. This design of experiments resulted in 60 communities
+(\(5~SR*20~CHV\)) and 240 simulations
 (\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
@@
 
-\subsection{Ecosystem answer analysis}\label{ecosystem-answer-analysis}
+\subsection{Ecosystem response
+analysis}\label{ecosystem-response-analysis}
 
@@
 to several ecosystem functions. We decided to describe simiulated
-tropical forests in two major functions: forest dynamic and forest
-production. Forest dynamic was represented by aboveground biomass
+tropical forests in two major functions: forest structure and forest
+functionning. Forest structure was represented by aboveground biomass
 (\(AGB\) in \(ton~C.ha^{-1}\)), basal area (\(BA\) in \(m^2.ha^{-1}\)),
@@
 (\(N10\)), and number of stem above \(30~cm\) diameter (\(N30\)). Forest
-production was represented by growth pimary productivity (\(GPP\) in
+functionning was represented by growth pimary productivity (\(GPP\) in
 \(MgC.ha^{-1}\)), net pimary productivity (\(NPP\) in \(MgC.ha^{-1}\)),
@@
 to equilibrium was calculated in a multi-dimensional space for the two
-functions described above: forest dynamic (AGB, BA, N, N10, and N30) and
-forest production (GPP, NPP, Rday, and Rnight). We then used cummulative
-integral over time of eulcidean distance to equilibrium to assess
+functions described above: forest strcuture (AGB, BA, N, N10, and N30)
+and forest functionning (GPP, NPP, Rday, and Rnight). We then used
+integrated eulcidean distance to equilibrium over time to assess
 simulations resilience.
@@
 monoculture \citep{Tobner2016}. But similarly to resilience measurement,
-biodiversity net effect \(NE\) can be in a dynamic equilibrium and vary
-over time without disturbance. So in order to correctly assess selection
-and complementarity effect in answer to disturbance, we normalized it by
+biodiversity net effect \(NE\) is a dynamic equilibrium and vary over
+time without disturbance. So in order to correctly assess selection and
+complementarity effect in answer to disturbance, we normalized it by
 undisturbed control ecosystem net effect \(NE_C\) to measure treatments
@@
 Most of functional traits had a significant long term influence on
-ecosystem outputs (figure \ref{fig:ftVar}). On the other hand, few
-functional traits influenced final ecosystem structure. Only specific
-maximum diameter \textbf{dmax} add higher diversity for greater orders
-implying better evenness in species distributions. Regarding functional
+ecosystem outputs (figure \ref{fig:ftVar}). Only specific maximum
+diameter \textbf{dmax} add higher diversity for greater orders implying
+better evenness in species distributions. Regarding functional
 composition, traits fixed to mean did not change other functional traits
@@
 variations over time. Number of trees with dbh above 10 cm (N10) and 30
-cm (N30), above ground biomass (AGB) and basal area (BA). Grey area
-represents the interval of control replicates whereas black line
-represents the mean of control replicates.}
+cm (N30), above ground biomass (AGB) and basal area (BA). Sensitivity of
+model to functional traits was performed by fixing species trait values
+to their mean. Grey area represents the interval of control replicates
+whereas black line represents the mean of control replicates, thus if
+ecosystems outputs are outside of grey area values the studied parameter
+is considered to have a significant influence on the model.}
 \end{figure}
@@
 monoculture simulations. The net effect was then partition between
-selection and complementarity effect. We normalized treatments effect by
-control net effect (see Table \ref{tab:tNE}) to measure their
-aboveground biomass resilience to disturbance (see Figure
-\ref{fig:gBE}). We found that complementarity effect was recovering net
-effect in the first decades untill it diminished after a century. On the
-contrary selection effect was reduced or even removed by the
-disturbance. It increased during the whole simulation and was greater
-than complementarity effect only after decades. The time lag for which
-complementarity effect was greater than selection effect was increasing
-with disturbance intensity. Finally 600 years after the disturbance
-event biodiversity net effect was still not recovered for a disturbance
-intensity greater than 25\% of basal area. We obtained those results for
-aboveground biomass. We found similar results but with an amlified
-signal for basal area (\(BA\)) and stem abundance (\(N\) but with an
-inverted signal because of the forest self thinning) (see
+selection and complementarity effect. We normalized complementarity and
+selection effect of disturbed simulations by biodiversity net effect of
+undisturbed control simulation (see Table \ref{tab:tNE}) to measure the
+resilience to disturbance of their aboveground biomass (see Figure
+\ref{fig:gBE}). We found that complementarity effect was recovering
+biodiversity net effect and was stronger than selection effect in the
+first decades. But after few decades the complementarity effect
+diminsihed toward a low value. On the contrary selection effect was
+reduced or even removed by the disturbance. It increased during the
+whole simulation and was greater than complementarity effect only after
+decades. The time lag for which complementarity effect was greater than
+selection effect was increasing with disturbance intensity. Finally 600
+years after the disturbance event biodiversity net effect was still not
+recovered for a disturbance intensity greater than 25\% of basal area.
+We obtained those results for aboveground biomass. We found similar
+results but with an aplified signal for basal area (\(BA\)) and stem
+abundance (\(N\) but with an inverted signal because of the forest self
+thinning) (see
 \protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
@@
 equilibrium close to the regional species frequency levied by the seed
-rain. On the contrary, low seed rain let the simulated forest works has
-a closed system with more system feedback but a lower stability through
+rain. On the contrary, low seed rain let the simulated forest works as a
+closed system with more system feedback but a lower stability through
 time with longer species diversity transitions. In order to study the
@@
 that diversity improve forest resilience. More particularly, functional
-diversity seems a major aspect of resilience if its enforced by a high
-functional evenness. Effectively high functional diversity needs
+diversity seems a major aspect of resilience if its strengthened by a
+high functional evenness. Effectively high functional diversity needs
 evenness in order to better answer disturbance, if not the diversity is
@@
 role of evenness in productivity and thus resilience following our
-hypothesis. Finally, species and functional richnesses is not directly
+hypothesis. Finally, species and functional richnesses are not directly
 increasing resilience. But increased species and functional richnesses
@@
 by TROLL is the light, the main facilitation will be light shading of
-post-pioneer species by pioneer species in disturbance gaps. But the
-complementarity effect is reducing through time, to let the selection
-effect insure forest resilience due to an inforced dominance of more
-productive species. The diminishing of complementarity effect is due to
-competitive selection though time in gaps succession. But the study
-scale matters as shwon by \citet{Chisholm2013}. Here we look at
-processes to a 16 ha scale. We do not have topography reducing
+post-pioneer species by pioneer species in disturbance gaps. We found
+similar results to \citet{Morin2011} regarding complementarity and light
+facilitation. But the complementarity effect is reducing through time,
+to let the selection effect insure forest resilience due to an inforced
+dominance of more productive species. The diminishing of complementarity
+effect is due to competitive selection though time in gaps succession.
+But the study scale matters as shwon by \citet{Chisholm2013}. Here we
+look at processes to a 16 ha scale. We do not have topography reducing
 micro-environment effects, but \citet{Chisholm2013} results suggest that
@@
 resilience to disturbance. We found that diversity improved forest
-resilience. Additionally we found that complementarity between species
-was insuring forest recovery in forest succession start with
-facilitation, before more productive species dominate the forest and
-insure forest recovery. Our results advocate for sustainable selective
-logging against monoculture stand: high diversity will increase forest
-resilience and thus improve logging cycles. Even if monoculture stand
-are naturalized they will not reach mature forest diversity, and
-consequently they will not reach its natural high resilience. But on the
-other hand selective logging in tropical forest needs to be sustainable,
-if not the diversity will slowly decrease after each cycles degrading
-forest ecosystem functions, and resilience will decrease due to negative
-feedbacks. \citet{Zimmerman2012} criticized the state of sylviculture in
-tropics, suggesting that ``we have not been able to reconcile these
-opposing biological and economic forces''. Still, they advocates for a
+resilience together with productivity \citep{Liang2016}. Additionally we
+found that complementarity between species was insuring forest recovery
+in forest succession start with facilitation, before more productive
+species dominate the forest and insure forest recovery. Our results
+advocate for sustainable selective logging against monoculture stand:
+high diversity will increase forest resilience and thus improve logging
+cycles. Even if monoculture stand are naturalized they will not reach
+mature forest diversity, and consequently they will not reach its
+natural high resilience. But on the other hand selective logging in
+tropical forest needs to be sustainable, if not the diversity will
+slowly decrease after each cycles degrading forest ecosystem functions,
+and resilience will decrease due to negative feedbacks.
+\citet{Zimmerman2012} criticized the state of sylviculture in tropics,
+suggesting that ``we have not been able to reconcile these opposing
+biological and economic forces''. Still, they advocates for a
 possibility of sustainable tropical logging, already existing in

### 2017-07-18 05-Results.Rmd
Sylviculture results

@@
 Most of functional traits had a significant long term influence on ecosystem outputs (figure \@ref(fig:ftVar)).
-Only specific maximum diameter **dmax** add higher diversity for greater orders implying better evenness in species distributions.
+Only specific maximum diameter dmax add higher diversity for greater orders implying better evenness in species distributions.
 Regarding functional composition, traits fixed to mean did not change other functional traits density distribution.
@@
   xlab('Time (years)') +
-  ylab('Distance to forest dynamic equilibrium\n(AGB, BA, N, N10 and N30)') +
+  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
   scale_color_discrete('Disturbance\nintensity\n(% of BA)')
@@
   xlab('Time (years)') +
-  ylab('Cummulative integral of\ndistance to forest dynamic equilibrium\n(AGB, BA, N, N10 and N30)') +
+  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
   scale_color_discrete('Disturbance\nintensity\n(% of BA)')
@@
 We compared cummulative integral after 600 years to communities taxonomic and functional diversity for each leavel of disturbance (see Figure \@ref(fig:gIeq)).
-We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years ($Ieq_{600}$).
+We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cummulative integral from ecosystem distance to forest structure equilibrium after 600 years ($Ieq_{600}$).
 In addition, functional evenness was complementary reducing $Ieq_{600}$.
@@
 Effectivelly, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$.
-We found similar results for all disturbance levels and other ecosystem functions (forest production and forest diversity, see [Appendix 5: Disturbance simulations]).
+We found similar results for all disturbance levels and forest functionning (see [Appendix 5: Disturbance simulations]).
 
 ```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
+Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
 Ieq %>% 
@@
   ylab('Cummulative integral from ecosystem 
-       distance to forest dynamic equilibrium after 600 years') +
+       distance to forest structure equilibrium after 600 years') +
   scale_shape_discrete('Species\nrichness\n(SR)') +
@@
 ```
+
+## Sylviculture
+
+```{r sylviculture, echo=FALSE, message=FALSE}
+library(dplyr)
+library(reshape2)
+library(ggplot2)
+path <- '~/Documents/ECOFOG/Results/disturbance/sylviculture/data/'
+load(file.path(path, 'DOE.Rdata'))
+load(file.path(path, 'resilience.Rdata'))
+load(file.path(path, 'BErecovery.Rdata'))
+deq <- data.frame(
+  time = resilience$time,
+  sim = resilience$sim,
+  structure = apply(resilience[c('agb', 'ba', 'n', 'n10', 'n30')], 1, 
+                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))),
+  production = apply(resilience[c('gpp', 'npp', 'Rday', 'Rnight')], 1, 
+                  function(x) sqrt(sum((rep(1, length(x)) - x)^2)))
+)
+sim <- unique(resilience$sim)
+Ieq <- sapply(sim, function(s){
+  data_s <- subset(deq, sim ==s)
+  return(cbind(data_s[c(1,2)], cumsum(data_s[-c(1,2)])))
+}, simplify = F) ; rm(sim)
+Ieq <- data.frame(do.call('rbind', Ieq))
+names(Ieq)[3:4] <- paste0('Ieq.', names(Ieq)[3:4])
+```
+
+We repeated design of experiments and simulations used for the disturbance simulations with the sylviculture module.
+But only 37 simulations included harvestable species.
+Moreover most of simulations included low harvestable volume resulting in a small disturbance from 0.5 to 2.0 $m^2.ha^{-1}$ (see Figure \@ref(fig:gIeqSylv)).
+Consequently most of disturbed volume was due to primary and secondary tracks and was not related to logged trees.
+Thus sylviculture module resulted in aggregated localized gaps but not in species selective disturbance.
+We found that, contrarily to previous results, functional richness was significantly ($p<0.01$) decreasing forest resilience see Figure \@ref(fig:gIeqSylv)).
+But the range of variation from resilience was really low ($2 < I_{eq_{600}} < 10$).
+And resilience was mainly due to the density of *Bocoa prouacensis* in the simulated mature forests.
+Effectively *Bocoa prouacensis* was often the main harvested species and possessed additionnally a quick recovery inside the model due to unrealisticaly advantageous parameters.
+Still we found that functional dispersion was decreasing resilience but not significantly (see [Appendix 6: Sylviculture simulations]).
+Finally, biodiversity net effect is almost undisturbed due to the low disturbance intensity from the sylviculture module and results did not show complementarity or selection effect in forest resilience (see [Appendix 6: Sylviculture simulations]).
+
+```{r gIeqSylv, echo=FALSE, warning=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity.
+Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years normalized by disturbed basal area was represented against functional richness [FRic, @villeger_new_2008].
+Dot color represents the species richness (nb) whereas dot size represents the disturbed basal area (m2/ha).
+Grey line represents the linear regression and grey area the confidence interval.', cache=TRUE}
+Ieq %>% 
+  left_join(., doe, by = 'sim') %>% 
+  filter(time == 600) %>% 
+  ggplot(aes(x = FRic, y = Ieq.structure/ba, label = sim)) + 
+  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
+  geom_point(aes(color = richness, size = ba/16)) +
+  xlab('Functional richness (FRic)') +
+  ylab('Cummulative integral from ecosystem 
+       distance to forest dynamic equilibrium after 600 years') +
+  scale_color_discrete('Species\nrichness\n(SR)') +
+  scale_size_continuous('Basla\narea\ndisturbed\n(m2/ha)')
+```

### 2017-07-18 06-Discussion.Rmd
Sylviculture results

@@
 In order to study the role of diversity in forest resilience we decided to remove the seed rain constant to get a closed system and look at the role of diversity when it maintains itself through feedbacks and not with immigration.
+
+Finally, sylviculture module implemented inside TROLL showed lacks of ability to reproduce selective logging with current design of experiments.
+Results did not seem to be usable and will not be discussed.
+The main issues was the ability of the model to simulate mature forest with correct abundancy of mature trees from commonly harvested species.
+Two solutions might be possible in the tunning of the model.
+First, forest inventories could be used to initialize the model instead of a bare soil resulting in realistic species abundances, especially for harvested species.
+Secondly, selective logging could focus on guilds of species meeting peculiar values of functional traits [as wood density, see @Huth2004; @Khler2004; @Ruger2008] allowing the model to harvest more volume in order to meet reality.
 

### 2017-07-18 11-Appendix5.Rmd
Thesis v1
Sylviculture results

@@
 
-Appendix 4 presents ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
+This appendix presents ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
 It encompass all functionnal diversity components [FRIC, FEve, FDiv, and FDis, @villeger_new_2008].
-And it presents results for both forest dynamic (Figure \@ref(fig:A5IeqStructureAll)) and forest production (Figure \@ref(fig:A5IeqProductionAll)).
+And it presents results for both forest structure (Figure \@ref(fig:A5IeqStructureAll)) and forest functionning (Figure \@ref(fig:A5IeqProductionAll)).
 
 ```{r A5IeqStructureAll, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
+Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
 library(dplyr)
@@
   xlab('Cummulative integral from ecosystem 
-       distance to forest dynamic equilibrium after 600 years') +
+       distance to forest structure equilibrium after 600 years') +
   scale_shape_discrete('Species\nrichness\n(SR)')
@@
 ```{r A5IeqProductionAll, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest production equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
+Cummulative integral from ecosystem distance to forest functionning equilibrium after 600 years was represented against functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness.', cache=TRUE, fig.height=8, fig.width=12}
 Ieq %>% 
@@
   xlab('Cummulative integral from ecosystem 
-       distance to forest production equilibrium after 600 years') +
+       distance to forest functionning equilibrium after 600 years') +
   scale_shape_discrete('Species\nrichness\n(SR)')

### 2017-07-18 12-Appendix6.Rmd
Sylviculture results

# Appendix 6: Sylviculture simulations

```{r A6data, echo=FALSE, cache=TRUE}
library(dplyr)
library(reshape2)
library(ggplot2)
path <- '~/Documents/ECOFOG/Results/disturbance/sylviculture/data/'
load(file.path(path, 'DOE.Rdata'))
load(file.path(path, 'resilience.Rdata'))
deq <- data.frame(
  time = resilience$time,
  sim = resilience$sim,
  structure = apply(resilience[c('agb', 'ba', 'n', 'n10', 'n30')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))),
  production = apply(resilience[c('gpp', 'npp', 'Rday', 'Rnight')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2)))
)
sim <- unique(resilience$sim)
Ieq <- sapply(sim, function(s){
  data_s <- subset(deq, sim ==s)
  return(cbind(data_s[c(1,2)], cumsum(data_s[-c(1,2)])))
}, simplify = F) ; rm(sim)
Ieq <- data.frame(do.call('rbind', Ieq))
names(Ieq)[3:4] <- paste0('Ieq.', names(Ieq)[3:4])
```

## Ecosystem functions

This appendix presents ecosystem resilience after 600 years with taxonomic and functional diversity after selective logging. It encompass all functionnal diversity components [FRIC, FEve, FDiv, and FDis, @villeger_new_2008]. And it presents results for both forest structure (Figure \@ref(fig:A6IeqStructureAll)) and forest functionning (Figure \@ref(fig:A6IeqProductionAll)).

```{r A6IeqStructureAll, echo=FALSE, warning=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity. Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years normalized by disturbed basal area was represented against functional functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008]. Dot color represents the species richness (nb) whereas dot size represents the disturbed basal area (m2/ha). Grey line represents the linear regression and grey area the confidence interval.', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>%
  select(Ieq.structure, ba, richness, FRic, FEve, FDiv, FDis) %>% 
  melt(id.vars = c('Ieq.structure', 'ba', 'richness')) %>% 
  ggplot(aes(x = Ieq.structure/ba, y = value)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(color = richness, size = ba/16)) +
  facet_wrap(~variable, scales = "free", labeller = 'label_both') + 
  ylab('Functional diversity') +
  xlab('Cummulative integral from ecosystem 
       distance to forest structure equilibrium after 600 years') +
  scale_color_discrete('Species\nrichness\n(SR)') +
  scale_size_continuous('Basla\narea\ndisturbed\n(m2/ha)')

```

```{r A6IeqProductionAll, echo=FALSE, warning=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity. Cummulative integral from ecosystem distance to forest functionning equilibrium after 600 years normalized by disturbed basal area was represented against functional functional diversity [FRIC, FEve, FDiv, and FDis, @villeger_new_2008]. Dot color represents the species richness (nb) whereas dot size represents the disturbed basal area (m2/ha). Grey line represents the linear regression and grey area the confidence interval.', cache=TRUE, fig.height=8, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>%
  select(Ieq.production, ba, richness, FRic, FEve, FDiv, FDis) %>% 
  melt(id.vars = c('Ieq.production', 'ba', 'richness')) %>% 
  ggplot(aes(x = Ieq.production/ba, y = value)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(color = richness, size = ba/16)) +
  facet_wrap(~variable, scales = "free", labeller = 'label_both') + 
  ylab('Functional diversity') +
  xlab('Cummulative integral from ecosystem 
       distance to forest functionning equilibrium after 600 years') +
  scale_color_discrete('Species\nrichness\n(SR)') +
  scale_size_continuous('Basla\narea\ndisturbed\n(m2/ha)')
```

## Biodiversity effect

Figure \@ref(fig:A6allBE) presents the resilience of complementarity and selection effects for different ecosystem metrics (AGB, BA, N, GPP and NPP).

```{r A6allBE, echo=FALSE, warning=FALSE, fig.cap='Resilience of complementarity and selection effects. Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time for different ecosystem variables (AGB, BA, N, GPP).', cache=TRUE, fig.height=8, fig.width=12}
load(file.path(path, 'BErecovery.Rdata'))
BEresilience %>% 
  select(BE, time, sim, agb, ba, n, gpp) %>% 
  melt(id.vars = c('BE', 'time', 'sim')) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(BE %in% c('SE', 'CE')) %>% 
  group_by(time, BE, variable) %>%
  summarise(qt1 =quantile(value, probs=0.05, na.rm = T),
            qt2 =quantile(value, probs=0.75, na.rm = T),
            mean=mean(value, na.rm = T)) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_wrap(~ variable, scales = "free", labeller = 'label_both') +
  xlab('Time (years)') +
  ylab('Recovery of variable net effect') + 
  ylim(-2.5,2.5)
```

### 2017-07-18 docs/master-thesis.tex
Sylviculture results

@@
 ecosystem outputs (figure \ref{fig:ftVar}). Only specific maximum
-diameter \textbf{dmax} add higher diversity for greater orders implying
-better evenness in species distributions. Regarding functional
-composition, traits fixed to mean did not change other functional traits
-density distribution. Moreover, seedrain did not seem to affect
-aboveground biomass and final ecosystem height and diameter structure.
-Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a
-similar effect. Lower seedrain implied faster decrease of stem above 10
-cm dbh and higher number of stem above 30 cm diameter at breast height
-after ecosystem resilience to disturbance (approximately 50 years).
-Lower seedrain than default decreased basal area over time. In addition,
-lower seedrain than default decreased equitability by increasing
-abundance of abundant species and decreasing abundance of less abundant
-species. See
+diameter dmax add higher diversity for greater orders implying better
+evenness in species distributions. Regarding functional composition,
+traits fixed to mean did not change other functional traits density
+distribution. Moreover, seedrain did not seem to affect aboveground
+biomass and final ecosystem height and diameter structure. Seedrain
+constant fixed to 2 or 20 seed per hectare seemed to have a similar
+effect. Lower seedrain implied faster decrease of stem above 10 cm dbh
+and higher number of stem above 30 cm diameter at breast height after
+ecosystem resilience to disturbance (approximately 50 years). Lower
+seedrain than default decreased basal area over time. In addition, lower
+seedrain than default decreased equitability by increasing abundance of
+abundant species and decreasing abundance of less abundant species. See
 \protect\hyperlink{appendix-4-sensitivity-analysis}{Appendix 4:
@@
 functional diversity \citep[FDiv,][]{villeger_new_2008} was reducing
-cummulative integral from ecosystem distance to forest dynamic
+cummulative integral from ecosystem distance to forest structure
 equilibrium after 600 years (\(Ieq_{600}\)). In addition, functional
@@
 consequently lower \(Ieq_{600}\). We found similar results for all
-disturbance levels and other ecosystem functions (forest production and
-forest diversity, see
+disturbance levels and forest functionning (see
 \protect\hyperlink{appendix-5-disturbance-simulations}{Appendix 5:
@@
 and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest dynamic
+Cummulative integral from ecosystem distance to forest structure
 equilibrium after 600 years was represented against functional diversity
@@
 
+\subsection{Sylviculture}\label{sylviculture-1}
+
+We repeated design of experiments and simulations used for the
+disturbance simulations with the sylviculture module. But only 37
+simulations included harvestable species. Moreover most of simulations
+included low harvestable volume resulting in a small disturbance from
+0.5 to 2.0 \(m^2.ha^{-1}\) (see Figure \ref{fig:gIeqSylv}). Consequently
+most of disturbed volume was due to primary and secondary tracks and was
+not related to logged trees. Thus sylviculture module resulted in
+aggregated localized gaps but not in species selective disturbance. We
+found that, contrarily to previous results, functional richness was
+significantly (\(p<0.01\)) decreasing forest resilience see Figure
+\ref{fig:gIeqSylv}). But the range of variation from resilience was
+really low (\(2 < I_{eq_{600}} < 10\)). And resilience was mainly due to
+the density of \emph{Bocoa prouacensis} in the simulated mature forests.
+Effectively \emph{Bocoa prouacensis} was often the main harvested
+species and possessed additionnally a quick recovery inside the model
+due to unrealisticaly advantageous parameters. Still we found that
+functional dispersion was decreasing resilience but not significantly
+(see \protect\hyperlink{appendix-6-sylviculture-simulations}{Appendix 6:
+Sylviculture simulations}). Finally, biodiversity net effect is almost
+undisturbed due to the low disturbance intensity from the sylviculture
+module and results did not show complementarity or selection effect in
+forest resilience (see
+\protect\hyperlink{appendix-6-sylviculture-simulations}{Appendix 6:
+Sylviculture simulations}).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/gIeqSylv-1.pdf}
+\caption{\label{fig:gIeqSylv}Ecosystem resilience after 600 years with
+taxonomic and functional diversity. Cummulative integral from ecosystem
+distance to forest structure equilibrium after 600 years normalized by
+disturbed basal area was represented against functional richness
+\citep[FRic,][]{villeger_new_2008}. Dot color represents the species
+richness (nb) whereas dot size represents the disturbed basal area
+(m2/ha). Grey line represents the linear regression and grey area the
+confidence interval.}
+\end{figure}
+
 \section{Discussion}\label{discussion}
@@
 when it maintains itself through feedbacks and not with immigration.
+
+Finally, sylviculture module implemented inside TROLL showed lacks of
+ability to reproduce selective logging with current design of
+experiments. Results did not seem to be usable and will not be
+discussed. The main issues was the ability of the model to simulate
+mature forest with correct abundancy of mature trees from commonly
+harvested species. Two solutions might be possible in the tunning of the
+model. First, forest inventories could be used to initialize the model
+instead of a bare soil resulting in realistic species abundances,
+especially for harvested species. Secondly, selective logging could
+focus on guilds of species meeting peculiar values of functional traits
+\citep[as wood density, see][]{Huth2004, Khler2004, Ruger2008} allowing
+the model to harvest more volume in order to meet reality.
 
@@
 FDiv, and FDis,][]{villeger_new_2008}. And it presents results for both
-forest dynamic (Figure \ref{fig:A5IeqStructureAll}) and forest
-production (Figure \ref{fig:A5IeqProductionAll}).
+forest structure (Figure \ref{fig:A5IeqStructureAll}) and forest
+functionning (Figure \ref{fig:A5IeqProductionAll}).
 
@@
 disturbance. Cummulative integral from ecosystem distance to forest
-dynamic equilibrium after 600 years was represented against functional
+structure equilibrium after 600 years was represented against functional
 diversity \citep[FRIC, FEve, FDiv, and FDis,][]{villeger_new_2008} for
@@
 disturbance. Cummulative integral from ecosystem distance to forest
-production equilibrium after 600 years was represented against
+functionning equilibrium after 600 years was represented against
 functional diversity \citep[FRIC, FEve, FDiv, and
@@
 \caption{\label{fig:A5allBE}Resilience of complementarity and selection
+effects. Complementarity effect (CE) and selection effect (SE) where
+normalized by control net effect (NEc), thus measuring their resilience
+over time for different ecosystem variables (AGB, BA, N, GPP).}
+\end{figure}
+
+\hypertarget{appendix-6-sylviculture-simulations}{\section{Appendix 6:
+Sylviculture simulations}\label{appendix-6-sylviculture-simulations}}
+
+\subsection{Ecosystem functions}\label{ecosystem-functions-3}
+
+This appendix presents ecosystem resilience after 600 years with
+taxonomic and functional diversity after selective logging. It encompass
+all functionnal diversity components \citep[FRIC, FEve, FDiv, and
+FDis,][]{villeger_new_2008}. And it presents results for both forest
+structure (Figure \ref{fig:A6IeqStructureAll}) and forest functionning
+(Figure \ref{fig:A6IeqProductionAll}).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A6IeqStructureAll-1.pdf}
+\caption{\label{fig:A6IeqStructureAll}Ecosystem resilience after 600 years
+with taxonomic and functional diversity. Cummulative integral from
+ecosystem distance to forest structure equilibrium after 600 years
+normalized by disturbed basal area was represented against functional
+functional diversity \citep[FRIC, FEve, FDiv, and
+FDis,][]{villeger_new_2008}. Dot color represents the species richness
+(nb) whereas dot size represents the disturbed basal area (m2/ha). Grey
+line represents the linear regression and grey area the confidence
+interval.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A6IeqProductionAll-1.pdf}
+\caption{\label{fig:A6IeqProductionAll}Ecosystem resilience after 600 years
+with taxonomic and functional diversity. Cummulative integral from
+ecosystem distance to forest functionning equilibrium after 600 years
+normalized by disturbed basal area was represented against functional
+functional diversity \citep[FRIC, FEve, FDiv, and
+FDis,][]{villeger_new_2008}. Dot color represents the species richness
+(nb) whereas dot size represents the disturbed basal area (m2/ha). Grey
+line represents the linear regression and grey area the confidence
+interval.}
+\end{figure}
+
+\subsection{Biodiversity effect}\label{biodiversity-effect-3}
+
+Figure \ref{fig:A6allBE} presents the resilience of complementarity and
+selection effects for different ecosystem metrics (AGB, BA, N, GPP and
+NPP).
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{master-thesis_files/figure-latex/A6allBE-1.pdf}
+\caption{\label{fig:A6allBE}Resilience of complementarity and selection
 effects. Complementarity effect (CE) and selection effect (SE) where

### 2017-07-19 00-Abstract.Rmd
Abstract 1

@@
 
-Écrire le résumé français ici...
+Les forêts tropicales font face à de nombreuses perturbations qui représentent la troisième source mondiale d'émission de gaz à effet de serre.
+La déforestation et la dégradation des forêts tropicales sont responsables de l'émission de 8.26 milliards de tonnes de dioxyde de carbone par an [@Pearson2017].
+La déforestation a retenu l'attention mondiale, mais la dégradation des forêts représente 20% des émissions de l'Amazonie brésilienne [@Asner2005].
+La gestion durable des forêts a été proposée comme réponse à la déforestation et la dégradation, malgré la remise en question de la durabilité de l'exploitation forestière [@Zimmerman2012].
+D'autre part, les forêts tropicales abritent plus de la moitié de la biodiversité terrestre mondiale [@Scheffers2012].
+Par conséquent, nous avons décidé d'étudier le rôle de la biodiversité dans la réponse des écosystèmes forestiers aux perturbations, en reliant diversité et fonctionnement de l'écosystème [@Loreau2010].
+Nous avons utilisé l'hypothèse que lors d'une perturbation, grâce à une productivité plus forte, une forêt plus diverse aura une meilleure résilience, en se basant sur la relation positive entre biodiversité et productivité.
+Nous avons relié cette hypothèse aux effets de complémentarité et de sélection [@Loreau2001].
+La complémentarité est la combinaison de la partition des ressources et de la facilitation, alors que l'effet de sélection est le résultat de la sélection compétitive.
+Nous avons ainsi centré l'étude sur les mécanismes impliqués dans la relation entre biodiversité et résilience des écosystème forestiers par une approche par simulation afin d'appréhender les processus à long terme.
+Nous avons utilisé le modèle TROLL [@Li] pour simuler 60 forêts matures aux diversités taxonomiques et fonctionnelles croissantes.
+Nous avons perturbé toutes les forêts et mesuré la résilience de leurs fonctions écosystémiques.
+En outre, nous avons mesuré la résilience de l'effet net de la biodiversité que l'on a décomposé en en effets de complémentarité et de sélection.
+Nous avons trouvé que la diversité améliore la résilience des forêts tropicales, particulièrement au travers de la diversité et l'équitabilité fonctionnelle.
+De plus, nous avons montré que la complémentarité entre les espèces assurait la résilience de la forêt en début de succession avant de laisser place à l'effet de sélection.
+Nos résultats suggèrent la possibilité d'une gestion durable des forêts tropicales grâce à une meilleure résilience avec une plus haute diversité.
+Mais cette conclusion n'a de sens que si l'exploitation sélective est durable [@Zimmerman2012].
+Au contraire, une gestion non durable des forêts tropicales entraînera des rétroactions négatives diminuant lentement la diversité et donc la résilience des forêts, aboutissant ultimement à la dégradation des forêts.
 
-Write the english abstract here...
+Forest disturbances are the third worldwide source of greenhous gas.
+Tropical deforestation and degradation emit 8.26 billion of tons of carbon dioxyde per year [@Pearson2017].
+Deforestation has retained much attention, but degradation from forest represents 20% of emissions in brazilian Amazon [@Asner2005].
+Sustainable forest management has been promoted as an answer to deforestation and degradation, besides logging sustainability has been questionned [@Zimmerman2012].
+On the other hand, tropical forest host over half of the Earth's biodiversity [@Scheffers2012].
+Consequently, we decided to study the role of biodiversity in forest ecosystem answer to disturbance, linking diversity to ecosystem functionning [@Loreau2010].
+We used the hypothesis that when a disturbance event happen, due to a higher productivity, a more diverse forest will be more resilient, based on the positive relationship between biodiversity and productivity.
+We linked that hypothesis to the complementarity and selection effects [@Loreau2001].
+Complementarity is the addition of ressource partitionning and facilitation, whereas selection effect is the result of competitive selection.
+We thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience with a simulation approach to assess long term processes.
+We used TROLL model [@Li] to simulate 60 matures forests with growing taxonomic and functional  diversities.
+We disturbed all forests and measured the resilience of their  ecosystem functions.
+Additionnally, we measured biodiversity net effect resilience partitioned into complementarity and selection effects.
+We found that diversity improved tropical forest resilience, particularly through functional diversity and eveness.
+Moreover, we showed that complementarity between species insured forest recovery in the beginning of the succession before being replaced by selection effect.
+Our results suggest the possibility for a sustainable management of tropical forest due to an increased resilience with an higher diversity.
+But this conclusion has meaning only if selective logging meet sustainability [@Zimmerman2012].
+On the contrary, unsustainable tropical forest management will lead to negative feedbacks slowly diminishing diversity and thus forest resilience, resulting ultimately in forest degradation.

### 2017-07-19 02-Introduction.Rmd
Abstract 1

@@
 
-Tropical forests perturbations, through deforestation and degradation, account for 8.26 billion of carbon dioxyde emissions per year [@Pearson2017].
+Tropical forests disturbances, through deforestation and degradation, account for 8.26 billion of tons of carbon dioxyde emissions per year [@Pearson2017].
 It represents the third source of greenhouse gas.
@@
 
-We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance highlighting tropical forest carbone resilience to disturbance.
+We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance.
 Resilience encompass several definitions but was summarized by @Oliver2015 as the degree of resistance or fast recovery from an ecosystem function to environmental disturbance.

### 2017-07-19 06-Discussion.Rmd
Abstract 1

@@
 But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012.
-Because if the harvesting is not sustainable negative feedbaks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.
+Because if the harvesting is not sustainable negative feedbacks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.
 

### 2017-07-19 docs/master-thesis.tex
Abstract 1

@@
 
-Écrire le résumé français ici\ldots{}
-
-Write the english abstract here\ldots{}
+Les forêts tropicales font face à de nombreuses perturbations qui
+représentent la troisième source mondiale d'émission de gaz à effet de
+serre. La déforestation et la dégradation des forêts tropicales sont
+responsables de l'émission de 8.26 milliards de tonnes de dioxyde de
+carbone par an \citep{Pearson2017}. La déforestation a retenu
+l'attention mondiale, mais la dégradation des forêts représente 20\% des
+émissions de l'Amazonie brésilienne \citep{Asner2005}. La gestion
+durable des forêts a été proposée comme réponse à la déforestation et la
+dégradation, malgré la remise en question de la durabilité de
+l'exploitation forestière \citep{Zimmerman2012}. D'autre part, les
+forêts tropicales abritent plus de la moitié de la biodiversité
+terrestre mondiale \citep{Scheffers2012}. Par conséquent, nous avons
+décidé d'étudier le rôle de la biodiversité dans la réponse des
+écosystèmes forestiers aux perturbations, en reliant diversité et
+fonctionnement de l'écosystème \citep{Loreau2010}. Nous avons utilisé
+l'hypothèse que lors d'une perturbation, grâce à une productivité plus
+forte, une forêt plus diverse aura une meilleure résilience, en se
+basant sur la relation positive entre biodiversité et productivité. Nous
+avons relié cette hypothèse aux effets de complémentarité et de
+sélection \citep{Loreau2001}. La complémentarité est la combinaison de
+la partition des ressources et de la facilitation, alors que l'effet de
+sélection est le résultat de la sélection compétitive. Nous avons ainsi
+centré l'étude sur les mécanismes impliqués dans la relation entre
+biodiversité et résilience des écosystème forestiers par une approche
+par simulation afin d'appréhender les processus à long terme. Nous avons
+utilisé le modèle TROLL \citep{Li} pour simuler 60 forêts matures aux
+diversités taxonomiques et fonctionnelles croissantes. Nous avons
+perturbé toutes les forêts et mesuré la résilience de leurs fonctions
+écosystémiques. En outre, nous avons mesuré la résilience de l'effet net
+de la biodiversité que l'on a décomposé en en effets de complémentarité
+et de sélection. Nous avons trouvé que la diversité améliore la
+résilience des forêts tropicales, particulièrement au travers de la
+diversité et l'équitabilité fonctionnelle. De plus, nous avons montré
+que la complémentarité entre les espèces assurait la résilience de la
+forêt en début de succession avant de laisser place à l'effet de
+sélection. Nos résultats suggèrent la possibilité d'une gestion durable
+des forêts tropicales grâce à une meilleure résilience avec une plus
+haute diversité. Mais cette conclusion n'a de sens que si l'exploitation
+sélective est durable \citep{Zimmerman2012}. Au contraire, une gestion
+non durable des forêts tropicales entraînera des rétroactions négatives
+diminuant lentement la diversité et donc la résilience des forêts,
+aboutissant ultimement à la dégradation des forêts.
+
+Forest disturbances are the third worldwide source of greenhous gas.
+Tropical deforestation and degradation emit 8.26 billion of tons of
+carbon dioxyde per year \citep{Pearson2017}. Deforestation has retained
+much attention, but degradation from forest represents 20\% of emissions
+in brazilian Amazon \citep{Asner2005}. Sustainable forest management has
+been promoted as an answer to deforestation and degradation, besides
+logging sustainability has been questionned \citep{Zimmerman2012}. On
+the other hand, tropical forest host over half of the Earth's
+biodiversity \citep{Scheffers2012}. Consequently, we decided to study
+the role of biodiversity in forest ecosystem answer to disturbance,
+linking diversity to ecosystem functionning \citep{Loreau2010}. We used
+the hypothesis that when a disturbance event happen, due to a higher
+productivity, a more diverse forest will be more resilient, based on the
+positive relationship between biodiversity and productivity. We linked
+that hypothesis to the complementarity and selection effects
+\citep{Loreau2001}. Complementarity is the addition of ressource
+partitionning and facilitation, whereas selection effect is the result
+of competitive selection. We thus focused on mechanisms involved in the
+relationship between biodiversity and forest ecosystem resilience with a
+simulation approach to assess long term processes. We used TROLL model
+\citep{Li} to simulate 60 matures forests with growing taxonomic and
+functional diversities. We disturbed all forests and measured the
+resilience of their ecosystem functions. Additionnally, we measured
+biodiversity net effect resilience partitioned into complementarity and
+selection effects. We found that diversity improved tropical forest
+resilience, particularly through functional diversity and eveness.
+Moreover, we showed that complementarity between species insured forest
+recovery in the beginning of the succession before being replaced by
+selection effect. Our results suggest the possibility for a sustainable
+management of tropical forest due to an increased resilience with an
+higher diversity. But this conclusion has meaning only if selective
+logging meet sustainability \citep{Zimmerman2012}. On the contrary,
+unsustainable tropical forest management will lead to negative feedbacks
+slowly diminishing diversity and thus forest resilience, resulting
+ultimately in forest degradation.
 
@@
 \item
-  Laurent
+  Camilla
 \item
@@
 \item
+  Laurent
+\item
   Jérôme
-\item
-  Camilla
 \item
@@
 \item
+  AMAP colleagues
+\item
+  Myriam
+\item
+  BIOGECO colleagues
+\item
+  Pascal Pe
+\item
   Paracou team
 \item
-  \ldots{}
+  Pascal Pa
+\item
+  Aurélie C
 \end{itemize}
@@
 
-Tropical forests perturbations, through deforestation and degradation,
-account for 8.26 billion of carbon dioxyde emissions per year
+Tropical forests disturbances, through deforestation and degradation,
+account for 8.26 billion of tons of carbon dioxyde emissions per year
 \citep{Pearson2017}. It represents the third source of greenhouse gas.
@@
 We decided to use the forest model TROLL to study the role of
-biodiversity in forest ecosystem answer to disturbance highlighting
-tropical forest carbone resilience to disturbance. Resilience encompass
-several definitions but was summarized by \citet{Oliver2015} as the
-degree of resistance or fast recovery from an ecosystem function to
-environmental disturbance.
+biodiversity in forest ecosystem answer to disturbance. Resilience
+encompass several definitions but was summarized by \citet{Oliver2015}
+as the degree of resistance or fast recovery from an ecosystem function
+to environmental disturbance.
 
@@
 following \citet{Zimmerman2012}. Because if the harvesting is not
-sustainable negative feedbaks will slowly diminish diversity and its
+sustainable negative feedbacks will slowly diminish diversity and its
 benefits for forest resilience, resulting in forest degradation.
@@
   \newpage
+  \scriptsize{
   \paragraph{Résumé :}
-  Écrire le résumé ici...
-  \paragraph{Mots clés :} mots clés
-  \newline\newline
+  Les forêts tropicales font face à de nombreuses perturbations qui représentent la troisième source mondiale d’émission de gaz à effet de serre.
+La déforestation et la dégradation des forêts tropicales sont responsables de l’émission de 8.26 milliards de tonnes de dioxyde de carbone par an (Pearson et al.
+2017).
+La déforestation a retenu l’attention mondiale, mais la dégradation des forêts représente 20\% des émissions de l’Amazonie brésilienne (Asner et al.
+2005).
+La gestion durable des forêts a été proposée comme réponse à la déforestation et la dégradation, malgré la remise en question de la durabilité de l’exploitation forestière (Zimmerman \& Kormos 2012).
+D’autre part, les forêts tropicales abritent plus de la moitié de la biodiversité terrestre mondiale (Scheffers et al.
+2012).
+Par conséquent, nous avons décidé d’étudier le rôle de la biodiversité dans la réponse des écosystèmes forestiers aux perturbations, en reliant diversité et fonctionnement de l’écosystème (Loreau 2010).
+Nous avons utilisé l’hypothèse que lors d’une perturbation, grâce à une productivité plus forte, une forêt plus diverse aura une meilleure résilience, en se basant sur la relation positive entre biodiversité et productivité.
+Nous avons relié cette hypothèse aux effets de complémentarité et de sélection (Loreau \& Hector 2001b).
+La complémentarité est la combinaison de la partition des ressources et de la facilitation, alors que l’effet de sélection est le résultat de la sélection compétitive.
+Nous avons ainsi centré l’étude sur les mécanismes impliqués dans la relation entre biodiversité et résilience des écosystème forestiers par une approche par simulation afin d’appréhender les processus à long terme.
+Nous avons utilisé le modèle TROLL (Maréchaux \& Chave) pour simuler 60 forêts matures aux diversités taxonomiques et fonctionnelles croissantes.
+Nous avons perturbé toutes les forêts et mesuré la résilience de leurs fonctions écosystémiques.
+En outre, nous avons mesuré la résilience de l’effet net de la biodiversité que l’on a décomposé en en effets de complémentarité et de sélection.
+Nous avons trouvé que la diversité améliore la résilience des forêts tropicales, particulièrement au travers de la diversité et l’équitabilité fonctionnelle.
+De plus, nous avons montré que la complémentarité entre les espèces assurait la résilience de la forêt en début de succession avant de laisser place à l’effet de sélection.
+Nos résultats suggèrent la possibilité d’une gestion durable des forêts tropicales grâce à une meilleure résilience avec une plus haute diversité.
+Mais cette conclusion n’a de sens que si l’exploitation sélective est durable (Zimmerman \& Kormos 2012).
+Au contraire, une gestion non durable des forêts tropicales entraînera des rétroactions négatives diminuant lentement la diversité et donc la résilience des forêts, aboutissant ultimement à la dégradation des forêts.
+  \paragraph{Mots clés :} Résilience, Biodiveristé, Exploitation sélective, Fonctionnement de l'écosystème
   \paragraph{Abstract:}
-  Write abstract here
-  \paragraph{Keywords:} keywords
+  Forest disturbances are the third worldwide source of greenhous gas.
+Tropical deforestation and degradation emit 8.26 billion of tons of carbon dioxyde per year (Pearson et al.
+2017).
+Deforestation has retained much attention, but degradation from forest represents 20\% of emissions in brazilian Amazon (Asner et al.
+2005).
+Sustainable forest management has been promoted as an answer to deforestation and degradation, besides logging sustainability has been questionned (Zimmerman \& Kormos 2012).
+On the other hand, tropical forest host over half of the Earth’s biodiversity (Scheffers et al.
+2012).
+Consequently, we decided to study the role of biodiversity in forest ecosystem answer to disturbance, linking diversity to ecosystem functionning (Loreau 2010).
+We used the hypothesis that when a disturbance event happen, due to a higher productivity, a more diverse forest will be more resilient, based on the positive relationship between biodiversity and productivity.
+We linked that hypothesis to the complementarity and selection effects (Loreau \& Hector 2001b).
+Complementarity is the addition of ressource partitionning and facilitation, whereas selection effect is the result of competitive selection.
+We thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience with a simulation approach to assess long term processes.
+We used TROLL model (Maréchaux \& Chave) to simulate 60 matures forests with growing taxonomic and functional diversities.
+We disturbed all forests and measured the resilience of their ecosystem functions.
+Additionnally, we measured biodiversity net effect resilience partitioned into complementarity and selection effects.
+We found that diversity improved tropical forest resilience, particularly through functional diversity and eveness.
+Moreover, we showed that complementarity between species insured forest recovery in the beginning of the succession before being replaced by selection effect.
+Our results suggest the possibility for a sustainable management of tropical forest due to an increased resilience with an higher diversity.
+But this conclusion has meaning only if selective logging meet sustainability (Zimmerman \& Kormos 2012).
+On the contrary, unsustainable tropical forest management will lead to negative feedbacks slowly diminishing diversity and thus forest resilience, resulting ultimately in forest degradation.
+  \paragraph{Keywords:}Resilience, Biodiversity, Selective logging, Ecosystem functionning
+  }
 

### 2017-07-19 latex/after_body.tex
Abstract 1

@@
 \newpage
+\scriptsize{
 \paragraph{Résumé :}
-Écrire le résumé ici...
-\paragraph{Mots clés :} mots clés
-\newline\newline
+Les forêts tropicales font face à de nombreuses perturbations qui représentent la troisième source mondiale d’émission de gaz à effet de serre.
+La déforestation et la dégradation des forêts tropicales sont responsables de l’émission de 8.26 milliards de tonnes de dioxyde de carbone par an (Pearson et al.
+2017).
+La déforestation a retenu l’attention mondiale, mais la dégradation des forêts représente 20\% des émissions de l’Amazonie brésilienne (Asner et al.
+2005).
+La gestion durable des forêts a été proposée comme réponse à la déforestation et la dégradation, malgré la remise en question de la durabilité de l’exploitation forestière (Zimmerman \& Kormos 2012).
+D’autre part, les forêts tropicales abritent plus de la moitié de la biodiversité terrestre mondiale (Scheffers et al.
+2012).
+Par conséquent, nous avons décidé d’étudier le rôle de la biodiversité dans la réponse des écosystèmes forestiers aux perturbations, en reliant diversité et fonctionnement de l’écosystème (Loreau 2010).
+Nous avons utilisé l’hypothèse que lors d’une perturbation, grâce à une productivité plus forte, une forêt plus diverse aura une meilleure résilience, en se basant sur la relation positive entre biodiversité et productivité.
+Nous avons relié cette hypothèse aux effets de complémentarité et de sélection (Loreau \& Hector 2001b).
+La complémentarité est la combinaison de la partition des ressources et de la facilitation, alors que l’effet de sélection est le résultat de la sélection compétitive.
+Nous avons ainsi centré l’étude sur les mécanismes impliqués dans la relation entre biodiversité et résilience des écosystème forestiers par une approche par simulation afin d’appréhender les processus à long terme.
+Nous avons utilisé le modèle TROLL (Maréchaux \& Chave) pour simuler 60 forêts matures aux diversités taxonomiques et fonctionnelles croissantes.
+Nous avons perturbé toutes les forêts et mesuré la résilience de leurs fonctions écosystémiques.
+En outre, nous avons mesuré la résilience de l’effet net de la biodiversité que l’on a décomposé en en effets de complémentarité et de sélection.
+Nous avons trouvé que la diversité améliore la résilience des forêts tropicales, particulièrement au travers de la diversité et l’équitabilité fonctionnelle.
+De plus, nous avons montré que la complémentarité entre les espèces assurait la résilience de la forêt en début de succession avant de laisser place à l’effet de sélection.
+Nos résultats suggèrent la possibilité d’une gestion durable des forêts tropicales grâce à une meilleure résilience avec une plus haute diversité.
+Mais cette conclusion n’a de sens que si l’exploitation sélective est durable (Zimmerman \& Kormos 2012).
+Au contraire, une gestion non durable des forêts tropicales entraînera des rétroactions négatives diminuant lentement la diversité et donc la résilience des forêts, aboutissant ultimement à la dégradation des forêts.
+\paragraph{Mots clés :} Résilience, Biodiveristé, Exploitation sélective, Fonctionnement de l'écosystème
 \paragraph{Abstract:}
-Write abstract here
-\paragraph{Keywords:} keywords
+Forest disturbances are the third worldwide source of greenhous gas.
+Tropical deforestation and degradation emit 8.26 billion of tons of carbon dioxyde per year (Pearson et al.
+2017).
+Deforestation has retained much attention, but degradation from forest represents 20\% of emissions in brazilian Amazon (Asner et al.
+2005).
+Sustainable forest management has been promoted as an answer to deforestation and degradation, besides logging sustainability has been questionned (Zimmerman \& Kormos 2012).
+On the other hand, tropical forest host over half of the Earth’s biodiversity (Scheffers et al.
+2012).
+Consequently, we decided to study the role of biodiversity in forest ecosystem answer to disturbance, linking diversity to ecosystem functionning (Loreau 2010).
+We used the hypothesis that when a disturbance event happen, due to a higher productivity, a more diverse forest will be more resilient, based on the positive relationship between biodiversity and productivity.
+We linked that hypothesis to the complementarity and selection effects (Loreau \& Hector 2001b).
+Complementarity is the addition of ressource partitionning and facilitation, whereas selection effect is the result of competitive selection.
+We thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience with a simulation approach to assess long term processes.
+We used TROLL model (Maréchaux \& Chave) to simulate 60 matures forests with growing taxonomic and functional diversities.
+We disturbed all forests and measured the resilience of their ecosystem functions.
+Additionnally, we measured biodiversity net effect resilience partitioned into complementarity and selection effects.
+We found that diversity improved tropical forest resilience, particularly through functional diversity and eveness.
+Moreover, we showed that complementarity between species insured forest recovery in the beginning of the succession before being replaced by selection effect.
+Our results suggest the possibility for a sustainable management of tropical forest due to an increased resilience with an higher diversity.
+But this conclusion has meaning only if selective logging meet sustainability (Zimmerman \& Kormos 2012).
+On the contrary, unsustainable tropical forest management will lead to negative feedbacks slowly diminishing diversity and thus forest resilience, resulting ultimately in forest degradation.
+\paragraph{Keywords:}Resilience, Biodiversity, Selective logging, Ecosystem functionning
+}
 

### 2017-07-20 01-Acknowledgments.Rmd
Thesis v1
Abstract 1
Akcnowledgments

@@
 
-I would like to thank...
-
-- Bruno
-- Stéphane
-- Laurent
-- Fabian
-- Isabelle
-- Jérôme
-- Camilla
-- Aurélie
-- Éric
-- ...
+I would like to thank Bruno Hérault and Stéphane Traissac for their guidance and advice during the whole internship.
+I congratulate Bruno to have been able to communicate and be reactive besides his numerous recent travels.
+I am grateful to Bruno and Camille Piponiot to have introduced me to the wonderful world of Bayesian statistics.
+All this work would not have been possible without TROLL model and the whole team that works on it : Jérôme Chave, Isabelle Maréchaux and Fabian Fischer.
+I thank all of the team to have introduced me to the model and guide me in its development from the opposite side of the Atlantic.
+I would also like to thank Laurent Descroix and people from National Forest Office who introduced me to sylviculture from french Guiana and helped me to build a realistic forest model.
+I also thank Aurélie Dourdain for her help to access Paracou data, Éric Marcon for his advice on biodiversity analysis, and Pascal Padolus for his patience and help wiht Tabebuia cluster.
+The study would not have been possible without Pascal Petronelli and all people who have contributed to Paracou and Guyafor data collection over the years.
+Finally I am grateful to Maxime Réjou-Méchain, Raphaël Pélissier, and AMAP colleagues, in addition to Myriam Heuertz and BIOGECO colleagues and Aurélie Cuvélier who helped me discuss the study results and build the present master thesis.
+To conclude, I thank everybody that played a role more or less important in the development of my internship and that I might have forgotten.

### 2017-07-20 02-Introduction.Rmd
Akcnowledgments

@@
 We assumed that when a disturbance event happen, due to a higher productivity, forest with an increased diversity will recover quicker and thus be more resilient.
-This theoritical expectation stand on two processes: complementarity and diversity effects [@Loreau2001].
-If we take a species pools with different productivity in monoculture.
-Their assemblage will allow theim an overall higher productivity due to a better ressources acquisition through ressource partitionning and niche differentiation.
+This theoritical expectation stand on two processes: complementarity and selection effects [@Loreau2001].
+If we take a species pool with different productivity in monoculture.
+Their assemblage will allow theim an overall higher productivity due to a better ressources acquisition through ressources partitionning and niche differentiation.
 Additionally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage.
@@
 This ressources preemption by more competitive species is the selection effect.
+Moreover, we expect the sampling effect to improve more diverse forest.
+A bigger initial sampling of the regional species pool in a rich forest, will allow more redundancy and a lower risk to lose important functional traits assemblages for the ecosystem when the disturbance happen.
 

### 2017-07-20 docs/master-thesis.tex
Akcnowledgments

@@
 
-I would like to thank\ldots{}
-
-\begin{itemize}
-\tightlist
-\item
-  Bruno
-\item
-  Stéphane
-\item
-  Camilla
-\item
-  Fabian
-\item
-  Isabelle
-\item
-  Laurent
-\item
-  Jérôme
-\item
-  Aurélie
-\item
-  Éric
-\item
-  Maxime
-\item
-  Raphaël
-\item
-  AMAP colleagues
-\item
-  Myriam
-\item
-  BIOGECO colleagues
-\item
-  Pascal Pe
-\item
-  Paracou team
-\item
-  Pascal Pa
-\item
-  Aurélie C
-\end{itemize}
+I would like to thank Bruno Hérault and Stéphane Traissac for their
+guidance and advice during the whole internship. I congratulate Bruno to
+have been able to communicate and be reactive besides his numerous
+recent travels. I am grateful to Bruno and Camille Piponiot to have
+introduced me to the wonderful world of Bayesian statistics. All this
+work would not have been possible without TROLL model and the whole team
+that works on it : Jérôme Chave, Isabelle Maréchaux and Fabian Fischer.
+I thank all of the team to have introduced me to the model and guide me
+in its development from the opposite side of the Atlantic. I would also
+like to thank Laurent Descroix and people from National Forest Office
+who introduced me to sylviculture from french Guiana and helped me to
+build a realistic forest model. I also thank Aurélie Dourdain for her
+help to access Paracou data, Éric Marcon for his advice on biodiversity
+analysis, and Pascal Padolus for his patience and help wiht Tabebuia
+cluster. The study would not have been possible without Pascal
+Petronelli and all people who have contributed to Paracou and Guyafor
+data collection over the years. Finally I am grateful to Maxime
+Réjou-Méchain, Raphaël Pélissier, and AMAP colleagues, in addition to
+Myriam Heuertz and BIOGECO colleagues and Aurélie Cuvélier who helped me
+discuss the study results and build the present master thesis. To
+conclude, I thank everybody that played a role more or less important in
+the development of my internship and that I might have forgotten.
 
@@
 This theoritical expectation stand on two processes: complementarity and
-diversity effects \citep{Loreau2001}. If we take a species pools with
+selection effects \citep{Loreau2001}. If we take a species pool with
 different productivity in monoculture. Their assemblage will allow theim
 an overall higher productivity due to a better ressources acquisition
-through ressource partitionning and niche differentiation. Additionally,
-some species will individually have an higher productivity in the
-assemblage than in monoculture due to facilitation from other species
-present in the assemblage. Complementarity effect is the addition of the
-better ressource acquisition and the facilitation. Now if we look at the
-evolution of the species assemblage over time, more competitive species
-will progressively dominate through competitive selection. And if
-competitive species are more productive, they will increase assemblage
-overall productivity. This ressources preemption by more competitive
-species is the selection effect.
+through ressources partitionning and niche differentiation.
+Additionally, some species will individually have an higher productivity
+in the assemblage than in monoculture due to facilitation from other
+species present in the assemblage. Complementarity effect is the
+addition of the better ressource acquisition and the facilitation. Now
+if we look at the evolution of the species assemblage over time, more
+competitive species will progressively dominate through competitive
+selection. And if competitive species are more productive, they will
+increase assemblage overall productivity. This ressources preemption by
+more competitive species is the selection effect. Moreover, we expect
+the sampling effect to improve more diverse forest. A bigger initial
+sampling of the regional species pool in a rich forest, will allow more
+redundancy and a lower risk to lose important functional traits
+assemblages for the ecosystem when the disturbance happen.
 
@@
 by TROLL is the light, the main facilitation will be light shading of
-post-pioneer species by pioneer species in disturbance gaps. We found
-similar results to \citet{Morin2011} regarding complementarity and light
-facilitation. But the complementarity effect is reducing through time,
-to let the selection effect insure forest resilience due to an inforced
-dominance of more productive species. The diminishing of complementarity
-effect is due to competitive selection though time in gaps succession.
-But the study scale matters as shwon by \citet{Chisholm2013}. Here we
-look at processes to a 16 ha scale. We do not have topography reducing
+post-pioneer species by pioneer species in disturbance gaps. Our results
+confirm the study of \citet{Morin2011} who also highlighted the
+importance of facilitation through light shading in forest resilience.
+But the complementarity effect is reducing through time, to let the
+selection effect insure forest resilience due to an inforced dominance
+of more productive species. The diminishing of complementarity effect is
+due to competitive selection though time in gaps succession. But the
+study scale matters as shwon by \citet{Chisholm2013}. Here we look at
+processes to a 16 ha scale. We do not have topography reducing
 micro-environment effects, but \citet{Chisholm2013} results suggest that

### 2017-07-24 02-Introduction.Rmd
Final soutenance

@@
 Their assemblage will allow theim an overall higher productivity due to a better ressources acquisition through ressources partitionning and niche differentiation.
-Additionally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage.
+Additionnally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage.
 Complementarity effect is the addition of the better ressource acquisition and the facilitation.
@@
 But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.
-
-In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
-We used a simulation approach using TROLL model to assess long term processes.

### 2017-07-24 03-Model.Rmd
Final soutenance

@@
 Each tree is consistently defined by its age, diameter at brest height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA) (see figure \@ref(fig:TROLLtree)).
-Tree geometry is calculated with allometric equations but leaf area vary dinamically within each crown following carbon allocations.
+Tree geometry is calculated with allometric equations but leaf area vary dynamically within each crown following carbon allocations.
 Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels.
@@
 We used an out of the bag method applied on a random forest to select variables with highest importance to explain leaf lifespan.
-We thus selected leaf mass per are $LMA$, leaf nitrogen content $N$ and wood specific gravity $wsg$.
+We thus selected leaf mass per area $LMA$, leaf nitrogen content $N$ and wood specific gravity $wsg$.
 We then used a bayesian approach to test different models with growing level of complexity.

### 2017-07-24 04-MnMs.Rmd
Final soutenance

@@
 This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]).
+Functional diversities of mature forests were assessed with @villeger_new_2008 indices (FRIC, FEve, FDiv, and FDis).
 Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the mature forest were simulated, and thus before disturbance.
@@
 Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions.
-We decided to describe simiulated tropical forests in two major functions: forest structure and forest functionning.
+We decided to describe simulated tropical forests in two major functions: forest structure and forest functionning.
 Forest structure was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$).
-Forest functionning was represented by growth pimary productivity ($GPP$ in $MgC.ha^{-1}$), net pimary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
+Forest functionning was represented by growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
 
@@
 This effect can be partitioned between complementarity effect $CE$, representing niche partitionning, positive interactions, and resource supply, and selectvie effect $SE$ due to dominant species pool driving the ecosystem.
+$N$ represents the total number of species, and $M$ the vector of monocultures performance.
 Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:
@@
 Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives).
-Whereas $SE$ become positive when dominant species outperform theimselves in mixture than in monoculture, and negative when less dominant species outperform theimselves in mixture than in monoculture [@Tobner2016].
+Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016].
 But similarly to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.

### 2017-07-24 05-Results.Rmd
Final soutenance

@@
 
-We measured all ecosystem outputs biodiversity net effect in disturbance simulations by comparing theim to their species corresponding monoculture simulations.
+We measured all ecosystem outputs biodiversity net effect in disturbance simulations by comparing them to their species corresponding monoculture simulations.
 The net effect was then partition between selection and complementarity effect.
@@
 We obtained those results for aboveground biomass.
-We found similar results but with an aplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]).
+We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]).
 Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 5: Disturbance simulations]).

### 2017-07-24 06-Discussion.Rmd
Akcnowledgments
Final soutenance

@@
 As the only ressource simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps.
-We found similar results to @Morin2011 regarding complementarity and light facilitation.
+Our results confirm the study of @Morin2011 who also highlighted the importance of facilitation through light shading in forest resilience.
 But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species.
 The diminishing of complementarity effect is due to competitive selection though time in gaps succession.
-But the study scale matters as shwon by @Chisholm2013.
+But the study scale matters as shown by @Chisholm2013.
 Here we look at processes to a 16 ha scale.

### 2017-07-24 docs/master-thesis.tex
Final soutenance

@@
 through ressources partitionning and niche differentiation.
-Additionally, some species will individually have an higher productivity
-in the assemblage than in monoculture due to facilitation from other
-species present in the assemblage. Complementarity effect is the
-addition of the better ressource acquisition and the facilitation. Now
-if we look at the evolution of the species assemblage over time, more
-competitive species will progressively dominate through competitive
+Additionnally, some species will individually have an higher
+productivity in the assemblage than in monoculture due to facilitation
+from other species present in the assemblage. Complementarity effect is
+the addition of the better ressource acquisition and the facilitation.
+Now if we look at the evolution of the species assemblage over time,
+more competitive species will progressively dominate through competitive
 selection. And if competitive species are more productive, they will
@@
 mechanisms involved in the relationship.
-
-In the present study, we focused on mechanisms involved in the
-relationship between biodiversity and forest ecosystem resilience. We
-used a simulation approach using TROLL model to assess long term
-processes.
 
@@
 geometry is calculated with allometric equations but leaf area vary
-dinamically within each crown following carbon allocations. Voxels
+dynamically within each crown following carbon allocations. Voxels
 resolution of 1 \(m^3\) allow the establishment of maximum one tree by
@@
 applied on a random forest to select variables with highest importance
-to explain leaf lifespan. We thus selected leaf mass per are \(LMA\),
+to explain leaf lifespan. We thus selected leaf mass per area \(LMA\),
 leaf nitrogen content \(N\) and wood specific gravity \(wsg\). We then
@@
 (\(60 ~communities*4~levels~of~disturbance\)) over 600 years (maturity
-being assumed after 500 years of regeneration \citep{Li}). Figure
+being assumed after 500 years of regeneration \citep{Li}). Functional
+diversities of mature forests were assessed with
+\citet{villeger_new_2008} indices (FRIC, FEve, FDiv, and FDis). Figure
 \ref{fig:DOE} presents the design of experiment for communities
@@
 Tropical forest ecosystems provides numerous ecosystem services linked
-to several ecosystem functions. We decided to describe simiulated
+to several ecosystem functions. We decided to describe simulated
 tropical forests in two major functions: forest structure and forest
@@
 (\(N10\)), and number of stem above \(30~cm\) diameter (\(N30\)). Forest
-functionning was represented by growth pimary productivity (\(GPP\) in
-\(MgC.ha^{-1}\)), net pimary productivity (\(NPP\) in \(MgC.ha^{-1}\)),
+functionning was represented by growth primary productivity (\(GPP\) in
+\(MgC.ha^{-1}\)), net primary productivity (\(NPP\) in \(MgC.ha^{-1}\)),
 tree autotrophic respiration in day (\(Rday\) in \(MgC.ha^{-1}\)) and
@@
 partitionning, positive interactions, and resource supply, and selectvie
-effect \(SE\) due to dominant species pool driving the ecosystem. Both
-metrics depend on the variation of relative ecosystem variable
-\(\Delta RX\):
+effect \(SE\) due to dominant species pool driving the ecosystem. \(N\)
+represents the total number of species, and \(M\) the vector of
+monocultures performance. Both metrics depend on the variation of
+relative ecosystem variable \(\Delta RX\):
 
@@
 positives). Whereas \(SE\) become positive when dominant species
-outperform theimselves in mixture than in monoculture, and negative when
-less dominant species outperform theimselves in mixture than in
+outperform themselves in mixture than in monoculture, and negative when
+less dominant species outperform themselves in mixture than in
 monoculture \citep{Tobner2016}. But similarly to resilience measurement,
@@
 We measured all ecosystem outputs biodiversity net effect in disturbance
-simulations by comparing theim to their species corresponding
-monoculture simulations. The net effect was then partition between
-selection and complementarity effect. We normalized complementarity and
-selection effect of disturbed simulations by biodiversity net effect of
+simulations by comparing them to their species corresponding monoculture
+simulations. The net effect was then partition between selection and
+complementarity effect. We normalized complementarity and selection
+effect of disturbed simulations by biodiversity net effect of
 undisturbed control simulation (see Table \ref{tab:tNE}) to measure the
@@
 We obtained those results for aboveground biomass. We found similar
-results but with an aplified signal for basal area (\(BA\)) and stem
+results but with an amplified signal for basal area (\(BA\)) and stem
 abundance (\(N\) but with an inverted signal because of the forest self
@@
 due to competitive selection though time in gaps succession. But the
-study scale matters as shwon by \citet{Chisholm2013}. Here we look at
+study scale matters as shown by \citet{Chisholm2013}. Here we look at
 processes to a 16 ha scale. We do not have topography reducing

### 2017-07-24 02-Introduction.Rmd
Final soutenance 2

@@
 But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.
+
+In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
+We used a simulation approach using TROLL model to assess long term processes for different types and levels of disturbances.
+We first assessed diversity effect on forest resilience of structure and functionning using numerous indices of taxonomic and functional diversities.
+Then, we measured biodiversity net effect, partitioned into complementarity and selection effects, resilience for several ecosystem metrics.

### 2017-07-24 docs/master-thesis.tex
Final soutenance 2

@@
 mechanisms involved in the relationship.
+
+In the present study, we focused on mechanisms involved in the
+relationship between biodiversity and forest ecosystem resilience. We
+used a simulation approach using TROLL model to assess long term
+processes for different types and levels of disturbances. We first
+assessed diversity effect on forest resilience of structure and
+functionning using numerous indices of taxonomic and functional
+diversities. Then, we measured biodiversity net effect, partitioned into
+complementarity and selection effects, resilience for several ecosystem
+metrics.
 

### 2017-07-25 docs/master-thesis.tex
Examinateurs

@@
     Dr Bruno HÉRAULT  \emph{Tuteur de stage} \\
-    Titre Prénom NOM  \emph{Examinateur} \\
-    Titre Prénom NOM  \emph{Examinateur} \\
+    Dr Éric MARCON  \emph{Examinateur} \\
+    Dr Heidy SCHIMANN  \emph{Examinateur} \\
     Dr Stéphane TRAISSAC  \emph{Enseignant-réferent} \\

### 2017-07-25 latex/before_body.tex
Discussion v1
Examinateurs

@@
   \large{Sujet :} \\
-  \Large{\textbf{Résilience des écosystèmes forestiers tropicaux après perturbation}} \\
+  \Large{\textbf{Rôle de la biodiversité dans la résilience des écosystèmes forestiers tropicaux après perturbation}} \\
   \vspace*{\fill}
@@
   Dr Bruno HÉRAULT  \emph{Tuteur de stage} \\
-  Titre Prénom NOM  \emph{Examinateur} \\
-  Titre Prénom NOM  \emph{Examinateur} \\
+  Dr Éric MARCON  \emph{Examinateur} \\
+  Dr Heidy SCHIMANN  \emph{Examinateur} \\
   Dr Stéphane TRAISSAC  \emph{Enseignant-réferent} \\

### 2017-07-26 01-Acknowledgments.Rmd
Typo

@@
 The study would not have been possible without Pascal Petronelli and all people who have contributed to Paracou and Guyafor data collection over the years.
-Finally I am grateful to Maxime Réjou-Méchain, Raphaël Pélissier, and AMAP colleagues, in addition to Myriam Heuertz and BIOGECO colleagues and Aurélie Cuvélier who helped me discuss the study results and build the present master thesis.
+Finally I am grateful to Maxime Réjou-Méchain, Raphaël Pélissier, and AMAP colleagues, in addition to Myriam Heuertz and BIOGECO colleagues and Aurélie Cuvelier who helped me discuss the study results and build the present master thesis.
 To conclude, I thank everybody that played a role more or less important in the development of my internship and that I might have forgotten.

### 2017-07-26 docs/master-thesis.tex
Typo

@@
 Réjou-Méchain, Raphaël Pélissier, and AMAP colleagues, in addition to
-Myriam Heuertz and BIOGECO colleagues and Aurélie Cuvélier who helped me
+Myriam Heuertz and BIOGECO colleagues and Aurélie Cuvelier who helped me
 discuss the study results and build the present master thesis. To

### 2018-01-26 S1.Rmd
Draft 0.1.9001 Init from the article draft branch

--- 2017-07-13 08-Appendix2.Rmd
@@
-# Appendix 2: Leaf lifespan model
+---
+title: "Supplementary material S1: Leaf lifespan model"
+date: '`r Sys.Date()`'
+output:
+  bookdown::pdf_document2:
+    number_sections: false
+    toc: false
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
+csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
+bibliography: /home/sylvain/Documents/Bibliography/library.bib
+link-citations: yes
+---
 
@@
 
-## Material and methods
+# Material and methods
 
@@
 
-## Results
+# Results
 
@@
 For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.
+
+# References

### 2018-01-26 S2.Rmd
Draft 0.1.9001 Init from the article draft branch

--- 2017-07-18 11-Appendix5.Rmd
@@
-# Appendix 5: Disturbance simulations
+---
+title: "Supplementary material S2: Disturbance simulations"
+date: '`r Sys.Date()`'
+output:
+  bookdown::pdf_document2:
+    number_sections: false
+    toc: false
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
+csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
+bibliography: /home/sylvain/Documents/Bibliography/library.bib
+link-citations: yes
+---
+
+```{r config, message=FALSE, warning=FALSE, include=FALSE}
+rm(list = ls()); invisible(gc())
+library(knitr)
+library(ggplot2)
+opts_chunk$set( echo = F, message = F, warning = F, 
+                fig.height = 8, fig.width = 8,
+                cache = T, cache.lazy = F)
+```
 
@@
 
-## Ecosystem functions
+# Ecosystem functions
 
@@
 
-## Biodiversity effect
+# Biodiversity effect
 
@@
 ```
+
+# References

### 2018-01-26 draft.Rmd
Draft 0.1.9001 Init from the article draft branch

---
title: Draft
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(ggplot2)
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 8,
                cache = T, cache.lazy = F)
```

# Abstract

**To be written (but eventually re-use [master-thesis one](https://sylvainschmitt.github.io/master-thesis/resume-et-abstract.html))**

# Introduction

Tropical forests disturbances, through deforestation and degradation, account for 8.26 billion of tons of carbon dioxyde emissions per year [@Pearson2017]. It represents the third source of greenhouse gas. Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics. Degradation from forest has been estimated to represents 10 time the deforestation area [@Herold2011] and represents 20% of greenhouse gas emissions in the brazilian Amazon [@Asner2005]. Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.

Sutainable forest management in the tropics (i.e. managed selective harvesting of timber) has been widely promoted internationnaly to combat tropical deforestation and other forms of degradations [@Zimmerman2012]. Currently logging from tropical forests accounts for one eighth of global timber production [@Blaser2011] and is still increasing. Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cutting cycle [@Martin2015].

On the other hand, tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978]. Effectively tropical forests host over half of the Earth's biodiversity [@Scheffers2012]. High biodiversity from tropical rainforests is the source of many ecosystem functions. Amongst others, tropical forests play a key role in biogeochemical cycles, including carbone storage [@Lewis2004]. Ecosystem functions from tropical forests support numerous ecosystem services, such as timber production and climate regulation. 
<!-- IM: À mettre avant ? -->

But several authors argue that selective logging represents a major threat to biodiversity [@Carreno-Rocabado2012; @DeAvila2015; @Gibson2013; @Martin2015; @Zimmerman2012], challenging the sustainable definition from current selective logging. We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better sylvicultural practices in order to reach sustainability.

The question of selective logging impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functionning [@Loreau2000]. Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to logging disturbance. And forest ecosystem response to logging disturbance will directly modify ecosystem functionning in both short and long term. Consequently assessing selective logging effect on tropical forest linking diversity and ecosystem functionning seems an obvious and promising way [@Loreau2010].

Negative short term impacts of selective logging have been assessed [@Carreno-Rocabado2012; @DeAvila2015;  but see @Martin2015]. Much less is known about the long term impact [@Osazuwa-Peters2015]. The main reason is the difficulty to conduct long term empirical study [but see @Herault2010], which can be completed by the use of forest simulators [@Huth2004; @Khler2004; @Ruger2008; @Tietjen2006]. Individual-based models of forest dynamics present the perfect framework to develop such joint biodiversity-ecosystem approaches [@Li]. Individual-based models describe forest accumulating carbon through time, assessing tree growth, or releasing carbon through gap opening [@Bugmann2001]. Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit [@Pacala1996]. Recently, the forest growth simulator TROLL [@Chave1999], an individual-based and spatially explicit forest model, was developped to introduce recent advances in plant physiological community. TROLL model relates physiological processes to species-specific functional traits [@Li]. Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functionning link response to logging disturbance.

We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance. Resilience encompass several definitions but was summarized by @Oliver2015 as the degree of resistance or fast recovery from an ecosystem function to environmental disturbance.

Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity. We assumed that when a disturbance event happen, due to a higher productivity, forest with an increased diversity will recover quicker and thus be more resilient. This theoritical expectation stand on two processes: complementarity and selection effects [@Loreau2001]. If we take a species pool with different productivity in monoculture. Their assemblage will allow theim an overall higher productivity due to a better ressources acquisition through ressources partitionning and niche differentiation. Additionnally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage. Complementarity effect is the addition of the better ressource acquisition and the facilitation. Now if we look at the evolution of the species assemblage over time, more competitive species will progressively dominate through competitive selection. And if competitive species are more productive, they will increase assemblage overall productivity. This ressources preemption by more competitive species is the selection effect. Moreover, we expect the sampling effect to improve more diverse forest. A bigger initial sampling of the regional species pool in a rich forest, will allow more redundancy and a lower risk to lose important functional traits assemblages for the ecosystem when the disturbance happen.

Generally, positive relationship between biodiversity and productivity were emprically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of tropical forests. One of the few studies was realized by @Chisholm2013 and has shown a positive significative relationship between species richness and wood productivity on a worlwide forest network. But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.

In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience. We used a simulation approach using TROLL model to assess long term processes for different types and levels of disturbances. We first assessed diversity effect on forest resilience of structure and functionning using numerous indices of taxonomic and functional diversities. Then, we measured biodiversity net effect, partitioned into complementarity and selection effects, resilience for several ecosystem metrics.

# Material and Methods

## Model

TROLL model each tree indivdually in a located environment. Thus TROLL model, alongside with SORTIE  [@Pacala1996; @Uriarte2009] and FORMIND [@Fischer2016; @Kohler1998], can be defined as an individual-based and spatially explicit forest growth model. TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production. Trees are growing in a spatialized light environment explicitly computed witin voxels of 1 $m^3$. Each tree is consistently defined by its age, diameter at brest height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA) (see figure \@ref(fig:TROLLtree)). Tree geometry is calculated with allometric equations but leaf area vary dynamically within each crown following carbon allocations. Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels. Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment. A species label is associated to a number of species specific parameters (see table \@ref(tab:traits)) related to functional trait values which can be sampled on the field.
Carbon assimilation is computed over half-hourly period of a representative day. Then allocation is computed to simulate tree growth from an explicit carbone balance (in contrast to previous models). Finally environment is updated at each timestep set to one month. Seedlings are not simulated explicitly but as a pool. In addition belowground processes, herbaceous plants, epiphytes and lianas are not simulated inside TROLL. The source code is written in C++ and available upon request. All modules of TROLL models are further detailed in [Appendix 1: TROLL model].

```{r traits, echo=FALSE}
table <- data.frame(
  Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breasth height threshold', 'asymptotic height', 'parameter of the tree-height-dbh allometry'),
  Units = c('$g.m^{-2}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$g.cm^{-3}$', '$m$', '$m$', '$m$')
)
knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Marechaux2017 Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.', format = 'pandoc')
```

Previous implementation of TROLL model used @Reich1991a allometry to infer leaf lifespan $LL$ from species leaf mass per area $LMA$ [@Li, see [Appendix 1: TROLL model]]. But the use of the allometrie from @Reich1991a with current implementation of the TROLL model resulted in an underestimation of leaf lifespan for low LMA species. Consequently in the following paragraph we suggest a new allometry.

Selective logging is defined as the targeted harvesting of timber from species of interest. Consequently, tropical sylviculture can be assimilated to a disturbance. The main difference between a disturbance and selective logging is the targetting of both species and individuals of interest. So we decided to first asses unselective disturbance effect on tropical forest ecosystem to subsequently better understand selective logging effect. First, we implemented a disturbance module inside TROLL model to simulate unselective disturbance. Secondly, we implemented a sylviculture module inside TROLL model to simulate selective logging in regards to french Guiana practices.

Disturbance module was designed in the simplest way in order to relate the ecosystem answer to volume loss without any individuals nor species targetting. For a given iteration $disturb_{iter}$, individuals are picked randomly with a uniform law on the number of trees. Selected individuals are then removed without trigerring a treefall to avoid any side effect. The operation is repeated untill the disturbance result in a defined lost basal area ($disturb_{intensity}$ in % of BA).

## Design of experiment

In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time. Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation using default parameters. Biodiversity was integrated with two of its components: taxonomic and functional diversities. We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species). Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean. Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess them after running the simulations, i.e. before applying the diturbance. Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$. For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means. Effectivelly, we did not wanted drastic change in community means that could have more effect than functional richness itself. This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]). Functional diversities of mature forests were assessed with @villeger_new_2008 indices (FRIC, FEve, FDiv, and FDis). Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the mature forest were simulated, and thus before disturbance. We obtained a broad range of both functionl dispersion $FDis$ and aboveground biomass $AGB$ for simulated forest ecosystems before disturbance.

```{r DOE, echo=FALSE, fig.cap='Experimental design before disturbance. Communities are implemented along a gradient of species richness (SR) and functional dispersion (FDis) resulting in a broad range of aboveground biomass (AGB). FDis was caluclated based on 4 functional traits (leaf mass per area, wood specific gravity, maximum diameter, and maximum height).', cache=TRUE}
# Data prep
load('~/Documents/ECOFOG/Results/disturbance/maturity/matureData.Rdata')
data <- datal$time
data <- data[data$time > 450,]
data <- data.frame(apply(data[c('agb','ba','n10','n30')],
                         2, function(x) tapply(x, data$sim, mean)))
data <- cbind(data.frame(sim = row.names(data)), data)
data$richness <- as.factor(floor(as.numeric(as.character(data$sim))))
species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt',
                      header=T, dec=".", sep="",
                      row.names = 1)[c('LMA', 'wsg', 'dmax', 'hmax')]
species <- species[row.names(datal$species),]
fd <- suppressWarnings(FD::dbFD(species, t(datal$species), messages = F))
data <- cbind(data, do.call('cbind',
                            fd[c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ')]))
data <- cbind(data, fd$CWM)
rm(fd, species, datal)

# Graph
g <- ggplot(data, aes(x = FDis, y = richness, 
                      label = sim, color = agb/1000)) + 
  geom_point(size = 3) + 
  ggrepel::geom_text_repel() +
  scale_color_continuous("Aboveground\nBiomass\n(AGB, ton C/ha)") +
  xlab("Functional dispersion (FDis)") +
  ylab("Species richness (SR)")

g ; rm(data)
```

## Ecosystem response analysis

### Ecosystem functions

Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions. We decided to describe simulated tropical forests in two major functions: forest structure and forest functionning. Forest structure was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$). Forest functionning was represented by growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).

The resilience of metrics values post disturbance were assessed through @Henry2012 formula:

\begin{equation}
  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X_T(t)}{X_C(t)}
  (\#eq:Resilience)
\end{equation}

The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$. But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic. Consequently, to encompass undisturbed ecosystem variations throught time, we simulated an undisturbed control ecosystem $C$. And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ . Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics. $R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value. Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$. Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest strcuture (AGB, BA, N, N10, and N30) and forest functionning (GPP, NPP, Rday, and Rnight). We then used integrated eulcidean distance to equilibrium over time to assess simulations resilience.

### Biodiversity effect

Biodiversity is not only a facet of the experimental design and an ecosystem output through forest diversity, but also interact on ecosystem functioning and consequently on its answer to disturbance. Biodiversity ecosystem functioning relation can be split in complementarity and selection effect with @Loreau2001 partitioning:

\begin{equation}
  \begin{array}{c}
    NE = X_O - X_E = CE + SE \\
    CE = N* \overline{\Delta RX} \overline{M}\\
    SE = N*cov(\Delta RX,M)
  \end{array}
  (\#eq:BiodivPart)
\end{equation}

Biodiversity net effect $NE$ is based on the difference between ecosystem variable $X$ observed value $X_O$ within the community mixture of species and its expected value $X_E$ if species performance were equal to their performance in monocultures. This effect can be partitioned between complementarity effect $CE$, representing niche partitionning, positive interactions, and resource supply, and selectvie effect $SE$ due to dominant species pool driving the ecosystem. $N$ represents the total number of species, and $M$ the vector of monocultures performance. Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:

\begin{equation}
  \Delta RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
  (\#eq:DeltaRY)
\end{equation}

$X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$. $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance. Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives). Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016]. But similarly to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance. So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:

\begin{equation}
  R(NE_T) = \frac{NE_T}{NE_C} = \frac{SE_T}{NE_C} + \frac{CE_T}{NE_C}
  (\#eq:RNE)
\end{equation}

Resilience trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$. In order to do that, the design of experiment was repeated for each species indivdually representing 652 simulations of monoculture.

# Resultss

## Disturbance

### Ecosystem functions

```{r resilience, echo=FALSE}
rich <- c(5,25,125)
cvh <- 1:20
dist <- c(0,25,50,75)
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
load(file.path(path, 'disturbanceData.Rdata'))
load(file.path(path, 'disturbanceResilience.Rdata'))
```

```{r Deq, cache=TRUE, echo=FALSE}
deq <- data.frame(
  time = resilience$time,
  sim = resilience$sim,
  structure = apply(resilience[c('agb', 'ba', 'n', 'n10', 'n30')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))),
  production = apply(resilience[c('gpp', 'npp', 'Rday', 'Rnight')], 1, 
                  function(x) sqrt(sum((rep(1, length(x)) - x)^2)))
)
```

```{r Ieq, cache=TRUE, echo=FALSE}
sim <- c(sapply(c(sapply(rich, function(r) paste0(r, '.', cvh))), function(x) paste0(x, '.', dist[-1])))
Ieq <- sapply(sim, function(s){
  data_s <- subset(deq, sim ==s)
  return(cbind(data_s[c(1,2)], cumsum(data_s[-c(1,2)])))
}, simplify = F) ; rm(sim)
Ieq <- data.frame(do.call('rbind', Ieq))
names(Ieq)[3:4] <- paste0('Ieq.', names(Ieq)[3:4])
```

We transformed all ecosystem outputs from the 240 disturbance simulations in resilience metrics normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A et B). We then gathered ecosystem outputs by main ecosystem functions (forest dynamic and forest production) to compute ecosystem distance to equilibrium (Figure \@ref(fig:datatrans) C). Finally, we integrated distance to equilibrium in a cummulative sum over time (Figure \@ref(fig:datatrans) D).

```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation. Ecosystem outputs (**A**) are normalized by the control value over time to calculate resilience (**B**); resilience of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance ot equilibirum is integrated over time in a cummulative sum (**D**). ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
library(dplyr)
g_agb <- ecosystem %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Aboveground biomass (tonC/ha)') +
  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
g_resilience <- resilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb, color = disturbance)) +
  geom_hline(yintercept = 1, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Resilience of aboveground biomass') +
  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
g_deq <- deq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, color = disturbance)) +
  geom_hline(yintercept = 0, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
g_Ieq <- Ieq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = Ieq.structure, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
```

The ranking was stable over time for the 240 simulations. So we used the cummulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience. We compared cummulative integral after 600 years to communities taxonomic and functional diversity for each leavel of disturbance (see Figure \@ref(fig:gIeq)). We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cummulative integral from ecosystem distance to forest structure equilibrium after 600 years ($Ieq_{600}$). In addition, functional evenness was complementary reducing $Ieq_{600}$. Finally species richness was not directly link to $Ieq_{600}$. Effectivelly, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$. We found similar results for all disturbance levels and forest functionning (see [Appendix 5: Disturbance simulations]).

```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>% 
  ggplot(aes(x = FDiv, y = Ieq.structure, label = sim)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(shape = richness, color = FEve)) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
  xlab('Functional diversity (FDiv)') +
  ylab('Cummulative integral from ecosystem 
       distance to forest structure equilibrium after 600 years') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\nevenness\n(FEve)')
```

### Biodiversity effect

```{r BE, echo=FALSE}
load(file.path(path, 'disturbanceBE.Rdata'))
load(file.path(path, 'disturbanceBErecovery.Rdata'))
```

We measured all ecosystem outputs biodiversity net effect in disturbance simulations by comparing them to their species corresponding monoculture simulations. The net effect was then partition between selection and complementarity effect. We normalized complementarity and selection effect of disturbed simulations by biodiversity net effect of undisturbed control simulation (see Table \@ref(tab:tNE)) to measure the resilience to disturbance of their aboveground biomass (see Figure \@ref(fig:gBE)). We found that complementarity effect was recovering biodiversity net effect and was stronger than selection effect in the first decades. But after few decades the complementarity effect diminsihed toward a low value. On the contrary selection effect was reduced or even removed by the disturbance. It increased during the whole simulation and was greater than complementarity effect only after decades. The time lag for which complementarity effect was greater than selection effect was increasing with disturbance intensity. Finally 600 years after the disturbance event biodiversity net effect was still not recovered for a disturbance intensity greater than 25% of basal area. We obtained those results for aboveground biomass. We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]). Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 5: Disturbance simulations]).

```{r tNE, echo=FALSE, cache=TRUE}
library(reshape2)
library(knitr)
BE %>%
  mutate(agb = agb/1000) %>% 
  filter(BE == 'NE') %>%
  melt(id.vars = c('BE', 'time', 'sim')) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>%
  filter(disturbance == 0) %>%
  group_by(variable) %>% 
  summarise(mean = mean(value, na.rm = T), 
            'standard deviation' = sd(value, na.rm = T)) %>% 
  filter(variable != 'n30') %>% 
  bind_cols(., data.frame(
    name = c('aboveground biomass', 'basal area', 'number of stems', 
             'number of stems above 10cm dbh', 'growth primary production', 
             'net primary production', 'autotrophic respiration during day',
             'autotrophic respiration during night'),
    unit = c('$tonC.ha^{-1}$', '$m^2.ha^{-1}$', '$n.ha^{-1}$', '$n.ha^{-1}$', 
             '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$'))) %>% 
  kable(caption = 'Table 1: Biodiversity net effect mean value and standard deviation for different ecosystem variable.', digits = 3, format = 'pandoc')
```

```{r gBE, echo=FALSE, fig.cap='Resilience of complementarity and selection effects. Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time.', cache=TRUE, fig.width=12}
BEresilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(BE %in% c('SE', 'CE')) %>% 
  group_by(time, BE, disturbance, richness) %>%
  summarise(qt1 =quantile(agb, probs=0.05),
            qt2 =quantile(agb, probs=0.75),
            mean=mean(agb),
            n=n()) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_grid(richness ~ disturbance, labeller = 'label_both') +
  xlab('Time (years)') +
  ylab('Recovery of aboveground biomass net effect')
```

# Discussion

In the limit of the model, we were able to show that diversity improved tropical forest resilience. More particularly, functional diversity and evenness are key components of diversity in forest recovery after disturbance. Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery. Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity. But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012. Because if the harvesting is not sustainable negative feedbacks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.

## TROLL limits

Belowground processes, herbaceaous plants, epyphytes and lianas are not simulated in TROLL but they not reprensent the only limit of the model. Other processes are simulated but simplified, and we used sensitivity analysis to assess their relative importance. We found that few functional traits were influencing whole forest structure and dynamic. In addition, we found that the seed rain constant had an important effect on species functional composition and diversity. High external seed rain resulted in a quick recovery of the system toward an equilibrium close to the regional species frequency levied by the seed rain. On the contrary, low seed rain let the simulated forest works as a closed system with more system feedback but a lower stability through time with longer species diversity transitions. In order to study the role of diversity in forest resilience we decided to remove the seed rain constant to get a closed system and look at the role of diversity when it maintains itself through feedbacks and not with immigration.

Finally, sylviculture module implemented inside TROLL showed lacks of ability to reproduce selective logging with current design of experiments. Results did not seem to be usable and will not be discussed. The main issues was the ability of the model to simulate mature forest with correct abundancy of mature trees from commonly harvested species. Two solutions might be possible in the tunning of the model. First, forest inventories could be used to initialize the model instead of a bare soil resulting in realistic species abundances, especially for harvested species. Secondly, selective logging could focus on guilds of species meeting peculiar values of functional traits [as wood density, see @Huth2004; @Khler2004; @Ruger2008] allowing the model to harvest more volume in order to meet reality.

## Diversity improve tropical forest resilience

Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectivelly). Thus we were able to show that diversity improve forest resilience. More particularly, functional diversity seems a major aspect of resilience if its strengthened by a high functional evenness. Effectively high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species. Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes. Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis. Finally, species and functional richnesses are not directly increasing resilience. But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998]. An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.

## Complementarity and selection insure forest resilience

We found that complementary effect was insuring forest resilience in the beginning of forest successions. We interpreted this results as the consequence of facilitation processes. As the only ressource simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps. Our results confirm the study of @Morin2011 who also highlighted the importance of facilitation through light shading in forest resilience. But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species. The diminishing of complementarity effect is due to competitive selection though time in gaps succession. But the study scale matters as shown by @Chisholm2013. Here we look at processes to a 16 ha scale. We do not have topography reducing micro-environment effects, but @Chisholm2013 results suggest that complementarity will be stronger at smaller scale and could explain its low value after forest recovery. Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada. They also found that selection effect was greater than complementarity in most of the cases. In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield). High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two cutting cycles in order to maintain both productive ecosystem and sustainable management.

## Conclusion

We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance. We found that diversity improved forest resilience together with productivity [@Liang2016]. Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery. Our results advocate for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles. Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience. But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedbacks. @Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces". Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy. This view correspond to the high resilience of tropical forest hyperdiverse systems, we were able to show, and let hope for a future sustainable selective logging in tropical forests.

# Supplementary materials

* Supplementary materials S1: Leaf lifespan model
    + [HTML](S1.html)
    + [PDF](S1.pdf)
* Supplementary materials S2: Disturbance simulations
    + [HTML](S2.html)
    + [PDF](S2.pdf)

# References

### 2018-01-31 draft.Rmd
Bruno's draft

@@
 output:
+  bookdown::pdf_document2:
+    includes:
+      before_body: doc_prefix.tex
+      in_header: preamble.tex
+    number_sections: false
+    toc: false
   bookdown::html_document2:
@@
     toc_float: yes
-  bookdown::pdf_document2:
-    number_sections: false
-    toc: false
 csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
@@
 library(knitr)
-library(ggplot2)
+library(tidyverse)
 opts_chunk$set( echo = F, message = F, warning = F, 
@@
 ```
-
-# Abstract
-
-**To be written (but eventually re-use [master-thesis one](https://sylvainschmitt.github.io/master-thesis/resume-et-abstract.html))**
-
-# Introduction
-
-Tropical forests disturbances, through deforestation and degradation, account for 8.26 billion of tons of carbon dioxyde emissions per year [@Pearson2017].
-It represents the third source of greenhouse gas.
-Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics.
-Degradation from forest has been estimated to represents 10 time the deforestation area [@Herold2011] and represents 20% of greenhouse gas emissions in the brazilian Amazon [@Asner2005].
-Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.
-
-Sutainable forest management in the tropics (i.e.
-managed selective harvesting of timber) has been widely promoted internationnaly to combat tropical deforestation and other forms of degradations [@Zimmerman2012].
-Currently logging from tropical forests accounts for one eighth of global timber production [@Blaser2011] and is still increasing.
-Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cutting cycle [@Martin2015].
-
-On the other hand, tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
-Effectively tropical forests host over half of the Earth's biodiversity [@Scheffers2012].
-High biodiversity from tropical rainforests is the source of many ecosystem functions.
-Amongst others, tropical forests play a key role in biogeochemical cycles, including carbone storage [@Lewis2004].
-Ecosystem functions from tropical forests support numerous ecosystem services, such as timber production and climate regulation.
-<!-- IM: À mettre avant ? -->
-
-But several authors argue that selective logging represents a major threat to biodiversity [@Carreno-Rocabado2012; @DeAvila2015; @Gibson2013; @Martin2015; @Zimmerman2012], challenging the sustainable definition from current selective logging.
-We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better sylvicultural practices in order to reach sustainability.
-
-The question of selective logging impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functionning [@Loreau2000].
-Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to logging disturbance.
-And forest ecosystem response to logging disturbance will directly modify ecosystem functionning in both short and long term.
-Consequently assessing selective logging effect on tropical forest linking diversity and ecosystem functionning seems an obvious and promising way [@Loreau2010].
-
-Negative short term impacts of selective logging have been assessed [@Carreno-Rocabado2012; @DeAvila2015;  but see @Martin2015].
-Much less is known about the long term impact [@Osazuwa-Peters2015].
-The main reason is the difficulty to conduct long term empirical study [but see @Herault2010], which can be completed by the use of forest simulators [@Huth2004; @Khler2004; @Ruger2008; @Tietjen2006].
-Individual-based models of forest dynamics present the perfect framework to develop such joint biodiversity-ecosystem approaches [@Li].
-Individual-based models describe forest accumulating carbon through time, assessing tree growth, or releasing carbon through gap opening [@Bugmann2001].
-Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit [@Pacala1996].
-Recently, the forest growth simulator TROLL [@Chave1999], an individual-based and spatially explicit forest model, was developped to introduce recent advances in plant physiological community.
-TROLL model relates physiological processes to species-specific functional traits [@Li].
-Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functionning link response to logging disturbance.
-
-We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance.
-Resilience encompass several definitions but was summarized by @Oliver2015 as the degree of resistance or fast recovery from an ecosystem function to environmental disturbance.
-
-Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity.
-We assumed that when a disturbance event happen, due to a higher productivity, forest with an increased diversity will recover quicker and thus be more resilient.
-This theoritical expectation stand on two processes: complementarity and selection effects [@Loreau2001].
-If we take a species pool with different productivity in monoculture.
-Their assemblage will allow theim an overall higher productivity due to a better ressources acquisition through ressources partitionning and niche differentiation.
-Additionnally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage.
-Complementarity effect is the addition of the better ressource acquisition and the facilitation.
-Now if we look at the evolution of the species assemblage over time, more competitive species will progressively dominate through competitive selection.
-And if competitive species are more productive, they will increase assemblage overall productivity.
-This ressources preemption by more competitive species is the selection effect.
-Moreover, we expect the sampling effect to improve more diverse forest.
-A bigger initial sampling of the regional species pool in a rich forest, will allow more redundancy and a lower risk to lose important functional traits assemblages for the ecosystem when the disturbance happen.
-
-Generally, positive relationship between biodiversity and productivity were emprically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of tropical forests.
-One of the few studies was realized by @Chisholm2013 and has shown a positive significative relationship between species richness and wood productivity on a worlwide forest network.
-But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.
-
-In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
-We used a simulation approach using TROLL model to assess long term processes for different types and levels of disturbances.
-We first assessed diversity effect on forest resilience of structure and functionning using numerous indices of taxonomic and functional diversities.
-Then, we measured biodiversity net effect, partitioned into complementarity and selection effects, resilience for several ecosystem metrics.
-
-# Material and Methods
-
-## Model
-
-TROLL model each tree indivdually in a located environment.
-Thus TROLL model, alongside with SORTIE  [@Pacala1996; @Uriarte2009] and FORMIND [@Fischer2016; @Kohler1998], can be defined as an individual-based and spatially explicit forest growth model.
-TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production.
-Trees are growing in a spatialized light environment explicitly computed witin voxels of 1 $m^3$.
-Each tree is consistently defined by its age, diameter at brest height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA) (see figure \@ref(fig:TROLLtree)).
-Tree geometry is calculated with allometric equations but leaf area vary dynamically within each crown following carbon allocations.
-Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels.
-Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
-A species label is associated to a number of species specific parameters (see table \@ref(tab:traits)) related to functional trait values which can be sampled on the field.
-Carbon assimilation is computed over half-hourly period of a representative day.
-Then allocation is computed to simulate tree growth from an explicit carbone balance (in contrast to previous models).
-Finally environment is updated at each timestep set to one month.
-Seedlings are not simulated explicitly but as a pool.
-In addition belowground processes, herbaceous plants, epiphytes and lianas are not simulated inside TROLL.
-The source code is written in C++ and available upon request.
-All modules of TROLL models are further detailed in [Appendix 1: TROLL model].
-
-```{r traits, echo=FALSE}
-table <- data.frame(
-  Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
-  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breasth height threshold', 'asymptotic height', 'parameter of the tree-height-dbh allometry'),
-  Units = c('$g.m^{-2}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$g.cm^{-3}$', '$m$', '$m$', '$m$')
-)
-knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Marechaux2017 Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.', format = 'pandoc')
-```
-
-Previous implementation of TROLL model used @Reich1991a allometry to infer leaf lifespan $LL$ from species leaf mass per area $LMA$ [@Li, see [Appendix 1: TROLL model]].
-But the use of the allometrie from @Reich1991a with current implementation of the TROLL model resulted in an underestimation of leaf lifespan for low LMA species.
-Consequently in the following paragraph we suggest a new allometry.
-
-Selective logging is defined as the targeted harvesting of timber from species of interest.
-Consequently, tropical sylviculture can be assimilated to a disturbance.
-The main difference between a disturbance and selective logging is the targetting of both species and individuals of interest.
-So we decided to first asses unselective disturbance effect on tropical forest ecosystem to subsequently better understand selective logging effect.
-First, we implemented a disturbance module inside TROLL model to simulate unselective disturbance.
-Secondly, we implemented a sylviculture module inside TROLL model to simulate selective logging in regards to french Guiana practices.
-
-Disturbance module was designed in the simplest way in order to relate the ecosystem answer to volume loss without any individuals nor species targetting.
-For a given iteration $disturb_{iter}$, individuals are picked randomly with a uniform law on the number of trees.
-Selected individuals are then removed without trigerring a treefall to avoid any side effect.
-The operation is repeated untill the disturbance result in a defined lost basal area ($disturb_{intensity}$ in % of BA).
-
-## Design of experiment
-
-In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time.
-Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation using default parameters.
-Biodiversity was integrated with two of its components: taxonomic and functional diversities.
-We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species).
-Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean.
-Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess them after running the simulations, i.e.
-before applying the diturbance.
-Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$.
-For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means.
-Effectivelly, we did not wanted drastic change in community means that could have more effect than functional richness itself.
-This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]).
-Functional diversities of mature forests were assessed with @villeger_new_2008 indices (FRIC, FEve, FDiv, and FDis).
-Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the mature forest were simulated, and thus before disturbance.
-We obtained a broad range of both functionl dispersion $FDis$ and aboveground biomass $AGB$ for simulated forest ecosystems before disturbance.
-
-```{r DOE, echo=FALSE, fig.cap='Experimental design before disturbance.
-Communities are implemented along a gradient of species richness (SR) and functional dispersion (FDis) resulting in a broad range of aboveground biomass (AGB).
-FDis was caluclated based on 4 functional traits (leaf mass per area, wood specific gravity, maximum diameter, and maximum height).', cache=TRUE}
-# Data prep
-load('~/Documents/ECOFOG/Results/disturbance/maturity/matureData.Rdata')
-data <- datal$time
-data <- data[data$time > 450,]
-data <- data.frame(apply(data[c('agb','ba','n10','n30')],
-                         2, function(x) tapply(x, data$sim, mean)))
-data <- cbind(data.frame(sim = row.names(data)), data)
-data$richness <- as.factor(floor(as.numeric(as.character(data$sim))))
-species <- read.table('~/Documents/ECOFOG/TROLL/inst/extdata/species.txt',
-                      header=T, dec=".", sep="",
-                      row.names = 1)[c('LMA', 'wsg', 'dmax', 'hmax')]
-species <- species[row.names(datal$species),]
-fd <- suppressWarnings(FD::dbFD(species, t(datal$species), messages = F))
-data <- cbind(data, do.call('cbind',
-                            fd[c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ')]))
-data <- cbind(data, fd$CWM)
-rm(fd, species, datal)
-
-# Graph
-g <- ggplot(data, aes(x = FDis, y = richness, 
-                      label = sim, color = agb/1000)) + 
-  geom_point(size = 3) + 
-  ggrepel::geom_text_repel() +
-  scale_color_continuous("Aboveground\nBiomass\n(AGB, ton C/ha)") +
-  xlab("Functional dispersion (FDis)") +
-  ylab("Species richness (SR)")
-
-g ; rm(data)
-```
-
-## Ecosystem response analysis
-
-### Ecosystem functions
-
-Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions.
-We decided to describe simulated tropical forests in two major functions: forest structure and forest functionning.
-Forest structure was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$).
-Forest functionning was represented by growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
-
-The resilience of metrics values post disturbance were assessed through @Henry2012 formula:
-
-\begin{equation}
-  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X_T(t)}{X_C(t)}
-  (\#eq:Resilience)
-\end{equation}
-
-The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$.
-But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic.
-Consequently, to encompass undisturbed ecosystem variations throught time, we simulated an undisturbed control ecosystem $C$.
-And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ .
-Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics.
-$R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value.
-Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$.
-Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest strcuture (AGB, BA, N, N10, and N30) and forest functionning (GPP, NPP, Rday, and Rnight).
-We then used integrated eulcidean distance to equilibrium over time to assess simulations resilience.
-
-### Biodiversity effect
-
-Biodiversity is not only a facet of the experimental design and an ecosystem output through forest diversity, but also interact on ecosystem functioning and consequently on its answer to disturbance.
-Biodiversity ecosystem functioning relation can be split in complementarity and selection effect with @Loreau2001 partitioning:
-
-\begin{equation}
-  \begin{array}{c}
-    NE = X_O - X_E = CE + SE \\
-    CE = N* \overline{\Delta RX} \overline{M}\\
-    SE = N*cov(\Delta RX,M)
-  \end{array}
-  (\#eq:BiodivPart)
-\end{equation}
-
-Biodiversity net effect $NE$ is based on the difference between ecosystem variable $X$ observed value $X_O$ within the community mixture of species and its expected value $X_E$ if species performance were equal to their performance in monocultures.
-This effect can be partitioned between complementarity effect $CE$, representing niche partitionning, positive interactions, and resource supply, and selectvie effect $SE$ due to dominant species pool driving the ecosystem.
-$N$ represents the total number of species, and $M$ the vector of monocultures performance.
-Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:
-
-\begin{equation}
-  \Delta RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
-  (\#eq:DeltaRY)
-\end{equation}
-
-$X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$.
-$P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance.
-Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives).
-Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016].
-But similarly to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
-So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:
-
-\begin{equation}
-  R(NE_T) = \frac{NE_T}{NE_C} = \frac{SE_T}{NE_C} + \frac{CE_T}{NE_C}
-  (\#eq:RNE)
-\end{equation}
-
-Resilience trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
-In order to do that, the design of experiment was repeated for each species indivdually representing 652 simulations of monoculture.
-
-# Resultss
-
-## Disturbance
-
-### Ecosystem functions
 
@@
 
-We transformed all ecosystem outputs from the 240 disturbance simulations in resilience metrics normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A et B).
-We then gathered ecosystem outputs by main ecosystem functions (forest dynamic and forest production) to compute ecosystem distance to equilibrium (Figure \@ref(fig:datatrans) C).
-Finally, we integrated distance to equilibrium in a cummulative sum over time (Figure \@ref(fig:datatrans) D).
+# Abstract
+
+**To be written (but eventually re-use [master-thesis one](https://sylvainschmitt.github.io/master-thesis/resume-et-abstract.html))**
+
+# Introduction
+
+Tropical forests disturbances, through deforestation and degradation, account for 8.26 billion of tons of carbon dioxide emissions per year [@Pearson2017].
+It represents the third source of greenhouse gas.
+Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics.
+Degradation from forest has been estimated to represents 10 time the deforestation area [@Herold2011] and represents 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].
+Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.
+
+Sustainable forest management in the tropics (i.e.
+managed selective harvesting of timber) has been widely promoted internationally to combat tropical deforestation and other forms of degradation [@Zimmerman2012].
+Currently logging from tropical forests accounts for one eighth of global timber production [@Blaser2011] and is still increasing.
+Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cutting cycle [@Martin2015].
+
+On the other hand, tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
+Effectively tropical forests host over half of the Earth's biodiversity [@Scheffers2012].
+High biodiversity from tropical rainforests is the source of many ecosystem functions.
+Among others, tropical forests play a key role in biogeochemical cycles, including carbon storage [@Lewis2004].
+Ecosystem functions from tropical forests support numerous ecosystem services, such as timber production and climate regulation.
+
+But several authors argue that selective logging represents a major threat to biodiversity [@Carreno-Rocabado2012; @DeAvila2015; @Gibson2013; @Martin2015; @Zimmerman2012], challenging the sustainable definition from current selective logging.
+@DeAvila2015 found that additional sylvicultural practices and logging damages had stronger negative effect on diversity than logging itself.
+@Carreno-Rocabado2012 found no direct effect of medium intensity logging on species diversity but a community drift towards 'fast' acquisitive species [@reich_world-wide_2014].
+However, even the short term impact on biodiversity is unclear, as sometimes short term increases in diversity can occur at low sylvicultural intensities [@Martin2015].
+We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better sylvicultural practices in order to reach sustainability.
+
+The question of selective logging impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functioning [@Loreau2000].
+Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to logging disturbance.
+And forest ecosystem response to logging disturbance will directly modify ecosystem functioning in both short and long term.
+Consequently assessing selective logging effect on tropical forest linking diversity and ecosystem functioning seems an obvious and promising way [@Loreau2010].
+
+Negative short term impacts of selective logging have been assessed [@Carreno-Rocabado2012; @DeAvila2015;  but see @Martin2015].
+Much less is known about the long term impact [@Osazuwa-Peters2015].
+The main reason is the difficulty to conduct long term empirical study [but see @Herault2010], which can be completed by the use of forest simulators [@Huth2004; @Khler2004; @Ruger2008; @Tietjen2006].
+Individual-based models of forest dynamics present the perfect framework to develop such joint biodiversity-ecosystem approaches [@Marechaux2017].
+Individual-based models describe forest accumulating carbon through time, assessing tree growth, or releasing carbon through gap opening [@Bugmann2001].
+Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit [@Pacala1996].
+Recently, the forest growth simulator TROLL [@Chave1999], an individual-based and spatially explicit forest model, was developed to introduce recent advances in plant ecophysiology.
+TROLL model relates physiological processes to species-specific functional traits [@Marechaux2017].
+Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functioning link response to logging disturbance.
+
+We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance.
+Resilience encompass several definitions but was summarized by @Oliver2015 as the degree of resistance or fast recovery from an ecosystem function to environmental disturbance.
+
+Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity.
+We assumed that when a disturbance event happen, due to a higher productivity, forest with an increased diversity will recover quicker and thus be more resilient, besides other mechanisms could be involved in an increased resilience from diverse forest.
+This theoretical expectation stand on two processes: complementarity and selection effects [@Loreau2001].
+If we take a species pool with different productivity in monoculture, their assemblage will allow them an overall higher productivity due to a better resources acquisition through resources partitioning and niche differentiation.
+Additionally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage.
+Complementarity effect is the addition of the better resource acquisition and the facilitation.
+Now if we look at the evolution of the species assemblage over time, more competitive species will progressively dominate through competitive selection.
+And if competitive species are more productive, they will increase assemblage overall productivity.
+This resources preemption by more competitive species is the selection effect.
+Moreover, we expect the sampling effect to improve resilience of species-rich forests.
+A bigger initial sampling of the regional species pool in a rich forest, will allow more redundancy and a lower risk to lose important functional traits assemblages for the ecosystem when the disturbance happen.
+
+Generally, positive relationship between biodiversity and productivity were empirically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of forests.
+One of the few studies was realized by @Chisholm2013 and has shown a positive significative relationship between species richness and wood productivity on a worldwide forest network.
+But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.
+
+We suggest to use a process-based model to overcome those limitations because TROLL model have the same key features that @Morin2011: (i) TROLL has been developed and validated independently from the biodiversity and ecosystem functioning issue, (ii) TROLL use 163 real tropical tree species characterized by their functional traits, (iii) TROLL allows to study long term effects with realized species richness for several centuries, (iv) TROLL allows to explore more species associations than real experiments, (v) TROLL allows to explore process resulting in observed relations.
+In addition, TROLL has been designed to be used in species-rich ecosystems such as tropical rainforests.
+
+In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
+We used a simulation approach using TROLL model to assess long term processes for different types and levels of disturbances.
+We first assessed diversity effect on forest resilience of structure and functioning using numerous indices of taxonomic and functional diversities.
+Then, we measured biodiversity net effect, partitioned into complementarity and selection effects, on resilience for several ecosystem metrics.
+
+# Material and Methods
+
+## Model
+
+TROLL model each tree individually in a located environment.
+Thus TROLL model, alongside with SORTIE  [@Pacala1996; @Uriarte2009] and FORMIND [@Fischer2016; @Kohler1998], can be defined as an individual-based and spatially explicit forest growth model.
+TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production.
+Trees are growing in a spatialized light environment explicitly computed within voxels of 1 $m^3$.
+Each tree is consistently defined by its age, diameter at breast height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
+Tree geometry is calculated with allometric equations but leaf area vary dynamically within each crown following carbon allocations.
+Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels.
+Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
+A species label is associated to a number of species specific parameters (see table \@ref(tab:traits)) related to functional trait values which can be sampled on the field.
+Carbon assimilation is computed over half-hourly period of a representative day.
+Then allocation is computed to simulate tree growth from an explicit carbon balance (in contrast to previous models).
+Finally environment is updated at each time step set to one month.
+Seedlings are not simulated explicitly but as a pool.
+In addition below-ground processes, herbaceous plants, epiphytes and lianas are not simulated inside TROLL.
+The source code is written in C++ and available upon request.
+All modules of TROLL models are further detailed in @Marechaux2017.
+
+```{r traits, echo=FALSE}
+table <- data.frame(
+  Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
+  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breasth height threshold', 'asymptotic height', 'parameter of the tree-height-dbh allometry'),
+  Units = c('$g.m^{-2}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$g.cm^{-3}$', '$m$', '$m$', '$m$')
+)
+knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Marechaux2017 Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.', format = 'pandoc')
+```
+
+Previous implementation of TROLL model used @Reich1991a allometry to infer leaf lifespan $LL$ from species leaf mass per area $LMA$ [@Marechaux2017].
+But the use of the allometrie from @Reich1991a with current implementation of the TROLL model resulted in an underestimation of leaf lifespan for low LMA species.
+Consequently we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1).
+
+We decided to explore the resilience of forest ecosystem following a generic disturbance.
+Disturbances impacting forest ecosystems are diverse as ...
+Disturbance module inside TROLL was designed in the simplest way in order to relate the ecosystem answer to volume loss without any individuals nor species targeting.
+For a given iteration $disturb_{iter}$, individuals are picked randomly with a uniform law on the number of trees.
+Selected individuals are then removed without triggering a treefall to avoid any side effect.
+The operation is repeated until the disturbance result in a defined lost basal area ($disturb_{intensity}$ in % of BA).
+
+## Design of experiment
+
+In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance and biodiversity.
+Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), using default parameters.
+Biodiversity was integrated with two of its components: taxonomic and functional diversities.
+We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species).
+Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean.
+Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess them after running the simulations, i.e.
+before applying the disturbance.
+Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$.
+For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means.
+Effectively, we did not wanted drastic change in community means that could have more effect than functional richness itself.
+This design of experiments resulted in 60 communities ($3~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Marechaux2017]).
+Functional diversities of mature forests were assessed with @villeger_new_2008 indices (FRIC, FEve, FDiv, and FDis).
+
+## Ecosystem response analysis
+
+### Ecosystem functions
+
+Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions.
+We decided to describe simulated tropical forests in two major functions: forest structure and forest functioning.
+Forest structure was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$).
+Forest functioning was represented by growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
+Variables representing both forest functioning and forest structure are sometimes highly correlated within each group.
+But multiple variables allow a better assessment of ecosystem equilibrium.
+
+The resilience of metrics values post disturbance were assessed through @Henry2012 formula:
+
+\begin{equation}
+  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X_T(t)}{X_C(t)}
+  (\#eq:Resilience)
+\end{equation}
+
+The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$.
+But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic.
+Consequently, to encompass undisturbed ecosystem variations through time, we simulated an undisturbed control ecosystem $C$.
+And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ (the control simulations being relatively stable due to a small stochasticity among TROLL simulations [@Marechaux2017]).
+Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics.
+$R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value.
+Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$.
+Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest structure (AGB, BA, N, N10, and N30) and forest functioning (GPP, NPP, Rday, and Rnight).
+We then used integrated euclidean distance to equilibrium over time to assess simulations resilience.
+
+We transformed all ecosystem outputs from the 240 disturbance simulations in resilience metrics normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A and B).
+We then gathered ecosystem outputs by main ecosystem functions (forest structure and forest functioning) to compute ecosystem distance to equilibrium (Figure \@ref(fig:datatrans) C).
+Finally, we integrated distance to equilibrium in a cumulative sum over time (Figure \@ref(fig:datatrans) D).
+The ranking of simulations regarding disturbance intensity was stable over time for the 240 simulations.
+So we used the cumulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience.
 
@@
 ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
-library(dplyr)
+cols <- RColorBrewer::brewer.pal(3, "YlOrRd")
 g_agb <- ecosystem %>% 
@@
   ylab('Aboveground biomass (tonC/ha)') +
-  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
+  scale_color_manual('Disturbance\nintensity\n(% of BA)',
+                       values=c("#999999", cols))
 g_resilience <- resilience %>% 
@@
   ylab('Resilience of aboveground biomass') +
-  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
+  scale_color_manual('Disturbance\nintensity\n(% of BA)',
+                     values = cols)
 g_deq <- deq %>% 
@@
   filter(cvh == 1) %>% 
-  ggplot(aes(x = time, y = structure, color = disturbance)) +
+  ggplot(aes(x = time, y = structure, 
+             color = disturbance, fill = disturbance)) +
   geom_hline(yintercept = 0, col = 'red') +
   geom_point() +
+  geom_area(alpha = 0.2) +
   xlab('Time (years)') +
   ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
-  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
+  scale_color_manual('Disturbance\nintensity\n(% of BA)',
+                     values = cols) +
+  scale_fill_manual('Disturbance\nintensity\n(% of BA)',
+                     values = cols)
 g_Ieq <- Ieq %>% 
@@
   ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
-  scale_color_discrete('Disturbance\nintensity\n(% of BA)')
+  scale_color_manual('Disturbance\nintensity\n(% of BA)',
+                     values = cols)
 cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
@@
 
-The ranking was stable over time for the 240 simulations.
-So we used the cummulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience.
-We compared cummulative integral after 600 years to communities taxonomic and functional diversity for each leavel of disturbance (see Figure \@ref(fig:gIeq)).
-We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cummulative integral from ecosystem distance to forest structure equilibrium after 600 years ($Ieq_{600}$).
+### Biodiversity effect
+
+Biodiversity is not only a facet of the experimental design and an ecosystem output through forest diversity, but also interact on ecosystem functioning and consequently on its answer to disturbance.
+Biodiversity ecosystem functioning relation can be split in complementarity and selection effect with @Loreau2001 partitioning:
+
+\begin{equation}
+  \begin{array}{c}
+    NE = X_O - X_E = CE + SE \\
+    CE = N* \overline{\Delta RX} \overline{M}\\
+    SE = N*cov(\Delta RX,M)
+  \end{array}
+  (\#eq:BiodivPart)
+\end{equation}
+
+Biodiversity net effect $NE$ is based on the difference between ecosystem variable $X$ observed value $X_O$ within the community mixture of species and its expected value $X_E$ if species performance were equal to their performance in monocultures.
+This effect can be partitioned between complementarity effect $CE$, representing niche partitioning, positive interactions, and resource supply, and selective effect $SE$ due to dominant species pool driving the ecosystem.
+$N$ represents the total number of species, and $M$ the vector of monocultures performance.
+Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:
+
+\begin{equation}
+  \Delta RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
+  (\#eq:DeltaRY)
+\end{equation}
+
+$X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$.
+$P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance.
+Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives).
+Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016].
+But similarly to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
+So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:
+
+\begin{equation}
+  R(NE_T) = \frac{NE_T}{NE_C} = \frac{SE_T}{NE_C} + \frac{CE_T}{NE_C}
+  (\#eq:RNE)
+\end{equation}
+
+Resilience trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
+In order to do that, the design of experiment was repeated for the each of the 163 species individually representing 652 simulations of monoculture.
+
+# Results
+
+## Ecosystem functions
+
+We compared cumulative integral after 600 years to communities taxonomic and functional diversity for each level of disturbance (see Figure \@ref(fig:gIeq)).
+We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cumulative integral from ecosystem distance to forest structure equilibrium after 600 years ($Ieq_{600}$).
 In addition, functional evenness was complementary reducing $Ieq_{600}$.
 Finally species richness was not directly link to $Ieq_{600}$.
-Effectivelly, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$.
-We found similar results for all disturbance levels and forest functionning (see [Appendix 5: Disturbance simulations]).
+Effectively, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$.
+We found similar results for all disturbance levels and forest functioning (see [Appendix 5: Disturbance simulations]).
 
@@
 
-### Biodiversity effect
+## Biodiversity effect
 
@@
 The net effect was then partition between selection and complementarity effect.
-We normalized complementarity and selection effect of disturbed simulations by biodiversity net effect of undisturbed control simulation (see Table \@ref(tab:tNE)) to measure the resilience to disturbance of their aboveground biomass (see Figure \@ref(fig:gBE)).
-We found that complementarity effect was recovering biodiversity net effect and was stronger than selection effect in the first decades.
-But after few decades the complementarity effect diminsihed toward a low value.
+We normalized complementarity and selection effect of disturbed simulations by biodiversity net effect of undisturbed control simulation (see Table \@ref(tab:tNE)) to measure the resilience of each ecosystem outputs, e.g.
+aboveground biomass (see Figure \@ref(fig:gBE)).
+We found that complementarity effect was recovering biodiversity net effect on aboverground biomass and was stronger than selection effect in the first decades.
+But after few decades the complementarity effect diminished toward a low value.
 On the contrary selection effect was reduced or even removed by the disturbance.
@@
 The time lag for which complementarity effect was greater than selection effect was increasing with disturbance intensity.
-Finally 600 years after the disturbance event biodiversity net effect was still not recovered for a disturbance intensity greater than 25% of basal area.
-We obtained those results for aboveground biomass.
-We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]).
-Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 5: Disturbance simulations]).
+Finally 600 years after the disturbance event biodiversity net effect on aboveground biomass was still not recovered for a disturbance intensity greater than 25% of basal area.
+We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see Supplementary materials S2).
+Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionally to disturbance), and its net effect was maintained by complementarity effect (see Supplementary materials S2).
 
@@
              '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$'))) %>% 
-  kable(caption = 'Table 1: Biodiversity net effect mean value and standard deviation for different ecosystem variable.', digits = 3, format = 'pandoc')
+  kable(caption = 'Table 1: Mean value and standard deviation of biodiversity net effect accross mature simulations before the disturbance event for different ecosystem variable.', digits = 3, format = 'pandoc')
 ```
@@
 Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity.
-But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012.
-Because if the harvesting is not sustainable negative feedbacks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.
-
-## TROLL limits
-
-Belowground processes, herbaceaous plants, epyphytes and lianas are not simulated in TROLL but they not reprensent the only limit of the model.
-Other processes are simulated but simplified, and we used sensitivity analysis to assess their relative importance.
-We found that few functional traits were influencing whole forest structure and dynamic.
-In addition, we found that the seed rain constant had an important effect on species functional composition and diversity.
-High external seed rain resulted in a quick recovery of the system toward an equilibrium close to the regional species frequency levied by the seed rain.
-On the contrary, low seed rain let the simulated forest works as a closed system with more system feedback but a lower stability through time with longer species diversity transitions.
-In order to study the role of diversity in forest resilience we decided to remove the seed rain constant to get a closed system and look at the role of diversity when it maintains itself through feedbacks and not with immigration.
-
-Finally, sylviculture module implemented inside TROLL showed lacks of ability to reproduce selective logging with current design of experiments.
-Results did not seem to be usable and will not be discussed.
-The main issues was the ability of the model to simulate mature forest with correct abundancy of mature trees from commonly harvested species.
-Two solutions might be possible in the tunning of the model.
-First, forest inventories could be used to initialize the model instead of a bare soil resulting in realistic species abundances, especially for harvested species.
-Secondly, selective logging could focus on guilds of species meeting peculiar values of functional traits [as wood density, see @Huth2004; @Khler2004; @Ruger2008] allowing the model to harvest more volume in order to meet reality.
+But this conclusion should met a sustainable definition of selective logging following @Zimmerman2012.
+Because if the harvesting is not sustainable negative feedback will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.
 
@@
 
-Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectivelly).
+Our results validated the hypothesis of a significative positive relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectively).
 Thus we were able to show that diversity improve forest resilience.
-More particularly, functional diversity seems a major aspect of resilience if its strengthened by a high functional evenness.
-Effectively high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species.
+More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008] seems a major aspect of resilience if its strengthened by a high functional evenness [regularity of the distribution of abundance in this volume, @villeger_new_2008].
+Indeed, high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species.
 Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes.
 Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis.
-Finally, species and functional richnesses are not directly increasing resilience.
-But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998].
+Finally, species and functional richness are not directly increasing resilience.
+But increased species and functional richness will increase chance for high functional diversity through the sampling effect [@Loreau1998].
 An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.
@@
 We interpreted this results as the consequence of facilitation processes.
-As the only ressource simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps.
+As the only resource currently simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps.
 Our results confirm the study of @Morin2011 who also highlighted the importance of facilitation through light shading in forest resilience.
-But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species.
-The diminishing of complementarity effect is due to competitive selection though time in gaps succession.
+But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an enforced dominance of more productive species.
+The diminishing of complementarity effect is attributed to competitive selection though time in gaps succession.
 But the study scale matters as shown by @Chisholm2013.
@@
 They also found that selection effect was greater than complementarity in most of the cases.
-In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield).
+In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in Guyana shield).
 High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two cutting cycles in order to maintain both productive ecosystem and sustainable management.
@@
 We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance.
-We found that diversity improved forest resilience together with productivity [@Liang2016].
+We found that diversity improved forest resilience, similarly to productivity [@Liang2016].
 Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery.
@@
 Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience.
-But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedbacks.
+But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedback.
 @Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces".
-Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy.
+Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the international community.
 This view correspond to the high resilience of tropical forest hyperdiverse systems, we were able to show, and let hope for a future sustainable selective logging in tropical forests.

### 2018-02-20 draft.Rmd
Draft 0.1.9002 logging removal, different disturbances & discussion

@@
 ---
-title: Draft
-author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
+title: Functional diversity improves tropical forest resilience to disturbance
+author: "Sylvain Schmitt, Bruno Hérault, et al"
 date: '`r Sys.Date()`'
@@
     includes:
-      before_body: doc_prefix.tex
-      in_header: preamble.tex
+      before_body: ./template/doc_prefix.tex
+      in_header: ./template/preamble.tex
     number_sections: false
@@
     toc_float: yes
+linestretch: 1.5
 csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
@@
 
-**To be written (but eventually re-use [master-thesis one](https://sylvainschmitt.github.io/master-thesis/resume-et-abstract.html))**
+**To be written**
 
@@
 
+Natural disturbance trough treefall is a main factor of forest dynamic [@Veblen1992].
+Still, climate and human activities are also at the origin of forest disturbances, with huge and underevaluated impacts.
+Human-induced forest disturbances will result either in forest degradation or in deforestation.
+Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.
+Degradation can be the result of forest logging and management, whereas deforestation can be realized through total afforestation or fragmentation resulting in land use change.
+@Pearson2017 encompassed selective timber harvest in native forests, woodfuel harvest and fire in tropical forest degradation.
 Tropical forests disturbances, through deforestation and degradation, account for 8.26 billion of tons of carbon dioxide emissions per year [@Pearson2017].
@@
 Degradation from forest has been estimated to represents 10 time the deforestation area [@Herold2011] and represents 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].
-Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.
+In addition to the direct impact of human on forests, a more insidious disturbance is climate change [@Herault2016].
 
-Sustainable forest management in the tropics (i.e.
-managed selective harvesting of timber) has been widely promoted internationally to combat tropical deforestation and other forms of degradation [@Zimmerman2012].
-Currently logging from tropical forests accounts for one eighth of global timber production [@Blaser2011] and is still increasing.
-Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cutting cycle [@Martin2015].
+Regarding deforestation, the ecosystem response to land use change is almost absent due to the negative feedback of the new ecosystem state after the tipping point was reached.
+Consequently, ecosystem response to disturbance should be studied in the different contexts of forest degradation.
+Concerning timber and woodfuel harvest, @DeAvila2015 found that additional sylvicultural practices and logging damages had stronger negative effect on diversity than logging itself.
+@Carreno-Rocabado2012 found no direct effect of medium intensity logging on species diversity but a community drift towards 'fast' acquisitive species [@reich_world-wide_2014].
+However, even the short term impact on biodiversity is unclear, as sometimes short term increases in diversity can occur at low sylvicultural intensities [@Martin2015].
+Forest fires have been shown to increase pioneers abundance, and if recurrent, to impoverish the forests in the eastern Amazon [@Cochrane1999a].
+Finally, climate change effect on tropical areas results in a significant risk of precipitation decrease [@Davidson2012].
+In addition, intense drought events has been shown to result in huge mortality rates and biomass loss [@Phillips2009].
 
-On the other hand, tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
+Tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
 Effectively tropical forests host over half of the Earth's biodiversity [@Scheffers2012].
@@
 
-But several authors argue that selective logging represents a major threat to biodiversity [@Carreno-Rocabado2012; @DeAvila2015; @Gibson2013; @Martin2015; @Zimmerman2012], challenging the sustainable definition from current selective logging.
-@DeAvila2015 found that additional sylvicultural practices and logging damages had stronger negative effect on diversity than logging itself.
-@Carreno-Rocabado2012 found no direct effect of medium intensity logging on species diversity but a community drift towards 'fast' acquisitive species [@reich_world-wide_2014].
-However, even the short term impact on biodiversity is unclear, as sometimes short term increases in diversity can occur at low sylvicultural intensities [@Martin2015].
-We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better sylvicultural practices in order to reach sustainability.
+The question of disturbance impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functioning [@Loreau2000].
+Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to disturbance.
+And forest ecosystem response to disturbance will directly modify ecosystem functioning in both short and long term.
+Consequently assessing disturbance effect on tropical forest linking diversity and ecosystem functioning seems an obvious and promising way [@Loreau2010].
 
-The question of selective logging impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functioning [@Loreau2000].
-Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to logging disturbance.
-And forest ecosystem response to logging disturbance will directly modify ecosystem functioning in both short and long term.
-Consequently assessing selective logging effect on tropical forest linking diversity and ecosystem functioning seems an obvious and promising way [@Loreau2010].
-
-Negative short term impacts of selective logging have been assessed [@Carreno-Rocabado2012; @DeAvila2015;  but see @Martin2015].
-Much less is known about the long term impact [@Osazuwa-Peters2015].
+Negative short term impacts of disturbances have been assessed for logging, fires and droughts ; but much less is known about the long term impact [@Osazuwa-Peters2015].
 The main reason is the difficulty to conduct long term empirical study [but see @Herault2010], which can be completed by the use of forest simulators [@Huth2004; @Khler2004; @Ruger2008; @Tietjen2006].
@@
 TROLL model relates physiological processes to species-specific functional traits [@Marechaux2017].
-Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functioning link response to logging disturbance.
+Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functioning link response to a disturbance.
 
@@
 In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
-We used a simulation approach using TROLL model to assess long term processes for different types and levels of disturbances.
+We used a simulation approach using TROLL model to assess long term processes for different levels of disturbances.
 We first assessed diversity effect on forest resilience of structure and functioning using numerous indices of taxonomic and functional diversities.
@@
 We decided to explore the resilience of forest ecosystem following a generic disturbance.
-Disturbances impacting forest ecosystems are diverse as ...
 Disturbance module inside TROLL was designed in the simplest way in order to relate the ecosystem answer to volume loss without any individuals nor species targeting.
@@
 ```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
+Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].
+Solid lines and grey areas show fitted relationships with their mean and variance of the form $Ieq_600 = FDiv + \\epsilon$.', cache=TRUE, fig.width=12}
 Ieq %>% 
@@
 ```{r gBE, echo=FALSE, fig.cap='Resilience of complementarity and selection effects.
-Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time.', cache=TRUE, fig.width=12}
+Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time.
+Solid lines represents the mean value whereas areas represents the interval of values among simulations.', cache=TRUE, fig.width=12}
 BEresilience %>% 
@@
 Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery.
-Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity.
-But this conclusion should met a sustainable definition of selective logging following @Zimmerman2012.
-Because if the harvesting is not sustainable negative feedback will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.
+We thus advocate the view that both diversity theory (with niche complementarity and facilitation) and mass ratio theory (represented by selection) represents the mechanisms at the origin of functional diversity effect on ecosystem functions depending on the spatial scale and the time since the disturbance event.
 
@@
 Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis.
-Finally, species and functional richness are not directly increasing resilience.
-But increased species and functional richness will increase chance for high functional diversity through the sampling effect [@Loreau1998].
+Finally, species and functional richness are not directly increasing resilience, thus supporting the idea that species diversity is often an inadequate surrogate of functional diversity [@Diaz2001; but see @Poorter2015].
+Still, increased species and functional richness will increase chance for high functional diversity through the sampling effect [@Loreau1998].
 An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.
@@
 The diminishing of complementarity effect is attributed to competitive selection though time in gaps succession.
-But the study scale matters as shown by @Chisholm2013.
-Here we look at processes to a 16 ha scale.
-We do not have topography reducing micro-environment effects, but @Chisholm2013 results suggest that complementarity will be stronger at smaller scale and could explain its low value after forest recovery.
-Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada.
-They also found that selection effect was greater than complementarity in most of the cases.
-In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in Guyana shield).
-High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two cutting cycles in order to maintain both productive ecosystem and sustainable management.
+
+In addition, the study scale matters as shown by @Chisholm2013 and @Sullivan2017.
+Here we look at processes to a 16-ha scale.
+On the other hand, our simulation does not include topography resulting in the absence of microenvironmental variation affecting both water and nutrient availability.
+And studies suggest that complementarity will be stronger at smaller scale [@Chisholm2013; @Sullivan2017], within plot itself (0.4-ha scale).
+This could explain the low value of complementarity after forest recovery.
+On the opposite, @Sullivan2017 suggested that environment at larger scale could obscure the positive relationship between functional diversity and ecosystem functions and explain why the relation is detected at 0.4-ha scale and not at 1-ha scale.
+Consequently, the absence of topography and environmental heterogeneity in our simulations could be the reason why we still observed an effect to a 16-ha scale, strengthening @Sullivan2017 interpretation.
+
+Our results follows several studies advocating for multiple mechanisms through which functional diversity affect ecosystem functions.
+Complementarity effect encompass niche complementarity theory (or diversity theory), which has been shown to indirectly improve forest biomass and coarse wood productivity in subtropical forest [@Chiang2016].
+Similarly, complementarity has been shown of primary importance in grassland functioning [@Cardinale2007; @Isbell2011].
+On the other hand, selection effect can be related to the mass ratio hypothesis, which proposes that ecosystem functions are regulated by dominant species of the community [@Grime1998].
+And mass ratio theory has been shown to primarily influence ecosystem functions with biomass accumulation (aboveground and shoots), productivity (coarse wood or aboveground), water use, and light interception for both grassland and subtropical forests [@Chiang2016; @Mokany2008].
+Consequently, concerning mechanisms responsible for the positive effect of functional diversity on tropical forest functions after disturbance, we also advocate for the interplay of both niche complementarity and mass ratio theories, depending on time since the disturbance event and the spatial scale.
+
+Finally, our findings confirm results from experimental studies.
+@Williams2017 also found experimentally an overyielding due to complementarity effect based on light and spatial complementarity with tree crown of young trees.
+In addition, @Tobner2016 studied 4-years stand in low diverse forests of Canada.
+They also found that selection effect was greater than complementarity in most of the cases, promoting the idea that mass ratio theory is the primary mechanisms involved in functional diversity effect on ecosystem functions [@Mokany2008].
 
@@
 We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance.
-We found that diversity improved forest resilience, similarly to productivity [@Liang2016].
+We found that diversity improved forest resilience, similarly to productivity [@Liang2016; @Poorter2015].
 Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery.
-Our results advocate for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles.
-Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience.
-But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedback.
-@Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces".
-Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the international community.
-This view correspond to the high resilience of tropical forest hyperdiverse systems, we were able to show, and let hope for a future sustainable selective logging in tropical forests.
 

### 2018-03-30 draft.Rmd
Update draft.Rmd
BH

@@
 title: Functional diversity improves tropical forest resilience to disturbance
-author: "Sylvain Schmitt, Bruno Hérault, et al"
+author: "Sylvain Schmitt, Jerome Chave, Fabien Fischer, Isabelle Maréchaux, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
 date: '`r Sys.Date()`'

### 2018-03-30 S1.Rmd
Draft 0.1.9003 md files for @BrunoHerault review

--- 2017-07-13 08-Appendix2.Rmd
@@
-# Appendix 2: Leaf lifespan model
+---
+title: "Supplementary material S1: Leaf lifespan model"
+date: '`r Sys.Date()`'
+output:
+  github_document: default
+  bookdown::pdf_document2:
+    number_sections: false
+    toc: false
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
+csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
+bibliography: /home/sylvain/Documents/Bibliography/library.bib
+link-citations: yes
+---
 

### 2018-03-30 S2.Rmd
Draft 0.1.9003 md files for @BrunoHerault review

@@
 output:
+  github_document: default
   bookdown::pdf_document2:

### 2018-03-30 draft.Rmd
Draft 0.1.9003 md files for @BrunoHerault review

--- 2018-02-20 draft.Rmd
@@
 output:
+  github_document: default
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     toc: false
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
+    keep_tex: true
 linestretch: 1.5

### 2018-03-30 draft.Rmd

--- 2018-02-20 draft.Rmd
@@
 title: Functional diversity improves tropical forest resilience to disturbance
-author: "Sylvain Schmitt, Bruno Hérault, et al"
+author: "Sylvain Schmitt, Jerome Chave, Fabien Fischer, Isabelle Maréchaux, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
 date: '`r Sys.Date()`'
 output:
+  github_document: default
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
+  bookdown::word_document2: default
   bookdown::pdf_document2:

### 2018-04-23 draft.Rmd
Draft 0.1.9004 Third version for IM review

--- 2018-03-30 draft.Rmd
@@
 title: Functional diversity improves tropical forest resilience to disturbance
-author: "Sylvain Schmitt, Bruno Hérault, et al"
+author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabien Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
 date: '`r Sys.Date()`'
 output:
-  github_document: default
   bookdown::html_document2:
@@
     toc_float: yes
-  bookdown::word_document2: default
+  github_document: default
   bookdown::pdf_document2:
@@
     keep_tex: true
+  bookdown::word_document2: default
 linestretch: 1.5
@@
 
-Natural disturbance trough treefall is a main factor of forest dynamic [@Veblen1992].
-Still, climate and human activities are also at the origin of forest disturbances, with huge and underevaluated impacts.
-Human-induced forest disturbances will result either in forest degradation or in deforestation.
-Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.
-Degradation can be the result of forest logging and management, whereas deforestation can be realized through total afforestation or fragmentation resulting in land use change.
-@Pearson2017 encompassed selective timber harvest in native forests, woodfuel harvest and fire in tropical forest degradation.
-Tropical forests disturbances, through deforestation and degradation, account for 8.26 billion of tons of carbon dioxide emissions per year [@Pearson2017].
-It represents the third source of greenhouse gas.
+Natural disturbances caused by the death of one or more trees, are the dominant forms of forest renewal in tropical forests as the creation of canopy gaps continuously modifies forest horizontal and vertical structure [@Veblen1992].
+Today, climate change and other direct human activities are modifying forest disturbance regimes, with huge and under evaluated impacts.
+When human induced disturbances are high, it is referred to as "degraded forests" characterized by a modification of the forest ecosystem functioning, possibly reducing the furnished ecosystem services, while keeping the ecosystem as a forest, contrary to deforestation [@Simula2009].
 Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics.
-Degradation from forest has been estimated to represents 10 time the deforestation area [@Herold2011] and represents 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].
-In addition to the direct impact of human on forests, a more insidious disturbance is climate change [@Herault2016].
+The main source of degradation around the tropics is forest logging that encompass selective timber harvest, in native forests, and wood fuel harvest, often in secondary forests [@Pearson2017].
+Tropical forests disturbances account for 2.07 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005], estimating to cover areas 10 or more times larger than deforestation itself [@Herold2011].
+Finally, disturbance in tropical forest dynamics induced by climate change likely results from significant changes in precipitation regime [@Davidson2012] with intense drought events resulting in huge mortality rates and biomass loss [@Phillips2009].
 
-Regarding deforestation, the ecosystem response to land use change is almost absent due to the negative feedback of the new ecosystem state after the tipping point was reached.
-Consequently, ecosystem response to disturbance should be studied in the different contexts of forest degradation.
-Concerning timber and woodfuel harvest, @DeAvila2015 found that additional sylvicultural practices and logging damages had stronger negative effect on diversity than logging itself.
-@Carreno-Rocabado2012 found no direct effect of medium intensity logging on species diversity but a community drift towards 'fast' acquisitive species [@reich_world-wide_2014].
-However, even the short term impact on biodiversity is unclear, as sometimes short term increases in diversity can occur at low sylvicultural intensities [@Martin2015].
-Forest fires have been shown to increase pioneers abundance, and if recurrent, to impoverish the forests in the eastern Amazon [@Cochrane1999a].
-Finally, climate change effect on tropical areas results in a significant risk of precipitation decrease [@Davidson2012].
-In addition, intense drought events has been shown to result in huge mortality rates and biomass loss [@Phillips2009].
+Tropical forest ecosystem response to disturbance depends first on the disturbance intensity itself.
+For instance, @DeAvila2015, focusing on timber and wood fuel harvest in Central Amazonia, found that when disturbance is high, e.g.
+induced by post logging sylvicultural practices had stronger negative effect on the post-disturbance system trajectory than logging itself.
+@Carreno-Rocabado2012 found no direct effect of disturbance intensity logging on community diversity itself but a community composition drift towards 'fast' acquisitive species [@reich_world-wide_2014].
+However, even the short-term impact on biodiversity is unclear, as sometimes short term increase in diversity can occur at low disturbance intensities [@Martin2015].
 
-Tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].
-Effectively tropical forests host over half of the Earth's biodiversity [@Scheffers2012].
-High biodiversity from tropical rainforests is the source of many ecosystem functions.
-Among others, tropical forests play a key role in biogeochemical cycles, including carbon storage [@Lewis2004].
-Ecosystem functions from tropical forests support numerous ecosystem services, such as timber production and climate regulation.
+The main reason of our lack of knowledge is the difficulty to conduct long term empirical study so that the use of forest simulators has become inevitable [@Huth2004].
+Individual-based models of forest dynamics with species-explicit model parameters present the perfect framework to develop such joint biodiversity-ecosystem functioning studies [@Marechaux2017].
+Starting from scratch, these models allow describing forest recovery through time, assessing individual tree recruitment, growth and mortality [@Bugmann2001].
+Managing diversity in such models is still challenging and many models have built open Plant Functional Types to manage the volume of model parameters.
+[@Pacala1996].
 
-The question of disturbance impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functioning [@Loreau2000].
-Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to disturbance.
-And forest ecosystem response to disturbance will directly modify ecosystem functioning in both short and long term.
-Consequently assessing disturbance effect on tropical forest linking diversity and ecosystem functioning seems an obvious and promising way [@Loreau2010].
-
-Negative short term impacts of disturbances have been assessed for logging, fires and droughts ; but much less is known about the long term impact [@Osazuwa-Peters2015].
-The main reason is the difficulty to conduct long term empirical study [but see @Herault2010], which can be completed by the use of forest simulators [@Huth2004; @Khler2004; @Ruger2008; @Tietjen2006].
-Individual-based models of forest dynamics present the perfect framework to develop such joint biodiversity-ecosystem approaches [@Marechaux2017].
-Individual-based models describe forest accumulating carbon through time, assessing tree growth, or releasing carbon through gap opening [@Bugmann2001].
-Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit [@Pacala1996].
-Recently, the forest growth simulator TROLL [@Chave1999], an individual-based and spatially explicit forest model, was developed to introduce recent advances in plant ecophysiology.
-TROLL model relates physiological processes to species-specific functional traits [@Marechaux2017].
-Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functioning link response to a disturbance.
-
-We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance.
-Resilience encompass several definitions but was summarized by @Oliver2015 as the degree of resistance or fast recovery from an ecosystem function to environmental disturbance.
+Recently the forest growth simulator TROLL [@Chave1999], an individual-based and spatially-explicit forest model has been explicitly developed to deal with high diversity levels by using plant ecophysiological traits as surrogates of model parameters [@Marechaux2017].
+Consequently, TROLL allow to fully simulate a highly-diverse system and offers the opportunity to study biodiversity-ecosystem functioning *in silico*.
 
 Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity.
-We assumed that when a disturbance event happen, due to a higher productivity, forest with an increased diversity will recover quicker and thus be more resilient, besides other mechanisms could be involved in an increased resilience from diverse forest.
+We assumed that when a disturbance event happen, due to a higher productivity, more diverse forests will recover quicker and thus be more resilient.
 This theoretical expectation stand on two processes: complementarity and selection effects [@Loreau2001].
-If we take a species pool with different productivity in monoculture, their assemblage will allow them an overall higher productivity due to a better resources acquisition through resources partitioning and niche differentiation.
-Additionally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage.
-Complementarity effect is the addition of the better resource acquisition and the facilitation.
-Now if we look at the evolution of the species assemblage over time, more competitive species will progressively dominate through competitive selection.
-And if competitive species are more productive, they will increase assemblage overall productivity.
-This resources preemption by more competitive species is the selection effect.
-Moreover, we expect the sampling effect to improve resilience of species-rich forests.
-A bigger initial sampling of the regional species pool in a rich forest, will allow more redundancy and a lower risk to lose important functional traits assemblages for the ecosystem when the disturbance happen.
+Complementarity effect is the addition of better resources acquisition, due to resources partitioning and niche differentiation, and facilitation in diverse communities.
+Selection effect results from resources preemption by more competitive species.
+Moreover, we expect the sampling effect to improve resilience of species-rich forest, with redundancy and a lower risk to lose important functional traits.
 
@@
 
-We suggest to use a process-based model to overcome those limitations because TROLL model have the same key features that @Morin2011: (i) TROLL has been developed and validated independently from the biodiversity and ecosystem functioning issue, (ii) TROLL use 163 real tropical tree species characterized by their functional traits, (iii) TROLL allows to study long term effects with realized species richness for several centuries, (iv) TROLL allows to explore more species associations than real experiments, (v) TROLL allows to explore process resulting in observed relations.
-In addition, TROLL has been designed to be used in species-rich ecosystems such as tropical rainforests.
-
-In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
+In the present study, we thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
 We used a simulation approach using TROLL model to assess long term processes for different levels of disturbances.
 We first assessed diversity effect on forest resilience of structure and functioning using numerous indices of taxonomic and functional diversities.
-Then, we measured biodiversity net effect, partitioned into complementarity and selection effects, on resilience for several ecosystem metrics.
+Then, we measured biodiversity net effect, partitioned into complementarity and selection effects, on resilience of simulated tropical forest communities to disturbance.
 
@@
 
-## Model
+## TROLL Model
 
-TROLL model each tree individually in a located environment.
-Thus TROLL model, alongside with SORTIE  [@Pacala1996; @Uriarte2009] and FORMIND [@Fischer2016; @Kohler1998], can be defined as an individual-based and spatially explicit forest growth model.
+TROLL is an individual-based and spatially explicit forest growth model.
 TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production.
@@
 Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
-A species label is associated to a number of species specific parameters (see table \@ref(tab:traits)) related to functional trait values which can be sampled on the field.
+A species label is associated to a number of species specific parameters (Table \@ref(tab:traits)) related to functional trait values.
 Carbon assimilation is computed over half-hourly period of a representative day.
-Then allocation is computed to simulate tree growth from an explicit carbon balance (in contrast to previous models).
-Finally environment is updated at each time step set to one month.
-Seedlings are not simulated explicitly but as a pool.
+Then allocation is computed to simulate tree growth from an explicit carbon balance.
+Environment is updated at each time step set to one month.
+Seed-rains and seedlings are not simulated explicitly but as a pool.
 In addition below-ground processes, herbaceous plants, epiphytes and lianas are not simulated inside TROLL.
@@
 All modules of TROLL models are further detailed in @Marechaux2017.
+Following an underestimation of leaf lifespan for low LMA species, we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1).
 
@@
 
-Previous implementation of TROLL model used @Reich1991a allometry to infer leaf lifespan $LL$ from species leaf mass per area $LMA$ [@Marechaux2017].
-But the use of the allometrie from @Reich1991a with current implementation of the TROLL model resulted in an underestimation of leaf lifespan for low LMA species.
-Consequently we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1).
-
-We decided to explore the resilience of forest ecosystem following a generic disturbance.
-Disturbance module inside TROLL was designed in the simplest way in order to relate the ecosystem answer to volume loss without any individuals nor species targeting.
-For a given iteration $disturb_{iter}$, individuals are picked randomly with a uniform law on the number of trees.
+The disturbance module inside TROLL was designed in the simplest way by, for a given iteration $disturb_{iter}$, picking individual trees randomly up with a uniform law and without any individuals nor species targeting.
 Selected individuals are then removed without triggering a treefall to avoid any side effect.
-The operation is repeated until the disturbance result in a defined lost basal area ($disturb_{intensity}$ in % of BA).
+The operation is repeated until the disturbance result in a defined lost basal area ($disturb_{intensity}$).
 
@@
 
-In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance and biodiversity.
-Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), using default parameters.
-Biodiversity was integrated with two of its components: taxonomic and functional diversities.
-We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species).
-Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean.
-Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess them after running the simulations, i.e.
-before applying the disturbance.
-Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$.
-For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means.
-Effectively, we did not wanted drastic change in community means that could have more effect than functional richness itself.
-This design of experiments resulted in 60 communities ($3~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Marechaux2017]).
-Functional diversities of mature forests were assessed with @villeger_new_2008 indices (FRIC, FEve, FDiv, and FDis).
+In order to assess the role of biodiversity in ecosystem response to disturbance, we  simulated post-disturbance forest trajectories, encompassing both variation of disturbance intensity and biodiversity levels.
+The gradient of disturbance intensity was represented by percentage of basal area loss (0%, 25%, 50% and 75%).
+Biodiversity was assessed with two perspectives: taxonomic and functional.
+We used species richness $SR$ to simulate initial assemblages with diverse taxonomic diversity (5, 25, and 125 species).
+Functional diversity was assessed after running the initial simulations, *i.e.* before applying the disturbance.
+We focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$.
+For each level of species richness $SR$, we simulated 20 communities with a large range of convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means.
+This design of experiments resulted in 60 tree communities ($3~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Marechaux2017]).
+Functional diversities at the end of simulations were assessed with @villeger_new_2008 indices (FRIC, FEve, FDiv, and FDis).
 
-## Ecosystem response analysis
+## Analysing Ecosystem Response
 
-### Ecosystem functions
+### Ecosystem structure and functioning
 
-Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions.
-We decided to describe simulated tropical forests in two major functions: forest structure and forest functioning.
-Forest structure was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$).
-Forest functioning was represented by growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).
-Variables representing both forest functioning and forest structure are sometimes highly correlated within each group.
-But multiple variables allow a better assessment of ecosystem equilibrium.
+We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).
 
-The resilience of metrics values post disturbance were assessed through @Henry2012 formula:
+The resilience of forest structure and functioning values was assessed with:
 
@@
 The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$.
-But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic.
-Consequently, to encompass undisturbed ecosystem variations through time, we simulated an undisturbed control ecosystem $C$.
-And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ (the control simulations being relatively stable due to a small stochasticity among TROLL simulations [@Marechaux2017]).
-Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics.
+But in our peculiar case of simulated forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ is dynamic in time so that we simulated an undisturbed control ecosystem $C$.
+And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ (the control simulations being relatively stable due to small stochasticity among TROLL simulations [@Marechaux2017]).
 $R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value.
-Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$.
-Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest structure (AGB, BA, N, N10, and N30) and forest functioning (GPP, NPP, Rday, and Rnight).
-We then used integrated euclidean distance to equilibrium over time to assess simulations resilience.
+Next, we calculated an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$.
+Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for (the two functions described above: (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight).
+We then integrate the euclidean distances over time to get our ecosystem resilience indicators.
 
-We transformed all ecosystem outputs from the 240 disturbance simulations in resilience metrics normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A and B).
-We then gathered ecosystem outputs by main ecosystem functions (forest structure and forest functioning) to compute ecosystem distance to equilibrium (Figure \@ref(fig:datatrans) C).
-Finally, we integrated distance to equilibrium in a cumulative sum over time (Figure \@ref(fig:datatrans) D).
-The ranking of simulations regarding disturbance intensity was stable over time for the 240 simulations.
+We then applied the following to each indicator:
+
+1. Normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A and B).
+2. Respectively gathering forest structure and forest functioning together to get synthetic indicators (Figure \@ref(fig:datatrans) C)
+3. Integration over time (Figure \@ref(fig:datatrans) D)
+4.
+The ranking of simulations regarding disturbance intensity was highly stable over time for the 240 simulations.
 So we used the cumulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience.
@@
 ```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation.
-Ecosystem outputs (**A**) are normalized by the control value over time to calculate resilience (**B**); resilience of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance ot equilibirum is integrated over time in a cummulative sum (**D**).
+Ecosystem outputs (**A**) are normalized by the control value over time to calculate resilience (**B**); resilience of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**).
 ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
@@
 
-Biodiversity is not only a facet of the experimental design and an ecosystem output through forest diversity, but also interact on ecosystem functioning and consequently on its answer to disturbance.
-Biodiversity ecosystem functioning relation can be split in complementarity and selection effect with @Loreau2001 partitioning:
+The effect of biodiversity on ecosystem resilience can be split in complementarity and selection effect using @Loreau2001 partitioning:
 
@@
 Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016].
-But similarly to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
+But similarly, to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
 So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:
@@
 Resilience trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
-In order to do that, the design of experiment was repeated for the each of the 163 species individually representing 652 simulations of monoculture.
+In order to do that, we simulated the 163 species individually representing 652 simulations of monoculture.
 
@@
 
-We compared cumulative integral after 600 years to communities taxonomic and functional diversity for each level of disturbance (see Figure \@ref(fig:gIeq)).
-We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cumulative integral from ecosystem distance to forest structure equilibrium after 600 years ($Ieq_{600}$).
+Species richness was not directly link to $Ieq_{600}$.
+Increased functional diversity reduced cumulative integral from ecosystem distance to equilibrium after 600 years ($Ieq_{600}$).
 In addition, functional evenness was complementary reducing $Ieq_{600}$.
-Finally species richness was not directly link to $Ieq_{600}$.
-Effectively, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$.
+Finally, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$.
 We found similar results for all disturbance levels and forest functioning (see [Appendix 5: Disturbance simulations]).
@@
 
-We measured all ecosystem outputs biodiversity net effect in disturbance simulations by comparing them to their species corresponding monoculture simulations.
-The net effect was then partition between selection and complementarity effect.
-We normalized complementarity and selection effect of disturbed simulations by biodiversity net effect of undisturbed control simulation (see Table \@ref(tab:tNE)) to measure the resilience of each ecosystem outputs, e.g.
+The net biodiversity effect was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation (see Table \@ref(tab:tNE)) to measure the resilience of each ecosystem measure, e.g.
 aboveground biomass (see Figure \@ref(fig:gBE)).
-We found that complementarity effect was recovering biodiversity net effect on aboverground biomass and was stronger than selection effect in the first decades.
-But after few decades the complementarity effect diminished toward a low value.
-On the contrary selection effect was reduced or even removed by the disturbance.
-It increased during the whole simulation and was greater than complementarity effect only after decades.
-The time lag for which complementarity effect was greater than selection effect was increasing with disturbance intensity.
-Finally 600 years after the disturbance event biodiversity net effect on aboveground biomass was still not recovered for a disturbance intensity greater than 25% of basal area.
+We found that complementarity effect was dominant in the biodiversity net effect in the first decades.
+But after few decades the complementarity effect diminished toward an extremely low value.
+On the contrary, selection effect was reduced in the initial times after disturbance.
+It increased during the first decades and was then greater.
+The time lag for which complementarity effect overcame the selection effect increased with disturbance intensity.
+Finally, 600 years after disturbance, the biodiversity net effect on aboveground biomass was still not at equilibrium for a disturbance intensity greater than 25% of basal area.
 We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see Supplementary materials S2).
@@
 
-In the limit of the model, we were able to show that diversity improved tropical forest resilience.
-More particularly, functional diversity and evenness are key components of diversity in forest recovery after disturbance.
-Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery.
-We thus advocate the view that both diversity theory (with niche complementarity and facilitation) and mass ratio theory (represented by selection) represents the mechanisms at the origin of functional diversity effect on ecosystem functions depending on the spatial scale and the time since the disturbance event.
+In the limit of the model, we were able to show that functional diversity improved tropical forest resilience after disturbance.
+More particularly, functional diversity and evenness are key components of diversity in forest recovery trajectory after disturbance.
+Moreover, we found that complementarity between species was insuring forest recovery in the early times before more productive species dominate the forest and insure late recovery.
+We thus advocate the view that both diversity theory (with niche complementarity and facilitation) and mas ratio theory (represented by selection) are implied in ecosystem resilience but with relative importance depending on the spatial grain and time since the disturbance event.
 
-## Diversity improve tropical forest resilience
+## Functional diversity improve tropical forest resilience
 
 Our results validated the hypothesis of a significative positive relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectively).
-Thus we were able to show that diversity improve forest resilience.
-More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008] seems a major aspect of resilience if its strengthened by a high functional evenness [regularity of the distribution of abundance in this volume, @villeger_new_2008].
-Indeed, high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species.
-Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes.
-Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis.
-Finally, species and functional richness are not directly increasing resilience, thus supporting the idea that species diversity is often an inadequate surrogate of functional diversity [@Diaz2001; but see @Poorter2015].
-Still, increased species and functional richness will increase chance for high functional diversity through the sampling effect [@Loreau1998].
+More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008] seems a major aspect of resilience if it's strengthened by a high functional evenness [regularity of the distribution of abundance in this volume, @villeger_new_2008].
+Indeed, high functional diversity combined to high evenness avoid hegemonic effect of a few dominant species.
+Our results thus confirms both the review of @Diaz2001,advocating for an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity and resilience of terrestrial ecosystems.
+Finally, species richness is not directly increasing resilience, thus supporting the idea that species diversity is often an inadequate surrogate of functional diversity [@Diaz2001; but see @Poorter2015].
+Still, increased species richness will increase chance for high functional diversity through the sampling effect [@Loreau1998].
 An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.
+
+<!-- OK good point. But you need to go more into the details of the diversity effect, perhaps highlighting the key choice of the traits? BH -->
 
@@
 We found that complementary effect was insuring forest resilience in the beginning of forest successions.
-We interpreted this results as the consequence of facilitation processes.
-As the only resource currently simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps.
-Our results confirm the study of @Morin2011 who also highlighted the importance of facilitation through light shading in forest resilience.
-But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an enforced dominance of more productive species.
-The diminishing of complementarity effect is attributed to competitive selection though time in gaps succession.
+We interpreted this result as the consequence of facilitation processes or lower competition due to different niche.
+As the only resource currently simulated by TROLL is the light, the main facilitation would be light shading of post-pioneer species by pioneer species in disturbance gaps, confirming the study of @Morin2011 who also highlighted the importance of facilitation through light shading in forest resilience.
+But the complementarity effect quickly reduced through time, to let the selection effect insure forest resilience due to an re-enforced dominance of more productive species.
+The diminishing of complementarity effect is attributed to the selection of the most competitive species though time.
 
 In addition, the study scale matters as shown by @Chisholm2013 and @Sullivan2017.
-Here we look at processes to a 16-ha scale.
-On the other hand, our simulation does not include topography resulting in the absence of microenvironmental variation affecting both water and nutrient availability.
-And studies suggest that complementarity will be stronger at smaller scale [@Chisholm2013; @Sullivan2017], within plot itself (0.4-ha scale).
-This could explain the low value of complementarity after forest recovery.
-On the opposite, @Sullivan2017 suggested that environment at larger scale could obscure the positive relationship between functional diversity and ecosystem functions and explain why the relation is detected at 0.4-ha scale and not at 1-ha scale.
-Consequently, the absence of topography and environmental heterogeneity in our simulations could be the reason why we still observed an effect to a 16-ha scale, strengthening @Sullivan2017 interpretation.
+In this work, we look at processes to a 16-ha scale, being limited by the computational effort and, our simulation did not include topography resulting in the absence of micro environmental variation affecting both water and nutrient availability.
+Because many studies suggest that, in mature ecosystems, complementarity will be stronger at smaller scale [@Chisholm2013; @Sullivan2017], this could explain the low value of complementarity after the initial stages of forest recovery.
+Still, the absence of topography and environmental heterogeneity in our simulations could explain why we observed an effect of complementarity to a 16-ha scale, strengthening @Sullivan2017 interpretation.
 
@@
 And mass ratio theory has been shown to primarily influence ecosystem functions with biomass accumulation (aboveground and shoots), productivity (coarse wood or aboveground), water use, and light interception for both grassland and subtropical forests [@Chiang2016; @Mokany2008].
-Consequently, concerning mechanisms responsible for the positive effect of functional diversity on tropical forest functions after disturbance, we also advocate for the interplay of both niche complementarity and mass ratio theories, depending on time since the disturbance event and the spatial scale.
+Consequently, concerning mechanisms responsible for the positive effect of functional diversity on tropical forest functions after disturbance, we also advocate for the interplay of both niche complementarity and mass ratio theories, depending on time since the disturbance event and the spatial grain.
 
 Finally, our findings confirm results from experimental studies.
-@Williams2017 also found experimentally an overyielding due to complementarity effect based on light and spatial complementarity with tree crown of young trees.
+@Williams2017 also found experimentally an over yielding due to complementarity effect based on light and spatial complementarity with tree crown of young trees.
 In addition, @Tobner2016 studied 4-years stand in low diverse forests of Canada.
@@
 We found that diversity improved forest resilience, similarly to productivity [@Liang2016; @Poorter2015].
-Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery.
+Additionally, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure the final forest recovery.
 

### 2018-07-13 S1-LL.Rmd
Draft 0.1.9005 Supplementary: RNE & LM

--- 2017-07-13 08-Appendix2.Rmd
@@
-# Appendix 2: Leaf lifespan model
+---
+title: "Supplementary material S1: Leaf lifespan model"
+date: '`r Sys.Date()`'
+output:
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
+  bookdown::pdf_document2:
+    number_sections: false
+    toc: false
+csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
+bibliography: /home/sylvain/Documents/Bibliography/library.bib
+link-citations: yes
+---
 

### 2018-07-13 S2-RNE.Rmd
Draft 0.1.9005 Supplementary: RNE & LM

---
title: "Supplementary material S2: Biodiversity net effect and resilience"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(tidyverse)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 8,
                cache = T, cache.lazy = F)
```

We want to merge two functions to analyse the output of ecosystem after disturbance: (i) the biodiversity net effect (NE) and its partitionning in complementarity (CE) and selection (SE) effects, and (ii) the resilience function of the ecosystem. More particularly we computed $R[NE(X)]$, and we want to know if $R[NE(X)]=NE[R(X)]$. In order to do that we will proceed in three times. First we need to define all mathematic functions. Then we will use a simulation approach to test for the equality (easy way to reject it). Finally if the simulation did not rejected the equality we will try to prove it mathematically.

# Mathematic definitions

We first need to define the resilience function, the biodiversity net effect functions, and their composed functions for any variable $X$ from the ecosystem.

## Resilience

Resilience has been defined by us as:

\begin{equation}
  R(t) = \frac{X_T(t)}{X_C(t)}
  (\#eq:R)
\end{equation}

where $X_T(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation and in the undisturbed control simulation.

## Biodiversity net effect

The net effect of biodiversity on ecosystem recovery (NE) is defined as the difference between the ecosystem recovery of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e. ecosystem recovery for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001]. We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:

\begin{equation}
  \begin{array}{c}
    NE = X_O - X_E = CE + SE \\
    CE(X) = N* \overline{\Delta _RX} \overline{M}\\
    SE(X) = N*cov(\Delta _RX,M) \\
    \Delta _R(X_{sp}) = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
  \end{array}
  (\#eq:NE)
\end{equation}

where $N$ represents the total number of species, and $M$ the vector of monoculture performance. Both metrics depend on the variation of relative ecosystem variable $\Delta _R(X)$. $X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$. $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance.

## Resilience of the biodiversity net effect $R[NE(X)]$

Using equations \@ref(eq:R) and \@ref(eq:NE) we can easily define resilience of the biodiversity net effect as:

\begin{equation}
  \begin{array}{c}
    R[NE(X)]=\frac{NE_T(X)}{NE_C(X)} \\
    R[NE(X)]=\frac{CE_T(X)}{NE_C(X)} + \frac{SE_T(X)}{NE_C(X)} \\
    CE(X) = N* \overline{\Delta _R(X)} \overline{M}\\
    SE(X) = N*cov(\Delta _R(X),M) \\
    \Delta _RX_{sp} = \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} - P_{sp}
  \end{array}
  (\#eq:RNE)
\end{equation}

## Biodiversity net effect of the resilience $NE[R(X)]$

Using equations \@ref(eq:R) and \@ref(eq:NE) definition of the biodiversity net effect of resilience is less straightforward. We first need to define the variation of relative ecosystem resilience $\Delta _R[R(X)]$ used to define both complementarity and selection effects. For that we need mathematic properties on $\Delta _R(X)$. We know that if the biodiversity net effect is null, it might corresponds to the case where each species yield as much in mixture than in monoculture:

\begin{equation}
  \begin{array}{c}
    NE = 0 \\
    \sum_{sp} \Delta _R(X)_{sp}=0  ~|~ \forall sp,  \Delta _R(X)_{sp}=0 \\
    \frac{X_{sp}(mixture)}{X_{sp}(monoculture)} = P_{sp} \\
    X_{sp}(mixture) = P_{sp}*X_{sp}(monoculture)
  \end{array}
  (\#eq:DeltaNull)
\end{equation}

So this property must be respected when defining $\Delta _R[R(X)]$.

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{R(X)_{sp}(mixture)}{R(X)_{sp}(monoculture)} - P_{sp} \\
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{       \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    R(M_{sp})= \frac{{M_{sp}}_T}{{M_{sp}}_C} \\
    CE[R(X)] = N* \overline{\Delta _R[R(X)]} \overline{R(M)}\\
    SE[R(X)] = N*cov(\Delta _R[R(X)],R(M)) \\
    NE[R(X)] = CE[R(X)] + SE[R(X)]  
  \end{array}
  (\#eq:NER)
\end{equation}

# Simulation test of the equality

Let's imagine a mixture ($C_T$) of three species $sp1$, $sp2$ and $sp3$ with an ecosystem variable $X$ evolving after a disturbance event, the corresponding three monoculture (${M_{sp1}}_T$, ${M_{sp2}}_T$, ${M_{sp3}}_T$) which encountered the same disturbance, and the corresponding control mixture ($C_C$) and three control monoculture (${M_{sp1}}_C$, ${M_{sp2}}_C$, ${M_{sp3}}_C$) which did not encoutered any disturbance.

## Generation of fake data

Let's first imagine the three species without disturbance in monoculture and let's assume their value is constant through time. We will use a time length of 100 steps. Let's imagine their value in the mixture to be increased by a factor $1.2$ (assuming this is the same for each species, which is obviously wrong). The mixture in control will be simply the sum of each weighted by their abundance, that we will define as 0.7, 0.2, and 0.1. And finally let's assume the value of the ecosyustem after disturbance to be a function of time $f(x) = {{\frac{x_0}{2}+\frac{x}{2}*(\frac{t}{100}})^2}$.

```{r data, echo=T}
data <- data.frame(
  t = 1:100,
  M.sp1.control = rep(100, 100),
  M.sp2.control = rep(150, 100),
  M.sp3.control = rep(50, 100)
) %>% 
  mutate(C.sp1.control = M.sp1.control*1.2,
         C.sp2.control = M.sp2.control*1.2,
         C.sp3.control = M.sp3.control*1.2) %>% 
  mutate(C.control = 0.7*C.sp1.control + 0.2*C.sp2.control + 0.1*C.sp3.control) %>% 
  mutate(M.sp1.treatment = first(M.sp1.control)/2 + (M.sp1.control/2)*(t/100)^2,
         M.sp2.treatment = first(M.sp2.control)/2 + (M.sp2.control/2)*(t/100)^2,
         M.sp3.treatment = first(M.sp3.control)/2 + (M.sp3.control/2)*(t/100)^2,
         C.sp1.treatment = first(C.sp1.control)/2 + (C.sp1.control/2)*(t/100)^2,
         C.sp2.treatment = first(C.sp2.control)/2 + (C.sp2.control/2)*(t/100)^2,
         C.sp3.treatment = first(C.sp3.control)/2 + (C.sp3.control/2)*(t/100)^2,
         C.treatment = 0.7*C.sp1.treatment + 0.2*C.sp2.treatment + 0.1*C.sp3.treatment)
```

```{r gData, fig.cap="Ecosystems through time.", fig.height=4}
data %>% 
  reshape2::melt(id.vars = "t",
                 variable.name = "ecosystem",
                 value.name = "X") %>% 
  ggplot(aes(t, X, color = ecosystem)) +
  geom_line() +
  ylim(0,NA)
```

## Defining functions in R

We define all the previous functions for simulations.

```{r functions, echo=T}
R <- function(Xt, Xc) Xt / Xc
NE <- function(CE, SE) CE + SE
CE <- function(DeltaR, M) length(M) * mean(DeltaR) * mean(M)
SE <-  function(DeltaR, M) length(M) * cov(DeltaR, M)
DeltaR <- function(Xmixture, Xmonoculture, Psp) Xmixture/Xmonoculture - Psp
```

## Simulations

### R[NE(X)]

```{r RNE, echo=T}
RNE <- data %>% 
  group_by(t) %>% 
  mutate(DeltaR.sp1.control = DeltaR(C.sp1.control, M.sp1.control, 0.7),
         DeltaR.sp2.control = DeltaR(C.sp2.control, M.sp2.control, 0.2),
         DeltaR.sp3.control = DeltaR(C.sp3.control, M.sp3.control, 0.1),
         DeltaR.sp1.treatment = DeltaR(C.sp1.treatment, M.sp1.treatment, 0.7),
         DeltaR.sp2.treatment = DeltaR(C.sp2.treatment, M.sp2.treatment, 0.2),
         DeltaR.sp3.treatment = DeltaR(C.sp3.treatment, M.sp3.treatment, 0.1)) %>% 
  mutate(CE.control = CE(c(DeltaR.sp1.control, DeltaR.sp2.control, DeltaR.sp3.control),
                         c(M.sp1.control, M.sp2.control, M.sp3.control)),
         CE.treatment = CE(c(DeltaR.sp1.treatment, DeltaR.sp2.treatment, DeltaR.sp3.treatment),
                         c(M.sp1.treatment, M.sp2.treatment, M.sp3.treatment))) %>% 
  mutate(SE.control = SE(c(DeltaR.sp1.control, DeltaR.sp2.control, DeltaR.sp3.control),
                         c(M.sp1.control, M.sp2.control, M.sp3.control)),
         SE.treatment = SE(c(DeltaR.sp1.treatment, DeltaR.sp2.treatment, DeltaR.sp3.treatment),
                           c(M.sp1.treatment, M.sp2.treatment, M.sp3.treatment))) %>% 
  mutate(NE.control = NE(CE.control, SE.control),
         NE.treatment = NE(CE.treatment, SE.treatment)) %>%
  mutate(RNE = R(NE.treatment, NE.control))
```

### NE[R(X)]

```{r NER, echo=T}
NER <- data %>% 
  group_by(t) %>% 
  mutate(C.sp1 = R(C.sp1.treatment, C.sp1.control),
         C.sp2 = R(C.sp2.treatment, C.sp2.control),
         C.sp3 = R(C.sp3.treatment, C.sp3.control),
         M.sp1 = R(M.sp1.treatment, M.sp1.control),
         M.sp2 = R(M.sp2.treatment, M.sp2.control),
         M.sp3 = R(M.sp3.treatment, M.sp3.control)) %>%
  mutate(DeltaR.sp1 = DeltaR(C.sp1, M.sp1, 0.7),
         DeltaR.sp2 = DeltaR(C.sp2, M.sp2, 0.2),
         DeltaR.sp3 = DeltaR(C.sp3, M.sp3, 0.1)) %>% 
  mutate(CE = CE(c(DeltaR.sp1, DeltaR.sp2, DeltaR.sp3),
                         c(M.sp1, M.sp2, M.sp3))) %>% 
  mutate(SE = SE(c(DeltaR.sp1, DeltaR.sp2, DeltaR.sp3),
                         c(M.sp1, M.sp2, M.sp3))) %>% 
  mutate(NER = NE(CE, SE))
```

## Result

Strangely simulations (see figure \@ref(fig:gNE)) were able to show that $R[NE(X)] \neq NE[R(X)]$ but that:

\begin{equation}
  R[NE(X)] = \frac{NE[R(X)]}{2}
  (\#eq:sim)
\end{equation}

This result (equation \@ref(eq:sim)) was not expected and should be shwon mathematically.

```{r gNE, fig.cap="Resilience and Biodiversity net effect combination."}
RNE %>% 
  left_join(NER, by = "t") %>% 
  select(t, RNE, NER) %>% 
  mutate(R = NER/RNE) %>% 
  rename_("NE[R(X)]" = "NER",
          "NE[R(X)]/R[NE(X)]" = "R",
          "R[NE(X)]" = "RNE") %>% 
  reshape2::melt(id.vars = "t") %>% 
  ggplot(aes(t, value, color = variable)) +
  geom_line() +
  ylim(0,NA) + facet_wrap(~variable, ncol = 1)
```

# Mathematic resolution of the equality

## Variation of relative ecosystem resilience $\Delta _R[R(X)]$

In the simplest case where $\Delta _R[R(X)]_{sp} = 0$:

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    \Delta _R[R(X)]_{sp} = 0 \\
    \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } = P_{sp} \\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{1}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} = P_{sp} * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} + P_{sp} - P_{sp} = P_{sp} * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
    {\Delta _R(X)_{sp}}_T + P_{sp} = P_{sp} * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
    {\Delta _R(X)_{sp}}_T + 1 = {\Delta _R(X)_{sp}}_C + P_{sp}, \forall P_{sp} \neq 0\\
    {\Delta _R(X)_{sp}}_T = {\Delta _R(X)_{sp}}_C + P_{sp} - 1 \\
    R[\Delta _R(X)_{sp}] = 1 + \frac{ P_{sp} - 1 }{ {\Delta _R(X)_{sp}}_C }, \forall {\Delta _R(X)_{sp}}_C \neq 0
  \end{array}
  (\#eq:DeltaR0)
\end{equation}

If we note $\Delta _R[R(X)]_{sp} = a, \forall a in \mathcal R$:

\begin{equation}
  \begin{array}{c}
    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
    \Delta _R[R(X)]_{sp} = a \\
    \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } = (P_{sp} + a) \\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{1}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} = (P_{sp} + a) * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} + P_{sp} - P_{sp} = (P_{sp} + a) * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
    {\Delta _R(X)_{sp}}_T + P_{sp} = (P_{sp} + a) * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
    {\Delta _R(X)_{sp}}_T = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + (P_{sp} + a)*P_{sp} - P_{sp} \\
    {\Delta _R(X)_{sp}}_T = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + P_{sp}*(P_{sp} + a - 1)\\
    R[\Delta _R(X)_{sp}] = P_{sp} + a + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0
  \end{array}
  (\#eq:DeltaRa)
\end{equation}

## Monoculture of resilience $M[R(X)]$

This one is straightforward, it only depends on the order of the indices.
\begin{equation}
  M[R(X)] = \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  (\#eq:M)
\end{equation}

## Complementarity effect of resilience $CE[R(X)]$

Using equations \@ref(eq:NE), \@ref(eq:DeltaRa) and \@ref(eq:M), we have:

\begin{equation}
  \begin{array}{c}
    CE[R(X)] = N* \overline{\Delta _R[R(X)]} \overline{R(M)}\\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    M[R(X)] = \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  \end{array}
  (\#eq:CER1)
\end{equation}

where $N$ stays $N$ and $\overline{R(M)} = \overline{M(R))}$ (equation \@ref(eq:M)). So we need only to calculate $\overline{\Delta _R[R(X)]}$:

\begin{equation}
  \begin{array}{c}
    \overline{\Delta _R[R(X)]} = \sum _{sp} \frac{\Delta _R[R(X)]_{sp}}{N} \\
    \overline{\Delta _R[R(X)]} = \sum _{sp} \frac{ R[\Delta _R(X)_{sp}] - P_{sp} - \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C }  }{N} \\
    \overline{\Delta _R[R(X)]} = R(\overline{\Delta _R(X)}) -\frac{1}{N}(1 + \sum _{sp} \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C }) \\
  \end{array}
  (\#eq:CER2)
\end{equation}

## Selection effect of resilience $SE[R(X)]$

Using equations \@ref(eq:NE), \@ref(eq:DeltaRa) and \@ref(eq:M), we have:

\begin{equation}
  \begin{array}{c}
    SE[R(X)] = N*cov(\Delta _R[R(X)],M[R(X)]) \\
    R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
    M[R(X)] = \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
  \end{array}
  (\#eq:SER1)
\end{equation}

We can develop covariance with following formula:

\begin{equation}
  cov(X,Y) = \frac{\sum _{i=1} ^N (X_i - \overline{X})(Y_i - \overline{Y})}{n-1}
  (\#eq:cov)
\end{equation}

## Biodiversity net effect of resilience $NE[R(X)]$

Using equations \@ref(eq:CER2) and \@ref(eq:SER1), we have:

\begin{equation}
  NE[R(X)] = CE[R(X)] + SE[R(X)]
  (\#eq:NER1)
\end{equation}

# Conclusion

I did not managed to solve mathematically the issue. Still simulations seems to show that $NE[R(X)]$ and $R[NE(X)]$ are mathematically linked through a factor $2$ (tested for several parameters not shown here).

# References

### 2018-07-13 S3-LM.Rmd
Draft 0.1.9005 Supplementary: RNE & LM

---
title: "Supplementary material S3: Linear models"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 10, fig.width = 12,
                cache = T, cache.lazy = F)
```

```{r resilience, echo=FALSE}
rich <- c(5,25,125)
cvh <- 1:20
dist <- c(0,25,50,75)
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
doe <- rename(doe, SR = richness)
load(file.path(path, 'disturbanceData.Rdata'))
load(file.path(path, 'disturbanceResilience.Rdata'))
```

```{r Deq, cache=TRUE, echo=FALSE}
Deq <- resilience %>% 
  group_by(time, sim) %>% 
  mutate_all(funs((1-.)^2)) %>% 
  mutate(structure = sqrt(sum(agb, ba, n, n10, n30))) %>% 
  mutate(production = sqrt(sum(gpp, npp, Rday, Rnight)))
```

```{r Ir, cache=TRUE, echo=FALSE}
Ir <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time), cumsum) %>% 
  filter(time == 600)
```

This supplementary materials is here check the validity of the linear model and to give additional ouptuts of the results not shown in the manuscript. It will be separated in three parts, (i) linear models, (ii) variables multicolinearity, (iii) validation of linear models, and (iii) results of linear models.

# Linear models

We want to test the effect of diversity on the resilience indice $I_R$. For that we will two models, one including only multi-trait diversity indices (model \@ref(eq:m1)) and one including single-trait diversity indices (model \@ref(eq:m2)):

\begin{equation}
    I_R = \beta_0 + \beta_{SR=5} + \beta_{SR=25} + \beta_{FDiv}*FDiv + beta_{FEve}*FEve + \epsilon \\
  (\#eq:m1)
\end{equation}

\begin{equation}
  \begin{split}
    I_R = \beta_0 + \beta_{SR=5} + \beta_{SR=25} + \beta_{FDiv}*FDiv + beta_{FEve}*FEve \\ 
    + beta_{FDis.LMA}*FDis.LMA beta_{FDis.wsg}*FDis.wsg \\
    + beta_{FDis.hmax}*FDis.hmax \epsilon
  \end{split}
  (\#eq:m2)
\end{equation}

The resilience indices is built either on multiple ecosystem variables as for forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and for forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)), or for each variable independantly.

This results in 2 models multiplied by 11 $I_R$ indices giving 22 regressions.

# Variables multicolinearity

We tested multicolinearity for all variables from the two models, and none of theim have shown multicolinearity issues (see figure \@ref(fig:multicolinearity)).

```{r multicolinearity, fig.cap="Pairs plot and multicolinearity between diversity indices."}
doe %>% 
  select(SR, FEve, FDiv, FDis.LMA, FDis.wsg, FDis.hmax) %>% 
  GGally::ggpairs()
```

# Models validation

We validated the 66 regressions with diagnostic plots (qqplot and Cook's leverage). Figures \@ref(fig:diag1Str), \@ref(fig:diag1Prod), \@ref(fig:diag2Str) and \@ref(fig:diag2Prod) shows relations between the $I_R$ indices for several ecosystem outputs and diversity indices. Most of theim did not seems to have issues. Still $n$, $n10$ and $n30$ showed strong outliers for every level of disturbance and disturbance intensity of 75% included those outliers in the $structure$ output. Similarly disturbance intensity of 75% included outliers for $production$, $gpp$, $npp$, $Rday$ and $Rnight$. Those outliers are linked to species with only 5 species strongly disturbed that might derived through an unexpected ecosystem state. Nevertheless those outliers hsould be taken into account in further interpretations.

## Multitrait only

### Structure

```{r diag1Str, fig.cap="Effect of multi-trait diversity indices on resilience index $I_R$ for ecosystem variables linked to structure (AGB, BA, N, N10 and N30)."}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         structure, agb, ba, n, n10, n30,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
  geom_smooth(method = "lm") +
  geom_point() +
  facet_wrap(disturbance~variable, scales = "free_y", nrow = 3)
```

### Production

```{r diag1Prod, fig.cap="Effect of multi-trait diversity indices on resilience index $I_R$ for ecosystem variables linked to production (GPP, NPP, Rday, and Rnight)."}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         production, gpp, npp, Rday, Rnight,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
  geom_smooth(method = "lm") +
  geom_point() +
  facet_wrap(disturbance~variable, scales = "free_y", nrow = 3)
```

## Multi- and mono-trait

### Structure

```{r diag2Str, fig.cap="Effect of mono-trait diversity indices on resilience index $I_R$ for structure ecosystem variables."}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         structure,
         FDis.LMA, FDis.wsg, FDis.hmax) %>% 
  reshape2::melt(id.vars = c("disturbance", "structure"), 
                 value.name = "FDis", variable.name = "Trait") %>% 
  ggplot(aes(FDis, structure, group = NA)) +
  geom_smooth(method = "lm") +
  geom_point() +
  facet_wrap(disturbance~Trait, scales = "free_y", nrow = 3)
```

### Production

```{r diag2Prod, fig.cap="Effect of mono-trait diversity indices on resilience index $I_R$ for production ecosystem variables."}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         production,
         FDis.LMA, FDis.wsg, FDis.hmax) %>% 
  reshape2::melt(id.vars = c("disturbance", "production"), 
                 value.name = "FDis", variable.name = "Trait") %>% 
  ggplot(aes(FDis, production, group = NA)) +
  geom_smooth(method = "lm") +
  geom_point() +
  facet_wrap(disturbance~Trait, scales = "free_y", nrow = 3)
```

# Models results

Tables \@ref(tab:fit1) and \@ref(tab:fit2) present the result of the 66 regressions for linear models \@ref(eq:m1) and \@ref(eq:m2).

## Multitrait only

Globally and as expected, $structure$ variables represents well $AGB$, $BA$, $N$, $N10$, with a strong negative effect of functional diversity $FDiv$ increasing with disturbance intensity and a significative effect of functional eveness $FEve$ only for the strongest distruabnce of 75%. $N30$ did not had any significant effect from diversity indices in multitrait. $SR25$ decreased $I_R$ for strong disturbance, only through its significative but low effect on $BA$.

Similarly, $production$ variable represented well $GPP$, $NPP$, $Rday$ and $Rnight$ with a significant effect of both functional diversity $FDiv$ and eveness $FEve$ increasing with all disturbance levels (except $FDiv$ for a disturbance of 75%). $SR25$ decreased $I_R$ for strong disturbance, only through its significative but low effect on $NPP$.

```{r fit1}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         structure, agb, ba, n, n10, n30,
         production, gpp, npp, Rday, Rnight,
         SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
  group_by(variable, disturbance) %>%
  do(fit = lm(Ir ~ SR + FEve + FDiv, data = .)) %>% 
  tidy(fit) %>% 
  ungroup() %>% 
  filter(term != "(Intercept)") %>% 
  mutate(`P-value` = ifelse(p.value < 0.001, "***",
                       ifelse(p.value < 0.01, "**",
                              ifelse(p.value < 0.05, "*",
                                     ifelse(p.value < 0.1, ".", "n.s."))))) %>% 
  mutate(Coefficient = round(estimate,2)) %>% 
  mutate(Recovery = variable) %>% 
  rename(Variable = term, `Disturbance intensity` = disturbance) %>%
  select(Recovery, `Disturbance intensity`, Variable, Coefficient, `P-value`) %>% 
  mutate(value = paste(Coefficient, `P-value`)) %>% 
    mutate(value = cell_spec(value, 
                           bold = ifelse(grepl("*", value, fixed = T), T, F))) %>%
  reshape2::dcast(Recovery + `Disturbance intensity` ~ Variable, value.var = "value") %>% 
  kable(caption = "Result of linear regression testing the effect of species richness and functional diversity indices on the recovery index by recovery type and disturbance intensity.", escape = F) %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

## Multi- and mono-trait

Globally mono-trait functional dispersion indices did not changed a lot the result due to absent or weak effects. Regarding $structure$, $FDis.hmax$ and $FDis.wsg$ had respectivelly significant small positive and negative effects on $BA$ and $N10$ but not on $structure$ itself, except for $FDis.wsg$ at low disturbance intensity (25%). On the opposite for $production$, $FDis.LMA$ had a significant positive effect increasing with disturbance intensity on $production$, $npp$, $Rday$ and $Rnight$.

```{r fit2}
Ir %>% 
  left_join(doe, by = "sim") %>% 
  ungroup() %>% 
  select(disturbance, 
         structure, agb, ba, n, n10, n30,
         production, gpp, npp, Rday, Rnight,
         SR, FEve, FDiv, FDis.LMA, FDis.wsg, FDis.hmax) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv", 
                             "FDis.LMA", "FDis.wsg", "FDis.hmax"), value.name = "Ir") %>% 
  group_by(variable, disturbance) %>%
  do(fit = lm(Ir ~ SR + FEve + FDiv + FDis.LMA + FDis.wsg + FDis.hmax, data = .)) %>% 
  tidy(fit) %>% 
  ungroup() %>% 
  filter(term != "(Intercept)") %>% 
  mutate(`P-value` = ifelse(p.value < 0.001, "***",
                       ifelse(p.value < 0.01, "**",
                              ifelse(p.value < 0.05, "*",
                                     ifelse(p.value < 0.1, ".", "n.s."))))) %>% 
  mutate(Coefficient = round(estimate,2)) %>% 
  mutate(Recovery = variable) %>% 
  rename(Variable = term, `Disturbance intensity` = disturbance) %>%
  select(Recovery, `Disturbance intensity`, Variable, Coefficient, `P-value`) %>%
  mutate(value = paste(Coefficient, `P-value`)) %>% 
  mutate(value = cell_spec(value, 
                           bold = ifelse(grepl("*", value, fixed = T), T, F))) %>%
  reshape2::dcast(Recovery + `Disturbance intensity` ~ Variable, value.var = "value") %>% 
  kable(caption = "Result of linear regression testing the effect of species richness and functional diversity indices on the recovery index by recovery type and disturbance intensity.", escape = F) %>% 
  kable_styling(full_width = F) %>% 
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

# Conclusion

We developped here two linear models to explore the result of our simulations and more particularly the linked between diversity indices and resilience of ecosystem variables. Our main issues is to summarize the global information of the 66 regressions in something more easy to assimilate for the reader.

First, I think that based on the hysteresis issue we can justify and use the aggregated variables $structure$ and $production$ to summarize all ecosystem variables. And I prefer to show and interpret those general ones, even if they are more abstrait, instead of picking one as an example.

Secondly tfor $structure$ and $production$ variables models \@ref(eq:m1) and \@ref(eq:m2) gives similar results, at the exception of the effect of $FDis.LMA$ in $production$ resilience. So I think we can keep only model \@ref(eq:m2) keeping in mind the effect of $FDis.LMA$ on $production$.

Consequently results can be summarized as follow. Functional diversity $FDiv$ have shown a significant strong negative effect increasing with disturbance intensity both on forest $structure$ and $production$. In addition, functional eveness $FEve$ have shown a significant strong negative effect increasing with disturbance intensity for forest $production$ and a significant negative effect on forest $structure$ only for the strongest level of disturbance (75%). Finally a species richness of 25 decreased the resilience indice of forest $structure$ and $production$ and $FDis.LMA$ have also shown a sinificant negative effect on forest $production$ only for model \@ref(eq:m2).

### 2018-07-13 S4-disturbance.Rmd
Draft 0.1.9005 Supplementary: RNE & LM

--- 2018-01-26 S2.Rmd
@@
 output:
-  bookdown::pdf_document2:
-    number_sections: false
-    toc: false
   bookdown::html_document2:
@@
     toc_float: yes
+  bookdown::pdf_document2:
+    number_sections: false
+    toc: false
 csl: /home/sylvain/Documents/Bibliography/csl/mee.csl

### 2018-07-13 draft.Rmd
Draft 0.1.9005 Supplementary: RNE & LM

@@
 ---
-title: Functional diversity improves tropical forest resilience to disturbance
-author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabien Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
+title: Functional diversity improves tropical forest recovery from disturbance shifting from complementarity to selection effect (to rework)
+author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
 date: '`r Sys.Date()`'
@@
     toc_float: yes
+  bookdown::word_document2: default
   github_document: default
@@
     keep_tex: true
-  bookdown::word_document2: default
 linestretch: 1.5
@@
 library(knitr)
+library(kableExtra)
+if(knitr:::is_html_output()) options(knitr.table.format = "html") 
+if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
 library(tidyverse)
+library(broom)
+theme_set(bayesplot::theme_default())
 opts_chunk$set( echo = F, message = F, warning = F, 
@@
 
-Natural disturbances caused by the death of one or more trees, are the dominant forms of forest renewal in tropical forests as the creation of canopy gaps continuously modifies forest horizontal and vertical structure [@Veblen1992].
-Today, climate change and other direct human activities are modifying forest disturbance regimes, with huge and under evaluated impacts.
-When human induced disturbances are high, it is referred to as "degraded forests" characterized by a modification of the forest ecosystem functioning, possibly reducing the furnished ecosystem services, while keeping the ecosystem as a forest, contrary to deforestation [@Simula2009].
-Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics.
-The main source of degradation around the tropics is forest logging that encompass selective timber harvest, in native forests, and wood fuel harvest, often in secondary forests [@Pearson2017].
-Tropical forests disturbances account for 2.07 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005], estimating to cover areas 10 or more times larger than deforestation itself [@Herold2011].
-Finally, disturbance in tropical forest dynamics induced by climate change likely results from significant changes in precipitation regime [@Davidson2012] with intense drought events resulting in huge mortality rates and biomass loss [@Phillips2009].
+Natural disturbances caused by tree death play a major role in tropical forest functioning through gap-phase regeneration [@Veblen1992].
+Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015].
+If deforestation has been rightly the focus of worldwide attention, forest degradation has been less studied and quantified through tropics [@Putz2010; @Simula2009].
+However, degraded forests cover areas over ten times larger than deforestation [@Herold2011], a majority of which is being logged for timber production [@Blaser2011].
+Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 20.7 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].
+Understanding the drivers of tropical forest resilience to disturbance is thus critical.
 
-Tropical forest ecosystem response to disturbance depends first on the disturbance intensity itself.
-For instance, @DeAvila2015, focusing on timber and wood fuel harvest in Central Amazonia, found that when disturbance is high, e.g.
-induced by post logging sylvicultural practices had stronger negative effect on the post-disturbance system trajectory than logging itself.
-@Carreno-Rocabado2012 found no direct effect of disturbance intensity logging on community diversity itself but a community composition drift towards 'fast' acquisitive species [@reich_world-wide_2014].
-However, even the short-term impact on biodiversity is unclear, as sometimes short term increase in diversity can occur at low disturbance intensities [@Martin2015].
+Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010].
+Both exogenous (e.g.
+climate, disturbance features) and endogenous (e.g.
+forest structure and composition) factors can drive ecosystem recovery after disturbance.
+Among the former, logging intensity has been reported to strongly control and increase time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and change in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015].
+Among endogenous drivers, forest pre-disturbance composition has been identified as a key driver of forest recovery [@Herault2018], as different species respond differently to perturbations depending on their ecological strategies and functional properties [@herault_functional_2011].
+Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@Cadotte2011; @Ives2007; @Loreau2013].
 
-The main reason of our lack of knowledge is the difficulty to conduct long term empirical study so that the use of forest simulators has become inevitable [@Huth2004].
-Individual-based models of forest dynamics with species-explicit model parameters present the perfect framework to develop such joint biodiversity-ecosystem functioning studies [@Marechaux2017].
-Starting from scratch, these models allow describing forest recovery through time, assessing individual tree recruitment, growth and mortality [@Bugmann2001].
-Managing diversity in such models is still challenging and many models have built open Plant Functional Types to manage the volume of model parameters.
-[@Pacala1996].
+Two types of mechanisms can underlie an effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001].
+Complementarity effect results from resources partitioning and niche differentiation, and facilitation, leading to a more efficient use of available resources.
+Selection effect results from a higher probability of more diverse community to include particularly efficient species regarding ecosystem process or stability.
+Moreover the sampling effect additionally implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].
 
-Recently the forest growth simulator TROLL [@Chave1999], an individual-based and spatially-explicit forest model has been explicitly developed to deal with high diversity levels by using plant ecophysiological traits as surrogates of model parameters [@Marechaux2017].
-Consequently, TROLL allow to fully simulate a highly-diverse system and offers the opportunity to study biodiversity-ecosystem functioning *in silico*.
+The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2009], or observations along natural gradient of biodiversity [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017].
+However, experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Schnitzer2016], rarely include logged forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018].
+Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014; @Sakschewski2016], but the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].
 
-Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity.
-We assumed that when a disturbance event happen, due to a higher productivity, more diverse forests will recover quicker and thus be more resilient.
-This theoretical expectation stand on two processes: complementarity and selection effects [@Loreau2001].
-Complementarity effect is the addition of better resources acquisition, due to resources partitioning and niche differentiation, and facilitation in diverse communities.
-Selection effect results from resources preemption by more competitive species.
-Moreover, we expect the sampling effect to improve resilience of species-rich forest, with redundancy and a lower risk to lose important functional traits.
+In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest, that allow to deal with high diversity levels through a parameterization using plant functional traits [@Marechaux2017].
+Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning *in silico* in tropical forest communities.
 
-Generally, positive relationship between biodiversity and productivity were empirically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of forests.
-One of the few studies was realized by @Chisholm2013 and has shown a positive significative relationship between species richness and wood productivity on a worldwide forest network.
-But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.
-
-In the present study, we thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience.
-We used a simulation approach using TROLL model to assess long term processes for different levels of disturbances.
-We first assessed diversity effect on forest resilience of structure and functioning using numerous indices of taxonomic and functional diversities.
-Then, we measured biodiversity net effect, partitioned into complementarity and selection effects, on resilience of simulated tropical forest communities to disturbance.
+We used a simulation approach to assess the effect of species richness and functional composition on long-term forest recovery from disturbances of different levels.
+We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance.
+In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned it into complementarity and selection effects.
+Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance due to an increased productivity.
 
@@
 
-TROLL is an individual-based and spatially explicit forest growth model.
-TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production.
-Trees are growing in a spatialized light environment explicitly computed within voxels of 1 $m^3$.
-Each tree is consistently defined by its age, diameter at breast height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
-Tree geometry is calculated with allometric equations but leaf area vary dynamically within each crown following carbon allocations.
-Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels.
-Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
-A species label is associated to a number of species specific parameters (Table \@ref(tab:traits)) related to functional trait values.
-Carbon assimilation is computed over half-hourly period of a representative day.
-Then allocation is computed to simulate tree growth from an explicit carbon balance.
-Environment is updated at each time step set to one month.
-Seed-rains and seedlings are not simulated explicitly but as a pool.
-In addition below-ground processes, herbaceous plants, epiphytes and lianas are not simulated inside TROLL.
-The source code is written in C++ and available upon request.
-All modules of TROLL models are further detailed in @Marechaux2017.
+TROLL is an individual-based and spatially explicit forest dynamic model in which the life cycle of individual trees (recruitment, growth, seed production and death) $>1cm$ diameter is simulated.
+Trees grows in a forest light environment explicitly computed within a $1m^3$ resolution voxel grid.
+Seedlings below the $1cm$ size class are not modeled explicitly, but are part of a seed/seedlings pool.
+A maximum of one tree can establish in each $1*1m$ pixel.
+Each tree is defined by its age, diameter at breast height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
+Tree geometry is calculated with allometric relationships.
+Additionally, each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
+To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which processes are simulated using up-to-date established relationships drawn from the literature.
+The model thus represents tree biodiversity in a realistic way, where data requirement and model outputs parallel current trait collection efforts and field inventories, respectively.
+
+Carbon assimilation is computed over half-hourly period of a representative day per month using the Farquhar, von Caemmerer, and Berry model of photosynthesis [@Farquhar1980].
+Photosynthetic capacities, as well as leaf respiration rate, are computed from species-specific leaf concentration of nitrogen (N) and phosphorous (P), and leaf mass per area [LMA; @Atkin2015; @Domingues2010].
+The net amount of assimilated carbon available for growth is then partitioned among the different plant organs using empirical factors, and converted into increments of stem volume and leaf area using species-specific wood-density and LMA respectively.
+Leaf demography is simulated using species-specific leaf lifespan derived from LMA.
 Following an underestimation of leaf lifespan for low LMA species, we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1).
+Natural tree deaths are stochastic events whose background rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010].
+Additionally, tree can die due to carbon starvation or treefalls.
+Below-ground processes, herbaceous plants, epiphytes and lianas are not simulated in this version of TROLL.
+The source code is written in C++ and available at https://github.com/TROLL-code/TROLL.
+A further detailed description of the model is available in @Marechaux2017.
 
@@
 
-The disturbance module inside TROLL was designed in the simplest way by, for a given iteration $disturb_{iter}$, picking individual trees randomly up with a uniform law and without any individuals nor species targeting.
+A new disturbance module was developed for this study.
+At a given iteration $disturb_{iter}$, individual trees are randomly picking following a uniform law and without any individuals nor species targeting.
 Selected individuals are then removed without triggering a treefall to avoid any side effect.
-The operation is repeated until the disturbance result in a defined lost basal area ($disturb_{intensity}$).
+The operation is repeated until the disturbance result in a defined removed basal area ($disturb_{intensity}$).
 
-## Design of experiment
+## Simulation experiment
 
-In order to assess the role of biodiversity in ecosystem response to disturbance, we  simulated post-disturbance forest trajectories, encompassing both variation of disturbance intensity and biodiversity levels.
-The gradient of disturbance intensity was represented by percentage of basal area loss (0%, 25%, 50% and 75%).
-Biodiversity was assessed with two perspectives: taxonomic and functional.
-We used species richness $SR$ to simulate initial assemblages with diverse taxonomic diversity (5, 25, and 125 species).
-Functional diversity was assessed after running the initial simulations, *i.e.* before applying the disturbance.
-We focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$.
-For each level of species richness $SR$, we simulated 20 communities with a large range of convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means.
-This design of experiments resulted in 60 tree communities ($3~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Marechaux2017]).
-Functional diversities at the end of simulations were assessed with @villeger_new_2008 indices (FRIC, FEve, FDiv, and FDis).
+In order to assess the role of biodiversity on ecosystem recovery from disturbance, we  simulated forest trajectories for different levels of disturbance intensity and tree community compositions.
+For each simulation, we first simulated forest regeneration from bare soil for 600 years [pre-disturbance step, @Marechaux2017], then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step).
+Disturbance intensity was quantified by the percentage of basal area loss, here fixed at 0% (control), 25%, 50% and 75%.
+For each level of disturbance, we performed simulations differing in the original number of species ($SR=15,25,125$) and composition.
+For each level of species richness, we created 1000 virtual communities by randomly picking desired number of species among the regional species pool.
+We used 20 communities for simulations with a large range of convex hull volume $CVH$ but with a community weighted mean close to the regional species pool community weighted mean.
+This led to 240 simulations varying in community composition and disturbance intensity ($3~SR*20~CHV*4~ levels~of~disturbance$).
 
-## Analysing Ecosystem Response
+All simulations were made with a parameterization for an Amazonian forest of French Guiana, with 163 parameterized species, as in @Marechaux2017.
 
-### Ecosystem structure and functioning
+## Analysis
 
-We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).
+### Quantifying simulated community diversity
 
-The resilience of forest structure and functioning values was assessed with:
+We focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$ to define simulation communities.
+Whereas species and functional diversities were assessed for each simulation at the end of the pre-disturbance step with species richness ($SR$) and @villeger_dynamique_2008 indices ($FRIC$, $FEve$, and $FDiv$).
+
+### Ecosystem recovery from disturbance
+
+**We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).**
+
+The recovery of the simulated forest ecosystem at a time $t(R(t))$ regarding a given properties $X$ was defined ad follows:
 
 \begin{equation}
-  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X_T(t)}{X_C(t)}
-  (\#eq:Resilience)
+  R\left(t\right)= \frac{X_T(t)}{X_C(t)}
+  (\#eq:Recovery)
 \end{equation}
 
-The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$.
-But in our peculiar case of simulated forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ is dynamic in time so that we simulated an undisturbed control ecosystem $C$.
-And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ (the control simulations being relatively stable due to small stochasticity among TROLL simulations [@Marechaux2017]).
-$R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value.
-Next, we calculated an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$.
-Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for (the two functions described above: (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight).
-We then integrate the euclidean distances over time to get our ecosystem resilience indicators.
-
-We then applied the following to each indicator:
-
-1. Normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A and B).
-2. Respectively gathering forest structure and forest functioning together to get synthetic indicators (Figure \@ref(fig:datatrans) C)
-3. Integration over time (Figure \@ref(fig:datatrans) D)
-4.
-The ranking of simulations regarding disturbance intensity was highly stable over time for the 240 simulations.
-So we used the cumulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience.
+where $X_T(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$).
+Note that demographic stochasticity in the model induces a low variability among replicated simulations with the same set-up so that **XX** [@Marechaux2017].
+When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery.
+We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$.
+**Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for (the two functions described above: (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight).
+We then integrate the euclidean distances over time to get our ecosystem recovery indicators.** (Figure \@ref(fig:datatrans)).
+We used the cumulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem recovery.
+A high value of $Ieq_{600}$ indicating a slow recovery.
+The influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem recovery index $Ieq_{600}$  was finally investigated by linear regression ($Ieq_{600} \sim SR + FEve + FDiv$).
 
 ```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation.
-Ecosystem outputs (**A**) are normalized by the control value over time to calculate resilience (**B**); resilience of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**).
+Ecosystem outputs (**A**) are normalized by the control value over time to calculate recovery (**B**); recovery of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**).
 ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
@@
 
-The effect of biodiversity on ecosystem resilience can be split in complementarity and selection effect using @Loreau2001 partitioning:
+The net effect of biodiversity on ecosystem recovery (NE) is defined as the difference between the ecosystem recovery of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e.
+ecosystem recovery for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001].
+We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:
 
@@
 
-Biodiversity net effect $NE$ is based on the difference between ecosystem variable $X$ observed value $X_O$ within the community mixture of species and its expected value $X_E$ if species performance were equal to their performance in monocultures.
-This effect can be partitioned between complementarity effect $CE$, representing niche partitioning, positive interactions, and resource supply, and selective effect $SE$ due to dominant species pool driving the ecosystem.
-$N$ represents the total number of species, and $M$ the vector of monocultures performance.
-Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:
+where $N$ represents the total number of species, and $M$ the vector of monoculture performance. Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:
 
@@
 
-$X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$.
+where $X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$.
 $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance.
@@
 Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016].
-But similarly, to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
-So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:
+But similarly, to recovery measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
+So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure treatments net effect recovery $R(NE_T)$:
 
@@
 
-Resilience trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
+Recovery trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
 In order to do that, we simulated the 163 species individually representing 652 simulations of monoculture.
@@
 
-Species richness was not directly link to $Ieq_{600}$.
-Increased functional diversity reduced cumulative integral from ecosystem distance to equilibrium after 600 years ($Ieq_{600}$).
-In addition, functional evenness was complementary reducing $Ieq_{600}$.
+Species richness was not significantly influencing $Ieq_{600}$ with forest structure and had a small positive effect on forest production (Table \@ref(tab:tgIeq)).
+Increased functional diversity reduced cumulative integral from ecosystem distance to equilibrium after 600 years ($Ieq_{600}$) and functional evenness was complementary reducing $Ieq_{600}$ (Figure \@ref(fig:gIeq) and Table \@ref(tab:tgIeq)).
 Finally, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$.
-We found similar results for all disturbance levels and forest functioning (see [Appendix 5: Disturbance simulations]).
+We found similar results for all disturbance levels (see [Appendix 5: Disturbance simulations]).
 
-```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].
+```{r gIeq, echo=FALSE, fig.cap='Ecosystem recovery after 600 years with taxonomic and functional diversity for different levels of disturbance.
+Cumulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].
 Solid lines and grey areas show fitted relationships with their mean and variance of the form $Ieq_600 = FDiv + \\epsilon$.', cache=TRUE, fig.width=12}
@@
 
+```{r tgIeq, cache=T}
+Ieq %>% 
+  left_join(., doe, by = 'sim') %>% 
+  mutate(SR = as.integer(richness)) %>% 
+  select(disturbance, Ieq.structure, Ieq.production, SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"),
+                 variable.name = "Ieq.type", value.name = "Ieq.value") %>% 
+  group_by(Ieq.type, disturbance) %>%
+  do(fit = lm(Ieq.value ~ SR + FEve + FDiv, data = .)) %>% 
+  tidy(fit) %>% 
+  ungroup() %>% 
+  filter(term != "(Intercept)") %>% 
+  mutate(`P-value` = ifelse(p.value < 0.001, "***",
+                       ifelse(p.value < 0.01, "**",
+                              ifelse(p.value < 0.05, "*",
+                                     ifelse(p.value < 0.1, ".", "n.s."))))) %>% 
+  mutate(Coefficient = round(estimate,2)) %>% 
+  mutate(Recovery = gsub("Ieq.", "", Ieq.type)) %>% 
+  rename(Variable = term, `Disturbance intensity` = disturbance) %>%
+  select(Recovery, `Disturbance intensity`, Variable, Coefficient, `P-value`) %>%
+  kable(caption = "Result of linear regression testing the effect of species richness and functional diversity indices on the recovery index by recovery type and disturbance intensity.") %>% 
+  kable_styling(full_width = F) %>%
+  column_spec(1, bold = T) %>%
+  collapse_rows(columns = 1:2)
+```
+
 ## Biodiversity effect
@@
 
-The net biodiversity effect was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation (see Table \@ref(tab:tNE)) to measure the resilience of each ecosystem measure, e.g.
+The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, see Table \@ref(tab:tNE)) for every ecosystem global variable, e.g.
 aboveground biomass (see Figure \@ref(fig:gBE)).
@@
              '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$'))) %>% 
-  kable(caption = 'Table 1: Mean value and standard deviation of biodiversity net effect accross mature simulations before the disturbance event for different ecosystem variable.', digits = 3, format = 'pandoc')
+  kable(caption = 'Mean value and standard deviation of biodiversity net effect accross mature simulations before the disturbance event for different ecosystem variable.', digits = 3, format = 'pandoc')
 ```
 
-```{r gBE, echo=FALSE, fig.cap='Resilience of complementarity and selection effects.
-Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their resilience over time.
+```{r gBE, echo=FALSE, fig.cap='recovery of complementarity and selection effects.
+Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their recovery over time.
 Solid lines represents the mean value whereas areas represents the interval of values among simulations.', cache=TRUE, fig.width=12}
@@
   filter(richness == 125) %>% 
-  filter(BE %in% c('SE', 'CE')) %>% 
+  # filter(BE %in% c('SE', 'CE')) %>% 
   group_by(time, BE, disturbance, richness) %>%
@@
 
-In the limit of the model, we were able to show that functional diversity improved tropical forest resilience after disturbance.
+In the limit of the model, we were able to show that functional diversity improved tropical forest recovery from disturbance.
 More particularly, functional diversity and evenness are key components of diversity in forest recovery trajectory after disturbance.
 Moreover, we found that complementarity between species was insuring forest recovery in the early times before more productive species dominate the forest and insure late recovery.
-We thus advocate the view that both diversity theory (with niche complementarity and facilitation) and mas ratio theory (represented by selection) are implied in ecosystem resilience but with relative importance depending on the spatial grain and time since the disturbance event.
+We thus advocate the view that both diversity theory (with niche complementarity and facilitation) and mass ratio theory (represented by selection) are implied in ecosystem recovery but with relative importance depending on the spatial grain and time since the disturbance event.
 
-## Functional diversity improve tropical forest resilience
+## Functional diversity improve tropical forest recovery
 
-Our results validated the hypothesis of a significative positive relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectively).
-More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008] seems a major aspect of resilience if it's strengthened by a high functional evenness [regularity of the distribution of abundance in this volume, @villeger_new_2008].
+Our results validated the hypothesis of a significant positive relationship between forest recovery and functional diversity and evenness.
+More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008] seems a major aspect of recovery if it's strengthened by a high functional evenness [regularity of the distribution of abundance in this volume, @villeger_new_2008].
 Indeed, high functional diversity combined to high evenness avoid hegemonic effect of a few dominant species.
@@
 
-## Complementarity and selection insure forest resilience
+## Complementarity and selection insure forest recovery
 
-We found that complementary effect was insuring forest resilience in the beginning of forest successions.
+We found that complementary effect was insuring forest recovery in the beginning of forest successions.
 We interpreted this result as the consequence of facilitation processes or lower competition due to different niche.
 As the only resource currently simulated by TROLL is the light, the main facilitation would be light shading of post-pioneer species by pioneer species in disturbance gaps, confirming the study of @Morin2011 who also highlighted the importance of facilitation through light shading in forest resilience.
-But the complementarity effect quickly reduced through time, to let the selection effect insure forest resilience due to an re-enforced dominance of more productive species.
+But the complementarity effect quickly reduced through time, to let the selection effect insure forest recovery due to an re-enforced dominance of more productive species.
 The diminishing of complementarity effect is attributed to the selection of the most competitive species though time.
@@
 In this work, we look at processes to a 16-ha scale, being limited by the computational effort and, our simulation did not include topography resulting in the absence of micro environmental variation affecting both water and nutrient availability.
-Because many studies suggest that, in mature ecosystems, complementarity will be stronger at smaller scale [@Chisholm2013; @Sullivan2017], this could explain the low value of complementarity after the initial stages of forest recovery.
+Because several studies suggest that, in mature ecosystems, complementarity will be stronger at smaller scale [@Chisholm2013; @Sullivan2017], this could explain the low value of complementarity after the initial stages of forest recovery.
 Still, the absence of topography and environmental heterogeneity in our simulations could explain why we observed an effect of complementarity to a 16-ha scale, strengthening @Sullivan2017 interpretation.
@@
 
-We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance.
-We found that diversity improved forest resilience, similarly to productivity [@Liang2016; @Poorter2015].
+We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest recovery from disturbance, and thus in tropical forest resilience.
+We found that diversity improved forest recovery, similarly to productivity [@Liang2016; @Poorter2015].
 Additionally, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure the final forest recovery.

### 2018-07-16 S2-Resilience.Rmd
Draft 0.1.9006 Supplementary: Resilience indice, Linear models, Biodiversity results

---
title: "Supplementary material S2: Resilience indice"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 8,
                cache = T, cache.lazy = F)
```

```{r resilience, echo=FALSE}
rich <- c(5,25,125)
cvh <- 1:20
dist <- c(0,25,50,75)
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
doe <- rename(doe, SR = richness)
load(file.path(path, 'disturbanceData.Rdata'))
load(file.path(path, 'disturbanceResilience.Rdata'))
load(file.path('~/Documents/ECOFOG/Results/disturbance/maturity', 'matureData.Rdata'))
mature <- datal$time %>% mutate(time = time - 600) ; rm(datal)
```

```{r Deq, cache=TRUE, echo=FALSE}
Deq <- resilience %>% 
  group_by(time, sim) %>% 
  mutate_all(funs((1-.)^2)) %>% 
  mutate(structure = sqrt(sum(agb, ba, n, n10, n30))) %>% 
  mutate(production = sqrt(sum(gpp, npp, Rday, Rnight)))
```

```{r Ir, cache=TRUE, echo=FALSE}
Ir <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time), cumsum) %>% 
  filter(time == 600)
```

```{r IrGraph, eval=F, include=F}
mature %>% 
  select(time, sim, agb) %>% 
  separate(sim, c("SR", "CVH"), remove = F) %>% 
  mutate(disturbance = 50) %>% 
  rbind(ecosystem %>% 
          select(time, sim, agb) %>% 
          separate(sim, c("SR", "CVH", "disturbance"), remove = F)) %>% 
  filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
  reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
  rename(control = `0`, treatment = `50`) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = control, color = "control")) +
  geom_line(aes(y = treatment, color = "disturbed")) +
  geom_ribbon(aes(ymax = control, ymin = treatment), 
              fill = "lightgrey", col = NA, alpha = 0.5) +
  xlab("Time (years)") +
  ylab("Aboveground biomass (tonC/ha)") +
  scale_color_discrete("Simulation")
```

This document introduce and explain the recovery function $R(X)$ and the resilience indice $I_R$ definitions (figure \@ref(fig:IrGraphFinal)).

```{r IrGraphFinal, fig.cap="Definition of recovery function and resilience indice shown on aboveground biomass dynamics of a control and disturbed ecosystem before and after the disturbance event."}
knitr::include_graphics("images/final.png")
```

# Recovery

The recovery of the simulated forest ecosystem at a time $t$, $R(t)$ regarding a given properties $X$ was defined ad follows:

\begin{equation}
  R[X(t)] = \frac{X_D(t)}{X_C(t)}
  (\#eq:Recovery)
\end{equation}

where $X_D(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$).

# Resilience

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$. Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance. Consequently, we can use the cummulated recovery as an indice of ecosystem resilience, further defined as $I_R$:

\begin{equation}
  I_R(X) = \int _{t=0} ^{t=600} \frac{X_D(t)}{X_C(t)}.dt
  (\#eq:Resilience)
\end{equation}

Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight). We then integrated the euclidean distances over time to get our ecosystem resilience indice. (Figure \@ref(fig:datatrans)). We used the cumulative integral after 600 years defined as $I_R$ as a measurement of ecosystem resilience. A high value of $I_R$ indicating a slow recovery.

```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation. Ecosystem outputs (**A**) are normalized by the control value over time to calculate recovery (**B**); recovery of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**). ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
cols <- RColorBrewer::brewer.pal(3, "YlOrRd")
g_agb <- ecosystem %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Aboveground biomass (tonC/ha)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                       values=c("#999999", cols))
g_resilience <- resilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb, color = disturbance)) +
  geom_hline(yintercept = 1, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Resilience of aboveground biomass') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_deq <- Deq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, 
             color = disturbance, fill = disturbance)) +
  geom_hline(yintercept = 0, col = 'red') +
  geom_point() +
  geom_area(alpha = 0.2) +
  xlab('Time (years)') +
  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols) +
  scale_fill_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_Ieq <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time), cumsum) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
```

```{r}

```

### 2018-07-16 S3-RNE.Rmd
Draft 0.1.9006 Supplementary: Resilience indice, Linear models, Biodiversity results

--- 2018-07-13 S2-RNE.Rmd
@@
 ---
-title: "Supplementary material S2: Biodiversity net effect and resilience"
+title: "Supplementary material S3: Biodiversity net effect and resilience"
 date: '`r Sys.Date()`'

### 2018-07-16 S4-LM.Rmd
Draft 0.1.9006 Supplementary: Resilience indice, Linear models, Biodiversity results

--- 2018-07-13 S3-LM.Rmd
@@
 ---
-title: "Supplementary material S3: Linear models"
+title: "Supplementary material S4: Linear models"
 date: '`r Sys.Date()`'
@@
 
-Secondly tfor $structure$ and $production$ variables models \@ref(eq:m1) and \@ref(eq:m2) gives similar results, at the exception of the effect of $FDis.LMA$ in $production$ resilience.
+Secondly, for $structure$ and $production$ variables models \@ref(eq:m1) and \@ref(eq:m2) gives similar results, at the exception of the effect of $FDis.LMA$ in $production$ resilience.
 So I think we can keep only model \@ref(eq:m2) keeping in mind the effect of $FDis.LMA$ on $production$.

### 2018-07-16 S5-Biodiversity.Rmd
Draft 0.1.9006 Supplementary: Resilience indice, Linear models, Biodiversity results

---
title: "Supplementary material S5: Biodiversity effect results"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 12,
                cache = T, cache.lazy = F)
```

```{r BE, echo=FALSE}
rich <- c(5,25,125)
cvh <- 1:20
dist <- c(0,25,50,75)
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
doe <- rename(doe, SR = richness)
load(file.path(path, 'disturbanceData.Rdata'))
load(file.path(path, 'disturbanceBE.Rdata'))
load(file.path(path, 'disturbanceBErecovery.Rdata'))
load(file.path(path, '../../maturity', 'matureBE.Rdata'))
load(file.path(path, '../../maturity', 'matureData.Rdata'))
mature <- datal$time ; rm(datal)
```

This document explore graphically the result of the biodiversity net effect partitionning. More especially we want to look if the representation from figure \@ref(fig:gBE) is correct, with the dynamic biodiversity net effect normalized by the control across ecosystems. In order to do that we will start by looking into the dynamic of a single ecosystem \@ref(fig:gBEsingle), we will then move to all ecosystem not normalized by the control \@ref(fig:gBEunorm), before validating our current representation \@ref(fig:gBE).

# Single ecosystem biodiversity dynamic

The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect. We are thus able for one community to look at aboveground biomass dynamic after and before disturbance in control and disturbed ecosystems and correspondig bioviersity net effect, selection effect and complementarity effect (figure \@ref(fig:gBEsingle)). Similarly to global aboveground biomass from the ecosystem the biodiversity net effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increase following the disturbance. Interestingly, we can see that this pattern of decrease of selection effect after disturbance and recovery after is similar to the effect of succession in forest early stages (first century, sub-figure on left-bottom on figure \@ref(fig:gBEsingle)).

```{r gBEsingle, fig.cap='Dynamic of complementarity and selection effects before and after the disturbance event for one ecosystem including 125 species. Solid lines represents the mean value whereas areas represents the interval of values among simulations.'}
g <- BE %>% 
  select(sim, BE, time, agb) %>% 
  separate(sim, c("SR", "CHV", "disturbance"), remove = F) %>% 
  rbind(ecosystem %>% 
          mutate(BE = "ecosystem") %>%
          select(sim, BE, time, agb) %>% 
          separate(sim, c("SR", "CHV", "disturbance"), remove = F)) %>% 
  rbind(BE_mature %>% 
          mutate(time = time - 600) %>% 
          select(sim, BE, time, agb) %>% 
          separate(sim, c("SR", "CHV"), remove = F) %>% 
          mutate(disturbance = "0-75") %>% 
          separate_rows(disturbance, sep = "-", convert = T)) %>% 
  rbind(mature %>% 
          mutate(time = time-600) %>% 
          mutate(BE = "ecosystem") %>% 
          select(sim, BE, time, agb) %>% 
          separate(sim, c("SR", "CHV"), remove = F) %>% 
          mutate(disturbance = "0-75") %>% 
          separate_rows(disturbance, sep = "-", convert = T)) %>% 
  filter(SR == 125, disturbance %in% c(0,75), CHV == 4) %>% 
  reshape2::dcast(SR + CHV + BE + time ~ disturbance, value.var = "agb") %>% 
  rename(control = `0`, disturbed = `75`) %>% 
  ggplot(aes(x = time, col = BE, group = BE)) +
  geom_line(aes(y = control, linetype = "control")) +
  geom_line(aes(y = disturbed, linetype = "disturbed")) + 
  xlab("Time (years)") + 
  ylab("Aboveground biomass (tonC/ha)") +
  scale_color_discrete("Biodiversity\neffect")
g0 <- g + 
  geom_line(aes(y = control, linetype = "control")) +
  geom_line(aes(y = disturbed, linetype = "disturbed")) + 
  geom_ribbon(aes(ymin = disturbed, ymax = control), 
              col = NA, fill = "lightgrey", alpha = 0.5)
gdist <- g + 
  geom_line(aes(y = control, linetype = "control"), lwd = 1.4) +
  geom_line(aes(y = disturbed, linetype = "disturbed"), lwd = 1.4) + 
  xlim(-10,50) + ylim(NA, 60000) +
  scale_color_discrete(guide = F) +
  scale_linetype_discrete(guide = F)
gstart <- g + 
  geom_line(aes(y = control, linetype = "control"), lwd = 1.4) +
  geom_line(aes(y = disturbed, linetype = "disturbed"), lwd = 1.4) + 
  xlim(-600,-500) + ylim(NA, 50000) +
  scale_color_discrete(guide = F) +
  scale_linetype_discrete(guide = F)
gridExtra::grid.arrange(g0, gstart, gdist, 
                        layout_matrix = rbind(c(1, 1), c(2, 3)))
```

# Non-normalized ecosystems biodiversity dynamic

If we aggregate the result of all community with 125 species, we still observe the same pattern (figure \@ref(fig:gBEunorm)). But due to an increase of biodiversity net effect through recovery time, complementarity importance in early recovery is less visible over a long period. Consequently normalizing by control biodiversity net effect can help avoiding this issue.

```{r gBEunorm, fig.cap='Dynamic of complementarity and selection effects for all ecosystems including 125 species. Solid lines represents the mean value whereas areas represents the interval of values among simulations.'}
BE %>% 
  rbind(mutate(ecosystem, BE = "ecosystem")) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125, disturbance == 75) %>% 
  group_by(time, BE, disturbance, SR) %>%
  summarise(qt1 =quantile(agb, probs=0.05),
            qt2 =quantile(agb, probs=0.75),
            mean=mean(agb),
            n=n()) %>% 
  ggplot(aes(x = time)) +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  xlab('Time (years)') +
  ylab('Aboveground biomass net effect')
```

# Normalized ecosystems biodiversity dynamic

The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, figure \@ref(fig:gBE)). Thanks to previous graphical exploration, we can affirm that normalizing by the control is not distorting the result. And this last figure brings the clearer message regarding complementarity and selection effect role in the resilience of the biodiveristy net effect.

```{r gBE, echo=FALSE, fig.cap='Recovery of complementarity and selection effects. Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their recovery over time. Solid lines represents the mean value whereas areas represents the interval of values among simulations.', cache=TRUE}
BEresilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  # filter(BE %in% c('SE', 'CE')) %>%
  group_by(time, BE, disturbance, SR) %>%
  summarise(qt1 =quantile(agb, probs=0.05),
            qt2 =quantile(agb, probs=0.75),
            mean=mean(agb),
            n=n()) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_grid(disturbance ~ SR, labeller = 'label_both') +
  xlab('Time (years)') +
  ylab('Recovery of aboveground biomass net effect')
```

### 2018-07-16 draft.Rmd
Draft 0.1.9006 Supplementary: Resilience indice, Linear models, Biodiversity results

@@
 opts_chunk$set( echo = F, message = F, warning = F, 
-                fig.height = 8, fig.width = 8,
+                fig.height = 8, fig.width = 12,
                 cache = T, cache.lazy = F)
@@
 Cumulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].
-Solid lines and grey areas show fitted relationships with their mean and variance of the form $Ieq_600 = FDiv + \\epsilon$.', cache=TRUE, fig.width=12}
+Solid lines and grey areas show fitted relationships with their mean and variance of the form $Ieq_600 = FDiv + \\epsilon$.', cache=TRUE}
 Ieq %>% 
@@
 Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their recovery over time.
-Solid lines represents the mean value whereas areas represents the interval of values among simulations.', cache=TRUE, fig.width=12}
+Solid lines represents the mean value whereas areas represents the interval of values among simulations.', cache=TRUE}
 BEresilience %>% 
@@
   geom_line(aes(y = mean, color = BE)) +
-  facet_grid(richness ~ disturbance, labeller = 'label_both') +
+  facet_grid(disturbance ~ richness, labeller = 'label_both') +
   xlab('Time (years)') +
@@
 * Supplementary materials S1: Leaf lifespan model
-    + [HTML](S1.html)
-    + [PDF](S1.pdf)
-* Supplementary materials S2: Disturbance simulations
-    + [HTML](S2.html)
-    + [PDF](S2.pdf)
+    + [HTML](S1-LL.html)
+    + [PDF](S1-LL.pdf)
+* Supplementary materials S2: Resilience indice
+    + [HTML](S2-Resilience.html)
+    + [PDF](S2-Resilience.pdf)
+* Supplementary materials S3: Biodiversity net effect and resilience 
+    + [HTML](S3-RNE.html)
+    + [PDF](S3-RNE.pdf)
+* Supplementary materials S4: Linear models
+    + [HTML](S4-LM.html)
+    + [PDF](S4-LM.pdf)
+* Supplementary materials S5: Biodiversity effect results
+    + [HTML](S5-Biodiversity.html)
+    + [PDF](S5-Biodiversity.pdf)
 

### 2018-11-06 S2-Resilience.Rmd
# Draft 0.1.9007 * Fourth version after @BrunoHerault meeting

@@
     toc: false
+  bookdown::word_document2: default
 csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
@@
   ylab("Aboveground biomass (tonC/ha)") +
-  scale_color_discrete("Simulation")
+  scale_color_discrete("Simulation") +
+  scale_color_discrete("Disturbance\nintensity\n(% of BA)", 
+                       labels = c("0 (control)", "50 (disturbed)")) + 
+  ggsave(filename = "./images/original.png", width = 20, height = 10, units = "cm")
 ```
 
-This document introduce and explain the recovery function $R(X)$ and the resilience indice $I_R$ definitions (figure \@ref(fig:IrGraphFinal)).
+This document introduces and explains the recovery function $R(X)$ and the resilience indice $I_R$ definitions (figure \@ref(fig:IrGraphFinal)).
+Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010].
 
@@
 
-The recovery of the simulated forest ecosystem at a time $t$, $R(t)$ regarding a given properties $X$ was defined ad follows:
+The recovery of the simulated forest ecosystem, i.e. the ecosystem capacity to recover from a disturbance event, at a time $t$, $R(t)$ regarding a given property $X$ was defined ad follows:
 
@@
 We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$.
-Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance.
-Consequently, we can use the cummulated recovery as an indice of ecosystem resilience, further defined as $I_R$:
+Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance, i.e.
+the ecosystem capacity to absorb a disturbance event,.
+Consequently **in this particular case**, we can use the cummulated recovery as an indice of ecosystem resilience, further defined as $I_R$:
 
@@
 
-Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight).
-We then integrated the euclidean distances over time to get our ecosystem resilience indice.
-(Figure \@ref(fig:datatrans)).
-We used the cumulative integral after 600 years defined as $I_R$ as a measurement of ecosystem resilience.
 A high value of $I_R$ indicating a slow recovery.
+Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure (AGB, BA, N, N10, and N30, e.g.
+figure \@ref(fig:datatrans)) and (ii) forest functioning (GPP, NPP, Rday, and Rnight).
 
@@
 
-```{r}
-
-```
+# References

### 2018-11-06 S3-RNE.Rmd
# Draft 0.1.9007 * Fourth version after @BrunoHerault meeting

@@
 
-We want to merge two functions to analyse the output of ecosystem after disturbance: (i) the biodiversity net effect (NE) and its partitionning in complementarity (CE) and selection (SE) effects, and (ii) the resilience function of the ecosystem.
-More particularly we computed $R[NE(X)]$, and we want to know if $R[NE(X)]=NE[R(X)]$.
+We want to merge two functions to analyse the output of ecosystem after disturbance: (i) the biodiversity net effect (NE) and its partitionning in complementarity (CE) and selection (SE) effects, and (ii) the resilience function of the ecosystem properties (R).
+More particularly we computed $R[NE(X)]$, and we want to know how $R[NE(X)]$ relate to $NE[R(X)]$.
 In order to do that we will proceed in three times.
@@
 
-We first need to define the resilience function, the biodiversity net effect functions, and their composed functions for any variable $X$ from the ecosystem.
+We first need to define the resilience function, the biodiversity net effect function, and their composed functions for any variable $X$ of the ecosystem.
 
@@
 
-Resilience has been defined by us as:
+We defined resilience as follows:
 
 \begin{equation}
-  R(t) = \frac{X_T(t)}{X_C(t)}
+  R(t) = \frac{X_D(t)}{X_C(t)}
   (\#eq:R)
@@
 
-where $X_T(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation and in the undisturbed control simulation.
+where $X_D(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation and in the undisturbed control simulation.
 
@@
 
-The net effect of biodiversity on ecosystem recovery (NE) is defined as the difference between the ecosystem recovery of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e.
-ecosystem recovery for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001].
-We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:
+The net effect of biodiversity on an ecosystem variable X (NE) is defined as the difference between the ecosystem variable of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e.
+ecosystem variable for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001].
+We then split this net biodiversity effect in complementarity and selection effects using @Loreau2001 partitioning:
 
@@
   \begin{array}{c}
-    R[NE(X)]=\frac{NE_T(X)}{NE_C(X)} \\
-    R[NE(X)]=\frac{CE_T(X)}{NE_C(X)} + \frac{SE_T(X)}{NE_C(X)} \\
+    R[NE(X)]=\frac{NE_D(X)}{NE_C(X)} \\
+    R[NE(X)]=\frac{CE_D(X)}{NE_C(X)} + \frac{SE_D(X)}{NE_C(X)} \\
     CE(X) = N* \overline{\Delta _R(X)} \overline{M}\\
@@
     \Delta _R[R(X)]_{sp} = \frac{R(X)_{sp}(mixture)}{R(X)_{sp}(monoculture)} - P_{sp} \\
-    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{       \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
-    R(M_{sp})= \frac{{M_{sp}}_T}{{M_{sp}}_C} \\
+    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{       \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
+    R(M_{sp})= \frac{{M_{sp}}_D}{{M_{sp}}_C} \\
     CE[R(X)] = N* \overline{\Delta _R[R(X)]} \overline{R(M)}\\
@@
 
-Let's imagine a mixture ($C_T$) of three species $sp1$, $sp2$ and $sp3$ with an ecosystem variable $X$ evolving after a disturbance event, the corresponding three monoculture (${M_{sp1}}_T$, ${M_{sp2}}_T$, ${M_{sp3}}_T$) which encountered the same disturbance, and the corresponding control mixture ($C_C$) and three control monoculture (${M_{sp1}}_C$, ${M_{sp2}}_C$, ${M_{sp3}}_C$) which did not encoutered any disturbance.
+Let's imagine a mixture ($C_D$) of three species $sp1$, $sp2$ and $sp3$ with an ecosystem variable $X$ evolving after a disturbance event, the corresponding three monoculture (${M_{sp1}}_D$, ${M_{sp2}}_D$, ${M_{sp3}}_D$) which encountered the same disturbance, and the corresponding control mixture ($C_C$) and three control monoculture (${M_{sp1}}_C$, ${M_{sp2}}_C$, ${M_{sp3}}_C$) which did not encoutered any disturbance.
 
@@
   \begin{array}{c}
-    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
+    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
     \Delta _R[R(X)]_{sp} = 0 \\
-    \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } = P_{sp} \\
-    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)}\\
-    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{1}{{X_{sp}}_C(monoculture)}\\
-    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} = P_{sp} * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
-    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} + P_{sp} - P_{sp} = P_{sp} * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
-    {\Delta _R(X)_{sp}}_T + P_{sp} = P_{sp} * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
-    {\Delta _R(X)_{sp}}_T + 1 = {\Delta _R(X)_{sp}}_C + P_{sp}, \forall P_{sp} \neq 0\\
-    {\Delta _R(X)_{sp}}_T = {\Delta _R(X)_{sp}}_C + P_{sp} - 1 \\
+    \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } = P_{sp} \\
+    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)}\\
+    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = P_{sp} * \frac{1}{{X_{sp}}_C(monoculture)}\\
+    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)} = P_{sp} * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
+    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)} + P_{sp} - P_{sp} = P_{sp} * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
+    {\Delta _R(X)_{sp}}_D + P_{sp} = P_{sp} * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
+    {\Delta _R(X)_{sp}}_D + 1 = {\Delta _R(X)_{sp}}_C + P_{sp}, \forall P_{sp} \neq 0\\
+    {\Delta _R(X)_{sp}}_D = {\Delta _R(X)_{sp}}_C + P_{sp} - 1 \\
     R[\Delta _R(X)_{sp}] = 1 + \frac{ P_{sp} - 1 }{ {\Delta _R(X)_{sp}}_C }, \forall {\Delta _R(X)_{sp}}_C \neq 0
@@
   \begin{array}{c}
-    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
+    \Delta _R[R(X)]_{sp} = \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } - P_{sp} \\
     \Delta _R[R(X)]_{sp} = a \\
-    \frac{ \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} } = (P_{sp} + a) \\
-    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)}\\
-    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{1}{{X_{sp}}_C(monoculture)}\\
-    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} = (P_{sp} + a) * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
-    \frac{{X_{sp}}_T(mixture)}{{X_{sp}}_T(monoculture)} + P_{sp} - P_{sp} = (P_{sp} + a) * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
-    {\Delta _R(X)_{sp}}_T + P_{sp} = (P_{sp} + a) * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
-    {\Delta _R(X)_{sp}}_T = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + (P_{sp} + a)*P_{sp} - P_{sp} \\
-    {\Delta _R(X)_{sp}}_T = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + P_{sp}*(P_{sp} + a - 1)\\
+    \frac{ \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} }{ \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} } = (P_{sp} + a) \\
+    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)}\\
+    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)}*\frac{1}{{X_{sp}}_C(mixture)} = (P_{sp} + a) * \frac{1}{{X_{sp}}_C(monoculture)}\\
+    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)} = (P_{sp} + a) * \frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)}\\
+    \frac{{X_{sp}}_D(mixture)}{{X_{sp}}_D(monoculture)} + P_{sp} - P_{sp} = (P_{sp} + a) * (\frac{{X_{sp}}_C(mixture)}{{X_{sp}}_C(monoculture)} + P_{sp} - P_{sp})\\
+    {\Delta _R(X)_{sp}}_D + P_{sp} = (P_{sp} + a) * ({\Delta _R(X)_{sp}}_C + P_{sp}) \\
+    {\Delta _R(X)_{sp}}_D = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + (P_{sp} + a)*P_{sp} - P_{sp} \\
+    {\Delta _R(X)_{sp}}_D = (P_{sp} + a)*{\Delta _R(X)_{sp}}_C + P_{sp}*(P_{sp} + a - 1)\\
     R[\Delta _R(X)_{sp}] = P_{sp} + a + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
@@
 \begin{equation}
-  M[R(X)] = \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
+  M[R(X)] = \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
   (\#eq:M)
@@
     R[\Delta _R(X)_{sp}] = \Delta _R[R(X)]_{sp} + P_{sp} + \frac{ P_{sp}*(P_{sp} + a - 1) }{ {\Delta _R(X)_{sp}}_C } \forall {\Delta _R(X)_{sp}}_C \neq 0 \\
-    M[R(X)] = \frac{{X_{sp}}_T(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
+    M[R(X)] = \frac{{X_{sp}}_D(monoculture)}{{X_{sp}}_C(monoculture)} = R[M(X)]
   \end{array}

### 2018-11-06 S4-LM.Rmd
# Draft 0.1.9007 * Fourth version after @BrunoHerault meeting

@@
   arrange(time) %>% 
-  mutate_at(vars(-time), cumsum) %>% 
+  mutate_at(vars(-time, -sim), cumsum) %>% 
   filter(time == 600)
@@
 
-This supplementary materials is here check the validity of the linear model and to give additional ouptuts of the results not shown in the manuscript.
-It will be separated in three parts, (i) linear models, (ii) variables multicolinearity, (iii) validation of linear models, and (iii) results of linear models.
+This supplementary material is here to check the validity of the linear model and to give additional ouptuts of the results not shown in the manuscript.
+It will be separated in four parts, (i) linear models, (ii) variables multicolinearity, (iii) validation of linear models, and (iv) results of linear models.
 
@@
 We want to test the effect of diversity on the resilience indice $I_R$.
-For that we will two models, one including only multi-trait diversity indices (model \@ref(eq:m1)) and one including single-trait diversity indices (model \@ref(eq:m2)):
+For that we will test two models, one including only multi-trait diversity indices (model \@ref(eq:m1)) and one including single-trait diversity indices (model \@ref(eq:m2)):
 
@@
 
-This results in 2 models multiplied by 11 $I_R$ indices giving 22 regressions.
+This results in two models multiplied by 11 $I_R$ indices giving 22 regressions.
 
@@
 
-We tested multicolinearity for all variables from the two models, and none of theim have shown multicolinearity issues (see figure \@ref(fig:multicolinearity)).
+We tested multicolinearity for all independent variables of the two models, and none of theim have shown multicolinearity issues (all Pearson's and/or Spearman's correlation coefficients below 0.6, see figure \@ref(fig:multicolinearity)).
 
-```{r multicolinearity, fig.cap="Pairs plot and multicolinearity between diversity indices."}
+```{r multicolinearity, fig.cap="Pairs plot and multicolinearity between diversity indices.
+Subplots lower the diagonal shows scatterplots between continuous variables or boxplot when a discrete variable is implied.
+Subplots on the diagonal show density plot for continuous variables and barplot for discrete variables.
+Subplots upper the diagonal indicates the Pearson's correlation coefficient between variables."}
 doe %>% 
@@
 We validated the 66 regressions with diagnostic plots (qqplot and Cook's leverage).
-Figures \@ref(fig:diag1Str), \@ref(fig:diag1Prod), \@ref(fig:diag2Str) and \@ref(fig:diag2Prod) shows relations between the $I_R$ indices for several ecosystem outputs and diversity indices.
-Most of theim did not seems to have issues.
-Still $n$, $n10$ and $n30$ showed strong outliers for every level of disturbance and disturbance intensity of 75% included those outliers in the $structure$ output.
+Figures \@ref(fig:diag1Str), \@ref(fig:diag1Prod), \@ref(fig:diag2Str) and \@ref(fig:diag2Prod) shows relations between the $I_R$ indices computed for several ecosystem outputs, and diversity indices.
+Most of theim met the model assumptions requirements.
+Still $n$, $n10$ and $n30$ showed strong outliers for every level of disturbance in the two models.
+Those outliers echoed in the $structure$ variable only at a disturbance intensity of 75%.
 Similarly disturbance intensity of 75% included outliers for $production$, $gpp$, $npp$, $Rday$ and $Rnight$.
-Those outliers are linked to species with only 5 species strongly disturbed that might derived through an unexpected ecosystem state.
-Nevertheless those outliers hsould be taken into account in further interpretations.
+Previously identified outliers corresponded to ecosystem simulations with only 5 species, which seemed to have derived toward an unexpected ecosystem state when strongly disturbed.
+Nevertheless those outliers should be taken into account in further interpretations.
 
-## Multitrait only
+## Multitrait
 
@@
 
-```{r diag1Str, fig.cap="Effect of multi-trait diversity indices on resilience index $I_R$ for ecosystem variables linked to structure (AGB, BA, N, N10 and N30)."}
+```{r diag1Str, fig.cap="Effect of multi-trait diversity indices on resilience index $I_R$ for ecosystem variables linked to structure (AGB, BA, N, N10 and N30).
+To be completed with significativity and regression outputs when the final figure is chosen."}
 Ir %>% 
@@
 
-## Multi- and mono-trait
+## Monotrait
 
 ### Structure
+
+#### FM
+
+```{r diag3Str, fig.cap="Effect of mono-trait diversity indices on resilience index $I_R$ for structure ecosystem variables."}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  select(disturbance, 
+         structure,
+         CWM.LMA, CWM.wsg, CWM.hmax) %>% 
+  reshape2::melt(id.vars = c("disturbance", "structure"), 
+                 value.name = "CWM", variable.name = "Trait") %>% 
+  ggplot(aes(CWM, structure, group = NA)) +
+  geom_smooth(method = "lm") +
+  geom_point() +
+  facet_wrap(disturbance~Trait, scales = "free", nrow = 3)
+```
+
+#### FD
 
@@
 ### Production
+
+#### FM
+
+```{r diag3Prod, fig.cap="Effect of mono-trait diversity indices on resilience index $I_R$ for production ecosystem variables."}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  select(disturbance, 
+         production,
+         CWM.LMA, CWM.wsg, CWM.hmax) %>% 
+  reshape2::melt(id.vars = c("disturbance", "production"), 
+                 value.name = "CWM", variable.name = "Trait") %>% 
+  ggplot(aes(CWM, production, group = NA)) +
+  geom_smooth(method = "lm") +
+  geom_point() +
+  facet_wrap(disturbance~Trait, scales = "free", nrow = 3)
+```
+
+#### FD
 
@@
 
-Globally and as expected, $structure$ variables represents well $AGB$, $BA$, $N$, $N10$, with a strong negative effect of functional diversity $FDiv$ increasing with disturbance intensity and a significative effect of functional eveness $FEve$ only for the strongest distruabnce of 75%.
+Globally and as expected, $structure$ variables represented well $AGB$, $BA$, $N$ and $N10$ behavior, with a strong negative effect of functional diversity $FDiv$ on $I_r$ increasing with disturbance intensity and a significative effect of functional eveness $FEve$ only for the strongest distruabnce of 75%.
 $N30$ did not had any significant effect from diversity indices in multitrait.
@@
 
-Similarly, $production$ variable represented well $GPP$, $NPP$, $Rday$ and $Rnight$ with a significant effect of both functional diversity $FDiv$ and eveness $FEve$ increasing with all disturbance levels (except $FDiv$ for a disturbance of 75%).
+Similarly, $production$ variable represented well $GPP$, $NPP$, $Rday$ and $Rnight$ behavior, with a significant effect of both functional diversity $FDiv$ and eveness $FEve$ on $I_r$ increasing with all disturbance levels (except $FDiv$ for a disturbance of 75%).
 $SR25$ decreased $I_R$ for strong disturbance, only through its significative but low effect on $NPP$.
@@
 On the opposite for $production$, $FDis.LMA$ had a significant positive effect increasing with disturbance intensity on $production$, $npp$, $Rday$ and $Rnight$.
+Increased dispersion in LMA traits thus tend to decrease ecosystem resilience of production variables.
 
@@
 
+## Monotrait only
+
+Adding $FM$  as the weighted sum of species traits = $CWM$.
+
+```{r fit3}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  select(disturbance, 
+         structure, agb, ba, n, n10, n30,
+         production, gpp, npp, Rday, Rnight,
+         SR, FDis.LMA, FDis.wsg, FDis.hmax,
+         CWM.LMA, CWM.wsg, CWM.hmax) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", 
+                             "FDis.LMA", "FDis.wsg", "FDis.hmax",
+                             "CWM.LMA", "CWM.wsg", "CWM.hmax"), value.name = "Ir") %>% 
+  group_by(variable, disturbance) %>%
+  do(fit = lm(Ir ~ SR + CWM.LMA + CWM.wsg + CWM.hmax + FDis.LMA + FDis.wsg + FDis.hmax, data = .)) %>% 
+  tidy(fit) %>% 
+  ungroup() %>% 
+  filter(term != "(Intercept)") %>% 
+  mutate(`P-value` = ifelse(p.value < 0.001, "***",
+                       ifelse(p.value < 0.01, "**",
+                              ifelse(p.value < 0.05, "*",
+                                     ifelse(p.value < 0.1, ".", "n.s."))))) %>% 
+  mutate(Coefficient = round(estimate,2)) %>% 
+  mutate(Recovery = variable) %>% 
+  rename(Variable = term, `Disturbance intensity` = disturbance) %>%
+  select(Recovery, `Disturbance intensity`, Variable, Coefficient, `P-value`) %>%
+  mutate(value = paste(Coefficient, `P-value`)) %>% 
+  mutate(value = cell_spec(value, 
+                           bold = ifelse(grepl("*", value, fixed = T), T, F))) %>%
+  reshape2::dcast(Recovery + `Disturbance intensity` ~ Variable, value.var = "value") %>% 
+  kable(caption = "Result of linear regression testing the effect of species richness and functional diversity indices on the recovery index by recovery type and disturbance intensity.", escape = F) %>%
+  kable_styling(full_width = F) %>% 
+  column_spec(1, bold = T) %>%
+  collapse_rows(columns = 1:2)
+```
+
 # Conclusion
 
-We developped here two linear models to explore the result of our simulations and more particularly the linked between diversity indices and resilience of ecosystem variables.
+We developped here two linear models to explore the result of our simulations and more particularly the links between diversity indices and resilience of ecosystem variables.
 Our main issues is to summarize the global information of the 66 regressions in something more easy to assimilate for the reader.
 
-First, I think that based on the hysteresis issue we can justify and use the aggregated variables $structure$ and $production$ to summarize all ecosystem variables.
+First, I think that based on the hysteresis issue, the fact that ecosystem variables can reach equilibrium value in a transient state whithout reaching equilibirum, we can justify and use the aggregated variables $structure$ and $production$ to summarize all ecosystem variables.
 And I prefer to show and interpret those general ones, even if they are more abstrait, instead of picking one as an example.
 
-Secondly, for $structure$ and $production$ variables models \@ref(eq:m1) and \@ref(eq:m2) gives similar results, at the exception of the effect of $FDis.LMA$ in $production$ resilience.
+Secondly, for $structure$ and $production$ variables models \@ref(eq:m1) and \@ref(eq:m2) gives similar results, at the exception of the effect of $FDis.LMA$ in $production$ resilience, meaning that in most of the case monotrait information is not useful.
 So I think we can keep only model \@ref(eq:m2) keeping in mind the effect of $FDis.LMA$ on $production$.

### 2018-11-06 S5-Biodiversity.Rmd
# Draft 0.1.9007 * Fourth version after @BrunoHerault meeting

@@
 
-This document explore graphically the result of the biodiversity net effect partitionning.
+This document explores graphically the result of the biodiversity net effect partitionning.
 More especially we want to look if the representation from figure \@ref(fig:gBE) is correct, with the dynamic biodiversity net effect normalized by the control across ecosystems.
-In order to do that we will start by looking into the dynamic of a single ecosystem \@ref(fig:gBEsingle), we will then move to all ecosystem not normalized by the control \@ref(fig:gBEunorm), before validating our current representation \@ref(fig:gBE).
+In order to do so, we will start by looking into the dynamics of a single ecosystem \@ref(fig:gBEsingle), we will then move to all ecosystems not normalized by the control \@ref(fig:gBEunorm), before validating our current representation \@ref(fig:gBE).
 
@@
 The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect.
-We are thus able for one community to look at aboveground biomass dynamic after and before disturbance in control and disturbed ecosystems and correspondig bioviersity net effect, selection effect and complementarity effect (figure \@ref(fig:gBEsingle)).
-Similarly to global aboveground biomass from the ecosystem the biodiversity net effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increase following the disturbance.
-Interestingly, we can see that this pattern of decrease of selection effect after disturbance and recovery after is similar to the effect of succession in forest early stages (first century, sub-figure on left-bottom on figure \@ref(fig:gBEsingle)).
+We are thus able for one community to look at aboveground biomass dynamic after and before disturbance in control and disturbed ecosystems and correspondig net biodiversity effect, selection effect and complementarity effect (figure \@ref(fig:gBEsingle)).
+Similarly to the aboveground biomass of the ecosystem, the net biodiversity effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increases following the disturbance.
+Interestingly, we can see that this pattern of decrease of selection effect after disturbance and recovery after is similar to the effect of succession in forest early stages, when the new cohort arrives selection effect suddenly decreases and complementarity effect insure a small net effect (first century, sub-figure on left-bottom on figure \@ref(fig:gBEsingle)).
+Anyway the pattern is stronger after a disturbance, maybe due to an increased diversity allowing a stronger diversity effect.
 
-```{r gBEsingle, fig.cap='Dynamic of complementarity and selection effects before and after the disturbance event for one ecosystem including 125 species.
+```{r gBEsingle, fig.cap='Dynamic of complementarity and selection effects before and after the disturbance event for one ecosystem simulation including 125 species.
 Solid lines represents the mean value whereas areas represents the interval of values among simulations.'}
@@
           separate_rows(disturbance, sep = "-", convert = T)) %>% 
-  filter(SR == 125, disturbance %in% c(0,75), CHV == 4) %>% 
+  filter(SR == 125, disturbance %in% c(0,75), CHV == 10) %>% # 10, 3
   reshape2::dcast(SR + CHV + BE + time ~ disturbance, value.var = "agb") %>% 
@@
   geom_line(aes(y = disturbed, linetype = "disturbed"), lwd = 1.4) + 
-  xlim(-10,50) + ylim(NA, 60000) +
+  xlim(-10,100) + ylim(-30000, 40000) +
   scale_color_discrete(guide = F) +
@@
   geom_line(aes(y = disturbed, linetype = "disturbed"), lwd = 1.4) + 
-  xlim(-600,-500) + ylim(NA, 50000) +
+  xlim(-600,-500) + ylim(-30000, 40000) +
   scale_color_discrete(guide = F) +
   scale_linetype_discrete(guide = F)
-gridExtra::grid.arrange(g0, gstart, gdist, 
-                        layout_matrix = rbind(c(1, 1), c(2, 3)))
+gridExtra::grid.arrange(g0, gstart, gdist, layout_matrix = rbind(c(1, 1), c(2, 3)))
 ```
@@
 
-If we aggregate the result of all community with 125 species, we still observe the same pattern (figure \@ref(fig:gBEunorm)).
+If we aggregate the result of all community with 125 species, we still observe the same pattern for all simulations compard to the single previous one (figure \@ref(fig:gBEunorm)).
 But due to an increase of biodiversity net effect through recovery time, complementarity importance in early recovery is less visible over a long period.
@@
 ```{r gBEunorm, fig.cap='Dynamic of complementarity and selection effects for all ecosystems including 125 species.
-Solid lines represents the mean value whereas areas represents the interval of values among simulations.'}
+Solid lines represents the mean value whereas areas represents the range of values among simulations.'}
 BE %>% 

### 2018-11-06 SX-mortality.Rmd
# Draft 0.1.9007 * Fourth version after @BrunoHerault meeting

---
title: "Nino mortality issue"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = T, message = F, warning = F, 
                fig.height = 8, fig.width = 12,
                cache = T, cache.lazy = F)
```

This document explore if the issue encoutered by Nino of over-mortality in years following the selective logging appeared in my simulations too in TROLL version 2.3.1 with my previous version of the sylviculture module.

```{r data}
path <- '~/Documents/ECOFOG/Results/disturbance/sylviculture/data/'
load(file.path(path, 'ecosystem.Rdata'))
ecosystem <- ecosystem %>% 
  separate(sim, c("SR", "CHV", "control"), remove = F, extra = "drop") %>% 
  filter(SR == 125, is.na(control))
Vkilled <- ecosystem %>% 
  group_by(sim) %>% 
  filter(time %in% 1:2) %>% 
  arrange(time) %>% 
  summarise(Vkilled = 1-last(agb)/first(agb)) %>% 
  select(Vkilled) %>% unlist()
```

```{r plot}
ggplot(ecosystem, aes(time, agb, color = sim)) +
  geom_point() +
  geom_line() +
  ggtitle(paste("Volume killed:", round(mean(Vkilled*100), 2), 
            "%", "(+/-", round(sd(Vkilled*100), 2), "%)")) +
  scale_color_discrete(guide = F) +
  xlim(0,10)
```

### 2018-11-06 draft.Rmd
# Draft 0.1.9007 * Fourth version after @BrunoHerault meeting

@@
 ---
-title: Functional diversity improves tropical forest recovery from disturbance shifting from complementarity to selection effect (to rework)
+title: Biodiversity improves tropical forest resilience, shifting from complementarity to selection effect
 author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
@@
 output:
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
-  bookdown::word_document2: default
-  github_document: default
   bookdown::pdf_document2:
@@
     keep_tex: true
+  bookdown::word_document2: default
+  github_document: default
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
 linestretch: 1.5
@@
 load(file.path(path, 'disturbanceDOE.Rdata'))
+doe <- rename(doe, SR = richness)
 load(file.path(path, 'disturbanceData.Rdata'))
@@
 ```{r Deq, cache=TRUE, echo=FALSE}
-deq <- data.frame(
-  time = resilience$time,
-  sim = resilience$sim,
-  structure = apply(resilience[c('agb', 'ba', 'n', 'n10', 'n30')], 1, 
-                  function(x) sqrt(sum((rep(1, length(x)) - x)^2))),
-  production = apply(resilience[c('gpp', 'npp', 'Rday', 'Rnight')], 1, 
-                  function(x) sqrt(sum((rep(1, length(x)) - x)^2)))
-)
+Deq <- resilience %>% 
+  group_by(time, sim) %>% 
+  mutate_all(funs((1-.)^2)) %>% 
+  mutate(structure = sqrt(sum(agb, ba, n, n10, n30))) %>% 
+  mutate(production = sqrt(sum(gpp, npp, Rday, Rnight)))
 ```
 
-```{r Ieq, cache=TRUE, echo=FALSE}
-sim <- c(sapply(c(sapply(rich, function(r) paste0(r, '.', cvh))), function(x) paste0(x, '.', dist[-1])))
-Ieq <- sapply(sim, function(s){
-  data_s <- subset(deq, sim ==s)
-  return(cbind(data_s[c(1,2)], cumsum(data_s[-c(1,2)])))
-}, simplify = F) ; rm(sim)
-Ieq <- data.frame(do.call('rbind', Ieq))
-names(Ieq)[3:4] <- paste0('Ieq.', names(Ieq)[3:4])
+```{r Ir, cache=TRUE, echo=FALSE}
+Ir <- Deq %>% 
+  group_by(sim) %>% 
+  arrange(time) %>% 
+  mutate_at(vars(-time, -sim), cumsum) %>% 
+  filter(time == 600)
 ```
@@
 Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 20.7 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].
+In addition to the direct impact of human on forests, a more insidious disturbance is climate change [@Herault2016].
 Understanding the drivers of tropical forest resilience to disturbance is thus critical.
+Here we explored the role of biodiversity in tropical forests answer to disturbances through a simulation approach.
 
@@
 Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@Cadotte2011; @Ives2007; @Loreau2013].
+Finally, climate change effect on tropical areas results in a significant risk of precipitation decrease [@Davidson2012], whereas intense drought events has been shown to result in huge mortality rates and biomass loss [@Phillips2009].
 
@@
 The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2009], or observations along natural gradient of biodiversity [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017].
-However, experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Schnitzer2016], rarely include logged forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018].
-Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014; @Sakschewski2016], but the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].
+However, experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Schnitzer2016], rarely include disturbed forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018].
+Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014; @Sakschewski2016] However, the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].
 
@@
 We used a simulation approach to assess the effect of species richness and functional composition on long-term forest recovery from disturbances of different levels.
+Generally, a disturbance can be characterized by a loss of individuals at a specific time.
+Individual loss can be more or less targeted on individual characteristics (i.e.
+species, diameter...) and time can be cyclic or not.
+To have an overview of disturbance effect without confounding effect of disturbance type specificity, we used untargeted disturbance happening only once.
 We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance.
 In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned it into complementarity and selection effects.
-Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance due to an increased productivity.
+Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity.
 
@@
 TROLL is an individual-based and spatially explicit forest dynamic model in which the life cycle of individual trees (recruitment, growth, seed production and death) $>1cm$ diameter is simulated.
-Trees grows in a forest light environment explicitly computed within a $1m^3$ resolution voxel grid.
+Trees grow in a forest light environment explicitly computed within a $1m^3$ resolution voxel grid.
 Seedlings below the $1cm$ size class are not modeled explicitly, but are part of a seed/seedlings pool.
@@
 A new disturbance module was developed for this study.
-At a given iteration $disturb_{iter}$, individual trees are randomly picking following a uniform law and without any individuals nor species targeting.
+In order to have a general overview of disturbance effect on forest recovery, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificity.
+At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law and without any individuals nor species targeting.
 Selected individuals are then removed without triggering a treefall to avoid any side effect.
@@
 In order to assess the role of biodiversity on ecosystem recovery from disturbance, we  simulated forest trajectories for different levels of disturbance intensity and tree community compositions.
-For each simulation, we first simulated forest regeneration from bare soil for 600 years [pre-disturbance step, @Marechaux2017], then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step).
+For each simulation, we first simulated forest regeneration from bare soil for 600 years (pre-disturbance step), then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step).
 Disturbance intensity was quantified by the percentage of basal area loss, here fixed at 0% (control), 25%, 50% and 75%.
 For each level of disturbance, we performed simulations differing in the original number of species ($SR=15,25,125$) and composition.
-For each level of species richness, we created 1000 virtual communities by randomly picking desired number of species among the regional species pool.
-We used 20 communities for simulations with a large range of convex hull volume $CVH$ but with a community weighted mean close to the regional species pool community weighted mean.
+We proceeded in two steps: (i) for each level of species richness, we created 1000 virtual communities by randomly picking desired number of species among the regional species pool; (ii) among those 1000 communities, we selected 20 communities for simulations with a large range of convex hull volume $CVH$ [@Cornwell2006a] but with a community weighted mean close to the regional species pool community weighted mean in order to observe change in functional diversity and not in functional composition.
 This led to 240 simulations varying in community composition and disturbance intensity ($3~SR*20~CHV*4~ levels~of~disturbance$).
+We focused on functional richness with convex hull volume $CVH$ and functional mean with community weighted mean $CWM$ to define simulation communities.
 
@@
 
-We focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$ to define simulation communities.
-Whereas species and functional diversities were assessed for each simulation at the end of the pre-disturbance step with species richness ($SR$) and @villeger_dynamique_2008 indices ($FRIC$, $FEve$, and $FDiv$).
+Species and functional diversities were assessed for each simulation at the end of the pre-disturbance step with species richness ($SR$) and @villeger_dynamique_2008 indices ($FDiv$, the volume of the functional space occupied by the community, and $FEve$, the regularity of the distribution of abundance in this volume) using the `FD` R package [@laliberte_distance-based_2010-1].
 
@@
 
-**We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).**
+We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).
 
-The recovery of the simulated forest ecosystem at a time $t(R(t))$ regarding a given properties $X$ was defined ad follows:
+The recovery of the simulated forest ecosystem, i.e.
+the ecosystem capacity to recover from a disturbance event, at a time $t$, $R(t)$ regarding a given property $X$ was defined ad follows (Figure \@ref(fig:IrGraphFinal)):
 
 \begin{equation}
-  R\left(t\right)= \frac{X_T(t)}{X_C(t)}
+  R[X(t)] = \frac{X_D(t)}{X_C(t)}
   (\#eq:Recovery)
@@
 
-where $X_T(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$).
-Note that demographic stochasticity in the model induces a low variability among replicated simulations with the same set-up so that **XX** [@Marechaux2017].
+where $X_D(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$).
+Note that demographic stochasticity in the model induces a low variability among replicated simulations with the same set-up so that we did not include repetition of virtual communities simulations [@Marechaux2017].
+
 When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery.
 We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$.
-**Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for (the two functions described above: (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight).
-We then integrate the euclidean distances over time to get our ecosystem recovery indicators.** (Figure \@ref(fig:datatrans)).
-We used the cumulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem recovery.
-A high value of $Ieq_{600}$ indicating a slow recovery.
-The influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem recovery index $Ieq_{600}$  was finally investigated by linear regression ($Ieq_{600} \sim SR + FEve + FDiv$).
+Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance, i.e.
+the ecosystem capacity to absorb a disturbance event.
+Consequently in this particular case, we can use the cummulated recovery as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):
 
-```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation.
-Ecosystem outputs (**A**) are normalized by the control value over time to calculate recovery (**B**); recovery of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**).
-', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
-cols <- RColorBrewer::brewer.pal(3, "YlOrRd")
-g_agb <- ecosystem %>% 
-  left_join(., doe, by = 'sim') %>%
-  filter(richness == 125) %>% 
-  filter(cvh == 1) %>% 
-  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
-  geom_point() +
-  xlab('Time (years)') +
-  ylab('Aboveground biomass (tonC/ha)') +
-  scale_color_manual('Disturbance\nintensity\n(% of BA)',
-                       values=c("#999999", cols))
-g_resilience <- resilience %>% 
-  left_join(., doe, by = 'sim') %>%
-  filter(richness == 125) %>% 
-  filter(cvh == 1) %>% 
-  ggplot(aes(x = time, y = agb, color = disturbance)) +
-  geom_hline(yintercept = 1, col = 'red') +
-  geom_point() +
-  xlab('Time (years)') +
-  ylab('Resilience of aboveground biomass') +
-  scale_color_manual('Disturbance\nintensity\n(% of BA)',
-                     values = cols)
-g_deq <- deq %>% 
-  left_join(., doe, by = 'sim') %>%
-  filter(richness == 125) %>% 
-  filter(cvh == 1) %>% 
-  ggplot(aes(x = time, y = structure, 
-             color = disturbance, fill = disturbance)) +
-  geom_hline(yintercept = 0, col = 'red') +
-  geom_point() +
-  geom_area(alpha = 0.2) +
-  xlab('Time (years)') +
-  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
-  scale_color_manual('Disturbance\nintensity\n(% of BA)',
-                     values = cols) +
-  scale_fill_manual('Disturbance\nintensity\n(% of BA)',
-                     values = cols)
-g_Ieq <- Ieq %>% 
-  left_join(., doe, by = 'sim') %>%
-  filter(richness == 125) %>% 
-  filter(cvh == 1) %>% 
-  ggplot(aes(x = time, y = Ieq.structure, color = disturbance)) +
-  geom_point() +
-  xlab('Time (years)') +
-  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
-  scale_color_manual('Disturbance\nintensity\n(% of BA)',
-                     values = cols)
-cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
+\begin{equation}
+  I_R(X) = \int _{t=0} ^{t=600} \frac{X_D(t)}{X_C(t)}.dt
+  (\#eq:Resilience)
+\end{equation}
+
+A high value of $I_R$ indicating a slow recovery.
+Ecosystem euclidean distance to equilibrium and resilience index $I_R$ was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).
+The influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem resilience index $I_R$  was finally investigated by linear regression ($I_R \sim SR + FEve + FDiv$).
+
+```{r IrGraphFinal, fig.cap="Definition of recovery function and resilience indice shown on aboveground biomass dynamics of a control and disturbed ecosystem before and after the disturbance event."}
+knitr::include_graphics("images/final.png")
 ```
@@
 
-The net effect of biodiversity on ecosystem recovery (NE) is defined as the difference between the ecosystem recovery of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e.
-ecosystem recovery for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001].
+The net effect of biodiversity ($NE$) on ecosystem property $X$ is defined as the difference between the ecosystem property $X$ of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e.
+ecosystem property $X$ for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001].
 We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:
@@
 But similarly, to recovery measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
-So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure treatments net effect recovery $R(NE_T)$:
+So in order to correctly assess selection and complementarity effect role in tropical forest resilience, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure treatments net effect recovery $R(NE_T)$:
 
@@
 
-Species richness was not significantly influencing $Ieq_{600}$ with forest structure and had a small positive effect on forest production (Table \@ref(tab:tgIeq)).
-Increased functional diversity reduced cumulative integral from ecosystem distance to equilibrium after 600 years ($Ieq_{600}$) and functional evenness was complementary reducing $Ieq_{600}$ (Figure \@ref(fig:gIeq) and Table \@ref(tab:tgIeq)).
-Finally, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$.
-We found similar results for all disturbance levels (see [Appendix 5: Disturbance simulations]).
+Species richness was not significantly influencing ecosystem resilience index $I_R$ for both forest structure and forest production (Table \@ref(tab:tgIr)).
+Increased functional diversity reduced  ecosystem resilience index $I_R$ and functional evenness was complementary reducing $I_R$ (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
+Finally, low species richness could result in variant $I_R$, but increased species richness resulted in increased functional diversity and consequently lower $I_R$.
+We found similar results for all disturbance levels (see [Appendix S4: Linear models]).
 
-```{r gIeq, echo=FALSE, fig.cap='Ecosystem recovery after 600 years with taxonomic and functional diversity for different levels of disturbance.
-Cumulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].
-Solid lines and grey areas show fitted relationships with their mean and variance of the form $Ieq_600 = FDiv + \\epsilon$.', cache=TRUE}
-Ieq %>% 
-  left_join(., doe, by = 'sim') %>% 
-  filter(time == 600) %>% 
-  ggplot(aes(x = FDiv, y = Ieq.structure, label = sim)) + 
+```{r gIr, echo=FALSE, fig.cap='Ecosystem resilience $I_R$ with taxonomic and functional diversity for different levels of disturbance.
+Resilience indice $I_R$ was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].
+Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.'}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  select(disturbance, 
+         structure, SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
+  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
   geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
-  geom_point(aes(shape = richness, color = FEve)) +
+  geom_point() +
     facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
   xlab('Functional diversity (FDiv)') +
-  ylab('Cummulative integral from ecosystem 
-       distance to forest structure equilibrium after 600 years') +
+  ylab('Resilience indice Ir') +
   scale_shape_discrete('Species\nrichness\n(SR)') +
@@
 
-```{r tgIeq, cache=T}
-Ieq %>% 
-  left_join(., doe, by = 'sim') %>% 
-  mutate(SR = as.integer(richness)) %>% 
-  select(disturbance, Ieq.structure, Ieq.production, SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"),
-                 variable.name = "Ieq.type", value.name = "Ieq.value") %>% 
-  group_by(Ieq.type, disturbance) %>%
-  do(fit = lm(Ieq.value ~ SR + FEve + FDiv, data = .)) %>% 
+```{r tgIr, cache=T}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  mutate(SR = as.numeric(SR)) %>% 
+  select(disturbance, structure, production,
+         SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
+  group_by(variable, disturbance) %>%
+  do(fit = lm(Ir ~ SR + FEve + FDiv, data = .)) %>% 
   tidy(fit) %>% 
@@
   mutate(Coefficient = round(estimate,2)) %>% 
-  mutate(Recovery = gsub("Ieq.", "", Ieq.type)) %>% 
+  mutate(Recovery = variable) %>% 
   rename(Variable = term, `Disturbance intensity` = disturbance) %>%
   select(Recovery, `Disturbance intensity`, Variable, Coefficient, `P-value`) %>% 
-  kable(caption = "Result of linear regression testing the effect of species richness and functional diversity indices on the recovery index by recovery type and disturbance intensity.") %>% 
+  mutate(value = paste(Coefficient, `P-value`)) %>% 
+    mutate(value = cell_spec(value, 
+                           bold = ifelse(grepl("*", value, fixed = T), T, F))) %>%
+  reshape2::dcast(Recovery + `Disturbance intensity` ~ Variable, value.var = "value") %>% 
+  kable(caption = "Result of linear regression testing the effect of species richness and functional diversity indices on the recovery index by recovery type and disturbance intensity.", escape = F) %>%
   kable_styling(full_width = F) %>% 
@@
 
-The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, see Table \@ref(tab:tNE)) for every ecosystem global variable, e.g.
+The recovery from net effect of biodiversity on ecosystem properties was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, see Table \@ref(tab:tNE)) for every ecosystem global variable, e.g.
 aboveground biomass (see Figure \@ref(fig:gBE)).
@@
 Finally, 600 years after disturbance, the biodiversity net effect on aboveground biomass was still not at equilibrium for a disturbance intensity greater than 25% of basal area.
-We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see Supplementary materials S2).
-Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionally to disturbance), and its net effect was maintained by complementarity effect (see Supplementary materials S2).
+We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning).
+Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionally to disturbance), and its net effect was maintained by complementarity effect.
 
@@
   left_join(., doe, by = 'sim') %>%
-  filter(richness == 125) %>%
+  filter(SR == 125) %>%
   filter(disturbance == 0) %>%
@@
 Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their recovery over time.
-Solid lines represents the mean value whereas areas represents the interval of values among simulations.', cache=TRUE}
+Solid lines represents the mean value whereas areas represents the interval of values among simulations.'}
 BEresilience %>% 
   left_join(., doe, by = 'sim') %>%
-  filter(richness == 125) %>% 
+  filter(SR == 125) %>% 
   # filter(BE %in% c('SE', 'CE')) %>%
-  group_by(time, BE, disturbance, richness) %>%
+  group_by(time, BE, disturbance, SR) %>%
   summarise(qt1 =quantile(agb, probs=0.05),
@@
   geom_line(aes(y = mean, color = BE)) +
-  facet_grid(disturbance ~ richness, labeller = 'label_both') +
+  facet_grid(disturbance ~ SR, labeller = 'label_both') +
   xlab('Time (years)') +
@@
 An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.
-
-<!-- OK good point. But you need to go more into the details of the diversity effect, perhaps highlighting the key choice of the traits? BH -->
 
@@
 
-We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest recovery from disturbance, and thus in tropical forest resilience.
-We found that diversity improved forest recovery, similarly to productivity [@Liang2016; @Poorter2015].
-Additionally, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure the final forest recovery.
+We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance.
+We found that diversity improved forest resilience, similarly to productivity [@Liang2016; @Poorter2015].
+Additionally, we found that complementarity between species was insuring forest resilience in forest succession start with facilitation, before more productive species dominate the forest and insure the final forest resilience.
+
+Our results advocate for an important interplay of diversity and dominant species in forest resilience depending on the recovery stage, similarly to findings of @Sakschewski2016.
+Moreover, we studied a generic form of disturbance, and our findings might be modulated depending on the disturbance type.
+For instance, selective logging will induce both species targeting diminishing selection effect if dominant species are chosen and cyclic perturbation, increasing the role of early recovery where complementarity effect play the most.
+Consequently, we could hypothesize that biodiversity will be more of primary importance regarding selective logging compared to less targeted disturbances with longer cycles such as drought, fires and hurricanes.
 
@@
     + [PDF](S1-LL.pdf)
-* Supplementary materials S2: Resilience indice
+* Supplementary materials S2: Resilience index
     + [HTML](S2-Resilience.html)

### 2018-11-15 B1-Biblio.Rmd
Draft 0.1.9008 Fifth version (IM references)

---
title: "B1: Bibliography"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 8,
                cache = T, cache.lazy = F)
```

# Biodiversity & Ecosystem resilience / stability

## @Morin2014a

* First our results show that variability in productivity between stands differing in species composition decreases as species richness and functional diversity increase.
* Stability increases with increasing diversity due to compensatory dynamics across species, supporting the biodiversity insurance hypothesis.

## Others

[x] @Sakschewski2016 already included

# Temporal dynamic of complementarity and selection effects

## @Fargione2007

* 10 years of grasslands
* effect of biodiversity on plant biomass production shifted from sampling to complementarity over time
* with more efficient N use at higher diversity

## @Cardinale2007

* Thus, much like the tradeoff between yield and stability in an investment portfolio, a select few species may be able to produce higher yield than a diverse community, but this might come at the expense of the stability of yield through time

# Glossary example & Introduction

[x] @Niklaus2017a

# Selection effect > Complementarity effect

## @Bauters2015

* Monospecific and two species mixture plots of a 77-year-old tree diversity experiment in Yangambi in the Congo basin
* the proportion of planted species in the total stand basal area (BApl) and effective species richnesswere identified as compositional parameters with a significant effect on the aboveground carbon (AGC), with BApl being more important
* no clear benefits, but also no drawbacks, for AGC of two speciesmixture plots over monospecific plots

## @Finegan2015

* Data are presented for three rain forests in Bolivia, Brazil and Costa Rica. Initial above-ground biomass and biomass increments of survivors, recruits and survivors + recruits (total) were estimated for trees ≥10 cm d.b.h. in 62 and 21 1.0-ha plots, respectively.
* We found no support for the niche complementarity hypothesis and support for the green soup hypothesis only for biomass increments of recruits. We have strong support for the biomass ratio hypothesis. CWM Hmax is a strong driver of ecosystem biomass and carbon storage and CWM SLA, and other CWM leaf traits are especially important for biomass increments and carbon sequestration.

## Others

* [x] @Chiang2016 already included

## @Lohbeck2016

* __stable ecosystem__
* human- modified tropical forest landscape in Southern Mexico.
* We quantified aboveground standing biomass, primary productivity, litter production, and wood decomposition at the landscape level, and evaluated the extent to which tree species contribute to these ecosystem functions.
* We found that dominance was more important than species traits in determining a species’ contribution to ecosystem functions.

## @Prado-Junior2016

* __stable ecosystem__
* above-ground net biomass change and its underlying biomass dynamics (biomassrecruitment, growth and mortality)
* soil fertility, biomass ratio, niche complementarity and vegetation quantity hypotheses.
* 5-year period in 200 permanent sample plots in 8 tropical dry forests in Brazil
* dominance of conser-vative species, rather than of acquisitive species, is associated with high biomass growth and storage
* Biomass stor-age was favoured by low- rather than high trait diversity, as most of the biomass is concentrated in species with large stem diameter and high wood density

## @Poorter2017

* aboveground biomass (AGB) and biomass dynamics (i.e., net biomass productivity and its underly-ing demographic drivers: biomass recruitment, growth and mortality) to forest attributes (tree diversity, community-mean traits and stand basal area) and environmental conditions (water avail-ability, soil fertility and disturbance)
* 26 sites, 201 1-ha plots and >92,000 trees distributed across the Neotropics
* vegetation attributes were more frequently associated significantly with biomass stocks and dynamics than environmental conditions (in 67 vs. 33% of the relationships)
* light disturbance increased biomass growth
* Rarefied tree species richness had consistent positive relationships with biomass stocks and dynamics
* Community-mean traits were good predictors of biomass stocks and dynamics

## @VanderSande2018

* effects of abiotic factors (light variation, caused by logging distur-bance, and soil fertility) and biotic factors (species richness and functional trait composition) on biomass stocks (above- ground biomass, fine root biomass), SOM and productivity in a relatively monodominant Guyanese tropical rainforest
* Disturbance increased productivity but decreased above- ground biomass stocks
* As expected, species richness did not affect productivity

# Others

## @DeAvila2018

* unclear how above- ground biomass (AGB) recovery of species- rich tropical forests is affected by disturbance intensity and post- disturbance (remain-ing) tree- community attributes
* ($\DeltaAGB$) decreases with management- related disturbance intensity (disturbance hypothesis), and increases with the diversity (niche- complementarity hypothesis) and the community-weighted mean (CWM) of acquisitive traits of dominant species (biomass- ratio hypothesis)
* AGB increment of survivors and of the total tree community increased with basal area remaining
* We did not find consistent support for the niche- complementarity and biomass- ratio hypotheses

## @VanderSande2017a

* __Review__
* empirical studies that analyzed the relationships between plant biodiversity attributes and carbon stocks and carbon uptake in tropical forests. Our results show that biodiversity attributes related to species, traits or structure signifi-cantly affect carbon stocks or uptake in 64% of the evaluated relationships
* biodi- versity is not only a co-benefit, but also a requirement for short- and long-term maintenance of carbon stocks and enhancement of lighting the importance of performing a comprehensive analysis that adequately accounts for environmental drivers. In summary, biodi- versity is not only a co-benefit, but also a requirement for short- and long-term maintenance of carbon stocks and enhancement of uptake

## @Huang2018

* manipulated tree species richness by planting more than 150,000 trees in plots with 1 to 16 species
* richness strongly increased stand-level productivity. After 8 years, 16-species mixtures had accumulated over twice the amount of carbon found in average monocultures and similar amounts as those of two commercial monocultures
* increasing complementarity effect with null to negative selection effect depending on species richness
* faster growth of live trees inmore diverse stands because tree survival rate did not increase with species richness
* 10% decrease in tree spe-cies richness would lead to a 2.7% decrease in for-est productivity on average (24), which is within the range of productivity decreases (2.1 to 3.1%) reported for the same tree species loss scenario in a recent observational study that used plot data covering a large part of the global forests (3). In that study, itwas estimated that such a loss would correspond to around $20 billion per year of com-mercial wood production.

# References

### 2018-11-15 draft.Rmd
Draft 0.1.9008 Fifth version (IM references)

@@
 output:
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
+  bookdown::word_document2: default
+  github_document: default
   bookdown::pdf_document2:
@@
     keep_tex: true
-  bookdown::word_document2: default
-  github_document: default
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
 linestretch: 1.5
@@
 Among endogenous drivers, forest pre-disturbance composition has been identified as a key driver of forest recovery [@Herault2018], as different species respond differently to perturbations depending on their ecological strategies and functional properties [@herault_functional_2011].
-Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@Cadotte2011; @Ives2007; @Loreau2013].
+Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@DeAvila2018; @Cadotte2011; @Ives2007; @Loreau2013].
 Finally, climate change effect on tropical areas results in a significant risk of precipitation decrease [@Davidson2012], whereas intense drought events has been shown to result in huge mortality rates and biomass loss [@Phillips2009].
@@
 
-The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2009], or observations along natural gradient of biodiversity [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017].
-However, experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Schnitzer2016], rarely include disturbed forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018].
-Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014; @Sakschewski2016] However, the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].
+The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], or observations along natural gradient of biodiversity [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017].
+Whereas several studies advocate only for the importance of few dominant species through selection effect to regulate ecosystem functionning in an old tree diversity experiment [@Bauters2015], in wet or dry natural forests [@Finegan2015; @Prado-Junior2016], or in human-modified forests [@DeAvila2018; @Lohbeck2016].
+Numerous authors highlighted diversity role through complementarity, alone or in addition to selection.
+@Morin2014a identified by simulation an increases of ecosystem stability with species richness.
+@Fargione2007 observed a shift from sampling to complementarity effect, when @Huang2018 observed an increased complementarity with a null selection effect in forest establishments experiments.
+And @Cardinale2007; @Poorter2017; @VanderSande2017a reviewed grasslands and forests experiments and study sites and highlighted the role of both complementarity and selection for ecosystem biomass stock and dynamic.
+However, experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Poorter2017; @Schnitzer2016], rarely include disturbed forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018].
+Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014a; @Sakschewski2016] However, the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].
 
@@
 We used a simulation approach to assess the effect of species richness and functional composition on long-term forest recovery from disturbances of different levels.
-Generally, a disturbance can be characterized by a loss of individuals at a specific time.
-Individual loss can be more or less targeted on individual characteristics (i.e.
-species, diameter...) and time can be cyclic or not.
-To have an overview of disturbance effect without confounding effect of disturbance type specificity, we used untargeted disturbance happening only once.
+A disturbance can be characterized by a loss of individuals, targeted or not (i.e.
+species, diameter...), at a specific time, cyclic or not.
+To have an overview of disturbance effect without confounding effect of disturbance specificities, we used untargeted disturbance happening only once.
 We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance.
 In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned it into complementarity and selection effects.
-Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity.
+Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018].
 
@@
 A new disturbance module was developed for this study.
-In order to have a general overview of disturbance effect on forest recovery, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificity.
+In order to have a general overview of disturbance effect on forest recovery, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificities.
 At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law and without any individuals nor species targeting.
@@
 
-In the limit of the model, we were able to show that functional diversity improved tropical forest recovery from disturbance.
+In the limit of the model, we were able to show that functional diversity improved tropical forest resilience from disturbance.
 More particularly, functional diversity and evenness are key components of diversity in forest recovery trajectory after disturbance.
 Moreover, we found that complementarity between species was insuring forest recovery in the early times before more productive species dominate the forest and insure late recovery.
-We thus advocate the view that both diversity theory (with niche complementarity and facilitation) and mass ratio theory (represented by selection) are implied in ecosystem recovery but with relative importance depending on the spatial grain and time since the disturbance event.
+We thus advocate the view that both diversity theory (with niche complementarity and facilitation) and mass ratio theory (represented by selection) are implied in ecosystem resilience but with relative importance depending on the spatial grain and time since the disturbance event.
 
@@
 Indeed, high functional diversity combined to high evenness avoid hegemonic effect of a few dominant species.
-Our results thus confirms both the review of @Diaz2001,advocating for an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity and resilience of terrestrial ecosystems.
-Finally, species richness is not directly increasing resilience, thus supporting the idea that species diversity is often an inadequate surrogate of functional diversity [@Diaz2001; but see @Poorter2015].
-Still, increased species richness will increase chance for high functional diversity through the sampling effect [@Loreau1998].
+Our results thus confirms both the review of @Daz2001a, advocating for an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity and resilience of terrestrial ecosystems.
+Finally, species richness is not directly increasing resilience, thus supporting the idea that species diversity is often an inadequate surrogate of functional diversity [@Daz2001a; but see @Poorter2015], which might explain the absence of diversity effect on resilience observed by @VanderSande2018.
+Still, increased species richness will increase chance for high functional diversity through the sampling effect [@Loreau1998; @Morin2014a].
 An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.
@@
 Our results follows several studies advocating for multiple mechanisms through which functional diversity affect ecosystem functions.
-Complementarity effect encompass niche complementarity theory (or diversity theory), which has been shown to indirectly improve forest biomass and coarse wood productivity in subtropical forest [@Chiang2016].
-Similarly, complementarity has been shown of primary importance in grassland functioning [@Cardinale2007; @Isbell2011].
+Complementarity effect encompass niche complementarity theory (or diversity theory), which has been shown to improve forest biomass [@Poorter2017; @Huang2018] and coarse wood productivity [@Chiang2016] in tropical and subtropical forests.
+Similarly, complementarity has been shown of primary importance in grassland functioning [@Cardinale2007; @Fargione2007; @Isbell2011].
 On the other hand, selection effect can be related to the mass ratio hypothesis, which proposes that ecosystem functions are regulated by dominant species of the community [@Grime1998].
-And mass ratio theory has been shown to primarily influence ecosystem functions with biomass accumulation (aboveground and shoots), productivity (coarse wood or aboveground), water use, and light interception for both grassland and subtropical forests [@Chiang2016; @Mokany2008].
+And mass ratio theory has been shown to primarily influence ecosystem functions with biomass accumulation (aboveground and shoots), productivity (coarse wood or aboveground), water use, and light interception for both grassland and forests [@Finegan2015; @Lohbeck2016; @Prado-Junior2016; @Chiang2016; @Mokany2008].
 Consequently, concerning mechanisms responsible for the positive effect of functional diversity on tropical forest functions after disturbance, we also advocate for the interplay of both niche complementarity and mass ratio theories, depending on time since the disturbance event and the spatial grain.
@@
 They also found that selection effect was greater than complementarity in most of the cases, promoting the idea that mass ratio theory is the primary mechanisms involved in functional diversity effect on ecosystem functions [@Mokany2008].
+Besides @Huang2018 highlighted increased complementarity over null selection effect resulting in higher productivity due to tree growth in an 8-years forest experiment of China.
 
@@
 Our results advocate for an important interplay of diversity and dominant species in forest resilience depending on the recovery stage, similarly to findings of @Sakschewski2016.
-Moreover, we studied a generic form of disturbance, and our findings might be modulated depending on the disturbance type.
+Moreover, we studied a generic form of disturbance, and our findings might be modulated depending on the disturbance specificities.
 For instance, selective logging will induce both species targeting diminishing selection effect if dominant species are chosen and cyclic perturbation, increasing the role of early recovery where complementarity effect play the most.

### 2018-11-15 draft.Rmd
spell check

@@
 The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], or observations along natural gradient of biodiversity [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017].
-Whereas several studies advocate only for the importance of few dominant species through selection effect to regulate ecosystem functionning in an old tree diversity experiment [@Bauters2015], in wet or dry natural forests [@Finegan2015; @Prado-Junior2016], or in human-modified forests [@DeAvila2018; @Lohbeck2016].
+Whereas several studies advocate only for the importance of few dominant species through selection effect to regulate ecosystem functioning in an old tree diversity experiment [@Bauters2015], in wet or dry natural forests [@Finegan2015; @Prado-Junior2016], or in human-modified forests [@DeAvila2018; @Lohbeck2016].
 Numerous authors highlighted diversity role through complementarity, alone or in addition to selection.
@@
 species, diameter...), at a specific time, cyclic or not.
-To have an overview of disturbance effect without confounding effect of disturbance specificities, we used untargeted disturbance happening only once.
+To have an overview of disturbance effect without confounding effect of disturbance specificity, we used untargeted disturbance happening only once.
 We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance.
@@
 A new disturbance module was developed for this study.
-In order to have a general overview of disturbance effect on forest recovery, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificities.
+In order to have a general overview of disturbance effect on forest recovery, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificity.
 At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law and without any individuals nor species targeting.
@@
 Our results advocate for an important interplay of diversity and dominant species in forest resilience depending on the recovery stage, similarly to findings of @Sakschewski2016.
-Moreover, we studied a generic form of disturbance, and our findings might be modulated depending on the disturbance specificities.
+Moreover, we studied a generic form of disturbance, and our findings might be modulated depending on the disturbance specificity.
 For instance, selective logging will induce both species targeting diminishing selection effect if dominant species are chosen and cyclic perturbation, increasing the role of early recovery where complementarity effect play the most.

### 2018-11-30 S2-Resilience.Rmd
Draft 0.1.9008 Final 1 Intro, M&M and 1/2 Res

@@
 
-```{r IrGraph, eval=F, include=F}
+```{r IrGraph1, eval=F, include=F}
 mature %>% 
@@
   filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
+  mutate(agb = agb/10^3) %>% 
   reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
@@
   ylab("Aboveground biomass (tonC/ha)") +
-  scale_color_discrete("Simulation") +
   scale_color_discrete("Disturbance\nintensity\n(% of BA)", 
                        labels = c("0 (control)", "50 (disturbed)")) +
-  ggsave(filename = "./images/original.png", width = 20, height = 10, units = "cm")
+  ggsave(filename = "./images/original1.png", width = 20, height = 10, units = "cm")
+```
+
+```{r IrGraph2, eval=F, include=F}
+bind_rows(
+  mature %>% 
+    select(time, sim, agb) %>% 
+    separate(sim, c("SR", "CVH"), remove = F) %>% 
+    mutate(disturbance = 50) %>% 
+    rbind(ecosystem %>% 
+            select(time, sim, agb) %>% 
+            separate(sim, c("SR", "CVH", "disturbance"), remove = F)) %>% 
+    filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
+    mutate(agb = agb/10^3) %>% 
+    reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
+    rename(control = `0`, treatment = `50`) %>% 
+    mutate(control = ifelse(is.na(control), treatment, control)) %>% 
+    mutate(type = "Raw"),
+  mature %>% 
+    select(time, sim, agb) %>% 
+    separate(sim, c("SR", "CVH"), remove = F) %>% 
+    mutate(disturbance = 50) %>% 
+    rbind(ecosystem %>% 
+            select(time, sim, agb) %>% 
+            separate(sim, c("SR", "CVH", "disturbance"), remove = F)) %>% 
+    filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
+    mutate(agb = agb/10^3) %>% 
+    reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
+    rename(control = `0`, treatment = `50`) %>% 
+    mutate(control = ifelse(is.na(control), treatment, control)) %>% 
+    mutate(treatment = treatment/control) %>% 
+    mutate(control = control/control) %>%
+    mutate(type = 'Normalized by control')
+) %>% 
+  mutate(type = factor(type, levels=c('Raw', 'Normalized by control'))) %>% 
+  ggplot(aes(x = time)) +
+  geom_line(aes(y = control, color = "control")) +
+  geom_line(aes(y = treatment, color = "disturbed")) +
+  geom_ribbon(aes(ymax = control, ymin = treatment), 
+              fill = "lightgrey", col = NA, alpha = 0.5) +
+  xlab("Time (years)") +
+  ylab("Aboveground biomass (tonC/ha)") +
+  scale_color_discrete("Disturbance\nintensity\n(% of BA)", 
+                       labels = c("0 (control)", "50 (disturbed)")) +
+  facet_wrap(~type, nrow = 2, scales = "free") +
+  ggsave(filename = "./images/original2.png", width = 20, height = 20, units = "cm")
 ```

### 2018-11-30 draft.Rmd
Draft 0.1.9008 Final 1 Intro, M&M and 1/2 Res

@@
 ---
-title: Biodiversity improves tropical forest resilience, shifting from complementarity to selection effect
+title: Biodiversity improves tropical forest resilience
 author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
@@
     toc_float: yes
-  bookdown::word_document2: default
-  github_document: default
   bookdown::pdf_document2:
@@
     keep_tex: true
+  github_document: default
+  bookdown::word_document2: default
 linestretch: 1.5
@@
   arrange(time) %>% 
-  mutate_at(vars(-time, -sim), cumsum) %>% 
+  mutate_at(vars(-time, -sim), funs(1/cumsum(.))) %>% 
   filter(time == 600)
@@
 
-Natural disturbances caused by tree death play a major role in tropical forest functioning through gap-phase regeneration [@Veblen1992].
-Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015].
-If deforestation has been rightly the focus of worldwide attention, forest degradation has been less studied and quantified through tropics [@Putz2010; @Simula2009].
-However, degraded forests cover areas over ten times larger than deforestation [@Herold2011], a majority of which is being logged for timber production [@Blaser2011].
-Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 20.7 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].
-In addition to the direct impact of human on forests, a more insidious disturbance is climate change [@Herault2016].
+Natural disturbances play a major role in tropical forest functioning through gap-phase regeneration [@Veblen1992].
+Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015] and while deforestation has been rightly the focus of worldwide attention, the consequences of human-induced disturbance on tropical forests has been less studied [@Putz2010; @Simula2009].
+However, degraded forests currently cover areas over ten times larger than deforested ones [@Herold2011], a majority of which is being unsustainably logged for timber production [@Blaser2011].
+Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 20.7 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing, for instance, 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].
+In addition to the direct impact of human on forests, a more insidious disturbance is climate change that modifies regional climates with, for instance, projected increases in the frequency and intensity of convective storms in the Amazon having the potential to increase wind-related tree mortality [@Negron-Juarez2018].
 Understanding the drivers of tropical forest resilience to disturbance is thus critical.
-Here we explored the role of biodiversity in tropical forests answer to disturbances through a simulation approach.
+Here we explored the role of biodiversity in tropical forest recovery after disturbance through a simulation approach.
 
-Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010].
+The concept of ecological resilience has met numerous definitions, but encompass both ecosystem ability to adsorb, the resistance component, and recover from disturbance [@Gunderson2010].
 Both exogenous (e.g.
@@
 Among endogenous drivers, forest pre-disturbance composition has been identified as a key driver of forest recovery [@Herault2018], as different species respond differently to perturbations depending on their ecological strategies and functional properties [@herault_functional_2011].
-Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@DeAvila2018; @Cadotte2011; @Ives2007; @Loreau2013].
-Finally, climate change effect on tropical areas results in a significant risk of precipitation decrease [@Davidson2012], whereas intense drought events has been shown to result in huge mortality rates and biomass loss [@Phillips2009].
+Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@Cadotte2011; @Ives2007; @Loreau2013].
+Finally, disturbance intensity itself has been shown to be a key component of forest resilience [@DeAvila2018].
 
-Two types of mechanisms can underlie an effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001].
-Complementarity effect results from resources partitioning and niche differentiation, and facilitation, leading to a more efficient use of available resources.
-Selection effect results from a higher probability of more diverse community to include particularly efficient species regarding ecosystem process or stability.
+Two types of mechanisms can underlie the effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001].
+Complementarity effect results from resources partitioning and niche differentiation, and facilitation, leading to a more efficient use of available resources at the community level.
+Selection effect results from a higher probability of more diverse community to include particularly efficient species regarding ecosystem processes.
 Moreover the sampling effect additionally implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].
 
-The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], or observations along natural gradient of biodiversity [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017].
-Whereas several studies advocate only for the importance of few dominant species through selection effect to regulate ecosystem functioning in an old tree diversity experiment [@Bauters2015], in wet or dry natural forests [@Finegan2015; @Prado-Junior2016], or in human-modified forests [@DeAvila2018; @Lohbeck2016].
-Numerous authors highlighted diversity role through complementarity, alone or in addition to selection.
-@Morin2014a identified by simulation an increases of ecosystem stability with species richness.
-@Fargione2007 observed a shift from sampling to complementarity effect, when @Huang2018 observed an increased complementarity with a null selection effect in forest establishments experiments.
-And @Cardinale2007; @Poorter2017; @VanderSande2017a reviewed grasslands and forests experiments and study sites and highlighted the role of both complementarity and selection for ecosystem biomass stock and dynamic.
-However, experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Poorter2017; @Schnitzer2016], rarely include disturbed forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018].
-Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014a; @Sakschewski2016] However, the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].
+The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], or direct observations along natural gradient [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017].
+Several studies advocate for the importance of few dominant species through selection effect to regulate ecosystem functioning in an old tree diversity experiment [@Bauters2015], in wet or dry natural forests [@Finegan2015; @Prado-Junior2016], or in human-modified forests [@DeAvila2018; @Lohbeck2016].
+Numerous authors have, however, highlighted the pivotal role of diversity in ecosystem stability [@Morin2014a] and highlighted the role of both complementarity and selection for ecosystem biomass stock and dynamic [@Cardinale2007; @Poorter2017; @VanderSande2017a].
+For instance, @Fargione2007 observed a shift from selection to complementarity effect over time in experimental grasslands, and @Huang2018 observed an increased complementarity over time with no selection effect in forest establishments experiments.
+
+Overall, the role of biodiversity in tropical forest recovery from disturbances and the mechanisms underlying the biodiversity effect on forest functioning along long-term dynamics remain under documented [@VanderSande2017].
+Experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Poorter2017; @Schnitzer2016], rarely include disturbed forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018].
+Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014a; @Sakschewski2016; @Bohn2017].
+However, the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].
 
 In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest, that allow to deal with high diversity levels through a parameterization using plant functional traits [@Marechaux2017].
-Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning *in silico* in tropical forest communities.
-
-We used a simulation approach to assess the effect of species richness and functional composition on long-term forest recovery from disturbances of different levels.
-A disturbance can be characterized by a loss of individuals, targeted or not (i.e.
-species, diameter...), at a specific time, cyclic or not.
-To have an overview of disturbance effect without confounding effect of disturbance specificity, we used untargeted disturbance happening only once.
+Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning  relationships *in silico* in tropical forest communities.
+More specifically, we used a simulation approach to assess the effect of species and functional diversity on long-term forest recovery from disturbances of different levels.
 We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance.
-In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned it into complementarity and selection effects.
+In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned ecosystem properties into complementarity and selection effects.
 Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018].
@@
 
-TROLL is an individual-based and spatially explicit forest dynamic model in which the life cycle of individual trees (recruitment, growth, seed production and death) $>1cm$ diameter is simulated.
-Trees grow in a forest light environment explicitly computed within a $1m^3$ resolution voxel grid.
-Seedlings below the $1cm$ size class are not modeled explicitly, but are part of a seed/seedlings pool.
+TROLL is an individual-based and spatially explicit forest dynamic model in which the life cycle of individual trees (recruitment, growth, seed production and death) $>1cm$ diameter at breast height (dbh) is simulated.
+Trees grow in a forest light environment computed within a $1m^3$ resolution voxel grid.
+Seedlings below the $1cm$ size class are not modeled on an individual basis, but are part of a seed/seedlings pool.
 A maximum of one tree can establish in each $1*1m$ pixel.
-Each tree is defined by its age, diameter at breast height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
+Each tree is defined by its age, dbh, height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
 Tree geometry is calculated with allometric relationships.
 Additionally, each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
-To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which processes are simulated using up-to-date established relationships drawn from the literature.
+To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which ecophysiological processes are simulated using up-to-date established relationships drawn from the literature.
 The model thus represents tree biodiversity in a realistic way, where data requirement and model outputs parallel current trait collection efforts and field inventories, respectively.
@@
 Following an underestimation of leaf lifespan for low LMA species, we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1).
-Natural tree deaths are stochastic events whose background rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010].
+Natural tree deaths are stochastic events whose background rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010] an locally [@Aubry-Kientz2013].
 Additionally, tree can die due to carbon starvation or treefalls.
@@
 A new disturbance module was developed for this study.
-In order to have a general overview of disturbance effect on forest recovery, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificity.
+In order to have a general overview of disturbance effect, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificity.
 At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law and without any individuals nor species targeting.
 Selected individuals are then removed without triggering a treefall to avoid any side effect.
-The operation is repeated until the disturbance result in a defined removed basal area ($disturb_{intensity}$).
+The operation is repeated until the cummulative basal area of removed trees reaches targeted threshold ($disturb_{intensity}$).
 
@@
 
-In order to assess the role of biodiversity on ecosystem recovery from disturbance, we  simulated forest trajectories for different levels of disturbance intensity and tree community compositions.
+In order to assess the role of biodiversity on ecosystem recovery from disturbance, we  simulated forest trajectories for different levels of disturbance intensity ($disturb_{intensity}$) and tree community compositions.
 For each simulation, we first simulated forest regeneration from bare soil for 600 years (pre-disturbance step), then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step).
-Disturbance intensity was quantified by the percentage of basal area loss, here fixed at 0% (control), 25%, 50% and 75%.
-For each level of disturbance, we performed simulations differing in the original number of species ($SR=15,25,125$) and composition.
-We proceeded in two steps: (i) for each level of species richness, we created 1000 virtual communities by randomly picking desired number of species among the regional species pool; (ii) among those 1000 communities, we selected 20 communities for simulations with a large range of convex hull volume $CVH$ [@Cornwell2006a] but with a community weighted mean close to the regional species pool community weighted mean in order to observe change in functional diversity and not in functional composition.
+Disturbance intensity, as quantified by the percentage of basal area loss ($disturb_{intensity}$), was fixed at 0% (control), 25%, 50% or 75%.
+For each level of disturbance, we performed 20 simulations differing in the original number of species (species richness, $SR=15,25,125$) and composition.
+We proceeded in two steps: (i) for each level of species richness, we created 1000 virtual communities by randomly picking the desired number of species among the regional species pool; (ii) among those 1000 communities, we selected 20 communities for simulations with a large range of convex hull volume $CVH$ [@Cornwell2006a] but with a community mean ($CM$) close to the one of the regional species pool ($CM_{LMA} = 115 \pm 19$, $CM_{wsg} = 0.73 \pm 0.07$, $CM_{dmax} = 0.37 \pm 0.1$, $CM_{hmax} = 40 \pm 11$) in order to get change in functional diversity and not in average functional composition.
 This led to 240 simulations varying in community composition and disturbance intensity ($3~SR*20~CHV*4~ levels~of~disturbance$).
@@
 
+```{r, eval=F}
+doe %>% 
+  select(starts_with("CWM")) %>% 
+  reshape2::melt() %>% 
+  group_by(variable) %>% 
+  summarise_all(funs(mean, sd, min, max))
+```
+
 ## Analysis
@@
 
-Species and functional diversities were assessed for each simulation at the end of the pre-disturbance step with species richness ($SR$) and @villeger_dynamique_2008 indices ($FDiv$, the volume of the functional space occupied by the community, and $FEve$, the regularity of the distribution of abundance in this volume) using the `FD` R package [@laliberte_distance-based_2010-1].
+Species and functional diversities were assessed for each simulation at the end of the pre-disturbance step with species richness ($SR$) and functional diversity indices [$FDiv$, the volume of the functional space occupied by the community, and $FEve$, the regularity of the distribution of abundance in this volume; @villeger_dynamique_2008] using the `FD` R package [@laliberte_distance-based_2010-1].
 
@@
 
-We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).
+Simulated tropical forests were characterized in terms of changes in ecosystem properties regarding (i) forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and (ii) forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic day-time respiration ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic night-time respiration ($Rnight$ in $MgC.ha^{-1}$)).
 
-The recovery of the simulated forest ecosystem, i.e.
-the ecosystem capacity to recover from a disturbance event, at a time $t$, $R(t)$ regarding a given property $X$ was defined ad follows (Figure \@ref(fig:IrGraphFinal)):
+The recovery of the simulated forest at a time $t$, $R(t)$ regarding a given property $X$ was defined as follows (Figure \@ref(fig:IrGraphFinal)):
 
@@
 
-where $X_D(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$).
-Note that demographic stochasticity in the model induces a low variability among replicated simulations with the same set-up so that we did not include repetition of virtual communities simulations [@Marechaux2017].
+where $X_D(t)$ and $X_C(t)$ are the values fo the ecosystem property $X$ at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$) respectively.
+Note that demographic stochasticity in the model induces a very low variability among replicated simulations with the same set-up so that we we assumed a given simulation is representative of the community properties and dynamics given an initial species pool and disturbance intensity [@Marechaux2017].
 
 When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery.
-We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$.
-Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance, i.e.
+We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Figure \@ref(fig:IrGraphFinal)).
+Because the intensity of disturbance was set up by the user, our analysis did not include any ecosystem resistance, i.e.
 the ecosystem capacity to absorb a disturbance event.
-Consequently in this particular case, we can use the cummulated recovery as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):
+Consequently in this particular case, we used the inverse of cumulated euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):
 
 \begin{equation}
-  I_R(X) = \int _{t=0} ^{t=600} \frac{X_D(t)}{X_C(t)}.dt
+  I_R(X) = \frac 1 {\int _{t=disturb_{iter}} ^{t=disturb_{iter}+600} d_{eq}(t).dt}
   (\#eq:Resilience)
@@
 
-A high value of $I_R$ indicating a slow recovery.
-Ecosystem euclidean distance to equilibrium and resilience index $I_R$ was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).
-The influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem resilience index $I_R$  was finally investigated by linear regression ($I_R \sim SR + FEve + FDiv$).
+A high value of $I_R$ indicates a fast resilience.
+Ecosystem euclidean distance to equilibrium and resilience index $I_R$ were calculated in a multi-dimensional space for the two set of ecosystem property values: (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).
 
-```{r IrGraphFinal, fig.cap="Definition of recovery function and resilience indice shown on aboveground biomass dynamics of a control and disturbed ecosystem before and after the disturbance event."}
-knitr::include_graphics("images/final.png")
+<!-- $$R[X(t)] = \frac{X_D(t)}{X_C(t)}$$ -->
+<!-- $$d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$$ -->
+<!-- $$I_R(X) = \frac1{\int _{t=disturb_{iter}} ^{t=disturb_{iter}+600} d_{eq}(t).dt}$$ -->
+
+```{r IrGraphFinal, fig.cap="Definition of recovery function R and resilience indice $I_R$ shown on aboveground biomass dynamics for a simulation with $disturb_{intensity}=50%$."}
+knitr::include_graphics("images/figure3.png")
 ```
@@
 
-The net effect of biodiversity ($NE$) on ecosystem property $X$ is defined as the difference between the ecosystem property $X$ of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e.
-ecosystem property $X$ for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001].
+We tested the influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem resilience index $I_R$ for both forest structure and functioning using one-way ANOVA for species richness ($SR$) and linear regression for functional indices ($I_R \sim FEve + FDiv$).
+To meet the assumption of normality ecosystem resilience index $I_R$ was log-transformed.
+
+For each ecosystem property $X$, we further investigated the effect of biodiversity and its variation along post-disturbance trajectories.
+The net effect of biodiversity ($NE$) on ecosystem property $X$ is defined as the difference between the ecosystem property $X$ of the simulated mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e.
+that ecosystem property $X$ for the mixed-species community equals the relative abundance-weighted sum of that in monoculture [@Loreau2001].
+In order to do so, we made simulations for each of the 163 species in monoculture and for each level of disturbance leading to 652 simulations of monoculutre.
+
 We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:
@@
   \begin{array}{c}
-    NE = X_O - X_E = CE + SE \\
+    NE = CE + SE \\
     CE = N* \overline{\Delta RX} \overline{M}\\
@@
 
-where $N$ represents the total number of species, and $M$ the vector of monoculture performance. Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:
+where $N$ represents the total number of species, and $M$ the vector of $X$ for the 163 simulations in monoculture and $\Delta RX$:
 
@@
 $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance.
-Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives).
-Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016].
-But similarly, to recovery measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance.
-So in order to correctly assess selection and complementarity effect role in tropical forest resilience, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure treatments net effect recovery $R(NE_T)$:
+Similarly to recovery measurement, biodiversity net effect $NE$ can vary over time without disturbance.
+So in order to correctly assess selection and complementarity effect, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure the recovery of the net biodiversity effect $R(NE_T)$:
 
@@
 
-Recovery trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$.
-In order to do that, we simulated the 163 species individually representing 652 simulations of monoculture.
-
 # Results
 
-## Ecosystem functions
+## Biodiversity effect on forest resilience
 
-Species richness was not significantly influencing ecosystem resilience index $I_R$ for both forest structure and forest production (Table \@ref(tab:tgIr)).
-Increased functional diversity reduced  ecosystem resilience index $I_R$ and functional evenness was complementary reducing $I_R$ (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
-Finally, low species richness could result in variant $I_R$, but increased species richness resulted in increased functional diversity and consequently lower $I_R$.
-We found similar results for all disturbance levels (see [Appendix S4: Linear models]).
+Species richness ($SR$) did not significantly influence ecosystem resilience index $I_R$ for forest structure but significantly increased forest functioning resilience (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
+Both functional diversity ($FDiv$) and functional evenness ($FEve$) were positively related with $I_R$, i.e.
+they increased forest resilience (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
+Finally, low species richness resulted in variable $I_R$, but increased species richness resulted in increased functional diversity and consequently higher $I_R$ (one-way ANOVA $FDiv \sim SR$, $p-value < 0.001$).
+We found similar results for all disturbance levels, with increased significance for higher disturbance intensity;
 
-```{r gIr, echo=FALSE, fig.cap='Ecosystem resilience $I_R$ with taxonomic and functional diversity for different levels of disturbance.
-Resilience indice $I_R$ was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].
-Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.'}
+```{r gIr, echo=FALSE, fig.cap='Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance.
+Resilience indice $I_R$ was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance ($disturb_intensity$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness [FEve, @villeger_new_2008].
+Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.width=10, fig.height=15}
 Ir %>% 
@@
   geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
-  geom_point() +
-    facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
+  geom_point(size = 4) +
+  facet_wrap(~disturbance, scales = "free", labeller = 'label_both', 
+             nrow = 3) + 
   xlab('Functional diversity (FDiv)') +
@@
   scale_shape_discrete('Species\nrichness\n(SR)') +
-  scale_color_continuous('Functional\nevenness\n(FEve)')
+  scale_color_continuous('Functional\nevenness\n(FEve)') +
+  theme(text = element_text(size=20),
+        legend.key.size=unit(4,"line"))
 ```
@@
   ungroup() %>% 
+  mutate(SR = as.factor(SR)) %>% 
+  select(disturbance, structure, production,
+         SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
+                 variable.name = "Ir") %>% 
+  group_by(Ir, disturbance) %>%
+  do(fit = aov(lm(log(value) ~ SR, data = .))) %>% 
+  tidy(fit) %>% 
+  ungroup() %>% 
+  filter(term != "Residuals") %>% 
+  mutate(SR = ifelse(p.value < 0.001, "***",
+                            ifelse(p.value < 0.01, "**",
+                                   ifelse(p.value < 0.05, "*",
+                                          ifelse(p.value < 0.1, ".", "n.s."))))) %>% 
+  mutate(SR = cell_spec(SR,
+                         bold = ifelse(grepl("*", SR, fixed = T), T, F))) %>%
+  select(Ir, disturbance, SR) %>% 
+  left_join(Ir %>% 
+              left_join(doe, by = "sim") %>% 
+              ungroup() %>% 
               mutate(SR = as.numeric(SR)) %>% 
@@
                      SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
-  group_by(variable, disturbance) %>%
-  do(fit = lm(Ir ~ SR + FEve + FDiv, data = .)) %>% 
+              reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
+                             variable.name = "Ir") %>% 
+              group_by(Ir, disturbance) %>%
+              do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
               tidy(fit) %>% 
@@
                                                ifelse(p.value < 0.05, "*",
-                                     ifelse(p.value < 0.1, ".", "n.s."))))) %>% 
+                                                      ifelse(p.value < 0.1, ".", 
+                                                             "n.s."))))) %>% 
               mutate(Coefficient = round(estimate,2)) %>% 
-  mutate(Recovery = variable) %>% 
-  rename(Variable = term, `Disturbance intensity` = disturbance) %>%
-  select(Recovery, `Disturbance intensity`, Variable, Coefficient, `P-value`) %>% 
               mutate(value = paste(Coefficient, `P-value`)) %>%
               mutate(value = cell_spec(value,
-                           bold = ifelse(grepl("*", value, fixed = T), T, F))) %>%
-  reshape2::dcast(Recovery + `Disturbance intensity` ~ Variable, value.var = "value") %>% 
-  kable(caption = "Result of linear regression testing the effect of species richness and functional diversity indices on the recovery index by recovery type and disturbance intensity.", escape = F) %>%
+                                       bold = ifelse(grepl("*", value,
+                                                           fixed = T), T, F))) %>%
+              reshape2::dcast(Ir + disturbance ~ term, value.var = "value")) %>% 
+  left_join(Ir %>% 
+              left_join(doe, by = "sim") %>% 
+              ungroup() %>% 
+              mutate(SR = as.factor(SR)) %>% 
+              select(disturbance, structure, production,
+                     SR, FEve, FDiv) %>% 
+              reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
+                             variable.name = "Ir") %>% 
+              group_by(Ir, disturbance) %>%
+              do(fit = lm(log(value) ~ SR + FEve + FDiv, data = .)) %>% 
+              glance(fit) %>% 
+              mutate(r.squared = round(r.squared, 2)) %>% 
+              select(Ir, disturbance, r.squared)) %>% 
+  kable(caption = "Effects of taxonomic and functional diversity on forest resilience.
+Results of the linear regression testing the effect of species richness (SR) and functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functionning and for each level of disturbance intensity ($disturb_{intensity}$).
+To meet the assumption of normality $I_R$ was log-transformed.
+Test of SR effect was done by one-way ANOVA with SR as a fix factor.
+Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independant variables.
+Values of FDiv and FEve are coefficients from linear regressions.
+$R^2$ is computed for the full model.
+Significant $p-values$ are shown in bold.
+ns, non‐significant; .
+p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.", escape = F,
+        col.names = c("$I_R$", "Disturbance intensity", "SR", 
+                      "FDiv", "FEve", "$R^2$")) %>% 
   kable_styling(full_width = F) %>% 
@@
 
-## Biodiversity effect
+```{r anova, eval=F}
+# summary(aov(lm(FDiv ~ SR, left_join(Ir, doe, by = "sim"))))[[1]]$`Pr(>F)`[1]
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  mutate(SR = as.factor(SR)) %>% 
+  select(disturbance, structure, production,
+         SR, FEve, FDiv) %>% 
+  rename(Ir.structure = structure, Ir.production = production) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR")) %>% 
+  ggplot(aes(SR, value, fill = SR)) +
+  geom_boxplot() +
+  facet_grid(variable ~ disturbance, scales = "free") +
+  ggpubr::stat_compare_means(method = "anova",
+                             aes(label = paste0("Anova ", ..p.signif..))) +
+  ggpubr::stat_compare_means(comparisons = list(c("5", "25"), c("25", "125"), 
+                                                c("5", "125")),
+                             aes(label = ..p.signif..)) + 
+  theme(axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank())
+  # ggsave(filename = "./images/anova.png", width = 30, height = 30, units = "cm")
+```
+
+```{r residualsHist, eval=F}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  mutate(SR = as.factor(SR)) %>% 
+  select(disturbance, structure, production,
+         SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
+                 variable.name = "Ir") %>% 
+  group_by(Ir, disturbance) %>%
+  do(residuals = residuals(lm(log(value) ~ SR + FEve + FDiv, data = .))) %>% 
+  tidy(residuals) %>% 
+  ggplot(aes(x)) +
+  geom_histogram(bin = 20) +
+  geom_vline(xintercept = 0, col = "red") +
+  facet_grid(~ Ir + disturbance, scales = "free") +
+  xlab("Residuals")
+```
+
+```{r residualsTest, eval=F}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  mutate(SR = as.factor(SR)) %>% 
+  select(disturbance, structure, production,
+         SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
+                 variable.name = "Ir") %>% 
+  group_by(Ir, disturbance) %>%
+  do(residuals = residuals(lm(log(value) ~ SR + FEve + FDiv, data = .))) %>% 
+  tidy(residuals) %>% 
+  group_by(Ir, disturbance) %>% 
+  do(normality = shapiro.test(.$x)) %>%
+  tidy(normality) %>% 
+  transmute(p.value = ifelse(p.value < 0.001, "***",
+                             ifelse(p.value < 0.01, "**",
+                                    ifelse(p.value < 0.05, "*",
+                                           ifelse(p.value < 0.1, ".", 
+                                                  "n.s.")))))
+```
+
+## Post-disturbance trajectories of biodiversity effect
 
@@
 
-The recovery from net effect of biodiversity on ecosystem properties was partitioned between selection and complementarity effect after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, see Table \@ref(tab:tNE)) for every ecosystem global variable, e.g.
-aboveground biomass (see Figure \@ref(fig:gBE)).
-We found that complementarity effect was dominant in the biodiversity net effect in the first decades.
-But after few decades the complementarity effect diminished toward an extremely low value.
-On the contrary, selection effect was reduced in the initial times after disturbance.
-It increased during the first decades and was then greater.
-The time lag for which complementarity effect overcame the selection effect increased with disturbance intensity.
-Finally, 600 years after disturbance, the biodiversity net effect on aboveground biomass was still not at equilibrium for a disturbance intensity greater than 25% of basal area.
-We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning).
-Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionally to disturbance), and its net effect was maintained by complementarity effect.
+The net effect of biodiversity on each ecosystem property was quantified and partitioned between selection and complementarity effects after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, see Table \@ref(tab:tNE) and Figure \@ref(fig:gBE)).
+We found that complementarity effect sharply increased immediately after disturbance and was then the dominant effect of biodiversity during the first decades (Figure \@ref(fig:gBE)).
+Thereafter, the complementarity effect diminished toward a low value close to zeo.
+On the contrary, selection effect sharply decreased right after disturbance.
+It then progressively increased to finally represent the majority of the net biodiveristy effect about a century after disturbance.
+The dominance of complementarity effect was more important and lasted longer after disturbance with increasing disturbance intensity.
+Finally, 600 years after disturbance, the biodiversity net effect on aboveground biomass was still lower than the control for higher disturbance intensity (Figure \@ref(fig:gBE)).
+We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$).
+Finally, forest growth primary productivity ($GPP$) recovered only in few years with a dominant complementarity effect.
 
@@
   summarise(mean = mean(value, na.rm = T), 
-            'standard deviation' = sd(value, na.rm = T)) %>% 
+            sd = sd(value, na.rm = T)) %>% 
   filter(variable != 'n30') %>% 
@@
              '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$'))) %>% 
-  kable(caption = 'Mean value and standard deviation of biodiversity net effect accross mature simulations before the disturbance event for different ecosystem variable.', digits = 3, format = 'pandoc')
+  mutate(NE = paste0("$", round(mean,2), "\\pm", round(sd,2), "$")) %>% 
+  select(name, NE, unit) %>% 
+  kable(caption = 'Mean and standard deviation of the biodiversity net effect across mature simulations before the disturbance event (i.e.
+after 600 years of regeneration) for different ecosystem variables.',
+        format = 'pandoc', col.names = c("Ecosystem property", "Net effect", "Unit")) %>% 
+  kable_styling(full_width = F)
 ```
 
-```{r gBE, echo=FALSE, fig.cap='recovery of complementarity and selection effects.
-Complementarity effect (CE) and selection effect (SE) where normalized by control net effect (NEc), thus measuring their recovery over time.
-Solid lines represents the mean value whereas areas represents the interval of values among simulations.'}
+```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects.
+The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0%$).
+Solid lines represents the mean value whereas areas represents the range of values across the 20 simulations with a species richness ($SR$) of 125.
+Time axis has been square-root transformed to highlight first decades dynamic.'}
 BEresilience %>% 
@@
   xlab('Time (years)') +
-  ylab('Recovery of aboveground biomass net effect')
+  ylab('Post-disturbance trajectories of aboveground biomass net effect') +
+  scale_x_sqrt() +
+  theme(text = element_text(size=20),
+        legend.key.size=unit(2,"line"))
 ```

### 2018-12-11 draft.Rmd
Draft 0.1.9009 Final 1 Discussion with BEF review

@@
 output:
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
   github_document: default
-  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: false
+    toc: true
+    toc_float: yes
 linestretch: 1.5
@@
 Selected individuals are then removed without triggering a treefall to avoid any side effect.
-The operation is repeated until the cummulative basal area of removed trees reaches targeted threshold ($disturb_{intensity}$).
+The operation is repeated until the cumulative basal area of removed trees reaches targeted threshold ($disturb_{intensity}$).
 
@@
 
-where $X_D(t)$ and $X_C(t)$ are the values fo the ecosystem property $X$ at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$) respectively.
+where $X_D(t)$ and $X_C(t)$ are the values for the ecosystem property $X$ at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$) respectively.
 Note that demographic stochasticity in the model induces a very low variability among replicated simulations with the same set-up so that we we assumed a given simulation is representative of the community properties and dynamics given an initial species pool and disturbance intensity [@Marechaux2017].
@@
 
-We tested the influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem resilience index $I_R$ for both forest structure and functioning using one-way ANOVA for species richness ($SR$) and linear regression for functional indices ($I_R \sim FEve + FDiv$).
+We tested the influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem resilience index $I_R$ for both forest structure and functioning using linear regression ($I_R \sim FEve + FDiv$).
+We tested variable correlation by Pearson's correlation coefficient for continuous variables and one-way ANOVA for species richness $SR$.
+In case of significant correlation, we selected the variable resulting in the model with the higher Aikaike information criterion in order to have independent variables.
 To meet the assumption of normality ecosystem resilience index $I_R$ was log-transformed.
@@
 Similarly to recovery measurement, biodiversity net effect $NE$ can vary over time without disturbance.
-So in order to correctly assess selection and complementarity effect, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure the recovery of the net biodiversity effect $R(NE_T)$:
+So in order to correctly assess selection and complementarity effect, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure the recovery of the net biodiversity effect $R(NE)$:
 
 \begin{equation}
-  R(NE_T) = \frac{NE_T}{NE_C} = \frac{SE_T}{NE_C} + \frac{CE_T}{NE_C}
+  R(NE) = \frac{NE}{NE_C} = \frac{SE}{NE_C} + \frac{CE}{NE_C}
   (\#eq:RNE)
@@
 
-Species richness ($SR$) did not significantly influence ecosystem resilience index $I_R$ for forest structure but significantly increased forest functioning resilience (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
+```{r, eval=F}
+doe %>% 
+  select(SR, FDiv, FEve) %>% 
+  reshape2::melt(id.vars = "SR") %>% 
+  mutate(variable = paste0("$", variable, "$")) %>% 
+  group_by(variable, SR) %>% 
+  summarise(mean =  paste0("$", round(mean(value),2), 
+                           "\\pm", round(sd(value),2), "$")) %>% 
+  arrange(variable, desc(SR)) %>% 
+  kable(caption = 'Mean and standard deviation of the predisturbance functional diversity indices for each level of species richness.',
+        format = 'pandoc', col.names = c("Variable", "Species richness ($SR$)", 
+                                         "Mean")) %>% 
+  kable_styling(full_width = F)
+```
+
+Functional diversity ($FDiv$) and evenness ($FEve$) varied respectively from $0.76\pm0.18$ and $0.46\pm0.24$ to $0.96\pm0.02$ and $0.64\pm0.02$ with increasing species richness ($SR$) from 5 to 125 species.
+Species richness ($SR$) was significantly correlated to functional indices (one-way ANOVA $p-value < 0.001$ for both $FDiv$ and $FEve$) and model $AIC$ was higher without $SR$, consequently we used only functional indices for linear regressions.
 Both functional diversity ($FDiv$) and functional evenness ($FEve$) were positively related with $I_R$, i.e.
 they increased forest resilience (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
-Finally, low species richness resulted in variable $I_R$, but increased species richness resulted in increased functional diversity and consequently higher $I_R$ (one-way ANOVA $FDiv \sim SR$, $p-value < 0.001$).
-We found similar results for all disturbance levels, with increased significance for higher disturbance intensity;
+We found similar results for all disturbance levels, with increased significance for higher disturbance intensity.
 
@@
   ungroup() %>% 
-  mutate(SR = as.factor(SR)) %>% 
-  select(disturbance, structure, production,
-         SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
-                 variable.name = "Ir") %>% 
-  group_by(Ir, disturbance) %>%
-  do(fit = aov(lm(log(value) ~ SR, data = .))) %>% 
-  tidy(fit) %>% 
-  ungroup() %>% 
-  filter(term != "Residuals") %>% 
-  mutate(SR = ifelse(p.value < 0.001, "***",
-                            ifelse(p.value < 0.01, "**",
-                                   ifelse(p.value < 0.05, "*",
-                                          ifelse(p.value < 0.1, ".", "n.s."))))) %>% 
-  mutate(SR = cell_spec(SR,
-                         bold = ifelse(grepl("*", SR, fixed = T), T, F))) %>%
-  select(Ir, disturbance, SR) %>% 
-  left_join(Ir %>% 
-              left_join(doe, by = "sim") %>% 
-              ungroup() %>% 
   mutate(SR = as.numeric(SR)) %>% 
@@
   do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
+  # do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
   tidy(fit) %>% 
@@
   filter(term != "(Intercept)") %>% 
+  filter(!grepl("SR", term)) %>% 
   mutate(`P-value` = ifelse(p.value < 0.001, "***",
@@
                                                fixed = T), T, F))) %>%
-              reshape2::dcast(Ir + disturbance ~ term, value.var = "value")) %>% 
+  reshape2::dcast(Ir + disturbance ~ term, value.var = "value") %>% 
   left_join(Ir %>% 
@@
               group_by(Ir, disturbance) %>%
-              do(fit = lm(log(value) ~ SR + FEve + FDiv, data = .)) %>% 
+              do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
               glance(fit) %>% 
@@
               select(Ir, disturbance, r.squared)) %>% 
-  kable(caption = "Effects of taxonomic and functional diversity on forest resilience.
-Results of the linear regression testing the effect of species richness (SR) and functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functionning and for each level of disturbance intensity ($disturb_{intensity}$).
+  mutate(Ir = ifelse(Ir == "production", "functioning", "structure")) %>% 
+  kable(caption = "Effects of functional diversity on forest resilience.
+Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functionning and for each level of disturbance intensity ($disturb_{intensity}$).
 To meet the assumption of normality $I_R$ was log-transformed.
-Test of SR effect was done by one-way ANOVA with SR as a fix factor.
 Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independant variables.
 Values of FDiv and FEve are coefficients from linear regressions.
-$R^2$ is computed for the full model.
 Significant $p-values$ are shown in bold.
 ns, non‐significant; .
-p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.", escape = F,
-        col.names = c("$I_R$", "Disturbance intensity", "SR", 
+p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.",
+        escape = F, col.names = c("$I_R$", "Disturbance intensity",
                                   "FDiv", "FEve", "$R^2$")) %>% 
@@
 ```{r anova, eval=F}
-# summary(aov(lm(FDiv ~ SR, left_join(Ir, doe, by = "sim"))))[[1]]$`Pr(>F)`[1]
-Ir %>% 
-  left_join(doe, by = "sim") %>% 
+tHSD <- doe %>% 
   ungroup() %>% 
   mutate(SR = as.factor(SR)) %>% 
-  select(disturbance, structure, production,
+  select(disturbance,
          SR, FEve, FDiv) %>% 
-  rename(Ir.structure = structure, Ir.production = production) %>% 
   reshape2::melt(id.vars = c("disturbance", "SR")) %>% 
-  ggplot(aes(SR, value, fill = SR)) +
+  group_by(variable) %>% 
+  do(anova = aov(value ~ SR, data = .)) %>% 
+  mutate(tHSD = TukeyHSD(anova, ordered = F, conf.level = 0.95)) %$%
+  lapply(tHSD, function(x) multcompView::multcompLetters(x[,4])) %>% 
+  lapply(function(x) data.frame(SR = names(x$Letters), group = x$Letters)) %>% 
+  bind_rows() %>% 
+  mutate(axis = c(rep("FEve", 3), rep("FDiv", 3)))
+doe %>% 
+  ungroup() %>% 
+  mutate(SR = as.factor(SR)) %>% 
+  select(disturbance,
+         SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR")) %>% 
+  left_join(tHSD) %>% 
+  ggplot(aes(SR, value, fill = SR, label = group)) +
   geom_boxplot() +
-  facet_grid(variable ~ disturbance, scales = "free") +
-  ggpubr::stat_compare_means(method = "anova",
-                             aes(label = paste0("Anova ", ..p.signif..))) +
-  ggpubr::stat_compare_means(comparisons = list(c("5", "25"), c("25", "125"), 
-                                                c("5", "125")),
-                             aes(label = ..p.signif..)) + 
+  geom_text(aes(y = 1.1)) +
+  ggpubr::stat_compare_means(method = "anova", aes(label = paste("Anova", ..p.signif..))) +
+  facet_wrap(~ variable, scales = "free") +
   theme(axis.title.x=element_blank(),
@@
         axis.ticks.x=element_blank())
-  # ggsave(filename = "./images/anova.png", width = 30, height = 30, units = "cm")
 ```
@@
   group_by(Ir, disturbance) %>%
-  do(residuals = residuals(lm(log(value) ~ SR + FEve + FDiv, data = .))) %>% 
+  do(residuals = residuals(lm(log(value) ~ FEve + FDiv, data = .))) %>% 
   tidy(residuals) %>% 
@@
 We found that complementarity effect sharply increased immediately after disturbance and was then the dominant effect of biodiversity during the first decades (Figure \@ref(fig:gBE)).
-Thereafter, the complementarity effect diminished toward a low value close to zeo.
+Thereafter, the complementarity effect diminished toward a low value close to zero.
 On the contrary, selection effect sharply decreased right after disturbance.
-It then progressively increased to finally represent the majority of the net biodiveristy effect about a century after disturbance.
+It then progressively increased to finally represent the majority of the net biodiversity effect about a century after disturbance.
 The dominance of complementarity effect was more important and lasted longer after disturbance with increasing disturbance intensity.
@@
 
-In the limit of the model, we were able to show that functional diversity improved tropical forest resilience from disturbance.
-More particularly, functional diversity and evenness are key components of diversity in forest recovery trajectory after disturbance.
-Moreover, we found that complementarity between species was insuring forest recovery in the early times before more productive species dominate the forest and insure late recovery.
-We thus advocate the view that both diversity theory (with niche complementarity and facilitation) and mass ratio theory (represented by selection) are implied in ecosystem resilience but with relative importance depending on the spatial grain and time since the disturbance event.
+We investigated the effect of biodiversity on tropical forest resilience to disturbance and the variation of biodiversity effect on ecosystem functioning along trajectory of recovery from disturbance by means of a virtual experiment.
+We found that functional diversity improved tropical forest resilience from disturbance.
+Additionally, the mechanisms that underlie the biodiversity effect on forest functioning changed in dominance as trajectories of recovery unfolded.
+Complememtarity between species was the dominant effect of biodiversity in the early times after disturbance.
+Complementarity was then progressively overtaken by a selection effect as more competitive species dominate the forest community.
+This shift was all the more pronounced that disturbance increased in intensity.
+Our results suggest that ecosystem state and past history can strongly influence the mechanisms by which biodiversity impacts ecosystem processes.
+This can help reconciling previous contrasted results obtained when dealing with snapshots of ecosystem state.
 
-## Functional diversity improve tropical forest recovery
+## Functional diversity improves tropical forest recovery
 
-Our results validated the hypothesis of a significant positive relationship between forest recovery and functional diversity and evenness.
-More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008] seems a major aspect of recovery if it's strengthened by a high functional evenness [regularity of the distribution of abundance in this volume, @villeger_new_2008].
+Our results validated the hypothesis of a significant positive relationship between forest recovery and functional diversity.
+More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008] appeared as a major aspect of resilience if it is strengthened by a high functional evenness [regularity of the distribution of abundance in this volume, @villeger_new_2008].
 Indeed, high functional diversity combined to high evenness avoid hegemonic effect of a few dominant species.
-Our results thus confirms both the review of @Daz2001a, advocating for an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity and resilience of terrestrial ecosystems.
-Finally, species richness is not directly increasing resilience, thus supporting the idea that species diversity is often an inadequate surrogate of functional diversity [@Daz2001a; but see @Poorter2015], which might explain the absence of diversity effect on resilience observed by @VanderSande2018.
-Still, increased species richness will increase chance for high functional diversity through the sampling effect [@Loreau1998; @Morin2014a].
-An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.
+Functional diversity positive effect on forest resilience can be due to (i) temporal complementarity with the diversity of species response to disturbances and response asynchronicity, resulting in ecological sorting [@Loreau2013; @Sakschewski2016; @Morin2014a], and (ii) functional complementarity reducing competition strength [@Loreau2013] advocating both for the biodiversity insurance hypothesis [@Yachi1999].
+Our results thus confirms both the review of @Daz2001a, advocating against an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity of forest ecosystems.
+In addition, increased species richness will de facto increase chance for high functional diversity through the sampling effect [@Loreau1998; @Morin2014a].
+A higher sampling of regional species pool will thus allow a greater chance to pick more functionally diverse species.
 
-## Complementarity and selection insure forest recovery
+## Mechanisms underlying the biodiveristy effect on forest ecosystem change along trajectories of recovery from disturbances
 
-We found that complementary effect was insuring forest recovery in the beginning of forest successions.
-We interpreted this result as the consequence of facilitation processes or lower competition due to different niche.
-As the only resource currently simulated by TROLL is the light, the main facilitation would be light shading of post-pioneer species by pioneer species in disturbance gaps, confirming the study of @Morin2011 who also highlighted the importance of facilitation through light shading in forest resilience.
-But the complementarity effect quickly reduced through time, to let the selection effect insure forest recovery due to an re-enforced dominance of more productive species.
-The diminishing of complementarity effect is attributed to the selection of the most competitive species though time.
+We found that complementarity among the species was the prevailing effect of biodiversity on forest functioning during the beginning of forest post-disturbance trajectory.
+We interpreted this result as the consequence of niche partitioning.
+By creating additional gaps in the simulated mature forest plots, the simulated disturbances resulted in an increased heterogeneity in the availability in light, which is the main limited resource in dense tropical forest.
+Species are known to vary widely in their growth response to gap creation depending on their functional strategy [@Herault2010; @Laurans2012; @Ruger2009].
+Pioneer species with productive leaf tissue and light wood typically quickly close gaps, hence creating a shaded environment where late successional species with more conservative leaf tissue are more competitive.
+By promoting a higher diversity of shade tolerance and growth response to light, species richness hence foster forest response to small-scale perturbations [@Morin2011].
+This complementarity effect quickly reduced through time letting the selection effect progressively gain in dominance.
+The diminishing of complementarity effect is attributed to the increased abundance of the most competitive species in more stable and less heterogeneous environment.
 
-In addition, the study scale matters as shown by @Chisholm2013 and @Sullivan2017.
-In this work, we look at processes to a 16-ha scale, being limited by the computational effort and, our simulation did not include topography resulting in the absence of micro environmental variation affecting both water and nutrient availability.
-Because several studies suggest that, in mature ecosystems, complementarity will be stronger at smaller scale [@Chisholm2013; @Sullivan2017], this could explain the low value of complementarity after the initial stages of forest recovery.
-Still, the absence of topography and environmental heterogeneity in our simulations could explain why we observed an effect of complementarity to a 16-ha scale, strengthening @Sullivan2017 interpretation.
+## Perspectives in forest ecosystem modelling
 
-Our results follows several studies advocating for multiple mechanisms through which functional diversity affect ecosystem functions.
-Complementarity effect encompass niche complementarity theory (or diversity theory), which has been shown to improve forest biomass [@Poorter2017; @Huang2018] and coarse wood productivity [@Chiang2016] in tropical and subtropical forests.
-Similarly, complementarity has been shown of primary importance in grassland functioning [@Cardinale2007; @Fargione2007; @Isbell2011].
-On the other hand, selection effect can be related to the mass ratio hypothesis, which proposes that ecosystem functions are regulated by dominant species of the community [@Grime1998].
-And mass ratio theory has been shown to primarily influence ecosystem functions with biomass accumulation (aboveground and shoots), productivity (coarse wood or aboveground), water use, and light interception for both grassland and forests [@Finegan2015; @Lohbeck2016; @Prado-Junior2016; @Chiang2016; @Mokany2008].
-Consequently, concerning mechanisms responsible for the positive effect of functional diversity on tropical forest functions after disturbance, we also advocate for the interplay of both niche complementarity and mass ratio theories, depending on time since the disturbance event and the spatial grain.
+In this study we took advantage of a forest simulator able to jointly simulate biodiversity and ecosystem functioning to explore the effect of biodiversity on forest resilience on a long temporal scale, that remains out of reach of most experimental approaches.
+By virtually manipulating the number and identity of species across a large number of simulations, we were able to evidence the role of diversity in forest resilience and the changing driver of biodiversity effect on forest functioning along trajectory of recovery.
 
-Finally, our findings confirm results from experimental studies.
-@Williams2017 also found experimentally an over yielding due to complementarity effect based on light and spatial complementarity with tree crown of young trees.
-In addition, @Tobner2016 studied 4-years stand in low diverse forests of Canada.
-They also found that selection effect was greater than complementarity in most of the cases, promoting the idea that mass ratio theory is the primary mechanisms involved in functional diversity effect on ecosystem functions [@Mokany2008].
-Besides @Huang2018 highlighted increased complementarity over null selection effect resulting in higher productivity due to tree growth in an 8-years forest experiment of China.
+As any modelling approaches, it also comes with simplifications and assumptions.
+In our simulations, light was the unique resource whose variability was explicitly accounted for [@Marechaux2017].
+Our results should thus be restricted to situations where aboveground competition for light overcomes the effect of belowground competition for water and nutrients, a reasonable assumption in humid tropical forests, even in early successional stages [@VanBreugel2012].
+However this assumption may be less suitable under the predicted increase in drought intensity and frequency.
+We did not include the effect of topography, known to result in higher micro environmental heterogeneity.
+Overall, we expect a stronger role of complementarity effect through hydrological and edaphic niche partitioning as water become more limited and in more heterogeneous landscapes [@Baltzer2005; @Jucker2018; @Silvertown2004a].
+Additionally, we simulated a generic form of disturbance, and our findings might be modulated depending on the disturbance specificity.
+For instance, selective logging will induce both species targeting which could lower the selection effect if dominant species are chosen and cyclic perturbation, increasing the prevalence of early recovery stages where complementarity effect plays the most.
+Consequently, we could hypothesize that species complementarity and functional diversity will be of primary importance regarding selective logging compared to less targeted disturbances with longer cycles such as drought, fires and hurricanes.
 
-## Conclusion
+## Unifying biodiversity-ecosystem functioning studies
 
-We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance.
-We found that diversity improved forest resilience, similarly to productivity [@Liang2016; @Poorter2015].
-Additionally, we found that complementarity between species was insuring forest resilience in forest succession start with facilitation, before more productive species dominate the forest and insure the final forest resilience.
+Our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depends on the ecosystem state and its history of perturbation, with dominant complementarity effect in recently disturbed mature ecosystem whereas selection effect takes over complementarity in more stable systems with older disturbances.
+In agreement with this, studies in recently disturbed mature forest did found a stronger complementarity effect [@Huang2018; @Poorter2017], whereas a dominant selection effect has been found both in undisturbed forests our new plantations [@Bauters2015; @Finegan2015; @Marechaux2017; @Tobner2016] but see [@DeAvila2018; @Lohbeck2016].
 
-Our results advocate for an important interplay of diversity and dominant species in forest resilience depending on the recovery stage, similarly to findings of @Sakschewski2016.
-Moreover, we studied a generic form of disturbance, and our findings might be modulated depending on the disturbance specificity.
-For instance, selective logging will induce both species targeting diminishing selection effect if dominant species are chosen and cyclic perturbation, increasing the role of early recovery where complementarity effect play the most.
-Consequently, we could hypothesize that biodiversity will be more of primary importance regarding selective logging compared to less targeted disturbances with longer cycles such as drought, fires and hurricanes.
+Moreover, we here focus on the effect of biodiversity along trajectory of post-disturbance recovery, and our simulations and analyses were all made at a 16-ha scale.
+However, in addition to the influence of disturbance intensity and temporal trajectory we evidenced here, @Chisholm2013 and @Sullivan2017 found that biodiversity effect vary across spatial scale, with typically a stronger effect at smaller scale (e.g.
+0.04-ha).
+Finally, ecological constraints may differ between ecosystems, for instance pioneer species (e.g.
+hard-wood pioneer species in dry forest [@Prado-Junior2016]), light importance (e.g.
+nitrogen importance in grasslands [@Fargione2007; @Prado-Junior2016]), and maturity concept (e.g.
+temperate versus tropical forests) can differ from tropical wet forest to another system, that may at the origin of biodiversity effects differences.
+
+Together, this paves the way to a unifying framework to reconcile the apparent contradictions between studies advocating for the sole effect of complementarity or selection effects, and call for an accurate description of the ecosystem spatial-temporal and ecological context in future studies to improve global understanding of biodiversity and ecosystem functioning relations, and thus further open perspectives to a better prediction of ecosystem resilience in a context of increasing disturbances and biodiversity crisis.
+
+```{r LiteratureBiodiversityEffect, eval=F}
+literature <- data.frame(
+  Study = c("@Bauters2015", "@Finegan2015", "@Prado-Junior2016",
+            "@Lohbeck2016", "@DeAvila2018", "@Marechaux2017", "@Tobner2016",
+            "@Morin2014a", "@Cardinale2007", "@VanderSande2017a",
+            "@Poorter2017", "@Huang2018", "@Niklaus2017a",
+            "@Fargione2007"),
+  StudyType = c("Observation", "Observation", "Observation",
+                "Observation", "Experiment", "Simulation", "Experiment",
+                "Simulation", "Review", "Review",
+                "Observation", "Experiment", "Experiment", 
+                "Experiment"),
+  Type = c("Forest", "Forest", "Forest", 
+           "Forest", "Forest", "Forest", "Forest",
+           "Forest", "Grasslands-Tundra-Estuaries", "Forest",
+           "Forest", "Forest", "Forest", 
+           "Grassland"),
+  Biome = c("Tropical", "Tropical", "Tropical",
+            "Tropical", "Tropical", "Tropical", "Temperate",
+            "Temperate", "Temperate", "Tropical",
+            "Tropical", "SubTropical", "SubToprical", 
+            ""),
+  SubBiome = c("Wet", "Wet-Dry", "Dry",
+               "Wet?", "Wet", "Wet", "",
+               "", "", "",
+               "Wet", "", "", 
+               ""),
+  Effect = c("SE", "SE", "SE",
+             "SE", "SE", "SE", "SE",
+             "CE", "CE", "CE",
+             "CE", "CE", "CE", 
+             "SE-CE"),
+  Disturbed = c("Y", "N?", "Y",
+                "Y", "Y", "N", "Y",
+                "N", "?", "?",
+                "50%Y", "Y", "", 
+                "Y"),
+  Age = c("77yo", "", "12yo",
+          "1-29yo", "21yo", "600yo", "8yo",
+          "200yo", "", "",
+          "14-32yo", "8yo", "18yo", 
+          "10yo"),
+  DisturbationType = c("Plantation", "", "Grazing/Logging",
+                       "Secondary forest from fallow", "Logging-Thinning-Fire", "",
+                       "Plantation",
+                       "", "", "",
+                       "Logging", "Plantation", "Plantation", 
+                       "Plantation")
+)
+literature %>% 
+  select(Effect, Disturbed, Age, DisturbationType, Study, 
+         StudyType, Type, Biome, SubBiome) %>% 
+  arrange(Effect,
+          factor(Disturbed, levels = c("Y", "50%Y", "N", "N?", "?")),
+          factor(Age, levels = c("8yo", "10yo", "12yo", "18yo", "1-29yo",
+                                 "21yo", "14-32yo", "77yo", "200yo", "600yo", " "))) %>% 
+  kable() %>% 
+  kable_styling(full_width = F)
+```
 

### 2018-12-13 A01-mortality.Rmd
Draft 0.1.9010 Co-authors sharing

--- 2018-11-06 SX-mortality.Rmd
@@
 ---
-title: "Nino mortality issue"
+title: "A01 - Nino mortality issue"
 date: '`r Sys.Date()`'

### 2018-12-13 A02-Resilience.Rmd
Draft 0.1.9010 Co-authors sharing

--- 2018-11-30 S2-Resilience.Rmd
@@
 ---
-title: "Supplementary material S2: Resilience indice"
+title: "A02 - Resilience index"
 date: '`r Sys.Date()`'
@@
 
-This document introduces and explains the recovery function $R(X)$ and the resilience indice $I_R$ definitions (figure \@ref(fig:IrGraphFinal)).
+This document introduces and explains the recovery function $R(X)$ and the resilience index $I_R$ definitions (figure \@ref(fig:IrGraphFinal)).
 Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010].
@@
 ```{r IrGraphFinal, fig.cap="Definition of recovery function and resilience indice shown on aboveground biomass dynamics of a control and disturbed ecosystem before and after the disturbance event."}
-knitr::include_graphics("images/final.png")
+knitr::include_graphics("images/figure3.png")
 ```
@@
 the ecosystem capacity to absorb a disturbance event,.
-Consequently **in this particular case**, we can use the cummulated recovery as an indice of ecosystem resilience, further defined as $I_R$:
+Consequently **in this particular case**, we can use the cummulated recovery as an index of ecosystem resilience, further defined as $I_R$:
 
@@
 g_agb <- ecosystem %>% 
+  ungroup() %>% 
   left_join(., doe, by = 'sim') %>%
@@
 g_resilience <- resilience %>% 
+  ungroup() %>%
   left_join(., doe, by = 'sim') %>%
@@
   left_join(., doe, by = 'sim') %>%
+  ungroup() %>%
   filter(SR == 125) %>% 

### 2018-12-13 A03-RNE.Rmd
Draft 0.1.9010 Co-authors sharing

--- 2018-11-06 S3-RNE.Rmd
@@
 ---
-title: "Supplementary material S3: Biodiversity net effect and resilience"
+title: "A03 - Biodiversity net effect and resilience"
 date: '`r Sys.Date()`'

### 2018-12-13 A04-LM.Rmd
Draft 0.1.9010 Co-authors sharing

--- 2018-11-06 S4-LM.Rmd
@@
 ---
-title: "Supplementary material S4: Linear models"
+title: "A04: Linear models"
 date: '`r Sys.Date()`'

### 2018-12-13 A05-Biodiversity.Rmd
Draft 0.1.9010 Co-authors sharing

--- 2018-11-06 S5-Biodiversity.Rmd
@@
 ---
-title: "Supplementary material S5: Biodiversity effect results"
+title: "A05: Biodiversity effect results"
 date: '`r Sys.Date()`'
@@
 
-This document explores graphically the result of the biodiversity net effect partitionning.
+This document explores graphically the result of the biodiversity net effect partitioning.
 More especially we want to look if the representation from figure \@ref(fig:gBE) is correct, with the dynamic biodiversity net effect normalized by the control across ecosystems.
@@
 The net effect of biodiversity on ecosystem recovery was partitioned between selection and complementarity effect.
-We are thus able for one community to look at aboveground biomass dynamic after and before disturbance in control and disturbed ecosystems and correspondig net biodiversity effect, selection effect and complementarity effect (figure \@ref(fig:gBEsingle)).
+We are thus able for one community to look at aboveground biomass dynamic after and before disturbance in control and disturbed ecosystems and corresponding net biodiversity effect, selection effect and complementarity effect (figure \@ref(fig:gBEsingle)).
 Similarly to the aboveground biomass of the ecosystem, the net biodiversity effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increases following the disturbance.
@@
 
-If we aggregate the result of all community with 125 species, we still observe the same pattern for all simulations compard to the single previous one (figure \@ref(fig:gBEunorm)).
+If we aggregate the result of all community with 125 species, we still observe the same pattern for all simulations compared to the single previous one (figure \@ref(fig:gBEunorm)).
 But due to an increase of biodiversity net effect through recovery time, complementarity importance in early recovery is less visible over a long period.
@@
 Thanks to previous graphical exploration, we can affirm that normalizing by the control is not distorting the result.
-And this last figure brings the clearer message regarding complementarity and selection effect role in the resilience of the biodiveristy net effect.
+And this last figure brings the clearer message regarding complementarity and selection effect role in the resilience of the biodiversity net effect.
 

### 2018-12-13 S1-LL.Rmd
Draft 0.1.9010 Co-authors sharing

@@
 output:
+  bookdown::word_document2: default
   bookdown::html_document2:
-    number_sections: false
-    toc: true
+    number_sections: no
+    toc: yes
     toc_float: yes
   bookdown::pdf_document2:
-    number_sections: false
-    toc: false
+    includes:
+      before_body: ./template/doc_prefix.tex
+      in_header: ./template/preamble.tex
+    number_sections: no
+    toc: no
+    keep_tex: yes
 csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
@@
 Dataset defined as origin corresponded to leaf lifespan ($LL$) and most of the time to leaf mass per area ($LMA$) and leaf nitrogen content per leaf dry mass ($Nmass$).
-We measured variable importance in functionnal traits to explain leaf lifespan with an out-of the bag method applied on a random forest.
-Then, we used a bayesian apporach to test different models with growing level of complexity.
-We retained the model with the best tradeoff between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean sqaure error of prediciton $RMSEP$).
+We measured variable importance in functional traits to explain leaf lifespan with an out-of the bag method applied on a random forest.
+Then, we used a Bayesian approach to test different models with growing level of complexity.
+We retained the model with the best trade-off between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean square error of prediction $RMSEP$).
 We finally tested the new allometry obtained with the selected model with TROLL simulations.
@@
 
-Leaf lifespan $LL$ follows a lognormal law with location infered from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
+Leaf lifespan $LL$ follows a log-normal law with location inferred from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
 Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$.
-All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without presemption following a gamma law.
+All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without preemption following a gamma law.
 $d$ represents the dataset random effects and encompass environmental and protocol variations.
@@
 Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early death of individuals inside simulations.
-For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.
+For instance with this allometry *Symphonia sp 1* (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.
 

### 2018-12-13 draft.Rmd
Draft 0.1.9010 Co-authors sharing

@@
 title: Biodiversity improves tropical forest resilience
-author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
+author: Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille
+  Piponiot, Stéphane Traissac & Bruno Hérault
 date: '`r Sys.Date()`'
 output:
-  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: no
+    toc: yes
+    toc_float: yes
   bookdown::pdf_document2:
@@
       in_header: ./template/preamble.tex
-    number_sections: false
-    toc: false
-    keep_tex: true
-  github_document: default
-  bookdown::html_document2:
-    number_sections: false
-    toc: true
-    toc_float: yes
+    number_sections: no
+    toc: no
+    keep_tex: yes
+  gdoc::gdoc: default
+  bookdown::word_document2: default
 linestretch: 1.5
@@
 link-citations: yes
+gdoc_id: 1yRJitcL7DfCwTmZlUEuQEJg4-QcTfsznYqJ999VV-p8
 ---
@@
 The concept of ecological resilience has met numerous definitions, but encompass both ecosystem ability to adsorb, the resistance component, and recover from disturbance [@Gunderson2010].
-Both exogenous (e.g.
-climate, disturbance features) and endogenous (e.g.
-forest structure and composition) factors can drive ecosystem recovery after disturbance.
+Both exogenous (*e.g.* climate, disturbance features) and endogenous (*e.g.* forest structure and composition) factors can drive ecosystem recovery after disturbance.
 Among the former, logging intensity has been reported to strongly control and increase time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and change in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015].
@@
 
-```{r IrGraphFinal, fig.cap="Definition of recovery function R and resilience indice $I_R$ shown on aboveground biomass dynamics for a simulation with $disturb_{intensity}=50%$."}
+```{r IrGraphFinal, fig.cap="Definition of recovery function R and resilience indice $I_R$ shown on aboveground biomass dynamics for a simulation with $disturb_{intensity}=50\\%$."}
 knitr::include_graphics("images/figure3.png")
@@
 Significant $p-values$ are shown in bold.
-ns, non‐significant; .
+ns, non-significant; .
 p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.",
@@
 ```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects.
-The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0%$).
+The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\\%$).
 Solid lines represents the mean value whereas areas represents the range of values across the 20 simulations with a species richness ($SR$) of 125.
@@
 
-We investigated the effect of biodiversity on tropical forest resilience to disturbance and the variation of biodiversity effect on ecosystem functioning along trajectory of recovery from disturbance by means of a virtual experiment.
-We found that functional diversity improved tropical forest resilience from disturbance.
-Additionally, the mechanisms that underlie the biodiversity effect on forest functioning changed in dominance as trajectories of recovery unfolded.
-Complememtarity between species was the dominant effect of biodiversity in the early times after disturbance.
-Complementarity was then progressively overtaken by a selection effect as more competitive species dominate the forest community.
-This shift was all the more pronounced that disturbance increased in intensity.
-Our results suggest that ecosystem state and past history can strongly influence the mechanisms by which biodiversity impacts ecosystem processes.
-This can help reconciling previous contrasted results obtained when dealing with snapshots of ecosystem state.
+Using an *in silico* experiment, we investigated the effect of biodiversity on tropical forest resilience to disturbance and the variation of biodiversity effect on ecosystem functioning along the recovery trajectory.
+We found that functional diversity improved tropical forest resilience from disturbance with underlying ecological mechanisms changing in dominance with times.
+Complementarity effect was dominant in the early times after disturbance and was progressively overtaken by a selection effect as more competitive species dominate the forest community, a pattern that increased with disturbance intensity.
+Overall, our results suggest that past-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests.
+This can help reconciling previous contrasted results obtained with limited snapshots of ecosystem state.
 
-## Functional diversity improves tropical forest recovery
+## Functional diversity accelerates tropical forest recovery
 
-Our results validated the hypothesis of a significant positive relationship between forest recovery and functional diversity.
-More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008] appeared as a major aspect of resilience if it is strengthened by a high functional evenness [regularity of the distribution of abundance in this volume, @villeger_new_2008].
-Indeed, high functional diversity combined to high evenness avoid hegemonic effect of a few dominant species.
-Functional diversity positive effect on forest resilience can be due to (i) temporal complementarity with the diversity of species response to disturbances and response asynchronicity, resulting in ecological sorting [@Loreau2013; @Sakschewski2016; @Morin2014a], and (ii) functional complementarity reducing competition strength [@Loreau2013] advocating both for the biodiversity insurance hypothesis [@Yachi1999].
+Our results validated the hypothesis of a significant positive relationship between functional diversity and forest recovery rate.
+More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008], and especially its evenness component, appeared as a major aspect of the system resilience by avoiding the hegemonic effect of a few dominant species.
+This can be due to (i) temporal complementarity with the diversity of species response to disturbances and response asynchronicity, resulting in ecological sorting [@Loreau2013; @Sakschewski2016; @Morin2014a], and (ii) functional complementarity reducing competition strength [@Loreau2013] advocating both for the biodiversity insurance hypothesis [@Yachi1999].
 Our results thus confirms both the review of @Daz2001a, advocating against an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity of forest ecosystems.
-In addition, increased species richness will de facto increase chance for high functional diversity through the sampling effect [@Loreau1998; @Morin2014a].
-A higher sampling of regional species pool will thus allow a greater chance to pick more functionally diverse species.
+In addition, increased species richness will *de facto* increase chance for high functional diversity through the sampling effect [@Loreau1998; @Morin2014a].
 
-## Mechanisms underlying the biodiveristy effect on forest ecosystem change along trajectories of recovery from disturbances
+## First complementarity, then selection insure biodiversity effect after disturbance
 
-We found that complementarity among the species was the prevailing effect of biodiversity on forest functioning during the beginning of forest post-disturbance trajectory.
-We interpreted this result as the consequence of niche partitioning.
-By creating additional gaps in the simulated mature forest plots, the simulated disturbances resulted in an increased heterogeneity in the availability in light, which is the main limited resource in dense tropical forest.
-Species are known to vary widely in their growth response to gap creation depending on their functional strategy [@Herault2010; @Laurans2012; @Ruger2009].
+We found that complementarity among the species was the prevailing effect of biodiversity on forest functioning during the beginning of post-disturbance ecosystem trajectory.
+We interpreted this result as the direct consequence of the simulation of natural gaps that resulted in an increased heterogeneity in the availability in light, the main limited resource in dense tropical forest, opening the way for niche partitioning.
+Species are known to vary widely in their demography response to light-enrichment depending on their functional strategy [@Herault2010; @Laurans2012; @Ruger2009].
 Pioneer species with productive leaf tissue and light wood typically quickly close gaps, hence creating a shaded environment where late successional species with more conservative leaf tissue are more competitive.
-By promoting a higher diversity of shade tolerance and growth response to light, species richness hence foster forest response to small-scale perturbations [@Morin2011].
+By inducing a higher diversity of shade tolerance and growth response to light, species richness hence fosters forest response to perturbations [@Morin2011].
 This complementarity effect quickly reduced through time letting the selection effect progressively gain in dominance.
@@
 We did not include the effect of topography, known to result in higher micro environmental heterogeneity.
-Overall, we expect a stronger role of complementarity effect through hydrological and edaphic niche partitioning as water become more limited and in more heterogeneous landscapes [@Baltzer2005; @Jucker2018; @Silvertown2004a].
+Overall, we expect a stronger role of complementarity effect through hydrological and edaphic niche partitioning as water become locally more limited in more heterogeneous landscapes [@Baltzer2005; @Jucker2018; @Silvertown2004a].
 Additionally, we simulated a generic form of disturbance, and our findings might be modulated depending on the disturbance specificity.
 For instance, selective logging will induce both species targeting which could lower the selection effect if dominant species are chosen and cyclic perturbation, increasing the prevalence of early recovery stages where complementarity effect plays the most.
-Consequently, we could hypothesize that species complementarity and functional diversity will be of primary importance regarding selective logging compared to less targeted disturbances with longer cycles such as drought, fires and hurricanes.
+Consequently, we could hypothesize that species complementarity and functional diversity will be of primary importance regarding the system trajectory after selective logging as compared to less targeted disturbances with longer cycles such as drought, fires and hurricanes.
 
@@
 
-Our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depends on the ecosystem state and its history of perturbation, with dominant complementarity effect in recently disturbed mature ecosystem whereas selection effect takes over complementarity in more stable systems with older disturbances.
-In agreement with this, studies in recently disturbed mature forest did found a stronger complementarity effect [@Huang2018; @Poorter2017], whereas a dominant selection effect has been found both in undisturbed forests our new plantations [@Bauters2015; @Finegan2015; @Marechaux2017; @Tobner2016] but see [@DeAvila2018; @Lohbeck2016].
+Our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depends on the current ecosystem state and its previous history of perturbations, with dominant complementarity effect in recently disturbed mature ecosystem whereas selection effect takes over complementarity in more stable systems with older disturbances.
+In agreement with this, studies in recently disturbed mature forest did found a stronger complementarity effect [@Huang2018; @Poorter2017], whereas a dominant selection effect has been found both in undisturbed forests and new plantations [@Bauters2015; @Finegan2015; @Marechaux2017; @Tobner2016] but see [@DeAvila2018; @Lohbeck2016].
+Dominant selection effect of plantation might be explained by the absence of remaining diversity to insure complementarity as opposed to disturbed mature forest (results unshown).
 
 Moreover, we here focus on the effect of biodiversity along trajectory of post-disturbance recovery, and our simulations and analyses were all made at a 16-ha scale.
-However, in addition to the influence of disturbance intensity and temporal trajectory we evidenced here, @Chisholm2013 and @Sullivan2017 found that biodiversity effect vary across spatial scale, with typically a stronger effect at smaller scale (e.g.
-0.04-ha).
-Finally, ecological constraints may differ between ecosystems, for instance pioneer species (e.g.
-hard-wood pioneer species in dry forest [@Prado-Junior2016]), light importance (e.g.
-nitrogen importance in grasslands [@Fargione2007; @Prado-Junior2016]), and maturity concept (e.g.
-temperate versus tropical forests) can differ from tropical wet forest to another system, that may at the origin of biodiversity effects differences.
+However, in addition to the influence of disturbance intensity and temporal trajectory we evidenced here, @Chisholm2013 and @Sullivan2017 found that biodiversity effect vary across spatial scale, with typically a stronger effect at smaller scale (*e.g.* 0.04-ha).
+Finally, ecological constraints may differ between ecosystems, for instance pioneer species (*e.g.* hard-wood pioneer species in dry forest [@Prado-Junior2016]), light importance (*e.g.* nitrogen importance in grasslands [@Fargione2007; @Prado-Junior2016]), and maturity definition (*e.g.* temperate versus tropical forests) can differ from tropical wet forest to another system, that might originate biodiversity effects differences.
 
-Together, this paves the way to a unifying framework to reconcile the apparent contradictions between studies advocating for the sole effect of complementarity or selection effects, and call for an accurate description of the ecosystem spatial-temporal and ecological context in future studies to improve global understanding of biodiversity and ecosystem functioning relations, and thus further open perspectives to a better prediction of ecosystem resilience in a context of increasing disturbances and biodiversity crisis.
+Together, this paves the way to a unifying framework to reconcile the apparent contradictions between studies advocating for the sole effect of complementarity or selection effects, and call for an accurate description of the ecosystem spatial-temporal and ecological context in future studies to improve global understanding of biodiversity and ecosystem functioning relations.
+This further open perspectives to a better prediction of ecosystem resilience in a context of increasing disturbances and biodiversity crisis.
 
@@
     + [PDF](S1-LL.pdf)
-* Supplementary materials S2: Resilience index
-    + [HTML](S2-Resilience.html)
-    + [PDF](S2-Resilience.pdf)
-* Supplementary materials S3: Biodiversity net effect and resilience 
-    + [HTML](S3-RNE.html)
-    + [PDF](S3-RNE.pdf)
-* Supplementary materials S4: Linear models
-    + [HTML](S4-LM.html)
-    + [PDF](S4-LM.pdf)
-* Supplementary materials S5: Biodiversity effect results
-    + [HTML](S5-Biodiversity.html)
-    + [PDF](S5-Biodiversity.pdf)
 

### 2018-12-13 S1-LL.Rmd
Draft 0.1.9010 Co-authors sharing

@@
 ---
-title: "Supplementary material S1: Leaf lifespan model"
+title: 'Supplementary material S1: Leaf lifespan model'
 date: '`r Sys.Date()`'
@@
     keep_tex: yes
+  gdoc::gdoc: default
 csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
@@
 link-citations: yes
+gdoc_id: 1G5GQy7yxmwSsUCcC_Exky_cPYDhwHxNcVXnnILdPAUs
 ---

### 2018-12-13 draft.Rmd
Draft 0.1.9010 Co-authors sharing

@@
 output:
-  bookdown::html_document2:
-    number_sections: no
-    toc: yes
-    toc_float: yes
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
   gdoc::gdoc: default
-  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: no
+    toc: yes
+    toc_float: yes
 linestretch: 1.5

### 2019-01-03 A02-Resilience.Rmd
Draft 0.1.9011 JC and CP review except discussion last paragraph

@@
   scale_color_discrete("Disturbance\nintensity\n(% of BA)", 
-                       labels = c("0 (control)", "50 (disturbed)")) +
-  ggsave(filename = "./images/original1.png", width = 20, height = 10, units = "cm")
+                       labels = c("0 (control)", "50 (disturbed)")) #+
+  # ggsave(filename = "./images/original1.png", width = 20, height = 10, units = "cm")
 ```
@@
     mutate(control = ifelse(is.na(control), treatment, control)) %>% 
-    mutate(type = "Raw"),
+    mutate(type = "Above-ground biomass (MgC/ha)"),
   mature %>% 
@@
     mutate(control = control/control) %>%
-    mutate(type = 'Normalized by control')
+    mutate(type = 'AGB recovery')
 ) %>%
-  mutate(type = factor(type, levels=c('Raw', 'Normalized by control'))) %>% 
   ggplot(aes(x = time)) +
@@
   xlab("Time (years)") +
-  ylab("Aboveground biomass (tonC/ha)") +
+  ylab("") +
   scale_color_discrete("Disturbance\nintensity\n(% of BA)", 

### 2019-01-03 draft.Rmd
Draft 0.1.9011 JC and CP review except discussion last paragraph

@@
 title: Biodiversity improves tropical forest resilience
-author: Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille
-  Piponiot, Stéphane Traissac & Bruno Hérault
-date: '`r Sys.Date()`'
+author: Sylvain Schmitt$^{1,*}$, Isabelle Maréchaux$^2$, Jerome Chave $^3$, Fabian Fischer$^3$, Camille  Piponiot$^4$, Stéphane Traissac$^5$, Bruno Hérault$^6$
 output:
-  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: no
+    toc: yes
+    toc_float: yes
   bookdown::pdf_document2:
@@
   gdoc::gdoc: default
-  bookdown::html_document2:
-    number_sections: no
-    toc: yes
-    toc_float: yes
+  bookdown::word_document2: default
 linestretch: 1.5
-csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
+csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
 bibliography: /home/sylvain/Documents/Bibliography/library.bib
@@
 
+$^1$ UMR BIOGECO, Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
+
+$^2$ AMAP, INRA, University of Montpellier, IRD, CIRAD, CNRS, , Montpellier, France
+
+$^3$ Laboratoire Evolution et Diversité Biologique, UMR5174, CNRS, Université Paul Sabatier, IRD, Toulouse Cedex 9, France
+
+$^4$ Université de Guyane, UMR EcoFoG (Agroparistech, CNRS, Inra, Université des
+Antilles, Cirad), Kourou, French Guiana
+
+$^5$ AgroParisTech, UMR ECOlogie des FOrêts de Guyane, 97387 Kourou, France
+
+$^6$ INPHB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast
+
+$^*$ Corresponding author: sylvain.schmitt@agroparistech.fr
+
 # Abstract
 
-**To be written**
+*The Abstract should be written to be understood in isolation from the rest of the paper and must not exceed 350 words. Using typically four simple, factual, numbered statements, it should:*
+
+- *give the conceptual context*
+- *state the methodological approach*
+- *report the main results and conclusions*
+- *include a final point headed ‘Synthesis’, which sums up the paper’s key message in generic terms that can be understood by non-specialists, indicating clearly how this study has advanced ecological understanding.*
+
+*Advice for optimizing the Abstract (and Title) so that the paper is more likely to be found in online searches is provided at http://authorservices.wiley.com/bauthor/seo.asp .*
+
+## Keywords
+
+*A list in alphabetical order not exceeding eight words or short phrases. More [advice on selecting good keywords can be found here](https://methodsblog.com/2015/12/18/seo/).*
 
@@
 
-Natural disturbances play a major role in tropical forest functioning through gap-phase regeneration [@Veblen1992].
 Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015] and while deforestation has been rightly the focus of worldwide attention, the consequences of human-induced disturbance on tropical forests has been less studied [@Putz2010; @Simula2009].
-However, degraded forests currently cover areas over ten times larger than deforested ones [@Herold2011], a majority of which is being unsustainably logged for timber production [@Blaser2011].
-Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 20.7 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing, for instance, 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005].
-In addition to the direct impact of human on forests, a more insidious disturbance is climate change that modifies regional climates with, for instance, projected increases in the frequency and intensity of convective storms in the Amazon having the potential to increase wind-related tree mortality [@Negron-Juarez2018].
+Degraded forests currently cover areas over ten times larger than deforested ones [@Herold2011], a majority of which is being unsustainably logged for timber production [@Blaser2011].
+Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 2.1 billion of tons of carbon dioxide emissions per year representing 25% of greenhouse gas emissions related to deforestation and forest degradation globally [@Pearson2017].
+In addition to the direct impact of human on forests, climate change modifies regional climates with, for instance, projected increases in the frequency and intensity of convective storms in the Amazon having the potential to increase wind-related tree mortality [@Negron-Juarez2018].
 Understanding the drivers of tropical forest resilience to disturbance is thus critical.
@@
 
-The concept of ecological resilience has met numerous definitions, but encompass both ecosystem ability to adsorb, the resistance component, and recover from disturbance [@Gunderson2010].
-Both exogenous (*e.g.* climate, disturbance features) and endogenous (*e.g.* forest structure and composition) factors can drive ecosystem recovery after disturbance.
-Among the former, logging intensity has been reported to strongly control and increase time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and change in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015].
-Among endogenous drivers, forest pre-disturbance composition has been identified as a key driver of forest recovery [@Herault2018], as different species respond differently to perturbations depending on their ecological strategies and functional properties [@herault_functional_2011].
-Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@Cadotte2011; @Ives2007; @Loreau2013].
-Finally, disturbance intensity itself has been shown to be a key component of forest resilience [@DeAvila2018].
+Ecological resilience refers to the ecosystem ability to resist and to recover from disturbance [@Gunderson2010].
+Following a disturbance, ecosystem recovery is mediated by both exogenous (*e.g.* climate, disturbance features) and endogenous (*e.g.* forest structure and composition) factors.
+Logging intensity (the quantity of extracted timber and subsequent biomass loss) determines to a large extent the time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and shifts in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015].
+Among endogenous drivers, baseline floristic composition appears to be a key driver of forest recovery [@Herault2018], as species respond differently to disturbances depending on their ecological strategies and functional properties [@herault_functional_2011].
+Beyond species identity, tree community diversity may also foster forest resilience to disturbances [@Cadotte2011; @Ives2007; @Loreau2013].
 
 Two types of mechanisms can underlie the effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001].
-Complementarity effect results from resources partitioning and niche differentiation, and facilitation, leading to a more efficient use of available resources at the community level.
-Selection effect results from a higher probability of more diverse community to include particularly efficient species regarding ecosystem processes.
-Moreover the sampling effect additionally implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].
+Complementarity effect results from resource partitioning and niche differentiation, and to facilitation, leading to a more efficient use of available resources at the community level.
+Selection effect results from a higher probability of including functionnaly efficient species in more diverse community.
+Moreover the sampling effect implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].
+Previous theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], and direct observations along natural gradients [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017] suggest that both complementarity and selection mechanisms play a crucial role in ecosystem biomass stock and dynamic [@Cardinale2007; @Poorter2017; @VanderSande2017a].
+The relative importance of each mechanism can change in time: for instance @Fargione2007 observed a shift from selection to complementarity effect over time in experimental grasslands, and @Huang2018 observed an increased complementarity over time with no selection effect in forest establishments experiments.
 
-The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], or direct observations along natural gradient [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017].
-Several studies advocate for the importance of few dominant species through selection effect to regulate ecosystem functioning in an old tree diversity experiment [@Bauters2015], in wet or dry natural forests [@Finegan2015; @Prado-Junior2016], or in human-modified forests [@DeAvila2018; @Lohbeck2016].
-Numerous authors have, however, highlighted the pivotal role of diversity in ecosystem stability [@Morin2014a] and highlighted the role of both complementarity and selection for ecosystem biomass stock and dynamic [@Cardinale2007; @Poorter2017; @VanderSande2017a].
-For instance, @Fargione2007 observed a shift from selection to complementarity effect over time in experimental grasslands, and @Huang2018 observed an increased complementarity over time with no selection effect in forest establishments experiments.
+In tropical forests, the role of biodiversity in ecosystem recovery and the underlying mechanisms remain under-documented [@VanderSande2017].
+Studies relying on forest stand inventories often lack repeatability [@Poorter2017; @Schnitzer2016], rarely include disturbed forests [@Sist2015], and temporal coverage remains too low to assess long-term effects [@Herault2018].
+Forest simulators have thus proven useful to investigate the effect of biodiversity on forest functioning through virtual biodiversity experiments [@Morin2011; @Morin2014a; @Sakschewski2016; @Bohn2017].
+However, the problem of biodiversity representation beyond a few plant functional types has long been an impediment to apply this approach to species-rich tropical forests [@Kazmierczak2014; @Kohler1998; @Purves2008].
 
-Overall, the role of biodiversity in tropical forest recovery from disturbances and the mechanisms underlying the biodiversity effect on forest functioning along long-term dynamics remain under documented [@VanderSande2017].
-Experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Poorter2017; @Schnitzer2016], rarely include disturbed forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018].
-Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014a; @Sakschewski2016; @Bohn2017].
-However, the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].
-
-In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest, that allow to deal with high diversity levels through a parameterization using plant functional traits [@Marechaux2017].
+In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest, that simulates high-diversity levels through a species-specific parameterization using plant functional traits [@Marechaux2017].
 Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning  relationships *in silico* in tropical forest communities.
 More specifically, we used a simulation approach to assess the effect of species and functional diversity on long-term forest recovery from disturbances of different levels.
-We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance.
+We use this approach to manipulate species number and identity across a large number of simulations, and measured the net effect of biodiversity on simulated forest recovery after disturbance.
 In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned ecosystem properties into complementarity and selection effects.
-Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018].
+Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018 for no effect of species richness on productivity].
 
@@
 
-TROLL is an individual-based and spatially explicit forest dynamic model in which the life cycle of individual trees (recruitment, growth, seed production and death) $>1cm$ diameter at breast height (dbh) is simulated.
-Trees grow in a forest light environment computed within a $1m^3$ resolution voxel grid.
-Seedlings below the $1cm$ size class are not modeled on an individual basis, but are part of a seed/seedlings pool.
-A maximum of one tree can establish in each $1*1m$ pixel.
+TROLL is an individual-based and spatially explicit forest dynamic model.
+A detailed description of the model is available in @Marechaux2017, and is only summarized here.
+In TROLL the life cycle (recruitment, growth, seed production and death) is simulated for individual trees $>1\textrm{cm}$ diameter at breast height (dbh).
+Trees grow in a forest light environment computed within a $1\textrm{m}^3$ resolution voxel grid.
+Seedlings below the $1\textrm{cm}$ size class are not modeled on an individual basis, but are part of a seed/seedlings pool.
+A maximum of one tree can establish in each $1\times1\textrm{m}$ pixel.
 Each tree is defined by its age, dbh, height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
@@
 Additionally, each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
-To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which ecophysiological processes are simulated using up-to-date established relationships drawn from the literature.
+To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which ecophysiological processes are simulated using relationships drawn from the literature.
 The model thus represents tree biodiversity in a realistic way, where data requirement and model outputs parallel current trait collection efforts and field inventories, respectively.
@@
 Photosynthetic capacities, as well as leaf respiration rate, are computed from species-specific leaf concentration of nitrogen (N) and phosphorous (P), and leaf mass per area [LMA; @Atkin2015; @Domingues2010].
-The net amount of assimilated carbon available for growth is then partitioned among the different plant organs using empirical factors, and converted into increments of stem volume and leaf area using species-specific wood-density and LMA respectively.
+The net amount of assimilated carbon allocated to growth is then partitioned among the different plant organs using empirical factors, and the fraction allocated to woody growth is converted into increments of stem volume using species-specific wood density; while the fraction allocated to leaves is converted into leaf area increment using LMA.
 Leaf demography is simulated using species-specific leaf lifespan derived from LMA.
-Following an underestimation of leaf lifespan for low LMA species, we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1).
-Natural tree deaths are stochastic events whose background rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010] an locally [@Aubry-Kientz2013].
-Additionally, tree can die due to carbon starvation or treefalls.
+Following an underestimation of leaf lifespan for low LMA species.
+Here we did not use the same relationship between LMA and leaf lifespan as in @Marechaux2017, as we found that it generated an underestimation of leaf lifespan for low LMA species.
+We thus designed a new leaf lifespan allometry, specific to tropical tree species, used throughout this study (see supplementary materials S1).
+Natural tree deaths are stochastic events and their rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010] an also locally [@Aubry-Kientz2013].
+Additionally, trees can die due to carbon starvation or treefalls.
 Below-ground processes, herbaceous plants, epiphytes and lianas are not simulated in this version of TROLL.
-The source code is written in C++ and available at https://github.com/TROLL-code/TROLL.
+The source code is written in C++ and version 2.3.1 used in this study are available at https://github.com/TROLL-code/TROLL.
 A further detailed description of the model is available in @Marechaux2017.
@@
 
-A new disturbance module was developed for this study.
-In order to have a general overview of disturbance effect, we used untargeted disturbance happening only once to be free of confounding factors due to disturbance type specificity.
-At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law and without any individuals nor species targeting.
+In order to have a general overview of disturbance effect, a generic disturbance module was developed for this study.
+At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law, without any individuals nor species targeting to be free of confounding factors due to disturbance type specificity.
 Selected individuals are then removed without triggering a treefall to avoid any side effect.
@@
 In order to assess the role of biodiversity on ecosystem recovery from disturbance, we  simulated forest trajectories for different levels of disturbance intensity ($disturb_{intensity}$) and tree community compositions.
-For each simulation, we first simulated forest regeneration from bare soil for 600 years (pre-disturbance step), then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step).
-Disturbance intensity, as quantified by the percentage of basal area loss ($disturb_{intensity}$), was fixed at 0% (control), 25%, 50% or 75%.
-For each level of disturbance, we performed 20 simulations differing in the original number of species (species richness, $SR=15,25,125$) and composition.
-We proceeded in two steps: (i) for each level of species richness, we created 1000 virtual communities by randomly picking the desired number of species among the regional species pool; (ii) among those 1000 communities, we selected 20 communities for simulations with a large range of convex hull volume $CVH$ [@Cornwell2006a] but with a community mean ($CM$) close to the one of the regional species pool ($CM_{LMA} = 115 \pm 19$, $CM_{wsg} = 0.73 \pm 0.07$, $CM_{dmax} = 0.37 \pm 0.1$, $CM_{hmax} = 40 \pm 11$) in order to get change in functional diversity and not in average functional composition.
+For each simulation, we first conducted a spin-up run, where we simulated forest regeneration from bare soil for 600 years (pre-disturbance step).
+We then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step).
+Disturbance intensity ($disturb_{intensity}$) was fixed at 0% (control), 25%, 50% or 75% of basal area loss.
+For each level of disturbance, we performed 20 simulations differing in the original number of species (species richness $SR\{15,25,125\}$) and in floristic composition.
+For each level of species richness, we first created 1000 virtual communities by randomly picking the desired number of species among the regional species pool.
+Then of those 1000 communities, we selected 20 communities with a large range of convex hull volume $CVH$ [@Cornwell2006a] but with a community mean ($CM$) close to the one of the regional species pool ($CM_{LMA} = 115 \pm 19 ~\textrm{g.m}^{-2}$, $CM_{wsg} = 0.73 \pm 0.07~\textrm{g.cm}^{-3}$, $CM_{dmax} = 0.37 \pm 0.1~\textrm{m}$, $CM_{hmax} = 40 \pm 11~\textrm{m}$) in order to control for average functional composition.
 This led to 240 simulations varying in community composition and disturbance intensity ($3~SR*20~CHV*4~ levels~of~disturbance$).
@@
 
-where $X_D(t)$ and $X_C(t)$ are the values for the ecosystem property $X$ at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$) respectively.
-Note that demographic stochasticity in the model induces a very low variability among replicated simulations with the same set-up so that we we assumed a given simulation is representative of the community properties and dynamics given an initial species pool and disturbance intensity [@Marechaux2017].
+where $X_D(t)$ and $X_C(t)$ are the values for the ecosystem property $X$ at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the control simulation ($disturb_{intensity} = 0\%$), respectively.
+Note that variability among replicated simulations with the same set-up, due to demographic stochasticity in the model, is low enough to assume a given simulation is representative of the community properties and dynamics given an initial species pool and disturbance intensity [@Marechaux2017].
 
 When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery.
-We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Figure \@ref(fig:IrGraphFinal)).
+We then calculated the Euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Figure \@ref(fig:IrGraphFinal)).
 Because the intensity of disturbance was set up by the user, our analysis did not include any ecosystem resistance, i.e.
 the ecosystem capacity to absorb a disturbance event.
-Consequently in this particular case, we used the inverse of cumulated euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):
+Consequently, in this particular case, we used the inverse of the cumulative Euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):
 
@@
 
-A high value of $I_R$ indicates a fast resilience.
-Ecosystem euclidean distance to equilibrium and resilience index $I_R$ were calculated in a multi-dimensional space for the two set of ecosystem property values: (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).
+A high value of $I_R$ indicates a high resilience.
+Distance to equilibrium $d_{eq}$ and resilience index $I_R$ were calculated in a multi-dimensional space for the two set of ecosystem property values: (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).
 
@@
 
-```{r IrGraphFinal, fig.cap="Definition of recovery function R and resilience indice $I_R$ shown on aboveground biomass dynamics for a simulation with $disturb_{intensity}=50\\%$."}
+```{r IrGraphFinal, fig.cap="Definition of recovery function R and resilience index $I_R$ shown for aboveground biomass dynamics for a simulation with $disturb_{intensity}=50\\%$."}
 knitr::include_graphics("images/figure3.png")
@@
 
-We tested the influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem resilience index $I_R$ for both forest structure and functioning using linear regression ($I_R \sim FEve + FDiv$).
+We tested the influence of three factors on the resilience index $I_R$: pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) using linear regression ($I_R \sim SR + FEve + FDiv$).
+The analysis was conducted twice for each disturbance level (25%, 50% and 75% of basal area loss): one with forest structure properties and one with forest functioning properties.
 We tested variable correlation by Pearson's correlation coefficient for continuous variables and one-way ANOVA for species richness $SR$.
 In case of significant correlation, we selected the variable resulting in the model with the higher Aikaike information criterion in order to have independent variables.
-To meet the assumption of normality ecosystem resilience index $I_R$ was log-transformed.
+To meet the assumption of normality, the ecosystem resilience index $I_R$ was log-transformed.
 
@@
 that ecosystem property $X$ for the mixed-species community equals the relative abundance-weighted sum of that in monoculture [@Loreau2001].
-In order to do so, we made simulations for each of the 163 species in monoculture and for each level of disturbance leading to 652 simulations of monoculutre.
+In order to do so, we simulated monoculture of each of the 163 species and for each level of disturbance leading to 652 simulations of monocultures.
 
-We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:
+We then partitioned this net biodiversity effect into complementarity and selection effect using @Loreau2001's partitioning procedure:
 
@@
 
-where $N$ represents the total number of species, and $M$ the vector of $X$ for the 163 simulations in monoculture and $\Delta RX$:
+where $N$ represents the total number of species, and $M$ is the vector of $X$ for the 163 simulations in monoculture. Finally, $\Delta RX$ is as follows:
 
@@
 where $X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$.
-$P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance.
-Similarly to recovery measurement, biodiversity net effect $NE$ can vary over time without disturbance.
-So in order to correctly assess selection and complementarity effect, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure the recovery of the net biodiversity effect $R(NE)$:
+$P_{sp}$ is the relative abundance of the species in the mixture.
+Biodiversity net effect $NE$ can vary over time due to stochasticity, in the absence of disturbance.
+To correctly assess selection and complementarity effect, we normalized $NE$ by the value $NE_C$, calculated from the control simulation ($disturb_{intensity} = 0\%$) to measure the recovery of the net biodiversity effect $R(NE)$:
 
@@
 
-Functional diversity ($FDiv$) and evenness ($FEve$) varied respectively from $0.76\pm0.18$ and $0.46\pm0.24$ to $0.96\pm0.02$ and $0.64\pm0.02$ with increasing species richness ($SR$) from 5 to 125 species.
-Species richness ($SR$) was significantly correlated to functional indices (one-way ANOVA $p-value < 0.001$ for both $FDiv$ and $FEve$) and model $AIC$ was higher without $SR$, consequently we used only functional indices for linear regressions.
+As species richness ($SR$) increased from 5 to 125 species, functional diversity varied from $0.76\pm0.18$ to $0.96\pm0.02$ and evenness ($FEve$) varied from $0.46\pm0.24$ to $0.64\pm0.02$.
+Species richness ($SR$) was significantly correlated to functional indices (one-way ANOVA $p-value < 0.001$ for both $FDiv$ and $FEve$) and model $AIC$ was higher without $SR$, consequently we kept only $FDiv$ and $FEve$ within linear regressions.
 Both functional diversity ($FDiv$) and functional evenness ($FEve$) were positively related with $I_R$, i.e.
-they increased forest resilience (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
-We found similar results for all disturbance levels, with increased significance for higher disturbance intensity.
+forest resilience increased with both functional diversity and functional evenness (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
+Similar results were obtained for all disturbance levels, with increased significance for higher disturbance intensity.
 
 ```{r gIr, echo=FALSE, fig.cap='Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance.
-Resilience indice $I_R$ was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance ($disturb_intensity$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness [FEve, @villeger_new_2008].
+Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_intensity$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$).
 Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.width=10, fig.height=15}
@@
   kable(caption = "Effects of functional diversity on forest resilience.
-Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functionning and for each level of disturbance intensity ($disturb_{intensity}$).
+Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$).
 To meet the assumption of normality $I_R$ was log-transformed.
-Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independant variables.
+Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables.
 Values of FDiv and FEve are coefficients from linear regressions.
@@
 
-The net effect of biodiversity on each ecosystem property was quantified and partitioned between selection and complementarity effects after normalizing using undisturbed control simulation ($disturb_{intensity}=0\%$, see Table \@ref(tab:tNE) and Figure \@ref(fig:gBE)).
+The net effect of biodiversity on each ecosystem property was quantified and partitioned between selection and complementarity effects after normalizing  as explained in the Methods section.
 We found that complementarity effect sharply increased immediately after disturbance and was then the dominant effect of biodiversity during the first decades (Figure \@ref(fig:gBE)).
-Thereafter, the complementarity effect diminished toward a low value close to zero.
-On the contrary, selection effect sharply decreased right after disturbance.
-It then progressively increased to finally represent the majority of the net biodiversity effect about a century after disturbance.
-The dominance of complementarity effect was more important and lasted longer after disturbance with increasing disturbance intensity.
-Finally, 600 years after disturbance, the biodiversity net effect on aboveground biomass was still lower than the control for higher disturbance intensity (Figure \@ref(fig:gBE)).
+Thereafter, the complementarity effect progressively vanished.
+On the contrary, selection effect sharply decreased just after disturbance, and then progressively increased to dominate the net biodiversity effect ca.
+a century after disturbance.
+The effect of complementarity increased and lasted longer after disturbance with increasing disturbance intensity.
+Finally at high disturbance intensity, the net effect of biodiversity on aboveground biomass was still lower than the control 600 years after disturbance (Figure \@ref(fig:gBE)).
 We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$).
-Finally, forest growth primary productivity ($GPP$) recovered only in few years with a dominant complementarity effect.
+Finally, forest growth primary productivity ($GPP$) recovered only within a few years with a dominant complementarity effect.
 
@@
 
-```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects.
+```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss).
 The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\\%$).
-Solid lines represents the mean value whereas areas represents the range of values across the 20 simulations with a species richness ($SR$) of 125.
+Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125.
 Time axis has been square-root transformed to highlight first decades dynamic.'}
@@
   facet_grid(disturbance ~ SR, labeller = 'label_both') +
-  xlab('Time (years)') +
-  ylab('Post-disturbance trajectories of aboveground biomass net effect') +
+  xlab('Time since disturbance (years)') +
+  ylab('Biodiversity effect on aboveground biomass') +
   scale_x_sqrt() +
@@
 Using an *in silico* experiment, we investigated the effect of biodiversity on tropical forest resilience to disturbance and the variation of biodiversity effect on ecosystem functioning along the recovery trajectory.
-We found that functional diversity improved tropical forest resilience from disturbance with underlying ecological mechanisms changing in dominance with times.
-Complementarity effect was dominant in the early times after disturbance and was progressively overtaken by a selection effect as more competitive species dominate the forest community, a pattern that increased with disturbance intensity.
-Overall, our results suggest that past-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests.
+We found that functional diversity improved tropical forest resilience from disturbance with underlying ecological mechanisms changing in dominance with time.
+Complementarity effect was dominant soon after disturbance and was progressively overtaken by a selection effect as more competitive species dominated the forest community, a pattern that increased with disturbance intensity.
+Overall, our results suggest that post-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests.
 This can help reconciling previous contrasted results obtained with limited snapshots of ecosystem state.
@@
 
-Our results validated the hypothesis of a significant positive relationship between functional diversity and forest recovery rate.
-More particularly, functional diversity [volume of the functional space occupied by the community, @villeger_new_2008], and especially its evenness component, appeared as a major aspect of the system resilience by avoiding the hegemonic effect of a few dominant species.
-This can be due to (i) temporal complementarity with the diversity of species response to disturbances and response asynchronicity, resulting in ecological sorting [@Loreau2013; @Sakschewski2016; @Morin2014a], and (ii) functional complementarity reducing competition strength [@Loreau2013] advocating both for the biodiversity insurance hypothesis [@Yachi1999].
-Our results thus confirms both the review of @Daz2001a, advocating against an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity of forest ecosystems.
-In addition, increased species richness will *de facto* increase chance for high functional diversity through the sampling effect [@Loreau1998; @Morin2014a].
+Our results validate the hypothesis of a significant positive relationship between functional diversity and forest recovery rate.
+In particular, functional diversity, defined as the volume of the functional space occupied by the community [@villeger_new_2008], and its evenness component, both turned out to be major predictors of system resilience, as they helped avoid the hegemonic effect of a few dominant species.
+This can be due to (i) temporal complementarity, with the diversity of species response to disturbances and response asynchronicity, resulting in ecological sorting [@Loreau2013; @Sakschewski2016; @Morin2014a], and (ii) functional dissimilarity reducing competition strength [@Loreau2013].
+Our results thus confirm @Daz2001a, advocating against an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity of forest ecosystems.
+In addition, increased species richness will *de facto* increase the chances for high functional diversity through the sampling effect [@Loreau1998; @Morin2014a].
 
-## First complementarity, then selection insure biodiversity effect after disturbance
+## Shifting effects along the successional trajectory insure forest recovery
 
-We found that complementarity among the species was the prevailing effect of biodiversity on forest functioning during the beginning of post-disturbance ecosystem trajectory.
-We interpreted this result as the direct consequence of the simulation of natural gaps that resulted in an increased heterogeneity in the availability in light, the main limited resource in dense tropical forest, opening the way for niche partitioning.
-Species are known to vary widely in their demography response to light-enrichment depending on their functional strategy [@Herault2010; @Laurans2012; @Ruger2009].
-Pioneer species with productive leaf tissue and light wood typically quickly close gaps, hence creating a shaded environment where late successional species with more conservative leaf tissue are more competitive.
-By inducing a higher diversity of shade tolerance and growth response to light, species richness hence fosters forest response to perturbations [@Morin2011].
-This complementarity effect quickly reduced through time letting the selection effect progressively gain in dominance.
-The diminishing of complementarity effect is attributed to the increased abundance of the most competitive species in more stable and less heterogeneous environment.
+We found that complementarity among the species was the prevailing effect of biodiversity on forest functioning just after a disturbance.
+This is a direct consequence of the appearance of natural gaps and an increased heterogeneity in light availability, the main limiting resource in dense tropical forest.
+Species are known to vary widely in their demographic response to light availability depending on their functional strategy [@Herault2010; @Laurans2012; @Ruger2009].
+Pioneer species with productive leaf tissue and light wood tend to occupy gaps fast after their opening, hence creating a shaded environment where late successional species with more conservative leaf tissue later establish.
+By inducing a higher diversity of shade tolerance and growth response to light in the species pool, species richness hence fosters forest response to perturbations [@Morin2011].
+This complementarity effect quickly decreased through time while the selection effect progressively gained in dominance.
+The diminishing of complementarity effect is attributed to the increased abundance of the most competitive species in more stable and less heterogeneous environments.
 
-## Perspectives in forest ecosystem modelling
+## Implications of modelling choices
 
-In this study we took advantage of a forest simulator able to jointly simulate biodiversity and ecosystem functioning to explore the effect of biodiversity on forest resilience on a long temporal scale, that remains out of reach of most experimental approaches.
+In this study we used an individual-based forest simulator to jointly simulate biodiversity and ecosystem functioning, and we explored the effect of biodiversity on forest resilience over centuries, a timeframe that remains out of reach of most experimental approaches.
 By virtually manipulating the number and identity of species across a large number of simulations, we were able to evidence the role of diversity in forest resilience and the changing driver of biodiversity effect on forest functioning along trajectory of recovery.
 
-As any modelling approaches, it also comes with simplifications and assumptions.
+As any modelling approach, our study comes with simplifications and assumptions.
 In our simulations, light was the unique resource whose variability was explicitly accounted for [@Marechaux2017].
-Our results should thus be restricted to situations where aboveground competition for light overcomes the effect of belowground competition for water and nutrients, a reasonable assumption in humid tropical forests, even in early successional stages [@VanBreugel2012].
-However this assumption may be less suitable under the predicted increase in drought intensity and frequency.
+Our analysis thus simulates a situation where aboveground competition for light overcomes the effect of belowground competition for water and nutrients, a reasonable assumption in humid tropical forests, even in early successional stages [@VanBreugel2012].
+However, this assumption may be less suitable under the predicted increase in drought intensity and frequency.
 We did not include the effect of topography, known to result in higher micro environmental heterogeneity.
-Overall, we expect a stronger role of complementarity effect through hydrological and edaphic niche partitioning as water become locally more limited in more heterogeneous landscapes [@Baltzer2005; @Jucker2018; @Silvertown2004a].
-Additionally, we simulated a generic form of disturbance, and our findings might be modulated depending on the disturbance specificity.
-For instance, selective logging will induce both species targeting which could lower the selection effect if dominant species are chosen and cyclic perturbation, increasing the prevalence of early recovery stages where complementarity effect plays the most.
-Consequently, we could hypothesize that species complementarity and functional diversity will be of primary importance regarding the system trajectory after selective logging as compared to less targeted disturbances with longer cycles such as drought, fires and hurricanes.
+We expect a stronger role of complementarity effect through hydrological and edaphic niche partitioning as water become locally more limited in more heterogeneous landscapes [@Baltzer2005; @Jucker2018; @Silvertown2004a].
+
+Additionally, we simulated a generic form of disturbance, and our findings might differ with more specific types of disturbance.
+For instance, selective logging target a few species, and this could lower the selection effect if dominant species are chosen.
+Repeated disturbances such as periodic harvests would increase the prevalence of early recovery stages during which complementarity effect plays the largest role.
+Consequently, species complementarity could have an even bigger importance in forest recovery after repeated disturbances that target dominant species, as can be the case with selective logging.
 
@@
 
-Our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depends on the current ecosystem state and its previous history of perturbations, with dominant complementarity effect in recently disturbed mature ecosystem whereas selection effect takes over complementarity in more stable systems with older disturbances.
-In agreement with this, studies in recently disturbed mature forest did found a stronger complementarity effect [@Huang2018; @Poorter2017], whereas a dominant selection effect has been found both in undisturbed forests and new plantations [@Bauters2015; @Finegan2015; @Marechaux2017; @Tobner2016] but see [@DeAvila2018; @Lohbeck2016].
+Our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depend on the current ecosystem state and its previous history of disturbances, with dominant complementarity effect in recently disturbed ecosystems whereas selection effect takes over complementarity in more anciently disturbed systems.
+In agreement with this, studies in recently disturbed forest did find a stronger complementarity effect [@Huang2018; @Poorter2017], whereas a dominant selection effect has been found both in undisturbed forests and new plantations [@Bauters2015; @Finegan2015; @Marechaux2017; @Tobner2016] but see [@DeAvila2018; @Lohbeck2016].
 Dominant selection effect of plantation might be explained by the absence of remaining diversity to insure complementarity as opposed to disturbed mature forest (results unshown).
 
-Moreover, we here focus on the effect of biodiversity along trajectory of post-disturbance recovery, and our simulations and analyses were all made at a 16-ha scale.
-However, in addition to the influence of disturbance intensity and temporal trajectory we evidenced here, @Chisholm2013 and @Sullivan2017 found that biodiversity effect vary across spatial scale, with typically a stronger effect at smaller scale (*e.g.* 0.04-ha).
-Finally, ecological constraints may differ between ecosystems, for instance pioneer species (*e.g.* hard-wood pioneer species in dry forest [@Prado-Junior2016]), light importance (*e.g.* nitrogen importance in grasslands [@Fargione2007; @Prado-Junior2016]), and maturity definition (*e.g.* temperate versus tropical forests) can differ from tropical wet forest to another system, that might originate biodiversity effects differences.
+We here focus on the effect of biodiversity along a trajectory of post-disturbance recovery, and our simulations and analyses were all at a 16-ha scale.
+However, @Chisholm2013 and @Sullivan2017 found that the biodiversity effect varies across spatial scales, with typically stronger effects at smaller scales (*e.g.* 0.04-ha).
+Ecological constraints may also differ among ecosystems, for instance pioneer species (*e.g.* hard-wood pioneer species in dry forest [@Prado-Junior2016]), light importance (*e.g.* nitrogen importance in grasslands [@Fargione2007; @Prado-Junior2016]), and maturity definition (*e.g.* temperate versus tropical forests) can differ from tropical wet forest to another system, that might cause differences in biodiversity effects.
 
@@
 * Supplementary materials S1: Leaf lifespan model
-    + [HTML](S1-LL.html)
-    + [PDF](S1-LL.pdf)
+
+# Acknowledgements
+
+* METIST
+* GFClim
+
+# Authors’ contributions
+
+SS, BH and ST conceived the ideas, designed methodology and analysed model outputs; SS, IM and BH led the writing of the manuscript.
+All authors contributed critically to the drafts and gave final approval for publication.
 

### 2019-01-07 S1-LL.Rmd
Draft 0.1.9012 Co-authors sharing v2

--- 2018-12-13 S1-LL.Rmd
@@
 ---
-title: "Supplementary material S1: Leaf lifespan model"
-date: '`r Sys.Date()`'
+title: 'Supplementary material S1: Leaf lifespan model'
 output:
   bookdown::word_document2: default
-  bookdown::html_document2:
-    number_sections: no
-    toc: yes
-    toc_float: yes
   bookdown::pdf_document2:
@@
     keep_tex: yes
-csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
+  bookdown::html_document2:
+    number_sections: no
+    toc: yes
+    toc_float: yes
+csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
 bibliography: /home/sylvain/Documents/Bibliography/library.bib
@@
 
-TROLL model previous implementation encompass Reich's 1991 and 1997 and Wright's 2004 allometries to estimate leaf lifespan with  [@Reich1991a; @Reich1997; @wright_worldwide_2004].
-But we have shown that Reich's allometries are underestimating leaf lifespan for low LMA species.
-Moreover simulations estimated unrealistically low aboveground biomass for low LMA species.
-We assumed Reich's allometries underestimation of leaf lifespan for low LMA species being the source of unrealistically low aboveground biomass inside TROLL simulations.
+TROLL model previous implementation encompasses Reich's 1991 and 1997 and Wright's 2004 allometries to estimate leaf lifespan with  [@Reich1991a; @Reich1997; @wright_worldwide_2004].
+But we have shown that Reich's allometries underestimates leaf lifespan for low LMA species.
+As a consequence, simulations estimated unrealistically low aboveground biomass due to an early death of individual with low LMA.
 We decided to find a better allometry with @wright_worldwide_2004 GLOPNET dataset.
@@
 We measured variable importance in functional traits to explain leaf lifespan with an out-of the bag method applied on a random forest.
-Then, we used a Bayesian approach to test different models with growing level of complexity.
+Then, we used a Bayesian approach to test different models with growing levels of complexity.
 We retained the model with the best trade-off between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean square error of prediction $RMSEP$).
@@
 
-The selected model had a maximum likelihood of 13.6 and a RMSEP of 12 months:
+We selected the following model:
 
@@
 
-Leaf lifespan $LL$ follows a log-normal law with location inferred from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
-Each ${\beta_i}_d$ is following a normal law located on $\beta_i$ with a scale of $\sigma_i$.
-All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without preemption following a gamma law.
+Leaf lifespan $LL$ follows a log-normal distribution with location inferred from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
+Each ${\beta_i}_d$ follows a normal distribution located on $\beta_i$ with a scale of $\sigma_i$.
+All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without preemption following a gamma distribution.
 $d$ represents the dataset random effects and encompass environmental and protocol variations.
 
-```{r A2LLpred, echo=FALSE, fig.cap='Leaflifespan predictions for the selected model with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values.
+```{r A2LLpred, echo=FALSE, fig.cap='Leaf lifespan predictions for the selected model with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values.
 Leaf lifespan (LL) is predicted with model M10 fit.
@@
 
-Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early death of individuals inside simulations.
-For instance with this allometry *Symphonia sp 1* (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.
+Simulations with this new allometry have realistic estimations of above-ground biomass even with low-LMA species, resolving our initial issue.
+For instance with this allometry *Symphonia sp 1* (a low LMA species) is now reaching a realistic aboveground biomass above 400  $\textrm{tonC.ha}^{-1}$ and realistic diameter and age distribution inside the final population.
 

### 2019-01-07 S2-Biodiversity.Rmd
Draft 0.1.9012 Co-authors sharing v2

---
title: 'Supplementary material S2: Biodiversity trajectories'
output:
  bookdown::word_document2: default
  bookdown::pdf_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 12,
                cache = T, cache.lazy = F)
```

```{r BE, echo=FALSE}
rich <- c(5,25,125)
cvh <- 1:20
dist <- c(0,25,50,75)
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
doe <- rename(doe, SR = richness)
load(file.path(path, 'disturbanceData.Rdata'))
load(file.path(path, 'disturbanceBE.Rdata'))
load(file.path(path, 'disturbanceBErecovery.Rdata'))
load(file.path(path, '../../maturity', 'matureBE.Rdata'))
load(file.path(path, '../../maturity', 'matureData.Rdata'))
mature <- datal$time ; rm(datal)
```

This document explore the biodiversity trajectories of aboveground biomass partitioned into complementarity and selection effects from forest establishment from a bare soil to 600 years after the disturbance event (Figure \@ref(fig:gBE)). Similarly to the aboveground biomass of the ecosystem, the net biodiversity effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increases following the disturbance (sub-figure on right-bottom on figure \@ref(fig:gBE)). Interestingly, we can see that this pattern of decrease of selection effect after disturbance and recovery after is similar to the effect of succession in forest early stages, when the new cohort arrives selection effect suddenly decreases and complementarity effect insure a small net effect (first century, sub-figure on left-bottom on figure \@ref(fig:gBE)). Anyway the pattern is stronger after a disturbance, maybe due to an increased diversity allowing a stronger diversity effect compared to the absence of already existing diversity when forest established from bare soil.

```{r gBE, fig.cap='Dynamic of complementarity and selection effects before and after the disturbance event for one ecosystem simulation including 125 species. Solid lines represents the mean value whereas shaded areas represents the interval of values among simulations.', fig.height=10}
g <- BE %>% 
  select(sim, BE, time, agb) %>% 
  separate(sim, c("SR", "CHV", "disturbance"), remove = F) %>% 
  rbind(ecosystem %>% 
          mutate(BE = "ecosystem") %>%
          select(sim, BE, time, agb) %>% 
          separate(sim, c("SR", "CHV", "disturbance"), remove = F)) %>% 
  rbind(BE_mature %>% 
          mutate(time = time - 600) %>% 
          select(sim, BE, time, agb) %>% 
          separate(sim, c("SR", "CHV"), remove = F) %>% 
          mutate(disturbance = "0-75") %>% 
          separate_rows(disturbance, sep = "-", convert = T)) %>% 
  rbind(mature %>% 
          mutate(time = time-600) %>% 
          mutate(BE = "ecosystem") %>% 
          select(sim, BE, time, agb) %>% 
          separate(sim, c("SR", "CHV"), remove = F) %>% 
          mutate(disturbance = "0-75") %>% 
          separate_rows(disturbance, sep = "-", convert = T)) %>% 
  filter(SR == 125, disturbance %in% c(0,75), CHV == 10) %>% # 10, 3
  reshape2::dcast(SR + CHV + BE + time ~ disturbance, value.var = "agb") %>% 
  rename(control = `0`, disturbed = `75`) %>% 
  ggplot(aes(x = time, col = BE, group = BE)) +
  geom_line(aes(y = control, linetype = "control")) +
  geom_line(aes(y = disturbed, linetype = "disturbed")) + 
  xlab("Time (years)") + 
  ylab("Aboveground biomass (tonC/ha)") +
  theme(text = element_text(size=20),
        legend.key.size=unit(2,"line"))
g0 <- g + 
  geom_line(aes(y = control, linetype = "control")) +
  geom_line(aes(y = disturbed, linetype = "disturbed")) + 
  geom_ribbon(aes(ymin = disturbed, ymax = control), 
              col = NA, fill = "lightgrey", alpha = 0.5) +
  scale_color_discrete("Biodiversity\neffect") +
  xlab("")
gdist <- g + 
  geom_line(aes(y = control, linetype = "control"), lwd = 1.4) +
  geom_line(aes(y = disturbed, linetype = "disturbed"), lwd = 1.4) + 
  xlim(-10,100) + ylim(-30000, 40000) +
  scale_color_discrete(guide = F) +
  scale_linetype_discrete(guide = F) +
  ylab("")
gstart <- g + 
  geom_line(aes(y = control, linetype = "control"), lwd = 1.4) +
  geom_line(aes(y = disturbed, linetype = "disturbed"), lwd = 1.4) + 
  xlim(-600,-500) + ylim(-30000, 40000) +
  scale_color_discrete(guide = F) +
  scale_linetype_discrete(guide = F) +
  ylab("")
gridExtra::grid.arrange(g0, gstart, gdist, layout_matrix = rbind(c(1, 1), c(2, 3)))
```

### 2019-01-07 draft.Rmd
Draft 0.1.9012 Co-authors sharing v2

@@
     keep_tex: yes
-  gdoc::gdoc: default
-  bookdown::word_document2: default
+  bookdown::html_document2:
+    number_sections: no
+    toc: yes
+    toc_float: yes
 linestretch: 1.5
@@
 link-citations: yes
-gdoc_id: 1yRJitcL7DfCwTmZlUEuQEJg4-QcTfsznYqJ999VV-p8
 ---
@@
 
-*The Abstract should be written to be understood in isolation from the rest of the paper and must not exceed 350 words. Using typically four simple, factual, numbered statements, it should:*
+1.
+Climate change and direct human activities are currently deeply modifying forest disturbance regimes, highlighting the critical need to better understand the drivers of forest resilience.
+In tropical forests, the role of biodiversity in ecosystem recovery and the underlying mechanisms remain under-documented.
+Studies relying on forest stand inventories often lack repeatability, rarely include disturbed forests, and temporal coverage remains too low to assess long-term effects.
+Forest simulators have thus proven useful to investigate the effect of biodiversity on forest functioning through virtual biodiversity experiments.
+However, the problem of biodiversity representation beyond a few plant functional types has long been an impediment in applying this approach to species-rich tropical forests, until now.
+2.
+Using an individual-based and spatially-explicit model of tropical forest simulating high-diversity levels, we assessed *in silico* the effect of species and functional diversity on long-term forest recovery from disturbances of different levels.
+We used this approach to manipulate species number and identity across a large number of simulations, and measured the net effect of biodiversity on simulated forest recovery after disturbance.
+In order to further investigate the mechanisms underlying the effect of biodiversity, we partitioned ecosystem properties into complementarity and selection effects.
+3.
+We found that functional diversity improved tropical forest resilience from disturbance with underlying ecological mechanisms changing in dominance with time.
+Complementarity effect was dominant soon after disturbance and was progressively overtaken by a selection effect as more competitive species dominated the forest community, a pattern that increased with disturbance intensity.
+4.
+*Synthesis*.
+We found that the mechanisms through which biodiversity influence forest functioning strongly depend on the current ecosystem state and its previous history of disturbances, shifting from dominant complementarity effect in recently disturbed ecosystems to selection effect in anciently disturbed systems.
+Our results suggest that pre-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests and can help reconciling previous contrasted results obtained with limited snapshots of ecosystem state.
 
-- *give the conceptual context*
-- *state the methodological approach*
-- *report the main results and conclusions*
-- *include a final point headed ‘Synthesis’, which sums up the paper’s key message in generic terms that can be understood by non-specialists, indicating clearly how this study has advanced ecological understanding.*
-
-*Advice for optimizing the Abstract (and Title) so that the paper is more likely to be found in online searches is provided at http://authorservices.wiley.com/bauthor/seo.asp .*
+<!-- *The Abstract should be written to be understood in isolation from the rest of the paper and must not exceed 350 words.
+Using typically four simple, factual, numbered statements, it should: (i) give the conceptual context, (ii) state the methodological approach, (iii) report the main results and conclusions, and (iv) include a final point headed ‘Synthesis’, which sums up the paper’s key message in generic terms that can be understood by non-specialists, indicating clearly how this study has advanced ecological understanding.* -->
 
@@
 
-*A list in alphabetical order not exceeding eight words or short phrases. More [advice on selecting good keywords can be found here](https://methodsblog.com/2015/12/18/seo/).*
+Resilience; Recovery; Functional diversity; Tropical forest; Complementarity effect; Selection effect; Biodiversity-ecosystem functioning;
 
@@
 
-Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015] and while deforestation has been rightly the focus of worldwide attention, the consequences of human-induced disturbance on tropical forests has been less studied [@Putz2010; @Simula2009].
+Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015], and while deforestation have been rightly the focus of worldwide attention, the consequences of human-induced disturbance on tropical forests has been less studied [@Putz2010; @Simula2009].
 Degraded forests currently cover areas over ten times larger than deforested ones [@Herold2011], a majority of which is being unsustainably logged for timber production [@Blaser2011].
 Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 2.1 billion of tons of carbon dioxide emissions per year representing 25% of greenhouse gas emissions related to deforestation and forest degradation globally [@Pearson2017].
-In addition to the direct impact of human on forests, climate change modifies regional climates with, for instance, projected increases in the frequency and intensity of convective storms in the Amazon having the potential to increase wind-related tree mortality [@Negron-Juarez2018].
+In addition to direct human impact on forests, climate change modifies regional climates with, for instance, projected increases in the frequency and intensity of convective storms in the Amazon having the potential to increase wind-related tree mortality [@Negron-Juarez2018].
 Understanding the drivers of tropical forest resilience to disturbance is thus critical.
@@
 Ecological resilience refers to the ecosystem ability to resist and to recover from disturbance [@Gunderson2010].
-Following a disturbance, ecosystem recovery is mediated by both exogenous (*e.g.* climate, disturbance features) and endogenous (*e.g.* forest structure and composition) factors.
+Following a disturbance, ecosystem recovery is determined by both exogenous (*e.g.* climate, disturbance features) and endogenous (*e.g.* forest structure and composition) factors.
 Logging intensity (the quantity of extracted timber and subsequent biomass loss) determines to a large extent the time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and shifts in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015].
@@
 Two types of mechanisms can underlie the effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001].
-Complementarity effect results from resource partitioning and niche differentiation, and to facilitation, leading to a more efficient use of available resources at the community level.
-Selection effect results from a higher probability of including functionnaly efficient species in more diverse community.
+The complementarity effect results from resource partitioning and niche differentiation, and to facilitation, leading to a more efficient use of available resources at the community level.
+The selection effect describes how more diverse communities have a higher probability of including functionally efficient species.
 Moreover the sampling effect implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].
-Previous theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], and direct observations along natural gradients [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017] suggest that both complementarity and selection mechanisms play a crucial role in ecosystem biomass stock and dynamic [@Cardinale2007; @Poorter2017; @VanderSande2017a].
-The relative importance of each mechanism can change in time: for instance @Fargione2007 observed a shift from selection to complementarity effect over time in experimental grasslands, and @Huang2018 observed an increased complementarity over time with no selection effect in forest establishments experiments.
+Previous theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], and direct observations along natural gradients [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017] suggest that both complementarity and selection mechanisms play a crucial role in ecosystem biomass stock and dynamics [@Cardinale2007; @Poorter2017; @VanderSande2017a].
+The relative importance of each mechanism varies in time: for instance @Fargione2007 observed a shift from selection to complementarity effect over time in experimental grasslands, and @Huang2018 observed an increased complementarity over time with no selection effect in forest establishments experiments.
 
@@
 Forest simulators have thus proven useful to investigate the effect of biodiversity on forest functioning through virtual biodiversity experiments [@Morin2011; @Morin2014a; @Sakschewski2016; @Bohn2017].
-However, the problem of biodiversity representation beyond a few plant functional types has long been an impediment to apply this approach to species-rich tropical forests [@Kazmierczak2014; @Kohler1998; @Purves2008].
+However, the problem of biodiversity representation beyond a few plant functional types has long been an impediment in applying this approach to species-rich tropical forests [@Kazmierczak2014; @Kohler1998; @Purves2008].
 
@@
 More specifically, we used a simulation approach to assess the effect of species and functional diversity on long-term forest recovery from disturbances of different levels.
-We use this approach to manipulate species number and identity across a large number of simulations, and measured the net effect of biodiversity on simulated forest recovery after disturbance.
-In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned ecosystem properties into complementarity and selection effects.
-Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018 for no effect of species richness on productivity].
+We used this approach to manipulate species number and identity across a large number of simulations, and measured the net effect of biodiversity on simulated forest recovery after disturbance.
+In order to further investigate the mechanisms underlying the effect of biodiversity, we partitioned ecosystem properties into complementarity and selection effects.
+Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest would recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018 for no effect of species richness on productivity].
 
@@
 TROLL is an individual-based and spatially explicit forest dynamic model.
-A detailed description of the model is available in @Marechaux2017, and is only summarized here.
+The model is described in detail in @Marechaux2017, and only summarized here.
 In TROLL the life cycle (recruitment, growth, seed production and death) is simulated for individual trees $>1\textrm{cm}$ diameter at breast height (dbh).
-Trees grow in a forest light environment computed within a $1\textrm{m}^3$ resolution voxel grid.
+Trees grow in a light environment computed within a $1\textrm{m}^3$ voxel grid.
 Seedlings below the $1\textrm{cm}$ size class are not modeled on an individual basis, but are part of a seed/seedlings pool.
-A maximum of one tree can establish in each $1\times1\textrm{m}$ pixel.
-Each tree is defined by its age, dbh, height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
+A maximum of one tree can establish per squaremeter.
+Once established, the tree is defined by age, dbh, height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
 Tree geometry is calculated with allometric relationships.
-Additionally, each tree is flagged with a species label inherited from the parent tree through the seedling recruitment.
+Additionally, each tree is flagged with a species label inherited from the parent tree through seedling recruitment.
 To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which ecophysiological processes are simulated using relationships drawn from the literature.
-The model thus represents tree biodiversity in a realistic way, where data requirement and model outputs parallel current trait collection efforts and field inventories, respectively.
+The model thus represents tree biodiversity in a realistic way, where data requirement and model outputs parallel current trait collection efforts and field inventories.
 
@@
 Leaf demography is simulated using species-specific leaf lifespan derived from LMA.
-Following an underestimation of leaf lifespan for low LMA species.
 Here we did not use the same relationship between LMA and leaf lifespan as in @Marechaux2017, as we found that it generated an underestimation of leaf lifespan for low LMA species.
 We thus designed a new leaf lifespan allometry, specific to tropical tree species, used throughout this study (see supplementary materials S1).
-Natural tree deaths are stochastic events and their rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010] an also locally [@Aubry-Kientz2013].
+Natural tree deaths are stochastic events and their rate is negatively correlated with wood density within TROLL, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010] an also locally [@Aubry-Kientz2013].
 Additionally, trees can die due to carbon starvation or treefalls.
@@
   Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
-  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breasth height threshold', 'asymptotic height', 'parameter of the tree-height-dbh allometry'),
-  Units = c('$g.m^{-2}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$g.cm^{-3}$', '$m$', '$m$', '$m$')
+  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breast height threshold', 'asymptotic height', 'parameter of the tree height-dbh allometry'),
+  Units = c('$\\textrm{g.m}^{-2}$', '$\\textrm{mg.g}^{-1}$', '$\\textrm{mg.g}^{-1}$', '$\\textrm{g.cm}^{-3}$', '$\\textrm{m}$', '$\\textrm{m}$', '$\\textrm{m}$')
 )
@@
 In order to have a general overview of disturbance effect, a generic disturbance module was developed for this study.
-At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law, without any individuals nor species targeting to be free of confounding factors due to disturbance type specificity.
-Selected individuals are then removed without triggering a treefall to avoid any side effect.
+At a given iteration $disturb_{iter}$, individual trees are randomly picked following a uniform law, targeting neither individuals nor species.
+This ensures that there are no confounding factors due to disturbance type specificity.
+Selected individuals are then removed without triggering a treefall to control disturbed biomass volume.
 The operation is repeated until the cumulative basal area of removed trees reaches targeted threshold ($disturb_{intensity}$).
@@
 
-All simulations were made with a parameterization for an Amazonian forest of French Guiana, with 163 parameterized species, as in @Marechaux2017.
+All simulations were made with a parameterization for an Amazonian forest of French Guiana, picked in a pool of 163 parameterized species, as in @Marechaux2017.
 
@@
 
-Simulated tropical forests were characterized in terms of changes in ecosystem properties regarding (i) forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and (ii) forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic day-time respiration ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic night-time respiration ($Rnight$ in $MgC.ha^{-1}$)).
+Simulated tropical forests were characterized in terms of changes in ecosystem properties regarding (i) forest structure (aboveground biomass ($AGB$ in $\textrm{Mg~C.ha}^{-1}$), basal area ($BA$ in $\textrm{m}^2.\textrm{ha}^{-1}$), total number of stem ($N$), number of stem above $10~\textrm{cm}$ diameter ($N10$), and number of stem above $30~\textrm{cm}$ diameter ($N30$)) and (ii) forest functioning (growth primary productivity ($GPP$ in $\textrm{MgC.ha}^{-1}$), net primary productivity ($NPP$ in $\textrm{MgC.ha}^{-1}$), tree autotrophic day-time respiration ($Rday$ in $\textrm{MgC.ha}^{-1}$) and tree autotrophic night-time respiration ($Rnight$ in $\textrm{MgC.ha}^{-1}$)).
 
@@
 We then calculated the Euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Figure \@ref(fig:IrGraphFinal)).
-Because the intensity of disturbance was set up by the user, our analysis did not include any ecosystem resistance, i.e.
+Because we set up the intensity of disturbance, our analysis did not include any ecosystem resistance, i.e.
 the ecosystem capacity to absorb a disturbance event.
@@
 A high value of $I_R$ indicates a high resilience.
-Distance to equilibrium $d_{eq}$ and resilience index $I_R$ were calculated in a multi-dimensional space for the two set of ecosystem property values: (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).
+Distance to equilibrium $d_{eq}$ and resilience index $I_R$ were calculated in a multi-dimensional space for the two sets of ecosystem property values: (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).
 
@@
 
-For each ecosystem property $X$, we further investigated the effect of biodiversity and its variation along post-disturbance trajectories.
-The net effect of biodiversity ($NE$) on ecosystem property $X$ is defined as the difference between the ecosystem property $X$ of the simulated mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e.
+For each ecosystem property $X$, we investigated the effect of biodiversity and its variation along post-disturbance trajectories.
+The net effect of biodiversity ($NE$) on ecosystem property $X$ is defined as the difference between the ecosystem property $X$ of the simulated mixed-species community and the one expected under the null hypothesis.
+The latter would mean that there is no effect of biodiversity, i.e.
 that ecosystem property $X$ for the mixed-species community equals the relative abundance-weighted sum of that in monoculture [@Loreau2001].
@@
 forest resilience increased with both functional diversity and functional evenness (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
-Similar results were obtained for all disturbance levels, with increased significance for higher disturbance intensity.
+Similar results were obtained for all disturbance levels, with smaller p-values for higher disturbance intensity.
 
 ```{r gIr, echo=FALSE, fig.cap='Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance.
-Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_intensity$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$).
+Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_{intensity}$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$).
 Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.width=10, fig.height=15}
@@
 
-The net effect of biodiversity on each ecosystem property was quantified and partitioned between selection and complementarity effects after normalizing  as explained in the Methods section.
-We found that complementarity effect sharply increased immediately after disturbance and was then the dominant effect of biodiversity during the first decades (Figure \@ref(fig:gBE)).
-Thereafter, the complementarity effect progressively vanished.
-On the contrary, selection effect sharply decreased just after disturbance, and then progressively increased to dominate the net biodiversity effect ca.
+The net effect of biodiversity on each ecosystem property was quantified and partitioned into selection and complementarity effects after normalizing, as explained in the Methods section.
+We found that the complementarity effect sharply increased immediately after disturbance, was the dominant effect of biodiversity during the first decades and then progressively vanished (Figure \@ref(fig:gBE)).
+When increasing the disturbance intensity, the effect of complementarity increased further and dominated even longer.
+The selection effect, on the other hand, sharply decreased just after disturbance, and only gradually came to play a more important role, dominating the net biodiversity effect ca.
 a century after disturbance.
-The effect of complementarity increased and lasted longer after disturbance with increasing disturbance intensity.
-Finally at high disturbance intensity, the net effect of biodiversity on aboveground biomass was still lower than the control 600 years after disturbance (Figure \@ref(fig:gBE)).
-We found similar results but with an amplified signal for basal area ($BA$) and stem abundance ($N$).
+Finally, at high disturbance intensity, the net effect of biodiversity on aboveground biomass was still lower than the control 600 years after disturbance (Figure \@ref(fig:gBE)).
+We found similar results, but with an amplified signal, for basal area ($BA$) and stem abundance ($N$).
 Finally, forest growth primary productivity ($GPP$) recovered only within a few years with a dominant complementarity effect.
@@
     name = c('aboveground biomass', 'basal area', 'number of stems', 
-             'number of stems above 10cm dbh', 'growth primary production', 
+             'number of stems above 10 cm dbh', 'growth primary production', 
              'net primary production', 'autotrophic respiration during day',
              'autotrophic respiration during night'),
-    unit = c('$tonC.ha^{-1}$', '$m^2.ha^{-1}$', '$n.ha^{-1}$', '$n.ha^{-1}$', 
-             '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$', '$MgC.ha^{-1}$'))) %>% 
+    unit = c('$\\textrm{MgC.ha}^{-1}$', '$\\textrm{m}^2.\\textrm{ha}^{-1}$', '$\\textrm{n.ha}^{-1}$', '$\\textrm{n.ha}^{-1}$', 
+             '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$'))) %>% 
   mutate(NE = paste0("$", round(mean,2), "\\pm", round(sd,2), "$")) %>% 
@@
 We found that functional diversity improved tropical forest resilience from disturbance with underlying ecological mechanisms changing in dominance with time.
-Complementarity effect was dominant soon after disturbance and was progressively overtaken by a selection effect as more competitive species dominated the forest community, a pattern that increased with disturbance intensity.
-Overall, our results suggest that post-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests.
+Complementarity effect was dominant soon after disturbance and was progressively overtaken by a selection effect, as more competitive species dominated the forest community, a pattern that increased with disturbance intensity.
+Overall, our results suggest that pre-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests.
 This can help reconciling previous contrasted results obtained with limited snapshots of ecosystem state.
@@
 
-Our results validate the hypothesis of a significant positive relationship between functional diversity and forest recovery rate.
-In particular, functional diversity, defined as the volume of the functional space occupied by the community [@villeger_new_2008], and its evenness component, both turned out to be major predictors of system resilience, as they helped avoid the hegemonic effect of a few dominant species.
+Our results provide strong evidence of a significant positive relationship between functional diversity and forest recovery rate.
+In particular, functional diversity, defined as the volume of the functional space occupied by the community [@villeger_new_2008], and its evenness component, both turned out to be major predictors of system resilience, as they helped avoid an overreliance on a few dominant species and the consequent vulnerability to their loss.
 This can be due to (i) temporal complementarity, with the diversity of species response to disturbances and response asynchronicity, resulting in ecological sorting [@Loreau2013; @Sakschewski2016; @Morin2014a], and (ii) functional dissimilarity reducing competition strength [@Loreau2013].
@@
 
-## Shifting effects along the successional trajectory insure forest recovery
+## Shifting effects along the successional trajectory ensure forest recovery
 
@@
 By inducing a higher diversity of shade tolerance and growth response to light in the species pool, species richness hence fosters forest response to perturbations [@Morin2011].
-This complementarity effect quickly decreased through time while the selection effect progressively gained in dominance.
-The diminishing of complementarity effect is attributed to the increased abundance of the most competitive species in more stable and less heterogeneous environments.
+This complementarity effect decreased through time while the selection effect progressively gained in dominance.
+The diminished complementarity effect can be attributed to the increased abundance of the most competitive species in more stable and less heterogeneous environments.
 
@@
 In this study we used an individual-based forest simulator to jointly simulate biodiversity and ecosystem functioning, and we explored the effect of biodiversity on forest resilience over centuries, a timeframe that remains out of reach of most experimental approaches.
-By virtually manipulating the number and identity of species across a large number of simulations, we were able to evidence the role of diversity in forest resilience and the changing driver of biodiversity effect on forest functioning along trajectory of recovery.
+By virtually manipulating the number and identity of species across a large number of simulations, we were able to demonstrate the role of diversity in forest resilience and the changing drivers of biodiversity effects on forest functioning along the trajectory of recovery.
 
@@
 Our analysis thus simulates a situation where aboveground competition for light overcomes the effect of belowground competition for water and nutrients, a reasonable assumption in humid tropical forests, even in early successional stages [@VanBreugel2012].
-However, this assumption may be less suitable under the predicted increase in drought intensity and frequency.
-We did not include the effect of topography, known to result in higher micro environmental heterogeneity.
+However, this assumption may be less suitable under predicted increases in drought intensity and frequency.
+Furthermore, we did not include the effect of topography, known to result in higher micro environmental heterogeneity.
 We expect a stronger role of complementarity effect through hydrological and edaphic niche partitioning as water become locally more limited in more heterogeneous landscapes [@Baltzer2005; @Jucker2018; @Silvertown2004a].
 
-Additionally, we simulated a generic form of disturbance, and our findings might differ with more specific types of disturbance.
+Additionally, our simulations did not take into account the secondary effects of disturbances on community structure, such as damage to surrounding trees or secondary treefalls.
+These might introduce more spatial clustering (i.e.
+larger gaps) and play on the spatial grain.
+Effectively, @Chisholm2013 and @Sullivan2017 found that the biodiversity effect varies across spatial scales, with typically stronger effects at smaller scales (*e.g.* 0.04-ha).
+
+Finally, we simulated a generic form of disturbance, and our findings might differ with more specific types of disturbance.
 For instance, selective logging target a few species, and this could lower the selection effect if dominant species are chosen.
-Repeated disturbances such as periodic harvests would increase the prevalence of early recovery stages during which complementarity effect plays the largest role.
+Repeated disturbances such as periodic harvests, would increase the prevalence of early recovery stages during which complementarity effect plays the largest role.
 Consequently, species complementarity could have an even bigger importance in forest recovery after repeated disturbances that target dominant species, as can be the case with selective logging.
 
-## Unifying biodiversity-ecosystem functioning studies
+## Understanding ecosystem temporal dynamics unifies biodiversity-ecosystem functioning studies
 
-Our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depend on the current ecosystem state and its previous history of disturbances, with dominant complementarity effect in recently disturbed ecosystems whereas selection effect takes over complementarity in more anciently disturbed systems.
-In agreement with this, studies in recently disturbed forest did find a stronger complementarity effect [@Huang2018; @Poorter2017], whereas a dominant selection effect has been found both in undisturbed forests and new plantations [@Bauters2015; @Finegan2015; @Marechaux2017; @Tobner2016] but see [@DeAvila2018; @Lohbeck2016].
-Dominant selection effect of plantation might be explained by the absence of remaining diversity to insure complementarity as opposed to disturbed mature forest (results unshown).
+Previous biodiversity-ecosystem functioning studies are not consistent on dominant biodiversity effects.
+The complementarity effect has been highlighted as dominant experimentally in diverse subtropical forest plantation [@Huang2018; @Niklaus2017a] and observed within logged forest through species richness [@Poorter2017].
+The selection effect, on the other hand, has been shown as prevailing (i) experimentally in old plantations [@Bauters2015], young plantations [@Tobner2016], (ii) simulating old-growth tropical forest [@Marechaux2017], and (iii) observing natural undisturbed forest [@Finegan2015], logged, thinned and burnt forests [@DeAvila2018] and secondary forests [@Lohbeck2016].
 
-We here focus on the effect of biodiversity along a trajectory of post-disturbance recovery, and our simulations and analyses were all at a 16-ha scale.
-However, @Chisholm2013 and @Sullivan2017 found that the biodiversity effect varies across spatial scales, with typically stronger effects at smaller scales (*e.g.* 0.04-ha).
-Ecological constraints may also differ among ecosystems, for instance pioneer species (*e.g.* hard-wood pioneer species in dry forest [@Prado-Junior2016]), light importance (*e.g.* nitrogen importance in grasslands [@Fargione2007; @Prado-Junior2016]), and maturity definition (*e.g.* temperate versus tropical forests) can differ from tropical wet forest to another system, that might cause differences in biodiversity effects.
+However, our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depend on the current ecosystem state and its previous history of disturbances, with dominant complementarity effect in recently disturbed ecosystems and selection effect taking over complementarity in anciently disturbed systems.
+Effectively, @Huang2018, @Niklaus2017a and @Poorter2017 did find a stronger complementarity effect in species-rich ecosystems recently disturbed either by plantation or logging (8 years-old, 14 to 32 years-old, and 18 years old respectively).
+Contrarily, anciently disturbed or undisturbed forests showed dominant selection effect (77 years old planted forest for @Bauters2015; undisturbed forest for @Finegan2015; 500 years old simulated forest for @Marechaux2017).
+In addition, the dominant selection effect observed in plantations by @Tobner2016 or in secondary forest by @Lohbeck2016 might be explained by the recent ecosystem conversion, from agricultural field and fallow respectively, resulting in the absence of already existing tree diversity to insure complementarity as opposed to planted or disturbed forest (see supplementary materials S2).
 
-Together, this paves the way to a unifying framework to reconcile the apparent contradictions between studies advocating for the sole effect of complementarity or selection effects, and call for an accurate description of the ecosystem spatial-temporal and ecological context in future studies to improve global understanding of biodiversity and ecosystem functioning relations.
+Together, this paves the way to a unifying framework to reconcile the apparent contradictions between previous studies advocating for the sole effect of complementarity or selection effects, and call for an accurate description of the ecosystem temporal context and past-history in future studies to improve global understanding of biodiversity and ecosystem functioning relations.
 This further open perspectives to a better prediction of ecosystem resilience in a context of increasing disturbances and biodiversity crisis.
@@
 * Supplementary materials S1: Leaf lifespan model
+* Supplementary materials S2: Biodiversity trajectories
 

### 2019-01-08 Cover letter.Rmd
Draft 0.1.9013 Main document & Cover letter

---
author: 
- Sylvain Schmitt,
- on the behalf of the authors
opening: Dear Editor-in-chief,
closing: Sincerely,
address:
- "__Object__: Submission of a manuscript to Journal of Ecology"
return-address:
- Sylvain Schmitt
- Campus agronomique
- BP 316
- 97379 Kourou, French Guiana
output: linl::linl
---

We would like to submit an original research article entitled “Biodiversity improves tropical forest resilience” for consideration by Journal of Ecology. In this manuscript, we present an original research assessing *in silico* the effect of species and functional diversity on long-term forest recovery from disturbances.

In tropical forests, the role of biodiversity in ecosystem resilience remain under-documented. Using an individual-based and spatially-explicit model of tropical forest simulating high-diversity levels, we assessed the biodiversity effect on forest resilience and underlying mechanisms. We found that functional diversity improved tropical forest resilience from disturbance with underlying ecological mechanisms changing in dominance with time. In addition, our results suggest that pre-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests and can help reconciling previous contrasted results obtained with limited snapshots of ecosystem state.

We believe that this manuscript is appropriate for publication by Journal of Ecology because the journal recently published several articles related to forest or ecosystem resilience focusing on ecological mechanisms, and more generally to the field of Biodiversity and Ecosystem Functioning. Moreover, we believe that our specific research is likely to attract a large audience because it shed new lights on previous contrasted biodiversity-ecosystem functioning studies.

We have no conflicts of interest to disclose. This work is all original research and is not being considered for publication elsewhere

Thank you for your consideration.

### 2019-01-08 Main Document.Rmd
Draft 0.1.9013 Main document & Cover letter

--- 2019-01-07 draft.Rmd
@@
 title: Biodiversity improves tropical forest resilience
-author: Sylvain Schmitt$^{1,*}$, Isabelle Maréchaux$^2$, Jerome Chave $^3$, Fabian Fischer$^3$, Camille  Piponiot$^4$, Stéphane Traissac$^5$, Bruno Hérault$^6$
+author: Sylvain Schmitt$^{*,1,2}$, Isabelle Maréchaux$^3$, Jerome Chave $^4$, Fabian Fischer$^4$, Camille  Piponiot$^5$, Stéphane Traissac$^6$, Bruno Hérault$^{7,8}$
 output:
-  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     keep_tex: yes
-  bookdown::html_document2:
-    number_sections: no
-    toc: yes
-    toc_float: yes
 linestretch: 1.5
@@
 library(kableExtra)
-if(knitr:::is_html_output()) options(knitr.table.format = "html") 
-if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
+options(knitr.table.format = "latex") 
 library(tidyverse)
@@
 
-$^1$ UMR BIOGECO, Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
+```{r BE, echo=FALSE}
+load(file.path(path, 'disturbanceBE.Rdata'))
+load(file.path(path, 'disturbanceBErecovery.Rdata'))
+```
 
-$^2$ AMAP, INRA, University of Montpellier, IRD, CIRAD, CNRS, , Montpellier, France
+$^1$ CNRS, UMR EcoFoG (Agroparistech, Inra, Université des Antilles, Université de la Guyane, Cirad), Campus Agronomique, 97310 Kourou, French Guiana
 
-$^3$ Laboratoire Evolution et Diversité Biologique, UMR5174, CNRS, Université Paul Sabatier, IRD, Toulouse Cedex 9, France
+$^2$ UMR BIOGECO, Université de Bordeaux, 69 route d’Arcachon, F-33612 Cestas cedex France
 
-$^4$ Université de Guyane, UMR EcoFoG (Agroparistech, CNRS, Inra, Université des
+$^3$ AMAP, INRA, University of Montpellier, IRD, CIRAD, CNRS, , Montpellier, France
+
+$^4$ Laboratoire Evolution et Diversité Biologique, UMR5174, CNRS, Université Paul Sabatier, IRD, Toulouse Cedex 9, France
+
+$^5$ Université de Guyane, UMR EcoFoG (Agroparistech, CNRS, Inra, Université des
 Antilles, Cirad), Kourou, French Guiana
 
-$^5$ AgroParisTech, UMR ECOlogie des FOrêts de Guyane, 97387 Kourou, France
+$^6$ AgroParistech, UMR EcoFoG (CNRS, Inra, Université des Antilles, Université de la Guyane, Cirad), Campus Agronomique, 97310 Kourou, French Guiana
 
-$^6$ INPHB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast
+$^7$ Cirad, UMR EcoFoG (AgroParistech, CNRS, Inra, Université des Antilles, Université de la Guyane), Campus Agronomique, 97310 Kourou, French Guiana
 
-$^*$ Corresponding author: sylvain.schmitt@agroparistech.fr
+$^8$ INPHB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast
+
+__$^*$ Corresponding author:__ sylvain.schmitt@agroparistech.fr
+
+\newpage
 
@@
 Resilience; Recovery; Functional diversity; Tropical forest; Complementarity effect; Selection effect; Biodiversity-ecosystem functioning;
+
+\newpage
 
@@
 
-```{r traits, echo=FALSE}
-table <- data.frame(
-  Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
-  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breast height threshold', 'asymptotic height', 'parameter of the tree height-dbh allometry'),
-  Units = c('$\\textrm{g.m}^{-2}$', '$\\textrm{mg.g}^{-1}$', '$\\textrm{mg.g}^{-1}$', '$\\textrm{g.cm}^{-3}$', '$\\textrm{m}$', '$\\textrm{m}$', '$\\textrm{m}$')
-)
-knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Marechaux2017 Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.', format = 'pandoc')
-```
-
 In order to have a general overview of disturbance effect, a generic disturbance module was developed for this study.
@@
 
-```{r, eval=F}
-doe %>% 
-  select(starts_with("CWM")) %>% 
-  reshape2::melt() %>% 
-  group_by(variable) %>% 
-  summarise_all(funs(mean, sd, min, max))
-```
-
 ## Analysis
@@
 
-The recovery of the simulated forest at a time $t$, $R(t)$ regarding a given property $X$ was defined as follows (Figure \@ref(fig:IrGraphFinal)):
+The recovery of the simulated forest at a time $t$, $R(t)$ regarding a given property $X$ was defined as follows (Fig. \@ref(fig:IrGraphFinal)):
 
@@
 When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery.
-We then calculated the Euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Figure \@ref(fig:IrGraphFinal)).
+We then calculated the Euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$ (Fig.
+\@ref(fig:IrGraphFinal)).
 Because we set up the intensity of disturbance, our analysis did not include any ecosystem resistance, i.e.
 the ecosystem capacity to absorb a disturbance event.
-Consequently, in this particular case, we used the inverse of the cumulative Euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Figure \@ref(fig:IrGraphFinal)):
+Consequently, in this particular case, we used the inverse of the cumulative Euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Fig.
+\@ref(fig:IrGraphFinal)):
 
@@
 Distance to equilibrium $d_{eq}$ and resilience index $I_R$ were calculated in a multi-dimensional space for the two sets of ecosystem property values: (i) forest structure ($AGB$, $BA$, $N$, $N10$, and $N30$) and (ii) forest functioning ($GPP$, $NPP$, $Rday$, and $Rnight$).
-
-<!-- $$R[X(t)] = \frac{X_D(t)}{X_C(t)}$$ -->
-<!-- $$d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$$ -->
-<!-- $$I_R(X) = \frac1{\int _{t=disturb_{iter}} ^{t=disturb_{iter}+600} d_{eq}(t).dt}$$ -->
-
-```{r IrGraphFinal, fig.cap="Definition of recovery function R and resilience index $I_R$ shown for aboveground biomass dynamics for a simulation with $disturb_{intensity}=50\\%$."}
-knitr::include_graphics("images/figure3.png")
-```
 
@@
 
-```{r, eval=F}
-doe %>% 
-  select(SR, FDiv, FEve) %>% 
-  reshape2::melt(id.vars = "SR") %>% 
-  mutate(variable = paste0("$", variable, "$")) %>% 
-  group_by(variable, SR) %>% 
-  summarise(mean =  paste0("$", round(mean(value),2), 
-                           "\\pm", round(sd(value),2), "$")) %>% 
-  arrange(variable, desc(SR)) %>% 
-  kable(caption = 'Mean and standard deviation of the predisturbance functional diversity indices for each level of species richness.',
-        format = 'pandoc', col.names = c("Variable", "Species richness ($SR$)", 
-                                         "Mean")) %>% 
-  kable_styling(full_width = F)
-```
-
 As species richness ($SR$) increased from 5 to 125 species, functional diversity varied from $0.76\pm0.18$ to $0.96\pm0.02$ and evenness ($FEve$) varied from $0.46\pm0.24$ to $0.64\pm0.02$.
@@
 Both functional diversity ($FDiv$) and functional evenness ($FEve$) were positively related with $I_R$, i.e.
-forest resilience increased with both functional diversity and functional evenness (Figure \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
+forest resilience increased with both functional diversity and functional evenness (Fig.
+\@ref(fig:gIr) and Table \@ref(tab:tgIr)).
 Similar results were obtained for all disturbance levels, with smaller p-values for higher disturbance intensity.
-
-```{r gIr, echo=FALSE, fig.cap='Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance.
-Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_{intensity}$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$).
-Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.width=10, fig.height=15}
-Ir %>% 
-  left_join(doe, by = "sim") %>% 
-  ungroup() %>% 
-  select(disturbance, 
-         structure, SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
-  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
-  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
-  geom_point(size = 4) +
-  facet_wrap(~disturbance, scales = "free", labeller = 'label_both', 
-             nrow = 3) + 
-  xlab('Functional diversity (FDiv)') +
-  ylab('Resilience indice Ir') +
-  scale_shape_discrete('Species\nrichness\n(SR)') +
-  scale_color_continuous('Functional\nevenness\n(FEve)') +
-  theme(text = element_text(size=20),
-        legend.key.size=unit(4,"line"))
-```
-
-```{r tgIr, cache=T}
-Ir %>% 
-  left_join(doe, by = "sim") %>% 
-  ungroup() %>% 
-  mutate(SR = as.numeric(SR)) %>% 
-  select(disturbance, structure, production,
-         SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
-                 variable.name = "Ir") %>% 
-  group_by(Ir, disturbance) %>%
-  do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
-  # do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
-  tidy(fit) %>% 
-  ungroup() %>% 
-  filter(term != "(Intercept)") %>% 
-  filter(!grepl("SR", term)) %>% 
-  mutate(`P-value` = ifelse(p.value < 0.001, "***",
-                            ifelse(p.value < 0.01, "**",
-                                   ifelse(p.value < 0.05, "*",
-                                          ifelse(p.value < 0.1, ".", 
-                                                 "n.s."))))) %>% 
-  mutate(Coefficient = round(estimate,2)) %>% 
-  mutate(value = paste(Coefficient, `P-value`)) %>%
-  mutate(value = cell_spec(value,
-                           bold = ifelse(grepl("*", value,
-                                               fixed = T), T, F))) %>%
-  reshape2::dcast(Ir + disturbance ~ term, value.var = "value") %>% 
-  left_join(Ir %>% 
-              left_join(doe, by = "sim") %>% 
-              ungroup() %>% 
-              mutate(SR = as.factor(SR)) %>% 
-              select(disturbance, structure, production,
-                     SR, FEve, FDiv) %>% 
-              reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
-                             variable.name = "Ir") %>% 
-              group_by(Ir, disturbance) %>%
-              do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
-              glance(fit) %>% 
-              mutate(r.squared = round(r.squared, 2)) %>% 
-              select(Ir, disturbance, r.squared)) %>% 
-  mutate(Ir = ifelse(Ir == "production", "functioning", "structure")) %>% 
-  kable(caption = "Effects of functional diversity on forest resilience.
-Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$).
-To meet the assumption of normality $I_R$ was log-transformed.
-Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables.
-Values of FDiv and FEve are coefficients from linear regressions.
-Significant $p-values$ are shown in bold.
-ns, non-significant; .
-p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.",
-        escape = F, col.names = c("$I_R$", "Disturbance intensity",
-                                  "FDiv", "FEve", "$R^2$")) %>% 
-  kable_styling(full_width = F) %>% 
-  column_spec(1, bold = T) %>%
-  collapse_rows(columns = 1:2)
-```
-
-```{r anova, eval=F}
-tHSD <- doe %>% 
-  ungroup() %>% 
-  mutate(SR = as.factor(SR)) %>% 
-  select(disturbance,
-         SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR")) %>% 
-  group_by(variable) %>% 
-  do(anova = aov(value ~ SR, data = .)) %>% 
-  mutate(tHSD = TukeyHSD(anova, ordered = F, conf.level = 0.95)) %$%
-  lapply(tHSD, function(x) multcompView::multcompLetters(x[,4])) %>% 
-  lapply(function(x) data.frame(SR = names(x$Letters), group = x$Letters)) %>% 
-  bind_rows() %>% 
-  mutate(axis = c(rep("FEve", 3), rep("FDiv", 3)))
-doe %>% 
-  ungroup() %>% 
-  mutate(SR = as.factor(SR)) %>% 
-  select(disturbance,
-         SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR")) %>% 
-  left_join(tHSD) %>% 
-  ggplot(aes(SR, value, fill = SR, label = group)) +
-  geom_boxplot() +
-  geom_text(aes(y = 1.1)) +
-  ggpubr::stat_compare_means(method = "anova", aes(label = paste("Anova", ..p.signif..))) +
-  facet_wrap(~ variable, scales = "free") +
-  theme(axis.title.x=element_blank(),
-        axis.text.x=element_blank(),
-        axis.ticks.x=element_blank())
-```
-
-```{r residualsHist, eval=F}
-Ir %>% 
-  left_join(doe, by = "sim") %>% 
-  ungroup() %>% 
-  mutate(SR = as.factor(SR)) %>% 
-  select(disturbance, structure, production,
-         SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
-                 variable.name = "Ir") %>% 
-  group_by(Ir, disturbance) %>%
-  do(residuals = residuals(lm(log(value) ~ FEve + FDiv, data = .))) %>% 
-  tidy(residuals) %>% 
-  ggplot(aes(x)) +
-  geom_histogram(bin = 20) +
-  geom_vline(xintercept = 0, col = "red") +
-  facet_grid(~ Ir + disturbance, scales = "free") +
-  xlab("Residuals")
-```
-
-```{r residualsTest, eval=F}
-Ir %>% 
-  left_join(doe, by = "sim") %>% 
-  ungroup() %>% 
-  mutate(SR = as.factor(SR)) %>% 
-  select(disturbance, structure, production,
-         SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
-                 variable.name = "Ir") %>% 
-  group_by(Ir, disturbance) %>%
-  do(residuals = residuals(lm(log(value) ~ FEve + FDiv, data = .))) %>% 
-  tidy(residuals) %>% 
-  group_by(Ir, disturbance) %>% 
-  do(normality = shapiro.test(.$x)) %>%
-  tidy(normality) %>% 
-  transmute(p.value = ifelse(p.value < 0.001, "***",
-                             ifelse(p.value < 0.01, "**",
-                                    ifelse(p.value < 0.05, "*",
-                                           ifelse(p.value < 0.1, ".", 
-                                                  "n.s.")))))
-```
 
@@
 
-```{r BE, echo=FALSE}
-load(file.path(path, 'disturbanceBE.Rdata'))
-load(file.path(path, 'disturbanceBErecovery.Rdata'))
-```
-
 The net effect of biodiversity on each ecosystem property was quantified and partitioned into selection and complementarity effects after normalizing, as explained in the Methods section.
-We found that the complementarity effect sharply increased immediately after disturbance, was the dominant effect of biodiversity during the first decades and then progressively vanished (Figure \@ref(fig:gBE)).
+We found that the complementarity effect sharply increased immediately after disturbance, was the dominant effect of biodiversity during the first decades and then progressively vanished (Fig.
+\@ref(fig:gBE)).
 When increasing the disturbance intensity, the effect of complementarity increased further and dominated even longer.
@@
 a century after disturbance.
-Finally, at high disturbance intensity, the net effect of biodiversity on aboveground biomass was still lower than the control 600 years after disturbance (Figure \@ref(fig:gBE)).
+Finally, at high disturbance intensity, the net effect of biodiversity on aboveground biomass was still lower than the control 600 years after disturbance (Fig.
+\@ref(fig:gBE)).
 We found similar results, but with an amplified signal, for basal area ($BA$) and stem abundance ($N$).
 Finally, forest growth primary productivity ($GPP$) recovered only within a few years with a dominant complementarity effect.
-
-```{r tNE, echo=FALSE, cache=TRUE}
-library(reshape2)
-library(knitr)
-BE %>%
-  mutate(agb = agb/1000) %>% 
-  filter(BE == 'NE') %>%
-  melt(id.vars = c('BE', 'time', 'sim')) %>% 
-  left_join(., doe, by = 'sim') %>%
-  filter(SR == 125) %>%
-  filter(disturbance == 0) %>%
-  group_by(variable) %>% 
-  summarise(mean = mean(value, na.rm = T), 
-            sd = sd(value, na.rm = T)) %>% 
-  filter(variable != 'n30') %>% 
-  bind_cols(., data.frame(
-    name = c('aboveground biomass', 'basal area', 'number of stems', 
-             'number of stems above 10 cm dbh', 'growth primary production', 
-             'net primary production', 'autotrophic respiration during day',
-             'autotrophic respiration during night'),
-    unit = c('$\\textrm{MgC.ha}^{-1}$', '$\\textrm{m}^2.\\textrm{ha}^{-1}$', '$\\textrm{n.ha}^{-1}$', '$\\textrm{n.ha}^{-1}$', 
-             '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$'))) %>% 
-  mutate(NE = paste0("$", round(mean,2), "\\pm", round(sd,2), "$")) %>% 
-  select(name, NE, unit) %>% 
-  kable(caption = 'Mean and standard deviation of the biodiversity net effect across mature simulations before the disturbance event (i.e.
-after 600 years of regeneration) for different ecosystem variables.',
-        format = 'pandoc', col.names = c("Ecosystem property", "Net effect", "Unit")) %>% 
-  kable_styling(full_width = F)
-```
-
-```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss).
-The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\\%$).
-Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125.
-Time axis has been square-root transformed to highlight first decades dynamic.'}
-BEresilience %>% 
-  left_join(., doe, by = 'sim') %>%
-  filter(SR == 125) %>% 
-  # filter(BE %in% c('SE', 'CE')) %>%
-  group_by(time, BE, disturbance, SR) %>%
-  summarise(qt1 =quantile(agb, probs=0.05),
-            qt2 =quantile(agb, probs=0.75),
-            mean=mean(agb),
-            n=n()) %>% 
-  ggplot(aes(x = time)) +
-  geom_hline(yintercept = 1, color = 'red') +
-  geom_hline(yintercept = 0, color = 'blue') +
-  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
-  geom_line(aes(y = mean, color = BE)) +
-  facet_grid(disturbance ~ SR, labeller = 'label_both') +
-  xlab('Time since disturbance (years)') +
-  ylab('Biodiversity effect on aboveground biomass') +
-  scale_x_sqrt() +
-  theme(text = element_text(size=20),
-        legend.key.size=unit(2,"line"))
-```
 
@@
 
-```{r LiteratureBiodiversityEffect, eval=F}
-literature <- data.frame(
-  Study = c("@Bauters2015", "@Finegan2015", "@Prado-Junior2016",
-            "@Lohbeck2016", "@DeAvila2018", "@Marechaux2017", "@Tobner2016",
-            "@Morin2014a", "@Cardinale2007", "@VanderSande2017a",
-            "@Poorter2017", "@Huang2018", "@Niklaus2017a",
-            "@Fargione2007"),
-  StudyType = c("Observation", "Observation", "Observation",
-                "Observation", "Experiment", "Simulation", "Experiment",
-                "Simulation", "Review", "Review",
-                "Observation", "Experiment", "Experiment", 
-                "Experiment"),
-  Type = c("Forest", "Forest", "Forest", 
-           "Forest", "Forest", "Forest", "Forest",
-           "Forest", "Grasslands-Tundra-Estuaries", "Forest",
-           "Forest", "Forest", "Forest", 
-           "Grassland"),
-  Biome = c("Tropical", "Tropical", "Tropical",
-            "Tropical", "Tropical", "Tropical", "Temperate",
-            "Temperate", "Temperate", "Tropical",
-            "Tropical", "SubTropical", "SubToprical", 
-            ""),
-  SubBiome = c("Wet", "Wet-Dry", "Dry",
-               "Wet?", "Wet", "Wet", "",
-               "", "", "",
-               "Wet", "", "", 
-               ""),
-  Effect = c("SE", "SE", "SE",
-             "SE", "SE", "SE", "SE",
-             "CE", "CE", "CE",
-             "CE", "CE", "CE", 
-             "SE-CE"),
-  Disturbed = c("Y", "N?", "Y",
-                "Y", "Y", "N", "Y",
-                "N", "?", "?",
-                "50%Y", "Y", "", 
-                "Y"),
-  Age = c("77yo", "", "12yo",
-          "1-29yo", "21yo", "600yo", "8yo",
-          "200yo", "", "",
-          "14-32yo", "8yo", "18yo", 
-          "10yo"),
-  DisturbationType = c("Plantation", "", "Grazing/Logging",
-                       "Secondary forest from fallow", "Logging-Thinning-Fire", "",
-                       "Plantation",
-                       "", "", "",
-                       "Logging", "Plantation", "Plantation", 
-                       "Plantation")
-)
-literature %>% 
-  select(Effect, Disturbed, Age, DisturbationType, Study, 
-         StudyType, Type, Biome, SubBiome) %>% 
-  arrange(Effect,
-          factor(Disturbed, levels = c("Y", "50%Y", "N", "N?", "?")),
-          factor(Age, levels = c("8yo", "10yo", "12yo", "18yo", "1-29yo",
-                                 "21yo", "14-32yo", "77yo", "200yo", "600yo", " "))) %>% 
-  kable() %>% 
-  kable_styling(full_width = F)
-```
-
 # Supplementary materials
@@
 
-# References
+# References {-}
+
+<div id="refs"></div>
+
+\newpage
+
+## Table captions
+
+__Table \@ref(tab:traits)__ Species-specific parameters used in TROLL from @Marechaux2017 Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.
+
+__Table \@ref(tab:tgIr)__ Effects of functional diversity on forest resilience.
+Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$).
+To meet the assumption of normality $I_R$ was log-transformed.
+Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables.
+Values of FDiv and FEve are coefficients from linear regressions.
+Significant $p-values$ are shown in bold.
+ns, non-significant; .
+p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.
+
+__Table \@ref(tab:tNE)__ Mean and standard deviation of the biodiversity net effect across mature simulations before the disturbance event (i.e.
+after 600 years of regeneration) for different ecosystem variables.
+
+## Figure captions
+
+__Figure \@ref(fig:IrGraphFinal)__ Definition of recovery function R and resilience index $I_R$ shown for aboveground biomass dynamics for a simulation with $disturb_{intensity}=50\%$.
+
+__Figure \@ref(fig:gIr)__ Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance.
+Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_{intensity}$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$).
+Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \epsilon$.
+
+__Figure \@ref(fig:gBE)__ Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss).
+The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\%$).
+Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125.
+Time axis has been square-root transformed to highlight first decades dynamic.
+
+\newpage
+
+<!-- tables -->
+
+```{r traits, echo=FALSE}
+table <- data.frame(
+  Abbreviation = c('$LMA$', '$N_m$', '$P_m$', '$wsg$', '$dbh_{thresh}$', '$h_{lim}$', '$a_h$'),
+  Description = c('leaf mass per area', 'leaf nitrogen content per dry mass', 'leaf phosphorous content per dry mass', 'wood specific gravity', 'diameter at breast height threshold', 'asymptotic height', 'parameter of the tree height-dbh allometry'),
+  Units = c('$\\textrm{g.m}^{-2}$', '$\\textrm{mg.g}^{-1}$', '$\\textrm{mg.g}^{-1}$', '$\\textrm{g.cm}^{-3}$', '$\\textrm{m}$', '$\\textrm{m}$', '$\\textrm{m}$')
+)
+knitr::kable(table, caption = 'Species-specific parameters used in TROLL from @Marechaux2017 Data originates from the BRIDGE [@Baraloto2010] and TRY [@Kattge2011] datasets.', format = 'pandoc')
+```
+
+```{r tgIr}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  mutate(SR = as.numeric(SR)) %>% 
+  select(disturbance, structure, production,
+         SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
+                 variable.name = "Ir") %>% 
+  group_by(Ir, disturbance) %>%
+  do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
+  tidy(fit) %>% 
+  ungroup() %>% 
+  filter(term != "(Intercept)") %>% 
+  filter(!grepl("SR", term)) %>% 
+  mutate(`P-value` = ifelse(p.value < 0.001, "***",
+                            ifelse(p.value < 0.01, "**",
+                                   ifelse(p.value < 0.05, "*",
+                                          ifelse(p.value < 0.1, ".", 
+                                                 "n.s."))))) %>% 
+  mutate(Coefficient = round(estimate,2)) %>% 
+  mutate(value = paste(Coefficient, `P-value`)) %>%
+  mutate(value = cell_spec(value,
+                           bold = ifelse(grepl("*", value,
+                                               fixed = T), T, F))) %>%
+  reshape2::dcast(Ir + disturbance ~ term, value.var = "value") %>% 
+  left_join(Ir %>% 
+              left_join(doe, by = "sim") %>% 
+              ungroup() %>% 
+              mutate(SR = as.factor(SR)) %>% 
+              select(disturbance, structure, production,
+                     SR, FEve, FDiv) %>% 
+              reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
+                             variable.name = "Ir") %>% 
+              group_by(Ir, disturbance) %>%
+              do(fit = lm(log(value) ~ FEve + FDiv, data = .)) %>% 
+              glance(fit) %>% 
+              mutate(r.squared = round(r.squared, 2)) %>% 
+              select(Ir, disturbance, r.squared)) %>% 
+  mutate(Ir = ifelse(Ir == "production", "functioning", "structure")) %>% 
+  kable(caption = "Effects of functional diversity on forest resilience.
+Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$).
+To meet the assumption of normality $I_R$ was log-transformed.
+Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables.
+Values of FDiv and FEve are coefficients from linear regressions.
+Significant $p-values$ are shown in bold.
+ns, non-significant; .
+p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.",
+        col.names = c("$I_R$", "Disturbance intensity",
+                      "FDiv", "FEve", "$R^2$"), escape = F, format = "pandoc") %>% 
+  kable_styling(full_width = F) %>% 
+  column_spec(1, bold = T) %>%
+  collapse_rows(columns = 1:2)
+```
+
+```{r tNE, echo=FALSE, cache=TRUE}
+library(reshape2)
+library(knitr)
+BE %>%
+  mutate(agb = agb/1000) %>% 
+  filter(BE == 'NE') %>%
+  melt(id.vars = c('BE', 'time', 'sim')) %>% 
+  left_join(., doe, by = 'sim') %>%
+  filter(SR == 125) %>%
+  filter(disturbance == 0) %>%
+  group_by(variable) %>% 
+  summarise(mean = mean(value, na.rm = T), 
+            sd = sd(value, na.rm = T)) %>% 
+  filter(variable != 'n30') %>% 
+  bind_cols(., data.frame(
+    name = c('aboveground biomass', 'basal area', 'number of stems', 
+             'number of stems above 10 cm dbh', 'growth primary production', 
+             'net primary production', 'autotrophic respiration during day',
+             'autotrophic respiration during night'),
+    unit = c('$\\textrm{MgC.ha}^{-1}$', '$\\textrm{m}^2.\\textrm{ha}^{-1}$', '$\\textrm{n.ha}^{-1}$', '$\\textrm{n.ha}^{-1}$', 
+             '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$'))) %>% 
+  mutate(NE = paste0("$", round(mean,2), "\\pm", round(sd,2), "$")) %>% 
+  select(name, NE, unit) %>% 
+  kable(caption = 'Mean and standard deviation of the biodiversity net effect across mature simulations before the disturbance event (i.e.
+after 600 years of regeneration) for different ecosystem variables.',
+        format = 'pandoc', col.names = c("Ecosystem property", "Net effect", "Unit")) %>% 
+  kable_styling(full_width = F)
+```
+
+<!-- figures -->
+
+```{r IrGraphFinal, fig.cap="Definition of recovery function R and resilience index $I_R$ shown for aboveground biomass dynamics for a simulation with $disturb_{intensity}=50\\%$."}
+knitr::include_graphics("images/figure3.png") 
+```
+
+```{r gIr, echo=FALSE, fig.cap='Variation of forest resilience with taxonomic and functional diversity for different levels of disturbance.
+Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_{intensity}$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$).
+Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.height=12, dpi=600}
+Ir %>% 
+  left_join(doe, by = "sim") %>% 
+  ungroup() %>% 
+  select(disturbance, 
+         structure, SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
+  ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
+  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
+  geom_point(size = 4) +
+  facet_wrap(~disturbance, scales = "free", labeller = 'label_both', 
+             nrow = 3) + 
+  xlab('Functional diversity (FDiv)') +
+  ylab('Resilience indice Ir') +
+  scale_shape_discrete('Species\nrichness\n(SR)') +
+  scale_color_continuous('Functional\nevenness\n(FEve)') +
+  theme(text = element_text(size=20),
+        legend.key.size=unit(4,"line"))
+```
+
+```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss).
+The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\\%$).
+Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125.
+Time axis has been square-root transformed to highlight first decades dynamic.', fig.height=12, dpi=600}
+BEresilience %>% 
+  left_join(., doe, by = 'sim') %>%
+  filter(SR == 125) %>% 
+  # filter(BE %in% c('SE', 'CE')) %>%
+  group_by(time, BE, disturbance, SR) %>%
+  summarise(qt1 =quantile(agb, probs=0.05),
+            qt2 =quantile(agb, probs=0.75),
+            mean=mean(agb),
+            n=n()) %>% 
+  ggplot(aes(x = time)) +
+  geom_hline(yintercept = 1, color = 'red') +
+  geom_hline(yintercept = 0, color = 'blue') +
+  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
+  geom_line(aes(y = mean, color = BE)) +
+  facet_grid(disturbance ~ SR, labeller = 'label_both') +
+  xlab('Time since disturbance (years)') +
+  ylab('Biodiversity effect on aboveground biomass') +
+  scale_x_sqrt() +
+  theme(text = element_text(size=20),
+        legend.key.size=unit(2,"line"))
+```

### 2019-01-11 Cover letter.Rmd
# Draft 0.1.9014 * Submission

@@
 author: 
-- Sylvain Schmitt,
+- Sylvain Schmitt & Bruno Hérault,
 - on the behalf of the authors
@@
 return-address:
-- Sylvain Schmitt
+- Sylvain Schmitt & Bruno Hérault
 - Campus agronomique

### 2019-01-11 Main Document.Rmd
# Draft 0.1.9014 * Submission

@@
 title: Biodiversity improves tropical forest resilience
-author: Sylvain Schmitt$^{*,1,2}$, Isabelle Maréchaux$^3$, Jerome Chave $^4$, Fabian Fischer$^4$, Camille  Piponiot$^5$, Stéphane Traissac$^6$, Bruno Hérault$^{7,8}$
+author: Sylvain Schmitt$^{*,1,2}$, Isabelle Maréchaux$^3$, Jerome Chave $^4$, Fabian Jörg Fischer$^4$, Camille  Piponiot$^5$, Stéphane Traissac$^6$, Bruno Hérault$^{*,7,8}$
 output:
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
 
-$^3$ AMAP, INRA, University of Montpellier, IRD, CIRAD, CNRS, , Montpellier, France
+$^3$ AMAP, INRA, Univ Montpellier, IRD, CIRAD, CNRS, F-34000 Montpellier, France
 
@@
 $^5$ Université de Guyane, UMR EcoFoG (Agroparistech, CNRS, Inra, Université des
-Antilles, Cirad), Kourou, French Guiana
+Antilles, Cirad), Campus Agronomique, 97310 Kourou, French Guiana
 
@@
 
-$^7$ Cirad, UMR EcoFoG (AgroParistech, CNRS, Inra, Université des Antilles, Université de la Guyane), Campus Agronomique, 97310 Kourou, French Guiana
+$^7$ Cirad, Univ Montpellier, UR Forests & Societies, Montpellier, France
 
-$^8$ INPHB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast
+$^8$ INP-HB, Institut National Polytechnique Félix Houphouët-Boigny, BP 1093, Yamoussoukro, Ivory Coast
 
-__$^*$ Corresponding author:__ sylvain.schmitt@agroparistech.fr
+__$^*$ Corresponding authorw:__ sylvain.schmitt@agroparistech.fr ; bruno.herault@cirad.fr
 
@@
 1.
-Climate change and direct human activities are currently deeply modifying forest disturbance regimes, highlighting the critical need to better understand the drivers of forest resilience.
-In tropical forests, the role of biodiversity in ecosystem recovery and the underlying mechanisms remain under-documented.
-Studies relying on forest stand inventories often lack repeatability, rarely include disturbed forests, and temporal coverage remains too low to assess long-term effects.
-Forest simulators have thus proven useful to investigate the effect of biodiversity on forest functioning through virtual biodiversity experiments.
-However, the problem of biodiversity representation beyond a few plant functional types has long been an impediment in applying this approach to species-rich tropical forests, until now.
+Climate change and direct human activities are modifying the disturbance regimes of tropical forests.
+Understanding the role of their high biological diversity in ecosystem recovery pathway and the underlying ecological mechanisms is of primary importance to predict the fate of tropical forests .
+Studies relying on permanent forest plots often lack repeatability, rarely include disturbed forests, and their temporal coverage remains too low to assess forest long-term dynamics.
+Forest simulators can thus prove useful to investigate the effect of biodiversity on forest functioning through virtual biodiversity experiments.
+However, the problem of biodiversity representation beyond a few plant functional types has long been an impediment in simulating species-rich tropical forests.
 2.
-Using an individual-based and spatially-explicit model of tropical forest simulating high-diversity levels, we assessed *in silico* the effect of species and functional diversity on long-term forest recovery from disturbances of different levels.
-We used this approach to manipulate species number and identity across a large number of simulations, and measured the net effect of biodiversity on simulated forest recovery after disturbance.
-In order to further investigate the mechanisms underlying the effect of biodiversity, we partitioned ecosystem properties into complementarity and selection effects.
+Using an individual-based and trait-based model of tropical forest, we assessed the effect of species and functional diversity on long-term forest recovery from disturbances of different levels by manipulating species number and functional assemblages across a large number of simulations.
+We further partitioned the net effect of biodiversity on ecosystem properties into complementarity and selection effects to investigate the ecological mechanisms that underlie the effect of biodiversity on forest functioning along recovery pathways.
 3.
-We found that functional diversity improved tropical forest resilience from disturbance with underlying ecological mechanisms changing in dominance with time.
-Complementarity effect was dominant soon after disturbance and was progressively overtaken by a selection effect as more competitive species dominated the forest community, a pattern that increased with disturbance intensity.
+We found that functional diversity improved tropical forest resilience after disturbance.
+The ecological mechanisms underlying the effect of biodiversity on forest functioning changed in dominance along post-disturbance trajectories.
+The complementarity effect was dominant soon after disturbance and was progressively overtaken by a selection effect as more competitive species dominated the forest community, a pattern that increased with disturbance intensity.
 4.
 *Synthesis*.
-We found that the mechanisms through which biodiversity influence forest functioning strongly depend on the current ecosystem state and its previous history of disturbances, shifting from dominant complementarity effect in recently disturbed ecosystems to selection effect in anciently disturbed systems.
-Our results suggest that pre-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests and can help reconciling previous contrasted results obtained with limited snapshots of ecosystem state.
-
-<!-- *The Abstract should be written to be understood in isolation from the rest of the paper and must not exceed 350 words.
-Using typically four simple, factual, numbered statements, it should: (i) give the conceptual context, (ii) state the methodological approach, (iii) report the main results and conclusions, and (iv) include a final point headed ‘Synthesis’, which sums up the paper’s key message in generic terms that can be understood by non-specialists, indicating clearly how this study has advanced ecological understanding.* -->
+We found that the mechanisms through which biodiversity influences forest functioning depend on the ecosystem state, shifting from a dominant complementarity effect in recently disturbed systems to a selection effect in anciently disturbed systems.
+Our results suggest that disturbance history is thus a key to our understanding of the biodiversity-ecosystem functioning relationships in tropical forests and can help reconcile contrasting results obtained previously with snapshots of ecosystem state in empirical studies.
 
@@
 
-Resilience; Recovery; Functional diversity; Tropical forest; Complementarity effect; Selection effect; Biodiversity-ecosystem functioning;
+Resilience; Recovery; Functional diversity; Tropical forest; Complementarity effect; Selection effect; Biodiversity-ecosystem functioning.
 
@@
 
-Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015], and while deforestation have been rightly the focus of worldwide attention, the consequences of human-induced disturbance on tropical forests has been less studied [@Putz2010; @Simula2009].
+Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015].
+While deforestation has been the focus of worldwide attention, the consequences of human-induced disturbance on tropical forests have been less studied [@Putz2010; @Simula2009].
 Degraded forests currently cover areas over ten times larger than deforested ones [@Herold2011], a majority of which is being unsustainably logged for timber production [@Blaser2011].
-Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 2.1 billion of tons of carbon dioxide emissions per year representing 25% of greenhouse gas emissions related to deforestation and forest degradation globally [@Pearson2017].
-In addition to direct human impact on forests, climate change modifies regional climates with, for instance, projected increases in the frequency and intensity of convective storms in the Amazon having the potential to increase wind-related tree mortality [@Negron-Juarez2018].
-Understanding the drivers of tropical forest resilience to disturbance is thus critical.
+Forest degradation triggers severe biodiversity loss [@Burivalova2014] and is estimated to account for 2.1 billion of tons of carbon dioxide emissions per year representing 25% of greenhouse gas emissions related to deforestation and forest degradation globally [@Pearson2017].
+In addition to direct human impact on forests, climate change modifies regional climates with, for instance, projected increases in the frequency and intensity of convective storms in the Amazon increasing wind-related tree mortality [@Negron-Juarez2018].
+Understanding the drivers of tropical forest resilience to disturbances is thus critical.
 Here we explored the role of biodiversity in tropical forest recovery after disturbance through a simulation approach.
 
-Ecological resilience refers to the ecosystem ability to resist and to recover from disturbance [@Gunderson2010].
-Following a disturbance, ecosystem recovery is determined by both exogenous (*e.g.* climate, disturbance features) and endogenous (*e.g.* forest structure and composition) factors.
-Logging intensity (the quantity of extracted timber and subsequent biomass loss) determines to a large extent the time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and shifts in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015].
-Among endogenous drivers, baseline floristic composition appears to be a key driver of forest recovery [@Herault2018], as species respond differently to disturbances depending on their ecological strategies and functional properties [@herault_functional_2011].
-Beyond species identity, tree community diversity may also foster forest resilience to disturbances [@Cadotte2011; @Ives2007; @Loreau2013].
+Ecological resilience refers to the ecosystem ability to resist and recover from disturbance [@Gunderson2010].
+Post-disturbance recovery is determined by various factors, among which (i) the disturbance intensity which determines to a large extent the time to recover the ecosystem biomass, lost biodiversity, and tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015], (ii)  the pre-disturbance tree assemblage  as species respond differently to disturbances depending on their ecological strategies and functional properties [@herault_functional_2011] and (iii) tree community diversity which also fosters forest resilience to disturbances [@Cadotte2011; @Ives2007; @Loreau2013].
 
-Two types of mechanisms can underlie the effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001].
-The complementarity effect results from resource partitioning and niche differentiation, and to facilitation, leading to a more efficient use of available resources at the community level.
+Two types of mechanisms can underlie the effect of biodiversity on ecosystem properties: complementarity and selection [@Loreau2001].
+The complementarity effect results from resource partitioning and niche differentiation, and from facilitation, leading to a more efficient use of available resources at the community level.
 The selection effect describes how more diverse communities have a higher probability of including functionally efficient species.
-Moreover the sampling effect implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].
-Previous theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], and direct observations along natural gradients [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017] suggest that both complementarity and selection mechanisms play a crucial role in ecosystem biomass stock and dynamics [@Cardinale2007; @Poorter2017; @VanderSande2017a].
-The relative importance of each mechanism varies in time: for instance @Fargione2007 observed a shift from selection to complementarity effect over time in experimental grasslands, and @Huang2018 observed an increased complementarity over time with no selection effect in forest establishments experiments.
+Moreover, the sampling effect implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].
+Previous theoretical work [@Loreau2010], experiments [@Cardinale2007; @Cardinale2009; @Niklaus2017a; @Huang2018], and direct observations along natural gradients [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017] suggest that both complementarity and selection mechanisms may play a role in ecosystem biomass stock and dynamics [@VanderSande2017a], and their  relative importance can vary in time [@Fargione2007; @Huang2018].
 
@@
 
-In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest, that simulates high-diversity levels through a species-specific parameterization using plant functional traits [@Marechaux2017].
-Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning  relationships *in silico* in tropical forest communities.
-More specifically, we used a simulation approach to assess the effect of species and functional diversity on long-term forest recovery from disturbances of different levels.
-We used this approach to manipulate species number and identity across a large number of simulations, and measured the net effect of biodiversity on simulated forest recovery after disturbance.
-In order to further investigate the mechanisms underlying the effect of biodiversity, we partitioned ecosystem properties into complementarity and selection effects.
-Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest would recover quicker to a disturbance event due to an increased productivity [but see @VanderSande2018 for no effect of species richness on productivity].
+In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest that simulates high-diversity levels through a species-specific parameterization using plant functional traits [@Marechaux2017].
+Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning  relationships *in silico*.
+We used a simulation approach to assess the effect of species and functional diversity on long-term forest recovery from disturbances of different levels.
+We manipulated species number and functional identity across a large number of simulations, and measured the net effect of biodiversity on forest preperties along recovery pathways.
+To further investigate the underlying ecological mechanisms, we partitioned this net effect into complementarity and selection effects.
+Based on the general trend of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forests would recover quicker from a disturbance event due to an increased productivity.
 
@@
 
-TROLL is an individual-based and spatially explicit forest dynamic model.
-The model is described in detail in @Marechaux2017, and only summarized here.
-In TROLL the life cycle (recruitment, growth, seed production and death) is simulated for individual trees $>1\textrm{cm}$ diameter at breast height (dbh).
+TROLL is an individual-based and spatially explicit forest dynamic model (details in @Marechaux2017).
+In TROLL, the life cycle (recruitment, growth, seed production, and death) is simulated for individual trees $>1\textrm{cm}$ diameter at breast height (dbh).
 Trees grow in a light environment computed within a $1\textrm{m}^3$ voxel grid.
 Seedlings below the $1\textrm{cm}$ size class are not modeled on an individual basis, but are part of a seed/seedlings pool.
-A maximum of one tree can establish per squaremeter.
+A maximum of one tree can establish per square meter.
 Once established, the tree is defined by age, dbh, height (h), crown radius (CR), crown depth (CD) and leaf area (LA).
@@
 Photosynthetic capacities, as well as leaf respiration rate, are computed from species-specific leaf concentration of nitrogen (N) and phosphorous (P), and leaf mass per area [LMA; @Atkin2015; @Domingues2010].
-The net amount of assimilated carbon allocated to growth is then partitioned among the different plant organs using empirical factors, and the fraction allocated to woody growth is converted into increments of stem volume using species-specific wood density; while the fraction allocated to leaves is converted into leaf area increment using LMA.
+The net amount of assimilated carbon allocated to growth is then partitioned among the different plant organs using empirical factors.
+Yhe fraction allocated to woody growth is converted into increments of stem volume using species-specific wood density, while the fraction allocated to leaves is converted into leaf area increment using LMA.
 Leaf demography is simulated using species-specific leaf lifespan derived from LMA.
@@
 We thus designed a new leaf lifespan allometry, specific to tropical tree species, used throughout this study (see supplementary materials S1).
-Natural tree deaths are stochastic events and their rate is negatively correlated with wood density within TROLL, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010] an also locally [@Aubry-Kientz2013].
+Natural tree deaths are stochastic events and their rate is negatively correlated with wood density within TROLL, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010] and locally [@Aubry-Kientz2013].
 Additionally, trees can die due to carbon starvation or treefalls.
-Below-ground processes, herbaceous plants, epiphytes and lianas are not simulated in this version of TROLL.
-The source code is written in C++ and version 2.3.1 used in this study are available at https://github.com/TROLL-code/TROLL.
-A further detailed description of the model is available in @Marechaux2017.
+Below-ground processes, herbaceous plants, epiphytes and lianas are not simulated.
+The source code is written in C++ and version 2.3.1 used in this study is available at https://github.com/TROLL-code/TROLL.
 
@@
 Selected individuals are then removed without triggering a treefall to control disturbed biomass volume.
-The operation is repeated until the cumulative basal area of removed trees reaches targeted threshold ($disturb_{intensity}$).
+The operation is repeated until the cumulative basal area of removed trees reaches the targeted threshold ($disturb_{intensity}$).
 
@@
 For each level of species richness, we first created 1000 virtual communities by randomly picking the desired number of species among the regional species pool.
-Then of those 1000 communities, we selected 20 communities with a large range of convex hull volume $CVH$ [@Cornwell2006a] but with a community mean ($CM$) close to the one of the regional species pool ($CM_{LMA} = 115 \pm 19 ~\textrm{g.m}^{-2}$, $CM_{wsg} = 0.73 \pm 0.07~\textrm{g.cm}^{-3}$, $CM_{dmax} = 0.37 \pm 0.1~\textrm{m}$, $CM_{hmax} = 40 \pm 11~\textrm{m}$) in order to control for average functional composition.
+Then of those 1000 communities, we selected 20 communities with a large range of functional trait convex hull volume $CVH$ [@Cornwell2006a] but with a community mean ($CM$) close to the one of the regional species pool ($CM_{LMA} = 115 \pm 19 ~\textrm{g.m}^{-2}$, $CM_{wsg} = 0.73 \pm 0.07~\textrm{g.cm}^{-3}$, $CM_{dmax} = 0.37 \pm 0.1~\textrm{m}$, $CM_{hmax} = 40 \pm 11~\textrm{m}$) in order to control for average functional composition.
 This led to 240 simulations varying in community composition and disturbance intensity ($3~SR*20~CHV*4~ levels~of~disturbance$).
-We focused on functional richness with convex hull volume $CVH$ and functional mean with community weighted mean $CWM$ to define simulation communities.
 
@@
 where $X_D(t)$ and $X_C(t)$ are the values for the ecosystem property $X$ at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the control simulation ($disturb_{intensity} = 0\%$), respectively.
-Note that variability among replicated simulations with the same set-up, due to demographic stochasticity in the model, is low enough to assume a given simulation is representative of the community properties and dynamics given an initial species pool and disturbance intensity [@Marechaux2017].
+Note that variability among replicated simulations with the same set-up, due to demographic stochasticity in the model, is low enough to assume that a given simulation is representative of the community properties and dynamics given an initial species pool and disturbance intensity [@Marechaux2017].
 
@@
 the ecosystem capacity to absorb a disturbance event.
-Consequently, in this particular case, we used the inverse of the cumulative Euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Fig.
+We used the inverse of the cumulative Euclidean distance to the equilibrium as an index of ecosystem resilience, further defined as $I_R$ (Fig.
 \@ref(fig:IrGraphFinal)):
@@
 
-We tested the influence of three factors on the resilience index $I_R$: pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) using linear regression ($I_R \sim SR + FEve + FDiv$).
+We tested the influence of three factors on the resilience index $I_R$: pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$):
+$$log(I_R) \sim \mathcal{N} (\Theta _{SR} + \Theta _{FEve} \times Feve + \Theta _{FDiv} \times Fdiv; \sigma^2)$$
+where $\Theta _{SR}$ is a vector of 3 parameters corresponding to  the 3 species richness levels, $\Theta _{FEve}$ and $\Theta _{FDiv}$ respectively model the effect of $FEve$ and $FDiv$ on $I_R$  under a normal law of variance $\sigma^2$.
 The analysis was conducted twice for each disturbance level (25%, 50% and 75% of basal area loss): one with forest structure properties and one with forest functioning properties.
-We tested variable correlation by Pearson's correlation coefficient for continuous variables and one-way ANOVA for species richness $SR$.
-In case of significant correlation, we selected the variable resulting in the model with the higher Aikaike information criterion in order to have independent variables.
-To meet the assumption of normality, the ecosystem resilience index $I_R$ was log-transformed.
+We used the AIC to select the final, most-parsimonious, models.
+The ecosystem resilience index IR was log-transformed to improve residuals’ normality.
 
@@
 that ecosystem property $X$ for the mixed-species community equals the relative abundance-weighted sum of that in monoculture [@Loreau2001].
-In order to do so, we simulated monoculture of each of the 163 species and for each level of disturbance leading to 652 simulations of monocultures.
+In order to do so, we simulated a monoculture of each of the 163 species and for each level of disturbance leading to 652 simulations of monocultures.
 
@@
 As species richness ($SR$) increased from 5 to 125 species, functional diversity varied from $0.76\pm0.18$ to $0.96\pm0.02$ and evenness ($FEve$) varied from $0.46\pm0.24$ to $0.64\pm0.02$.
-Species richness ($SR$) was significantly correlated to functional indices (one-way ANOVA $p-value < 0.001$ for both $FDiv$ and $FEve$) and model $AIC$ was higher without $SR$, consequently we kept only $FDiv$ and $FEve$ within linear regressions.
+Species richness ($SR$) was never retained by the AIC selection procedure, whatever the disturbance intensity.
 Both functional diversity ($FDiv$) and functional evenness ($FEve$) were positively related with $I_R$, i.e.
@@
 \@ref(fig:gIr) and Table \@ref(tab:tgIr)).
-Similar results were obtained for all disturbance levels, with smaller p-values for higher disturbance intensity.
+Similar results were obtained for all disturbance levels, with higher significance for higher disturbance intensity.
+Model R2s were slightly lower for forest structure (from 0.18 to 0.27) than for forest functioning (from 0.27 to 0.51) variables.
+While FDiv was the most significant variable for forest structure models, Feve became prominent for forest functioning models (Table \@ref(tab:tgIr)).
 
@@
 
-The net effect of biodiversity on each ecosystem property was quantified and partitioned into selection and complementarity effects after normalizing, as explained in the Methods section.
-We found that the complementarity effect sharply increased immediately after disturbance, was the dominant effect of biodiversity during the first decades and then progressively vanished (Fig.
+The net effect of biodiversity on each ecosystem property was quantified and partitioned into selection and complementarity effects (Table \@ref(tab:tNE)).
+The net effect was positive across most ecosystem variables, apart from 2, *i.e.* the number of stems above 10cm DBH and the net primary production for which the average net effects were negative, albeit not significantly different from 0.
+Regarding the aboveground biomass, we found that the complementarity effect sharply increased immediately after disturbance, was the dominant effect of biodiversity during the first decades and then progressively vanished (Fig.
 \@ref(fig:gBE)).
@@
 We found similar results, but with an amplified signal, for basal area ($BA$) and stem abundance ($N$).
-Finally, forest growth primary productivity ($GPP$) recovered only within a few years with a dominant complementarity effect.
+Finally, forest growth primary production ($GPP$) recovered only within a few years with a dominant complementarity effect.
 
@@
 Using an *in silico* experiment, we investigated the effect of biodiversity on tropical forest resilience to disturbance and the variation of biodiversity effect on ecosystem functioning along the recovery trajectory.
-We found that functional diversity improved tropical forest resilience from disturbance with underlying ecological mechanisms changing in dominance with time.
+We found that functional diversity improved tropical forest recovery from disturbance with underlying ecological mechanisms changing in dominance with time.
 Complementarity effect was dominant soon after disturbance and was progressively overtaken by a selection effect, as more competitive species dominated the forest community, a pattern that increased with disturbance intensity.
-Overall, our results suggest that pre-disturbance history is a key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests.
+Overall, our results suggest that pre-disturbance history is key to our understanding of the ecological mechanisms that underlie the resilience of tropical forests.
 This can help reconciling previous contrasted results obtained with limited snapshots of ecosystem state.
@@
 Our results provide strong evidence of a significant positive relationship between functional diversity and forest recovery rate.
-In particular, functional diversity, defined as the volume of the functional space occupied by the community [@villeger_new_2008], and its evenness component, both turned out to be major predictors of system resilience, as they helped avoid an overreliance on a few dominant species and the consequent vulnerability to their loss.
+In particular, the volume of the functional space occupied by the community [@villeger_new_2008], and its evenness component, both turned out to be major predictors of system resilience, as they helped avoid an overreliance on a few dominant species and the consequent vulnerability to their loss.
 This can be due to (i) temporal complementarity, with the diversity of species response to disturbances and response asynchronicity, resulting in ecological sorting [@Loreau2013; @Sakschewski2016; @Morin2014a], and (ii) functional dissimilarity reducing competition strength [@Loreau2013].
-Our results thus confirm @Daz2001a, advocating against an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity of forest ecosystems.
-In addition, increased species richness will *de facto* increase the chances for high functional diversity through the sampling effect [@Loreau1998; @Morin2014a].
+Our results thus confirm @Daz2001a, advocating for an under-evaluated importance of plant functional diversity in ecosystem processes, and @Zhang2012 who highlighted the role of evenness in productivity of forest ecosystems.
+With our model simulations, we provide strong evidence that the high diversity of functional traits in tropical forests may represent an insurance against human-induced changes in disturbance regimes by maintaining a high ecosystem resilience.
 
@@
 
-We found that complementarity among the species was the prevailing effect of biodiversity on forest functioning just after a disturbance.
+We found that complementarity among species was the prevailing effect of biodiversity on forest functioning just after a disturbance.
 This is a direct consequence of the appearance of natural gaps and an increased heterogeneity in light availability, the main limiting resource in dense tropical forest.
-Species are known to vary widely in their demographic response to light availability depending on their functional strategy [@Herault2010; @Laurans2012; @Ruger2009].
+Species are known to vary widely in their demographic response to light availability depending on their functional strategy [@Herault2010;@Ruger2009].
 Pioneer species with productive leaf tissue and light wood tend to occupy gaps fast after their opening, hence creating a shaded environment where late successional species with more conservative leaf tissue later establish.
@@
 In this study we used an individual-based forest simulator to jointly simulate biodiversity and ecosystem functioning, and we explored the effect of biodiversity on forest resilience over centuries, a timeframe that remains out of reach of most experimental approaches.
-By virtually manipulating the number and identity of species across a large number of simulations, we were able to demonstrate the role of diversity in forest resilience and the changing drivers of biodiversity effects on forest functioning along the trajectory of recovery.
 
@@
 In our simulations, light was the unique resource whose variability was explicitly accounted for [@Marechaux2017].
-Our analysis thus simulates a situation where aboveground competition for light overcomes the effect of belowground competition for water and nutrients, a reasonable assumption in humid tropical forests, even in early successional stages [@VanBreugel2012].
-However, this assumption may be less suitable under predicted increases in drought intensity and frequency.
-Furthermore, we did not include the effect of topography, known to result in higher micro environmental heterogeneity.
-We expect a stronger role of complementarity effect through hydrological and edaphic niche partitioning as water become locally more limited in more heterogeneous landscapes [@Baltzer2005; @Jucker2018; @Silvertown2004a].
-
+Our analysis thus simulated a situation where aboveground competition for light overcomes the effect of belowground competition for water and nutrients, a reasonable assumption in humid tropical forests, even in early successional stages [@VanBreugel2012].
+However, this assumption may be less suitable under predicted increases in drought intensity and frequency in many tropical areas [@Duffy2015].
+Furthermore, we did not include the effect of topography, known to result in high local heterogeneity.
+We would expect a stronger role of complementarity effect through hydrological and edaphic niche partitioning as water become locally more limited in more heterogeneous landscapes [@Baltzer2005; @Jucker2018; @Silvertown2004a].
 Additionally, our simulations did not take into account the secondary effects of disturbances on community structure, such as damage to surrounding trees or secondary treefalls.
@@
 Finally, we simulated a generic form of disturbance, and our findings might differ with more specific types of disturbance.
-For instance, selective logging target a few species, and this could lower the selection effect if dominant species are chosen.
-Repeated disturbances such as periodic harvests, would increase the prevalence of early recovery stages during which complementarity effect plays the largest role.
-Consequently, species complementarity could have an even bigger importance in forest recovery after repeated disturbances that target dominant species, as can be the case with selective logging.
+For instance, selective logging target a few species, and this could lower the selection effect in the long-term if dominant species are repeatedly chosen.
+Recurring disturbances such as periodic harvests, would increase the prevalence of early recovery stages during which complementarity effect plays the largest role.
+Consequently, species complementarity could have an even bigger importance in forest recovery after repeated disturbances that target most frequent species.
 
-## Understanding ecosystem temporal dynamics unifies biodiversity-ecosystem functioning studies
+## Looking at temporal dynamics unifies biodiversity-ecosystem functioning studies
 
 Previous biodiversity-ecosystem functioning studies are not consistent on dominant biodiversity effects.
-The complementarity effect has been highlighted as dominant experimentally in diverse subtropical forest plantation [@Huang2018; @Niklaus2017a] and observed within logged forest through species richness [@Poorter2017].
-The selection effect, on the other hand, has been shown as prevailing (i) experimentally in old plantations [@Bauters2015], young plantations [@Tobner2016], (ii) simulating old-growth tropical forest [@Marechaux2017], and (iii) observing natural undisturbed forest [@Finegan2015], logged, thinned and burnt forests [@DeAvila2018] and secondary forests [@Lohbeck2016].
+The complementarity effect has been highlighted as the dominant effect in diverse experimental plantation [@Huang2018; @Niklaus2017a] and reported within disturbed-by-logging forests [@Poorter2017].
+The selection effect, on the other hand, has been shown as prevailing (i) in young [@Tobner2016] and old [@Bauters2015] experimental plantations, (ii) in simulated old-growth tropical forests [@Marechaux2017], and (iii) in natural undisturbed [@Finegan2015], secondary [@Lohbeck2016] or logged, thinned and burnt [@DeAvila2018] forests.
 
-However, our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depend on the current ecosystem state and its previous history of disturbances, with dominant complementarity effect in recently disturbed ecosystems and selection effect taking over complementarity in anciently disturbed systems.
+However, our results suggest that the mechanisms through which biodiversity influence forest functioning strongly depend on the current ecosystem state and its previous history of disturbances, with complementarity effect dominating in recently disturbed ecosystems and selection effect taking over complementarity in anciently disturbed systems.
 Effectively, @Huang2018, @Niklaus2017a and @Poorter2017 did find a stronger complementarity effect in species-rich ecosystems recently disturbed either by plantation or logging (8 years-old, 14 to 32 years-old, and 18 years old respectively).
 Contrarily, anciently disturbed or undisturbed forests showed dominant selection effect (77 years old planted forest for @Bauters2015; undisturbed forest for @Finegan2015; 500 years old simulated forest for @Marechaux2017).
-In addition, the dominant selection effect observed in plantations by @Tobner2016 or in secondary forest by @Lohbeck2016 might be explained by the recent ecosystem conversion, from agricultural field and fallow respectively, resulting in the absence of already existing tree diversity to insure complementarity as opposed to planted or disturbed forest (see supplementary materials S2).
+In addition, the dominant selection effect observed in plantations by @Tobner2016 or in secondary forest by @Lohbeck2016 might be explained by the recent ecosystem conversion, from agricultural field and fallow respectively, resulting in the absence of already existing tree diversity (see supplementary materials S2).
 
-Together, this paves the way to a unifying framework to reconcile the apparent contradictions between previous studies advocating for the sole effect of complementarity or selection effects, and call for an accurate description of the ecosystem temporal context and past-history in future studies to improve global understanding of biodiversity and ecosystem functioning relations.
+Together, our results pave the way to a unifying framework to reconcile the apparent contradictions between previous studies advocating for the sole effect of complementarity or selection effects, and call for an accurate description of the ecosystem temporal context and past-history in future studies aiming at improving our understanding of biodiversity and ecosystem functioning relations.
 This further open perspectives to a better prediction of ecosystem resilience in a context of increasing disturbances and biodiversity crisis.
@@
 
-* METIST
-* GFClim
+This study was funded by the GFclim project (FEDER 20142020, Project GY0006894) and an Investissement d’Avenir grant of the ANR (CEBA: ANR-10-LABEX-0025).
 
@@
 __Table \@ref(tab:tgIr)__ Effects of functional diversity on forest resilience.
-Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$).
+Results of the linear models testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$).
 To meet the assumption of normality $I_R$ was log-transformed.
-Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables.
-Values of FDiv and FEve are coefficients from linear regressions.
-Significant $p-values$ are shown in bold.
+Values of FDiv and FEve are the model coefficients.
+Significant $P-values$ are shown in bold.
 ns, non-significant; .
@@
   kable(caption = "Effects of functional diversity on forest resilience.
-Results of the linear regression testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$).
+Results of the linear models testing the effect of functional diversity indices (FDiv and FEve) on the resilience index ($I_R$) for forest structure and functioning and for each level of disturbance intensity ($disturb_{intensity}$).
 To meet the assumption of normality $I_R$ was log-transformed.
-Test of FDiv and FEve effects was done by linear regression with FDiv and FEve as independent variables.
-Values of FDiv and FEve are coefficients from linear regressions.
-Significant $p-values$ are shown in bold.
+Values of FDiv and FEve are the model coefficients.
+Significant $P-values$ are shown in bold.
 ns, non-significant; .
 p < 0.1; * p < 0.05; ** p < 0.01; *** p < 0.001.",
-        col.names = c("$I_R$", "Disturbance intensity",
+        col.names = c("$log(I_R)$", "Disturbance intensity",
                       "FDiv", "FEve", "$R^2$"), escape = F, format = "pandoc") %>% 
@@
     unit = c('$\\textrm{MgC.ha}^{-1}$', '$\\textrm{m}^2.\\textrm{ha}^{-1}$', '$\\textrm{n.ha}^{-1}$', '$\\textrm{n.ha}^{-1}$', 
-             '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$', '$\\textrm{MgC.ha}^{-1}$'))) %>% 
+             '$\\textrm{MgC.ha}^{-1}\\textrm{.yr}^{-1}$', '$\\textrm{MgC.ha}^{-1}\\textrm{.yr}^{-1}$', 
+             '$\\textrm{MgC.ha}^{-1}\\textrm{.yr}^{-1}$', '$\\textrm{MgC.ha}^{-1}$\\textrm{.yr}^{-1}'))) %>% 
   mutate(NE = paste0("$", round(mean,2), "\\pm", round(sd,2), "$")) %>% 
@@
 Resilience index $I_R$ was represented against functional diversity ($FDiv$) for different levels of disturbance ($disturb_{intensity}$} = 25, 50 and 75% of total basal area); dot shapes varies with species richness whereas dot color indicates functional evenness ($FEve$).
-Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.height=12, dpi=600}
+Solid lines and grey areas show fitted relationships with their mean and variance of the form $I_R = FDiv + \\epsilon$.', fig.height=12, fig.width=12, dpi=600}
 Ir %>% 
@@
   select(disturbance, 
-         structure, SR, FEve, FDiv) %>% 
-  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), value.name = "Ir") %>% 
+         structure, production,
+         SR, FEve, FDiv) %>% 
+  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"), 
+                 variable.name = "Resilience", value.name = "Ir") %>% 
   ggplot(aes(FDiv, Ir, shape = SR, col = FEve, group = NA)) +
@@
   geom_point(size = 4) +
-  facet_wrap(~disturbance, scales = "free", labeller = 'label_both', 
+  facet_wrap(disturbance~Resilience, scales = "free", labeller = 'label_both',
              nrow = 3) + 

### 2019-01-11 S1-LL.Rmd
# Draft 0.1.9014 * Submission

@@
 
-TROLL model previous implementation encompasses Reich's 1991 and 1997 and Wright's 2004 allometries to estimate leaf lifespan with  [@Reich1991a; @Reich1997; @wright_worldwide_2004].
-But we have shown that Reich's allometries underestimates leaf lifespan for low LMA species.
-As a consequence, simulations estimated unrealistically low aboveground biomass due to an early death of individual with low LMA.
-We decided to find a better allometry with @wright_worldwide_2004 GLOPNET dataset.
-
-# Material and methods
-
-We compiled functional traits from GLOPNET [@wright_worldwide_2004], TRY [@Kattge2011], and DRYAD [@chave_towards_2009] databases (see \@ref(tab:A2traits)).
-We kept dataset given by GLOPNET as origin dataset for observations.
-Dataset defined as origin corresponded to leaf lifespan ($LL$) and most of the time to leaf mass per area ($LMA$) and leaf nitrogen content per leaf dry mass ($Nmass$).
-We measured variable importance in functional traits to explain leaf lifespan with an out-of the bag method applied on a random forest.
-Then, we used a Bayesian approach to test different models with growing levels of complexity.
-We retained the model with the best trade-off between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean square error of prediction $RMSEP$).
-We finally tested the new allometry obtained with the selected model with TROLL simulations.
-
-```{r A2traits, echo=FALSE}
-knitr::kable(data.frame(
-  Name = c('LL', 'SLA', 'N', 'P', 'wsg'),
-  Trait = c('Leaf lifespan (longevity)',
-            'Leaf area per leaf dry mass (specific leaf area, SLA)',
-            'Leaf nitrogen (N) content per leaf dry mass',
-            'Leaf phosphorus (P) content per leaf dry mass',
-            'Stem dry mass per stem fresh volume (stem specific density)'),
-  Unit = c('month',' $m^2.kg^{-1}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$mg.mm^{-3}$'),
-  TRYcode = c(12, 11, 15, 14, 4)),
-  caption = 'Functional traits gathered with TRY.', format = 'pandoc')
-```
-
 ```{r A2data, include=FALSE, echo=FALSE, cache=TRUE}
@@
 library(BIOMASS)
+library(tidyverse)
 LES <- read.csv("~/Documents/ECOFOG/TROLL/inst/extdata/LL/GLOPNET.csv", skip = 10)
@@
 
+# Material and methods
+
+We compiled functional traits from GLOPNET [@wright_worldwide_2004] and DRYAD [@chave_towards_2009] databases (43 individuals, see Table \@ref(tab:A2traits)).
+We measured the importance of functional traits to explain leaf lifespan with an out-of-bag method applied on a random forest.
+Then, we used a Bayesian approach to test different models with growing levels of complexity.
+We retained the model with the best trade-off between model complexity (number of parameters), convergence, likelihood, and prediction quality (root mean square error of predictions).
+
+```{r A2traits, echo=FALSE}
+data %>% 
+  dplyr::select(LL, LMA, Nm, wsg) %>% 
+  rename(N = Nm) %>% 
+  melt(variable.name= "Abbreviation") %>% 
+  group_by(Abbreviation) %>% 
+  summarise(Range = paste(round(min(value),2), "-", round(max(value),2))) %>% 
+  mutate(Trait = c('Leaf lifespan',
+            'Leaf mass per area',
+            'Leaf nitrogen',
+            'Wood specific gravity')) %>% 
+  mutate(Unit = c('month',' $m^2.kg^{-1}$', '$mg.g^{-1}$', '$mg.mm^{-3}$')) %>% 
+  select(Trait, Abbreviation, Unit, Range) %>% 
+  knitr::kable(caption = 'Gathered functional traits with abbreviation, unit, and range.', format = 'pandoc')
+```
+
 # Results
 
-Out of the bag method applied on a random forest highlighted the importance of leaf nitrogen content per leaf dry mass ($N_m$) to model leaf lifespan (see \@ref(tab:A2OOB)).
-$N_m$ importance was higher than leaf mass per area (158 against 96 percent of mean square error increase) which was used as a proxy for leaf lifespan in previous models.
-Finally, wood specific gravity ($wsg$) add also a significant importance in leaf lifespan estimation.
+Leaf nitrogen content ($N$) was the most important trait to model leaf lifespan (see Table \@ref(tab:A2OOB)).
+In particular, $N$ importance was higher than leaf mass per area (158 against 96 percent of mean square error increase), which was used as a unique predictor of leaf lifespan in previous models [@Reich1991a; @Reich1997; @wright_worldwide_2004].
+Finally, wood specific gravity (wsg) also improved leaf lifespan estimation.
 
 ```{r A2OOB, echo=FALSE}
-knitr::kable(randomForest::randomForest(LL ~ LMA + Nm + LMA + wsg,
-                   data = data, na.action = na.omit,
+randomForest::randomForest(LL ~ LMA + N + LMA + wsg,
+                   data = rename(data, N = Nm), na.action = na.omit,
                    ntree = 1000, replace = F , importance = T,
-                   do.trace = F, keep.forest = T, keep.inbag = T)$importance,
-             caption = 'Variable importance calculated with out-of the bag method applied on a random forest.
-First column represents the mean decrease in mean square error (%IncMSE) whereas second column represents the total decrease in node impurities, measured by the Gini Index  (IncNodePurety).
-Leaf lifespan (LL) is taken in GLOPNET database from @wright_worldwide_2004.
-Leaf mass per area (LMA), and leaf nitrogen content (Nmass) are taken both in TRY (https://www.try-db.org) and GLOPNET databases.
-Wood specific gravity (wsg) is taken both in TRY and DRYAD databases.', format = 'pandoc')
+                   do.trace = F, keep.forest = T, keep.inbag = T)$importance %>% 
+  as.data.frame() %>% 
+  rename(MSE = `%IncMSE`) %>% 
+  select(MSE) %>% 
+  knitr::kable(caption = 'Variable importance calculated with an out-of-bag method applied on a random forest.
+Variable importance is represented by the mean increase in mean square error (MSE).', format = 'pandoc', col.names = "MSE increase (%)")
 ```
@@
 
-Leaf lifespan $LL$ follows a log-normal distribution with location inferred from leaf lifespan $LMA$, nitrogen content $N$ and wood specific gravity $wsg$ and a scale $\sigma$.
-Each ${\beta_i}_d$ follows a normal distribution located on $\beta_i$ with a scale of $\sigma_i$.
+Leaf lifespan $LL$ follows a log-normal distribution.
+Each ${\beta_i}_d$ follows a normal distribution centered on $\beta_i$.
 All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without preemption following a gamma distribution.
-$d$ represents the dataset random effects and encompass environmental and protocol variations.
+$d$ represents the dataset origin within GLOPNET database and is integrated as a random effect that encompass environmental and protocol variations.
 
-```{r A2LLpred, echo=FALSE, fig.cap='Leaf lifespan predictions for the selected model with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values.
-Leaf lifespan (LL) is predicted with model M10 fit.
-Leaf mass per area (LMA) and leaf nitrogen content (Nmass), and wood specific gravity (wsg) are taken in a composite dataset of GLOPNET, TRY and DRYAD  datasets.Warning LMA (resp.
-Nmass and wsg) is not constant and depend on the closest point value for right (resp.
-center and left) graph.', cache=TRUE, fig.height=12, fig.width=12}
+```{r A2LLpred, echo=FALSE, fig.cap='Leaf lifespan predictions using the selected model with leaf mass per area (LMA), leaf nitrogen content (N), and wood specific gravity (wsg), and predicted versus observed values.
+Points represent the observed values whereas red line represent the mean predicted values and shaded areas the 5-95% confidence interval for the predicted value.', cache=TRUE, fig.height=12, fig.width=12}
 library(ggplot2)

### 2019-01-11 S2-Biodiversity.Rmd
# Draft 0.1.9014 * Submission

@@
 ---
-title: 'Supplementary material S2: Biodiversity trajectories'
+title: 'Supplementary material S2: Biodiversity effect trajectories'
 output:
@@
 
-This document explore the biodiversity trajectories of aboveground biomass partitioned into complementarity and selection effects from forest establishment from a bare soil to 600 years after the disturbance event (Figure \@ref(fig:gBE)).
+This document explores the trajectories of biodiversity effect on aboveground biomass partitioned into complementarity and selection effects from forest establishment from a bare soil to 600 years after the disturbance event (Figure \@ref(fig:gBE)).
 Similarly to the aboveground biomass of the ecosystem, the net biodiversity effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increases following the disturbance (sub-figure on right-bottom on figure \@ref(fig:gBE)).
-Interestingly, we can see that this pattern of decrease of selection effect after disturbance and recovery after is similar to the effect of succession in forest early stages, when the new cohort arrives selection effect suddenly decreases and complementarity effect insure a small net effect (first century, sub-figure on left-bottom on figure \@ref(fig:gBE)).
-Anyway the pattern is stronger after a disturbance, maybe due to an increased diversity allowing a stronger diversity effect compared to the absence of already existing diversity when forest established from bare soil.
+Interestingly, this pattern of post-disturbance decrease of selection effect, followed by a recovery phase, is similar to the effect of succession in forest early stages: when the new cohort arrives selection effect suddenly decreases and complementarity effect insures a small net effect (first century, sub-figure on left-bottom on figure \@ref(fig:gBE)).
+The pattern is stronger after a disturbance, maybe due to an increased diversity allowing a stronger diversity effect compared to the absence of already existing diversity when forest established from bare soil.
 

### 2019-01-11 S3-Partitioning.Rmd
# Draft 0.1.9014 * Submission

---
title: 'Supplementary material S3: ?'
output:
  bookdown::word_document2: default
  bookdown::pdf_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 12,
                cache = T, cache.lazy = F)
```

```{r BE, echo=FALSE}
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
doe <- rename(doe, SR = richness)
load(file.path(path, 'disturbanceBE.Rdata'))
load(file.path(path, 'disturbanceBErecovery.Rdata'))
```

This document explores the trajectories of biodiversity effect on aboveground biomass partitioned into complementarity and selection effects from forest establishment from a bare soil to 600 years after the disturbance event (Figure \@ref(fig:gBE)). Similarly to the aboveground biomass of the ecosystem, the net biodiversity effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increases following the disturbance (sub-figure on right-bottom on figure \@ref(fig:gBE)). Interestingly, this pattern of post-disturbance decrease of selection effect, followed by a recovery phase, is similar to the effect of succession in forest early stages: when the new cohort arrives selection effect suddenly decreases and complementarity effect insures a small net effect (first century, sub-figure on left-bottom on figure \@ref(fig:gBE)). The pattern is stronger after a disturbance, maybe due to an increased diversity allowing a stronger diversity effect compared to the absence of already existing diversity when forest established from bare soil.

```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss). The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\\%$). Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125. Time axis has been square-root transformed to highlight first decades dynamic.', fig.height=12, dpi=600}
BEresilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  select(time, BE, disturbance, SR,
         agb, ba, n, n10, n30,
         gpp, npp, Rday, Rnight) %>% 
  reshape2::melt(id.vars = c("time", "BE", "disturbance", "SR")) %>% 
  group_by(time, BE, disturbance, SR, variable) %>%
  filter(!is.na(value)) %>% 
  summarise(qt1 =quantile(value, probs=0.05),
            qt2 =quantile(value, probs=0.75),
            mean=mean(value),
            n=n()) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_grid(variable ~ disturbance, labeller = 'label_both', scales = "free") +
  xlab('Time since disturbance (years)') +
  ylab('Biodiversity effect') +
  scale_x_sqrt() +
  theme(text = element_text(size=20),
        legend.key.size=unit(2,"line")) +
  ylim(-0.5, 1.5)
```
